### 2015-04-30 Paper/paper.tex
adding paper outline

\documentclass[aps,prl,amsmath,floats,floatfix, twocolumn,
superscriptaddress,nofootinbib,showpacs]{revtex4-1}

\usepackage[colorlinks, pdfborder={0 0 0}, plainpages=false]{hyperref}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[usenames,dvipsnames]{color}
\usepackage{amssymb}

\newcommand{\roughly}{\mathchar"5218\relax} % Different from \sim in spacing

% Macros for text changes
\newcommand{\red}{\textcolor{red}}
\newcommand{\dan}[1]{\textcolor{WildStrawberry}{#1}}
\newcommand{\Mark}[1]{\textcolor{Cerulean}{#1}}
\newcommand{\larry}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\bela}[1]{\textcolor{Blue}{#1}}
\newcommand{\saul}[1]{\textcolor{Orange}{#1}}

% Macros for text notes and comments
\newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}

\newcommand{\Caltech}{\affiliation{Theoretical Astrophysics 350-17,
    California Institute of Technology, Pasadena, CA 91125, USA}}
\newcommand{\Cornell}{\affiliation{Center for Radiophysics and Space
    Research, Cornell University, Ithaca, New York 14853, USA}}
\newcommand{\CITA}{\affiliation{Canadian Institute for Theoretical
    Astrophysics, 60 St.~George Street, University of Toronto,
    Toronto, ON M5S 3H8, Canada}} %
\newcommand{\GWPAC}{\affiliation{Gravitational Wave Physics and
    Astronomy Center, California State University Fullerton,
    Fullerton, California 92834, USA}} %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{
Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO
}

\author{People across the Atlantic}

\date{\today}

\begin{abstract}
Abstract
\end{abstract}

\pacs{}
% 04.25.D- Numerical relativity
% 04.25.dg Numerical studies of black holes and black-hole binaries
% 04.25.Nx Post-Newtonian approximation; perturbation theory; related approximations 
% 04.30.-w Gravitational waves (see also 04.80.Nn Gravitational wave detectors and experiments)
% 04.30.Db Wave generation and sources 
% 02.70.Hm Spectral methods

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Introduction

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modeling tidal deformation during inspiral}
Describe the PN terms used here

\subsection{Modeling tidal disruption near merger}
Describe the Lackey et al model

\subsection{MCMC methods}
Describe the emcee-based PE code

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SNR loss from ignoring NS tidal effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
See what is the loss in SNR if we recover NSBH signals with a template bank of
BH-BH waveforms? It is very likely that the answer would be: $<1\%$. But it
is very interesting to show that for spinning BHs!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Systematic bias in recovered parameters}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability. While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.

\begin{enumerate}
\item Above what SNR values, below what mass-ratios, above what BH spins, etc, 
do we begin to care about NS deformability?\newline
``one interesting plot to make would be the parameter bias (max-Likelihood
parameters minus true parameters) vs SNR, as functionals of q/sBH/Lambda, when
using T signals recovered with N templates.''
\end{enumerate}

Could have subsections for aLIGO and ET ?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using templates with tidal effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Having shown in the previous section that we begin to care for NS tidal effects
for signals in the XXX corner of the paraemter space, with SNRs above YY, here
we investigate the effect of the improvement in the accuracy of the recovered
parameters when tidal effects are included in the templates.

\begin{enumerate}
\item What is the reduction in the bias of the maximum likelihood parameters
when using tidal templates?\newline
``Plot the ratio of the bias between N templates and T templates (against T
signals), as a function of SNR.''
\item What is the reduction in the uncertainty in binary mass and spin, if
any?\newline
``show how the the $95\%$ confidence intervals shrink, as a function of SNR, 
when we go from using N templates to T templates.''
\end{enumerate}

Could have subsections for aLIGO and ET ?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Discussion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgments
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgments}
Acknowledgments
\end{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{References/References}

\end{document}

### 2015-05-18 Paper/paper.tex
minor change
added first figures

@@
-\documentclass[aps,prl,amsmath,floats,floatfix, twocolumn,
+\documentclass[aps,prd,amsmath,floats,floatfix, twocolumn,
 superscriptaddress,nofootinbib,showpacs]{revtex4-1}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[h]
+\centering 
+\textbf{Without BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+\textbf{Without NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
+\caption{These figures show the unfaithfulness of the two EOB models we consider
+in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
+SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
+higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
+We find that both models have high agreement with NR for smaller mass-ratios 
+and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
+waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
+SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
+magnitudes, especially when both spins are aligned with the orbital angular momentum.
+}
+\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering    
+\textbf{Without BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
+\textbf{Without NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
+\caption{These figures show the unfaithfulness of the two EOB models we consider
+in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
+SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
+higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
+We find that both models have high agreement with NR for smaller mass-ratios 
+and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
+waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
+SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
+magnitudes, especially when both spins are aligned with the orbital angular momentum.
+}
+\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering    
+\textbf{Without BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
+\textbf{Without NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
+\caption{These figures show the unfaithfulness of the two EOB models we consider
+in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
+SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
+higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
+We find that both models have high agreement with NR for smaller mass-ratios 
+and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
+waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
+SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
+magnitudes, especially when both spins are aligned with the orbital angular momentum.
+}
+\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering    
+\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
+\caption{These figures show the unfaithfulness of the two EOB models we consider
+in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
+SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
+higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
+We find that both models have high agreement with NR for smaller mass-ratios 
+and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
+waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
+SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
+magnitudes, especially when both spins are aligned with the orbital angular momentum.
+}
+\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\end{figure*}
+% 
+% 
 Advanced LIGO searches and parameter estimation efforts aimed at binaries 

### 2015-05-18 emcee_PE/scripts/sample_analysis.py
modified script to make some plots

@@ -27,8 +27,8 @@ perc_int_99_7 = [0.13, 99.87] # 3 sigma
 
 # Figure settings
 ppi=72.0
-aspect=0.75
-size=4.0 # was 6
+aspect=(5.**0.5 - 1) * 0.5
+size=4.0 * 2# was 6
 figsize=(size,aspect*size)
 plt.rcParams.update({'legend.fontsize':9, 'axes.labelsize':11,'font.family':'serif','font.size':11,'xtick.labelsize':11,'ytick.labelsize':11,'figure.subplot.bottom':0.2,'figure.figsize':figsize, 'savefig.dpi': 150.0, 'figure.autolayout': True})
 
@@ -117,7 +117,7 @@ def PercentileInterval(s, pc=95.0):
 
 def get_bias(basedir = '/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set001/',\
   simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
-  M_inj = 2*1.35, eta_inj = 2./9., chi2_inj=0.5, SNR_inj = 60):
+  M_inj = 2*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, SNR_inj = 60):
     """
     Load samples for a given physical configuration, by first decoding the location
     of the corresponding run. Compute the 90% confidence interval and the median
@@ -126,16 +126,17 @@ def get_bias(basedir = '/home/prayush/projects/nsbh/TidalParameterEstimation/Par
     test_dir = os.path.join(basedir, simdir)
     m1_inj, m2_inj = pnutils.mchirp_eta_to_mass1_mass2(M_inj * eta_inj**0.6, eta_inj)
     params_inj = {'eta' : eta_inj, 'Mtot' : M_inj, 'Mc' : M_inj * eta_inj**0.6,
-                  'chi2' : chi2_inj, 'm1' : m1_inj, 'm2' : m2_inj}
+                  'chi1' : chi1_inj, 'chi2' : chi2_inj, 
+                  'm1' : m1_inj, 'm2' : m2_inj}
     
     match = {}
     match['samples'] = load_samples(test_dir, SNR_inj)
     
     bias, CI90, width90 = {}, {}, {}
-    for param in ['eta', 'Mc', 'chi2', 'm1', 'm2', 'Mtot']:
[…]

### 2015-05-19 Paper/paper.tex
changing figure captions

--- 2015-04-30 Paper/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[h]
+\centering 
+\textbf{With BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+\textbf{With NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
+\caption{These figures show the (fractional) difference between the median value of
+recovered posterior distribution for binary chirp mass and the actual injected chirp mass,
+as a function of the signal-to-noise-ratio (SNR) of the injected signal.
+The top and bottom panels show results for parameter estimation performed using
+waveform templates without and with NS matter effects included, respectively.
+In each row, the left and right panel correspond to 
+$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
+In each panel, the color of the curves correspond to the value of NS deformability 
+parameter, and line style correspond to the dimensionless spin on the BH 
+($\chi_\mathrm{BH}\equiv \dfrac{a}{m_\mathrm{BH}^2}$).
+}
+\label{fig:TNT_chirpMassBias_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering    
+\textbf{With BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
+\textbf{With NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
+\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+with the difference that we consider the systematic bias in the dimensionless
+mass-ratio $\eta$ here.}
+\label{fig:TNT_EtaBias_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering    
+\textbf{With BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
+\textbf{With NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
+\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+with the differences that we consider the systematic bias in the dimensionless
+spin on the BH here, and that the differences shown are absolute and not
+fractions of the injected spin.}
+\label{fig:TNT_BHspinBias_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering
+\textbf{With NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
+\caption{This figure is similar to the bottom row of
+Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
+consider the NS tidal parameter $\Lambda$ here.}
+\label{fig:TT_NSLambdaBias_vs_SNR_q23}
+\end{figure*}
+% 
+% 
 Advanced LIGO searches and parameter estimation efforts aimed at binaries 

### 2015-05-19 Paper/paper.tex
added plots showing the width of 90% confidence interval for different recovered parameters

--- 2015-05-18 Paper/paper.tex
@@
 \centering 
-\textbf{Without BH-BH Templates}
+\textbf{With BH-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-\textbf{Without NS-BH Templates}
+\textbf{With NS-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
-\caption{These figures show the unfaithfulness of the two EOB models we consider
-in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
-SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
-higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
-We find that both models have high agreement with NR for smaller mass-ratios 
-and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
-waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
-SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
-magnitudes, especially when both spins are aligned with the orbital angular momentum.
+\caption{These figures show the (fractional) difference between the median value of
+recovered posterior distribution for binary chirp mass and the actual injected chirp mass,
+as a function of the signal-to-noise-ratio (SNR) of the injected signal.
+The top and bottom panels show results for parameter estimation performed using
+waveform templates without and with NS matter effects included, respectively.
+In each row, the left and right panel correspond to 
+$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
+In each panel, the color of the curves correspond to the value of NS deformability 
+parameter, and line style correspond to the dimensionless spin on the BH 
+($\chi_\mathrm{BH}\equiv \dfrac{a}{m_\mathrm{BH}^2}$).
 }
-\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\label{fig:TNT_chirpMassBias_vs_SNR_q23}
 \end{figure*}
@@
 	\centering 
-\textbf{Without BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
-\textbf{Without NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
-\caption{These figures show the unfaithfulness of the two EOB models we consider
-in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
-SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
-higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
-We find that both models have high agreement with NR for smaller mass-ratios 
-and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
-waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
-SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
-magnitudes, especially when both spins are aligned with the orbital angular momentum.
+	\textbf{With BH-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+	\textbf{With NS-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
+	\caption{These figures show the narrowing of the $90\%$ confidence 
+		interval (recovered) for chirp mass, as a function of the 
+		signal-to-noise-ratio (SNR). Note that the confidence interval
+                width is normalized by the true parameter value.
+                The binary parameters, i.e. black hole spin,
+                mass-ratio, NS tidal deformability, are all depicted in the 
+                same way as in Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}.
 	}
-\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+	\label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
 \end{figure*}
@@
 \centering    
-\textbf{Without BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
-\textbf{Without NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
-\caption{These figures show the unfaithfulness of the two EOB models we consider
-in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
-SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
-higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
-We find that both models have high agreement with NR for smaller mass-ratios 
-and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
-waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
-SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
-magnitudes, especially when both spins are aligned with the orbital angular momentum.
-}
-\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\textbf{With BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
+\textbf{With NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
+\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+with the difference that we consider the systematic bias in the dimensionless
+mass-ratio $\eta$ here.}
+\label{fig:TNT_EtaBias_vs_SNR_q23}
 \end{figure*}
@@
 	\centering    
+	\textbf{With BH-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
+	\textbf{With NS-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}%\quad
+	\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+		with the difference that we consider the confidence intervals in the recovery
+		of the dimensionless
+		mass-ratio $\eta$ here.}
+	\label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}[h]
+\centering    
+\textbf{With BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
+\textbf{With NS-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
+\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+with the differences that we consider the systematic bias in the dimensionless
+spin on the BH here, and that the differences shown are absolute and not
+fractions of the injected spin.}
+\label{fig:TNT_BHspinBias_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}[h]
+	\centering    
+	\textbf{With BH-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+	\textbf{With NS-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}%\quad
+	\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+		with the difference that we consider the confidence intervals in the recovery
+		of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.}
+	\label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}
+\centering
+\textbf{With NS-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
-\caption{These figures show the unfaithfulness of the two EOB models we consider
-in this paper against NR, with the top panel corresponding to SEOBNRv1 and the bottom panel to
-SEOBNRv2. Note that the SEOBNRv1 model does not include component spins
-higher than $+0.6$, we cannot compute its faithfulness in that region of binary parameter space.
-We find that both models have high agreement with NR for smaller mass-ratios 
-and/or small spin magnitudes. The SEOBNRv1 model significantly deviates from the NR 
-waveform when the spin of the primay black hole reaches $\sim +0.6$ for $q=3$, while
-SEOBNRv2 appears to be unfaithful to NR when \textit{both} black holes have high spin 
-magnitudes, especially when both spins are aligned with the orbital angular momentum.
-}
-\label{fig:SEOB_unfaith_TotalMass_Spin1z_Spin2z}
+\caption{This figure is similar to the bottom row of
+Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
+consider the NS tidal parameter $\Lambda$ here.}
+\label{fig:TT_NSLambdaBias_vs_SNR_q23}
+\end{figure*}
+% 
+\begin{figure*}
+	\centering
+	\textbf{With NS-BH Templates}
+	\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
+	\caption{This figure is similar to the bottom row of
+		Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  we
+		consider the NS tidal parameter $\Lambda$ here.}
+	\label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
 \end{figure*}

### 2015-05-19 emcee_PE/scripts/sample_analysis.py
changed plot script to plot the parameter confidence intervals

@@ -150,7 +150,8 @@ def get_bias(basedir = '/home/prayush/projects/nsbh/TidalParameterEstimation/Par
 linestyles = ['-', '--', '-.', '-x', '--o']
 linecolors = ['r', 'g', 'b', 'k', 'm', 'y']
 
-def make_bias_plot(plotparam='Mc', normalize='', xlabel='$\\rho$', ylabel=None, savefig='plots/plot.png'):
+def make_bias_plot(plotparam='Mc', plotquantity='bias', normalize='',\
+                    xlabel='$\\rho$', ylabel=None, savefig='plots/plot.png'):
   # parameter bias
   plt.figure(int(1e7 * np.random.random()))
   for pltid, q in enumerate(qvec):
@@ -159,7 +160,7 @@ def make_bias_plot(plotparam='Mc', normalize='', xlabel='$\\rho$', ylabel=None, 
       for Lidx, Lambda in enumerate(Lambdavec):
         if 'Lambda' in normalize: norm = Lambda
         else: norm = 1.
-        param_bias = {}
+        param_bias, param_CIwidth90 = {}, {}
         for SNR in SNRvec:
           for Ns in Nsamples:
             for Nw in Nwalkers:
@@ -174,7 +175,9 @@ def make_bias_plot(plotparam='Mc', normalize='', xlabel='$\\rho$', ylabel=None, 
                       chi1_inj=chi1val, chi2_inj=chi2val)
               for kk in bias:
                 if kk in param_bias: param_bias[kk].append(bias[kk])
-                else: param_bias[kk] = [bias[kk]]            
+                else: param_bias[kk] = [bias[kk]]
+                if kk in param_CIwidth90: param_CIwidth90[kk].append(width90[kk])
+                else: param_CIwidth90[kk] = [width90[kk]]
         # Make the plot legend
         if pltid == 0:
           if Lidx == 0:
@@ -187,11 +190,14 @@ def make_bias_plot(plotparam='Mc', normalize='', xlabel='$\\rho$', ylabel=None, 
             labelstring = '$\Lambda = %.0f$' % (Lambda)
           else: labelstring = ''
         # Make the acual plot
+        if plotquantity == 'bias': 
+          plotvec = np.array(param_bias[plotparam])/norm
[…]

### 2015-05-20 Paper/paper.tex
added brief description of current status of resuls

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Introduction
+Current status:
 
+\begin{enumerate}
+\item We choose the following values of different NS-BH binary parameters:
+\begin{enumerate}
+\item $m_\mathrm{NS}=1.35M_\odot$;
+\item $q=m_\mathrm{BH}/\mathrm{NS}=\{2,3\}$;
+\item $\chi_\mathrm{NS}=0$;
+\item $\chi_\mathrm{BH}=\{-0.5, 0, +0.5\}$;
+\item $\Lambda =\{500,1000,2000\}$.
+\end{enumerate}
+
+\item For each system comprised of all unique combinations of the above 
+parameter choices, we carry out two parameter estimation tests: 
+\begin{enumerate}
+\item NS-BH signals injected, recovered with BH-BH waveforms, and
+\item NS-BH signals injected, recovered with NS-BH waveforms.
+\end{enumerate}
+Injections were made at different SNR values: 
+\begin{enumerate}
+\item $\rho = \{20, 30, 40, 50, 60, 70, 80, 90, 100\}$.
+\end{enumerate}
+Choice of emcee parameters:
+\begin{enumerate}
+\item $N_\mathrm{samples}=6000$; 
+\item $N_\mathrm{walkers}=100$; 
+\item $N_\mathrm{burn-in}=500$; 
+\end{enumerate}
+and choice of prior boundaries:
+\begin{enumerate} 
+\item $m_\mathrm{BH}\in [1.2, 25]M_\odot$;
+\item $m_\mathrm{NS}\in [1.2, 15]M_\odot$;
+\item $\Lambda\leq 2000$; 
+\item $\sigma\left(\Lambda\right) = 100$ (for chains where templates are NS-BH waveforms).
+\end{enumerate}
+
+\item Results are shown in
+Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}-~\ref{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}.
+
+\item We deduce from
+\begin{enumerate}
+\item large chirp mass biases, 
+\item lack of convergence of $90\%$ confidence intervals with $\rho$ (except for
+$\Lambda$, noticeably), and
+\item jagged-ness of parameter bias curves against $\rho$
+\end{enumerate}
+that the MCMC chains have not yet converged.
+\end{enumerate}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-01 emcee_PE/scripts/sample_analysis.py
add ability to read data from intermediate files for chian/loglike data

@@ -71,6 +71,137 @@ def chi_PN(eta, chi1, chi2):
     return chi_s * (1.0 - 76.0/113.0 * eta) + delta * chi_a # Ajith's PN reduced spin parameter
     #}}}
 
+# Function to remove duplicate entries from a list
+def f7(seq):
+    seen = set()
+    seen_add = seen.add
+    return [ x for x in seq if not (x in seen or seen_add(x))]
+
+def combine_pickles(all_data_files):
+    #{{{
+    all_pickles = [np.load(fnam) for fnam in all_data_files]
+    pid = 0
+    for pickle in all_pickles:
+      if np.shape(np.shape(pickle))[0] != 3:
+        if verbose: 
+          print >>sys.stdout, "Skipping subfile %s" % all_data_files[pid]
+        continue
+      if pid == 0:
+        combined = pickle
+      else: combined = np.append( combined, pickle, axis=0 )
+    #
+    return combined
+    #}}}
+
+def read_run_part_names(dataDir, SNR, burnin=500, useMaxNRun=True, \
+                      chain_number=-1, verbose=True):
+    #{{{
+    import glob
+    # Figure out the random integer associated with each run
+    all_files = glob.glob(dataDir + "/chain*-*.npy")
+    chain_nums= [int(f.split('/')[-1].split('-')[0].strip('chain')) for f in all_files]
+    chain_nums = f7(chain_nums)
+    # Get all continuation files for each
+    part_files = {}
+    for chnum in chain_nums:
+      tmp = glob.glob(dataDir + ("/chain%d-?.npy" % chnum))
+      tmp.sort()
+      part_files[chnum] = tmp
+      tmp = glob.glob(dataDir + ("/chain%d-??.npy" % chnum))
+      tmp.sort()
+      part_files[chnum].extend( tmp )
+      tmp = glob.glob(dataDir + ("/chain%d-???.npy" % chnum))
+      tmp.sort()
+      part_files[chnum].extend( tmp )
+    # Which run to use?
+    if useMaxNRun:
+      last_len, last_chnum = -1, -1
+      for chnum in part_files:
+        if len(part_files[chnum]) > last_len:
+          last_len = len(part_files[chnum])
+          last_chnum = chnum
+    else: 
+      last_chnum, last_len = chain_number, len(part_files[chain_number])
[…]

### 2015-07-01 emcee_PE/src/emcee_match_aligned.py

@@ -169,6 +169,56 @@ def Mchirpfun(M, eta):
 def Mfun(Mc, eta):
   return Mc*eta**(-3.0/5.0)
 
+# Function to remove duplicate entries from a list
+def f7(seq):
+    seen = set()
+    seen_add = seen.add
+    return [ x for x in seq if not (x in seen or seen_add(x))]
+
+def read_run_part_names(dataDir, SNR, burnin=500, useMaxNRun=True, \
+                      chain_number=-1, verbose=True):
+    #{{{
+    import glob
+    # Figure out the random integer associated with each run
+    all_files = glob.glob(dataDir + "/chain*-*.npy")
+    chain_nums= [int(f.split('/')[-1].split('-')[0].strip('chain')) for f in all_files]
+    chain_nums = f7(chain_nums)
+    # Get all continuation files for each
+    part_files = {}
+    for chnum in chain_nums:
+      tmp = glob.glob(dataDir + ("/chain%d-?.npy" % chnum))
+      tmp.sort()
+      part_files[chnum] = tmp
+      tmp = glob.glob(dataDir + ("/chain%d-??.npy" % chnum))
+      tmp.sort()
+      part_files[chnum].extend( tmp )
+      tmp = glob.glob(dataDir + ("/chain%d-???.npy" % chnum))
+      tmp.sort()
+      part_files[chnum].extend( tmp )
+    # Which run to use?
+    if useMaxNRun:
+      last_len, last_chnum = -1, -1
+      for chnum in part_files:
+        if len(part_files[chnum]) > last_len:
+          last_len = len(part_files[chnum])
+          last_chnum = chnum
+    else: 
+      last_chnum, last_len = chain_number, len(part_files[chain_number])
+    if verbose: 
+      print >>sys.stdout, "run %d, %d subfiles" % (last_chnum, last_len)
+    # If no runs are found
+    if last_len <= 0 or last_chnum <= 0:
+      return [], []
+    # Now open each continuation file for the chosen run and gather samples
+    print part_files[last_chnum]
+    part_files_loglike = [fnam.replace('chain','loglike') for fnam in part_files[last_chnum]]    
+    #chain = combine_pickles( part_files[last_chnum] )
+    #loglike = combine_pickles( part_files_loglike )
+    chain = np.load(part_files[last_chnum][last_len-1])
+    loglike = np.load(part_files_loglike[last_len-1])
+    #
+    return chain, loglike
+    #}}}
 
 # Parameter settings
[…]

### 2015-07-01 emcee_PE/src/emcee_match_tidal.py

@@ -167,6 +167,56 @@ def Mchirpfun(M, eta):
 def Mfun(Mc, eta):
   return Mc*eta**(-3.0/5.0)
 
+# Function to remove duplicate entries from a list
+def f7(seq):
+    seen = set()
+    seen_add = seen.add
+    return [ x for x in seq if not (x in seen or seen_add(x))]
+
+def read_run_part_names(dataDir, SNR, burnin=500, useMaxNRun=True, \
+                      chain_number=-1, verbose=True):
+    #{{{
+    import glob
+    # Figure out the random integer associated with each run
+    all_files = glob.glob(dataDir + "/chain*-*.npy")
+    chain_nums= [int(f.split('/')[-1].split('-')[0].strip('chain')) for f in all_files]
+    chain_nums = f7(chain_nums)
+    # Get all continuation files for each
+    part_files = {}
+    for chnum in chain_nums:
+      tmp = glob.glob(dataDir + ("/chain%d-?.npy" % chnum))
+      tmp.sort()
+      part_files[chnum] = tmp
+      tmp = glob.glob(dataDir + ("/chain%d-??.npy" % chnum))
+      tmp.sort()
+      part_files[chnum].extend( tmp )
+      tmp = glob.glob(dataDir + ("/chain%d-???.npy" % chnum))
+      tmp.sort()
+      part_files[chnum].extend( tmp )
+    # Which run to use?
+    if useMaxNRun:
+      last_len, last_chnum = -1, -1
+      for chnum in part_files:
+        if len(part_files[chnum]) > last_len:
+          last_len = len(part_files[chnum])
+          last_chnum = chnum
+    else: 
+      last_chnum, last_len = chain_number, len(part_files[chain_number])
+    if verbose: 
+      print >>sys.stdout, "run %d, %d subfiles" % (last_chnum, last_len)
+    # If no runs are found
+    if last_len <= 0 or last_chnum <= 0:
+      return [], []
+    # Now open each continuation file for the chosen run and gather samples
+    print part_files[last_chnum]
+    part_files_loglike = [fnam.replace('chain','loglike') for fnam in part_files[last_chnum]]    
+    #chain = combine_pickles( part_files[last_chnum] )
+    #loglike = combine_pickles( part_files_loglike )
+    chain = np.load(part_files[last_chnum][last_len-1])
+    loglike = np.load(part_files_loglike[last_len-1])
+    #
+    return chain, loglike
+    #}}}
 
 # Parameter settings
[…]

### 2015-07-01 Paper/paper.tex
new plots

@@
 \centering 
-\textbf{With BH-BH Templates}
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=2\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-\textbf{With NS-BH Templates}
+\textbf{NS-BH Injection; NS-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
@@
 	\centering 
-	\textbf{With BH-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
+	\textbf{BH-BH Injection; BH-BH Templates}
+	\includegraphics[width=1.85\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+	\textbf{NS-BH Injection; BH-BH Templates}
+	\includegraphics[width=1.85\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+	\textbf{NS-BH Injection; NS-BH Templates}
+	\includegraphics[width=1.85\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
 	\caption{These figures show the narrowing of the $90\%$ confidence 
@@
 \centering    
-\textbf{With BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
-\textbf{With NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
 \caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
@@
 \centering    
-	\textbf{With BH-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}%\quad
 \caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
   with the difference that we consider the confidence intervals in the recovery
-		of the dimensionless
-		mass-ratio $\eta$ here.}
+  of the dimensionless mass-ratio $\eta$ here.}
 \label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
@@
 \centering    
-\textbf{With BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
-\textbf{With NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
 \caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
@@
 \centering    
-	\textbf{With BH-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.85\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
 \caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
@@
 \centering
-\textbf{With NS-BH Templates}
+\textbf{NS-BH Injection; NS-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
@@
 \centering
-	\textbf{With NS-BH Templates}
+\textbf{NS-BH Injection; NS-BH Templates}
 \includegraphics[width=2\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
 \caption{This figure is similar to the bottom row of
-		Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  we
-		consider the NS tidal parameter $\Lambda$ here.}
+Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
+we consider the NS tidal parameter $\Lambda$ here.}
 \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}

### 2015-07-05 Paper/paper.tex
updates to the paper

--- 2015-05-20 Paper/paper.tex
@@
 \item For each system comprised of all unique combinations of the above 
-parameter choices, we carry out two parameter estimation tests: 
+parameter choices, we carry out three parameter estimation tests: 
 \begin{enumerate}
-\item NS-BH signals injected, recovered with BH-BH waveforms, and
+\item BH-BH signals injected, recovered with BH-BH waveforms (control),
+\item NS-BH signals injected, recovered with BH-BH waveforms, (current approach) and
 \item NS-BH signals injected, recovered with NS-BH waveforms.
@@
 \begin{enumerate}
-\item $\rho = \{20, 30, 40, 50, 60, 70, 80, 90, 100\}$.
+\item $\rho = \{20, 30, 50, 70, 90, 120\}$.
 \end{enumerate}
@@
 \begin{enumerate}
-\item $N_\mathrm{samples}=6000$; 
+\item $N_\mathrm{samples}=150,000$; 
 \item $N_\mathrm{walkers}=100$; 
@@
 \item $m_\mathrm{NS}\in [1.2, 15]M_\odot$;
-\item $\Lambda\leq 2000$; 
+\item $\Lambda\leq 4000$; 
 \item $\sigma\left(\Lambda\right) = 100$ (for chains where templates are NS-BH waveforms).
@@
 
-\item We deduce from
-\begin{enumerate}
-\item large chirp mass biases, 
-\item lack of convergence of $90\%$ confidence intervals with $\rho$ (except for
-$\Lambda$, noticeably), and
-\item jagged-ness of parameter bias curves against $\rho$
-\end{enumerate}
-that the MCMC chains have not yet converged.
 \end{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{SNR loss from ignoring NS tidal effects}
+\section{How are detection searches affected by not including NS matter effects}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Systematic bias in recovered parameters}
+\section{How is PE affected by not including NS matter effects?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[h]
+\begin{figure*}
 \centering 
-\textbf{With BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-\textbf{With NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
-\caption{These figures show the (fractional) difference between the median value of
-recovered posterior distribution for binary chirp mass and the actual injected chirp mass,
+% \textbf{BH-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
+\caption{These figures show the (fractional) difference between the chirp mass 
+value corresponding to the median of its posterior probability distribution,
+and the actual injected chirp mass,
 as a function of the signal-to-noise-ratio (SNR) of the injected signal.
-The top and bottom panels show results for parameter estimation performed using
-waveform templates without and with NS matter effects included, respectively.
+% In the top row, NS matter effects on the inspiral and merger waveform
+% are neithed included in templates nor in the injected signal. 
+In the top row, NS matter effects (inspiral and merger) are included in the
+injections but not in templates. This corresponds to the current 
+plan for aLIGO parameter estimation studies. The bottom row shows the effect
+of additionally including NS matter effects in the templates.
+% 
 In each row, the left and right panel correspond to 
 $q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
-In each panel, the color of the curves correspond to the value of NS deformability 
-parameter, and line style correspond to the dimensionless spin on the BH 
-($\chi_\mathrm{BH}\equiv \dfrac{a}{m_\mathrm{BH}^2}$).
+In each panel, the color of each curve corresponds to the value of NS 
+deformability parameter $(\Lambda=\{500,1000,2000\})$, 
+and line-style corresponds to the value of
+the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
 }
@@
 % 
-\begin{figure*}[h]
+\begin{figure*}
 \centering 
-	\textbf{With BH-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
-	\caption{These figures show the narrowing of the $90\%$ confidence 
-		interval (recovered) for chirp mass, as a function of the 
-		signal-to-noise-ratio (SNR). Note that the confidence interval
-                width is normalized by the true parameter value.
-                The binary parameters, i.e. black hole spin,
-                mass-ratio, NS tidal deformability, are all depicted in the 
-                same way as in Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}.
+% \textbf{BH-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
+\caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
+with the difference that here we show the width of the $90\%$ confidence 
+interval (recovered) for chirp mass, as a function of the injected SNR. 
+Note that the confidence interval's width is normalized by the injected
+parameter value.
 }
@@
 % 
-\begin{figure*}[h]
+\newpage
+\newpage
+
+\begin{figure*}[!t]
 \centering    
-\textbf{With BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
-\textbf{With NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
-\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
+\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
 with the difference that we consider the systematic bias in the dimensionless
-mass-ratio $\eta$ here.}
+mass-ratio $\eta$ here. In the bottom row we show control runs where both
+signal and template waveforms lack matter effects.}
 \label{fig:TNT_EtaBias_vs_SNR_q23}
@@
 % 
-\begin{figure*}[h]
+\begin{figure*}
 \centering    
-	\textbf{With BH-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}%\quad
-	\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
+\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
   with the difference that we consider the confidence intervals in the recovery
-		of the dimensionless
-		mass-ratio $\eta$ here.}
+  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
+  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
+  confidence interval is shown here.}
 \label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
@@
 % 
-\begin{figure*}[h]
+\begin{figure*}
 \centering    
-\textbf{With BH-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
-\textbf{With NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
-\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
+\caption{The top two rows in this figure are
+similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
 with the differences that we consider the systematic bias in the dimensionless
 spin on the BH here, and that the differences shown are absolute and not
-fractions of the injected spin.}
+fractions of the injected spin. In the bottom row, we show a control run where
+both injected and template waveforms correspond to binary black holes.}
 \label{fig:TNT_BHspinBias_vs_SNR_q23}
@@
 % 
-\begin{figure*}[h]
+\begin{figure*}
 \centering    
-	\textbf{With BH-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}%\quad
-	\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+\textbf{BH-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; BH-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
+\caption{The top two rows in the figure are similar to
+Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
 with the difference that we consider the confidence intervals in the recovery
-		of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.}
+of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
+As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
+width of the confidence interval without normalizing with the injected spin.
+The bottom row show a control run where matter effects are ignored both 
+in the injected signal and the template waveforms.}
 \label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
@@
 \centering
-\textbf{With NS-BH Templates}
-\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
 \caption{This figure is similar to the bottom row of
@@
 \centering
-	\textbf{With NS-BH Templates}
-	\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
+\textbf{NS-BH Injection; NS-BH Templates}
+\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
 \caption{This figure is similar to the bottom row of
-		Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  we
-		consider the NS tidal parameter $\Lambda$ here.}
+Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
+we consider the NS tidal parameter $\Lambda$ here.}
 \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
@@
 containing NS and stellar-mass BHs are poised to use waveform models that do not
-include the effects of the NS tidal deformability. While this is not expected to
+include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
+While this is not expected to
 be the dominant source of error at low SNRs, it will likely introduce a
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Using templates with tidal effects}
+\section{What do we gain by using templates that include NS matter effects?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 any?\newline
-``show how the the $95\%$ confidence intervals shrink, as a function of SNR, 
+``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
 when we go from using N templates to T templates.''
@@
 
-Could have subsections for aLIGO and ET ?
+\red{Could have subsections for aLIGO and ET ?}
 

### 2015-11-08 emcee_PE/scripts/AnalyzeSamples.py
script to filter out paraemter biases etc into an HDF file

@@ -1,0 +1,448 @@
+#!/usr/bin/env python
+# Tools for analysis of MCMC samples from emcee for spin estimation project
+#
+# MP 04/2015, PK 2015
+import os, sys
+import commands as cmd
+import numpy as np
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt
+plt.rcParams.update({'text.usetex' : True})
+
+from pydoc import help
+from scipy.stats.stats import pearsonr
+from scipy.stats import gaussian_kde
+from scipy import stats
+
+import make_runs as mr
+from utils import *
+import pycbc.pnutils as pnutils
+
+
+# Constants
+quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
+quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
+percentiles_68 = 100*np.array(quantiles_68)
+percentiles_95 = 100*np.array(quantiles_95)
+perc_int_95 = [2.28, 97.72]
+perc_int_99_7 = [0.13, 99.87] # 3 sigma
+
+# Figure settings
+ppi=72.0
+aspect=(5.**0.5 - 1) * 0.5
+size=4.0 * 2# was 6
+figsize=(size,aspect*size)
+plt.rcParams.update({\
+'legend.fontsize':9, \
+'axes.labelsize':11,\
+'font.family':'serif',\
+'font.size':11,\
+'xtick.labelsize':11,\
+'ytick.labelsize':11,\
+'figure.subplot.bottom':0.2,\
+'figure.figsize':figsize, \
+'savefig.dpi': 300.0, \
+'figure.autolayout': True})
+
+# hist(PercentileInterval(s1[0], pc=99.0), 100);
+
+######################################################
+# Function to get the bias in the recovered median 
+# value for different parameters
+######################################################
+def get_bias(\
+  basedir='/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/TN/',\
+  simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
+  M_inj = 3*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, SNR_inj = 60):
+    """
+    Load samples for a given physical configuration, 
+    1. decode the location of the corresponding run. 
+    2. Compute the 90% confidence interval
[…]

### 2015-11-08 emcee_PE/scripts/ProcessSamples.py

@@ -1,0 +1,270 @@
+#!/usr/bin/env python
+# Tools for analysis of MCMC samples from emcee for spin estimation project
+#
+# MP 04/2015, PK 2015
+import os, sys
+import commands as cmd
+import numpy as np
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt
+plt.rcParams.update({'text.usetex' : True})
+
+from pydoc import help
+from scipy.stats.stats import pearsonr
+from scipy.stats import gaussian_kde
+from scipy import stats
+
+import make_runs as mr
+from utils import *
+import pycbc.pnutils as pnutils
+
+
+# Constants
+quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
+quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
+percentiles_68 = 100*np.array(quantiles_68)
+percentiles_95 = 100*np.array(quantiles_95)
+perc_int_95 = [2.28, 97.72]
+perc_int_99_7 = [0.13, 99.87] # 3 sigma
+
+# Figure settings
+ppi=72.0
+aspect=(5.**0.5 - 1) * 0.5
+size=4.0 * 2# was 6
+figsize=(size,aspect*size)
+plt.rcParams.update({\
+'legend.fontsize':9, \
+'axes.labelsize':11,\
+'font.family':'serif',\
+'font.size':11,\
+'xtick.labelsize':11,\
+'ytick.labelsize':11,\
+'figure.subplot.bottom':0.2,\
+'figure.figsize':figsize, \
+'savefig.dpi': 300.0, \
+'figure.autolayout': True})
+
+# hist(PercentileInterval(s1[0], pc=99.0), 100);
+
+######################################################
+# Function to get the bias in the recovered median 
+# value for different parameters
+######################################################
+def get_bias(\
+  basedir='/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/TN/',\
+  simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
+  M_inj = 3*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, SNR_inj = 60):
+    """
+    Load samples for a given physical configuration, 
+    1. decode the location of the corresponding run. 
+    2. Compute the 90% confidence interval
[…]

### 2015-11-08 emcee_PE/src/utils.py

@@ -344,9 +344,10 @@ def load_samples_join(dataDir, SNR, burnin=500, useMaxNRun=True, \
     #{{{
     import glob
     chain, loglike = read_run_part_names(dataDir, SNR, burnin=burnin, \
-                                        useMaxNRun=useMaxNRun, \
-                                        chain_number=chain_number, \
-                                        verbose=verbose )
+            useMaxNRun=useMaxNRun, chain_number=chain_number, verbose=verbose )
+    
+    print "datadir = ", dataDir, " SNR = ", SNR
+    #print chain, loglike
     if np.shape(chain)[0] != np.shape(loglike)[1] or np.shape(chain)[1] != np.shape(loglike)[0]:
       raise RuntimeError("Chain and loglikelihood file not the same length")
     if len(chain) == 0 or len(loglike) == 0:

### 2015-11-09 emcee_PE/scripts/AnalyzeSamples.py
Improved plotting script

@@ -1,12 +1,14 @@
 #!/usr/bin/env python
 # Tools for analysis of MCMC samples from emcee for spin estimation project
-#
-# MP 04/2015, PK 2015
+# Copyright 2015
+# 
+__author__ = "Prayush Kumar <prayush.kumar@ligo.org>"
+
 import os, sys
 import commands as cmd
 import numpy as np
-import matplotlib as mpl
-mpl.use('Agg')
+from matplotlib import mlab, cm, use
+#use('Agg')
 import matplotlib.pyplot as plt
 plt.rcParams.update({'text.usetex' : True})
 
@@ -14,19 +16,35 @@ from pydoc import help
 from scipy.stats.stats import pearsonr
 from scipy.stats import gaussian_kde
 from scipy import stats
+from scipy.interpolate import UnivariateSpline
+from scipy.optimize import minimize_scalar
 
-import make_runs as mr
 from utils import *
 import pycbc.pnutils as pnutils
+from PlotOverlapsNR import make_contour_plot_multrow
+import h5py
 
 
 # Constants
+verbose  = True
+vverbose = True
+
 quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
 quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
 percentiles_68 = 100*np.array(quantiles_68)
 percentiles_95 = 100*np.array(quantiles_95)
 perc_int_95 = [2.28, 97.72]
 perc_int_99_7 = [0.13, 99.87] # 3 sigma
+
+CILevels=[90.0, 68.26895, 95.44997, 99.73002]
+
+######################################################
+# Function to make parameter bias plots
+######################################################
+linestyles = ['-', '--', '-.', '-x', '--o']
+linecolors = ['r', 'g', 'b', 'k', 'm', 'y']
+plotdir = 'plots/'
+gmean = (5**0.5 + 1)/2.
 
 # Figure settings
 ppi=72.0
@@ -38,181 +56,157 @@ plt.rcParams.update({\
 'axes.labelsize':11,\
 'font.family':'serif',\
 'font.size':11,\
-'xtick.labelsize':11,\
-'ytick.labelsize':11,\
+'xtick.labelsize':9,\
+'ytick.labelsize':9,\
 'figure.subplot.bottom':0.2,\
 'figure.figsize':figsize, \
-'savefig.dpi': 300.0, \
[…]

### 2015-11-09 Paper/paper.tex
multi-row contour plots now workcd ../..

@@
 \usepackage{amssymb}
+\usepackage[dvipsnames]{xcolor}
 
@@
 \newcommand{\saul}[1]{\textcolor{Orange}{#1}}
+\newcommand{\prayush}{\textcolor{red!40!black}}
 
@@
 
+\newcommand{\NS}{\mathrm{NS}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Current status:
+
+The Advanced LIGO (aLIGO) observatories began their first observing run ``O1'' in 
+September, 2015. With both detectors a factor of $3-4$ more sensitive than
+their first-generation counterparts, we have access to $\approx 50$ times more 
+spacial volume of the universe at every instant. Within a 
+$3$ month observing run, these detectors would have access to as
+many binary mergers as initial-LIGO detectors did in $\sim 6$ years of 
+coincident operation. Through gradual updates, both of these instruments will
+reach their design sensitivity by $2019$, at which point they will like detect
+around $70$ binary mergers a year~\cite{Abadie:2010cf}.
+
+Coalescing binaries containing a neutron star (NS) and a black hole (BH) are
+one of the most important sources of gravitational waves (GW) with the aLIGO
+and Virgo instruments, which expect to detect $\mathcal{O}(10)$
+of such systems every 
+year~\cite{Abadie:2010cf}. Conventional astronomical methods have, to date, made 
+observations of NSs with masses between $XX-YY M_\odot$ and spin
+$|\vec{\chi}\lesssim 0.05$. On the other hand, indirect observations of BHs 
+have placed the masses of stellar-collapse BHs between $5-35M_\odot$, with 
+spin angular momenta ranging from low to nearly extremal values (i.e., nearly
+as fast as possible - see,
+e.g.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for examples of 
+nearly extremal estimates of BH spins, and see
+Refs.~\cite{McClintock:2013vwa,Reynolds:2013qqa} for recent reviews of
+astrophysical BH spin measurements). \prayush{[Do we want to emphasize on BH
+spins here?]}.
+
+One question that aLIGO is likely to probe is `what is the nature of
+matter at nuclear densities', i.e. what is the equation of state (EoS) of 
+neutron stars. The EoS determines the compactness of NSs, which
+determines how early in the coalescence will the NS be tidally disrupted
+by the gravitational field of the BH, resulting in the formation
+of an accretion disk and/or ejection of unbound matter. This makes NSBH
+mergers are one of the prime targets of the program for joint detection of
+GW and electromagnetic signals. 
+% 
+The difficulty in discerning the matter effects in an NSBH merger GW signal
+lies in distinguishing it reliably from a BHBH merger signal. The earlier the
+NS disrupts, the more orbits there are over which the effect of NS disruption
+is visible. Where exactly does disruption happen is governed both by the 
+mass-ratio of the binary, as well as the spin on the BH. Larger BHs (with
+$q=m_\mathrm{BH}/m_\mathrm{NS}\gtrsim 7$), as well as BHs with large retrograde
+spins ($\chi_\mathrm{BH} \lesssim 0$), tend to swallow the NS whole due to a
+rapid merger (i.e. without it disrupting before the inner-most stable circular
+orbit (ISCO) is reached). On the other hand, binaries with mass-ratios 
+$q=2-5$ with large pro-grade $\chi_\mathrm{BH}$ can disrupt much before 
+merger, leaving their imprint in the GW signal emitted close to $1.7-2$~kHz.
+%
+
+The goal of this paper is to probe the distinguishability of BHBH and NSBH
+mergers, focusing at the effects of the BH mass, BH spin, and NS tidal 
+deformability. We also investigate the accuracy with which aLIGO can measure
+the tidal deformability parameter $\Lambda_\NS\sim(R/M)^5_\NS$. For both of 
+these studies, we will use the zero-detuning high-power noise curve for the
+aLIGO detectors. \prayush{In addition, we also study the effect of tuning aLIGO
+over a narrow frequency range aimed at NSBH mergers.}
+We use the models published in Ref.~\cite{Lackey:2013axa,Pannarale:2015jka} to
+simulate NSBH merger waveforms.
+
+\vspace{2cm}
+
+\textit{\prayush{Current status:}}
+
+Below we explain the outline the different calculations performed in this 
+paper:
 
 \begin{enumerate}
-\item We choose the following values of different NS-BH binary parameters:
+\item We choose the following values of different NSBH binary parameters:
 \begin{enumerate}
@@
 \item $\chi_\mathrm{BH}=\{-0.5, 0, +0.5\}$;
-\item $\Lambda =\{500,1000,2000\}$.
+\item $\Lambda =\{500,1000,2000\}$ ($\Lambda=0\implies \mathrm{BHBH}$).
 \end{enumerate}
 
-\item For each system comprised of all unique combinations of the above 
-parameter choices, we carry out three parameter estimation tests: 
+\item For each system, corresponding to one of all possible unique combinations
+of the above parameter choices, we carry out three parameter estimation tests: 
 \begin{enumerate}
@@
 parameters minus true parameters) vs SNR, as functionals of q/sBH/Lambda, when
-using T signals recovered with N templates.''
+using T signals recovered with N templates.''\newline
+`` another interesting plot to make would be the fitting-factors recovered with
+a discrete (search oriented) template bank of BBH waveforms, for NSBH signals of
+loudness, deformability, spins, mass-ratios, etc. Which parameter plays the 
+dominant role in paramter regions where the bank does not suffice to recover
+the desired threshold of optimal SNR?''
 \end{enumerate}

### 2015-11-09 emcee_PE/scripts/AnalyzeSamples.py

@@ -11,6 +11,7 @@ from matplotlib import mlab, cm, use
 #use('Agg')
 import matplotlib.pyplot as plt
 plt.rcParams.update({'text.usetex' : True})
+from mpl_toolkits.axes_grid1 import ImageGrid
 
 from pydoc import help
 from scipy.stats.stats import pearsonr
@@ -52,12 +53,13 @@ aspect=(5.**0.5 - 1) * 0.5
 size=4.0 * 2# was 6
 figsize=(size,aspect*size)
 plt.rcParams.update({\
-'legend.fontsize':9, \
+'legend.fontsize':11, \
+'text.fontsize':11,\
 'axes.labelsize':11,\
 'font.family':'serif',\
 'font.size':11,\
-'xtick.labelsize':9,\
-'ytick.labelsize':9,\
+'xtick.labelsize':11,\
+'ytick.labelsize':11,\
 'figure.subplot.bottom':0.2,\
 'figure.figsize':figsize, \
 'savefig.dpi': 500.0, \
@@ -72,29 +74,102 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
                 levelspacing=0.25, vmin=None, vmax=None,\
                 xmin=None, xmax=None, ymin=None, ymax=None,\
                 colorbartype='simple', figname='plot.png'):
+  """
+  Function to plot arbitrary numbers of contour plots in a single figure
+  """
   if colorbartype != 'simple':
     raise IOError("Colorbar type %s not supported" % colorbartype)
-  colwidth = 2.8*1.5
+  colwidth = 1.7
   if np.shape(X)[:2] != np.shape(Y)[:2] or np.shape(X)[:2] != np.shape(Z2d)[:2]:
     raise IOError("X, Y and Z arrays have different number of sets to plot")
   
   nrow, ncol, _ = np.shape(X)
   if vverbose: print "Making plot with %d rows, %d cols" % (nrow, ncol)
-  pltid = 1
-  allaxes = []
-  fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=(2.*gmean*colwidth*ncol, colwidth*nrow))
+  pltid = 0
   
+  fig = plt.figure(int(1e7 * np.random.random()), \
+              figsize=((2.*gmean*ncol+0.3)*colwidth, colwidth*nrow))
+  fig.clf()
[…]

### 2015-11-09 emcee_PE/scripts/AnalyzeSamples.py
adding Lambda CI width plots

@@ -20,6 +20,11 @@ from scipy import stats
 from scipy.interpolate import UnivariateSpline
 from scipy.optimize import minimize_scalar
 
+sys.path.append( os.path.dirname(os.path.realpath(__file__)) + '/../src/' )
+sys.path.append( os.path.dirname(os.path.realpath(__file__)) )
+sys.path.append( '/home/prayush/src/UseNRinDA/scripts/' )
+sys.path.append( '/home/prayush/src/UseNRinDA/plotting' )
+
 from utils import *
 import pycbc.pnutils as pnutils
 from PlotOverlapsNR import make_contour_plot_multrow
@@ -28,7 +33,7 @@ import h5py
 
 # Constants
 verbose  = True
-vverbose = True
+vverbose = False
 
 quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
 quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
@@ -88,12 +93,12 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
   pltid = 0
   
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((2.*gmean*ncol+0.3)*colwidth, colwidth*nrow))
+              figsize=((2.*gmean*ncol+0.3)*colwidth, 1.1*colwidth*nrow))
   fig.clf()
   grid = ImageGrid(fig, 111, nrows_ncols=(nrow, ncol), \
             share_all=True,\
             cbar_mode="single", cbar_location="right",\
-            cbar_pad=0.1, cbar_size="5%", \
+            cbar_pad=0.1, cbar_size="4%", \
             aspect=True,\
             add_all=True)
   for idx in range(nrow):
@@ -668,46 +673,60 @@ The bias shown is a fraction of the injected Lambda, and different figures are
 made for different LAmbda values
 """
 
-lambdaBiases = {}
+lambdaBiases, lambdaCIwidths = {}, {}
 for Lambda in Lambdavec:
-  lambdaBiases[Lambda] = {}
+  lambdaBiases[Lambda], lambdaCIwidths[Lambda] = {}, {}
   for snr in SNRvec:
-    lambdaBiases[Lambda][snr] = {}
[…]

### 2015-11-09 emcee_PE/scripts/AnalyzeSamples.py
adding Lambda CI width plots for different CIs

@@ -697,35 +697,36 @@ for Lambda in Lambdavec:
 if plot_LambdaBias:
   plotSNRvec = [30, 50, 90]
   plotCI = 0
-  Xarray, Yarray, Zarray1, Zarray2 = [], [], [], []
-  titles = []
+  for plotCI in [0,1,2,3]:
+    Xarray, Yarray, Zarray1, Zarray2 = [], [], [], []
+    titles = []
+    
+    for Lambda in Lambdavec:
+      Xtmp, Ytmp, Ztmp1, Ztmp2 = [], [], [], []
+      ttmp = []
+      for snr in plotSNRvec:
+        Xtmp.append(np.array(chi2vec))
+        Ytmp.append(mNS * np.array(qvec))
+        Ztmp1.append(lambdaBiases[Lambda][snr][plotCI] * 100)
+        Ztmp2.append(lambdaCIwidths[Lambda][snr][plotCI] * 100)
+        ttmp.append('$\Lambda_\mathrm{NS}=%.1f, \\rho=%.1f$' % (Lambda, snr))
+      Xarray.append(Xtmp)
+      Yarray.append(Ytmp)
+      Zarray1.append(Ztmp1)
+      Zarray2.append(Ztmp2)
+      titles.append(ttmp)
+      
+    make_contour_array(Xarray, Yarray, Zarray2, \
+      xlabel='Black-hole spin', ylabel='Black-hole mass $(M_\odot)$', \
+      xmin=-0.5, xmax=0.75, ymin=2*mNS, ymax=4*mNS, titles=titles, \
+      clabel="$(\Delta\Lambda_\mathrm{NS})^{%.1f \%%}/\Lambda_\mathrm{NS}^\mathrm{Injected}\\times 100$" % CILevels[plotCI], levelspacing=0.5, \
+      figname=os.path.join(plotdir,'TT_LambdaCIWidths%.1f_Lambda_SNR.png' % CILevels[plotCI]))
   
-  for Lambda in Lambdavec:
-    Xtmp, Ytmp, Ztmp1, Ztmp2 = [], [], [], []
-    ttmp = []
-    for snr in plotSNRvec:
-      Xtmp.append(np.array(chi2vec))
-      Ytmp.append(mNS * np.array(qvec))
-      Ztmp1.append(lambdaBiases[Lambda][snr][plotCI] * 100)
[…]

### 2015-11-10 emcee_PE/scripts/AnalyzeSamples.py
rearrange some functions, and added plots from TT runs

@@ -371,6 +371,39 @@ def make_bias_plot(plotparam='Mc', plotquantity='bias', normalize='',\
 # value for different parameters
 ######################################################
 
+def get_bias(\
+  basedir='/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/TN/',\
+  simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
+  M_inj = 3*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, SNR_inj = 60):
+    """
+    Load samples for a given physical configuration, 
+    1. decode the location of the corresponding run. 
+    2. Compute the 90% confidence interval
+    3. Compute the median value of different parameters fromt the posterior.
+    """
+    #{{{
+    test_dir = os.path.join(basedir, simdir)
+    m1_inj, m2_inj = pnutils.mchirp_eta_to_mass1_mass2(M_inj * eta_inj**0.6, eta_inj)
+    params_inj = {'eta' : eta_inj, 'Mtot' : M_inj, 'Mc' : M_inj * eta_inj**0.6,
+                  'chi1' : chi1_inj, 'chi2' : chi2_inj, 
+                  'm1' : m1_inj, 'm2' : m2_inj}
+    
+    match = {}
+    match['samples'] = load_samples_join(test_dir, SNR_inj)
+    
+    bias, CI90, width90 = {}, {}, {}
+    for param in ['eta', 'Mc', 'chi1', 'chi2', 'm1', 'm2', 'Mtot']:
+      S = match['samples'][param]
+      if S == None: return None, [None, None], None
+      param_inj = params_inj[param]
+      if 'chi' in param: bias[param] = (np.median(S) - param_inj) / 1.
+      else: bias[param] = (np.median(S) - param_inj) / param_inj
+      CI90[param] = [np.percentile(S, 5), np.percentile(S, 95)]
+      if 'chi' in param: width90[param] = CI90[param][1] - CI90[param][0]
+      else: width90[param] = (CI90[param][1] - CI90[param][0]) / param_inj    
+    #print "bias = %f, width90 = %f" % (bias, width90)
+    return bias, CI90, width90
+    #}}}
 
 # Function to get the exact parameter bias/whatever given the injection
 # parameters as input
[…]

### 2015-11-10 emcee_PE/scripts/ProcessSamples.py

@@ -51,40 +51,6 @@ plt.rcParams.update({\
 # Function to get the bias in the recovered median 
 # value for different parameters
 ######################################################
-def get_bias(\
-  basedir='/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/TN/',\
-  simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
-  M_inj = 3*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, SNR_inj = 60):
-    """
-    Load samples for a given physical configuration, 
-    1. decode the location of the corresponding run. 
-    2. Compute the 90% confidence interval
-    3. Compute the median value of different parameters fromt the posterior.
-    """
-    #{{{
-    test_dir = os.path.join(basedir, simdir)
-    m1_inj, m2_inj = pnutils.mchirp_eta_to_mass1_mass2(M_inj * eta_inj**0.6, eta_inj)
-    params_inj = {'eta' : eta_inj, 'Mtot' : M_inj, 'Mc' : M_inj * eta_inj**0.6,
-                  'chi1' : chi1_inj, 'chi2' : chi2_inj, 
-                  'm1' : m1_inj, 'm2' : m2_inj}
-    
-    match = {}
-    match['samples'] = load_samples_join(test_dir, SNR_inj)
-    
-    bias, CI90, width90 = {}, {}, {}
-    for param in ['eta', 'Mc', 'chi1', 'chi2', 'm1', 'm2', 'Mtot']:
-      S = match['samples'][param]
-      if S == None: return None, [None, None], None
-      param_inj = params_inj[param]
-      if 'chi' in param: bias[param] = (np.median(S) - param_inj) / 1.
-      else: bias[param] = (np.median(S) - param_inj) / param_inj
-      CI90[param] = [np.percentile(S, 5), np.percentile(S, 95)]
-      if 'chi' in param: width90[param] = CI90[param][1] - CI90[param][0]
-      else: width90[param] = (CI90[param][1] - CI90[param][0]) / param_inj    
-    #print "bias = %f, width90 = %f" % (bias, width90)
-    return bias, CI90, width90
-    #}}}
-
 def calculate_bias(\
[…]

### 2015-11-11 Paper/paper.tex
add figures showing the SNR levels above which we can make meaningful statements on NS Lambda

@@
 \newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}
+\newcommand{\lambdans}{\Lambda_\mathrm{NS}}
+\newcommand{\chibh}{\chi_\mathrm{BH}}
+\newcommand{\mbh}{m_\mathrm{BH}}
+\newcommand{\mns}{m_\mathrm{NS}}
 
@@
 % 
-\begin{figure*}
-\centering
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
-\caption{This figure is similar to the bottom row of
-Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
-consider the NS tidal parameter $\Lambda$ here.}
-\label{fig:TT_NSLambdaBias_vs_SNR_q23}
-\end{figure*}
-% 
-\begin{figure*}
-\centering
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
-\caption{This figure is similar to the bottom row of
-Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
-we consider the NS tidal parameter $\Lambda$ here.}
-\label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
-\end{figure*}
-% 
 % 
@@
 
-Could have subsections for aLIGO and ET ?
+\prayush{Could have subsections for different frequency-optimized aLIGO noise curves ?}
+
+\hspace{5mm}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % 
+% \begin{figure*}
+% \centering
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
+% \caption{This figure is similar to the bottom row of
+% Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
+% consider the NS tidal parameter $\Lambda$ here.}
+% \label{fig:TT_NSLambdaBias_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
+% \caption{This figure is similar to the bottom row of
+% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
+% we consider the NS tidal parameter $\Lambda$ here.}
+% \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=2\columnwidth]{plots-TT/TT_LambdaBiases_CI90_0_Lambda_SNR.png}
+% \caption{This figure shows.}
+% \label{fig:TT_LambdaBiases_CI90_0_Lambda_SNR}
+% \end{figure*}
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=2.1\columnwidth]{plots-TT/TT_LambdaCIWidths68_3_Lambda_SNR.png}\\
+\includegraphics[width=2.1\columnwidth]{plots-TT/TT_LambdaCIWidths90_0_Lambda_SNR.png}
+\caption{This figure shows.}
+\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
+\end{figure*}
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
+\caption{This figure shows.}
+\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
+\end{figure*}
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
+\caption{This figure shows.}
+\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
+\end{figure*}
+
 Having shown in the previous section that we begin to care for NS tidal effects
@@
 \begin{enumerate}
+\item Where in the parameter space can we actually make a statement about 
+$\Lambda_\mathrm{NS}$ ?
 \item What is the reduction in the bias of the maximum likelihood parameters
@@
 
-\red{Could have subsections for aLIGO and ET ?}
+\subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}
+
+With the enhanced sensitivity of aLIGO, we are likely to be able to measure or
+constrain deviations in the observed GW signals, especially close to binary
+merger, from their expected form under the point-particle approximation.
+The tidal disruption of the NS close to merger, and outside of the 
+inner-most circular orbit, will drastically change the morphology of the emitted
+GW signal, the measurement of which will allow us to constrain the equation 
+of state (EoS) of nuclear matter.
+The earlier in the orbit the NS disrupts, the sharper is the drop in the emitted
+GW signal for frequencies above the disruption frequency. In this section we
+study the ability of aLIGO to constrain the 
+$\lambdans\sim \left(R/M\right)_\mathrm{NS}^5$ parameter, which
+the combination that parametrizes the deviation in the emitted GW signal.
+Past studies~\cite{}
+have studied a limited sample of systems with distinguishability criterion that
+are typically applicable in the very high signal-to-noise ratio cases. 
+In what follows, we will use stochastic parameter estimation algorithms to show
+aLIGO's capability of extracting EoS information out from GW signals with 
+more realistic (likely) SNRs.
+
+As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
+fixed at $1.35M_\odot$. We allow BH spin to take the values 
+$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
+parameter combinations, we inject a tidal-corrected waveform using the 
+Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
+posterior probability distribution for different binary parameters with (a) BHBH
+templates, and (b) NSBH templates. We will discuss the results of (b) here.
 

### 2015-11-11 emcee_PE/scripts/AnalyzeSamples.py

@@ -33,7 +33,8 @@ import h5py
 
 # Constants
 verbose  = True
-vverbose = False
+vverbose = True
+debug    = False
 
 quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
 quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
@@ -67,7 +68,7 @@ plt.rcParams.update({\
 'ytick.labelsize':11,\
 'figure.subplot.bottom':0.2,\
 'figure.figsize':figsize, \
-'savefig.dpi': 500.0, \
+'savefig.dpi': 300.0, \
 'figure.autolayout': True})
 
 # hist(PercentileInterval(s1[0], pc=99.0), 100);
@@ -76,7 +77,7 @@ plt.rcParams.update({\
 # Plotting functions
 def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
                 title='', titles=[], \
-                levelspacing=0.25, vmin=None, vmax=None,\
+                levelspacing=0.25, vmin=None, vmax=None, cmap=cm.rainbow,\
                 xmin=None, xmax=None, ymin=None, ymax=None,\
                 colorbartype='simple', figname='plot.png'):
   """
@@ -88,17 +89,30 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
   if np.shape(X)[:2] != np.shape(Y)[:2] or np.shape(X)[:2] != np.shape(Z2d)[:2]:
     raise IOError("X, Y and Z arrays have different number of sets to plot")
   
+  plt.rcParams.update({\
+    'legend.fontsize':16, \
+    'text.fontsize':16,\
+    'axes.labelsize':16,\
+    'font.family':'serif',\
+    'font.size':16,\
+    'xtick.labelsize':16,\
+    'ytick.labelsize':16,\
+    'figure.subplot.bottom':0.2,\
+    'figure.figsize':figsize, \
+    'savefig.dpi': 300.0, \
+    'figure.autolayout': True})
+
   nrow, ncol, _ = np.shape(X)
   if vverbose: print "Making plot with %d rows, %d cols" % (nrow, ncol)
   pltid = 0
   
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((2.*gmean*ncol+0.3)*colwidth, 1.1*colwidth*nrow))
[…]

### 2015-11-13 Paper/paper.tex
adding chirp mass bias plots

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{figure*}
+% \centering 
+% % \textbf{BH-BH Injection; BH-BH Templates}
+% % \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
+% \caption{These figures show the (fractional) difference between the chirp mass 
+% value corresponding to the median of its posterior probability distribution,
+% and the actual injected chirp mass,
+% as a function of the signal-to-noise-ratio (SNR) of the injected signal.
+% % In the top row, NS matter effects on the inspiral and merger waveform
+% % are neithed included in templates nor in the injected signal. 
+% In the top row, NS matter effects (inspiral and merger) are included in the
+% injections but not in templates. This corresponds to the current 
+% plan for aLIGO parameter estimation studies. The bottom row shows the effect
+% of additionally including NS matter effects in the templates.
+% % 
+% In each row, the left and right panel correspond to 
+% $q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
+% In each panel, the color of each curve corresponds to the value of NS 
+% deformability parameter $(\Lambda=\{500,1000,2000\})$, 
+% and line-style corresponds to the value of
+% the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
+% }
+% \label{fig:TNT_chirpMassBias_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering 
+% % \textbf{BH-BH Injection; BH-BH Templates}
+% % \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
+% \caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
+% with the difference that here we show the width of the $90\%$ confidence 
+% interval (recovered) for chirp mass, as a function of the injected SNR. 
+% Note that the confidence interval's width is normalized by the injected
+% parameter value.
+% }
+% \label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
+% \end{figure*}
+% % 
 \begin{figure*}
 \centering 
-% \textbf{BH-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
+\includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}\\ 
 \caption{These figures show the (fractional) difference between the chirp mass 
@@
 \centering
-% \textbf{BH-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
+\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_MchirpCIWidths90_0_Lambda_SNR.pdf}
 \caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}

### 2015-11-13 emcee_PE/scripts/AnalyzeSamples-TT.py

@@ -76,10 +76,10 @@ plt.rcParams.update({\
 
 # Plotting functions
 def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
-        title='', titles=[], \
-        levelspacing=0.25, vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f',\
-        xmin=None, xmax=None, ymin=None, ymax=None,\
-        colorbartype='simple', figname='plot.png'):
+                title='', titles=[], \
+                levelspacing=0.25, vmin=None, vmax=None, cmap=cm.rainbow,\
+                xmin=None, xmax=None, ymin=None, ymax=None,\
+                colorbartype='simple', figname='plot.png'):
   """
   Function to plot arbitrary numbers of contour plots in a single figure
   """
@@ -107,7 +107,7 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
   pltid = 0
   
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((2.1*gmean*ncol+1.25)*colwidth, 1.2*colwidth*nrow))
+              figsize=((2.1*gmean*ncol+0.3)*colwidth, 1.2*colwidth*nrow))
   fig.clf()
   grid = ImageGrid(fig, 111, nrows_ncols=(nrow, ncol), \
             share_all=True,\
@@ -115,22 +115,6 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
             cbar_pad=0.05, cbar_size="2%", \
             aspect=True,\
             add_all=True)
-  # Find the maximum Z-value of all subplots
-  #VMIN, VMAX = np.inf, -np.inf
-  #for idx in range(nrow):
-  #  for jdx in range(ncol):
-  #    VMIN = min(VMIN, Z2d[idx][jdx].min())
-  #    VMAX = max(VMAX, Z2d[idx][jdx].max())
-  VMINmax, VMAXmin = np.array([]), np.array([])
-  for idx in range(nrow):
-    for jdx in range(ncol):
-      VMINmax = np.append( VMINmax, Z2d[idx][jdx].min() )
-      VMAXmin = np.append( VMAXmin, Z2d[idx][jdx].max() )
-  VMIN, VMAX = VMINmax.min(), VMAXmin.max()
-  print "VMIN = %e, VMAX = %e" % (VMIN, VMAX)
[…]

### 2015-11-13 Paper/paper.tex
added figures showing TN/NN results

@@
 % % 
+% 
+% \begin{figure*}[!t]
+% \centering    
+% \textbf{NS-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
+% \textbf{BH-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
+% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+% with the difference that we consider the systematic bias in the dimensionless
+% mass-ratio $\eta$ here. In the bottom row we show control runs where both
+% signal and template waveforms lack matter effects.}
+% \label{fig:TNT_EtaBias_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering    
+% \textbf{NS-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
+% \textbf{BH-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
+% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+%   with the difference that we consider the confidence intervals in the recovery
+%   of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
+%   Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
+%   confidence interval is shown here.}
+% \label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
+% \end{figure*}
+% \begin{figure*}
+% \centering    
+% \textbf{NS-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
+% \textbf{BH-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
+% \caption{The top two rows in this figure are
+% similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
+% with the differences that we consider the systematic bias in the dimensionless
+% spin on the BH here, and that the differences shown are absolute and not
+% fractions of the injected spin. In the bottom row, we show a control run where
+% both injected and template waveforms correspond to binary black holes.}
+% \label{fig:TNT_BHspinBias_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering    
+% \textbf{BH-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; BH-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
+% \caption{The top two rows in the figure are similar to
+% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+% with the difference that we consider the confidence intervals in the recovery
+% of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
+% As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
+% width of the confidence interval without normalizing with the injected spin.
+% The bottom row show a control run where matter effects are ignored both 
+% in the injected signal and the template waveforms.}
+% \label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
+% \end{figure*}
+
+% #################
 \begin{figure*}
 \centering 
-\includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}\\ 
+\includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}
 \caption{These figures show the (fractional) difference between the chirp mass 
@@
 \centering    
-\textbf{NS-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
-\textbf{BH-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
+\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaBiases_Lambda_SNR.pdf}
 \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
@@
 \centering    
-\textbf{NS-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
-\textbf{BH-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
-\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
+\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaCIWidths68_3_Lambda_SNR.pdf}
+\caption{This figure is similar to
+Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
   with the difference that we consider the confidence intervals in the recovery
@@
 \centering
-\textbf{NS-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
-\textbf{BH-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
+\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHBiases_CI90_0_Lambda_SNR.pdf}
 \caption{The top two rows in this figure are
@@
 \centering    
-\textbf{BH-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; BH-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
-\textbf{NS-BH Injection; NS-BH Templates}
-\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
+\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHCIWidths90_0_Lambda_SNR.pdf}
 \caption{The top two rows in the figure are similar to

### 2015-11-13 emcee_PE/scripts/AnalyzeSamples.py

@@ -748,10 +748,10 @@ PLUS, the first 2 columns in each row of the dataset are:
     converted to LogLikelihood as LogL = 1/2 SNR^2 MATCH^2
 '''
 
-plot_MchirpBias = True
+plot_MchirpBias = False
 plot_EtaBias    = False
 plot_QBias      = False
-plot_sBHBias    = False
+plot_sBHBias    = True
 plot_mBHBias    = False
 plot_LambdaBias = False
 
@@ -768,7 +768,6 @@ print "MAKING PLOTS NOW.."
 ######################################################
 print " MAKING Plots with/for CHIRP MASS "
 ######################################################
-
 if plot_MchirpBias:
   print \
 """
@@ -871,11 +870,204 @@ made for different LAmbda values
 ######################################################
 print " MAKING Plots with/for MASS RATIO "
 ######################################################
+if plot_EtaBias:
+  print \
+"""
+Plotting the fractional bias in MASS-RATIO recovery at different SNR levels, as 
+a function of the BH mass and spin. In addition also plotting the width of the
+confidence interval for chirp-mass, i.e the statistical uncertainty around the
+maximum likelihood value.
+
+The bias shown is a fraction of the injected Lambda, and different figures are
+made for different LAmbda values
+"""
+  etaBiases, etaCIwidths = {}, {}
+  for Lambda in Lambdavec:
+    etaBiases[Lambda], etaCIwidths[Lambda] = {}, {}
+    for snr in SNRvec:
+      etaBiases[Lambda][snr], etaCIwidths[Lambda][snr] = {}, {}
+      for CI in range( len(CILevels) ):
+        etaBiases[Lambda][snr][CI] = np.zeros( (len(qvec), len(chi2vec)) )
+        etaCIwidths[Lambda][snr][CI] = np.zeros( (len(qvec), len(chi2vec)) )
+        for i, q in enumerate(qvec):
+          for j, chiBH in enumerate(chi2vec):
+            if vverbose:
+              print "getting bias in eta for q=%f, chiBh=%f, Lambda=%f at SNR = %f" %\
[…]

### 2015-11-13 Paper/paper.tex
fig labels added to s4

@@
 \newcommand{\mns}{m_\mathrm{NS}}
+\newcommand{\mchirp}{\mathcal{M}_c}
 
@@
 \begin{abstract}
-Abstract
+We attempt to answer the following questions: 
+\begin{enumerate}
+ \item Will ignoring the effect of NS deformability during inspiral and NS disruption during merger (in an NSBH binary) change the efficiency of our searches?
+ \item Will not modeling NS disruption during NSBH mergers change the systematic and 
+ statistical biases in the recovery of binary parameters?
+ \item If we use templates which models NS disruption, what information can we recover about
+ the $\Lambda_\mathrm{BH}\sim (R/M)^5_\mathrm{NS}$ parameter?
+\end{enumerate}
 \end{abstract}
@@
 \includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}
-\caption{These figures show the (fractional) difference between the chirp mass 
-value corresponding to the median of its posterior probability distribution,
-and the actual injected chirp mass,
-as a function of the signal-to-noise-ratio (SNR) of the injected signal.
-% In the top row, NS matter effects on the inspiral and merger waveform
-% are neithed included in templates nor in the injected signal. 
-In the top row, NS matter effects (inspiral and merger) are included in the
-injections but not in templates. This corresponds to the current 
-plan for aLIGO parameter estimation studies. The bottom row shows the effect
-of additionally including NS matter effects in the templates.
-% 
-In each row, the left and right panel correspond to 
-$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
-In each panel, the color of each curve corresponds to the value of NS 
-deformability parameter $(\Lambda=\{500,1000,2000\})$, 
-and line-style corresponds to the value of
-the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
+\caption{Systematic fractional biases in recovered chirp-mass $\mchirp$, as a
+fraction of the injected (true) value, for NSBH binaries, shown here as a 
+function of the mass and spin of the BH (each panel), $\lambdans$ (across 
+rows), and injection's SNR $\rho$ (across columns). Tidal effects during
+inspiral and the effects of NS disruption close to merger are modelled in the
+signal, but \textit{not} in the templates. 
 }
@@
 \includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_MchirpCIWidths90_0_Lambda_SNR.pdf}
-\caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
-with the difference that here we show the width of the $90\%$ confidence 
-interval (recovered) for chirp mass, as a function of the injected SNR. 
-Note that the confidence interval's width is normalized by the injected
-parameter value.
+\caption{$90\%$ confidence intervals for the recovered chirp-mass $\mchirp$, as a
+fraction of the injected (true) value, for NSBH binaries, shown here as a 
+function of the mass and spin of the BH (each panel), $\lambdans$ (across 
+rows), and injection's SNR $\rho$ (across columns). Tidal effects during
+inspiral and the effects of NS disruption close to merger are modelled in the
+signal, but \textit{not} in the templates. 
 }
@@
 \includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaBiases_Lambda_SNR.pdf}
-\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
-with the difference that we consider the systematic bias in the dimensionless
-mass-ratio $\eta$ here. In the bottom row we show control runs where both
-signal and template waveforms lack matter effects.}
+\caption{Systematic fractional biases in the recovered dimensionless mass-ratio
+$\eta = m_1 m_2 / (m_1+m_2)^2$, as a
+fraction of the injected (true) value, for NSBH binaries, shown here as a 
+function of the mass and spin of the BH (each panel), $\lambdans$ (across 
+rows), and injection's SNR $\rho$ (across columns). Tidal effects during
+inspiral and the effects of NS disruption close to merger are modelled in the
+signal, but \textit{not} in the templates. }
 \label{fig:TNT_EtaBias_vs_SNR_q23}
@@
 \includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaCIWidths68_3_Lambda_SNR.pdf}
-\caption{This figure is similar to
-Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
-  with the difference that we consider the confidence intervals in the recovery
-  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
-  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
-  confidence interval is shown here.}
+\caption{$90\%$ confidence intervals for the recovered dimensionless mass-ratio
+$\eta = m_1 m_2 / (m_1+m_2)^2$, as a
+fraction of the injected (true) value, for NSBH binaries, shown here as a 
+function of the mass and spin of the BH (each panel), $\lambdans$ (across 
+rows), and injection's SNR $\rho$ (across columns). Tidal effects during
+inspiral and the effects of NS disruption close to merger are modelled in the
+signal, but \textit{not} in the templates. }
 \label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
@@
 \includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHBiases_CI90_0_Lambda_SNR.pdf}
-\caption{The top two rows in this figure are
-similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
-with the differences that we consider the systematic bias in the dimensionless
-spin on the BH here, and that the differences shown are absolute and not
-fractions of the injected spin. In the bottom row, we show a control run where
-both injected and template waveforms correspond to binary black holes.}
+\caption{Systematic fractional biases in the recovered value of the BH spin $\chibh$,
+as a fraction of the injected (true) value, for NSBH binaries, shown here as a
+function of the mass and spin of the BH (each panel), $\lambdans$ (across 
+rows), and injection's SNR $\rho$ (across columns). Tidal effects during
+inspiral and the effects of NS disruption close to merger are modelled in the
+signal, but \textit{not} in the templates. }
 \label{fig:TNT_BHspinBias_vs_SNR_q23}
@@
 \includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHCIWidths90_0_Lambda_SNR.pdf}
-\caption{The top two rows in the figure are similar to
-Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
-with the difference that we consider the confidence intervals in the recovery
-of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
-As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
-width of the confidence interval without normalizing with the injected spin.
-The bottom row show a control run where matter effects are ignored both 
-in the injected signal and the template waveforms.}
+\caption{$90\%$ confidence intervals for the recovered value of the BH spin $\chibh$,
+as a fraction of the injected (true) value, for NSBH binaries, shown here as a 
+function of the mass and spin of the BH (each panel), $\lambdans$ (across 
+rows), and injection's SNR $\rho$ (across columns). Tidal effects during
+inspiral and the effects of NS disruption close to merger are modelled in the
+signal, but \textit{not} in the templates.}
 \label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}

### 2015-11-13 emcee_PE/scripts/AnalyzeSamples.py

@@ -33,7 +33,7 @@ import h5py
 
 # Constants
 verbose  = True
-vverbose = True
+vverbose = False
 debug    = False
 
 quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
@@ -751,10 +751,11 @@ PLUS, the first 2 columns in each row of the dataset are:
 plot_MchirpBias = False
 plot_EtaBias    = False
 plot_QBias      = False
-plot_sBHBias    = True
+plot_sBHBias    = False
 plot_mBHBias    = False
 plot_LambdaBias = False
 
+# Plots showing Lambda recovery
 plot_SNRcrit    = False
 plot_LambdaCrit = False
 
@@ -967,6 +968,103 @@ made for different LAmbda values
       levelspacing=0.4,  cfmt='%.1f',\
       figname=os.path.join(plotdir,simstring+('EtaBiases_CI%.1f_Lambda_SNR' % CILevels[plotCI]).replace('.','_')+'.pdf'))
 
+if plot_QBias:
+  print \
+"""
+Plotting the fractional bias in MASS-RATIO recovery at different SNR levels, as 
+a function of the BH mass and spin. In addition also plotting the width of the
+confidence interval for chirp-mass, i.e the statistical uncertainty around the
+maximum likelihood value.
+
+The bias shown is a fraction of the injected Lambda, and different figures are
+made for different LAmbda values
+"""
+  etaBiases, etaCIwidths = {}, {}
+  for Lambda in Lambdavec:
+    etaBiases[Lambda], etaCIwidths[Lambda] = {}, {}
+    for snr in SNRvec:
+      etaBiases[Lambda][snr], etaCIwidths[Lambda][snr] = {}, {}
+      for CI in range( len(CILevels) ):
+        etaBiases[Lambda][snr][CI] = np.zeros( (len(qvec), len(chi2vec)) )
+        etaCIwidths[Lambda][snr][CI] = np.zeros( (len(qvec), len(chi2vec)) )
+        for i, q in enumerate(qvec):
+          for j, chiBH in enumerate(chi2vec):
+            if vverbose:
+              print "getting bias in eta for q=%f, chiBh=%f, Lambda=%f at SNR = %f" %\
+                    (q, chiBH, Lambda, snr)
+            try:
[…]

### 2015-11-13 emcee_PE/scripts/ProcessSamples.py

@@ -51,7 +51,7 @@ plt.rcParams.update({\
 # Function to get the bias in the recovered median 
 # value for different parameters
 ######################################################
-def calculate_bias(\
+def calculate_bias_withoutQ(\
   basedir='/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/TN/',\
   simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
   M_inj = 3*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, Lambda_inj=0, SNR_inj = 60,\
@@ -133,6 +133,88 @@ def calculate_bias(\
     return summary_data
     #}}}
 
+def calculate_bias(\
+  basedir='/home/prayush/projects/nsbh/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/TN/',\
+  simdir='TN_q2.00_mNS1.35_chiBH0.50_Lambda500.0_SNR60.0/NW100_NS6000/',\
+  M_inj = 3*1.35, eta_inj = 2./9., chi1_inj=0, chi2_inj=0.5, Lambda_inj=0, SNR_inj = 60,\
+  biastype='fractional', recover_tidal=False, \
+  confidence_levels=[90.0, 68.26895, 95.44997, 99.73002]):
+    """
+    Load samples for a given physical configuration, 
+    1. decode the location of the corresponding run. 
+    2. Compute 68%, 90%,.. confidence interval for different sampled parameters
+    3. Compute the median value of different parameters fromt the posterior.
+    
+    We plan to store 5 quantities for each parameter k:
+    k+0: Median recovered parameter value from posterior samples
+    k+1: Bias = (X(median) - X(inj))/X(inj)
+    k+2: Confidence level lower bound = X(confidence_level_low)
+    k+3: Confidence level upper bound = X(confidence_level_up)
+    k+4: Confidence interval = (X(confidence_level_up) - X(confidence_level_low))/X(inj)
+    
+    PLUS
+    0 : Confidence interval probability
+    1 : max LogLikelihood value
+    
+    """
+    #{{{
+    test_dir = os.path.join(basedir, simdir)
[…]

### 2015-11-13 emcee_PE/src/utils.py

@@ -355,7 +355,7 @@ def load_samples_join(dataDir, SNR, burnin=500, useMaxNRun=True, \
         print "Empty lists gotten for chain and loglike for %s" % dataDir
       return {'eta' : None, 'Mc' : None, 'chieff' : None, 'chiPN' : None,
               'chi1' : None, 'chi2' : None, 'Mtot' : None,
-              'm1' : None, 'm2' : None, 'match' : None}
+              'm1' : None, 'm2' : None, 'q' : None, 'match' : None}
       #raise RuntimeError("Empty lists gotten for chain and loglike")
     #
     # Here onwards code remains the same as for load_samples
@@ -387,7 +387,7 @@ def load_samples_join(dataDir, SNR, burnin=500, useMaxNRun=True, \
     
     return {'eta' : etaval, 'Mc' : Mcval, 'chieff' : chieffval, 'chiPN' : chiPNval, 
             'chi1' : chi1val, 'chi2' : chi2val, 'Mtot' : Mval, 
-            'm1' : m1val, 'm2' : m2val, 'match' : matchval}
+            'm1' : m1val, 'm2' : m2val, 'q' : qval, 'match' : matchval}
     #}}}
 
 

### 2015-11-13 emcee_PE/src/emcee_match_aligned.py
redirecting corner from triangle to corner

@@ -23,7 +23,7 @@ from matplotlib import rc
 rc('text', usetex=True)
 import matplotlib.pyplot as pl
 from matplotlib.ticker import MaxNLocator
-import triangle
+import triangle, corner
 from optparse import OptionParser
 from scipy.stats import norm
 import scipy.interpolate as ip
@@ -706,17 +706,17 @@ quantiles=quantiles_95
 samples_combined = np.array([qval, Mval, chi1val, chi2val]).T
 print np.shape(samples_combined)
 print samples_combined
-fig = triangle.corner(samples_combined, labels=["$q$", "$M$", "$\chi_1$", "$\chi_2$"], truths=[q_true, M_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
+fig = corner.corner(samples_combined, labels=["$q$", "$M$", "$\chi_1$", "$\chi_2$"], truths=[q_true, M_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
 fig.savefig("corner_q_M_chi1_chi2.png")
 pl.close(fig)
 
 samples_combined2 = np.array([etaval, Mcval, chi1val, chi2val]).T
-fig = triangle.corner(samples_combined2, labels=["$\eta$", "$\mathcal{M}$", "$\chi_1$", "$\chi_2$"], truths=[eta_true, Mc_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
+fig = corner.corner(samples_combined2, labels=["$\eta$", "$\mathcal{M}$", "$\chi_1$", "$\chi_2$"], truths=[eta_true, Mc_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
 fig.savefig("corner_eta_Mc_chi1_chi2.png")
 pl.close(fig)
 
 samples_combined3 = np.array([m1val, m2val, chi1val, chi2val]).T
[…]

### 2015-11-13 emcee_PE/src/emcee_match_tidal.py

@@ -23,7 +23,7 @@ from matplotlib import rc
 rc('text', usetex=True)
 import matplotlib.pyplot as pl
 from matplotlib.ticker import MaxNLocator
-import triangle
+import triangle, corner
 from optparse import OptionParser
 from scipy.stats import norm
 import scipy.interpolate as ip
@@ -687,17 +687,17 @@ quantiles=quantiles_95
 samples_combined = np.array([qval, Mval, chi2val, Lambdaval]).T
 print np.shape(samples_combined)
 print samples_combined
-fig = triangle.corner(samples_combined, labels=["$q$", "$M$", "$\chi_2$", "$\Lambda$"], truths=[q_true, M_true, chi2_true, Lambda_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
+fig = corner.corner(samples_combined, labels=["$q$", "$M$", "$\chi_2$", "$\Lambda$"], truths=[q_true, M_true, chi2_true, Lambda_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
 fig.savefig("corner_q_M_chi2_Lambda.png")
 pl.close(fig)
 
 samples_combined2 = np.array([etaval, Mcval, chi2val, Lambdaval]).T
-fig = triangle.corner(samples_combined2, labels=["$\eta$", "$\mathcal{M}$", "$\chi_2$", "$\Lambda$"], truths=[eta_true, Mc_true, chi2_true, Lambda_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
+fig = corner.corner(samples_combined2, labels=["$\eta$", "$\mathcal{M}$", "$\chi_2$", "$\Lambda$"], truths=[eta_true, Mc_true, chi2_true, Lambda_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
 fig.savefig("corner_eta_Mc_chi2_Lambda.png")
 pl.close(fig)
 
 samples_combined3 = np.array([m1val, m2val, chi2val, Lambdaval]).T
[…]

### 2015-11-16 emcee_PE/src/emcee_match_aligned.py
change the match cutoff on samples, required for resuming chains

@@ -454,10 +454,11 @@ if not post_process:
       # use all good samples from last 50 iterations and extract as many as we need to restart (nwalkers)
       k = min(50, len(loglike[:,0]))
       print "Using good samples from last %d iterations" %(k)
-      match_cut = 1 - CR_99p9pc_threshold
+      match_cut = 1 - CR_99pc_threshold
       loglike_ok = loglike[-k:] # use samples from last k iterations
       print "len(loglike_ok) = ", np.shape(loglike_ok)
       matches_ok = np.sqrt(2*loglike_ok)/SNR
+      if not np.any( matches_ok > match_cut ): match_cut = 0.99 # CHECKME: Arbitrary condition to enable resume 
       print "len(matches_ok) = ", len(matches_ok)
       sel = np.isfinite(loglike_ok) & (matches_ok > match_cut)
       print "k = ", k, "\n shape and No of True elements in sel = ", np.shape(sel), np.count_nonzero(sel)

### 2015-11-16 emcee_PE/src/emcee_match_aligned.py
changing back default match cut fo 99.9%

@@ -454,7 +454,7 @@ if not post_process:
       # use all good samples from last 50 iterations and extract as many as we need to restart (nwalkers)
       k = min(50, len(loglike[:,0]))
       print "Using good samples from last %d iterations" %(k)
-      match_cut = 1 - CR_99pc_threshold
+      match_cut = 1 - CR_99p9pc_threshold
       loglike_ok = loglike[-k:] # use samples from last k iterations
       print "len(loglike_ok) = ", np.shape(loglike_ok)
       matches_ok = np.sqrt(2*loglike_ok)/SNR

### 2015-11-16 emcee_PE/src/emcee_match_aligned.py
adding more checks to enable resume

@@ -458,9 +458,14 @@ if not post_process:
       loglike_ok = loglike[-k:] # use samples from last k iterations
       print "len(loglike_ok) = ", np.shape(loglike_ok)
       matches_ok = np.sqrt(2*loglike_ok)/SNR
-      if not np.any( matches_ok > match_cut ): match_cut = 0.99 # CHECKME: Arbitrary condition to enable resume 
+      #if not np.any( matches_ok > match_cut ): match_cut = 0.99 # CHECKME: Arbitrary condition to enable resume 
       print "len(matches_ok) = ", len(matches_ok)
       sel = np.isfinite(loglike_ok) & (matches_ok > match_cut)
+      # CHECKME: Arbitrary condition to enable resume 
+      while np.count_nonzero(sel) < nwalkers:
+        match_cut -= 0.001
+        print " trying match cut = ", match_cut
+        sel = np.isfinite(loglike_ok) & (matches_ok > match_cut)
       print "k = ", k, "\n shape and No of True elements in sel = ", np.shape(sel), np.count_nonzero(sel)
       print "shape of map(lambda i: chain[:,-k:,i].T[sel], range(ndim)) = ", np.shape(map(lambda i: chain[:,-k:,i].T[sel], range(ndim)))
       print "shape of chain[:,-k:,i] = ", np.shape(chain[:,-k:,0])

### 2015-11-16 emcee_PE/src/emcee_match_tidal.py
change cutoff match for choosing points while auto resuming

@@ -449,6 +449,11 @@ if not post_process:
       matches_ok = np.sqrt(2*loglike_ok)/SNR
       print "len(matches_ok) = ", len(matches_ok)
       sel = np.isfinite(loglike_ok) & (matches_ok > match_cut)
+      # CHECKME: Arbitrary condition to enable resume 
+      while np.count_nonzero(sel) < nwalkers:
+        match_cut -= 0.001
+        print " trying match cut = ", match_cut
+        sel = np.isfinite(loglike_ok) & (matches_ok > match_cut)
       print "k = ", k, "\n shape and No of True elements in sel = ", np.shape(sel), np.count_nonzero(sel)
       print "shape of map(lambda i: chain[:,-k:,i].T[sel], range(ndim)) = ", np.shape(map(lambda i: chain[:,-k:,i].T[sel], range(ndim)))
       print "shape of chain[:,-k:,i] = ", np.shape(chain[:,-k:,0])

### 2015-11-17 Paper/paper.tex
text to bank effectualness section

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-See what is the loss in SNR if we recover NSBH signals with a template bank of
-BH-BH waveforms? It is very likely that the answer would be: $<1\%$. But it
-is very interesting to show that for spinning BHs!
+
+Near future detection searches with aLIGO are poised to search for NSBH
+signals using BHBH templates. In this section, we examine the effect of ignoring
+the effects of tidal deformation and disruption of the NS during inspiral and 
+merger phases of search templates. We construct a template bank $\mathcal{B}$ 
+restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
+$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
+using a PN-based geometric lattice placement method~\cite{Harry:2013tca}.
+
+Before investigating its SNR recovery against tidal NSBH signals, we test the 
+effectualness of this bank to point-particle NSBH signals,
+in order to make sure that the placement method does not leave any region of
+the binary parameter-space under-covered. We simulate $50,000$ signals with 
+component-mass and BH spin values
+sampled uniformly over their allowed ranges, and filter each of them against
+the entire bank. For both signals and templates, we use the frequency-domain 
+reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
+implemented within the LIGO algorithms Library~\cite{lal}. We record the 
+fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
+signals, which measures the fraction of optimal SNR that the bank is able
+to recover. Fig.~\ref{} shows these FFs as a function of the injected
+BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
+values. From these figures, we conclude that our bank is effectual as intended.
+
+Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
+detection searches, we perform a similar calculation as above, with the
+difference that the signals are now modelled using the model of tidal 
+corrections to IMR amplitude and phasing published in 
+Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.
 

### 2015-11-17 emcee_PE/src/TidalWaveforms.py

@@ -1,9 +1,10 @@
-#!/usr/bin/env python
+#!/usr/bin/nv python
 
 import os
 import sys
 #import matplotlib.pyplot as plt
 import numpy as np
+from scipy.optimize import minimize_scalar
 
 import lal
 from pycbc.waveform import *
@@ -154,3 +155,248 @@ class tidalWavs():
         return hp, hc
     #}}}
 
+
+def windowing_function(f, f0, d, sgn=+1):
+  """
+  Returns waveform windowing_function, for the input frequencies
+  """
+  return 0.5*(1. + sgn*np.tanh(4.*(f - f0)/d))
+
+class PNcoeffs():
+  def __init__(verbose=True):
+    """
+    This class is for storing all PN coefficients, as well for solving for 
+    orbital quantities.
+    """
+    self.verbose = verbose
+    self.initialize_coefficients_fd()
+    return
+  #
+  def initialize_coefficients_fd(self):
+    return
+  #
+  def GetWaveformPhase(self):
+    return
+  #
+  def GetWaveformAmplitude(self):
+    return
+
+class tidalWavsFP():
+    #{{{
+    def __init__(self, approx='IMRPhenomC',\
+          massBH = -1, spinBH = 2,\
+          massNS_B = -1, massNS = -1, radiusNS = 0, \
+          f_lower = 15., delta_f = 1./256, \
+          verbose=True):
+      """
+      This class returns the waveform generated using the model in P1.
+      
+      References:
+      P1 : arXiv:1509.00512
+      P2 : arXiv:1311.5931
+      """
+      if massBH <= 0 or massNS <= 0 or spinBH > 1 or spinBH < -1 or \
+          massNS_B < 0 or radiusNS <= 0:
+        raise IOError("Provide correct input to waveform generation please")
+        
+      self.massBH = massBH
+      self.spinBH = spinBH
+      
+      self.massNS = massNS
+      # The following two have to be determined from the EoS ..
+      self.radiusNS= radiusNS
+      self.massNS_B = massNS_B
+      
+      self.mtotal = self.massNS + self.massBH
+      self.eta = self.massNS * self.massBH / self.mtotal**2
+      
+      self.f_lower = f_lower
+      self.delta_f = delta_f
+      
+      self.verbose = verbose
+      self.approx  = approx
+      #
+      # Initialize constants
+      return
[…]

### 2015-11-18 emcee_PE/src/TidalWaveforms.py
first take at FP's NSBH model

@@ -7,6 +7,7 @@ import numpy as np
 from scipy.optimize import minimize_scalar
 
 import lal
+from pycbc import pnutils
 from pycbc.waveform import *
 from pycbc.types import *
 import pycbc.psd as ppsd
@@ -158,30 +159,177 @@ class tidalWavs():
 
 def windowing_function(f, f0, d, sgn=+1):
   """
-  Returns waveform windowing_function, for the input frequencies
+  Returns waveform windowing_function w_{f0,d}(f), for the input frequencies
   """
   return 0.5*(1. + sgn*np.tanh(4.*(f - f0)/d))
 
+def lorentzian(f, f0, d): return d**2 / ((f - f0)**2 + d**2 / 4.)
+  
 class PNcoeffs():
-  def __init__(verbose=True):
+  def __init__(self, mass1=5, mass2=5, spin1z=0, spin2z=0, verbose=True):
     """
     This class is for storing all PN coefficients, as well for solving for 
     orbital quantities.
     """
-    self.verbose = verbose
+    if not hasattr(self, 'verbose'): self.verbose = verbose
+    
+    self.total_mass = mass1 + mass2
+    self.eta = mass1 * mass2 / self.total_mass**2
+    self.spin1z = spin1z
+    self.spin2z = spin2z
+    self.chi = (mass1 * spin1z + mass2 * spin2z)/self.total_mass
+    
+    self.coeffs = {}
     self.initialize_coefficients_fd()
     return
   #
   def initialize_coefficients_fd(self):
+    """
+    Initialize PN coefficients. Currently implemented are
+    - 3PN amplitude corrections, for aligned-spins. given by Eq.A5 of 1005.3306
+    """
+    eta, chi1, chi2, chi = self.eta, self.spin1z, self.spin2z, self.chi
+    pi2 = np.pi * np.pi
+    eta2 = eta * eta
+    piM = self.total_mass * np.pi * lal.MTSUN_SI
+    chisum = chi + chi    ## This is the chisum in phenomC, is this right?
+    chiprod = chi * chi   ## This is the chiprod in phenomC, is this right?
+    chi2 = chi * chi      ## This is the chi**2 in phenomC, is this right?
+    
[…]

### 2015-11-19 emcee_PE/src/TidalWaveforms.py
improving FP model's code

@@ -344,6 +344,13 @@ class tidalWavsFP(PNcoeffs):
       P2 : arXiv:1311.5931
       P3 : arXiv:0512160
       """
+      print """\
+      (WARNING) --
+      This code has no knowledge of NS EoS yet. It assigns arbitrary values to
+      the baryonic matter in the NS, as well as its radius.
+      """
+      
+      
       PNcoeffs.__init__(self, mass1=massBH, mass2=massNS, spin1z=spinBH, \
                         spin2z=0, verbose=verbose)
       
@@ -554,6 +561,7 @@ class tidalWavsFP(PNcoeffs):
       if hasattr(self, 'fRD') and hasattr(self, 'QRD'):
         return self.fRD, self.QRD
       
+      total_mass = self.mtotal 
       mf = self.massBH_final()
       af = self.spinBH_final()
       
@@ -563,7 +571,7 @@ class tidalWavsFP(PNcoeffs):
       j = np.abs(af) # WHAT IS THIS ?
       
       self.mfRD = f1 + f2 * (1. - j)**f3
-      self.fRD  = self.mfRD / (mf * lal.MTSUN_SI)
+      self.fRD  = self.mfRD / (total_mass * lal.MTSUN_SI)
       self.QRD  = q1 + q2 * (1. - j)**q3
       
       return self.fRD, self.QRD
@@ -730,7 +738,8 @@ class tidalWavsFP(PNcoeffs):
       self.coeffs['TidalModel'] = coeffs
       return coeffs
     #
-    def getWaveformAmplitude(self, mtotal=None, eta=None, sBH=None, Lambda=None,\
+    def getWaveformAmplitude(self, mtotal=None, eta=None, sBH=None, 
+                    massNS_b=0.95, radiusNS=4,\
                     distance=1e6,\
                     f_lower=15., f_final=4096., \
                     delta_t=1./8192., delta_f=1./256, tidal=True):
@@ -754,11 +763,17 @@ class tidalWavsFP(PNcoeffs):
       
       Please go through Sec.IV of P1 for all details.
       """
-      if mtotal is None: mtotal = self.mtotal
-      if eta is None: eta = self.eta
-      if sBH is None: sBH = self.spinBH
-      if Lambda is None: Lambda = self.lambdaNS
-      
+      if mtotal is None and eta is None and sBH is None and Lambda is None:
+        Lambda = self.lambdaNS
[…]

### 2015-11-19 emcee_PE/src/TidalWaveforms.py
first waveform-like waveform made!

@@ -303,8 +303,10 @@ class PNcoeffs():
     logv = np.log(v)
     
     # Get the amplitude 
+    for kk in xdc.keys(): xdc[kk] = np.float128(xdc[kk])
     xdot = 0 + 0j
     xdot += 1. + xdc['a2']*v2 + xdc['a3']*v3 + xdc['a4']*v4
+    #print type(logv.data), type(v5), type(xdc['a6'])
     xdot += xdc['a5']*v5 + (xdc['a6'] + xdc['a6log']*2.*logv)*v6
     xdot += xdc['a7'] * v7
     xdot *= (xdc['aN'] * v10)
@@ -789,6 +791,7 @@ class tidalWavsFP(PNcoeffs):
       
       # Calculate RD freq and Quality factor
       fRD, QRD = self.get_ringdown_quality_frequency()
+      mfRD = self.mfRD
       
       # Calculate frequency of the onset of mass-shedding
       fTide = self.freqTide()
@@ -799,32 +802,39 @@ class tidalWavsFP(PNcoeffs):
       # Get frequency range
       N = int(np.round( 1./delta_t/delta_f ))
       frequencies = np.arange(N/2 + 1) * delta_f
+      mfrequencies = frequencies * mtotal * lal.MTSUN_SI
       
       # Calculate APN
-      APN, amp0 = self.GetWaveformAmplitudeFD(frequencies, mtotal, distance)
-      APN *= windowing_function(frequencies, coeffs['f1'], coeffs['d1'], sgn=-1)
+      APN, amp0 = self.GetWaveformAmplitudeFD(frequencies, mtotal, distance) # only this uses Hz
+      APN *= windowing_function(mfrequencies, coeffs['f1'] * mtotal * lal.MTSUN_SI, coeffs['d1'], sgn=-1)
       
       # Calculate AM
       AM = 0
       if coeffs['c2'] != 0:
-        AM = 1.25 * gamma1 * coeffs['c2'] * frequencies**(5./6.) 
-        AM *= windowing_function(frequencies * mtotal * lal.MTSUN_SI,\
-                                      coeffs['f2'], coeffs['d2'], sgn=-1)
+        AM = 1.25 * gamma1 * coeffs['c2'] * mfrequencies**(5./6.) 
+        AM *= windowing_function(mfrequencies, coeffs['f2'] * mtotal * lal.MTSUN_SI, coeffs['d2'], sgn=-1)
+        AM *= amp0
       
[…]

### 2015-11-19 emcee_PE/src/TidalWaveforms.py
remove lambda from input options for now

@@ -765,8 +765,8 @@ class tidalWavsFP(PNcoeffs):
       
       Please go through Sec.IV of P1 for all details.
       """
-      if mtotal is None and eta is None and sBH is None and Lambda is None:
-        Lambda = self.lambdaNS
+      if mtotal is None and eta is None and sBH is None:
+        #Lambda = self.lambdaNS
         mtotal = self.mtotal
         eta = self.eta
         sBH = self.spinBH

### 2015-12-07 emcee_PE/scripts/ProcessSamples.py
minor documentation change

@@ -220,13 +220,13 @@ def calculate_bias(\
 def calculate_store_biases(qvec=None, chi2vec=None, Lambdavec=None, SNRvec=None,\
             Nsamples=[150000], Nwalkers=[100], outfile='output.h5', mNS = 1.35,\
             recover_tidal=True):
-  #{{{
   ''' 
   This function loops over all parameters for which ranges are given as input,
   reads the posterior samples for each injection combination, and computes 
   biases in different parameters' recovery for each injected parameter 
   combination.  
   '''
+  #{{{
   fout = h5py.File(outfile, 'w')
   
   for qidx, q in enumerate(qvec):

### 2015-12-08 Paper/paper.tex
minor changes to intro

@@
 \newcommand{\chibh}{\chi_\mathrm{BH}}
+\newcommand{\chins}{\chi_\mathrm{NS}}
 \newcommand{\mbh}{m_\mathrm{BH}}
@@
 \begin{enumerate}
- \item Will ignoring the effect of NS deformability during inspiral and NS disruption during merger (in an NSBH binary) change the efficiency of our searches?
- \item Will not modeling NS disruption during NSBH mergers change the systematic and 
- statistical biases in the recovery of binary parameters?
- \item If we use templates which models NS disruption, what information can we recover about
+ \item Will ignoring the effect of NS deformability during inspiral and NS 
+ disruption during merger (in an NSBH binary) change the efficiency of our 
+ searches?
+ \item What systematic biases are introduced by neglecting the NS tidal
+ disruption close to merger? 
+ \item At what SNR do the systematic biases dominate over the statistical 
+ errors in the recovery of binary parameters?
+ \item By using templates which models NS deformation (over inspiral) and 
+ NS disruption (close to merger), what information can we recover about
  the $\Lambda_\mathrm{BH}\sim (R/M)^5_\mathrm{NS}$ parameter?
@@
 
-The Advanced LIGO (aLIGO) observatories began their first observing run ``O1'' in 
-September, 2015. With both detectors a factor of $3-4$ more sensitive than
-their first-generation counterparts, we have access to $\approx 50$ times more 
-spacial volume of the universe at every instant. Within a 
-$3$ month observing run, these detectors would have access to as
-many binary mergers as initial-LIGO detectors did in $\sim 6$ years of 
-coincident operation. Through gradual updates, both of these instruments will
-reach their design sensitivity by $2019$, at which point they will like detect
-around $70$ binary mergers a year~\cite{Abadie:2010cf}.
+The Advanced LIGO (aLIGO) observatories began their first observing run ``O1''
+mid-September 2015, operating at a factor of $3-4$ higher gravitational-wave (GW) 
+strain sensitivity than their first-generation counterparts~\cite{Shoemaker2009}.
+Coalescing binaries of
+compact objects, i.e. Black Holes (BH) and/or Neutron Stars (NS), are the most
+likely targets for GW detection with these ground-based instruments. Over a
+$3$-month O1, these detectors would have access to as
+many binary mergers as initial-LIGO detectors did over approximately $6$ years
+of coincident operation. Both of these instruments will be further upgraded in
+phases to reach their design sensitivity by $2018-19$, at which point we expect to
+observe $\sim 70$ binary mergers a year~\cite{Abadie:2010cf}.
 
-Coalescing binaries containing a neutron star (NS) and a black hole (BH) are
-one of the most important sources of gravitational waves (GW) with the aLIGO
-and Virgo instruments, which expect to detect $\mathcal{O}(10)$
-of such systems every 
-year~\cite{Abadie:2010cf}. Conventional astronomical methods have, to date, made 
-observations of NSs with masses between $XX-YY M_\odot$ and spin
-$|\vec{\chi}\lesssim 0.05$. On the other hand, indirect observations of BHs 
-have placed the masses of stellar-collapse BHs between $5-35M_\odot$, with 
-spin angular momenta ranging from low to nearly extremal values (i.e., nearly
-as fast as possible - see,
+NS-BH binaries are one of the most important sources of gravitational waves (GW)
+with these ground-based instruments, and we expect to detect $\mathcal{O}(10)$
+of such systems every year~\cite{Abadie:2010cf}. 
+%
+One question that aLIGO will likely probe is `what is the nature of
+matter at nuclear densities', i.e. what is the equation of state (EoS) of 
+neutron stars. The EoS determines the compactness of NSs, which determines
+their deformation over the course of inspiral (albeit weakly, at $5$PN order in 
+phasing~\prayush{CITE PN TIDAL RESULTS}), as well as how early in the coalescence
+process will the NS be disrupted by the tidal field of the BH. Once disrupted,
+the NS forms an accretion disk around the BH. Most of this matter disk spirals 
+into the BH within a few milliseconds, with a small fraction becoming unbounded
+and gets ejected~\prayush{CITE BHNS NR}. It is believed that an accompaning
+short (lasting less than $2$ seconds) Gamma-ray Burst (sGRB) might be present
+under certain physical conditions~\prayush{CITE NSBH-GRB},
+making these systems one of the prime targets for joint electromagnetic-GW
+astronomy searches~\prayush{CITE EM-GW}.
+In such a situtation, the emitted GW signal
+is heavily damped post-disruption, and the quasi-normal modes of the resulting
+BH are not excited. Such a GW signal is qualitatively very different from a
+BH-BH merger, and one of the main goals of this work is to determine the 
+signal strength at which this difference will be quantifiabe by aLIGO.
+%
+However, studies have shown that for mass-ratio $\mbh/\mns\gtrsim 5$ the BH
+will need to have fairly large and positively-aligned (with the orbit) spins
+in order for its tidal field to be strong enough to disrupt the NS before the
+innermost stable circular orbit (ISCO) is reached~\cite{Foucart:2013psa}.
+
+Conventional astronomical 
+methods have, to date, made observations of NSs with masses between 
+$XX-YY M_\odot$ and spins $|\vec{\chi}_\mathrm{NS}|\lesssim 0.05$. On the other hand, 
+indirect observations of stellar-mass BHs place their masses between 
+$5-35M_\odot$, with their spin angular momenta $|\vec{\chi}_\mathrm{BH}|$ 
+ranging from small to nearly extremal (Kerr) values (see,
 e.g.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for examples of 
@@
 Refs.~\cite{McClintock:2013vwa,Reynolds:2013qqa} for recent reviews of
-astrophysical BH spin measurements). \prayush{[Do we want to emphasize on BH
-spins here?]}.
+astrophysical BH spin measurements).
 
-One question that aLIGO is likely to probe is `what is the nature of
-matter at nuclear densities', i.e. what is the equation of state (EoS) of 
-neutron stars. The EoS determines the compactness of NSs, which
-determines how early in the coalescence will the NS be tidally disrupted
-by the gravitational field of the BH, resulting in the formation
-of an accretion disk and/or ejection of unbound matter. This makes NSBH
-mergers are one of the prime targets of the program for joint detection of
-GW and electromagnetic signals. 
-% 
 The difficulty in discerning the matter effects in an NSBH merger GW signal

### 2015-12-09 emcee_PE/scripts/ProcessSamples.py
very important bugfixes in emcee_match_* scripts

@@ -15,7 +15,6 @@ from scipy.stats.stats import pearsonr
 from scipy.stats import gaussian_kde
 from scipy import stats
 
-import make_runs as mr
 from utils import *
 import pycbc.pnutils as pnutils
 
@@ -243,7 +242,7 @@ def calculate_store_biases(qvec=None, chi2vec=None, Lambdavec=None, SNRvec=None,
           dset_l1 = 'SNR%.1f.dat' % SNR
           for Ns in Nsamples:
             for Nw in Nwalkers:
-              simdir = mr.get_simdirname(q, mNS, chi2, Lambda, SNR, Nw, Ns)
+              simdir = get_simdirname(q, mNS, chi2, Lambda, SNR, Nw, Ns)
               print "\n\n Trying to read in %s" % simdir
               #
               summ_data = calculate_bias(basedir=cmd.getoutput('pwd -P'),\
@@ -263,7 +262,7 @@ def calculate_store_biases(qvec=None, chi2vec=None, Lambdavec=None, SNRvec=None,
 chi1 = 0.   # small BH
 chi2vec = [-0.5, 0, 0.5, 0.74999]  # larger BH
 mNS = 1.35
-qvec = [2, 3, 4]
+qvec = [5]
 Lambdavec = [0]
 SNRvec = [20, 30, 50, 70, 90, 120]
 
@@ -302,7 +301,7 @@ simstring = sigstring + tmpstring + '_'
 if inject_tidal: Lambdavec = [500, 800, 1000]
 
 Nwalkers = [100]
-Nsamples = [150000]
+Nsamples = [2000]
 Nburnin  = 500
 
 #qvec = [2]
@@ -315,4 +314,5 @@ print "Storing data in HDF file for ", qvec, chi2vec, Lambdavec, SNRvec
 print "\n\n\n"
 calculate_store_biases(qvec=qvec, chi2vec=chi2vec, Lambdavec=Lambdavec,\
       SNRvec=SNRvec, recover_tidal=recover_tidal,\
+      Nsamples=Nsamples, Nwalkers=Nwalkers,\
       outfile=simstring+'ParameterBiasesAndConfidenceIntervals.h5')

### 2015-12-09 emcee_PE/src/emcee_match_aligned.py

@@ -563,7 +563,7 @@ if not post_process:
     outidx= 0
   print " >> will write to output file %d-%d" % (unique_id, outidx)
   print "Dumping chain and logposterior to .npy files every %d iterations." %nout
-  ii=0
+  ii=1
   #for result in sampler.sample(p0, iterations=nsamples, storechain=False):
   for result in sampler.sample(p0, iterations=nsamples, storechain=True):
     position = result[0]

### 2015-12-09 emcee_PE/src/emcee_match_tidal.py

@@ -553,7 +553,7 @@ if not post_process:
     outidx= 0
   print " >> will write to output file %d-%d" % (unique_id, outidx)
   print "Dumping chain and logposterior to .npy files every %d iterations." %nout
-  ii=0
+  ii=1
   #for result in sampler.sample(p0, iterations=nsamples, storechain=False):
   for result in sampler.sample(p0, iterations=nsamples, storechain=True):
     position = result[0]

### 2015-12-09 emcee_PE/src/utils.py

@@ -441,5 +441,11 @@ def load_samples(dataDir, SNR, burnin=500):
             'm1' : m1val, 'm2' : m2val, 'match' : matchval}
     #}}}
 
+def get_simdirname(q, mNS, chi2, Lambda, SNR, Nw, Ns):
+  #{{{
+  return 'q%.2f_mNS%.2f_chiBH%.2f_Lambda%.1f_SNR%.1f/NW%d_NS%d'\
+              % (q, mNS, chi2, Lambda, SNR, Nw, Ns)
+  #}}}
+
 
 

### 2016-01-18 emcee_PE/scripts/AnalyzeSamples.py
updating post-processing

@@ -668,19 +668,21 @@ if len(sys.argv) >= 5:
   if int(sys.argv[4]) != 0: plot_width = True
   else: plot_width = False
 
+figtype = 'pdf'
+
 ######################################################
 # Set up parameters of signal
 ######################################################
 chi1 = 0.                           # small BH
 chi2vec = np.array([-0.5, 0, 0.5, 0.74999])  # larger BH
 mNS = 1.35
-qvec = np.array([2, 3, 4])
+qvec = np.array([2, 3, 4, 5])
 etavec = qvec / (1. + qvec)**2
 mtotalvec = mNS + mNS * qvec
 mchirpvec = mtotalvec * etavec**0.6
 Lambdavec = np.array([0])#[500, 800, 1000])
 SNRvec = np.array([20, 30, 50, 70, 90, 120])
-if inject_tidal: Lambdavec = np.array([0, 500, 800, 1000])
+if inject_tidal: Lambdavec = np.array([500, 800, 1000, 1500, 2000])
 
 ######################################################
 # Set up parameters of templates
@@ -701,7 +703,8 @@ if inject_tidal: sigstring = 'T'
 else: sigstring = 'N'
 if recover_tidal: tmpstring = 'T'
 else: tmpstring = 'N'
-simstring = sigstring + tmpstring + '_'
+simstring = sigstring + tmpstring
+simstring = sigstring + tmpstring
 
 Nwalkers = [100]
 Nsamples = [150000]
@@ -710,13 +713,19 @@ Nburnin  = 500
 ######################################################
 # Read in parameter biases and other data
 ######################################################
-#datafile = simstring + 'ParameterBiasesAndConfidenceIntervals.h5'
+datadir  = '/home/prayush/research/NSBH/TidalParameterEstimation/FinalCombinedPEData/'
+#datadir  = '/home/prayush/research/NSBH/TidalParameterEstimation/ParameterBiasVsSnr/SEOBNRv2/set005/' + simstring
+datafile = simstring + '_ParameterBiasesAndConfidenceIntervals.h5'
+datafile = os.path.join(datadir, datafile)
[…]

### 2016-01-18 emcee_PE/scripts/ProcessSamples.py
processing script also stores maxL parameters

@@ -144,12 +144,14 @@ def calculate_bias(\
     2. Compute 68%, 90%,.. confidence interval for different sampled parameters
     3. Compute the median value of different parameters fromt the posterior.
     
-    We plan to store 5 quantities for each parameter k:
+    We plan to store 7 quantities for each parameter k:
     k+0: Median recovered parameter value from posterior samples
-    k+1: Bias = (X(median) - X(inj))/X(inj)
-    k+2: Confidence level lower bound = X(confidence_level_low)
-    k+3: Confidence level upper bound = X(confidence_level_up)
-    k+4: Confidence interval = (X(confidence_level_up) - X(confidence_level_low))/X(inj)
+    k+1: MaxL   recovered parameter value from posterior samples
+    k+2: Bias = (X(median) - X(inj))/X(inj)
+    k+3: Bias = (X(maxL) - X(inj))/X(inj)
+    k+4: Confidence level lower bound = X(confidence_level_low)
+    k+5: Confidence level upper bound = X(confidence_level_up)
+    k+6: Confidence interval = (X(confidence_level_up) - X(confidence_level_low))/X(inj)
     
     PLUS
     0 : Confidence interval probability
@@ -174,14 +176,14 @@ def calculate_bias(\
     else:
       parameters = ['m1', 'm2', 'Mc', 'Mtot', 'eta', 'q', 'chi2']
     
-    num_of_data_fields = 5
+    num_of_data_fields = 7
     summary_data = np.zeros(( len(confidence_levels), len(parameters)*num_of_data_fields + 1 + 1 ))
     
     # Populate first column of summary data : confidence level probabilities
     for idx in range(len(confidence_levels)):
       summary_data[idx,0] = confidence_levels[idx] / 100.
       summary_data[idx,1] = np.max( match['samples']['match'] ) 
-    
+   match['samples'][param][np.where(match['samples']['match'] == np.max(match['samples']['match']))[0][0]]
     idx = 2
     for param in parameters:
       # get the posterior samples
@@ -196,19 +198,22 @@ def calculate_bias(\
         
         # Fractional bias
         summary_data[jdx, idx] = np.median(S)
-        summary_data[jdx, idx+1] = np.median(S) - param_inj
[…]

### 2016-01-20 emcee_PE/scripts/AnalyzeSamples.py
adding line plots + changing default L range

@@ -17,7 +17,7 @@ from pydoc import help
 from scipy.stats.stats import pearsonr
 from scipy.stats import gaussian_kde
 from scipy import stats
-from scipy.interpolate import UnivariateSpline
+from scipy.interpolate import UnivariateSpline, interp1d
 from scipy.optimize import minimize_scalar
 
 sys.path.append( os.path.dirname(os.path.realpath(__file__)) + '/../src/' )
@@ -48,9 +48,8 @@ CILevels=[90.0, 68.26895, 95.44997, 99.73002]
 ######################################################
 # Function to make parameter bias plots
 ######################################################
-linestyles = ['-', '--', '-.', '-x', '--o']
-linecolors = ['r', 'g', 'b', 'k', 'm', 'y']
-plotdir = 'plots/'
+linestyles = ['-', '-.', '--', ':', '--o']
+linecolors = ['r', 'olivedrab', 'k', 'b', 'm', 'y']
 gmean = (5**0.5 + 1)/2.
 
 # Figure settings
@@ -267,7 +266,7 @@ def make_contour(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', title='',\
                 levelspacing=0.25, vmin=None, vmax=None, cmap=cm.Reds_r,\
                 xmin=None, xmax=None, ymin=None, ymax=None,\
                 cbfmt='%.1f', figname='plot.png'):
-  colwidth = 2.8
+  colwidth = 4.8
   plt.figure(int(1e7 * np.random.random()), figsize=(1.*gmean*colwidth, colwidth))
   norm = cm.colors.Normalize(vmax=Z2d.max(), vmin=Z2d.min())
   #cmap = cm.rainbow
@@ -316,6 +315,71 @@ def make_hexbin(X, Y, Z, xlabel='', ylabel='', clabel='', vmin=-1, vmax=None,\
   cb.set_clim(vmin=vmin, vmax=vmax)
   plt.grid()
   plt.savefig(figname,dpi=400)
+
+
+def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='',\
+                levelspacing=0.25, vmin=None, vmax=None, cmap=cm.Reds_r,\
+                pcolors=linecolors, pmarkers=linestyles,\
+                xmin=None, xmax=None, ymin=None, ymax=None,\
+                single_legend=True,\
+                cbfmt='%.1f', figname='plot.png'):
+  #
+  colwidth = 4.8  
[…]

### 2016-01-20 emcee_PE/scripts/make_runs.py

@@ -211,7 +211,7 @@ chi2min = -0.99
 chi2max = 0.98
 chi2only = True
 
-LambdaMax = 4000
+LambdaMax = 4000*2
 Lambdastdev = 100
 
 recover_tidal = True

### 2016-01-21 Paper/paper.tex
updated AnalyzeSamples to include many new plots

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{What do we gain by using templates that include NS matter effects?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% % 
+% \begin{figure*}
+% \centering
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
+% \caption{This figure is similar to the bottom row of
+% Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
+% consider the NS tidal parameter $\Lambda$ here.}
+% \label{fig:TT_NSLambdaBias_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering
+% \textbf{NS-BH Injection; NS-BH Templates}
+% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
+% \caption{This figure is similar to the bottom row of
+% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
+% we consider the NS tidal parameter $\Lambda$ here.}
+% \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
+% \end{figure*}
+% % 
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=2\columnwidth]{plots-TT/TT_LambdaBiases_CI90_0_Lambda_SNR.png}
+% \caption{This figure shows.}
+% \label{fig:TT_LambdaBiases_CI90_0_Lambda_SNR}
+% \end{figure*}
+
+\begin{figure*}
+\centering    
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
+\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
+\caption{This figure shows.}
+\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
+\end{figure*}
+
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=2.1\columnwidth]{plots-TT/TTLambdaCIWidths68_3_Lambda_SNR.png}\\
+\includegraphics[width=2.1\columnwidth]{plots-TT/TTLambdaCIWidths90_0_Lambda_SNR.png}
+\caption{This figure shows.}
+\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
+\end{figure*}
+% 
+%\begin{figure*}
+%\centering    
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
+%\caption{This figure shows.}
+%\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
+%\end{figure*}
+%% 
+%\begin{figure*}
+%\centering    
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
+%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
+%\caption{This figure shows.}
+%\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
+%\end{figure*}
+
+Having shown in the previous section that we begin to care for NS tidal effects
+for signals in the XXX corner of the paraemter space, with SNRs above YY, here
+we investigate the effect of the improvement in the accuracy of the recovered
+parameters when tidal effects are included in the templates.
+
+\begin{enumerate}
+\item Where in the parameter space can we actually make a statement about 
+$\Lambda_\mathrm{NS}$ ?
+\item What is the reduction in the bias of the maximum likelihood parameters
+when using tidal templates?\newline
+``Plot the ratio of the bias between N templates and T templates (against T
+signals), as a function of SNR.''
+\item What is the reduction in the uncertainty in binary mass and spin, if
+any?\newline
+``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
+when we go from using N templates to T templates.''
+\end{enumerate}
+
+\subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}
+
+With the enhanced sensitivity of aLIGO, we are likely to be able to measure or
+constrain deviations in the observed GW signals, especially close to binary
+merger, from their expected form under the point-particle approximation.
+The tidal disruption of the NS close to merger, and outside of the 
+inner-most circular orbit, will drastically change the morphology of the emitted
+GW signal, the measurement of which will allow us to constrain the equation 
+of state (EoS) of nuclear matter.
+The earlier in the orbit the NS disrupts, the sharper is the drop in the emitted
+GW signal for frequencies above the disruption frequency. In this section we
+study the ability of aLIGO to constrain the 
+$\lambdans\sim \left(R/M\right)_\mathrm{NS}^5$ parameter, which
+the combination that parametrizes the deviation in the emitted GW signal.
+Past studies~\cite{}
+have studied a limited sample of systems with distinguishability criterion that
+are typically applicable in the very high signal-to-noise ratio cases. 
+In what follows, we will use stochastic parameter estimation algorithms to show
+aLIGO's capability of extracting EoS information out from GW signals with 
+more realistic (likely) SNRs.
+
+As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
+fixed at $1.35M_\odot$. We allow BH spin to take the values 
+$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
+parameter combinations, we inject a tidal-corrected waveform using the 
+Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
+posterior probability distribution for different binary parameters with (a) BHBH
+templates, and (b) NSBH templates. We will discuss the results of (b) here.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{How is PE affected by not including NS matter effects?}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{What do we gain by using templates that include NS matter effects?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% % 
-% \begin{figure*}
-% \centering
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
-% \caption{This figure is similar to the bottom row of
-% Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
-% consider the NS tidal parameter $\Lambda$ here.}
-% \label{fig:TT_NSLambdaBias_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
-% \caption{This figure is similar to the bottom row of
-% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
-% we consider the NS tidal parameter $\Lambda$ here.}
-% \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=2\columnwidth]{plots-TT/TT_LambdaBiases_CI90_0_Lambda_SNR.png}
-% \caption{This figure shows.}
-% \label{fig:TT_LambdaBiases_CI90_0_Lambda_SNR}
-% \end{figure*}
-% 
-\begin{figure*}
-\centering    
-\includegraphics[width=2.1\columnwidth]{plots-TT/TT_LambdaCIWidths68_3_Lambda_SNR.png}\\
-\includegraphics[width=2.1\columnwidth]{plots-TT/TT_LambdaCIWidths90_0_Lambda_SNR.png}
-\caption{This figure shows.}
-\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
-\end{figure*}
-% 
-\begin{figure*}
-\centering    
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
-\caption{This figure shows.}
-\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
-\end{figure*}
-% 
-\begin{figure*}
-\centering    
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
-\caption{This figure shows.}
-\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
-\end{figure*}
-
-Having shown in the previous section that we begin to care for NS tidal effects
-for signals in the XXX corner of the paraemter space, with SNRs above YY, here
-we investigate the effect of the improvement in the accuracy of the recovered
-parameters when tidal effects are included in the templates.
-
-\begin{enumerate}
-\item Where in the parameter space can we actually make a statement about 
-$\Lambda_\mathrm{NS}$ ?
-\item What is the reduction in the bias of the maximum likelihood parameters
-when using tidal templates?\newline
-``Plot the ratio of the bias between N templates and T templates (against T
-signals), as a function of SNR.''
-\item What is the reduction in the uncertainty in binary mass and spin, if
-any?\newline
-``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
-when we go from using N templates to T templates.''
-\end{enumerate}
-
-\subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}
-
-With the enhanced sensitivity of aLIGO, we are likely to be able to measure or
-constrain deviations in the observed GW signals, especially close to binary
-merger, from their expected form under the point-particle approximation.
-The tidal disruption of the NS close to merger, and outside of the 
-inner-most circular orbit, will drastically change the morphology of the emitted
-GW signal, the measurement of which will allow us to constrain the equation 
-of state (EoS) of nuclear matter.
-The earlier in the orbit the NS disrupts, the sharper is the drop in the emitted
-GW signal for frequencies above the disruption frequency. In this section we
-study the ability of aLIGO to constrain the 
-$\lambdans\sim \left(R/M\right)_\mathrm{NS}^5$ parameter, which
-the combination that parametrizes the deviation in the emitted GW signal.
-Past studies~\cite{}
-have studied a limited sample of systems with distinguishability criterion that
-are typically applicable in the very high signal-to-noise ratio cases. 
-In what follows, we will use stochastic parameter estimation algorithms to show
-aLIGO's capability of extracting EoS information out from GW signals with 
-more realistic (likely) SNRs.
-
-As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
-fixed at $1.35M_\odot$. We allow BH spin to take the values 
-$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
-parameter combinations, we inject a tidal-corrected waveform using the 
-Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
-posterior probability distribution for different binary parameters with (a) BHBH
-templates, and (b) NSBH templates. We will discuss the results of (b) here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}

### 2016-01-21 emcee_PE/scripts/AnalyzeSamples.py

@@ -46,10 +46,13 @@ perc_int_99_7 = [0.13, 99.87] # 3 sigma
 CILevels=[90.0, 68.26895, 95.44997, 99.73002]
 
 ######################################################
+# PLOTTING :
 # Function to make parameter bias plots
 ######################################################
-linestyles = ['-', '-.', '--', ':', '--o']
+linestyles = ['solid', 'dashdot', 'dashed', 'dotted']
 linecolors = ['r', 'olivedrab', 'k', 'b', 'm', 'y']
+linemarkers= ['', 'x', 'o', '^']
+
 gmean = (5**0.5 + 1)/2.
 
 # Figure settings
@@ -75,8 +78,8 @@ plt.rcParams.update({\
 
 # Plotting functions
 def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
-        title='', titles=[], \
-        levelspacing=0.25, vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f',\
+        title='', titles=[], levellines=[100,200], levelspacing=0.25,\
+        vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f',\
         xmin=None, xmax=None, ymin=None, ymax=None,\
         colorbartype='simple', figname='plot.png'):
   """
@@ -126,7 +129,7 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
       VMINmax = np.append( VMINmax, Z2d[idx][jdx].min() )
       VMAXmin = np.append( VMAXmin, Z2d[idx][jdx].max() )
   VMIN, VMAX = VMINmax.min(), VMAXmin.max()
-  print "VMIN = %e, VMAX = %e" % (VMIN, VMAX)
+  if vverbose: print "VMIN = %e, VMAX = %e" % (VMIN, VMAX)
   if VMIN > VMAX: raise IOError("Cant fit all data on a single colorbar")
   
   # Now make contour plots
@@ -151,6 +154,10 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
               levels=levels, \
               cmap = cm.get_cmap(cmap, len(levels)-1), norm=norm,\
               alpha=0.9, vmin=VMIN, vmax=VMAX)
+      cset = ax.contour(xx, yy, zz, levels=levellines,\
[…]

### 2016-01-23 emcee_PE/scripts/AnalyzeSamples.py
add plots for lambda err contours

@@ -44,13 +44,15 @@ perc_int_95 = [2.28, 97.72]
 perc_int_99_7 = [0.13, 99.87] # 3 sigma
 
 CILevels=[90.0, 68.26895, 95.44997, 99.73002]
+ErrThresh=[100, 200]
 
 ######################################################
 # PLOTTING :
 # Function to make parameter bias plots
 ######################################################
-linestyles = ['solid', 'dashdot', 'dashed', 'dotted']
-linecolors = ['r', 'olivedrab', 'k', 'b', 'm', 'y']
+linestyles = ['solid', 'dashed', 'dashdot', 'dashed', 'dotted']
+#linecolors = ['crimson', 'olivedrab', 'k', 'b', 'm', 'y']
+linecolors = ['crimson', 'darkorange', 'olivedrab', 'royalblue', 'purple', 'k']
 linemarkers= ['', 'x', 'o', '^']
 
 gmean = (5**0.5 + 1)/2.
@@ -85,6 +87,15 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
   """
   Function to plot arbitrary numbers of contour plots in a single figure
   """
+  print """\
+
+Remember that the returned set of contours are indexed by integer pairs that
+set coordinates for [row, col]:-
+
+row corresponds to the higher level group in the input Z array, 
+col corresponds to the lower level group
+
+  """
   if colorbartype != 'simple':
     raise IOError("Colorbar type %s not supported" % colorbartype)
   colwidth = 1.7
@@ -109,7 +120,7 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
   pltid = 0
   
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((2.1*gmean*ncol+1.25)*colwidth, 1.2*colwidth*nrow))
+              figsize=((1.5*gmean*ncol+1.25)*colwidth, 1.2*colwidth*nrow))
   fig.clf()
   grid = ImageGrid(fig, 111, nrows_ncols=(nrow, ncol), \
             share_all=True,\
@@ -133,7 +144,9 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
[…]

### 2016-01-29 emcee_PE/scripts/AnalyzeSamples.py
minor changes to plotting script

@@ -8,7 +8,7 @@ import os, sys
 import commands as cmd
 import numpy as np
 from matplotlib import mlab, cm, use
-#use('Agg')
+use('pdf')
 import matplotlib.pyplot as plt
 plt.rcParams.update({'text.usetex' : True})
 from mpl_toolkits.axes_grid1 import ImageGrid
@@ -346,7 +346,7 @@ def make_hexbin(X, Y, Z, xlabel='', ylabel='', clabel='', vmin=-1, vmax=None,\
 
 
 def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='',\
-                levelspacing=0.25, vmin=None, vmax=None, cmap=cm.Reds_r,\
+                levelspacing=0.25, vmin=None, vmax=None, background_color=None,\
                 pcolors=linecolors, pmarkers=linestyles, lw=1.5,\
                 xmin=None, xmax=None, ymin=None, ymax=None,\
                 markerfirst=[True, False], sort_keys=False,\
@@ -375,8 +375,7 @@ def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='
     
   #fig = plt.figure(int(1e7 * np.random.random()), \
   #            figsize=((2.1*gmean*ncol+1.25)*colwidth, 1.2*colwidth*nrow))
-  plt.figure(int(1e7 * np.random.random()), figsize=(1.2*gmean*colwidth, colwidth))
-  
+  fig = plt.figure(int(1e7 * np.random.random()), figsize=(1.2*gmean*colwidth, colwidth))
   # FIrst make all lines
   all_lines = []
   xykeys = XY.keys()
@@ -391,6 +390,7 @@ def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='
       grp_lines.append( line )
     all_lines.append( grp_lines )
   #
+  if background_color is not None: plt.gca().get_axes().set_axis_bgcolor(background_color)
   # Now make legends, one to indicate each group's characteristic
   if single_legend:
     harray = [all_lines[i][0] for i in range(ngrp1)]
@@ -990,15 +990,15 @@ plotdir = 'plots' + simstring + '/'
 figtype = 'pdf'
 
 # Plots showing Mass/Spin/Lambda recovery biases
-plot_MchirpBias      = False
-plot_MchirpErrorsAll = False
-plot_EtaBias         = False
-plot_QBias           = False
[…]

### 2016-01-29 emcee_PE/scripts/make_runs.py

@@ -189,8 +189,8 @@ def get_simdirname(q, mNS, chi2, Lambda, SNR, Nw, Ns):
 chi1 = 0.   # small BH
 chi2vec = [-0.5, 0, 0.5, 0.74999]  # larger BH
 mNS = 1.35
-qvec = [2, 3, 4]
-Lambdavec = [500, 800, 1000]
+qvec = [2, 3, 4, 5]
+Lambdavec = [500, 800, 1000, 1500, 2000]
 SNRvec = [20, 30, 50, 70, 90, 120]
 
 inject_tidal = True
@@ -214,7 +214,7 @@ chi2only = True
 LambdaMax = 4000*2
 Lambdastdev = 100
 
-recover_tidal = True
+recover_tidal = False
 
 ######################################################
 # Set up RUN parameters

### 2016-01-29 emcee_PE/src/emcee_match_aligned.py

@@ -712,9 +712,11 @@ quantiles=quantiles_95
 samples_combined = np.array([qval, Mval, chi1val, chi2val]).T
 print np.shape(samples_combined)
 print samples_combined
-fig = corner.corner(samples_combined, labels=["$q$", "$M$", "$\chi_1$", "$\chi_2$"], truths=[q_true, M_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
-fig.savefig("corner_q_M_chi1_chi2.png")
-pl.close(fig)
+try:
+  fig = corner.corner(samples_combined, labels=["$q$", "$M$", "$\chi_1$", "$\chi_2$"], truths=[q_true, M_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)
+  fig.savefig("corner_q_M_chi1_chi2.png")
+  pl.close(fig)
+except: pass
 
 samples_combined2 = np.array([etaval, Mcval, chi1val, chi2val]).T
 fig = corner.corner(samples_combined2, labels=["$\eta$", "$\mathcal{M}$", "$\chi_1$", "$\chi_2$"], truths=[eta_true, Mc_true, chi1_true, chi2_true], quantiles=quantiles, show_titles=True, title_args={"fontsize": 12}, verbose=True, plot_contours=True, plot_datapoints=True)

### 2016-01-29 Paper/paper.tex
added new figures. IMP

@@
 \begin{enumerate}
- \item Will ignoring the effect of NS deformability during inspiral and NS 
- disruption during merger (in an NSBH binary) change the efficiency of our 
- searches?
  \item What systematic biases are introduced by neglecting the NS tidal
@@
 phasing~\prayush{CITE PN TIDAL RESULTS}), as well as how early in the coalescence
-process will the NS be disrupted by the tidal field of the BH. Once disrupted,
+process will the NS be disrupted by the tidal field of the BH~\prayush{[CITE]}.
+Once disrupted,
 the NS forms an accretion disk around the BH. Most of this matter disk spirals 
 into the BH within a few milliseconds, with a small fraction becoming unbounded
-and gets ejected~\prayush{CITE BHNS NR}. It is believed that an accompaning
+and gets ejected~\prayush{CITE BHNS NR}. It is believed that an accompanying
 short (lasting less than $2$ seconds) Gamma-ray Burst (sGRB) might be present
@@
 
-Conventional astronomical 
-methods have, to date, made observations of NSs with masses between 
-$XX-YY M_\odot$ and spins $|\vec{\chi}_\mathrm{NS}|\lesssim 0.05$. On the other hand, 
-indirect observations of stellar-mass BHs place their masses between 
-$5-35M_\odot$, with their spin angular momenta $|\vec{\chi}_\mathrm{BH}|$ 
-ranging from small to nearly extremal (Kerr) values (see,
-e.g.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for examples of 
-nearly extremal estimates of BH spins, and see
-Refs.~\cite{McClintock:2013vwa,Reynolds:2013qqa} for recent reviews of
-astrophysical BH spin measurements).
+Conventional astronomical methods have, to date, made observations of ~2500
+NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
+Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
+stellarcollapsemass}, although
+tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
+%
+Observations of over 2500 NSs have shown that the spin of NSs
+$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
+%
+On the other hand, indirect observations of stellar-mass BHs place their
+masses between $5-35M_\odot$, with their spin angular momenta 
+$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
+(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
+examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
+Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
+and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
 
@@
 The goal of this paper is to probe the distinguishability of BHBH and NSBH
-mergers, focusing at the effects of the BH mass, BH spin, and NS tidal 
-deformability. We also investigate the accuracy with which aLIGO can measure
-the tidal deformability parameter $\Lambda_\NS\sim(R/M)^5_\NS$. For both of 
+mergers, focusing at the effects of the BH mass, BH spin, and the tidal 
+deformability parameter $\Lambda_\NS\sim(R/M)^5_\NS$. We also investigate the
+accuracy with which aLIGO can measure $\Lambda_\NS$ itself. For both of 
 these studies, we will use the zero-detuning high-power noise curve for the
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How are detection searches affected by not including NS matter effects}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Near future detection searches with aLIGO are poised to search for NSBH
-signals using BHBH templates. In this section, we examine the effect of ignoring
-the effects of tidal deformation and disruption of the NS during inspiral and 
-merger phases of search templates. We construct a template bank $\mathcal{B}$ 
-restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
-$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
-using a PN-based geometric lattice placement method~\cite{Harry:2013tca}.
-
-Before investigating its SNR recovery against tidal NSBH signals, we test the 
-effectualness of this bank to point-particle NSBH signals,
-in order to make sure that the placement method does not leave any region of
-the binary parameter-space under-covered. We simulate $50,000$ signals with 
-component-mass and BH spin values
-sampled uniformly over their allowed ranges, and filter each of them against
-the entire bank. For both signals and templates, we use the frequency-domain 
-reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
-implemented within the LIGO algorithms Library~\cite{lal}. We record the 
-fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
-signals, which measures the fraction of optimal SNR that the bank is able
-to recover. Fig.~\ref{} shows these FFs as a function of the injected
-BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
-values. From these figures, we conclude that our bank is effectual as intended.
-
-Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
-detection searches, we perform a similar calculation as above, with the
-difference that the signals are now modelled using the model of tidal 
-corrections to IMR amplitude and phasing published in 
-Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{How are detection searches affected by not including NS matter effects}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%Near future detection searches with aLIGO are poised to search for NSBH
+%signals using BHBH templates. In this section, we examine the effect of ignoring
+%the effects of tidal deformation and disruption of the NS during inspiral and 
+%merger phases of search templates. We construct a template bank $\mathcal{B}$ 
+%restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
+%$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
+%using a PN-based geometric lattice placement method~\cite{Harry:2013tca}. 
+%
+%
+%Before investigating its SNR recovery against tidal NSBH signals, we test the 
+%effectualness of this bank to point-particle NSBH signals,
+%in order to make sure that the placement method does not leave any region of
+%the binary parameter-space under-covered. We simulate $50,000$ signals with 
+%component-mass and BH spin values
+%sampled uniformly over their allowed ranges, and filter each of them against
+%the entire bank. For both signals and templates, we use the frequency-domain 
+%reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
+%implemented within the LIGO algorithms Library~\cite{lal}. We record the 
+%fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
+%signals, which measures the fraction of optimal SNR that the bank is able
+%to recover. Fig.~\ref{} shows these FFs as a function of the injected
+%BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
+%values. From these figures, we conclude that our bank is effectual as intended.
+%
+%
+%Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
+%detection searches, we perform a similar calculation as above, with the
+%difference that the signals are now modelled using the model of tidal 
+%corrections to IMR amplitude and phasing published in 
+%Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% % 
-% \begin{figure*}
-% \centering
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
-% \caption{This figure is similar to the bottom row of
-% Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
-% consider the NS tidal parameter $\Lambda$ here.}
-% \label{fig:TT_NSLambdaBias_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
-% \caption{This figure is similar to the bottom row of
-% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
-% we consider the NS tidal parameter $\Lambda$ here.}
-% \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=2\columnwidth]{plots-TT/TT_LambdaBiases_CI90_0_Lambda_SNR.png}
-% \caption{This figure shows.}
-% \label{fig:TT_LambdaBiases_CI90_0_Lambda_SNR}
-% \end{figure*}
-
-\begin{figure*}
-\centering    
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
-\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
-\caption{This figure shows.}
-\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
-\end{figure*}
-
-% 
-\begin{figure*}
-\centering    
-\includegraphics[width=2.1\columnwidth]{plots-TT/TTLambdaCIWidths68_3_Lambda_SNR.png}\\
-\includegraphics[width=2.1\columnwidth]{plots-TT/TTLambdaCIWidths90_0_Lambda_SNR.png}
-\caption{This figure shows.}
-\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
-\end{figure*}
-% 
 %\begin{figure*}
@@
 %\end{figure*}
+% 
+\begin{figure*}
+\centering    
+\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
+%\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
+\caption{This figure shows the statistical uncertainty in the measurement of
+$\lambdans$, as a percentage of the injected signal's $\lambdans$. In each panel, 
+the same is shown as a function of the BH mass and spin, keeping the 
+$\lambdans$ and injection's SNR $\rho$ fixed (given in the panel). Each row contains
+panels with the same value of $\lambdans$, while each column contains panels 
+with the same value of $\rho$. Two contours are drawn in each panel demarkating
+that the statistical measurement uncertainty for $\lambda$ is equal to (or twice
+of) its injected $\lambdans$ value, i.e. where we can place a factor of $1-2$ 
+bound on $\lambdans$.
+% 
+We note that, as expected, the measurement accuracy for $\lambdans$ improves with (i) increasing
+SNR, (ii) decreasing BH mass, (iii) increasing BH spin, and 
+(iv) increasing $\lambdans$, i.e. the tidal deformability of the neutron star.
+}
+\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
+\end{figure*}
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
+\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
+\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
+% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
+% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR90_CI90_0.pdf}
+% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR120_CI90_0.pdf}
+\caption{We demarkate here the regions of BH mass and spin plane where we can 
+constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
+Panels correspond to the injected SNR, with signal strength increasing
+from left to right and top to bottom. Within each panel, $\lambdans$ measurement uncertainty
+contours are shown as functions of BH mass and spin, with different colors corresponding to the
+injected $\lambdans$. We note that this information is, in principle, contained in 
+Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
+the effect of NS's deformability itself on its measurement. As from Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability of NS matter
+effects improves with all factors that enhances the signature of the NS's disruption on
+the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
+set by the detectors).
+}
+\label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
+\end{figure*}
+%
+\begin{figure*}
+\centering    
+\includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
+\includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_m1_m2_chi2_Lambda.pdf}
+\caption{These figure shows the $1-$ and $2-$dimensional 
+\textit{a posteriori} probability distributions 
+for different physical parameters describing a BHNS system including NS matter 
+effects, as measured by a Bayesian Markov-chain Monte Carlo using tidal templates.
+The true parameters of the injected signal are: 
+$\mathcal{M}=1.6406M_\odot, \eta=0.2222, \chibh=\chi_2=0.75$ and $\lambdans=1500$,
+with the injected SNR $\rho=20$. We show this case as typical of those with the largest
+tidal signatures imprinted on the emitted GWs (due to the system being comparable
+mass-ratio with a large-aligned spin on the BH and a fairly deformable NS), yet 
+with low SNR.
+% 
+We note a few things:
+(i) chirp mass is well constrained, with a nearly-Gaussian probability distribution;
+while (ii) the same for the other $3$ less constrained parameters rails against
+prior boundaries. The specific boudary conditions that we run into are (a) for $\eta$,
+we are limited by the requirement that $m_\mathrm{NS}\geq 1.2M_\odot$ 
+of our templates; (b) for BH spin, we are limited by the $\chibh\leq+0.75$ requirement
+that results from a template model limitation~\cite{Lackey:2013axa}; and (c) for 
+$\lambdans$, we
+restrict it to $[0, 4000]$ for the same reason as (b). These prior boundaries
+limit the exploration of the entire likelihood surfaces. While (a) is astrophysically
+justified, given that we have not observed NSs with masses below 
+$1.25M_\odot$, (b) and (c) will be removed in future studies as more evolved template
+models become available. See, also, Fig.~\ref{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}.
+% 
+Finally, we note that the left panel corresponds to the actual sampling parameters,
+while the right panel to derived ones.
+}
+\label{fig:TT_pdf_q2_chi75_L1500_SNR20}
+\end{figure*}
+%
+\begin{figure*}
+\centering    
+% \includegraphics[width=.66\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
+\includegraphics[width=\columnwidth]{plots/TT_logLike_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
+\includegraphics[width=\columnwidth]{plots/TT_logProb_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
+\caption{These figures show the effect of limitation (a) (described in the caption
+of Fig.~\ref{fig:TT_pdf_q2_chi75_L1500_SNR20}) on the likelihood surface portion
+that is actually accessible to our Bayesian inferencing algorithm. On the left we show the
+entire likelihood surface in the $\mbh-\mns$ plane, while on the left is the 
+portion not forbidden by the requirement $\mns\geq 1.2M_\odot$. Similarly, our
+priors on BH spin and NS deformability further curtail the portion of sampled
+parameter space, albeit to much less impact than this mass cut.}
+\label{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}
+\end{figure*}
+
+%
 %% 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{figure*}
-% \centering 
-% % \textbf{BH-BH Injection; BH-BH Templates}
-% % \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
-% \caption{These figures show the (fractional) difference between the chirp mass 
-% value corresponding to the median of its posterior probability distribution,
-% and the actual injected chirp mass,
-% as a function of the signal-to-noise-ratio (SNR) of the injected signal.
-% % In the top row, NS matter effects on the inspiral and merger waveform
-% % are neithed included in templates nor in the injected signal. 
-% In the top row, NS matter effects (inspiral and merger) are included in the
-% injections but not in templates. This corresponds to the current 
-% plan for aLIGO parameter estimation studies. The bottom row shows the effect
-% of additionally including NS matter effects in the templates.
-% % 
-% In each row, the left and right panel correspond to 
-% $q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
-% In each panel, the color of each curve corresponds to the value of NS 
-% deformability parameter $(\Lambda=\{500,1000,2000\})$, 
-% and line-style corresponds to the value of
-% the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
-% }
-% \label{fig:TNT_chirpMassBias_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering 
-% % \textbf{BH-BH Injection; BH-BH Templates}
-% % \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
-% \caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
-% with the difference that here we show the width of the $90\%$ confidence 
-% interval (recovered) for chirp mass, as a function of the injected SNR. 
-% Note that the confidence interval's width is normalized by the injected
-% parameter value.
-% }
-% \label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
-% \end{figure*}
-% % 
-% 
-% \begin{figure*}[!t]
-% \centering    
-% \textbf{NS-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
-% \textbf{BH-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
-% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
-% with the difference that we consider the systematic bias in the dimensionless
-% mass-ratio $\eta$ here. In the bottom row we show control runs where both
-% signal and template waveforms lack matter effects.}
-% \label{fig:TNT_EtaBias_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering    
-% \textbf{NS-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
-% \textbf{BH-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
-% \caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
-%   with the difference that we consider the confidence intervals in the recovery
-%   of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
-%   Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
-%   confidence interval is shown here.}
-% \label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
-% \end{figure*}
-% \begin{figure*}
-% \centering    
-% \textbf{NS-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
-% \textbf{BH-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
-% \caption{The top two rows in this figure are
-% similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
-% with the differences that we consider the systematic bias in the dimensionless
-% spin on the BH here, and that the differences shown are absolute and not
-% fractions of the injected spin. In the bottom row, we show a control run where
-% both injected and template waveforms correspond to binary black holes.}
-% \label{fig:TNT_BHspinBias_vs_SNR_q23}
-% \end{figure*}
-% % 
-% \begin{figure*}
-% \centering    
-% \textbf{BH-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; BH-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
-% \textbf{NS-BH Injection; NS-BH Templates}
-% \includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
-% \caption{The top two rows in the figure are similar to
-% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
-% with the difference that we consider the confidence intervals in the recovery
-% of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
-% As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
-% width of the confidence interval without normalizing with the injected spin.
-% The bottom row show a control run where matter effects are ignored both 
-% in the injected signal and the template waveforms.}
-% \label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
-% \end{figure*}
 
@@
 \centering 
-\includegraphics[trim={{0.2\columnwidth} 0 0 0},width=2.1\columnwidth]{plots-TN/TN_MchirpBiases_Lambda_SNR.pdf}
-\caption{Systematic fractional biases in recovered chirp-mass $\mchirp$, as a
-fraction of the injected (true) value, for NSBH binaries, shown here as a 
-function of the mass and spin of the BH (each panel), $\lambdans$ (across 
-rows), and injection's SNR $\rho$ (across columns). Tidal effects during
-inspiral and the effects of NS disruption close to merger are modelled in the
-signal, but \textit{not} in the templates. 
+%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
+\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}.pdf
+\caption{We show here (Log of) the ratio of the systematic and statistical
+measurement uncertainties for the binary chirp mass over the NSBH parameter 
+space. Each panel shows the same as a function of BH mass and spin. Across
+each row, we can see the effect of increasing the signal strength (i.e. SNR)
+with the NS's deformability fixed. Down each column, we can see 
+the effect of the increasing NS's tidal deformability at fixed SNR. We also
+show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
+or $100\%$ (labelled ``0'').
+% 
+Generally, these two sources of uncertainty are different and for BBHs, the
+statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
+that for BHNS binaries, its not much different until we get to high SNRs
+$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
+measurement uncertainty for $\mchirp$ could become dominated by the exclusion
+of tidal effects in filtering templates.
 }
-\label{fig:TNT_chirpMassBias_vs_SNR_q23}
+\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
 \end{figure*}
+%
+\begin{figure*}[!t]
+\centering    
+%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
+\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
+\caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
+with the difference that the quantity in question here is the symmetric 
+mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
+NS tidal effects on GW emission can become the dominant source of error. However,
+for quieter signals with $\rho\leq 30$, we find that not including NS matter 
+effects in parameter estimation templates has minimal effects on the measurement
+of $\eta$. We remind the reader that the SNRs here are all single detector values.
+}
+\label{fig:TN_EtaBias_vs_Lambda_SNR}
+\end{figure*}
+% 
 % 
@@
 \centering
-\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_MchirpCIWidths90_0_Lambda_SNR.pdf}
-\caption{$90\%$ confidence intervals for the recovered chirp-mass $\mchirp$, as a
-fraction of the injected (true) value, for NSBH binaries, shown here as a 
-function of the mass and spin of the BH (each panel), $\lambdans$ (across 
-rows), and injection's SNR $\rho$ (across columns). Tidal effects during
-inspiral and the effects of NS disruption close to merger are modelled in the
-signal, but \textit{not} in the templates. 
+\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
+\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
+\caption{These figures show the ratio of the systematic and statistical
+measurement errors for BH spins, for all of our injected cases. The left
+panel corresponds to $1-\sigma$ statistical confidence level, while the right panel
+corresponds to $90\%$ confidence level. We note that below $\rho\approx 30$, ignoring
+tidal effects in templates introduces systematic effects that remain subdominant
+to the statistical measurement uncertainties.
 }
-\label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
-\end{figure*}
-% 
-\newpage
-\newpage
-
-\begin{figure*}[!t]
-\centering    
-\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaBiases_Lambda_SNR.pdf}
-\caption{Systematic fractional biases in the recovered dimensionless mass-ratio
-$\eta = m_1 m_2 / (m_1+m_2)^2$, as a
-fraction of the injected (true) value, for NSBH binaries, shown here as a 
-function of the mass and spin of the BH (each panel), $\lambdans$ (across 
-rows), and injection's SNR $\rho$ (across columns). Tidal effects during
-inspiral and the effects of NS disruption close to merger are modelled in the
-signal, but \textit{not} in the templates. }
-\label{fig:TNT_EtaBias_vs_SNR_q23}
-\end{figure*}
-% 
-\begin{figure*}
-\centering    
-\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_EtaCIWidths68_3_Lambda_SNR.pdf}
-\caption{$90\%$ confidence intervals for the recovered dimensionless mass-ratio
-$\eta = m_1 m_2 / (m_1+m_2)^2$, as a
-fraction of the injected (true) value, for NSBH binaries, shown here as a 
-function of the mass and spin of the BH (each panel), $\lambdans$ (across 
-rows), and injection's SNR $\rho$ (across columns). Tidal effects during
-inspiral and the effects of NS disruption close to merger are modelled in the
-signal, but \textit{not} in the templates. }
-\label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
-\end{figure*}
-% 
-\begin{figure*}
-\centering
-\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHBiases_CI90_0_Lambda_SNR.pdf}
-\caption{Systematic fractional biases in the recovered value of the BH spin $\chibh$,
-as a fraction of the injected (true) value, for NSBH binaries, shown here as a
-function of the mass and spin of the BH (each panel), $\lambdans$ (across 
-rows), and injection's SNR $\rho$ (across columns). Tidal effects during
-inspiral and the effects of NS disruption close to merger are modelled in the
-signal, but \textit{not} in the templates. }
 \label{fig:TNT_BHspinBias_vs_SNR_q23}
-\end{figure*}
-% 
-\begin{figure*}
-\centering    
-\includegraphics[trim={{0.4\columnwidth} 0 0 0},width=2.2\columnwidth]{plots-TN/TN_ChiBHCIWidths90_0_Lambda_SNR.pdf}
-\caption{$90\%$ confidence intervals for the recovered value of the BH spin $\chibh$,
-as a fraction of the injected (true) value, for NSBH binaries, shown here as a 
-function of the mass and spin of the BH (each panel), $\lambdans$ (across 
-rows), and injection's SNR $\rho$ (across columns). Tidal effects during
-inspiral and the effects of NS disruption close to merger are modelled in the
-signal, but \textit{not} in the templates.}
-\label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
 \end{figure*}

### 2016-02-02 Paper/paper.tex
changes to paper

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{How is PE affected by not including NS matter effects?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% #################
+\begin{figure*}
+\centering 
+%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
+\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}.pdf
+\caption{We show here (Log of) the ratio of the systematic and statistical
+measurement uncertainties for the binary chirp mass over the NSBH parameter 
+space. Each panel shows the same as a function of BH mass and spin. Across
+each row, we can see the effect of increasing the signal strength (i.e. SNR)
+with the NS's deformability fixed. Down each column, we can see 
+the effect of the increasing NS's tidal deformability at fixed SNR. We also
+show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
+or $100\%$ (labelled ``0'').
+% 
+Generally, these two sources of uncertainty are different and for BBHs, the
+statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
+that for BHNS binaries, its not much different until we get to high SNRs
+$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
+measurement uncertainty for $\mchirp$ could become dominated by the exclusion
+of tidal effects in filtering templates.
+}
+\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
+\end{figure*}
+%
+\begin{figure*}[!t]
+\centering    
+%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
+\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
+\caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
+with the difference that the quantity in question here is the symmetric 
+mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
+NS tidal effects on GW emission can become the dominant source of error. However,
+for quieter signals with $\rho\leq 30$, we find that not including NS matter 
+effects in parameter estimation templates has minimal effects on the measurement
+of $\eta$. We remind the reader that the SNRs here are all single detector values.
+}
+\label{fig:TN_EtaBias_vs_Lambda_SNR}
+\end{figure*}
+% 
+% 
+\begin{figure*}
+\centering
+\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
+\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
+\caption{These figures show the ratio of the systematic and statistical
+measurement errors for BH spins, for all of our injected cases. The left
+panel corresponds to $1-\sigma$ statistical confidence level, while the right panel
+corresponds to $90\%$ confidence level. We note that below $\rho\approx 30$, ignoring
+tidal effects in templates introduces systematic effects that remain subdominant
+to the statistical measurement uncertainties.
+}
+\label{fig:TNT_BHspinBias_vs_SNR_q23}
+\end{figure*}
+% 
+% 
+Advanced LIGO searches and parameter estimation efforts aimed at binaries 
+containing NS and stellar-mass BHs are poised to use waveform models that do not
+include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
+While this is not expected to
+be the dominant source of error at low SNRs, it will likely introduce a
+systematic bias in the physical parameters that we recover from GW observations.
+In this section, we study the impact of the same for binaries where the BH spins
+are aligned to the orbital angular momentum.
+
+\begin{enumerate}
+\item Above what SNR values, below what mass-ratios, above what BH spins, etc, 
+do we begin to care about NS deformability?\newline
+``one interesting plot to make would be the parameter bias (max-Likelihood
+parameters minus true parameters) vs SNR, as functionals of q/sBH/Lambda, when
+using T signals recovered with N templates.''\newline
+\end{enumerate}
+
+\prayush{Could have subsections for different frequency-optimized aLIGO noise curves ?}
+
+\hspace{5mm}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{What do we gain by using templates that include NS matter effects?}
@@
 (i) chirp mass is well constrained, with a nearly-Gaussian probability distribution;
-while (ii) the same for the other $3$ less constrained parameters rails against
+while (ii) the other $3$ less constrained parameters rail against
 prior boundaries. The specific boudary conditions that we run into are (a) for $\eta$,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How is PE affected by not including NS matter effects?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% #################
-\begin{figure*}
-\centering 
-%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}.pdf
-\caption{We show here (Log of) the ratio of the systematic and statistical
-measurement uncertainties for the binary chirp mass over the NSBH parameter 
-space. Each panel shows the same as a function of BH mass and spin. Across
-each row, we can see the effect of increasing the signal strength (i.e. SNR)
-with the NS's deformability fixed. Down each column, we can see 
-the effect of the increasing NS's tidal deformability at fixed SNR. We also
-show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
-or $100\%$ (labelled ``0'').
-% 
-Generally, these two sources of uncertainty are different and for BBHs, the
-statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
-that for BHNS binaries, its not much different until we get to high SNRs
-$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
-measurement uncertainty for $\mchirp$ could become dominated by the exclusion
-of tidal effects in filtering templates.
-}
-\label{fig:TN_chirpMassBias_vs_Lambda_SNR}
-\end{figure*}
-%
-\begin{figure*}[!t]
-\centering    
-%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
-\caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
-with the difference that the quantity in question here is the symmetric 
-mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
-NS tidal effects on GW emission can become the dominant source of error. However,
-for quieter signals with $\rho\leq 30$, we find that not including NS matter 
-effects in parameter estimation templates has minimal effects on the measurement
-of $\eta$. We remind the reader that the SNRs here are all single detector values.
-}
-\label{fig:TN_EtaBias_vs_Lambda_SNR}
-\end{figure*}
-% 
-% 
-\begin{figure*}
-\centering
-\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
-\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
-\caption{These figures show the ratio of the systematic and statistical
-measurement errors for BH spins, for all of our injected cases. The left
-panel corresponds to $1-\sigma$ statistical confidence level, while the right panel
-corresponds to $90\%$ confidence level. We note that below $\rho\approx 30$, ignoring
-tidal effects in templates introduces systematic effects that remain subdominant
-to the statistical measurement uncertainties.
-}
-\label{fig:TNT_BHspinBias_vs_SNR_q23}
-\end{figure*}
-% 
-% 
-Advanced LIGO searches and parameter estimation efforts aimed at binaries 
-containing NS and stellar-mass BHs are poised to use waveform models that do not
-include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
-While this is not expected to
-be the dominant source of error at low SNRs, it will likely introduce a
-systematic bias in the physical parameters that we recover from GW observations.
-In this section, we study the impact of the same for binaries where the BH spins
-are aligned to the orbital angular momentum.
-
-\begin{enumerate}
-\item Above what SNR values, below what mass-ratios, above what BH spins, etc, 
-do we begin to care about NS deformability?\newline
-``one interesting plot to make would be the parameter bias (max-Likelihood
-parameters minus true parameters) vs SNR, as functionals of q/sBH/Lambda, when
-using T signals recovered with N templates.''\newline
-`` another interesting plot to make would be the fitting-factors recovered with
-a discrete (search oriented) template bank of BBH waveforms, for NSBH signals of
-loudness, deformability, spins, mass-ratios, etc. Which parameter plays the 
-dominant role in paramter regions where the bank does not suffice to recover
-the desired threshold of optimal SNR?''
-\end{enumerate}
-
-\prayush{Could have subsections for different frequency-optimized aLIGO noise curves ?}
-
-\hspace{5mm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}

### 2016-02-03 LAL/TidalTest_MP.py
Fix bugs in C implementation and update patch

@@ -1,0 +1,232 @@
+#!/usr/bin/env python
+
+import os
+import sys
+import matplotlib.pyplot as plt
+import numpy as np
+
+import lal
+from pycbc.waveform import *
+from pycbc.types import *
+import pycbc.psd as ppsd
+from pycbc.filter import match
+
+class tidalWavs():
+    #{{{
+    def __init__(self, approx='IMRPhenomC', verbose=True):
+        self.verbose = verbose
+        self.approx = approx
+        # Amplitude correction factors
+        # Define constants as per Eq 33 of Lackey et al
+        if 'PhenomC' in approx:
+            self.b0, self.b1, self.b2 = -64.985, -2521.8, 555.17
+            self.c0, self.c1, self.c2 = -8.8093, 30.533, 0.64960
+        elif 'SEOBNR' in approx:
+            self.b0, self.b1, self.b2 = -1424.2, 6423.4, 0.84203
+            self.c0, self.c1, self.c2 = -9.7628, 33.939, 1.0971
+        # Phase correction factors
+        if 'PhenomC' in approx:
+            self.g0, self.g1, self.g2, self.g3 =\
+                    -1.9051, 15.564, -0.41109, 5.7044
+        elif 'SEOBNR' in approx: 
+            self.g0, self.g1, self.g2, self.g3 =\
+                    -4.6339, 27.719, 10.268, -41.741
+    #
+    def mtot_eta_to_m1_m2(self, M, eta):
+        SqrtOneMinus4Eta = np.sqrt(1.-4.*eta)
+        m1 = M * (1. + SqrtOneMinus4Eta) / 2.
+        m2 = M * (1. - SqrtOneMinus4Eta) / 2.
+        return m1, m2
+    # 
+    def tidalCorrectionAmplitude(self, mf, eta, sBH, tidalLambda,\
+                                mfA=0.01):
+        if mf <= mfA: return 1
+        # Impose cutoffs on mass-ratio, and BH spins
+        if eta < 6./49.: 
+            print eta, 6./49.
+            raise IOError("Eta too small")
+        if sBH > 0.75: raise IOError("BH spin too large")
+        if sBH < -0.75: raise IOError("BH spin too small")
[…]

### 2016-02-23 emcee_PE/scripts/make_runs.py
add debug options, and post-processing options

@@ -235,7 +235,7 @@ simstring = sigstring + tmpstring + '_'
 PWD = cmd.getoutput('pwd')
 filename = 'run.sub'
 Nwalkers = [100]
-Nsamples = [150000]
+Nsamples = [2000]
 Nburnin  = 500
 postprocess_only = False
 

### 2016-02-23 emcee_PE/src/emcee_match_tidal.py

@@ -9,9 +9,9 @@ L(s, theta) = exp(+(rho*match(s,h(theta)))^2/2)
 where rho is the desired snr.
 """
 
-__author__ = "Michael Puerrer"
+__author__ = "Michael Puerrer, Prayush Kumar"
 __copyright__ = "Copyright 2015"
-__email__ = "Michael.Puerrer@ligo.org"
+__email__ = "Michael.Puerrer@ligo.org, prkumar@cita.utoronto.ca"
 
 import lal, lalsimulation
 import numpy as np
@@ -75,6 +75,8 @@ parser.add_option("-s", "--snr", dest="SNR", type="float",
                   help="Signal SNR.", metavar="SNR")
 parser.add_option("-r", "--burnin", dest="burnin", type="int",
                   help="How many samples to discard at the beginning of each chain.", metavar="burnin")
+parser.add_option("--burnend", dest="burnend", type="int",
+                  help="How many samples to discard at the end of each chain.", metavar="burnend")
 parser.add_option("-S", "--signal_approximant", dest="signal_approximant", type="string",
                   help="Which approximant to use as a signal.", metavar="signal_approximant")
 parser.add_option("-T", "--template_approximant", dest="template_approximant", type="string",
@@ -116,7 +118,7 @@ parser.add_option("-Y", "--Lambda_stdev_init", dest="Lambda_stdev_init", type="f
 
 parser.set_defaults(nsamples=1000, nwalkers=100, q_signal=4.0, M_signal=100.0, 
   chi2_signal=0.0, f_min=10, f_max=2048, deltaF=0.5, 
-  m1_min=6.0, m1_max=100.0, m2_min=6.0, m2_max=300.0, Mtot_max=500.0, SNR=15, burnin=250, eta_min=0.01,
+  m1_min=6.0, m1_max=100.0, m2_min=6.0, m2_max=300.0, Mtot_max=500.0, SNR=15, burnin=250, burnend=0, eta_min=0.01,
   signal_approximant='lalsimulation.SEOBNRv2_ROM_DoubleSpin', 
   template_approximant='lalsimulation.SEOBNRv2_ROM_DoubleSpin',
   lalsim_psd='lalsimulation.SimNoisePSDaLIGOZeroDetHighPower',
[…]

### 2016-02-23 LAL/TidalTest_MP_PKtest.py
removing file already in git

@@ -1,390 +1,0 @@
-#!/usr/bin/env python
-
-import os
-import sys
-import matplotlib.pyplot as plt
-import numpy as np
-
-import lal
-import lalsimulation as LS
-from pycbc.waveform import *
-from pycbc.types import *
-import pycbc.psd as ppsd
-from pycbc.filter import match
-
-
-class tidalWavs():
-    #{{{
-    def __init__(self, approx='IMRPhenomC', verbose=True):
-        self.verbose = verbose
-        self.approx = approx
-        # Amplitude correction factors
-        # Define constants as per Eq 33 of Lackey et al
-        if 'PhenomC' in approx:
-            self.b0, self.b1, self.b2 = -64.985, -2521.8, 555.17
-            self.c0, self.c1, self.c2 = -8.8093, 30.533, 0.64960
-        elif 'SEOBNR' in approx:
-            self.b0, self.b1, self.b2 = -1424.2, 6423.4, 0.84203
-            self.c0, self.c1, self.c2 = -9.7628, 33.939, 1.0971
-        # Phase correction factors
-        if 'PhenomC' in approx:
-            self.g0, self.g1, self.g2, self.g3 =\
-                    -1.9051, 15.564, -0.41109, 5.7044
-        elif 'SEOBNR' in approx: 
-            self.g0, self.g1, self.g2, self.g3 =\
-                    -4.6339, 27.719, 10.268, -41.741
-    #
-    def mtot_eta_to_m1_m2(self, M, eta):
-        SqrtOneMinus4Eta = np.sqrt(1.-4.*eta)
-        m1 = M * (1. + SqrtOneMinus4Eta) / 2.
-        m2 = M * (1. - SqrtOneMinus4Eta) / 2.
-        return m1, m2
-    # 
-    def tidalCorrectionAmplitude(self, mf, eta, sBH, tidalLambda,\
-                                mfA=0.01):
-        if mf <= mfA: return 1
-        # Impose cutoffs on mass-ratio, and BH spins
-        if eta < 6./49.: 
-            # print eta, 6./49.
-            raise IOError("Eta too small")
-        if sBH > 0.75: raise IOError("BH spin too large")
[…]

### 2016-02-23 LAL/TidalTest_MP_PKtestVersion.py
add version of test script that uses local tidalWavs

@@ -2,6 +2,10 @@
 
 import os
 import sys
+sys.path.append('../emcee_PE/src/')
+from TidalWaveforms import tidalWavs
+print "Imported tidalWavs from ", tidalWavs.__module__
+
 import matplotlib.pyplot as plt
 import numpy as np
 
@@ -13,7 +17,7 @@ import pycbc.psd as ppsd
 from pycbc.filter import match
 
 
-class tidalWavs():
+class local_tidalWavs():
     #{{{
     def __init__(self, approx='IMRPhenomC', verbose=True):
         self.verbose = verbose

### 2016-02-23 LAL/TidalTest_MP_PKtest.py
vv minor change

@@ -377,7 +377,7 @@ for qq in q_range:
   fout.write('%.12e\t%.12e\n' % (qq, m))
   fout.flush()
   matches.append(m)
-  print " .. match = ", m
+  print " .. match = %.12f" % m
       
 fout.close()
 
@@ -387,4 +387,5 @@ semilogy(q_range[:len(mismatchQvary)], mismatchQvary, 'k-o')
 xlabel('MASS-RATIO =m1/m2 (m2=1.35 fixed)')
 ylabel('1 - OVERLAP')
 grid()
+savefig('plot.png')
 ###########################################################

### 2016-02-23 Paper/paper.tex

@@
 \newcommand{\mchirp}{\mathcal{M}_c}
+\newcommand{\LL}{\mathcal{L}}
 
@@
 \subsection{MCMC methods}
-Describe the emcee-based PE code
+% % Describe the emcee-based PE code
+
+The monte carlo algorithm samples the parameter space in proportion with 
+the likelihood $\LL(\vec{\theta}) = p(d|\vec{\theta})$, with a 
+uniform prior in parameters $p(\vec{\theta}) = \Theta(\vec{\theta}-\vec{\theta}_L) + \Theta(\vec{\theta}_H-\vec{\theta})$
+\begin{equation}
+ p(\vec{\theta}|d) = p(d|\vec{\theta}) p(\vec{\theta})
+\end{equation}
 

### 2016-02-23 LAL/TidalTest_MP.py
Fix patch for C implementation; TidalTest_MP_PKtest.py now uses correct tidalWaves class in emcee_PE/src
Simply replacing Mf by MfP in the calls:

  double psiT = tidalPNPhase(MfP, a0, a1, eta);
  double DpsiT= (Mf - MfP) * tidalPNPhaseDeriv(MfP, a0, a1, eta);

@@ -1,321 +1,0 @@
-#!/usr/bin/env python
-
-import os
-import sys
-import matplotlib.pyplot as plt
-import numpy as np
-
-import lal
-import lalsimulation as LS
-from pycbc.waveform import *
-from pycbc.types import *
-import pycbc.psd as ppsd
-from pycbc.filter import match
-
-
-class tidalWavs():
-    #{{{
-    def __init__(self, approx='IMRPhenomC', verbose=True):
-        self.verbose = verbose
-        self.approx = approx
-        # Amplitude correction factors
-        # Define constants as per Eq 33 of Lackey et al
-        if 'PhenomC' in approx:
-            self.b0, self.b1, self.b2 = -64.985, -2521.8, 555.17
-            self.c0, self.c1, self.c2 = -8.8093, 30.533, 0.64960
-        elif 'SEOBNR' in approx:
-            self.b0, self.b1, self.b2 = -1424.2, 6423.4, 0.84203
-            self.c0, self.c1, self.c2 = -9.7628, 33.939, 1.0971
-        # Phase correction factors
-        if 'PhenomC' in approx:
-            self.g0, self.g1, self.g2, self.g3 =\
-                    -1.9051, 15.564, -0.41109, 5.7044
-        elif 'SEOBNR' in approx: 
-            self.g0, self.g1, self.g2, self.g3 =\
-                    -4.6339, 27.719, 10.268, -41.741
-    #
-    def mtot_eta_to_m1_m2(self, M, eta):
-        SqrtOneMinus4Eta = np.sqrt(1.-4.*eta)
-        m1 = M * (1. + SqrtOneMinus4Eta) / 2.
-        m2 = M * (1. - SqrtOneMinus4Eta) / 2.
-        return m1, m2
-    # 
-    def tidalCorrectionAmplitude(self, mf, eta, sBH, tidalLambda,\
-                                mfA=0.01):
-        if mf <= mfA: return 1
-        # Impose cutoffs on mass-ratio, and BH spins
-        if eta < 6./49.: 
-            # print eta, 6./49.
-            raise IOError("Eta too small")
-        if sBH > 0.75: raise IOError("BH spin too large")
[…]

### 2016-02-23 LAL/TidalTest_MP_PKtest.py

@@ -12,134 +12,9 @@ from pycbc.types import *
 import pycbc.psd as ppsd
 from pycbc.filter import match
 
-
-class tidalWavs():
-    #{{{
-    def __init__(self, approx='IMRPhenomC', verbose=True):
-        self.verbose = verbose
-        self.approx = approx
-        # Amplitude correction factors
-        # Define constants as per Eq 33 of Lackey et al
-        if 'PhenomC' in approx:
-            self.b0, self.b1, self.b2 = -64.985, -2521.8, 555.17
-            self.c0, self.c1, self.c2 = -8.8093, 30.533, 0.64960
-        elif 'SEOBNR' in approx:
-            self.b0, self.b1, self.b2 = -1424.2, 6423.4, 0.84203
-            self.c0, self.c1, self.c2 = -9.7628, 33.939, 1.0971
-        # Phase correction factors
-        if 'PhenomC' in approx:
-            self.g0, self.g1, self.g2, self.g3 =\
-                    -1.9051, 15.564, -0.41109, 5.7044
-        elif 'SEOBNR' in approx: 
-            self.g0, self.g1, self.g2, self.g3 =\
-                    -4.6339, 27.719, 10.268, -41.741
-    #
-    def mtot_eta_to_m1_m2(self, M, eta):
-        SqrtOneMinus4Eta = np.sqrt(1.-4.*eta)
-        m1 = M * (1. + SqrtOneMinus4Eta) / 2.
-        m2 = M * (1. - SqrtOneMinus4Eta) / 2.
-        return m1, m2
-    # 
-    def tidalCorrectionAmplitude(self, mf, eta, sBH, tidalLambda,\
-                                mfA=0.01):
-        if mf <= mfA: return 1
-        # Impose cutoffs on mass-ratio, and BH spins
-        if eta < 6./49.: 
-            # print eta, 6./49.
-            raise IOError("Eta too small")
-        if sBH > 0.75: raise IOError("BH spin too large")
-        if sBH < -0.75: raise IOError("BH spin too small")
-        # Generate the amplitude factor
-        C = np.exp(self.b0 + self.b1*eta + self.b2 *sBH) \
-            + tidalLambda * np.exp(self.c0 + self.c1*eta + self.c2*sBH)
-        D = 3.
-        B = C * (mf - mfA)**D
[…]

### 2016-02-26 Paper/paper.tex
added discussion on multiple observations of identical sources

@@
 \usepackage[dvipsnames]{xcolor}
+\usepackage{placeins}
 
@@
 \newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}
+\newcommand{\D}{\mathrm{d}}
 \newcommand{\lambdans}{\Lambda_\mathrm{NS}}
@@
 \newcommand{\LL}{\mathcal{L}}
+\newcommand{\deff}{D_\mathrm{eff}}
 
@@
 
-\subsection{MCMC methods}
+\subsection{Bayesian methods}
 % % Describe the emcee-based PE code
@@
 
+\pagebreak
+\newpage
+\FloatBarrier
+\section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
+% 
+% 
+\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
+% 
+% 
+Here we show a crude calculation which tells us that approximately $8$ observations
+of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
+give us similarly accurate a measurement of $\lambdans$ as would a single detection
+with $\rho=50$. We follow the outline in Ref.~\cite{Markakis:2010mp}.
+
+We want to distribute signals in effective distance $\deff$, which is a combination of 
+luminosity distance, inclination angle of the binary's orbital angular momentum
+with respect to the detector, and its various sky angles. The signal strength 
+scales inversely with $\deff$, if it comprises only of the dominant $(2,2)$ mode, allowing
+for various angle-dependent factors to scale out beside the luminosity distance.
+% 
+In order to consider multiple events, lets imagine a population uniformly distributed
+in effective volume, i.e. within a sphere of radius $\deff^\mathrm{max}$. The 
+radius of this sphere is set by the lowest SNR that is distinguishable from noise
+by LIGO searches. Now, dividie the sphere into $I$ shells of equal width. The radius
+of the $i$-th sphere would then be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The 
+measurement uncertainty in $\lambdans$ scales inversely with the SNR, and hence
+directly with the effective distance to the source. Therefore, if we have a 
+measurement error $\sigma_0$ for a source located at $\deff = D_0$, the same error
+for the same source located within the $i$-th shell would be 
+$$
+\sigma_i = \sigma_0 \frac{D_i}{D_0}.
+$$
+Independent measurements of $\lambdans$ for identical sources at different distances
+would have their combined error given by
+\begin{equation}\label{eq:1oversigma}
+\frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
+\end{equation}
+where $N_i$ is the number of sources detected in the $i$-th shell, and is a random
+variable, with its probability proportional to the volume of the shell, i.e.
+$$
+p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
+$$
+Therefore, the expected number of detections in the $i$-th shell would be
+$$
+\langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
+$$
+where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
+to be Poisson distributed around the mean detection rate
+$\mathcal{R}\equiv\langle N\rangle$. Let us assume that we have $\mathcal{R}_d$ detections
+per year per unit effective volume, where $\mathcal{R}_d$ can vary from $0.6-1000$ per
+$\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. Then 
+\begin{equation}
+\langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi D^2 \mathcal{R}_d \D D = \frac{4\pi}{3} \mathcal{R}_d (\deff^\mathrm{max})^3,
+\end{equation}
+and
+\begin{eqnarray}
+ \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi\mathcal{R}_d D^2 }{D^2}\D D\\
+ &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi \mathcal{R}_d \deff^\mathrm{max},
+\end{eqnarray}
+where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
+to an integral. So the root-mean-square averaged measurement error from $N$ sources 
+distributed uniformly in effective volume within a sphere of radius $\deff^\mathrm{max}$
+would be
+\begin{equation}\label{eq:rmsSigmaIdenticalSources}
+ \sigma_{avg} \equiv \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{3\mathcal{R}_d}.
+\end{equation}
+Let us analyze the implications of Eq.~\ref{eq:rmsSigmaIdenticalSources}. It is
+straightforward to deduce that the same measurement certainty as afforded by a single
+observation with a high SNR $\alpha$ can be obtained from $N = \alpha^2/300$ observations
+uniformly distributed in effective volume with SNRs $\geq 10$. E.g., to get to the level 
+of certainty afforded at SNR$=70$, we would need $49/3\approx 16-17$ realistic detections.
+
+There are two main caveats to this calculation, (i) it applies to identical sources, 
+which is astrophysically next to impossible; (ii) the average measurement error 
+$\sigma_{avg}$ has been defined as in the first equality in 
+Eq.~\ref{eq:rmsSigmaIdenticalSources}.
+
+% 
+\subsection{Astrophysical source population}\label{s2:astro_multiple}
+%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-02-29 Paper/paper.tex
stuff about multiple observations

@@
 % 
-Here we show a crude calculation which tells us that approximately $8$ observations
+Here we show a crude calculation which tells us that approximately $8$ (or $16$) observations
 of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
 give us similarly accurate a measurement of $\lambdans$ as would a single detection
-with $\rho=50$. We follow the outline in Ref.~\cite{Markakis:2010mp}.
+with $\rho=50$ (or $\rho=70$).
 
@@
 
-There are two main caveats to this calculation, (i) it applies to identical sources, 
-which is astrophysically next to impossible; (ii) the average measurement error 
-$\sigma_{avg}$ has been defined as in the first equality in 
-Eq.~\ref{eq:rmsSigmaIdenticalSources}.
+There are two main caveats to this rudimentary calculation, which was outlined in 
+Ref.~\cite{Markakis:2010mp}, (i) it applies to identical sources,
+which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
+approximately; (ii) the average measurement error $\sigma_{avg}$ has been defined as in 
+the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}.
 
@@
 % 
+Lets try a respectable Bayesian approach. Imagine we have $N$ stretches of data,
+$d_1, d_2, \cdots, d_N$, each containing a single observation of an NSBH binary. Let $K$
+denote our collective prior knowledge, except for expectations on binary parameters
+themselves, which will enter the following calculation explicitly. Let the gravitational wave
+signal from each binary be characterized by the parameters 
+$\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}$ and $\lambdans$. Note that we have folded
+in the luminosity distance, inclination angle of the binary's orbital angular momentum with
+respect to the plane of the detector, and its sky location angles into a single parameter $\rho$,
+which is the SNR of the received signal. Note also that $\vec{\theta}$ does not include
+$\lambdans$ as the latter would be treated independently. In addition, we assume
+that all observed NS's have the same equation of state, that uniquely determines their
+$\lambdans$ parameter. In other words, we assume that we know its mass perfectly.
+% 
+Using Bayes' theorem, the measured probability distribution function for the $\lambdans$
+parameter is given as
+\begin{equation}\label{eq:lambdaMultiple}
+ p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{\alpha}\prod_{i=1}^N p(\lambdans | d_i, K),
+\end{equation}
+where $p(\lambdans|K)$ is the prior expectation for $\lambdans$, and we have assumed
+that all of the $N$ observations are independent. [Note: $\alpha$ will be calculated 
+soon, it should be a function only of N.]. A priori, we will assume that no particular value
+of $\lambdans$ is more preferred over another, and $\lambdans\leq 4000$, ie 
+$$
+p(\lambdans | K) = \mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right)
+$$
+% 
+The second set of terms $p(\lambdans | d_i, K)$ are the inferred posterior distribution
+functions for $\lambdans$ from \textit{each} observation independently, and are 
+calculated by marginalizing over all other binary parameters, i.e.
+\begin{equation}
+ p(\lambdans | d_n, K) = \int \D\, \vec{\theta} p(\vec{\theta}, \lambdans | d_n, K).
+\end{equation}
+The joint inferred probability distribution of all parameters $\vec{\theta}\cup\{\lambdans\}$ 
+is given by
+\begin{equation}
+ p(\vec{\theta}, \lambdans | d_n, K) = \dfrac{p(d_n|\vec{\theta}, \lambdans, K)\,p(\vec{\theta}, \lambdans | K)}{p(d_n|K)}.
+\end{equation}
+Here, $p(\vec{\theta}, \lambdans | K)$ is the prior probability for binary parameters of taking 
+particular values. In this paper, we use a uniform prior on the individual masses within 
+pre-chosen ranges, as well as a uniform prior on the spin of the black hole. We set neutron star's
+spin identically $=0$, and its tidal deformability parameter $\lambdans$ to be equally likely
+between $[0, 4000]$. The prior probability of obtaining a particular realization of data
+$p(d_n|K)$ is absorbed into the overall normalization. Finally, the first term in the numerator
+$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given stretch of 
+data $d_n$ if we assume that a signal parameterized by $\vec{\theta}\cup\{\lambdans\}$ is buried
+in it, i.e.
+\begin{equation}\label{eq:likelihood}
+ \LL(\vec{\theta}, \lambdans) \equiv p(\vec{\theta}, \lambdans | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+\end{equation}
+where $h\equiv h(\vec{\theta}, \lambdans)$ is our waveform model, and the detector noise
+weighted inner-product $\langle a|b\rangle$ is defined as 
+$$
+\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
+$$
+where $\tilde{a}$ means the Fourier transform of the finite time series $a(t)$, and 
+$S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The definition of $\LL$ in 
+Eq.~\ref{eq:likelihood} assumes that the instrument noise is colored Gaussian.
+
+\prayush{%
+As described in the previous sections, we have measured the $p(\lambdans | d_i, K)$ for a 
+set of different binary parameters. Therefore, the problem of finding the cumulative measurement
+constraint on $\lambdans$ using Eq.~\ref{eq:lambdaMultiple} reduces to being able to choose
+an astrophysically motivated sample set of $N$ binaries over which the second term on the RHS 
+can be computed.}
+
+\prayush{%
+Get posterior measurements for low SNRs, i.e. $\rho=\{10, 15, 20, 25, 30\}$. 
+Fix black hole mass and spins. Compute the measurement certainty for $\lambdans$
+as a function of the number of detections for a population astrophysically distributed 
+in effective distance. Repeat for a few different combinations of black hole masses
+and spins. This would be a sub-set of the overall detected population for NSBH binaries,
+but what fraction of it would this comprise of is something that we can only probe with
+LIGO itself. We do not know the real distribution of low-mass black holes in the field.
+}
 

### 2016-03-31 Paper/paper.tex
updated Mchirp error figure

@@
 %\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}.pdf
+\includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}
 \caption{We show here (Log of) the ratio of the systematic and statistical
@@
 to be Poisson distributed around the mean detection rate
-$\mathcal{R}\equiv\langle N\rangle$. Let us assume that we have $\mathcal{R}_d$ detections
-per year per unit effective volume, where $\mathcal{R}_d$ can vary from $0.6-1000$ per
-$\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. Then 
+$\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
+$\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
+a year per unit effective volume, then
 \begin{equation}
-\langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi D^2 \mathcal{R}_d \D D = \frac{4\pi}{3} \mathcal{R}_d (\deff^\mathrm{max})^3,
+\langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
 \end{equation}
@@
 \begin{eqnarray}
- \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi\mathcal{R}_d D^2 }{D^2}\D D\\
- &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi \mathcal{R}_d \deff^\mathrm{max},
+ \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
+ &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
 \end{eqnarray}
@@
 \begin{equation}\label{eq:rmsSigmaIdenticalSources}
- \sigma_{avg} \equiv \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{3\mathcal{R}_d}.
+ \sigma_{avg} \equiv \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}}.
 \end{equation}

### 2016-03-31 Paper/paper.tex
good quality parameter bias figures

@@
 %\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-\includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}
+% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}
+\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
 \caption{We show here (Log of) the ratio of the systematic and statistical
@@
 %\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
+% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
+\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
 \caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
@@
 \centering
-\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
-\includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
+\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
+% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
+% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
 \caption{These figures show the ratio of the systematic and statistical

### 2016-03-31 emcee_PE/scripts/SingleObservation.py

@@ -81,8 +81,8 @@ plt.rcParams.update({\
 # Plotting functions
 def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
         title='', titles=[], levellines=[100,200], levelspacing=0.25,\
-        vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f',\
-        xmin=None, xmax=None, ymin=None, ymax=None,\
+        vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f', lfmt='%.1f',\
+        xmin=None, xmax=None, ymin=None, ymax=None, alpha=0.9,\
         colorbartype='simple', figname='plot.png'):
   """
   Function to plot arbitrary numbers of contour plots in a single figure
@@ -143,6 +143,9 @@ col corresponds to the lower level group
   if vverbose: print "VMIN = %e, VMAX = %e" % (VMIN, VMAX)
   if VMIN > VMAX: raise IOError("Cant fit all data on a single colorbar")
   
+  ## FIXME
+  #VMIN = 0.01
+
   # Now make contour plots
   contours_all = {}
   for idx in range(nrow):
@@ -166,12 +169,12 @@ col corresponds to the lower level group
       CS = ax.contourf( xx, yy, zz,\
               levels=levels, \
               cmap = cm.get_cmap(cmap, len(levels)-1), norm=norm,\
-              alpha=0.9, vmin=VMIN, vmax=VMAX)
+              alpha=alpha, vmin=VMIN, vmax=VMAX)
       contours_tmp = {}
       for lev in levellines:
-        cset = ax.contour(xx, yy, zz, levels=[lev], colors='k', ls='--', linewidths=2, hold="on")
-        for c in cset.collections: c.set_linestyle('dashed')
-        plt.clabel(cset, inline=1, fmt='%.0f', fontsize=12)
+        cset = ax.contour(xx, yy, zz, levels=[lev], colors='k', ls='--', linewidths=4, hold="on")
+        for c in cset.collections: c.set_linestyle('dotted')
+        plt.clabel(cset, colors='r', inline=1, fmt=lfmt, fontsize=16)
         contours_tmp[lev] = cset.collections[0].get_paths()
       if vverbose:
         print "for %s" % titles[idx][jdx], contours_tmp
[…]

### 2016-03-31 Paper/paper.tex
good quality lambda measurement bias figures

@@
 \centering    
-\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
-\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
-\includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
+% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
+\includegraphics[width=1.\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
+\includegraphics[width=1.\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
 % \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
@@
 %
-\begin{figure*}
-\centering    
-\includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
-\includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_m1_m2_chi2_Lambda.pdf}
-\caption{These figure shows the $1-$ and $2-$dimensional 
-\textit{a posteriori} probability distributions 
-for different physical parameters describing a BHNS system including NS matter 
-effects, as measured by a Bayesian Markov-chain Monte Carlo using tidal templates.
-The true parameters of the injected signal are: 
-$\mathcal{M}=1.6406M_\odot, \eta=0.2222, \chibh=\chi_2=0.75$ and $\lambdans=1500$,
-with the injected SNR $\rho=20$. We show this case as typical of those with the largest
-tidal signatures imprinted on the emitted GWs (due to the system being comparable
-mass-ratio with a large-aligned spin on the BH and a fairly deformable NS), yet 
-with low SNR.
-% 
-We note a few things:
-(i) chirp mass is well constrained, with a nearly-Gaussian probability distribution;
-while (ii) the other $3$ less constrained parameters rail against
-prior boundaries. The specific boudary conditions that we run into are (a) for $\eta$,
-we are limited by the requirement that $m_\mathrm{NS}\geq 1.2M_\odot$ 
-of our templates; (b) for BH spin, we are limited by the $\chibh\leq+0.75$ requirement
-that results from a template model limitation~\cite{Lackey:2013axa}; and (c) for 
-$\lambdans$, we
-restrict it to $[0, 4000]$ for the same reason as (b). These prior boundaries
-limit the exploration of the entire likelihood surfaces. While (a) is astrophysically
-justified, given that we have not observed NSs with masses below 
-$1.25M_\odot$, (b) and (c) will be removed in future studies as more evolved template
-models become available. See, also, Fig.~\ref{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}.
-% 
-Finally, we note that the left panel corresponds to the actual sampling parameters,
-while the right panel to derived ones.
-}
-\label{fig:TT_pdf_q2_chi75_L1500_SNR20}
-\end{figure*}
-%
-\begin{figure*}
-\centering    
-% \includegraphics[width=.66\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
-\includegraphics[width=\columnwidth]{plots/TT_logLike_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
-\includegraphics[width=\columnwidth]{plots/TT_logProb_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
-\caption{These figures show the effect of limitation (a) (described in the caption
-of Fig.~\ref{fig:TT_pdf_q2_chi75_L1500_SNR20}) on the likelihood surface portion
-that is actually accessible to our Bayesian inferencing algorithm. On the left we show the
-entire likelihood surface in the $\mbh-\mns$ plane, while on the left is the 
-portion not forbidden by the requirement $\mns\geq 1.2M_\odot$. Similarly, our
-priors on BH spin and NS deformability further curtail the portion of sampled
-parameter space, albeit to much less impact than this mass cut.}
-\label{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}
-\end{figure*}
+% \begin{figure*}
+% \centering    
+% \includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
+% \includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_m1_m2_chi2_Lambda.pdf}
+% \caption{These figure shows the $1-$ and $2-$dimensional 
+% \textit{a posteriori} probability distributions 
+% for different physical parameters describing a BHNS system including NS matter 
+% effects, as measured by a Bayesian Markov-chain Monte Carlo using tidal templates.
+% The true parameters of the injected signal are: 
+% $\mathcal{M}=1.6406M_\odot, \eta=0.2222, \chibh=\chi_2=0.75$ and $\lambdans=1500$,
+% with the injected SNR $\rho=20$. We show this case as typical of those with the largest
+% tidal signatures imprinted on the emitted GWs (due to the system being comparable
+% mass-ratio with a large-aligned spin on the BH and a fairly deformable NS), yet 
+% with low SNR.
+% % 
+% We note a few things:
+% (i) chirp mass is well constrained, with a nearly-Gaussian probability distribution;
+% while (ii) the other $3$ less constrained parameters rail against
+% prior boundaries. The specific boudary conditions that we run into are (a) for $\eta$,
+% we are limited by the requirement that $m_\mathrm{NS}\geq 1.2M_\odot$ 
+% of our templates; (b) for BH spin, we are limited by the $\chibh\leq+0.75$ requirement
+% that results from a template model limitation~\cite{Lackey:2013axa}; and (c) for 
+% $\lambdans$, we
+% restrict it to $[0, 4000]$ for the same reason as (b). These prior boundaries
+% limit the exploration of the entire likelihood surfaces. While (a) is astrophysically
+% justified, given that we have not observed NSs with masses below 
+% $1.25M_\odot$, (b) and (c) will be removed in future studies as more evolved template
+% models become available. See, also, Fig.~\ref{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}.
+% % 
+% Finally, we note that the left panel corresponds to the actual sampling parameters,
+% while the right panel to derived ones.
+% }
+% \label{fig:TT_pdf_q2_chi75_L1500_SNR20}
+% \end{figure*}
+% %
+% \begin{figure*}
+% \centering    
+% % \includegraphics[width=.66\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
+% \includegraphics[width=\columnwidth]{plots/TT_logLike_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
+% \includegraphics[width=\columnwidth]{plots/TT_logProb_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
+% \caption{These figures show the effect of limitation (a) (described in the caption
+% of Fig.~\ref{fig:TT_pdf_q2_chi75_L1500_SNR20}) on the likelihood surface portion
+% that is actually accessible to our Bayesian inferencing algorithm. On the left we show the
+% entire likelihood surface in the $\mbh-\mns$ plane, while on the left is the 
+% portion not forbidden by the requirement $\mns\geq 1.2M_\odot$. Similarly, our
+% priors on BH spin and NS deformability further curtail the portion of sampled
+% parameter space, albeit to much less impact than this mass cut.}
+% \label{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}
+% \end{figure*}
 

### 2016-03-31 emcee_PE/scripts/SingleObservation.py

@@ -50,7 +50,7 @@ ErrThresh=[100, 200]
 # PLOTTING :
 # Function to make parameter bias plots
 ######################################################
-linestyles = ['solid', 'dashed', 'dashdot', 'dashed', 'dotted']
+linestyles = ['dotted', 'dashed', 'solid', 'dashed', 'dotted', 'dashdot', 'dashed', 'dotted']
 #linecolors = ['crimson', 'olivedrab', 'k', 'b', 'm', 'y']
 linecolors = ['crimson', 'darkorange', 'olivedrab', 'royalblue', 'purple', 'k']
 linemarkers= ['', 'x', 'o', '^']
@@ -174,7 +174,7 @@ col corresponds to the lower level group
       for lev in levellines:
         cset = ax.contour(xx, yy, zz, levels=[lev], colors='k', ls='--', linewidths=4, hold="on")
         for c in cset.collections: c.set_linestyle('dotted')
-        plt.clabel(cset, colors='r', inline=1, fmt=lfmt, fontsize=16)
+        plt.clabel(cset, colors='b', inline=1, fmt=lfmt, fontsize=16)
         contours_tmp[lev] = cset.collections[0].get_paths()
       if vverbose:
         print "for %s" % titles[idx][jdx], contours_tmp

### 2016-04-01 Paper/paper.tex
placeholder multiple observation figures
fig captions in S4

@@
 \includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
-\caption{We show here (Log of) the ratio of the systematic and statistical
+\caption{We show here the ratio of systematic and statistical
 measurement uncertainties for the binary chirp mass over the NSBH parameter 
@@
 \includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
-\caption{These figures are similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
+\caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
 with the difference that the quantity in question here is the symmetric 
 mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
-NS tidal effects on GW emission can become the dominant source of error. However,
-for quieter signals with $\rho\leq 30$, we find that not including NS matter 
-effects in parameter estimation templates has minimal effects on the measurement
+not including the effects of NS's tidal deformation on GW emission can become the dominant
+source of error for astrophysical searches with Advanced LIGO. However,
+for quieter signals with $\rho\leq 30$, it will have minimal effects on the measurement
 of $\eta$. We remind the reader that the SNRs here are all single detector values.
@@
 % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
-\caption{These figures show the ratio of the systematic and statistical
-measurement errors for BH spins, for all of our injected cases. The left
-panel corresponds to $1-\sigma$ statistical confidence level, while the right panel
-corresponds to $90\%$ confidence level. We note that below $\rho\approx 30$, ignoring
-tidal effects in templates introduces systematic effects that remain subdominant
-to the statistical measurement uncertainties.
+\caption{This figure shows the ratio of the systematic and statistical
+measurement errors for BH spins, with other attributed identical to 
+Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}.
+Similar to the case of mass parameters, we find that below
+$\rho\approx 30$, ignoring tidal effects in templates introduces minor systematic effects,
+which remain subdominant to the statistical measurement uncertainties.
 }
@@
 \end{enumerate}
-
-\prayush{Could have subsections for different frequency-optimized aLIGO noise curves ?}
 
@@
 \caption{This figure shows the statistical uncertainty in the measurement of
-$\lambdans$, as a percentage of the injected signal's $\lambdans$. In each panel, 
+$\lambdans$, as a percentage of the injected signal's $\lambdans$ value. In each panel,
 the same is shown as a function of the BH mass and spin, keeping the 
 $\lambdans$ and injection's SNR $\rho$ fixed (given in the panel). Each row contains
-panels with the same value of $\lambdans$, while each column contains panels 
-with the same value of $\rho$. Two contours are drawn in each panel demarkating
-that the statistical measurement uncertainty for $\lambda$ is equal to (or twice
-of) its injected $\lambdans$ value, i.e. where we can place a factor of $1-2$ 
-bound on $\lambdans$.
+panels with the same value of $\lambdans$, with $\rho$ increasing from left to right.
+Each column contains panels with the same value of $\rho$, with $\lambdans$ 
+increasing from top to bottom.
+% 
+Contours are drawn in each panel demarkating regions where we can constrain the
+$\lambdans$ parameter well (within a factor of two of the injected value).
 % 
@@
 % \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
-\includegraphics[width=1.\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
-\includegraphics[width=1.\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
+\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
 % \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
@@
 % \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR120_CI90_0.pdf}
-\caption{We demarkate here the regions of BH mass and spin plane where we can 
+\caption{In these figures we highlight the regions of BH mass and spin plane where we can 
 constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
-Panels correspond to the injected SNR, with signal strength increasing
-from left to right and top to bottom. Within each panel, $\lambdans$ measurement uncertainty
-contours are shown as functions of BH mass and spin, with different colors corresponding to the
-injected $\lambdans$. We note that this information is, in principle, contained in 
-Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
-the effect of NS's deformability itself on its measurement. As from Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability of NS matter
+Panels correspond to different signal strength, with $\rho$ increasing from left to right.
+Within each panel, $\lambdans$ measurement uncertainty contours are shown as functions of BH 
+mass and spin, with line-types corresponding to different uncertainty levels, and
+colors showing the injected $\lambdans$.
+% We note that this information is, in principle, contained in 
+% Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
+% the effect of NS's deformability itself on its measurement. 
+As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
+of NS matter
 effects improves with all factors that enhances the signature of the NS's disruption on
@@
 % 
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=1.\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}
+\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
+\caption{HELLO
+}
+\label{fig:TT_Lambda_vs_N_L800_CI90_0}
+\end{figure*}
+%
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L2000.pdf}
+\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1500.pdf}\\
+\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1000.pdf}
+\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L500.pdf}
+\caption{HELLO
+}
+\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}
+\end{figure*}
+%
+% 
+% \begin{figure}
+% \centering    
+% \includegraphics[width=1.\columnwidth]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
+% \caption{HELLO
+% }
+% \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}
+% \end{figure}
+%
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=1.\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
+\includegraphics[width=1.\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
+\caption{HELLO
+}
+\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
+\end{figure*}
+%
+
+% 
+\begin{figure}
+\centering    
+\includegraphics[width=1.\columnwidth]{plots/LambdalErrorBars_vs_Lambda_N49_Log.pdf}
+\caption{HELLO
+}
+\label{fig:TT_Lambda_vs_Lambda_L500_2000_CI90_0_AllInOne}
+\end{figure}
+%
+
 % 

### 2016-04-01 emcee_PE/scripts/SingleObservation.py

@@ -367,7 +367,7 @@ def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='
     'ytick.labelsize':16,\
     'figure.subplot.bottom':0.2,\
     'figure.figsize':figsize, \
-    'savefig.dpi': 300.0, \
+    'savefig.dpi': 500.0, \
     'figure.autolayout': True})
   
   ngrp1 = len(XY.keys())
@@ -378,7 +378,7 @@ def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='
     
   #fig = plt.figure(int(1e7 * np.random.random()), \
   #            figsize=((2.1*gmean*ncol+1.25)*colwidth, 1.2*colwidth*nrow))
-  fig = plt.figure(int(1e7 * np.random.random()), figsize=(1.2*gmean*colwidth, colwidth))
+  fig = plt.figure(int(1e7 * np.random.random()), figsize=(1.2*gmean*colwidth*0.75, 0.75*colwidth))
   # FIrst make all lines
   all_lines = []
   xykeys = XY.keys()
@@ -398,14 +398,14 @@ def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='
   if single_legend:
     harray = [all_lines[i][0] for i in range(ngrp1)]
     harray.extend( [all_lines[0][i] for i in range(ngrp2)] )
-    first_legend = plt.legend(handles=harray, loc=1, ncol=2, framealpha=False, markerfirst=markerfirst[0])
+    first_legend = plt.legend(handles=harray, loc=1, ncol=2, framealpha=False, markerfirst=markerfirst[0], fontsize=10)
     ax = plt.gca().add_artist(first_legend)
   else:
     harray1 = [all_lines[i][0] for i in range(ngrp1)]
     harray2 = [all_lines[0][i] for i in range(ngrp2)]
-    first_legend  = plt.legend(handles=harray1, loc=leg_loc[0], ncol=1, framealpha=False, markerfirst=markerfirst[0])
+    first_legend  = plt.legend(handles=harray1, loc=leg_loc[0], ncol=1, framealpha=False, markerfirst=markerfirst[0], fontsize=10)
     ax = plt.gca().add_artist(first_legend)
-    second_legend = plt.legend(handles=harray2, loc=leg_loc[1], ncol=1, framealpha=False, markerfirst=markerfirst[1])
[…]

### 2016-04-01 Paper/paper.tex
adding captions for multiple obs figs

@@
 \includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
-\caption{HELLO
+\caption{{\it Left}:  Posterior probability distributions for $\lambdans$
+and associated $90\%$ confidence intervals, shown as a function of number of events 
+accumulated $N$.
+{\it Right}: Filled-region plots showing the median and $90\%$ confidence intervals
+for $\lambdans$, as a function of the number of observed events $N$.
+ Shown are two independent filled regions. 
+%  
+ One in green shows the recovered confidence interval
+ for $\lambdans$ (on the left y-axis), normalized by its true value,
+ as a function of the 
+ number of observed events $N$. The recovered median is shown by the
+ green line-circled curve. The pair of horizontal green dashed (dotted)
+ lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
+ the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
+% 
+ The second filled region, in grey, shows the recovered
+ confidence interval for neutron star compactness, without normalizing.
+ The corresponding y-values are shown on the *right* y-axis. The horizontal grey
+ dashed line traces the true value of $\lambdans$, and is what we expect
+ the filled region to approach with increasing number of observations.
+%  
+ Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
 }
@@
 \includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L500.pdf}
-\caption{HELLO
+\caption{Filled-region plots showing the median and $90\%$ confidence intervals
+for $\lambdans$ measurement, as a function of the number of observed events $N$.
+Different panels correspond to different injected $\lambdans$ values, which is shown
+in their title. In each panel are shown two independent filled regions. 
+%  
+ One in green shows the recovered confidence interval
+ for $\lambdans$ (on the left y-axis), normalized by its true value,
+ as a function of the 
+ number of observed events $N$. The recovered median is shown by the
+ green line-circled curve. The pair of horizontal green dashed (dotted)
+ lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
+ the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
+% 
+ The second filled region, in grey, shows the recovered
+ confidence interval for neutron star compactness, without normalizing.
+ The corresponding y-values are shown on the *right* y-axis. The horizontal grey
+ dashed line traces the true value of $\lambdans$, and is what we expect
+ the filled region to approach with increasing number of observations.
+%  
+ Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
 }
@@
 \centering    
+\includegraphics[width=1.\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
 \includegraphics[width=1.\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
-\includegraphics[width=1.\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
-\caption{HELLO
+\caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 
+confidence interval for $\lambdans$, normalized by its true value,
+ as a function of the number of events observed $N$. In the inset, we
+ show the same measurement uncertainty, but not normalized, and on a 
+ logarithmic axis for $N$ as well.
+% 
+ From both the inset and the main plot, we find that for the first $3-4$ events
+ our measurement is prior limited. As $N$ increases, the uncertainty falls 
+ roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
+ (than $1/\sqrt{N}$) narrowing of the confidence interval.
+%  
+ {\it Right}: This figure shows the fractional difference between the median
+ recovered value from the posterior probability distribution for $\lambdans$
+ as a function of $N$. Different curves correspond to different injected values
+ of $\lambdans$, and are shaded from light to dark in direct proportion. 
+ The horizontal dashed brown line marks $10\%$ deviation of the median from the 
+ true value.
+%  
+We find that within $\approx 10$ detections, the median measured value for
+$\lambdans$ would estimate the true value for the star with the remarkable 
+accuracy of $10\%$.
 }
@@
 \includegraphics[width=1.\columnwidth]{plots/LambdalErrorBars_vs_Lambda_N49_Log.pdf}
-\caption{HELLO
+\caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
+as a function of $\lambdans$ itself. The drawn curves are cumulative, in the sense 
+that the $N=n$ curve includes information from all of the first $n$ observed events.
+Those corresponding to the first $N\leq 5$ events are shown in fading shades of red,
+with the ones for higher value of $N$ being shown in grey (going from light to dark in
+proportion to $N$).
+Green dash-dotted lines are contours of $y=x^{-(1+1/5)}$.
+% 
+\prayush{We note that the measurement uncertainty from a population follows the
+equation for the green lines.}
+% 
 }

### 2016-04-01 Paper/paper.tex
add snr threshold plots for single obs
S3 written as first draft

@@
 }
-\label{fig:TNT_BHspinBias_vs_SNR_q23}
+\label{fig:TN_BHspinBias_vs_Lambda_SNR}
 \end{figure*}
@@
 % 
-Advanced LIGO searches and parameter estimation efforts aimed at binaries 
-containing NS and stellar-mass BHs are poised to use waveform models that do not
-include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
-While this is not expected to
-be the dominant source of error at low SNRs, it will likely introduce a
-systematic bias in the physical parameters that we recover from GW observations.
-In this section, we study the impact of the same for binaries where the BH spins
-are aligned to the orbital angular momentum.
+Past and future GW search and source characterization efforts with Advanced LIGO
+have used (or plan to) BH-BH waveform templates for NS-BH mergers, ignoring
+the effect of the tidal deformation of the neutron star due to its companion hole
+on the emitted GWs. In this section we present the first fully Bayesian Monte-Carlo
+study of the effects of this assumption on the recovery of other physical parameters
+of NS-BH sources. We do so by comparing the systematic biases introduced by this 
+assumption to the overall statistical uncertainties in the measurement of binary 
+mass and spin parameters. We find that unless the single-detector SNR of the signal
+exceeds $30$ (for design Advanced LIGO sensitivity), the bias accurued by ignoring
+tidal effects in filtering templates will remain smaller than the statistical 
+uncertainty of the same measurement.
 
-\begin{enumerate}
-\item Above what SNR values, below what mass-ratios, above what BH spins, etc, 
-do we begin to care about NS deformability?\newline
-``one interesting plot to make would be the parameter bias (max-Likelihood
-parameters minus true parameters) vs SNR, as functionals of q/sBH/Lambda, when
-using T signals recovered with N templates.''\newline
-\end{enumerate}
+We first take a set of NSBH binary waveforms, which include the modeling of spin and
+tidal effects. Binary mass-ratio is sampled
+uniformly over the range $q\in[2,5]$ taking $q=\{2,3,4,5\}$ as our test points. The
+mass of the NS is fixed to $\mns=1.35M_\odot$, and therefore knowing $q$ gives $\mbh$.
+The spin of the NS is fixed to zero. The spin of the BH is sampled uniformly over the
+range $\chibh\in[-0.75, +0.75]$, choosing $\chibh=\{-0.5, 0, +0.5, +0.75\}$ as our
+test points. Finally, we vary the deformability parameter uniformly over 
+$\lambdans\in[0,2000]$, choosing $\lambdans=\{500, 1000, 1500, 2000\}$ plus an additional
+$\lambdans=800$ case. Waveform are generated for all possible combinations of these
+parameters, and rescaled to different strengths with $\rho=\{20, 30, 50, 70\}$. For all
+of the resulting signals, we perform a fully Bayesian parameter estimation aimed at
+measuring binary masses and spins from the signal. Our analysis
+templates differ from the signal in only one sense: all tidal terms are set to zero
+in the former, i.e. they correspond to BHBH mergers.
 
-\hspace{5mm}
+In Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR} we show the ratio of the systematic
+bias in the measured binary chirp mass $\mchirp$ versus its true value, to the 
+statistical uncertainty of the measurement itself. $\mchirp$ is the mass combination
+that occurs at leading order in the perturbative expression for GW strain from 
+compact binaries, and is therefore very precisely measurable. Even so, we find that
+for $\rho=30$ the signal is not loud enough for the systematic bias in $\mchirp$
+to exceed a fraction of the statistical uncertainty. When we increase the SNR to 
+$\rho=50$ (middle column), we find that if we have two favorable conditions present:
+(a) that the BH spin be $\gtrsim 0.4$, {\it and} (b) the NS's true deformability
+be large enough that its $\lambdans \gtrsim 1000$, the systematic biases from
+ignoring tidal effects in templates reach between $50-100\%$ of the statistical
+uncertainty. Therefore at this signal strength, including terms involving the neutron
+star's internal structure in waveform models becomes increasingly gainful. At 
+even higher SNRs, say $\rho=70$ (right column), under the same favorable conditions
+as listed above, the systematic bias in $\mchirp$ measurement {\it exceeds} its 
+statistical measurement uncertainty. If the BH is even mildly spinning, say
+$\chibh\gtrsim 0.2$, with the NS mildly deformable, say $\lambdans\gtrsim 800$, 
+we still are affected in our $\mchirp$ measurement at the $50\%$ level by the 
+use of point-particle waveform templates. Finally, we note that the measured (median)
+$\mchirp$ is always higher than its true value. This is reasonable since the tidal 
+deformation of NSs close to merger reduces the GW signal power in high frequencies, so
+it would look like a BH-BH signal of a higher mass (and therefore lower merger or
+peak GW-power frequency).
+
+In the next figure, Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the same
+information (i.e. the ratio of systematic bias to statistical uncertainty in parameter
+measurement) for the symmetric mass-ratio $\eta$. Note that $\eta$
+is bounded from above and below to within $\eta\in[0, 1/4]$. We immediately observe
+that the measured (median) $\eta$ is always lowered below its true value due to the 
+lack of $\lambdans$-dependent terms in the filter template model. This bias, as for 
+the $\mchirp$ bias, decreases the merger frequency for the BH-BH template, which must
+better match the NS-BH merger with its quietened merger and ringdown. At SNR $\rho=30$
+the systematic bias in $\eta$ measurement stays mostly around $10-50\%$ of its statistical
+measurement uncertainty. For the sole case of a very deformable NS with $\lambdans=2000$,
+the missing tidal pieces in template models could cause errors that are comparable to 
+that due to noise (i.e. statistical uncertainties). For a still stronger signal with 
+$\rho=50$, we find that 
+when the following conditions are fulfilled, the measurement of $\eta$ is {\it completely
+compromised} due to template biases: (a) BH mass $\mbh\leq 5M_\odot$, (b) BH spin
+$\chibh\gtrsim +0.4$, and (c) $\lambdans\gtrsim 1500$. Even under more moderate restrictions
+on BH and NS parameters, we find that $\rho=50$ is loud enough to necessitate the 
+use of tidal-modeled templates. This observation is further strengthened if we divert 
+our attention to the third column in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR} that presents
+the case for even higher SNRs ($\rho=70$).
+
+Moving on from mass to spin parameters, we now consider the measurement of spin
+angular momentum of the BH $\chibh$. The ratio of the systematic bias introduced into
+and the statistical uncertainty associated with $\chibh$ measurement is shown in
+Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}. The presentation of information in this 
+figure is identical to that in Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
+and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. A diverging colormap is used because both 
+extremes of the colorbar range point to large systematic biases, while its zero (or
+small) value lies in the middle. For the lowest considered SNR $\rho=30$, $\chibh$
+bias is negligible, being below $50\%$ of its statistical measurement uncertainty.
+For the most deformable NS structure {\it and} with a low-mass BH, the statistical
+bias in $\chibh$ does become an important error. This is furthered at higher SNR, 
+considering $\rho=50$ next (middle column of Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}).
+We find that for low-mass BHs (with masses below $4.5M_\odot$), the systematic 
+measurement bias for BH spin exceeds its statistical uncertainty to become the 
+dominant source of error. Not surprisingly, the systematic bias at both spin 
+extremes pushes towards milder spins, since the prior excludes any more extreme
+spins than $[-0.75, +0.75]$ and that pushes the bulk of the posterior inwards.
+However as the signal's $\lambdans$ is increased from $500\rightarrow 2000$, the 
+low-mass BH-NS binaries with large spins on the BHs show very large systematic biases
+in spin values, that becomes larger than the statistical uncertinty for $\rho\geq30$
+and reaches up to $4\times$ the same quantity as $\rho\rightarrow 70$.
+
+Summarizing these results, we find that irrespective of system parameters, below a
+signal-to-noise ratio of $30$, our measurements of mass and spin parameters of 
+astrophysical BHNS binaries will remain limited by the intrinsic uncertainty due to 
+instrument noise, and not depend on whether we include tidal effects in template
+models. However, when the signal-to-noise ratio exceeds $30$ the systematic bias in
+binary mass and spin measurements become comparable to and exceed the uncertainty
+due to noise. Of the different parameters considered, we find that the measurement 
+of $\eta$ degrades worst (in a relative-error sense) amongst all other parameters
+of BHNS binaries, due to the use of BH-BH waveforms in deciphering an NSBH signal.
 
@@
 \label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
+\end{figure*}
+% 
+\begin{figure*}
+\centering    
+% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
+\includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
+% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
+% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR90_CI90_0.pdf}
+% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR120_CI90_0.pdf}
+\caption{In these figures we highlight the regions of BH mass and spin plane where we can 
+constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
+Panels correspond to different signal strength, with $\rho$ increasing from left to right.
+Within each panel, $\lambdans$ measurement uncertainty contours are shown as functions of BH 
+mass and spin, with line-types corresponding to different uncertainty levels, and
+colors showing the injected $\lambdans$.
+% We note that this information is, in principle, contained in 
+% Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
+% the effect of NS's deformability itself on its measurement. 
+As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
+of NS matter
+effects improves with all factors that enhances the signature of the NS's disruption on
+the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
+set by the detectors).
+}
+\label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
 \end{figure*}

### 2016-04-01 Paper/paper.tex
adding more snr threshold plots

@@
 \newcommand{\lambdans}{\Lambda_\mathrm{NS}}
+\newcommand{\dlambda}{\delta\Lambda}
+
 \newcommand{\chibh}{\chi_\mathrm{BH}}
@@
 spins than $[-0.75, +0.75]$ and that pushes the bulk of the posterior inwards.
-However as the signal's $\lambdans$ is increased from $500\rightarrow 2000$, the 
-low-mass BH-NS binaries with large spins on the BHs show very large systematic biases
-in spin values, that becomes larger than the statistical uncertinty for $\rho\geq30$
-and reaches up to $4\times$ the same quantity as $\rho\rightarrow 70$.
+However as the signal's $\lambdans$ is increased from $500\rightarrow 2000$,
+low-mass BH-NS binaries with large spins on the BH show very large systematic biases
+in spin values, that become larger than the statistical uncertainty for $\rho\geq 30$
+and reach up to $4\times$ the same quantity as $\rho\rightarrow 70$.
 
@@
 astrophysical BHNS binaries will remain limited by the intrinsic uncertainty due to 
-instrument noise, and not depend on whether we include tidal effects in template
+instrument noise, and do not depend on whether we include tidal effects in template
 models. However, when the signal-to-noise ratio exceeds $30$ the systematic bias in
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{figure*}
-%\centering    
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
-%\caption{This figure shows.}
-%\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
-%\end{figure*}
 % 
@@
 \centering    
-% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1000_0_CI90_0}
 \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
-\includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
-% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
-% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR90_CI90_0.pdf}
-% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR120_CI90_0.pdf}
+% \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
 \caption{In these figures we highlight the regions of BH mass and spin plane where we can 
@@
 }
-\label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
+\label{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}
 \end{figure*}
@@
 %
-% \begin{figure*}
-% \centering    
-% \includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
-% \includegraphics[trim={0 0 0 0},width=\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_m1_m2_chi2_Lambda.pdf}
-% \caption{These figure shows the $1-$ and $2-$dimensional 
-% \textit{a posteriori} probability distributions 
-% for different physical parameters describing a BHNS system including NS matter 
-% effects, as measured by a Bayesian Markov-chain Monte Carlo using tidal templates.
-% The true parameters of the injected signal are: 
-% $\mathcal{M}=1.6406M_\odot, \eta=0.2222, \chibh=\chi_2=0.75$ and $\lambdans=1500$,
-% with the injected SNR $\rho=20$. We show this case as typical of those with the largest
-% tidal signatures imprinted on the emitted GWs (due to the system being comparable
-% mass-ratio with a large-aligned spin on the BH and a fairly deformable NS), yet 
-% with low SNR.
-% % 
-% We note a few things:
-% (i) chirp mass is well constrained, with a nearly-Gaussian probability distribution;
-% while (ii) the other $3$ less constrained parameters rail against
-% prior boundaries. The specific boudary conditions that we run into are (a) for $\eta$,
-% we are limited by the requirement that $m_\mathrm{NS}\geq 1.2M_\odot$ 
-% of our templates; (b) for BH spin, we are limited by the $\chibh\leq+0.75$ requirement
-% that results from a template model limitation~\cite{Lackey:2013axa}; and (c) for 
-% $\lambdans$, we
-% restrict it to $[0, 4000]$ for the same reason as (b). These prior boundaries
-% limit the exploration of the entire likelihood surfaces. While (a) is astrophysically
-% justified, given that we have not observed NSs with masses below 
-% $1.25M_\odot$, (b) and (c) will be removed in future studies as more evolved template
-% models become available. See, also, Fig.~\ref{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}.
-% % 
-% Finally, we note that the left panel corresponds to the actual sampling parameters,
-% while the right panel to derived ones.
-% }
-% \label{fig:TT_pdf_q2_chi75_L1500_SNR20}
-% \end{figure*}
-% %
-% \begin{figure*}
-% \centering    
-% % \includegraphics[width=.66\columnwidth]{plots/TT_q2-00_mNS1-35_chiBH0-75_Lambda1500-0_SNR20-0_corner_eta_Mc_chi2_Lambda.pdf}
-% \includegraphics[width=\columnwidth]{plots/TT_logLike_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
-% \includegraphics[width=\columnwidth]{plots/TT_logProb_vs_m1_m2_eta0-22_Mc1-64_chi0-00_chi0-50_L1000-0_SNR30.pdf}
-% \caption{These figures show the effect of limitation (a) (described in the caption
-% of Fig.~\ref{fig:TT_pdf_q2_chi75_L1500_SNR20}) on the likelihood surface portion
-% that is actually accessible to our Bayesian inferencing algorithm. On the left we show the
-% entire likelihood surface in the $\mbh-\mns$ plane, while on the left is the 
-% portion not forbidden by the requirement $\mns\geq 1.2M_\odot$. Similarly, our
-% priors on BH spin and NS deformability further curtail the portion of sampled
-% parameter space, albeit to much less impact than this mass cut.}
-% \label{fig:TT_logLikeProb_q2_chi75_L1500_SNR20}
-% \end{figure*}
+% 
+In the previous section we showed that we begin to care for the effects of NS's
+deformation due to its companion hole's tidal field, if (a) the BH is sufficiently
+small, (b) BH spin is aligned and $\gtrsim +0.4$, and (c) the NS is
+not very compact, with $\lambdans\gtrsim 1000$. The first two conditions
+enhance tidal effects on emitted GWs by increasing the number of orbits at small
+separation, where the deviation of NSBH and BHBH dynamics is the largest. The third,
+i.e. (c), simply makes the NS more easily deformable. All of them reduce the onset 
+frequency for NS's disruption close to merger. Therefore, systems that fulfil the 
+above conditions have considerably (or, at least, the most) different GW morphology
+when compared to equivalent BHBH coalescences. Ultimately, the extent of 
+tidal modulations of emitted GWs depends on the equation of state of neutron star
+matter, which becomes an indirect measurable in GW astrophysics.
+% 
+However, if we do indeed receive GWs from such a binary source, what information can we
+possibly extract from it about the internal structure of neutron stars?
+In this section we consider this question in perspective of measuring the tidal 
+deformability parameter of neutron stars purely from gravitational waves they emits 
+while orbiting spinning black holes.
+% 
+We find that when the BH is sufficiently spinning in the prograde sense,
+with $\chibh\gtrsim +0.7$, even at a relatively possible signal strengths ($\rho=20$),
+we can begin to constraint $\lambdans$ at the level of $\pm 75\%$ errors with a $90\%$
+confidence level. If the signal is stronger, say $\rho=30$, we can contraint
+$\lambdans$ better with only $\pm 50\%$ errors, for binaries with large BH spins.
+This trend continues as we vary the SNR from $\rho=30-50$ upwards. In other words,
+with a single but moderately loud NSBH signal observation, Advanced LIGO will begin
+to put constraints on NS matter properties. These constraints can consequently
+be used to assess the likelihood of different equations of state for nuclear
+matter, and possibly narrow the range they span.
+
+  `
+
+We use the same set of systems as in the previous section, with $q=\{2,3,4,5\}$,
+$\chibh=\{-0.5,0,+0.5,+0.75\}$, $\lambdans=\{500, 800, 1000, 1500, 2000\}$ and SNRs
+$\rho=\{20,30,50,70\}$. For each unique combination of these parameters, we obtain a
+signal waveform using the tidal model of Lackey et al~\cite{Lackey:2013axa}.
+We use the same waveform model as input to our Bayesian parameter estimation
+method in order to obtain estimates for different binary parameters alongwith 
+their full probability distributions, including for the combined NS tidal
+deformability parameter $\lambdans$.
+
+The precision with which $\lambdans$ can be measured is given by the width of 
+its $90\%$ confidence interval, as a fraction of the true value. 
+In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the same as a function
+of binary parameters and signal strength. Each panel shows the percent statistical
+uncertainty $\dlambda$ in the measurement of $\lambdans$ as a function of BH 
+masses and spins, with NS parameters and signal strength being held fixed. 
+In each row, the signal strength is increased from
+left to right with $\rho = 20\rightarrow 70$. In each column, the deformability
+of the NS is increased from top to bottom, with $\lambdans=500\rightarrow 2000$.
+% 
+We find that at SNRs around $\rho=30$, we begin to constrain $\lambdans$ 
+meaningfully with $\dlambda\approx 100-150\%$. Amongst other parameters, 
+the BH spin and mass play a big role. A smaller BH with a larger spin always
+allows for a more precise measurement on $\lambdans$. Visually we can see that
+in each panel the bottom right corner, which corresponds to low-mass BHs
+with large spins, is the region of smallest measurement errors on $\lambdans$.
+The actual deformability of the NS also plays an important role on its own
+measurability. For e.g., when $\lambdans\leq 1000$, it is fairly difficult
+to meaningfully constrain $\lambdans$ without getting remarkably lucky and
+finding such a binary with $\rho\gtrsim 50$.
+
+In Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we take the case of
+faily deformable neutron stars, with actual $\lambdans = 1500$ (left panel) and
+$2000$ (right panel). In each case, we calculate the lowest SNR value $\rho_c$ 
+in order
+for the signal to have enough information that we can constrain $\dlambda$ below
+$100\%$. Note that this is a modest requirement on constrainability of $\lambdans$.
+In each panel of the figure, we show $\rho_c$ as a function of BH mass and spin.
+% 
+We immediately note that to fulfil this modest aim with a single observation, 
+we need an SNR of $30$ or higher in each detector. Given we have that, we can 
+bring $\dlambda$ below $100\%$ for the lowest BH masses and higher spins. To
+attain the same for other values of BH parameters, we will need a very loud
+signal with $\rho\gtrsim 50$.
 
 % 
-%% 
-%\begin{figure*}
-%\centering    
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
-%\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
-%\caption{This figure shows.}
-%\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
-%\end{figure*}
+% 
+% \subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}
+%
 
-Having shown in the previous section that we begin to care for NS tidal effects
-for signals in the XXX corner of the paraemter space, with SNRs above YY, here
-we investigate the effect of the improvement in the accuracy of the recovered
-parameters when tidal effects are included in the templates.
-
-\begin{enumerate}
-\item Where in the parameter space can we actually make a statement about 
-$\Lambda_\mathrm{NS}$ ?
-\item What is the reduction in the bias of the maximum likelihood parameters
-when using tidal templates?\newline
-``Plot the ratio of the bias between N templates and T templates (against T
-signals), as a function of SNR.''
-\item What is the reduction in the uncertainty in binary mass and spin, if
-any?\newline
-``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
-when we go from using N templates to T templates.''
-\end{enumerate}
-
-\subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}
-
-With the enhanced sensitivity of aLIGO, we are likely to be able to measure or
-constrain deviations in the observed GW signals, especially close to binary
-merger, from their expected form under the point-particle approximation.
-The tidal disruption of the NS close to merger, and outside of the 
-inner-most circular orbit, will drastically change the morphology of the emitted
-GW signal, the measurement of which will allow us to constrain the equation 
-of state (EoS) of nuclear matter.
-The earlier in the orbit the NS disrupts, the sharper is the drop in the emitted
-GW signal for frequencies above the disruption frequency. In this section we
-study the ability of aLIGO to constrain the 
-$\lambdans\sim \left(R/M\right)_\mathrm{NS}^5$ parameter, which
-the combination that parametrizes the deviation in the emitted GW signal.
-Past studies~\cite{}
-have studied a limited sample of systems with distinguishability criterion that
-are typically applicable in the very high signal-to-noise ratio cases. 
-In what follows, we will use stochastic parameter estimation algorithms to show
-aLIGO's capability of extracting EoS information out from GW signals with 
-more realistic (likely) SNRs.
-
-As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
-fixed at $1.35M_\odot$. We allow BH spin to take the values 
-$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
-parameter combinations, we inject a tidal-corrected waveform using the 
-Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
-posterior probability distribution for different binary parameters with (a) BHBH
-templates, and (b) NSBH templates. We will discuss the results of (b) here.
+% As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
+% fixed at $1.35M_\odot$. We allow BH spin to take the values 
+% $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
+% parameter combinations, we inject a tidal-corrected waveform using the 
+% Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
+% posterior probability distribution for different binary parameters with (a) BHBH
+% templates, and (b) NSBH templates. We will discuss the results of (b) here.
 

### 2016-04-01 Paper/paper.tex
first draft of text for S4

@@
 % \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
-\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
-\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR50_CI90_0.pdf}\\
+\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}\\
 % \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
@@
 In Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we take the case of
-faily deformable neutron stars, with actual $\lambdans = 1500$ (left panel) and
-$2000$ (right panel). In each case, we calculate the lowest SNR value $\rho_c$ 
-in order
-for the signal to have enough information that we can constrain $\dlambda$ below
-$100\%$. Note that this is a modest requirement on constrainability of $\lambdans$.
-In each panel of the figure, we show $\rho_c$ as a function of BH mass and spin.
+moderately deformable neutron stars, with actual $\lambdans = 1000$ (left 
+panel) and $1500$ (right panel), and ask how loud such a signal will have
+to be for us to put meaningful constraints on the tidal deformability of NSs.
+In the left panel, as a function BH mass and spin, 
+we show the lowest SNR value $\rho_c$ in order for the signal to have enough
+information that we can constrain $\dlambda$ below $200\%$. Note that 
+being able to constrain $\lambdans$ within a factor of $2$ interval 
+(i.e. within $\pm 100\%$) is a fairly modest requirement. We find promising 
+results, which suggest that there is a fairly wide range of BH spins, less
+wide (but non-trivial) range of BH masses, within
+which if we get a moderately loud NSBH signal with $\rho\approx 20$, we constrain
+$\lambdans$ within a factor of $2$ interval (i.e. within $\pm 100\%$) around
+the true value. While this is not very precise, it can possibly allow us to
+rule out candidates for the equation of state of neutron star matter.
 % 
-We immediately note that to fulfil this modest aim with a single observation, 
-we need an SNR of $30$ or higher in each detector. Given we have that, we can 
-bring $\dlambda$ below $100\%$ for the lowest BH masses and higher spins. To
-attain the same for other values of BH parameters, we will need a very loud
-signal with $\rho\gtrsim 50$.
+If we want to constrain $\lambdans$ better with a single observation, however,
+we need SNRs of $30$ or higher in each detector. This can be seen from the right
+panel of Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}, which is similar
+to the left but for a higher $\lambdans (=1500)$ system, with a more stringent
+constraint requirement that $\lambdans$ be measured within a factor of $1$ of 
+its true value (i.e. with $\pm 50\%$ errors, instead of $\pm 100\%$). Even for
+the most conducive cases, we {\it need} the SNR to be above $30$ to reach our 
+accuracy threshold (which is more than $3\times$ less likely to happen than
+getting a signal with $\rho=20$).
 
-% 
-% 
-% \subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}
-%
+Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
+conditions that BH parameters must fulfil beside having an SNR of $20$ (left 
+panel) or $30$ (right panel), to obtain meaningful contraints on $\lambdans$.
+At $\rho=20$, the left panel shows that we need high BH spins 
+$\chibh\gtrsim +0.6$ and a very deformable NS with $\lambdans\geq 1500$, in 
+order to put a factor of $2$ bound on $\lambdans$. At $\rho=30$, with
+$\lambdans\gtrsim 1000$ and $\chibh\geq+0.6$ we can put a similar factor of
+$2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
+NS with $\lambdans=200$ one could even put a $\pm 50\%$ errorbar on 
+$\lambdans$ measurement.
 
-% As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
-% fixed at $1.35M_\odot$. We allow BH spin to take the values 
-% $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
-% parameter combinations, we inject a tidal-corrected waveform using the 
-% Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
-% posterior probability distribution for different binary parameters with (a) BHBH
-% templates, and (b) NSBH templates. We will discuss the results of (b) here.
+In summary, with a single moderately loud BHNS observation, we could constrain
+the NS compactness parameter $\lambdans$ within $\pm 100\%$ of its true value.
+To measure better with one observation, we will need to get lucky with signals
+with $\rho\geq 30$ and high BH spins, or $\rho\geq 50$.
 
-\pagebreak
-\newpage
-\FloatBarrier
+% \pagebreak
+% \newpage
+% \FloatBarrier
 \section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}

### 2016-04-01 Paper/paper.tex
cleaning up text in S5

@@
 %
-
 % 
-\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
+\textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
 % 
 % 
-Here we show a crude calculation which tells us that approximately $8$ (or $16$) observations
+First we show a crude calculation which tells us that approximately $8$ (or $16$) 
+observations
 of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
@@
 radius of this sphere is set by the lowest SNR that is distinguishable from noise
-by LIGO searches. Now, dividie the sphere into $I$ shells of equal width. The radius
+by LIGO searches. Now, divide the sphere into $I$ shells of equal width. The radius
 of the $i$-th sphere would then be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The 
@@
 for the same source located within the $i$-th shell would be 
-$$
-\sigma_i = \sigma_0 \frac{D_i}{D_0}.
-$$
+$\sigma_i = \sigma_0 \dfrac{D_i}{D_0}.$
 Independent measurements of $\lambdans$ for identical sources at different distances
@@
 \end{equation}
-where $N_i$ is the number of sources detected in the $i$-th shell, and is a random
-variable, with its probability proportional to the volume of the shell, i.e.
-$$
-p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
-$$
-Therefore, the expected number of detections in the $i$-th shell would be
-$$
-\langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
-$$
-where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
-to be Poisson distributed around the mean detection rate
-$\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
-$\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
-a year per unit effective volume, then
-\begin{equation}
-\langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
+where $N_i$ is the number of sources detected in the $i$-th shell.
+% , and is a random
+% variable, with its probability proportional to the volume of the shell, i.e.
+% $$
+% p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
+% $$
+% Therefore, the expected number of detections in the $i$-th shell would be
+% $$
+% \langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
+% $$
+% where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
+% to be Poisson distributed around the mean detection rate
+% $\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
+% $\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
+% a year per unit effective volume, then
+% \begin{equation}
+% \langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
+% \end{equation}
+% and
+% \begin{eqnarray}
+%  \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
+%  &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
+% \end{eqnarray}
+% where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
+% to an integral. 
+Continuing on this line of inquiry, Ref.~\cite{Markakis:2010mp} calculates the 
+root-mean-square averaged measurement error from 
+$\langle N\rangle:=\langle\sum N_i\rangle$ sources distributed uniformly in
+effective volume within a sphere of radius $\deff^\mathrm{max}$ to be
+\begin{equation}\label{eq:rmsSigmaIdenticalSources}
+ \sigma_{avg} := \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}},
 \end{equation}
-and
-\begin{eqnarray}
- \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
- &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
-\end{eqnarray}
-where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
-to an integral. So the root-mean-square averaged measurement error from $N$ sources 
-distributed uniformly in effective volume within a sphere of radius $\deff^\mathrm{max}$
-would be
-\begin{equation}\label{eq:rmsSigmaIdenticalSources}
- \sigma_{avg} \equiv \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}}.
-\end{equation}
-Let us analyze the implications of Eq.~\ref{eq:rmsSigmaIdenticalSources}. It is
-straightforward to deduce that the same measurement certainty as afforded by a single
+given a fiducial pair $(\sigma_0, D_0)$.
+Further, it is straightforward to deduce from Eq.~\ref{eq:rmsSigmaIdenticalSources}
+that the same measurement certainty as afforded by a single
 observation with a high SNR $\alpha$ can be obtained from $N = \alpha^2/300$ observations
@@
 which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
-approximately; (ii) the average measurement error $\sigma_{avg}$ has been defined as in 
-the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}.
+approximately; (ii) the combined average measurement error $\sigma_{avg}$ has been defined
+as in the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}. To overcome both,
+we next perform a Bayesian population analysis.
 
+\textbf{Astrophysical source population: }\label{s2:astro_multiple}
 % 
-\subsection{Astrophysical source population}\label{s2:astro_multiple}
-% 
-Lets try a respectable Bayesian approach. Imagine we have $N$ stretches of data,
+Imagine we have $N$ stretches of data,
 $d_1, d_2, \cdots, d_N$, each containing a single observation of an NSBH binary. Let $K$

### 2016-04-01 Paper/paper.tex
added error level lines in CI vs N figure

@@
 \centering    
-% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
 \includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
-\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}\\
-% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR70_CI90_0.pdf}
-% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR90_CI90_0.pdf}
-% \includegraphics[width=.68\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR120_CI90_0.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
 \caption{In these figures we highlight the regions of BH mass and spin plane where we can 
@@
 
-% \pagebreak
-% \newpage
-% \FloatBarrier
 \section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
@@
 \centering    
-\includegraphics[width=1.\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
-\includegraphics[width=1.\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
+\includegraphics[width=1.025\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
 \caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 
@@
 that the same measurement certainty as afforded by a single
-observation with a high SNR $\alpha$ can be obtained from $N = \alpha^2/300$ observations
+observation with a high SNR $\rho_c$ can be obtained from $N = \rho_c^2/300$ observations
 uniformly distributed in effective volume with SNRs $\geq 10$. E.g., to get to the level 

### 2016-04-03 emcee_PE/scripts/MultipleObservations.py
fixed bug in population generation code, and plots

@@ -164,6 +164,17 @@ the observable volume, i.e. 4/3 pi D_Luminosity^3
     raise RuntimeError("SNR sampling not being done properly")
         
 
+def find_closest_match(tmp_rand, vec):
+    '''
+Take a random number in the range spanned by vec, and find the
+element in vec closest to it.
+    '''
+    diffs = np.abs(np.array(vec) - tmp_rand)
+    idx_min = np.where(diffs == diffs.min())[0][0]
+    if idx_min < 0 or idx_min >= len(vec):
+        raise RuntimeError("Could not find closest match!")
+    return vec[idx_min]
+    
 #######################################################
 ### CALCULATION CLASSES
 #######################################################
@@ -252,7 +263,8 @@ What it does:
     ###
     def generate_events(self, lambda_posterior_chains=None,\
                         NSLambda=None,\
-                        qvec=None, chi2vec=None, SNRvec=None):
+                        qvec=None, chi2vec=None, SNRvec=None,\
+                       qmin=2.0, qmax=5.0, chi2min=0.0, chi2max=1.0):
         '''
         Returns is a list of events. For each event, two objects are returned:
         1. array of posterior samples
@@ -280,8 +292,8 @@ What it does:
                 print >>sys.stdout, "  Event %d" % i
                 sys.stdout.flush()
             # DRAW TWO uniformly distributed random samples for (q, chiBH)
-            rnd_q = qvec[np.int(np.random.uniform(0, len(qvec)))]
-            rnd_chiBH = chi2vec[np.int(np.random.uniform(0, len(chi2vec)))]
+            rnd_q = find_closest_match(np.random.uniform(qmin, qmax), qvec)
+            rnd_chiBH = find_closest_match(np.random.uniform(chi2min, chi2max), chi2vec)
             #
             # Sample the SNR depending on source distribution specified
             if 'UniformInVolume' in self.source_distribution:

### 2016-04-04 Paper/paper.tex
adding text to describe LEA

@@
 
+\newcommand{\ii}{\mathrm{i}}
 \newcommand{\chibh}{\chi_\mathrm{BH}}
@@
 We attempt to answer the following questions:
-\begin{enumerate}
- \item What systematic biases are introduced by neglecting the NS tidal
- disruption close to merger? 
- \item At what SNR do the systematic biases dominate over the statistical 
- errors in the recovery of binary parameters?
- \item By using templates which models NS deformation (over inspiral) and 
- NS disruption (close to merger), what information can we recover about
- the $\Lambda_\mathrm{BH}\sim (R/M)^5_\mathrm{NS}$ parameter?
-\end{enumerate}
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
+\section{Introduction}\label{s1:introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\vspace{2cm}
+Finally, we note that our results are applicable for the design sensitivity
+LIGO instruments, that are expected to come online in $2018-19$. Also note
+that more a recent work that improves upon the waveform model we use in this
+study has been published~\cite{Pannarale:2015jka}, but it gives only an 
+amplitude model and needs to be (in future) augmented with a compatible phase
+model. In addition, accuracy of the underlying numerical simulations used to
+calibrate the waveform model used here has not been verified against 
+independent codes so far.
+It is therefore difficult to assess the combined modeling error and the 
+results presented in our study are limited by the limitations of the model.
+However, we expect the combined effect of modeling errors to not change our
+qualitative conclusions.
 
-\textit{\prayush{Current status:}}
+The remainder of the paper is organized as follows. 
+Sec.~\ref{s1:techniques} discusses data analysis techniques and resources 
+used in this paper, such as the waveform model, and parameter estimation 
+algorithm.
+Sec.~\ref{s1:PEwithnoNS} discusses the consequences of ignoring tidal 
+effects in parameter estimation waveform models.
+Sec.~\ref{s1:PEwithNS} discusses the measurability for the leading order
+tidal parameter $\lambdans$ at plausible SNR values.
+Sec.~\ref{s1:multiple_observations} discusses the improvement in our
+measurement of $\lambdans$ with successive (multiple) observations of
+NSBH mergers.
+Finally, in Sec.~\ref{s1:discussion} we summarize our results and discuss
+future prospects with Advanced LIGO.
 
-Below we explain the outline the different calculations performed in this 
-paper:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Techniques}\label{s1:techniques}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Waveform Models}\label{s2:waveforms}
 
-\begin{enumerate}
-\item We choose the following values of different NSBH binary parameters:
-\begin{enumerate}
-\item $m_\mathrm{NS}=1.35M_\odot$;
-\item $q=m_\mathrm{BH}/\mathrm{NS}=\{2,3\}$;
-\item $\chi_\mathrm{NS}=0$;
-\item $\chi_\mathrm{BH}=\{-0.5, 0, +0.5\}$;
-\item $\Lambda =\{500,1000,2000\}$ ($\Lambda=0\implies \mathrm{BHBH}$).
-\end{enumerate}
+In Ref.~\cite{Lackey:2013axa}, the authors presented a suite of $134$ numerical
+relativity simulations of neutron stars inspiraling into spinning black holes. Amongst
+the varied quantities were (a) the equation of state for the NS, (b) NS mass, 
+(c) mass-ratio and (d) black hole spins, taken as aligned with the orbit.
+A total of $21$ two-parameter EoSs were used for these simulations, which also 
+had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
+$m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
+authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
+NSBH coalescences. The GW strain predicted by this model (henceforth ``LEA''),
+can be written as
+% 
+\begin{equation}
+ \tilde{h}_\mathrm{NSBH}(f, \vec{\theta}) = \tilde{h}_\mathrm{BBH}(f, \vec{\theta} - \{\lambdans\})\,A(f, \vec{\theta})\,e^{\ii \Delta\Phi(f, \vec{\theta})},
+\end{equation}
+where $\vec{\theta}:=\{M, \eta, \chibh, \lambdans\}$ are the intrinsic binary
+parameters. The first term on the right hand side is an underlying
+BBH waveform model that is used as a base. In the original construction of LEA,
+this was taken to be the SEOBNRv1 model~\cite{Taracchini:2012} of the 
+Effective-one-body (EOB) family~\cite{Buonanno99}. The second term changes the 
+amplitude of the BBH model
+to match that of an NSBH merger of otherwise identical parameters, with NS-matter
+effects parametrized by $\lambdans$. During inspiral this term is set to
+unity, but is a sensitive function of $\lambdans$ close to merger. The third 
+term does the same as the second, but to the waveform phasing. During inspiral,
+$\Delta\Phi(f, \vec{\theta})$ is set to the PN tidal phasing corrections,
+at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
+additional phenomenological terms are added and calibrated to the $134$ available
+NR simulations.
 
-\item For each system, corresponding to one of all possible unique combinations
-of the above parameter choices, we carry out three parameter estimation tests: 
-\begin{enumerate}
-\item BH-BH signals injected, recovered with BH-BH waveforms (control),
-\item NS-BH signals injected, recovered with BH-BH waveforms, (current approach) and
-\item NS-BH signals injected, recovered with NS-BH waveforms.
-\end{enumerate}
-Injections were made at different SNR values: 
-\begin{enumerate}
-\item $\rho = \{20, 30, 50, 70, 90, 120\}$.
-\end{enumerate}
-Choice of emcee parameters:
-\begin{enumerate}
-\item $N_\mathrm{samples}=150,000$; 
-\item $N_\mathrm{walkers}=100$; 
-\item $N_\mathrm{burn-in}=500$; 
-\end{enumerate}
-and choice of prior boundaries:
-\begin{enumerate} 
-\item $m_\mathrm{BH}\in [1.2, 25]M_\odot$;
-\item $m_\mathrm{NS}\in [1.2, 15]M_\odot$;
-\item $\Lambda\leq 4000$; 
-\item $\sigma\left(\Lambda\right) = 100$ (for chains where templates are NS-BH waveforms).
-\end{enumerate}
+In this paper we use LEA for our signal and template modeling, but switch the 
+underlying BBH model to SEOBNRv2~\cite{Taracchini:2013rva},
+using its reduced-order frequency-domain version~\cite{Purrer:2015tud}. We
+expect this change to make our conclusions more robust because: (a) the 
+SEOBNRv2 model is more accurate~\cite{Kumar:2015tha,Kumar:2016dhh}, and (b)
+the differences between the two EOB models are caused by the
+inaccuracies of SEOBNRv1 during the {\it inspiral} phase, many orbits before 
+merger~\cite{Kumar:2015tha}.
+Since LEA only augments the inspiral phasing with PN tidal terms, our
+change in the underlying BBH model does not change LEA's construction, but
+increases the overall model accuracy during inspiral. Therefore the model used 
+here (``LEA+'') is a more robust version of LEA.
 
-\item Results are shown in
-Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}-~\ref{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}.
-
-\end{enumerate}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Techniques}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Modeling tidal deformation during inspiral}
-Describe the PN terms used here
-
-\subsection{Modeling tidal disruption near merger}
-Describe the Lackey et al model
-
-\subsection{Bayesian methods}
+\subsection{Bayesian methods}\label{s2:bayesian}
 % % Describe the emcee-based PE code
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{How are detection searches affected by not including NS matter effects}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%Near future detection searches with aLIGO are poised to search for NSBH
-%signals using BHBH templates. In this section, we examine the effect of ignoring
-%the effects of tidal deformation and disruption of the NS during inspiral and 
-%merger phases of search templates. We construct a template bank $\mathcal{B}$ 
-%restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
-%$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
-%using a PN-based geometric lattice placement method~\cite{Harry:2013tca}. 
-%
-%
-%Before investigating its SNR recovery against tidal NSBH signals, we test the 
-%effectualness of this bank to point-particle NSBH signals,
-%in order to make sure that the placement method does not leave any region of
-%the binary parameter-space under-covered. We simulate $50,000$ signals with 
-%component-mass and BH spin values
-%sampled uniformly over their allowed ranges, and filter each of them against
-%the entire bank. For both signals and templates, we use the frequency-domain 
-%reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
-%implemented within the LIGO algorithms Library~\cite{lal}. We record the 
-%fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
-%signals, which measures the fraction of optimal SNR that the bank is able
-%to recover. Fig.~\ref{} shows these FFs as a function of the injected
-%BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
-%values. From these figures, we conclude that our bank is effectual as intended.
-%
-%
-%Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
-%detection searches, we perform a similar calculation as above, with the
-%difference that the signals are now modelled using the model of tidal 
-%corrections to IMR amplitude and phasing published in 
-%Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How is PE affected by not including NS matter effects?}
+\section{How is PE affected by not including NS matter effects?}\label{s1:PEwithnoNS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{What do we gain by using templates that include NS matter effects?}
+\section{What do we gain by using templates that include NS matter effects?}\label{s1:PEwithNS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+\section{Discussion}\label{s1:discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-04-04 emcee_PE/scripts/MultipleObservations.py

@@ -234,6 +234,7 @@ What it does:
                  only_positive_aligned_spins=True,\
                  kernel='gau',\
                  bw_method='scott',\
+                 write_data=True,\
                  verbose=False):
         ### REMOVE BH spins < 0
         if verbose:
@@ -257,6 +258,10 @@ What it does:
         self.bw_method = bw_method
         #
         self.verbose = verbose
+        self.write_data = write_data
+        #
+        # TAG for storing data to disk
+        self.TAG = str(int(np.random.random()*1e5))
         #
         self.print_info()
         return
@@ -280,6 +285,7 @@ What it does:
         if NSLambda == None: NSLambda = self.NSLambda
         # Initiate primary data structure
         chain_set = []
+        chain_params = []
         ###
         ## Loop over as many times as many events are to be generated
         ###
@@ -329,8 +335,84 @@ What it does:
             ## Add the RAW DATA and KDE to the PRIMARY DATA STRUCTURE
             ##
             chain_set.append( [lambda_chain, lambda_kde.evaluate, lambda_kde_norm] )
+            chain_params.append( [rnd_q, rnd_chiBH, rnd_SNR, NSLambda] )
         #
-        self.chain_set = chain_set
+        self.chain_set      = chain_set
+        self.chain_params   = chain_params
+        #
+        if self.write_data:
+          np.savetxt(self.TAG + "_chain_params.dat", chain_params, delimiter='\t')
+        #
+        self.FULL_chain_set = self.chain_set
+        self.generate_cumulative_normalizations() # Normalizations
+        self.print_info()
+        return chain_set
+    #####
+    ###
+    def load_events(self, chain_set,\
+                      lambda_posterior_chains=None,\
+                      NSLambda=None):
+        '''
+        Returns is a list of events. For each event, two objects are returned:
+        1. array of posterior samples
+        2. fitted kernel density estimator's evaluate method
+        
+        Input:
+        chain_set: list of tuples, where each tuple is (q, chi, SNR) for 1 event
+        '''
+        ####
+        if lambda_posterior_chains == None:
+            lambda_posterior_chains = self.lambda_posterior_chains
+        if NSLambda == None: NSLambda = self.NSLambda
+        # Initiate primary data structure
+        chain_params = []
+        ###
[…]

### 2016-04-04 Paper/paper.tex
fleshing out s5

@@
 \end{equation}
-where $\vec{\theta}:=\{M, \eta, \chibh, \lambdans\}$ are the intrinsic binary
+where $\vec{\theta}:=\{\mbh, \mns, \chibh,\chins=0, \lambdans\}$ are the intrinsic binary
 parameters. The first term on the right hand side is an underlying
@@
 
-The monte carlo algorithm samples the parameter space in proportion with 
-the likelihood $\LL(\vec{\theta}) = p(d|\vec{\theta})$, with a 
-uniform prior in parameters $p(\vec{\theta}) = \Theta(\vec{\theta}-\vec{\theta}_L) + \Theta(\vec{\theta}_H-\vec{\theta})$
+The joint inferred probability distribution of all parameters $\vec{\theta}\cup\{\lambdans\}$ 
+is given by
 \begin{equation}
- p(\vec{\theta}|d) = p(d|\vec{\theta}) p(\vec{\theta})
+ p(\vec{\theta}, \lambdans | d_n, K) = \dfrac{p(d_n|\vec{\theta}, \lambdans, K)\,p(\vec{\theta}, \lambdans | K)}{p(d_n|K)}.
 \end{equation}
+Here, $p(\vec{\theta}, \lambdans | K)$ is the prior probability for binary parameters of taking 
+particular values. In this paper, we use a uniform prior on the individual masses within 
+pre-chosen ranges, as well as a uniform prior on the spin of the black hole. We set neutron star's
+spin identically $=0$, and its tidal deformability parameter $\lambdans$ to be equally likely
+between $[0, 4000]$. The prior probability of obtaining a particular realization of data
+$p(d_n|K)$ is absorbed into the overall normalization. Finally, the first term in the numerator
+$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given stretch of 
+data $d_n$ if we assume that a signal parameterized by $\vec{\theta}\cup\{\lambdans\}$ is buried
+in it, i.e.
+\begin{equation}\label{eq:likelihood}
+ \LL(\vec{\theta}, \lambdans) \equiv p(\vec{\theta}, \lambdans | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+\end{equation}
+where $h\equiv h(\vec{\theta}, \lambdans)$ is our waveform model, and the detector noise
+weighted inner-product $\langle a|b\rangle$ is defined as 
+$$
+\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
+$$
+where $\tilde{a}$ means the Fourier transform of the finite time series $a(t)$, and 
+$S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The definition of $\LL$ in 
+Eq.~\ref{eq:likelihood} assumes that the instrument noise is colored Gaussian.
 
@@
 which is the SNR of the received signal. Note also that $\vec{\theta}$ does not include
-$\lambdans$ as the latter would be treated independently. In addition, we assume
-that all observed NS's have the same equation of state, that uniquely determines their
-$\lambdans$ parameter. In other words, we assume that we know its mass perfectly.
+$\lambdans$ as the latter would be treated independently. In addition we assume that all
+NS's in a population will have the same deformability parameter $\lambdans$, and therefore
+its cumulative measurement is possible.
 % 
-Using Bayes' theorem, the measured probability distribution function for the $\lambdans$
-parameter is given as
+Using Bayes' theorem, the measured probability distribution function for $\lambdans$
+from $N$ observations is
 \begin{equation}\label{eq:lambdaMultiple}
- p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{\alpha}\prod_{i=1}^N p(\lambdans | d_i, K),
+ p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{1-N}\prod_{i=1}^N p(\lambdans | d_i, K),
 \end{equation}
 where $p(\lambdans|K)$ is the prior expectation for $\lambdans$, and we have assumed
-that all of the $N$ observations are independent. [Note: $\alpha$ will be calculated 
-soon, it should be a function only of N.]. A priori, we will assume that no particular value
-of $\lambdans$ is more preferred over another, and $\lambdans\leq 4000$, ie 
-$$
-p(\lambdans | K) = \mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right)
-$$
+that all of the $N$ observations are independent. A priori, we assume that no 
+particular value of $\lambdans$ is more preferred over another, and $\lambdans\leq 4000$,
+i.e.
+\begin{equation}\label{eq:lprior}
+ p(\lambdans | K) = \dfrac{1}{4000}\,\mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right)
+\end{equation}
 % 
-The second set of terms $p(\lambdans | d_i, K)$ are the inferred posterior distribution
-functions for $\lambdans$ from \textit{each} observation independently, and are 
-calculated by marginalizing over all other binary parameters, i.e.
+The second set of terms in Eq.~\ref{eq:lambdaMultiple} of the form $p(\lambdans | d_i, K)$
+is each the probability distribution for $\lambdans$ inferred {\it a posteriori} from
+the \textit{i}-th observation independently. 
+They are calculated by integrating
 \begin{equation}
- p(\lambdans | d_n, K) = \int \D\, \vec{\theta} p(\vec{\theta}, \lambdans | d_n, K).
+ p(\lambdans | d_i, K) = \int \D \vec{\theta}\, p(\vec{\theta}, \lambdans | d_i, K),
 \end{equation}
-The joint inferred probability distribution of all parameters $\vec{\theta}\cup\{\lambdans\}$ 
-is given by
-\begin{equation}
- p(\vec{\theta}, \lambdans | d_n, K) = \dfrac{p(d_n|\vec{\theta}, \lambdans, K)\,p(\vec{\theta}, \lambdans | K)}{p(d_n|K)}.
-\end{equation}
-Here, $p(\vec{\theta}, \lambdans | K)$ is the prior probability for binary parameters of taking 
-particular values. In this paper, we use a uniform prior on the individual masses within 
-pre-chosen ranges, as well as a uniform prior on the spin of the black hole. We set neutron star's
-spin identically $=0$, and its tidal deformability parameter $\lambdans$ to be equally likely
-between $[0, 4000]$. The prior probability of obtaining a particular realization of data
-$p(d_n|K)$ is absorbed into the overall normalization. Finally, the first term in the numerator
-$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given stretch of 
-data $d_n$ if we assume that a signal parameterized by $\vec{\theta}\cup\{\lambdans\}$ is buried
-in it, i.e.
-\begin{equation}\label{eq:likelihood}
- \LL(\vec{\theta}, \lambdans) \equiv p(\vec{\theta}, \lambdans | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
-\end{equation}
-where $h\equiv h(\vec{\theta}, \lambdans)$ is our waveform model, and the detector noise
-weighted inner-product $\langle a|b\rangle$ is defined as 
-$$
-\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
-$$
-where $\tilde{a}$ means the Fourier transform of the finite time series $a(t)$, and 
-$S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The definition of $\LL$ in 
-Eq.~\ref{eq:likelihood} assumes that the instrument noise is colored Gaussian.
+where $p(\vec{\theta}, \lambdans | d_n, K)$ is the inferred joint probability 
+distribution of all source parameters $\vec{\theta}\cup\{\lambdans\}$ for the 
+$i$-th event, as described in Sec.~\ref{s2:bayesian}.
 
-\prayush{%
-As described in the previous sections, we have measured the $p(\lambdans | d_i, K)$ for a 
-set of different binary parameters. Therefore, the problem of finding the cumulative measurement
-constraint on $\lambdans$ using Eq.~\ref{eq:lambdaMultiple} reduces to being able to choose
-an astrophysically motivated sample set of $N$ binaries over which the second term on the RHS 
-can be computed.}
+Our goal is to understand the improvement in our measurement of $\lambdans$
+with the number of recorded events. To do so, we simulate populations of $N$
+observations each, and quantify what we learn from each successive observation 
+using Eq.~\ref{eq:lambdaMultiple}. This allows us to determine how
+rapidly does our median estimate for $\lambdans$ converges to the true value,
+and how broad our uncertainty intervals for the same are, as $N$ increases.
+In order to generate each population, we first fix
+the NS properties: (i) NS mass $\mns=1.35M_\odot$, (ii) NS spin $\chins=0$
+and (iii) NS tidal deformability $\lambdans=$ some value that is chosen from
+$\{500,800,1000,1500,2000\}$. Next, we successively generate $N$ 
+independent events, uniformly sampling the remaining
+source parameters from the following ranges: (i) mass-ratio $q\in[2,5]$, (ii)
+BH spin $\chibh\in[0, 0.75]$, (iii) effective distance from within a 
+spherical volume whose radius is given by the lowest detectable
+SNR (which we take as $\rho=10$). We remind ourselves that
+once source parameters are fixed, there is a one-to-one mapping between
+effective distance and SNR. Finally, in order to mitigate computational
+cost, we make an additional simplification.
+We perform full MCMC parameter estimation for a set of systems with parameters
+laid on the regular hypercubic grid (henceforth ``G'') that spans
+$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+The final population is given by substituting all of the $N$ events, generated
+as above, with their respective nearest neighbours on the grid G.
 
-\prayush{%
-Get posterior measurements for low SNRs, i.e. $\rho=\{10, 15, 20, 25, 30\}$. 
-Fix black hole mass and spins. Compute the measurement certainty for $\lambdans$
-as a function of the number of detections for a population astrophysically distributed 
-in effective distance. Repeat for a few different combinations of black hole masses
-and spins. This would be a sub-set of the overall detected population for NSBH binaries,
-but what fraction of it would this comprise of is something that we can only probe with
-LIGO itself. We do not know the real distribution of low-mass black holes in the field.
-}
+In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show results for a population
+of $N=50$ events with $\lambdans=800$. In the left panel, we show the 
+probability distributions for $\lambdans$ inferred from $N$ events with $N$
+ranging from $1-50$. We also mark the $90\%$ confidence interval on $\lambdans$
+as dark grey vertical lines corresponding to each cumulative distribution.
+Distributions are normalized to unit area. We see that the information coming
+from the first few observations does not limit $\lambdans$ much more than our
+prior from Eq.~\ref{eq:lprior} does. However, the median of the distribution 
+tracks the true value of $\lambdans$ fast, reaching close within as few as $4$
+observations.
+% 
+In the right panel, we show our measurement uncertainty $\delta\Lambda/\Lambda$
+as a function of $N$, where we take $\delta\Lambda$ to be the width of the 
+$90\%$ confidence interval on $\lambdans$ after $N$ observations.
+This is given by the green filled region (left y-axis).
+The green filled circles show the median measured value of $\lambdans$
+as a function of $N$. The dashed line at y value $=1$ marks zero error, i.e.
+$\delta\Lambda/\Lambda=0$. The dotted and dashed lines around this denote
+$\pm 25\%$ and $\pm 50\%$ relative errors, respectively. Independently, the
+light-grey filled shown the absolute measurement uncertainty $\delta\Lambda$
+on the right y-axis , as a function of $N$. A horizontal grey dashed line
+marks the true value of $\lambdans$. Contours of $1/\sqrt{N}$ are shown as 
+dotted light grey curves to guide the eye while studying the dependence of
+$\delta\Lambda$ on $N$.
+%
 

### 2016-04-06 emcee_PE/scripts/MultipleObservations.py
more changes

@@ -261,7 +261,11 @@ What it does:
         self.write_data = write_data
         #
         # TAG for storing data to disk
-        self.TAG = str(int(np.random.random()*1e5))
+        tmp_dir = 'L%d_N%d_' % (NSLambda, N)
+        self.RND = int(np.random.random()*1e5)
+        self.TAG = tmp_dir + str(self.RND) + '/'
+        try: os.makedirs(self.TAG)
+        except: print "Warning: temporary dir %s already exists!!" % self.TAG
         #
         self.print_info()
         return
@@ -341,7 +345,7 @@ What it does:
         self.chain_params   = chain_params
         #
         if self.write_data:
-          np.savetxt(self.TAG + "_chain_params.dat", chain_params, delimiter='\t')
+          np.savetxt(self.TAG + "chain_params.dat", chain_params, delimiter='\t')
         #
         self.FULL_chain_set = self.chain_set
         self.generate_cumulative_normalizations() # Normalizations
@@ -349,7 +353,7 @@ What it does:
         return chain_set
     #####
     ###
-    def load_events(self, chain_set,\
+    def load_events(self, chain_set_number,\
                       lambda_posterior_chains=None,\
                       NSLambda=None):
         '''
@@ -364,15 +368,23 @@ What it does:
         if lambda_posterior_chains == None:
             lambda_posterior_chains = self.lambda_posterior_chains
         if NSLambda == None: NSLambda = self.NSLambda
+        
+        #### 
+        ## Read in the set of events
+        param_file = str(chain_set_number) + "_chain_params.dat"
+        if not os.path.exists(param_file):
+            raise IOError("Could not load data for chain set #%d" % chain_set_number)
+        chain_params = np.loadtxt(param_file)
+        
         # Initiate primary data structure
-        chain_params = []
+        chain_set = []
         ###
         ## Loop over as many times as many events are to be generated
         ###
         if self.verbose:
             print >>sys.stdout, "Loading Events >>"
             sys.stdout.flush()
-        
+        #
         N = 0
         for rnd_q, rnd_chiBH, rnd_SNR, NSLambda in chain_params:
             N += 1
@@ -409,6 +421,9 @@ What it does:
             ##
             chain_set.append( [lambda_chain, lambda_kde.evaluate, lambda_kde_norm] )
         #
+        ###
+        ###
+        
         self.N              = N
         self.NSLambda       = NSLambda
         self.chain_set      = chain_set
[…]

### 2016-04-06 Paper/paper.tex
more changes
first pass at abstract

@@
 \begin{abstract}
-We attempt to answer the following questions:
+Very recently, the Advanced LIGO (aLIGO) observatories announced the first
+terrestrial detection of gravitational waves. This heralds an era of 
+observational gravitational wave (GW) astronomy. Neutron star - black hole
+(NSBH) binaries are one of the primary sources targeted for observation
+with the aLIGO detectors.
+% 
+GWs from these sources will carry signatures of the tidal distortion of
+the NS by the BH during the inspiral, and of its tidal disruption (if that
+happens) at merger. In the NS is disrupted, the merger and ringdown signal
+is drastically suppressed, which distinguishes it from a BHBH merger.
+% 
+In this paper, we present a Bayesian study of the measurability of the 
+tidal deformability of neutron stars, allowing the companion hole to take 
+on a variety of masses and spins.
+% 
+First, we find that if BHBH templates are used to estimate source 
+parameters for an NSBH signal, a significant bias is introduced in the 
+inference if the signal-to-noise ratio (SNR) is greater than $\rho=30$.
+% 
+Second, we find that if we are to get a loud signal with $\rho\geq 23$, 
+we can put a factor of two bound on the NS tidal deformability parameter
+$\lambdans\propto (R/M)^{5}$.
+% 
+Finally, we study the improvement in our measurement of $\lambdans$ from 
+multiple observations. Using an approximately astrophysical population, 
+we find that (i) the median measured $\lambdans$ catches up to within
+$10\%$ of the true value in approximately $20$ observations; (ii) the $90\%$
+confidence interval for $\lambda$ shrinks to $\pm 50\%$ within the same
+number of observatoins.
+% 
+We find these results encouraging, and recommend that an effort to 
+measure $\lambdans$ be taken within the LIGO-Virgo Collaboration once
+we begin to detect NSBH coalescences.
 \end{abstract}
@@
 \centering    
+\includegraphics[width=1.025\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
 \includegraphics[width=1.025\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
-\includegraphics[width=1.025\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
 \caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 

### 2016-04-06 Paper/paper.tex
author name just 2
updating intro

@@
     Fullerton, California 92834, USA}} %
+\newcommand{\AEI}{\affiliation{Albert Einstein Institute,
+Am M\"uhlenberg, G\"olm, Germany}} %
 
@@
 
+\author{Prayush Kumar}\CITA\email{prkumar@cita.utoronto.ca}
+\author{Michael P\"urrer}\AEI
 \author{People across the Atlantic}
@@
 % 
-In this paper, we present a Bayesian study of the measurability of the 
-tidal deformability of neutron stars, allowing the companion hole to take 
+In this paper, we present the first Bayesian study of the measurability of
+the tidal deformability of neutron stars, allowing the companion hole to take 
 on a variety of masses and spins.
@@
 parameters for an NSBH signal, a significant bias is introduced in the 
-inference if the signal-to-noise ratio (SNR) is greater than $\rho=30$.
+inference if the signal-to-noise ratio (SNR) is greater than $\rho\simeq30$.
 % 
@@
 $10\%$ of the true value in approximately $20$ observations; (ii) the $90\%$
-confidence interval for $\lambda$ shrinks to $\pm 50\%$ within the same
-number of observatoins.
+confidence interval for $\lambdans$ shrinks to $\pm 50\%$ around the true 
+value with the same number of observations.
 % 
@@
 strain sensitivity than their first-generation counterparts~\cite{Shoemaker2009}.
-Coalescing binaries of
-compact objects, i.e. Black Holes (BH) and/or Neutron Stars (NS), are the most
-likely targets for GW detection with these ground-based instruments. Over a
-$3$-month O1, these detectors would have access to as
-many binary mergers as initial-LIGO detectors did over approximately $6$ years
-of coincident operation. Both of these instruments will be further upgraded in
-phases to reach their design sensitivity by $2018-19$, at which point we expect to
-observe $\sim 70$ binary mergers a year~\cite{Abadie:2010cf}.
+% 
+During O1, the first terrestrial detection of gravitational waves was made, as
+was announced earlier in 2016~\cite{LIGOVirgo2016a}. Emitted more than a billion
+year ago, these waves herald the era of observational gravitational wave (GW)
+astronomy as they pass through Earth.
+% 
+Over a $3$-month O1, aLIGO had access to more binary mergers than initial-LIGO
+over its entire coincident operation period. Both of aLIGO instruments will be
+further upgraded in phases to reach their design sensitivity by $2018-19$, at 
+which point we may expect to observe $\sim 70$ binary mergers a 
+year~\cite{Abadie:2010cf}.
 
-NS-BH binaries are one of the most important sources of gravitational waves (GW)
-with these ground-based instruments, and we expect to detect $\mathcal{O}(10)$
-of such systems every year~\cite{Abadie:2010cf}. 
+Coalescing binaries of compact objects, i.e. black holes (BH) and/or neutron 
+stars (NS), remain the most interestings targets for GW based observation with
+aLIGO. NSBH binaries form an important class of compact binaries, and we expect
+to see $\mathcal{O}(10)$ of these every year with design-sensitivity 
+aLIGO~\cite{Abadie:2010cf}.
 % 
-One question that aLIGO will likely probe is `what is the nature of
-matter at nuclear densities', i.e. what is the equation of state (EoS) of 
-neutron stars. The EoS determines the compactness of NSs, which determines
-their deformation over the course of inspiral (albeit weakly, at $5$PN order in 
-phasing~\prayush{CITE PN TIDAL RESULTS}), as well as how early in the coalescence
-process will the NS be disrupted by the tidal field of the BH~\prayush{[CITE]}.
-Once disrupted,
-the NS forms an accretion disk around the BH. Most of this matter disk spirals 
-into the BH within a few milliseconds, with a small fraction becoming unbounded
-and gets ejected~\prayush{CITE BHNS NR}. It is believed that an accompanying
-short (lasting less than $2$ seconds) Gamma-ray Burst (sGRB) might be present
-under certain physical conditions~\prayush{CITE NSBH-GRB},
-making these systems one of the prime targets for joint electromagnetic-GW
-astronomy searches~\prayush{CITE EM-GW}.
-In such a situtation, the emitted GW signal
-is heavily damped post-disruption, and the quasi-normal modes of the resulting
-BH are not excited. Such a GW signal is qualitatively very different from a
-BH-BH merger, and one of the main goals of this work is to determine the 
-signal strength at which this difference will be quantifiabe by aLIGO.
+NSBH mergers are of interest for various reasons. For instance, they have 
+been associated with (as the progenitors of) short Gamma-ray Bursts
+(GRBs)~\cite{eichler:89,1992ApJ...395L..83N,moch:93,Barthelmy:2005bx,
+2005Natur.437..845F,2005Natur.437..851G,Shibata:2005mz,Paschalidis2014,
+Tanvir:2013}. Plausible mechanisms have been proposed for such a 
+scenario~\cite{Paschalidis2014,Ciolfi:2014yla}. A joint detection of GWs
+from an NSBH system with a GRB would validate this hypothesis.
 % 
+Another question that NSBH mergers can be used to probe is `what is the 
+nature of matter at nuclear densities', i.e. what is the equation of state
+(EoS) of neutron stars. 
+% 
+The gravitational-waves emitted by NSBH binaries carry subtle hints of the 
+matter structure of the neutron star. During early inspiral, the (weak) tidal
+field of the companion hole deforms the star, thereby exciting its resonant
+oscillation modes. This drains energy from the orbit resulting in an
+increase of the inspiral rate. We know the leading and next-to-leading
+order terms in post-Newtonian (PN) theory that capture this 
+effect~\cite{Vines2011}, entering binary phasing at $5$PN. Closer to merger, 
+the neutron star could get disrupted by the tidal field of its companion.
+Once disrupted, the neutron star material may form an accretion disk around 
+the black hole. Most of this matter disk spirals in within a few 
+milliseconds, with a small fraction becoming unbounded and gets 
+ejected~\cite{Foucart2015,Lovelace:2013vma,Deaton2013,Foucart2012,
+Paschalidis2014} (under favorable conditions, and possible accompanied by 
+gamma-ray emission).
+The emitted GW signal is, as a result, heavily damped post-disruption, 
+and the quasi-normal modes of the resulting black hole are not excited. Such a
+GW signal will be qualitatively very different from that from an equivalent
+binary black hole (BBH) merger. The primary goal of this work is to understand
+what it requires to optimally extract information about neutron star structure
+from such NSBH signals.
+
 However, studies have shown that for mass-ratio $\mbh/\mns\gtrsim 5$ the BH
@@
 innermost stable circular orbit (ISCO) is reached~\cite{Foucart:2013psa}.
+% 
+It is worth noting
+that not all point-particle terms in PN theory are known up to such high order,
+and therefore we need to be careful when using PN waveforms to measure tidal
+effects.
+%
 

### 2016-04-06 Paper/paper.tex
added power law figures

@@
 \centering    
-\includegraphics[width=1.\columnwidth]{plots/LambdalErrorBars_vs_Lambda_N49_Log.pdf}
+\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_Lambda.pdf}
+\caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
+as a function of $\lambdans$ itself. The drawn curves are cumulative, in the sense 
+that the $N=n$ curve includes information from all of the first $n$ observed events.
+Those corresponding to the first $N\leq 5$ events are shown in fading shades of red,
+with the ones for higher value of $N$ being shown in grey (going from light to dark in
+proportion to $N$).
+Green dash-dotted lines are contours of $y=x^{-(1+1/5)}$.
+% 
+\prayush{We note that the measurement uncertainty from a population follows the
+equation for the green lines.}
+% 
+}
+\label{fig:TT_PowerLawLambdaErrorVsN}
+\end{figure}
+%
+
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=\columnwidth]{plots/LambdalErrorBars_vs_Lambda_N79_Log.pdf}
+\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N.pdf}
 \caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
@@
 \label{fig:TT_Lambda_vs_Lambda_L500_2000_CI90_0_AllInOne}
-\end{figure}
+\end{figure*}
 %

### 2016-04-07 Paper/paper.tex
para to transitioning from nsbh general to paper

@@
 The Advanced LIGO (aLIGO) observatories began their first observing run ``O1''
-mid-September 2015, operating at a factor of $3-4$ higher gravitational-wave (GW) 
-strain sensitivity than their first-generation counterparts~\cite{Shoemaker2009}.
+mid-September 2015, operating at a factor of $3-4$ higher gravitational-wave
+(GW) strain sensitivity than their first-generation 
+counterparts~\cite{Shoemaker2009}.
 % 
 During O1, the first terrestrial detection of gravitational waves was made, as
-was announced earlier in 2016~\cite{LIGOVirgo2016a}. Emitted more than a billion
-year ago, these waves herald the era of observational gravitational wave (GW)
-astronomy as they pass through Earth.
-% 
-Over a $3$-month O1, aLIGO had access to more binary mergers than initial-LIGO
-over its entire coincident operation period. Both of aLIGO instruments will be
-further upgraded in phases to reach their design sensitivity by $2018-19$, at 
-which point we may expect to observe $\sim 70$ binary mergers a 
-year~\cite{Abadie:2010cf}.
+was announced earlier in 2016~\cite{LIGOVirgo2016a}. Emitted more than a
+billion year ago, these waves herald the era of observational gravitational
+wave (GW) astronomy as they pass through Earth.
+% 
+Over the $3$-months of O1, aLIGO had access to more binary mergers than 
+initial-LIGO over its entire coincident operation period. Both of aLIGO 
+instruments will be further upgraded in phases to reach their design 
+sensitivity by $2018-19$, at which point we may expect to observe $\sim 70$ 
+binary mergers a year~\cite{Abadie:2010cf}.
 
@@
 stars (NS), remain the most interestings targets for GW based observation with
-aLIGO. NSBH binaries form an important class of compact binaries, and we expect
-to see $\mathcal{O}(10)$ of these every year with design-sensitivity 
-aLIGO~\cite{Abadie:2010cf}.
+aLIGO. NSBH binaries form an important sub-class of compact binaries, and we
+expect to observe $\mathcal{O}(10)$ of these every year with design 
+sensitivity aLIGO~\cite{Abadie:2010cf}.
 % 
@@
 2005Natur.437..845F,2005Natur.437..851G,Shibata:2005mz,Paschalidis2014,
-Tanvir:2013}. Plausible mechanisms have been proposed for such a 
-scenario~\cite{Paschalidis2014,Ciolfi:2014yla}. A joint detection of GWs
-from an NSBH system with a GRB would validate this hypothesis.
-% 
-Another question that NSBH mergers can be used to probe is `what is the 
+Tanvir:2013}. Under favorable conditions, close to merger the neutron star
+can get disrupted by the tidal field of its companion hole. Once disrupted,
+the neutron star material may form an accretion disk around the black hole. 
+Most of the disk matter spirals in within a few milliseconds, with a small 
+fraction becoming unbounded and getting ejected, with a possibled simultaneous
+burst of gamma-ray emission)~\cite{Foucart2015,Lovelace:2013vma,Deaton2013,
+Foucart2012,Shibata:2005mz,Paschalidis2014}. A coincident detection of GWs
+from an NSBH system with a GRB would therefore provide us insights into the
+central engine powering GRBs. 
+% 
+Another question that NSBH mergers can help answer is `what is the 
 nature of matter at nuclear densities', i.e. what is the equation of state
-(EoS) of neutron stars. 
+(EoS) of neutron stars. This is what we will focus on in this paper.
 % 
@@
 order terms in post-Newtonian (PN) theory that capture this 
-effect~\cite{Vines2011}, entering binary phasing at $5$PN. Closer to merger, 
-the neutron star could get disrupted by the tidal field of its companion.
-Once disrupted, the neutron star material may form an accretion disk around 
-the black hole. Most of this matter disk spirals in within a few 
-milliseconds, with a small fraction becoming unbounded and gets 
-ejected~\cite{Foucart2015,Lovelace:2013vma,Deaton2013,Foucart2012,
-Paschalidis2014} (under favorable conditions, and possible accompanied by 
-gamma-ray emission).
-The emitted GW signal is, as a result, heavily damped post-disruption, 
-and the quasi-normal modes of the resulting black hole are not excited. Such a
-GW signal will be qualitatively very different from that from an equivalent
-binary black hole (BBH) merger. The primary goal of this work is to understand
-what it requires to optimally extract information about neutron star structure
-from such NSBH signals.
-
-However, studies have shown that for mass-ratio $\mbh/\mns\gtrsim 5$ the BH
-will need to have fairly large and positively-aligned (with the orbit) spins
-in order for its tidal field to be strong enough to disrupt the NS before the
-innermost stable circular orbit (ISCO) is reached~\cite{Foucart:2013psa}.
-% 
-It is worth noting
-that not all point-particle terms in PN theory are known up to such high order,
-and therefore we need to be careful when using PN waveforms to measure tidal
-effects.
-%
-
-Conventional astronomical methods have, to date, made observations of ~2500
-NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
-Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
-stellarcollapsemass}, although
-tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
-%
-Observations of over 2500 NSs have shown that the spin of NSs
-$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
-%
-On the other hand, indirect observations of stellar-mass BHs place their
-masses between $5-35M_\odot$, with their spin angular momenta 
-$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
-(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
-examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
-Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
-and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
-
-The difficulty in discerning the matter effects in an NSBH merger GW signal
-lies in distinguishing it reliably from a BHBH merger signal. The earlier the
-NS disrupts, the more orbits there are over which the effect of NS disruption
-is visible. Where exactly does disruption happen is governed both by the 
-mass-ratio of the binary, as well as the spin on the BH. Larger BHs (with
-$q=m_\mathrm{BH}/m_\mathrm{NS}\gtrsim 7$), as well as BHs with large retrograde
-spins ($\chi_\mathrm{BH} \lesssim 0$), tend to swallow the NS whole due to a
-rapid merger (i.e. without it disrupting before the inner-most stable circular
-orbit (ISCO) is reached). On the other hand, binaries with mass-ratios 
-$q=2-5$ with large pro-grade $\chi_\mathrm{BH}$ can disrupt much before 
-merger, leaving their imprint in the GW signal emitted close to $1.7-2$~kHz.
-%
+effect~\cite{Vines2011}, entering binary phasing at $5$PN order. Close to
+merger, as the NS gets disrupted and forms long tails that feed into the 
+companion black hole, the quadrupole moment of the system falls monotonically.
+As a result, the emitted GWs are subdued significantly post-disruption
+(as compared to what they would be for a BBH),
+and the quasi-normal modes of the resulting black hole are not 
+excited~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart2015,
+Pannarale:2015jia}.
+% 
+% 
+The difficulty in discerning NS matter effects in an NSBH coalescence from its
+GW signal lies in distinguishing it reliably from a BBH signal, since the 
+difference between the two is either subtle (during inspiral) or short-lived
+(disruption and merger). The earlier in inspiral the NS disrupts, the more
+orbits there are over which its effect is visible. The mechanism of disruption
+is governed by the nature of the BH, i.e. its mass and spin. More massive BHs 
+(with $m_\mathrm{BH}\gtrsim 7\, m_\mathrm{NS}$), as well as BHs with
+(large) retrograde spins $(\chi_\mathrm{BH}\in[-1. 0])$, tend to swallow the 
+NS whole due to a rapid merger (i.e. without it disrupting before the
+inner-most stable circular orbit (ISCO) is reached). The quasi-normal
+oscillation modes of the final black hole's horizon are also 
+excited~\cite{Foucart:2013psa}. Therefore, the only available information 
+comes from the inspiral rate modification. On the other hand, less massive
+black holes spinning in the prograde sense (with mass-ratio 
+$q\in[2, 5]$, and $\chi_\mathrm{BH}\geq 0$) can cause their companion star to 
+disrupt much before merger, leaving a strong imprint on the emitted GW 
+signal~\cite{Shibata:2007zm,2010PhRvD..81f4026F,Lovelace:2013vma,
+Kawaguchi:2015}.
+
+A large fraction of past work aimed at measuring NS matter effects from GW
+signals has consisted of inquiries about binary neutron stars (BNS)~\cite{
+Lee1999a,Lee1999b,Lee2000,oechslin:07,Read:2008iy,Markakis:2010mp,Markakis:2011vd,
+stergioulas:11,East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}. While single 
+observations are not as informative, Bayesian studies have shown that with
+$20-30$ BNS observations, aLIGO would begin to distinguish between hard,
+moderate and soft candidate equations of state~\cite{DelPozzo:13,
+Chatziioannou:2015uea,Agathos:2015a}. These studies rely on the PN 
+description of BNS dynamics, which is reasonable since BNSs merge at very high
+frequencies and most of their signal power resides in inspiral.
+% 
+On the other hand, NSBHs merge at relatively lower frequencies than BNS, 
+and their disruption during late-inspiral occurs at frequencies where aLIGO
+has relatively more resolving power. This, combined with a shorter inspiral,
+means that the merger portion of NSBH signals is relatively more valuable for
+measuring $\lambdans$. Some past NSBH studies aimed at measuring tidal effects,
+however, have used PN inspiral-only waveforms~\cite{Maselli:2013rza}. In doing
+so, however, (i) they ignore the merger signal which could contain significant
+information for NSBHs, and (ii) the errors due to unknown vaccum terms in PN 
+waveforms could easily dominate over the tidal terms 
+themselves~\cite{Yagi:2014}, since there are unknown point-particle terms at 
+orders {\it lower} than that at which tidal effects enter. Other studies that
+use complete numerical simulations or derived models either use Fisher matrix 
+to estimate $\lambdans$ measurement error~\cite{Lackey2011,Lackey:2013axa},
+which is not reliable at realistic signal-to-noise ratios 
+(SNR)~\cite{Vallisneri:2007ev}, or are limited in the binary parameter space
+they sample~\cite{Foucart:2013psa}.
 
@@
 simulate NSBH merger waveforms.
+
+As this line of inquiry relies mainly on 
+classifying a GW signal as an NSBH binary, it can be pursued using non-tidal
+waveform models~\cite{}.
+
+% Conventional astronomical methods have, to date, made observations of ~2500
+% NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
+% Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
+% stellarcollapsemass}, although
+% tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
+% %
+% Observations of over 2500 NSs have shown that the spin of NSs
+% $|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
+% %
+% On the other hand, indirect observations of stellar-mass BHs place their
+% masses between $5-35M_\odot$, with their spin angular momenta 
+% $|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
+% (see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
+% examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
+% Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
+% and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
+%
 
@@
 \begin{equation}
- \tilde{h}_\mathrm{NSBH}(f, \vec{\theta}) = \tilde{h}_\mathrm{BBH}(f, \vec{\theta} - \{\lambdans\})\,A(f, \vec{\theta})\,e^{\ii \Delta\Phi(f, \vec{\theta})},
+ \tilde{h}_\mathrm{NSBH}(f, \vec{\theta}, \lambdans) = \tilde{h}_\mathrm{BBH}(f, \vec{\theta})\,A(f, \vec{\theta}, \lambdans)\,e^{\ii \Delta\Phi(f, \vec{\theta}, \lambdans)},
 \end{equation}
-where $\vec{\theta}:=\{\mbh, \mns, \chibh,\chins=0, \lambdans\}$ are the intrinsic binary
+where $\vec{\theta}:=\{\mbh, \mns, \chibh,\chins=0\}$ are the intrinsic binary
 parameters. The first term on the right hand side is an underlying
@@
 term does the same as the second, but to the waveform phasing. During inspiral,
-$\Delta\Phi(f, \vec{\theta})$ is set to the PN tidal phasing corrections,
+$\Delta\Phi(f, \vec{\theta}, \lambdans)$ is set to the PN tidal phasing corrections,
 at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
@@
 
-The joint inferred probability distribution of all parameters $\vec{\theta}\cup\{\lambdans\}$ 
+The joint inferred probability distribution of all parameters 
+$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ 
 is given by
 \begin{equation}
- p(\vec{\theta}, \lambdans | d_n, K) = \dfrac{p(d_n|\vec{\theta}, \lambdans, K)\,p(\vec{\theta}, \lambdans | K)}{p(d_n|K)}.
+ p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
 \end{equation}
-Here, $p(\vec{\theta}, \lambdans | K)$ is the prior probability for binary parameters of taking 
+Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of taking 
 particular values. In this paper, we use a uniform prior on the individual masses within 
@@
 \begin{equation}\label{eq:likelihood}
- \LL(\vec{\theta}, \lambdans) \equiv p(\vec{\theta}, \lambdans | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+ \LL(\vec{\Theta}) \equiv p(\vec{\Theta} | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
 \end{equation}
-where $h\equiv h(\vec{\theta}, \lambdans)$ is our waveform model, and the detector noise
-weighted inner-product $\langle a|b\rangle$ is defined as 
+where $h\equiv h(\vec{\Theta})$ is given by our waveform model LEA+, and the
+detector noise weighted inner-product $\langle a|b\rangle$ is defined as 
 $$

### 2016-04-07 Paper/paper.tex
minor changes

@@
 field of the companion hole deforms the star, thereby exciting its resonant
-oscillation modes. This drains energy from the orbit resulting in an
-increase of the inspiral rate. We know the leading and next-to-leading
-order terms in post-Newtonian (PN) theory that capture this 
+oscillation modes. The induced quadrupolar moment of the star depends on the
+tidal field $\mathcal{E}_{ij}$ due to its companion through the relation
+$Q_{ij} = -\lambda\mathcal{E}_{ij}$, where $\lambda$ is the EoS dependent
+tidal deformability parameter related to the neutron star's dimensionless
+Love number $k_2$ and radius $R_\mathrm{NS}$ as
+$\lambda:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$. This drains energy from the orbit
+resulting in an increase of the inspiral rate. We know the leading and next-to
+-leading order terms in post-Newtonian (PN) theory that capture this 
 effect~\cite{Vines2011}, entering binary phasing at $5$PN order. Close to
@@
 
-As this line of inquiry relies mainly on 
-classifying a GW signal as an NSBH binary, it can be pursued using non-tidal
-waveform models~\cite{}.
-
-% Conventional astronomical methods have, to date, made observations of ~2500
-% NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
-% Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
-% stellarcollapsemass}, although
-% tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
-% %
-% Observations of over 2500 NSs have shown that the spin of NSs
-% $|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
-% %
-% On the other hand, indirect observations of stellar-mass BHs place their
-% masses between $5-35M_\odot$, with their spin angular momenta 
-% $|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
-% (see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
-% examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
-% Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
-% and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
-%
-
 Finally, we note that our results are applicable for the design sensitivity
@@
 that more a recent work that improves upon the waveform model we use in this
-study has been published~\cite{Pannarale:2015jka}, but it gives only an 
-amplitude model and needs to be (in future) augmented with a compatible phase
-model. In addition, accuracy of the underlying numerical simulations used to
-calibrate the waveform model used here has not been verified against 
+study has been published~\cite{Pannarale:2015jka}, but it provides only an 
+amplitude model which needs to be (in future) augmented with a compatible
+phase model. In addition, accuracy of the underlying numerical simulations 
+used to calibrate the waveform model used here have not been verified against 
 independent codes so far.
-It is therefore difficult to assess the combined modeling error and the 
-results presented in our study are limited by the limitations of the model.
-However, we expect the combined effect of modeling errors to not change our
-qualitative conclusions.
+It is therefore difficult to assess the combined modeling error and its effect
+on our results. they are, therefore, limited by the limitations of our
+waveform model. However, we expect the combined effect of modeling errors to
+not change our qualitative conclusions.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Discussion
+
+Conventional astronomical methods have, to date, made observations of ~2500
+NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
+Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
+stellarcollapsemass}, although
+tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
+%
+Observations of over 2500 NSs have shown that the spin of NSs
+$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
+%
+On the other hand, indirect observations of stellar-mass BHs place their
+masses between $5-35M_\odot$, with their spin angular momenta 
+$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
+(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
+examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
+Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
+and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
 

### 2016-04-07 Paper/paper.tex
first results in intro

@@
 
-The goal of this paper is to probe the distinguishability of BHBH and NSBH
-mergers, focusing at the effects of the BH mass, BH spin, and the tidal 
-deformability parameter $\Lambda_\NS\sim(R/M)^5_\NS$. We also investigate the
-accuracy with which aLIGO can measure $\Lambda_\NS$ itself. For both of 
-these studies, we will use the zero-detuning high-power noise curve for the
-aLIGO detectors. \prayush{In addition, we also study the effect of tuning aLIGO
-over a narrow frequency range aimed at NSBH mergers.}
-We use the models published in Ref.~\cite{Lackey:2013axa,Pannarale:2015jka} to
-simulate NSBH merger waveforms.
+In this paper we study the measurability of neutron star's tidal deformability
+from GWs emitted by {\it disruptive} NSBH mergers using aLIGO. We also probe how tidal 
+effects affect the estimation of other binary parameters for the same class
+of systems. This study improves upon previous work in many ways. First, we
+include tidal effects during inspiral and merger in a consistent way, by using
+the waveform model of Ref.~\cite{Lackey:2013axa} (henceforth ``LEA+''). Second,
+we include the effect
+of black hole spin on tidal GW signals, in addition to the effect of BH mass 
+and NS's deformability itself. Third, we perform a fully Bayesian 
+analysis, instead of using a Fisher matrix approximation. Fourth, we explore 
+how our measurement errors decrease as we gain information from multiple 
+(realistic) events.
+
+We first probe the effect of not including tidal effects in parameter 
+estimation templates on the recovery of non-tidal binary parameters, such as 
+component masses and spins. This is the case of current and planned aLIGO
+efforts.
+To do so, we use the LEA+ model to generate a set of realistic signals, and
+use that as data from which we estimate binary mass and spin parameters with
+a Bayesian Monte-Carlo algorithm, using non-tidal (BBH) waveforms as filters.
+% 
+As in rest of the paper, we restrict here our parameter space to span 
+mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin $\chibh\in[-0.5, +0.75]$, 
+and dimensionless neutron star tidal deformability 
+$\lambdans:= G\left(\frac{c^2}{G \mns}\right)^5\lambda \in[500, 2000]$.
+The mass of NS for signals (and not templates) is fixed to $1.35M_\odot$ 
+throughout~\cite{stellarcollapsemass}, and is taken to be 
+non-spinning~\cite{Miller:2014aaa}.
+Each signal is rescaled to a range of SNR values $\rho\in[10, 70]$.
+% 
+We find that non-inclusion of tidal effects leads to systematic biases in
+the measurement of non-tidal parameters. Consider first the binary chirp 
+mass $\mchirp:=M\eta^{3/5}$ where $M$ is the binary total mass, and 
+$\eta:=\mbh\mns/M^2$ is the dimensionless mass-ratio. We find that for
+realistic (moderate) SNRs, i.e. $\rho\lesssim 30$, not including tidal
+effects introduces a barely measurable bias that remains below $50\%$
+of the statistical uncertainty of the $\mchirp$ measurement itself.
+For higher signal strengths, we find that the underlying statistical
+uncertainty decreases enough for the systematic bias to become comparable
+to it for $\rho=50-70$ and higher. This is expected because chirp mass is fairly
+accurately determined by the inspiral, where tidal effects are weak.
+Next, we consider the mass-ratio $\eta$. We find that only if the NS is quite
+deformable (with $\lambdans=2000$), {\it and} has a companion BH whose mass
+is $\mbh\lesssim 4.5M_\odot$ (i.e. in the astrophysical ``mass-gap''~\cite{
+Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}),
+will the systematic bias in $\eta$ measurement become comparable to its
+statistical uncertainties at SNRs $\rho\leq 30$. For SNRs $\rho\gtrsim30$,
+however, not including tidal effects is very likely to {\it significantly} bias
+the mass-ratio estimate for NSBH binaries. Finally, we consider the spin on the
+black hole $\chibh$. We find that $\chibh$ measurements are affected in a similar
+way to $\eta$. By not including tidal effects, we compromise our measurement of
+$\chibh$ significantly only if signal $\rho\gtrsim 30$. 
+%
+Overall, we learn an important thing. Bayesian parameter estimation algorithms
+aimed at identifying NSBH binaries with low latency in order to alert 
+electromagnetic (EM) observers in time to establish a possible coincident 
+observation of a GRB~\cite{2012A&A...541A.155A,Singer:2014qca,Singer:2015ema,
+Pankow:2015cra} can use BBH templates to look for NSBH systems with aLIGO.
+This is so because the primary requirement of the EM follow-up effort is rapid
+classification of a binary as an NSBH system, which can be achieved just as 
+easily with BBH templates, purely on the basis of the smaller component's mass.
+Another conclusion we draw is that detection searches are unaffected by the
+choice of ignoring tidal effects in matched-filtering templates.
 

### 2016-04-07 Paper/paper.tex
updated methods & other secs

@@
 
+Next, we study the ability of aLIGO to constrain neutron star tidal 
+deformability with a single observation of coalescing NSBHs. For this, we
+use the same setup and parameter choices for signal waveforms as before, but
+replace the template model with one that includes tidal effects from inspiral
+through to merger~\cite{Lackey:2013axa} (i.e. LEA+). First of all, this reduces
+the biases in the measurement of non-tidal parameters. Second, as we can see
+from Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, at signal-to-noise
+ratios $\simeq 20-30$, only under the most favorable circumstance (including a
+BH with mass within the ``mass-gap'') can we put a factor of $1.5-2$ bound on
+$\lambdans$ at $90\%$ confidence level. For similar signals with 
+$\rho\gtrsim 30$ we can constrain $\lambdans$ within $\pm 50\%$ of its true
+value. For more realistic SNRs and BH parameters, a single measurement is 
+for $\lambdans$ is accompanied with a $\sim\pm150+\%$. uncertainty. While these
+results are expected, multiple realistic observations is the next window that 
+we explore in order to obtain more information on $\lambdans$. These results 
+are detailed in Sec.~\ref{s1:PEwithNS}.
+
+Multiple realistic observations deliver far more information than what a single
+observation furnishes, as has been shown the case to be for binary neutron
+stars~\cite{DelPozzo:13}. As our final study, we perform a fully-Bayesian 
+analysis of how our estimation of $\lambdans$ changes as we accumulate more and
+more detections. The choices that determine the population are important here,
+and were made as follows. Neutron star mass is fixed for signals at
+$1.35M_\odot$, and its spin $\chins=0$. Black hole mass is sampled uniformly
+from the range $[2,5]\times 1.35=[2.7, 6.75]M_\odot$. Our choice here is given
+by the intersection set of the mass range that allows for neutron star disruption
+and the range supported by the waveform model LEA+~\cite{Foucart2012,
+Foucart:2013a,Lackey:2013axa}. Black hole spin is sampled uniformly from the
+{\it aligned} range, i.e. $\chibh\in[0,+1]$. The true deformability is fixed
+once per population, sampled uniformly from the range $\lambdans\in[0,2000]$.
+In order to keep the computational cost tractable, we make a further 
+approximation. We inject signals into zero noise at different SNRs and recover
+their parameters from Bayesian inferencing for a fixed set of binaries,
+whose parameters lie on a uniform grid given by:
+$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+Once the population parameters been selected, each event's parameters are
+replaced by their nearest neighbor in the above set. We make sure that this 
+approximation is conservative by keeping our population sampling ranges within the
+conservative side of the range spanned by the above set. With the caveats
+laid out, we report our results in Sec.~\ref{s1:multiple_observations}.
+% 
+We find \prayush{[...]}
+
 Finally, we note that our results are applicable for the design sensitivity
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this section we summarize various technical aspects of this paper. The model
+used to obtain tidal waveforms is described in Sec.~\ref{s2:waveforms}. The 
+process of inferring source parameters from a GW signal is summarized in 
+Sec.~\ref{s2:bayesian}. The process of generating and combining multiple events
+is instead detailed in the corresponding results section,
+Sec.~\ref{s2:astro_multiple}.
+
 \subsection{Waveform Models}\label{s2:waveforms}
@@
 
-The joint inferred probability distribution of all parameters 
-$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ 
-is given by
-\begin{equation}
+We perform Bayesian parameter estimation for tidal signals, using
+both tidal and non-tidal templates. The signal parameters are selected 
+to lie on a grid given by:
+$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+Corresponding signal waveforms are generated using the model described in the 
+previous sub-section, and each injected into a separate zero noise data segment
+$d_n$. Templates can be either tidal or non-tidal (LEA+ with $\lambdans=0$).
+
+Inferred from $d_n$, the joint probability distribution for binary parameters
+$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ is given by
+\begin{equation}\label{eq:postprob}
  p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
 \end{equation}
-Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of taking 
-particular values. In this paper, we use a uniform prior on the individual masses within 
-pre-chosen ranges, as well as a uniform prior on the spin of the black hole. We set neutron star's
-spin identically $=0$, and its tidal deformability parameter $\lambdans$ to be equally likely
-between $[0, 4000]$. The prior probability of obtaining a particular realization of data
-$p(d_n|K)$ is absorbed into the overall normalization. Finally, the first term in the numerator
-$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given stretch of 
-data $d_n$ if we assume that a signal parameterized by $\vec{\theta}\cup\{\lambdans\}$ is buried
-in it, i.e.
+Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of
+taking particular values. In this paper, we use a uniform prior on the
+individual masses within pre-chosen ranges, as well as a uniform prior on the
+spin of the black hole. We set neutron star's spin $=0$, and its tidal
+deformability parameter $\lambdans$ is sampled uniformly from $[0, 4000]$. In
+addition, we exclude mass-ratios between $[1,2]$ from our prior. This is done
+because of model constraints for LEA+. The prior probability of obtaining a 
+particular realization of data $p(d_n|K)$ is absorbed into the overall 
+normalization. Finally, the first term in the numerator
+$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given
+stretch of data $d_n$ if we assume that a signal parameterized by
+$\vec{\theta}\cup\{\lambdans\}$ is buried in it, i.e.
 \begin{equation}\label{eq:likelihood}
- \LL(\vec{\Theta}) \equiv p(\vec{\Theta} | K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+ \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
 \end{equation}
-where $h\equiv h(\vec{\Theta})$ is given by our waveform model LEA+, and the
-detector noise weighted inner-product $\langle a|b\rangle$ is defined as 
-$$
+where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
+and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
+as
+\begin{equation}
 \langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
-$$
-where $\tilde{a}$ means the Fourier transform of the finite time series $a(t)$, and 
-$S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The definition of $\LL$ in 
-Eq.~\ref{eq:likelihood} assumes that the instrument noise is colored Gaussian.
+\end{equation}
+% 
+where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
+and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The 
+definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
+noise is colored Gaussian.
+
+We sample the probability distribution given in Eq.~\ref{eq:postprob} using
+an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}
+package~. A total of $100$ independent walkers were used, each of which
+was allowed to collect $100,000$ samples. We measure the correlation time and keep
+only $20,000$ independent posterior points to further analyze the calculated 
+posterior probability distributions. One simplifying assumption made to mitigate 
+computational cost was setting the frequency sampling rate $\Delta f=0.4$~Hz, which
+we found to be sufficient to calculate likelihoods robustly~\cite{Purrer:2015nkh}.
+\prayush{TODO: MICHAEL}.
 
@@
 \centering    
-\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
+\includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
 %\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
@@
 \centering    
-\includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1000_0_CI90_0}
-\includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
+\includegraphics[width=0.9025\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1000_0_CI90_0}
+\includegraphics[width=0.9025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
 % \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
@@
 \centering    
-\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0.pdf}
-\includegraphics[width=1.025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0.pdf}
-\caption{In these figures we highlight the regions of BH mass and spin plane where we can 
-constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
-Panels correspond to different signal strength, with $\rho$ increasing from left to right.
-Within each panel, $\lambdans$ measurement uncertainty contours are shown as functions of BH 
-mass and spin, with line-types corresponding to different uncertainty levels, and
-colors showing the injected $\lambdans$.
+\includegraphics[width=0.9025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
+\includegraphics[width=0.9025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
+\caption{In these figures we highlight the regions of BH mass and spin plane 
+where we can constrain the $\lambdans$ of its companion NS within a factor of 
+$1-2$ of its true value. Panels correspond to different signal strength, with 
+$\rho$ increasing from left to right. Within each panel, $\lambdans$ 
+measurement uncertainty contours are shown as functions of BH mass and spin,
+with line-types corresponding to different uncertainty levels, and colors 
+showing the injected $\lambdans$. 
 % We note that this information is, in principle, contained in 
@@
 % the effect of NS's deformability itself on its measurement. 
-As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
-of NS matter
-effects improves with all factors that enhances the signature of the NS's disruption on
-the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
-set by the detectors).
+As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, the 
+measurability of NS tidal deformability improves with all factors that 
+enhances the signature of the NS's disruption on the \textit{detectable} 
+portion of the emitted GW signal (implying, within a frequency band set by the
+detectors).
 }
@@
 %
-% 
+
 In the previous section we showed that we begin to care for the effects of NS's
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
-% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % 
@@
 \centering    
-\includegraphics[width=1.025\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
-\includegraphics[width=1.025\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
+\includegraphics[width=1\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
+\includegraphics[width=1\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
 \caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 
@@
 %
-
 % 
@@
 \end{figure*}
-%
-% 
+
 \textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
-% 
 % 
@@
 signal from each binary be characterized by the parameters 
-$\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}$ and $\lambdans$. Note that we have folded
+$\,\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}$ and $\lambdans$. Note that we have folded
 in the luminosity distance, inclination angle of the binary's orbital angular momentum with

### 2016-04-07 emcee_PE/scripts/SingleObservation.py

@@ -174,7 +174,7 @@ col corresponds to the lower level group
       for lev in levellines:
         cset = ax.contour(xx, yy, zz, levels=[lev], colors='k', ls='--', linewidths=4, hold="on")
         for c in cset.collections: c.set_linestyle('dotted')
-        plt.clabel(cset, colors='b', inline=1, fmt=lfmt, fontsize=16)
+        plt.clabel(cset, colors='r', inline=1, fmt=lfmt, fontsize=16)
         contours_tmp[lev] = cset.collections[0].get_paths()
       if vverbose:
         print "for %s" % titles[idx][jdx], contours_tmp

### 2016-04-07 Paper/paper.tex
minor change to bayesian sec

@@
 definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
-noise is colored Gaussian.
-
-We sample the probability distribution given in Eq.~\ref{eq:postprob} using
+noise is colored Gaussian. The marginalized probability distribution for any
+single binary parameter (say $\alpha$) can be obtained by integrating
+Eq.~\ref{eq:postprob} over all other parameters, i.e. 
+\begin{equation}
+ p(\alpha | d_n, K) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, K),
+\end{equation}
+where $\vec{\Theta}_\alpha$ is the set of remaining parameters, i.e.
+$\vec{\Theta}_\alpha:=\vec{\Theta} - \{\alpha\}$.
+
+We sample the probability distribution of Eq.~\ref{eq:postprob} using
 an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}

### 2016-04-09 Paper/paper.tex
get post processing to work

@@
 % 
+\begin{figure}
+\centering    
+\includegraphics[width=1.05\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}
+% \includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
+\caption{%
+% {\it Left}:
+Posterior probability distributions for $\lambdans$ (colored curves), and 
+associated $90\%$ confidence intervals (grey vertical lines), shown for 
+different number of accumulated observations $N$.
+% {\it Right}: Filled-region plots showing the median and $90\%$ confidence intervals
+% for $\lambdans$, as a function of the number of observed events $N$.
+%  Shown are two independent filled regions. 
+% %  
+%  One in green shows the recovered confidence interval
+%  for $\lambdans$ (on the left y-axis), normalized by its true value,
+%  as a function of the 
+%  number of observed events $N$. The recovered median is shown by the
+%  green line-circled curve. The pair of horizontal green dashed (dotted)
+%  lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
+%  the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
+% % 
+%  The second filled region, in grey, shows the recovered
+%  confidence interval for neutron star compactness, without normalizing.
+%  The corresponding y-values are shown on the *right* y-axis. The horizontal grey
+%  dashed line traces the true value of $\lambdans$, and is what we expect
+%  the filled region to approach with increasing number of observations.
+% %  
+%  Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
+}
+\label{fig:TT_Lambda_vs_N_L800_CI90_0}
+\end{figure}
+%
+% 
 \begin{figure*}
 \centering    
-\includegraphics[width=1.\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}
-\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
-\caption{{\it Left}:  Posterior probability distributions for $\lambdans$
-and associated $90\%$ confidence intervals, shown as a function of number of events 
-accumulated $N$.
-{\it Right}: Filled-region plots showing the median and $90\%$ confidence intervals
-for $\lambdans$, as a function of the number of observed events $N$.
- Shown are two independent filled regions. 
-%  
- One in green shows the recovered confidence interval
- for $\lambdans$ (on the left y-axis), normalized by its true value,
- as a function of the 
- number of observed events $N$. The recovered median is shown by the
- green line-circled curve. The pair of horizontal green dashed (dotted)
- lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
- the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
-% 
- The second filled region, in grey, shows the recovered
- confidence interval for neutron star compactness, without normalizing.
- The corresponding y-values are shown on the *right* y-axis. The horizontal grey
- dashed line traces the true value of $\lambdans$, and is what we expect
- the filled region to approach with increasing number of observations.
-%  
- Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
-}
-\label{fig:TT_Lambda_vs_N_L800_CI90_0}
-\end{figure*}
-%
-% 
-\begin{figure*}
-\centering    
-\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L2000.pdf}
-\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1500.pdf}\\
-\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1000.pdf}
-\includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L500.pdf}
+\includegraphics[width=.9\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L2000.pdf}
+\includegraphics[width=.9\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1500.pdf}\\
+\includegraphics[width=.67\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1000.pdf}
+\includegraphics[width=.67\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
+\includegraphics[width=.67\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L500.pdf}
 \caption{Filled-region plots showing the median and $90\%$ confidence intervals
@@
 % 
-Imagine we have $N$ stretches of data,
-$d_1, d_2, \cdots, d_N$, each containing a single observation of an NSBH binary. Let $K$
-denote our collective prior knowledge, except for expectations on binary parameters
-themselves, which will enter the following calculation explicitly. Let the gravitational wave
-signal from each binary be characterized by the parameters 
-$\,\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}$ and $\lambdans$. Note that we have folded
-in the luminosity distance, inclination angle of the binary's orbital angular momentum with
-respect to the plane of the detector, and its sky location angles into a single parameter $\rho$,
-which is the SNR of the received signal. Note also that $\vec{\theta}$ does not include
-$\lambdans$ as the latter would be treated independently. In addition we assume that all
-NS's in a population will have the same deformability parameter $\lambdans$, and therefore
-its cumulative measurement is possible.
-% 
-Using Bayes' theorem, the measured probability distribution function for $\lambdans$
-from $N$ observations is
+Imagine that we have $N$ stretches of data, $d_1, d_2, \cdots, d_N$, each 
+containing a single signal emitted by an NSBH binary. Each of these signals can
+be characterized by the following binary parameters:
+$\,\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}\cup\{\lambdans\}$. Note
+that we have folded in the luminosity distance, inclination angle of the 
+binary's orbital angular momentum with respect to the plane of the detector,
+and source's sky location angles into a single parameter $\rho$, which is the 
+SNR of the received signal. This is possible because of our simplification of 
+using only the dominant $l=|m|=2$ modes of the GW strain~\footnote{The 
+$l=|m|=2$ multipoles of the GW strain, when decomposed in a spin $-2$ weighted
+spherical harmonic basis, contain more than $99\%$ of the total signal power}
+that makes various angle parameters completely degenerate with distance. 
+Further, let $K$ denote all of our collective prior knowledge, except for 
+expectations on binary parameters themselves, which enter the following 
+calculation explicitly. For instance, $K$ includes our assumption that all
+NS's in a single population have the same deformability parameter $\lambdans$, 
+and its cumulative measurement is therefore possible.
+% 
+Using Bayes' theorem, the measured probability distribution function for 
+$\lambdans$ from $N$ observations is given by~\footnote{%
+Eq.~\ref{eq:lambdaMultiple} includes the following implicit factor to normalize
+its right-hand side: $\dfrac{\prod_i p(d_i|K)}{\int p(\lambdans)\,p(d_1, d_2, \cdots, d_N | \lambdans; K)\,\D\lambdans}$}
 \begin{equation}\label{eq:lambdaMultiple}
@@
 \end{equation}
-where $p(\lambdans|K)$ is the prior expectation for $\lambdans$, and we have assumed
-that all of the $N$ observations are independent. A priori, we assume that no 
-particular value of $\lambdans$ is more preferred over another, and $\lambdans\leq 4000$,
-i.e.
+where $p(\lambdans|K)$ is the prior expectation for $\lambdans$, and all $N$ 
+observations are taken to be mutually independent. A priori, we assume
+that no particular value of $\lambdans$ is preferred over another, and 
+$\lambdans\leq 4000$, i.e.
 \begin{equation}\label{eq:lprior}
- p(\lambdans | K) = \dfrac{1}{4000}\,\mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right)
+ p(\lambdans | K) = \dfrac{1}{4000}\,\mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right).
 \end{equation}
 % 
-The second set of terms in Eq.~\ref{eq:lambdaMultiple} of the form $p(\lambdans | d_i, K)$
-is each the probability distribution for $\lambdans$ inferred {\it a posteriori} from
-the \textit{i}-th observation independently. 
+In the second set of terms in Eq.~\ref{eq:lambdaMultiple} (of the form 
+$p(\lambdans | d_i, K)$), each is the probability distribution for $\lambdans$
+inferred {\it a posteriori} from the \textit{i}-th observation in isolation. 
 They are calculated by integrating
-\begin{equation}
+\begin{equation}\label{eq:margpost}
  p(\lambdans | d_i, K) = \int \D \vec{\theta}\, p(\vec{\theta}, \lambdans | d_i, K),
@@
 distribution of all source parameters $\vec{\theta}\cup\{\lambdans\}$ for the 
-$i$-th event, as described in Sec.~\ref{s2:bayesian}.
+$i$-th event, as given by Eq.~\ref{eq:postprob}. By substituting
+Eq.~\ref{eq:lprior}-\ref{eq:margpost} into Eq.~\ref{eq:lambdaMultiple}, we
+calculate the probability distribution for $\lambdans$ as measured using $N$
+independent events.
 
@@
 observations each, and quantify what we learn from each successive observation 
-using Eq.~\ref{eq:lambdaMultiple}. This allows us to determine how
-rapidly does our median estimate for $\lambdans$ converges to the true value,
-and how broad our uncertainty intervals for the same are, as $N$ increases.
-In order to generate each population, we first fix
+using Eq.~\ref{eq:lambdaMultiple}. For each population, we determine how
+rapidly our median estimate for $\lambdans$ converges to the true value,
+and how rapidly our confidence intervals for the same shrink, with increasing
+$N$. In order to generate each population, we first fix
 the NS properties: (i) NS mass $\mns=1.35M_\odot$, (ii) NS spin $\chins=0$
-and (iii) NS tidal deformability $\lambdans=$ some value that is chosen from
-$\{500,800,1000,1500,2000\}$. Next, we successively generate $N$ 
-independent events, uniformly sampling the remaining
-source parameters from the following ranges: (i) mass-ratio $q\in[2,5]$, (ii)
-BH spin $\chibh\in[0, 0.75]$, (iii) effective distance from within a 
-spherical volume whose radius is given by the lowest detectable
-SNR (which we take as $\rho=10$). We remind ourselves that
-once source parameters are fixed, there is a one-to-one mapping between
+and (iii) NS tidal deformability $\lambdans=$ fixed value chosen from
+$\{500,800,1000,1500,2000\}$. Each event is generated independently, sampling
+the remaining source parameters uniformly from the following ranges: (i)
+mass-ratio $q\in[2,5]$, (ii) BH spin $\chibh\in[0, 0.75]$, (iii) effective 
+distance from within a spherical volume whose radius is given by the lowest 
+detectable SNR (which we take as $\rho=10$). We remind ourselves that
+once binary parameters are fixed, there is a one-to-one mapping between
 effective distance and SNR. Finally, in order to mitigate computational
 cost, we make an additional simplification.
-We perform full MCMC parameter estimation for a set of systems with parameters
-laid on the regular hypercubic grid (henceforth ``G'') that spans
+We perform full Bayesian parameter estimation for a set of simulated signals
+whose parameters are the vertices of the following regular hypercubic grid
+(henceforth ``G''):
 $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
 The final population is given by substituting all of the $N$ events, generated
-as above, with their respective nearest neighbours on the grid G.
+as above, with their respective nearest neighbours on the parameter grid G.
 
 In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show results for a population
-of $N=50$ events with $\lambdans=800$. In the left panel, we show the 
-probability distributions for $\lambdans$ inferred from $N$ events with $N$
-ranging from $1-50$. We also mark the $90\%$ confidence interval on $\lambdans$
-as dark grey vertical lines corresponding to each cumulative distribution.
-Distributions are normalized to unit area. We see that the information coming
-from the first few observations does not limit $\lambdans$ much more than our
-prior from Eq.~\ref{eq:lprior} does. However, the median of the distribution 
-tracks the true value of $\lambdans$ fast, reaching close within as few as $4$
+of $N=80$ events. The neutron star deformability is fixed at $\lambdans=800$
+for all events. Each curve shows the probability distributions for $\lambdans$
+as inferred from $N$ events, with $N$ ranging from $1-50$. We also mark the 
+$90\%$ confidence intervals for $\lambdans$ as dark grey vertical lines,
+one associated with each probability distribution curve. Distributions are 
+normalized to unit area. From the figure, we can visualize the information 
+information flow. The first few observations do not have enough information 
+to bound $\lambdans$ (with $90\%$ confidence) much more than our prior from
+Eq.~\ref{eq:lprior} does. However, the median of these cumulative distributions
+track the true value of $\lambdans$ rapidly, reaching close in as few as $4$
 observations.
-% 
-In the right panel, we show our measurement uncertainty $\delta\Lambda/\Lambda$
+
+In Fig.~\ref{fig:TT_Lambda_vs_N_L500_2000_CI90_0}, we show our measurement 
+uncertainty for $\lambdans$ for different populations, with $\lambdans$ alone
+changing between populations. Lets consider the top left panel first, which
+corresponds to a population of the most deformable neutron stars considered,
+with $\lambdans=2000$. The green filled region shows $\delta\Lambda/\Lambda$
 as a function of $N$, where we take $\delta\Lambda$ to be the width of the 
 $90\%$ confidence interval on $\lambdans$ after $N$ observations.
-This is given by the green filled region (left y-axis).
+%
 The green filled circles show the median measured value of $\lambdans$

### 2016-04-09 emcee_PE/scripts/MultipleObservations.py

@@ -9,6 +9,8 @@ import matplotlib.pyplot as plt
 plt.rcParams.update({'text.usetex' : True})
 from mpl_toolkits.axes_grid1 import ImageGrid
 
+from glob import glob
+
 from pydoc import help
 from scipy import stats
 from scipy.stats.stats import pearsonr
@@ -234,7 +236,10 @@ What it does:
                  only_positive_aligned_spins=True,\
                  kernel='gau',\
                  bw_method='scott',\
+                 data_prefix='data',\
                  write_data=True,\
+                 post_process=False,\
+                 RND=0,\
                  verbose=False):
         ### REMOVE BH spins < 0
         if verbose:
@@ -258,14 +263,17 @@ What it does:
         self.bw_method = bw_method
         #
         self.verbose = verbose
+        self.data_prefix = data_prefix
         self.write_data = write_data
         #
         # TAG for storing data to disk
-        tmp_dir = 'L%d_N%d_' % (NSLambda, N)
-        self.RND = int(np.random.random()*1e5)
-        self.TAG = tmp_dir + str(self.RND) + '/'
-        try: os.makedirs(self.TAG)
-        except: print "Warning: temporary dir %s already exists!!" % self.TAG
+        if not post_process:
+            tmp_dir = '%s/L%d_N%d_' % (data_prefix, NSLambda, N)
+            if RND: self.RND = int(RND)
+            else: self.RND = int(np.random.random()*1e5)
+            self.TAG = tmp_dir + ('%d/' % self.RND)
+            try: os.makedirs(self.TAG)
+            except: print "Warning: temporary dir %s already exists!!" % self.TAG
         #
         self.print_info()
         return
@@ -353,8 +361,21 @@ What it does:
         return chain_set
     #####
     ###
-    def load_events(self, chain_set_number,\
+    def find_chain_sets(self, NSLambda=None, N=None):
+        '''
+This function finds the indexes of possible population sets (chain sets) that
+are available in the given data directory. The data directory is initialized
+by the constructor.
+        '''
+        if NSLambda == None: NSLambda = self.NSLambda
+        if N == None: N = self.N
+        dir_tag = '%s/L%d_N%d_*' % (self.data_prefix, NSLambda, N)
+        return [int(x.split('_')[-1]) for x in glob(dir_tag)]
+    #####
+    ###
+    def load_events(self, chain_set_number, N=None,\
                       lambda_posterior_chains=None,\
+                      precalculate_norms=False,\
[…]

### 2016-04-09 emcee_PE/scripts/MultipleObservations.py
patched MultipleObservations to sort available chain sets by time

@@ -370,7 +370,9 @@ by the constructor.
         if NSLambda == None: NSLambda = self.NSLambda
         if N == None: N = self.N
         dir_tag = '%s/L%d_N%d_*' % (self.data_prefix, NSLambda, N)
-        return [int(x.split('_')[-1]) for x in glob(dir_tag)]
+        dir_list = glob(dir_tag)
+        dir_list.sort(key=os.path.getmtime)
+        return [int(x.split('_')[-1]) for x in dir_list]
     #####
     ###
     def load_events(self, chain_set_number, N=None,\

### 2016-04-09 emcee_PE/scripts/MultipleEventGeneration-N80-20.py
script to generate multiple events offline

@@ -1,0 +1,1340 @@
+#!/usr/bin/env python
+# Tools for analysis of MCMC samples from emcee for spin estimation project
+# Copyright 2015
+# 
+__author__ = "Prayush Kumar <prayush.kumar@ligo.org>"
+
+import os, sys, time
+import commands as cmd
+import numpy as np
+from matplotlib import mlab, cm, use
+import matplotlib.pyplot as plt
+plt.rcParams.update({'text.usetex' : True})
+from mpl_toolkits.axes_grid1 import ImageGrid
+
+#get_ipython().magic(u'pylab inline')
+from pylab import *
+
+
+from pydoc import help
+from scipy import stats
+from scipy.stats.stats import pearsonr
+from scipy.stats import gaussian_kde
+from scipy.stats import norm
+from scipy.interpolate import UnivariateSpline
+from scipy.optimize import minimize_scalar
+import scipy.integrate as si
+
+from sklearn.neighbors import KernelDensity
+from sklearn.grid_search import GridSearchCV
+
+from statsmodels.nonparametric.kde import KDEUnivariate
+from statsmodels.nonparametric.kernel_density import KDEMultivariate
+
+try:
+    sys.path.append( os.path.dirname(os.path.realpath(__file__)) + '/../src/' )
+    sys.path.append( os.path.dirname(os.path.realpath(__file__)) )
+except:
+    sys.path.append( os.path.dirname(cmd.getoutput('pwd -P')) + '/../src/' )
+    sys.path.append( os.path.dirname(cmd.getoutput('pwd -P')) )
+
+sys.path.append( '/home/prayush/src/UseNRinDA/scripts/' )
+sys.path.append( '/home/prayush/src/UseNRinDA/plotting' )
+
+sys.path.append('/home/prayush/src/nsbh_tidal/emcee_PE/src/')
+sys.path.append('/home/prayush/src/nsbh_tidal/emcee_PE/scripts/')
+
+from utils import *
+import pycbc.pnutils as pnutils
+from PlotOverlapsNR import make_contour_plot_multrow
+import h5py
+
+import MultipleObservations as MO
+#reload(MO)
+
+# Constants
+verbose  = True
+vverbose = True
+debug    = True
+
+quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
+quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
+percentiles_68 = 100*np.array(quantiles_68)
[…]

### 2016-04-09 emcee_PE/scripts/MultipleEventGeneration-N80-20.py
mending MultipleEventGeneration script ot run on Vulcan

@@ -139,7 +139,7 @@ Nburnin  = 2000
 # Read in parameter biases and other data
 ######################################################
 datafile_postfix = 'TT_ParameterBiasesAndConfidenceIntervals.h5'
-data_dir         = '/home/prayush/research/NSBH/TidalParameterEstimation/set006/'
+data_dir         = '/home/prayush/projects/nsbh/TidalParameterEstimation/nsbh_tidal/PEData/'
 
 # Load in all the Posterior distribution function samples
 all_posterior_chains   = {}
@@ -181,22 +181,10 @@ for q in qvec:
   fin.close()
 
 
+#########################
 # ## READ/ORGANIZE DATA
-
-# In[ ]:
-
-## INITIATE A SET OF MULTIPLE OBSERVATIONS
 NEVENTS = int(sys.argv[1])
-INDEX   = int(sys.argv[2])
 
-plotdir = 'plots/N%d-%d/' % (NEVENTS, INDEX)
-try: os.makedirs(plotdir)
-except: pass
-
-
-# In[ ]:
-
-fout = open(os.path.join(plotdir, 'chain_indices.dat'), 'w')
 #
 xx = {}
 for i, NSL in enumerate(Lambdavec):
@@ -208,6 +196,13 @@ for i, NSL in enumerate(Lambdavec):
                                           kernel='gau',\
                                           verbose=True)
         xx_prev = xx[NSL]
+        ## INITIATE A SET OF MULTIPLE OBSERVATIONS
+        INDEX   = xx_prev.RND #int(sys.argv[2])
+        #
+        plotdir = 'plots/N%d_%d/' % (NEVENTS, INDEX)
+        try: os.makedirs(plotdir)
+        except: pass
+        fout = open(os.path.join(plotdir, 'chain_indices.dat'), 'w')
     else:
         xx[NSL] = MO.multiple_observation_results(lambda_posterior_chains,\
                                           N=NEVENTS, NSLambda=NSL,\
@@ -222,12 +217,7 @@ for i, NSL in enumerate(Lambdavec):
 fout.close()
 
 
-# In[ ]:
 
-
-
-
-# In[ ]:
 
 ## GENERATE A SEQUENCE OF EVENTS
 for NSL in Lambdavec:

### 2016-04-09 emcee_PE/scripts/MultipleEventGeneration-N80-20.py
minor changes to script

@@ -198,6 +198,7 @@ for i, NSL in enumerate(Lambdavec):
         xx_prev = xx[NSL]
         ## INITIATE A SET OF MULTIPLE OBSERVATIONS
         INDEX   = xx_prev.RND #int(sys.argv[2])
+        print "\n\n>> USING RANDOM = %d\n\n" % INDEX
         #
         plotdir = 'plots/N%d_%d/' % (NEVENTS, INDEX)
         try: os.makedirs(plotdir)

### 2016-04-12 Paper/paper.tex
update errorbar vs N figure

@@
 \centering    
-\includegraphics[width=.9\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L2000.pdf}
-\includegraphics[width=.9\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1500.pdf}\\
-\includegraphics[width=.67\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L1000.pdf}
-\includegraphics[width=.67\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
-\includegraphics[width=.67\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L500.pdf}
+\includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L2000.pdf}
+\includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1500.pdf}\\
+\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1000.pdf}
+\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
 \caption{Filled-region plots showing the median and $90\%$ confidence intervals
@@
 
+The pioneering terrestrial observation of gravitational waves by Advanced LIGO
+harbingers the dawn of an era of gravitational-wave astronomy where 
+observations, more than theory, would play the driving role. Binaries of 
+compact stellar objects evolve under the radiation reaction force of emitting
+gravitational waves.
+
 Conventional astronomical methods have, to date, made observations of ~2500

### 2016-04-13 Paper/paper.tex
updated all mult obs figs
some new population figures

@@
 % 
-% \begin{figure}
-% \centering    
-% \includegraphics[width=1.\columnwidth]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
-% \caption{HELLO
-% }
-% \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}
-% \end{figure}
-%
-% 
-\begin{figure*}
+\begin{figure}
 \centering    
-\includegraphics[width=1\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
-\includegraphics[width=1\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
-\caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 
-confidence interval for $\lambdans$, normalized by its true value,
- as a function of the number of events observed $N$. In the inset, we
- show the same measurement uncertainty, but not normalized, and on a 
- logarithmic axis for $N$ as well.
-% 
- From both the inset and the main plot, we find that for the first $3-4$ events
- our measurement is prior limited. As $N$ increases, the uncertainty falls 
- roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
- (than $1/\sqrt{N}$) narrowing of the confidence interval.
-%  
- {\it Right}: This figure shows the fractional difference between the median
+\includegraphics[trim=20 0 0 0, width=1.05\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
+\caption{This figure shows the fractional difference between the median
  recovered value from the posterior probability distribution for $\lambdans$
@@
 }
+\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}
+\end{figure}
+%
+% 
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=1.03\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
+% \includegraphics[width=1.03\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
+% \caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 
+% confidence interval for $\lambdans$, normalized by its true value,
+%  as a function of the number of events observed $N$. In the inset, we
+%  show the same measurement uncertainty, but not normalized, and on a 
+%  logarithmic axis for $N$ as well.
+% % 
+%  From both the inset and the main plot, we find that for the first $3-4$ events
+%  our measurement is prior limited. As $N$ increases, the uncertainty falls 
+%  roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
+%  (than $1/\sqrt{N}$) narrowing of the confidence interval.
+% %  
+%  {\it Right}: This figure shows the fractional difference between the median
+%  recovered value from the posterior probability distribution for $\lambdans$
+%  as a function of $N$. Different curves correspond to different injected values
+%  of $\lambdans$, and are shaded from light to dark in direct proportion. 
+%  The horizontal dashed brown line marks $10\%$ deviation of the median from the 
+%  true value.
+% %  
+% We find that within $\approx 10$ detections, the median measured value for
+% $\lambdans$ would estimate the true value for the star with the remarkable 
+% accuracy of $10\%$.
+% }
+% \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
+% \end{figure*}
+%LambdaCIWidths_vs_N_AllPopulations_Log_L
+% 
+%
+% 
+\begin{figure*}
+\centering    
+\includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
+\includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}\\
+\includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
+\includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
+\includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
+\caption{ In the main panel is shown the width of the measured $90\%$ 
+confidence interval for $\lambdans$, normalized by its true value,
+ as a function of the number of events observed $N$. In the inset, we
+ show the same measurement uncertainty, but not normalized, and on a 
+ logarithmic axis for $N$ as well.
+% 
+ From both the inset and the main plot, we find that for the first $3-4$ events
+ our measurement is prior limited. As $N$ increases, the uncertainty falls 
+ roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
+ (than $1/\sqrt{N}$) narrowing of the confidence interval.
+}
 \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
@@
 % 
-% 
 \begin{figure}
 \centering    
-\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_Lambda.pdf}
+\includegraphics[width=1.05\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_N.pdf}
 \caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
@@
 % 
-\begin{figure*}
+\begin{figure}
 \centering    
-\includegraphics[width=\columnwidth]{plots/LambdalErrorBars_vs_Lambda_N79_Log.pdf}
-\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N.pdf}
+% \includegraphics[width=1.03\columnwidth]{plots/LambdaRelErrorBars_vs_Lambda_AllPopulations_N80_Log.pdf}
+\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N_AllPopulations.pdf}
 \caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
@@
 \label{fig:TT_Lambda_vs_Lambda_L500_2000_CI90_0_AllInOne}
-\end{figure*}
+\end{figure}
 

### 2016-04-13 emcee_PE/scripts/MultipleObservations.py

@@ -240,6 +240,7 @@ What it does:
                  write_data=True,\
                  post_process=False,\
                  RND=0,\
+                 verbose_info=True,\
                  verbose=False):
         ### REMOVE BH spins < 0
         if verbose:
@@ -274,8 +275,12 @@ What it does:
             self.TAG = tmp_dir + ('%d/' % self.RND)
             try: os.makedirs(self.TAG)
             except: print "Warning: temporary dir %s already exists!!" % self.TAG
+        else:
+            if RND is not None and RND >= 0: self.RND = int(RND)
+            else:
+                raise IOError("Please provide the chain number as RND")
         #
-        self.print_info()
+        if verbose_info: self.print_info()
         return
     ###
     def generate_events(self, lambda_posterior_chains=None,\

### 2016-04-13 Paper/paper.tex
alternative to multipanel fig.8
pdf plots for 80 events

@@
 \title{
-Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO
+Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO: a Bayesian study
 }
@@
 \caption{%
-% {\it Left}:
 Posterior probability distributions for $\lambdans$ (colored curves), and 
@@
 different number of accumulated observations $N$.
-% {\it Right}: Filled-region plots showing the median and $90\%$ confidence intervals
-% for $\lambdans$, as a function of the number of observed events $N$.
-%  Shown are two independent filled regions. 
-% %  
-%  One in green shows the recovered confidence interval
-%  for $\lambdans$ (on the left y-axis), normalized by its true value,
-%  as a function of the 
-%  number of observed events $N$. The recovered median is shown by the
-%  green line-circled curve. The pair of horizontal green dashed (dotted)
-%  lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
-%  the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
-% % 
-%  The second filled region, in grey, shows the recovered
-%  confidence interval for neutron star compactness, without normalizing.
-%  The corresponding y-values are shown on the *right* y-axis. The horizontal grey
-%  dashed line traces the true value of $\lambdans$, and is what we expect
-%  the filled region to approach with increasing number of observations.
-% %  
-%  Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
 }
@@
 %
+\begin{figure}
+\centering    
+\includegraphics[width=1.05\columnwidth]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
+\caption{%
+Posterior probability distributions for $\lambdans$ (colored curves), and 
+associated $90\%$ confidence intervals (grey vertical lines), shown for 
+different number of accumulated observations $N$.
+}
+\label{fig:TT_Lambda_vs_N_CI90_0}
+\end{figure}
 % 

### 2016-04-13 Paper/paper.tex
added no-mass-gap pop plots + updated old ones

@@
 % 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[trim=20 0 0 0, width=1.05\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
+\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
+\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}
 \caption{This figure shows the fractional difference between the median
@@
 \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}
-\end{figure}
+\end{figure*}
 %
@@
 \centering    
-\includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
-\includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}\\
-\includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
-\includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
-\includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L2000.pdf}\\
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1500.pdf}\\
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1000.pdf}\\
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L800.pdf}\\
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
+\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L500.pdf}
 \caption{ In the main panel is shown the width of the measured $90\%$ 
@@
 \end{figure*}
+% 
+% % 
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
+% \includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}\\
+% \includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
+% \includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
+% \includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
+% \caption{ In the main panel is shown the width of the measured $90\%$ 
+% confidence interval for $\lambdans$, normalized by its true value,
+%  as a function of the number of events observed $N$. In the inset, we
+%  show the same measurement uncertainty, but not normalized, and on a 
+%  logarithmic axis for $N$ as well.
+% % 
+%  From both the inset and the main plot, we find that for the first $3-4$ events
+%  our measurement is prior limited. As $N$ increases, the uncertainty falls 
+%  roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
+%  (than $1/\sqrt{N}$) narrowing of the confidence interval.
+% }
+% \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
+% \end{figure*}
+% % 
 % 

### 2016-04-14 Paper/paper.tex
s5 text first pass done

@@
 % 
-\begin{figure}
+\begin{figure*}
 \centering    
 \includegraphics[width=1.05\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}
-% \includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}
+\includegraphics[width=1.\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
 \caption{%
-Posterior probability distributions for $\lambdans$ (colored curves), and 
-associated $90\%$ confidence intervals (grey vertical lines), shown for 
-different number of accumulated observations $N$.
+These panels show the accumulation of information from a population whose
+$\lambdans=800$.
+{\it Left:} Posterior probability distributions for $\lambdans$ (colored 
+curves), and associated $90\%$ confidence intervals (grey vertical lines),
+shown for different number of accumulated observations $N$. Distributions are
+normalized to unit area.
+{\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
+associated $90\%$ confidence intervals (as the height of filled region), shown
+as a function of number of observations $N$. Solid horizontal line indicates
+the true value of $\lambdans$. Dashed and dotted horizontal lines (a pair for each
+line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error bounds.
 }
 \label{fig:TT_Lambda_vs_N_L800_CI90_0}
-\end{figure}
-%
-% 
-\begin{figure*}
-\centering    
-\includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L2000.pdf}
-\includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1500.pdf}\\
-\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1000.pdf}
-\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
-\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
-\caption{Filled-region plots showing the median and $90\%$ confidence intervals
-for $\lambdans$ measurement, as a function of the number of observed events $N$.
-Different panels correspond to different injected $\lambdans$ values, which is shown
-in their title. In each panel are shown two independent filled regions. 
-%  
- One in green shows the recovered confidence interval
- for $\lambdans$ (on the left y-axis), normalized by its true value,
- as a function of the 
- number of observed events $N$. The recovered median is shown by the
- green line-circled curve. The pair of horizontal green dashed (dotted)
- lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
- the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
-% 
- The second filled region, in grey, shows the recovered
- confidence interval for neutron star compactness, without normalizing.
- The corresponding y-values are shown on the *right* y-axis. The horizontal grey
- dashed line traces the true value of $\lambdans$, and is what we expect
- the filled region to approach with increasing number of observations.
-%  
- Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
-}
-\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}
 \end{figure*}
+%
+% % 
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L2000.pdf}
+% \includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1500.pdf}\\
+% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1000.pdf}
+% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
+% \caption{Filled-region plots showing the median and $90\%$ confidence intervals
+% for $\lambdans$ measurement, as a function of the number of observed events $N$.
+% Each panel corresponds a unique population with injected $\lambdans$ fixed to
+% the value shown in its title. In each panel, the measured median for $\lambdans$
+% is shown by filled circles, and the $90\%$ confidence interval is shown by the 
+% height of the filled region. In addition, the dashed and 
+% dotted lines (2 of each line-style) show the $\pm 25\%$ and $\pm 50\%$ error
+% bounds, with respect to the true value.
+% }
+% \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}
+% \end{figure*}
 %
@@
 \caption{%
-Posterior probability distributions for $\lambdans$ (colored curves), and 
-associated $90\%$ confidence intervals (grey vertical lines), shown for 
-different number of accumulated observations $N$.
+This filled-region figure shows how our measurement of $\lambdans$ improves with
+an increasing number of observed events $N$. Each color corresponds to an
+independent populations with the true value of $\lambdans$ fixed as given in
+the legend. For each population, we show the measured median (as filled
+circles) and $90\%$ confidence intervals for $\lambdans$ measurement (as the
+height of the filled region about the median). 
 }
@@
 \includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}
-\caption{This figure shows the fractional difference between the median
+\caption{%
+{\it Left:} This figure shows the fractional difference between the median
  recovered value from the posterior probability distribution for $\lambdans$
@@
 accuracy of $10\%$.
+{\it Right:} 
 }
-\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}
+\label{fig:TT_LambdaMedian_vs_N_AllInOne}
 \end{figure*}
@@
 }
-\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
+\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}
 \end{figure*}
@@
 }
-\label{fig:TT_Lambda_vs_Lambda_L500_2000_CI90_0_AllInOne}
+\label{fig:TT_PowerLawLambdaErrorVsLambda}
 \end{figure}
@@
 % 
-First we show a crude calculation which tells us that approximately $8$ (or $16$) 
+We start with a crude calculation which tells us that approximately $8$ (or $16$) 
 observations
@@
 
-We want to distribute signals in effective distance $\deff$, which is a combination of 
-luminosity distance, inclination angle of the binary's orbital angular momentum
-with respect to the detector, and its various sky angles. The signal strength 
-scales inversely with $\deff$, if it comprises only of the dominant $(2,2)$ mode, allowing
-for various angle-dependent factors to scale out beside the luminosity distance.
-% 
-In order to consider multiple events, lets imagine a population uniformly distributed
-in effective volume, i.e. within a sphere of radius $\deff^\mathrm{max}$. The 
-radius of this sphere is set by the lowest SNR that is distinguishable from noise
-by LIGO searches. Now, divide the sphere into $I$ shells of equal width. The radius
-of the $i$-th sphere would then be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The 
-measurement uncertainty in $\lambdans$ scales inversely with the SNR, and hence
-directly with the effective distance to the source. Therefore, if we have a 
-measurement error $\sigma_0$ for a source located at $\deff = D_0$, the same error
-for the same source located within the $i$-th shell would be 
-$\sigma_i = \sigma_0 \dfrac{D_i}{D_0}.$
-Independent measurements of $\lambdans$ for identical sources at different distances
-would have their combined error given by
+The approximation of using (only) the dominant $l=|m|=2$ multipoles of
+gravitational-wave strain to construct the signal, ignoring other $l\neq 2$
+multipoles, has the advantage of making various angle parameters 
+associated with the source degenerate with its distance. As a result, source
+distance, source sky location angles, and the inclination angle of
+the binary's orbital angular momentum with respect to the detector,
+all combine into a single ``effective distance'' $\deff$ that scales
+out as the constant amplitude factor $1/\deff$ (constant because none
+of the above parameters vary temporally for aligned-spin binaries).
+% 
+In order to consider multiple events, lets imagine a population uniformly
+distributed in effective volume, i.e. within a sphere of radius 
+$\deff^\mathrm{max}$. The radius of this sphere is set by the lowest SNR that 
+is distinguishable from noise by LIGO searches. Now, divide the sphere into
+$I$ shells of equal thickness. The radius of the $i$-th sphere would then
+be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The measurement uncertainty in
+$\lambdans$ scales inversely with SNR, and hence directly with the 
+effective distance to the source. Therefore, if we have a measurement error 
+$\sigma_0$ for a source located at $\deff = D_0$, the same error for the same
+source located within the $i$-th shell would be 
+$\sigma_i = \sigma_0 \dfrac{D_i}{D_0}$. The combined error from independent
+measurements of $\lambdans$ for identical sources at different distances is
+given by
 \begin{equation}\label{eq:1oversigma}
@@
 % to an integral. 
-Continuing on this line of inquiry, Ref.~\cite{Markakis:2010mp} calculates the 
+Continuing on this line of inquiry, Ref.~\cite{Markakis:2010mp} found the 
 root-mean-square averaged measurement error from 
@@
 \end{equation}
-given a fiducial pair $(\sigma_0, D_0)$.
-Further, it is straightforward to deduce from Eq.~\ref{eq:rmsSigmaIdenticalSources}
+given a fiducial pair $(\sigma_0, D_0)$. It is straightforward to deduce from
+Eq.~\ref{eq:rmsSigmaIdenticalSources}
 that the same measurement certainty as afforded by a single
@@
 
-There are two main caveats to this rudimentary calculation, which was outlined in 
-Ref.~\cite{Markakis:2010mp}, (i) it applies to identical sources,
+There are two main caveats to this rudimentary calculation, first outlined in 
+Ref.~\cite{Markakis:2010mp}: (i) it applies to identical sources,
 which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
@@
 once binary parameters are fixed, there is a one-to-one mapping between
-effective distance and SNR. Finally, in order to mitigate computational
-cost, we make an additional simplification.
-We perform full Bayesian parameter estimation for a set of simulated signals
-whose parameters are the vertices of the following regular hypercubic grid
-(henceforth ``G''):
+effective distance and SNR. Additionally, we consider an alternate population
+paradigm (paradigm B, with paradigm A being the standard one described 
+above), where populations are generated as above but further restricting
+mass-ratios such that source BH masses lie {\it outside} the astrophysical 
+mass-gap~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}.
+Finally, in order to mitigate computational cost, we make an additional 
+simplification. We perform full Bayesian parameter estimation for a set of 
+simulated signals whose parameters are the vertices of the following regular
+hypercubic grid (henceforth ``G''):
 $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
@@
 
-In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show results for a population
-of $N=80$ events. The neutron star deformability is fixed at $\lambdans=800$
-for all events. Each curve shows the probability distributions for $\lambdans$
-as inferred from $N$ events, with $N$ ranging from $1-50$. We also mark the 
-$90\%$ confidence intervals for $\lambdans$ as dark grey vertical lines,
-one associated with each probability distribution curve. Distributions are 
-normalized to unit area. From the figure, we can visualize the information 
-information flow. The first few observations do not have enough information 
-to bound $\lambdans$ (with $90\%$ confidence) much more than our prior from
-Eq.~\ref{eq:lprior} does. However, the median of these cumulative distributions
-track the true value of $\lambdans$ rapidly, reaching close in as few as $4$
-observations.
-
-In Fig.~\ref{fig:TT_Lambda_vs_N_L500_2000_CI90_0}, we show our measurement 
-uncertainty for $\lambdans$ for different populations, with $\lambdans$ alone
-changing between populations. Lets consider the top left panel first, which
-corresponds to a population of the most deformable neutron stars considered,
-with $\lambdans=2000$. The green filled region shows $\delta\Lambda/\Lambda$
-as a function of $N$, where we take $\delta\Lambda$ to be the width of the 
-$90\%$ confidence interval on $\lambdans$ after $N$ observations.
-%
-The green filled circles show the median measured value of $\lambdans$
-as a function of $N$. The dashed line at y value $=1$ marks zero error, i.e.
-$\delta\Lambda/\Lambda=0$. The dotted and dashed lines around this denote
-$\pm 25\%$ and $\pm 50\%$ relative errors, respectively. Independently, the
-light-grey filled shown the absolute measurement uncertainty $\delta\Lambda$
-on the right y-axis , as a function of $N$. A horizontal grey dashed line
-marks the true value of $\lambdans$. Contours of $1/\sqrt{N}$ are shown as 
-dotted light grey curves to guide the eye while studying the dependence of
-$\delta\Lambda$ on $N$.
-%
+In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show illustrative results
+for a single 
+population of $N=80$ events. The neutron star deformability is fixed at
+$\lambdans=800$ for all events. In the left panel, each curve shows the 
+probability distributions for $\lambdans$ as inferred from $N$ events, with $N$
+ranging from $1-80$. We also mark the $90\%$ confidence intervals associated
+with each of the probability distribution curves. From the figure, we
+can visualize information accumulation. The first few observations do not have
+enough information to bound $\lambdans$ (with $90\%$ confidence) much more than
+our prior from Eq.~\ref{eq:lprior} does. However, the median of these
+cumulative distributions appears to track the true value of $\lambdans$ 
+rapidly.
+% 
+In the right panel, we present information derived from the left panel.
+The line-circle curve shows the measured median from $N$ observations, with
+the latter on the $x$-axis. The pair of dashed (dotted) horizontal lines mark
+$\pm25\%$ ($\pm50\%$) error bars. At each $N$, the range spanned by the 
+filled region is the $90\%$ confidence interval deduced from the same 
+events. This figure somewhat quantifies the qualitative deductions we made
+from the left panel. We find that the measured median does track the true value
+quickly, reaching $10\%$ accuracy with $10-15$ observations. With the same
+information, our confidence intervals also shrink to $\pm 25\%$.
+% 
+In Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0} we show further results from $4$
+independent populations, with the source $\lambdans$ alone changing between
+them. The sampled values, i.e.$\lambdans=\{500,1000,1500,2000\}$, 
+range from the least to the most deformable neutron stars. As in the 
+right panel of Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}, the line-circle curves
+track the median measured $\lambdans$, while the filled regions show
+the associated $90\%$ confidence intervals. From the figure, we observe
+the following: (i) the shrinkage of confidence interval widths with increasing
+$N$ happens in a similar manner for all 
+populations, (ii) within $5-10$ observations, median measured $\lambdans$ for
+all populations lie within $\pm10\%$ of the true values, and (iii) it takes
+approximately $20$ events to distinguish definitively (with $90\%$ confidence)
+between deformable NSs with $\lambdans=2000$ and compact NSs with 
+$\lambdans=500$, or equivalently to distinguish between hard, moderate and soft
+nuclear equations of state. This is comparable to what has been found for
+monolithic binaries of neutron stars~\cite{DelPozzo:13,Chatziioannou:2015uea,
+Agathos:2015a}.
+
+So far we have discussed individual realizations of NSBH populations. The 
+underlying stochasticity of the population generation process makes it
+difficult to draw generalized inferences (from a single realization) about the
+measurability of $\lambdans$. We mitigate this by drawing multiple
+realizations of population generation (i.e. multiple identically constructed
+populations). Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} shows the first
+results for multiple populations. Lets focus on the left panel first. We
+show the median measured value of $\lambdans$ from $100$ independent 
+populations with fixed $\lambdans=2000$ as the top set of curves. The one 
+highlighted (in green) curve corresponds to the population discussed above.
+Similarly, we generate sets of $100$ populations with $\lambdans=500-1500$, and
+show the median measured $\lambdans$ as four distinct groups of curves in
+Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} (with the single highlighted
+curve in each group corresponding to
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+Dash-dotted horizontal lines demarcate $\pm50\%$ error intervals around
+the true $\lambdans$ values. All of these populations are drawn under
+paradigm A (i.e. with BH masses allowed to sample the mass-gap that spans
+$2-5M_\odot$).
+% 
+From the figure, we can conclude with a high likelihood that our measured
+$\lambdans$ values will be within $10\%$ of the real value in $15-20$ events
+for less deformable neutron stars with $\lambdans\leq 1000$, and in as few as
+$10$ events for more deformable stars with $\lambdans\geq 1500$.
+% 
+The right panel of the figure is identical to the left panel, but drawn under
+population paradigm B (no BHs alloewd within the mass-gap). Under this
+paradigm, we expectedly find that information accumulation is slower. It would
+take $20-40$ detections with $\rho\geq10$ under this paradigm, for our (median)
+measured $\lambdans$ to lie within $10\%$ of the true value (fewer for 
+more deformable NSs, as one would expect).
+% 
+This is a promising deduction, as one might expect to see $\mathcal{O}(10)$
+events over a few years' timescale with design aLIGO~\cite{Abadie:2010cfa}.
+
+Further, we investigate the statistical uncertainties associated with
+$\lambdans$ measurements. We use $90\%$ confidence intervals as our measure of
+the same. The left column of 
+Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne} shows results
+under paradigm A, where BH mass priors include the mass-gap. In each panel
+we show the width of the $90\%$ confidence interval measured from $N$ events
+(with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
+the population. In grey are shown results from each of $100$ population
+realizations, with the single population highlighted in color corresponding
+to the one we focused on in 
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+The inset shows the same, except that the ordinate is {\it not} normalized by
+the true value of $\lambdans$. From top to bottom, the population $\lambdans$
+decreases from $\lambdans=2000\rightarrow 500$, corresponding to decreasingly
+deformable NSs with softer equations of state. We observe the following:
+(i) for moderately-hard to hard equations of state with $\lambdans\geq 1000$,
+we can constrain $\lambdans$ within $\pm 50\%$ using only $10-20$ events, and
+within $\pm 25\%$ with $25-40$ events; (ii) for softer equations of state with
+$\lambdans<1000$, we will achieve the same accuracy with $20-30$ and $50+$ 
+events, respectively; and (iii) for the first $5$ or so observations, our
+measurement confidence spans the entire prior allowed range 
+$\lambdans\in[0,4000]$, as shown by the plateauing towards the left edge in
+each panel's inset.
+% 
+In the right column, each panel panel is identical to its corresponding left
+one, with the difference that populations are drawn under paradigm B, which
+does {\it not} allow for BH masses to fall within the mass-gap. We find that
+for NSs with $\lambdans\leq 1000$, it would take $25-40$ events to
+constrain $\lambdans$ within $\pm 50\%$ and $50+$ events to constrain it
+within $\pm 25\%$. This is somewhat slower than paradigm A, as is to be
+expected since here we preclude the lowest mass-ratios, which correspond to
+signals with largest tidal signatures. For $\lambdans>1000$ we find that we
+can constrain $\lambdans$ within $\pm 50\%$ with a similar number of events as
+for paradigm A, but will need more ($30-40$, as compared to $25-40$) events
+to further constrain it to within $\pm 25\%$ of the true value.
+% 
+These are some of the primary findings of this paper.
+
+Finally, we quantitatively explore the dependence of our statistical
+uncertainties for $\lambdans$ on the number of events, as well as on the true
+NS deformability itself. First, we will focus on the dependence on $N$. We
+assume a power-law dependence of the form
+$\delta\lambdans\propto\ 1/N^\alpha$. For each of the $100$ populations 
+for each of $\lambdans=500-2000$, we compute the exponent $\alpha$ as a
+function of the number of observed events $N$, and show it in 
+Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN}. There are $100\times5=500$ curves
+on the figure, with one highlighted for each value of population's $\lambdans$.
+These highlighted values are only special in the sense that they correspond to
+populations discussed earlier in this section (c.f.
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}).
+We immediately observe two things, (i) there is a globally similar dependence
+on $N$ for all populations, and (ii) information accumulates faster than
+$1/\sqrt{N}$. In fact, we find that if
+$\delta\lambdans\propto\frac{1}{N^\alpha}$, $\alpha$ lines in the range
+$0.7_{-0.2}^{+0.2}$.
+% 
+Next, we focus on the dependence of $\delta\lambdans$ on $\lambdans$ of the
+population itself. As suggested by Fisher-matrix studies~\cite{Lackey:2013axa},
+and as for $N$, we assume the form $\delta\lambdans\propto\lambdans^\beta$.
+From each set of $100$ populations with a given $\lambdans$ value, we draw one
+at random, and form a set of $5$ similarly drawn populations, one for each of
+$\lambdans=\{500,800,1000,1500,2000\}$. With each set, we determine $\beta$
+for different number of observed events $N$. In all, we make $100$ independent
+$5-$population sets and show the value of $\beta$ measured from each in 
+Fig.~\ref{fig:TT_PowerLawLambdaErrorVsLambda}. We find that the assumed
+relation $\delta\lambdans\propto\lambdans^\beta$ gets fairly robust for 
+larger values of $N$, with $\beta$ converging to $\beta=0.5^{+0.33}_{-0.33}$.
+The fact that $\beta<1$ implies that the relative error
+$\delta\lambdans/\lambdans$ decreases with increasing $\lambdans$, while the 
+absolute error increases.
+% 
+From these results, we conclude that the measurement uncertainty for
+$\lambdans$ after $N$ observations is
+\begin{equation}
+ \delta\lambdans\propto \dfrac{\lambdans^{0.5^{+0.33}_{-0.33}}}{N^{0.7_{-0.2}^{+0.2}}}.
+\end{equation}
+We also find that while these results are inferred from paradigm A populations,
+paradigm B gives very similar results.
+
+To summarize, in this section we study the improvement in our measurement of 
+NS deformability parameter $\lambdans$ with an increasing number of events. We
+do so by simulating plausible populations of disrupting NSBH binaries (with
+$\rho\geq 10$). We find that:
+(i) for more deformable neutron stars (harder equation of states), the median
+measured value of $\lambdans$ comes within $10\%$ of the true value with as 
+few as $10$ events, while achieving the same accuracy for softer equations of 
+state will take $15-20$ source detections; (ii) the statistical uncertainty
+associated with $\lambdans$ measurement shrinks to within $\pm50\%$ with
+$10-20$ events, and to within $\pm 25\%$ with $50+$ events, when source 
+$\lambdans\geq 1000$; (iii) for softer equations of state, the same could take
+$25-40$ and $50+$ events, respectively for the two uncertainty thresholds;
+and (iv) if BHs really do observe the astrophysical mass-gap, the information
+accumulation is somewhat slower than if they do not. We conclude that within
+$20-30$ observations, aLIGO would begin to place very interesting bounds on 
+the NS deformability, which would allow us to rule out or rank different
+equations of state for neutron star matter.
 

### 2016-04-14 emcee_PE/scripts/MultipleObservations.py

@@ -286,7 +286,7 @@ What it does:
     def generate_events(self, lambda_posterior_chains=None,\
                         NSLambda=None,\
                         qvec=None, chi2vec=None, SNRvec=None,\
-                       qmin=2.0, qmax=5.0, chi2min=0.0, chi2max=1.0):
+                       qmin=20./9., qmax=5.0, chi2min=0.0, chi2max=1.0):
         '''
         Returns is a list of events. For each event, two objects are returned:
         1. array of posterior samples

### 2016-04-14 Paper/paper.tex
fixed fig captions from fig.7-12

@@
 {\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
-associated $90\%$ confidence intervals (as the height of filled region), shown
-as a function of number of observations $N$. Solid horizontal line indicates
-the true value of $\lambdans$. Dashed and dotted horizontal lines (a pair for each
-line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error bounds.
+associated $90\%$ confidence intervals (as the vertical extent of filled
+region), shown as a function of number of observations $N$. Solid horizontal
+line indicates the true value of $\lambdans$. Dashed and dotted horizontal
+lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
+bounds.
 }
@@
 \caption{%
-This filled-region figure shows how our measurement of $\lambdans$ improves with
-an increasing number of observed events $N$. Each color corresponds to an
-independent populations with the true value of $\lambdans$ fixed as given in
-the legend. For each population, we show the measured median (as filled
-circles) and $90\%$ confidence intervals for $\lambdans$ measurement (as the
-height of the filled region about the median). 
+In this figure, the filled regions show how our measurement of $\lambdans$
+improves as the number of observed events ($N$, shown on $x$-axis) increases.
+Each color corresponds to an independent population with its true value of
+$\lambdans$ given in the legend. For each population, we show the median 
+measured $\lambdans$ value (as filled circles), as well as the associated
+$90\%$ confidence intervals for the measurement (as the vertical extent of the
+filled region about the median), as functions of $N$.
 }
@@
 \caption{%
-{\it Left:} This figure shows the fractional difference between the median
- recovered value from the posterior probability distribution for $\lambdans$
- as a function of $N$. Different curves correspond to different injected values
- of $\lambdans$, and are shaded from light to dark in direct proportion. 
- The horizontal dashed brown line marks $10\%$ deviation of the median from the 
- true value.
-%  
-We find that within $\approx 10$ detections, the median measured value for
-$\lambdans$ would estimate the true value for the star with the remarkable 
-accuracy of $10\%$.
-{\it Right:} 
+{\it Left:} This figure shows the median value of the recovered
+probability distribution for $\lambdans$, as a function of the number of events
+in the population $N$. There are four family of curves, one corresponding each
+to $\lambdans=\{500,1000,1500,2000\}$, with $100$ independent populations
+within each family. One curve in each family is highlighted in color, 
+representing that it is the same population as was illustrated in
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+In the same color we show $\pm 10\%$ error-bounds on $\lambdans$ with
+horizontal dash-dotted lines. We observe that within $10-25$ observations, 
+the median of the measured cumulative probability distribution for $\lambdans$
+will lie within $10\%$ of the true value.
+{\it Right:} This figure is identical to the left panel, with the only
+difference that the BH masses in each population are restricted to lie
+{\it outside} the astrophysical mass-gap (i.e. paradigm B). The difference that
+we observe under this paradigm is that we need more ($30+$) events to achieve 
+the same ($10\%$) measurement accuracy for populations with $\lambdans<1000$.
+For more deformable neutron stars, $10-25$ events would suffice.
 }
@@
 \centering    
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L2000.pdf}\\
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1500.pdf}\\
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1000.pdf}\\
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L800.pdf}\\
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
-\includegraphics[width=0.85\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L500.pdf}
-\caption{ In the main panel is shown the width of the measured $90\%$ 
-confidence interval for $\lambdans$, normalized by its true value,
- as a function of the number of events observed $N$. In the inset, we
- show the same measurement uncertainty, but not normalized, and on a 
- logarithmic axis for $N$ as well.
-% 
- From both the inset and the main plot, we find that for the first $3-4$ events
- our measurement is prior limited. As $N$ increases, the uncertainty falls 
- roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
- (than $1/\sqrt{N}$) narrowing of the confidence interval.
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L2000.pdf}\\
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1500.pdf}\\
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1000.pdf}\\
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L800.pdf}\\
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
+\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L500.pdf}
+\caption{%
+These figures show the width of the $90\%$ confidence interval for $\lambdans$
+(normalized by its true value), as a function of the number of observed events
+$N$. The left column panels show populations sampled under paradigm A, which
+allows BH masses to fall within the astrophysical mass-gap, while those on the
+right show populations drawn under paradigm B which respects the mass-gap.
+Each panel corresponds to a unique value of populations' $\lambdans$,
+decreasing from $2000\rightarrow 500$ as we go from top to bottom. In each panel,
+$100$ curves are shown, each corresponding to an independent population draw. The
+one colored curve in each panel highlights the population used in
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+The insets in each panel show the same information, with the ordinate {\it not}
+normalized by the true value of $\lambdans$.
+% 
+We find that with approximately $25$ or so events, we begin to put
+statistically meaningful constraints on $\lambdans$, restricting it to within
+$\pm 50\%$ of the true value. We can expect to achieve this with a few years
+of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of the
+confidence intervals will require $40+$ events.
 }
@@
 \includegraphics[width=1.05\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_N.pdf}
-\caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
-as a function of $\lambdans$ itself. The drawn curves are cumulative, in the sense 
-that the $N=n$ curve includes information from all of the first $n$ observed events.
-Those corresponding to the first $N\leq 5$ events are shown in fading shades of red,
-with the ones for higher value of $N$ being shown in grey (going from light to dark in
-proportion to $N$).
-Green dash-dotted lines are contours of $y=x^{-(1+1/5)}$.
-% 
-\prayush{We note that the measurement uncertainty from a population follows the
-equation for the green lines.}
-% 
+\caption{%
+Assuming a power-law dependence of the measurement error on the number of
+events: $\delta\lambdans\propto 1/N^\alpha$, we show $\alpha$ in this figure
+as a function of the number of observed events $N$. Shown are five families
+of $100$ population draws each, with each family corresponding to one of
+$\lambdans=\{500,800,1000,1500,2000\}$. Each grey curve corresponds to one
+of these $100\times5 = 500$ populations. The thicker curves, one from each
+family, shows the population we discussed in
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+We find that a power-law is a good approximation for the concerned dependence,
+and information accumulates {\it faster} than $1/\sqrt{N}$. We estimate
+$\alpha\simeq 0.7^{+0.2}_{-0.2}$.
 }
@@
 \includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N_AllPopulations.pdf}
-\caption{This figure shows the relative uncertainty in the measurement of $\lambdans$,
-as a function of $\lambdans$ itself. The drawn curves are cumulative, in the sense 
-that the $N=n$ curve includes information from all of the first $n$ observed events.
-Those corresponding to the first $N\leq 5$ events are shown in fading shades of red,
-with the ones for higher value of $N$ being shown in grey (going from light to dark in
-proportion to $N$).
-Green dash-dotted lines are contours of $y=x^{-(1+1/5)}$.
-% 
-\prayush{We note that the measurement uncertainty from a population follows the
-equation for the green lines.}
-% 
+\caption{%
+In this figure, which is similar to Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN},
+we quantify the dependence of $\delta\lambdans$ on $\lambdans$ itself. Of 
+the five families of simulated NSBH populations, we construct $100$
+independent sets taking one population from each family. With each of 
+these $100$ sets, and assuming a power-law dependence:
+$\delta\lambdans\propto\lambdans^\beta$, we estimate $\beta$ and show it in
+this figure as a function of the number of observed events $N$.
+% 
+We find that $\beta$ can be estimated to lie within $[1/6,5/6]$ with a
+likely value close to $1/2$. Since $0<\beta<1$, the relative error
+$\delta\lambdans/\lambdans$ {\it decreases} as the star gets more 
+deformable, while the absolute error $\delta\lambdans$ {\it increases}.
 }
@@
 We immediately observe two things, (i) there is a globally similar dependence
-on $N$ for all populations, and (ii) information accumulates faster than
+on $N$ for all populations, and (ii) information accumulates {\it faster} than
 $1/\sqrt{N}$. In fact, we find that if
@@
 larger values of $N$, with $\beta$ converging to $\beta=0.5^{+0.33}_{-0.33}$.
-The fact that $\beta<1$ implies that the relative error
-$\delta\lambdans/\lambdans$ decreases with increasing $\lambdans$, while the 
-absolute error increases.
+The fact that $0<\beta<1$ implies that the relative error
+$\delta\lambdans/\lambdans$ {\it decreases} with increasing $\lambdans$, while
+the absolute error {\it increases}.
 % 

### 2016-04-14 Paper/paper.tex
fixed typo in refs

@@
 representing that it is the same population as was illustrated in
-Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
 In the same color we show $\pm 10\%$ error-bounds on $\lambdans$ with
@@
 one colored curve in each panel highlights the population used in
-Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
 The insets in each panel show the same information, with the ordinate {\it not}
@@
 family, shows the population we discussed in
-Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
 We find that a power-law is a good approximation for the concerned dependence,
@@
 curve in each group corresponding to
-Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}).
 Dash-dotted horizontal lines demarcate $\pm50\%$ error intervals around
@@
 to the one we focused on in 
-Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
 The inset shows the same, except that the ordinate is {\it not} normalized by
@@
 populations discussed earlier in this section (c.f.
-Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}).
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}).
 We immediately observe two things, (i) there is a globally similar dependence

### 2016-04-14 Paper/paper.tex
include MO in intro

@@
 laid out, we report our results in Sec.~\ref{s1:multiple_observations}.
-% 
-We find \prayush{[...]}
+In addition, we probe two paradigms, one that allows for BH masses to lie
+within the astrophysical mass-gap (paradigm A), and one that does not (paradigm
+B).
+% 
+For paradigm A, we find the following: (i) for the softer equations of state
+that result in less deformable neutron stars, $15-20$ detections are sufficient
+for our median measured value of $\lambdans$ to lie within $10\%$ of the true
+value. (ii) For NSBH populations with more deformable NSs
+($\lambdans\geq 1000$), the same is achievable within as few as $10$ realistic
+observations. (iii) The statistical uncertainty associated with $\lambdans$
+measurement can be restricted to be within $\pm50\%$ using $10-20$ observations
+when $\lambdans\geq 1000$), and using $25-40$ observations for softer equations
+of state. All of the above is possible within a few years of design
+aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
+masses between $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
+restricting $\lambdans$ will require $50+$ NSBH observations.
+% 
+For paradigm B, we find the information accumulation to be somewhat slower.
+While the quantitative inferences for $\lambdans\lgeq1000$ populations are
+not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
+required to attain the same measurement accuracy as under paradigm A.
+% 
+We conclude that within as few as $20-30$ observations of disruptive NSBH
+mergers, aLIGO will begin to place interesting bounds on NS deformability.
+This, amongst other things, will allow us to rank different equations of 
+state for neutron star matter from most to least likely, with a few years'
+time-scale.
 

### 2016-04-14 Paper/paper.tex
placeholder text in discussion
updated paradigm B figs

@@
 For paradigm B, we find the information accumulation to be somewhat slower.
-While the quantitative inferences for $\lambdans\lgeq1000$ populations are
+While the quantitative inferences for $\lambdans\geq1000$ populations are
 not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
@@
 
+\prayush{[PK: To be continued after revision of the introduction.]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-04-18 Paper/paper.tex
updated author list and acknowledgments

@@
 Am M\"uhlenberg, G\"olm, Germany}} %
+\newcommand{\CIFAR}{\affiliation{Canadian Institute for Advanced Research, 
+    180 Dundas St.~West, Toronto, ON M5G 1Z8, Canada}} %
 
@@
 \author{Michael P\"urrer}\AEI
-\author{People across the Atlantic}
+\author{Harald P. Pfeiffer}\CITA\AEI
 
@@
 \begin{acknowledgments}
-Acknowledgments
+We thank Ben Lackey, Francesco Pannarale, Francois Foucart, and Duncan Brown
+    for helpful discussions. We gratefully acknowledge support
+  for this research at CITA from NSERC of Canada, the Ontario Early 
+  Researcher Awards Program, the Canada Research
+  Chairs Program, and the Canadian Institute for Advanced Research; at
+  Caltech from the Sherman Fairchild Foundation and NSF grants
+  PHY-1404569 and AST-1333520; at Cornell from the
+  Sherman Fairchild Foundation and NSF grants PHY-1306125 and
+  AST-1333129; and at Princeton from NSF grant PHY-1305682 and the
+  Simons Foundation.  Calculations were performed at the Vulcan
+  supercomputer at the Albert Einstein Institute;
+  H.P. and P.K. thank the Albert-Einstein Institute,
+  Potsdam, for hospitality during part of the time where this research
+  was completed.
 \end{acknowledgments}

### 2016-05-04 Paper/paper.tex
notes tex
making changes from comments

@@
 \author{Michael P\"urrer}\AEI
-\author{Harald P. Pfeiffer}\CITA\AEI
+\author{Harald P. Pfeiffer}\CITA\CIFAR\AEI
 
@@
 \begin{abstract}
-Very recently, the Advanced LIGO (aLIGO) observatories announced the first
-terrestrial detection of gravitational waves. This heralds an era of 
-observational gravitational wave (GW) astronomy. Neutron star - black hole
-(NSBH) binaries are one of the primary sources targeted for observation
-with the aLIGO detectors.
-% 
-GWs from these sources will carry signatures of the tidal distortion of
-the NS by the BH during the inspiral, and of its tidal disruption (if that
-happens) at merger. In the NS is disrupted, the merger and ringdown signal
-is drastically suppressed, which distinguishes it from a BHBH merger.
-% 
-In this paper, we present the first Bayesian study of the measurability of
-the tidal deformability of neutron stars, allowing the companion hole to take 
-on a variety of masses and spins.
+The pioneering discovery of gravitational waves (GW) by Advanced LIGO has ushered
+us into an era of observational GW astrophysics. Compact binaries remain the 
+primary target sources for GW observation, of which neutron star - black hole 
+(NSBH) binaries form an important subset.
+% 
+GWs from NSBH sources carry signatures of (a) the star's tidal distortion by
+the companion black hole (BH) during inspiral, and (b) its tidal disruption
+(if that happens) near merger.
+%
+In this paper, we present the first fully Bayesian study of the measurability 
+of neutron star (NS) tidal deformability $\lambdans\propto (R/M)_\mathrm{NS}^{5}$ from NSBH
+observation(s), where we
+take into account the crucial effect of BH spins as well.
 % 
@@
 parameters for an NSBH signal, a significant bias is introduced in the 
-inference if the signal-to-noise ratio (SNR) is greater than $\rho\simeq30$.
-% 
-Second, we find that if we are to get a loud signal with $\rho\geq 23$, 
-we can put a factor of two bound on the NS tidal deformability parameter
-$\lambdans\propto (R/M)^{5}$.
+inference if the signal-to-noise ratio (SNR) is greater than $\simeq30$.
+% 
+Second, we find that if we are to get a loud signal with SNR $\gtrsim 23$, 
+we can put a factor of two bound on $\lambdans$.
 % 
@@
 we find that (i) the median measured $\lambdans$ catches up to within
-$10\%$ of the true value in approximately $20$ observations; (ii) the $90\%$
-confidence interval for $\lambdans$ shrinks to $\pm 50\%$ around the true 
-value with the same number of observations.
+$10\%$ of the true value in approximately $20$ observations; (ii) in the same
+number of observations, the associated $90\%$ confidence intervals shrink to
+$\pm 50\%$ of the true value.
 % 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Advanced LIGO (aLIGO) observatories began their first observing run ``O1''
-mid-September 2015, operating at a factor of $3-4$ higher gravitational-wave
+% 
+% Over the $3$-months of O1, aLIGO had access to more binary mergers than 
+% initial-LIGO over its entire coincident operation period.
+
+The Advanced LIGO (aLIGO) observatories completed their first observing run ``O1''
+early-2016, operating at a factor of $3-4$ higher gravitational-wave
 (GW) strain sensitivity than their first-generation 
@@
 % 
-During O1, the first terrestrial detection of gravitational waves was made, as
-was announced earlier in 2016~\cite{LIGOVirgo2016a}. Emitted more than a
-billion year ago, these waves herald the era of observational gravitational
-wave (GW) astronomy as they pass through Earth.
-% 
-Over the $3$-months of O1, aLIGO had access to more binary mergers than 
-initial-LIGO over its entire coincident operation period. Both of aLIGO 
-instruments will be further upgraded in phases to reach their design 
-sensitivity by $2018-19$, at which point we may expect to observe $\sim 70$ 
-binary mergers a year~\cite{Abadie:2010cf}.
-
-Coalescing binaries of compact objects, i.e. black holes (BH) and/or neutron 
-stars (NS), remain the most interestings targets for GW based observation with
-aLIGO. NSBH binaries form an important sub-class of compact binaries, and we
-expect to observe $\mathcal{O}(10)$ of these every year with design 
-sensitivity aLIGO~\cite{Abadie:2010cf}.
-% 
-NSBH mergers are of interest for various reasons. For instance, they have 
-been associated with (as the progenitors of) short Gamma-ray Bursts
-(GRBs)~\cite{eichler:89,1992ApJ...395L..83N,moch:93,Barthelmy:2005bx,
+During O1, aLIGO made the first terrestrial observation of gravitational 
+waves~\cite{LIGOVirgo2016a}. Emitted by a pair of coalescing black holes, these
+waves
+heralded an era of observational GW astrophysics as they traveled through Earth.
+% 
+By $2018-19$, we expect the aLIGO instruments to reach their design
+sensitivity, at which point we may expect to observe $\sim 70$ binary mergers a
+year~\cite{Abadie:2010cf}.
+
+Coalescing binaries of compact objects, i.e. black holes (BH) and/or
+neutron stars (NS) are the primary target sources for 
+aLIGO~\cite{Abbott:2016nhf}. An important sub-class of these sources are NSBH
+binaries, of which we expect $\mathcal{O}(10)$ GW observations every 
+year~\cite{Abadie:2010cf}.
+NSBH binaries are of interest for multiple reasons. For instance,
+they have been long associated with (as possible progenitors of) short
+Gamma-ray Bursts (GRBs)~\cite{eichler:89,1992ApJ...395L..83N,moch:93,Barthelmy:2005bx,
 2005Natur.437..845F,2005Natur.437..851G,Shibata:2005mz,Paschalidis2014,
-Tanvir:2013}. Under favorable conditions, close to merger the neutron star
-can get disrupted by the tidal field of its companion hole. Once disrupted,
-the neutron star material may form an accretion disk around the black hole. 
-Most of the disk matter spirals in within a few milliseconds, with a small 
-fraction becoming unbounded and getting ejected, with a possibled simultaneous
-burst of gamma-ray emission)~\cite{Foucart2015,Lovelace:2013vma,Deaton2013,
-Foucart2012,Shibata:2005mz,Paschalidis2014}. A coincident detection of GWs
-from an NSBH system with a GRB would therefore provide us insights into the
-central engine powering GRBs. 
+Tanvir:2013}. Under favorable conditions, the NS can get disrupted by
+the tidal field of its companion BH. Once disrupted, most of the NS
+material falls into the BH over an $\mathcal{O}(1$ms$)$ time-scale,
+with the rest forming an accretion disk around the BH. This disk-BH system,
+which survives for $0.1-1$s, is hypothesized to drive SGRBs through the
+production of relativistic jets~\cite{Foucart2015,
+Lovelace:2013vma,Deaton2013,Foucart2012,Shibata:2005mz,Paschalidis2014}.
+A {\it coincident} detection of both gravitational and electromagnetic
+waves (gamma rays) from an NSBH system is therefore a {\it unique}
+possibility to confirm this hypothesis~\cite{Abbott:2016wya}.
+% 
+% Most of the matter falls into the BH in ~1ms. But the disk itself has an
+% accretion timescale of ~100ms-1s. The disk-BH system is the expected engine
+% for a SGRB, through the production of a relativistic jet. The unbound material,
+% on the other hand, is responsible for an EM signal in the infrared/optical, 
+% powered by radioactive decays in the ejecta. And possibly a radio afterglow
+% through interactions with the ISM. But not gamma-rays (unless you are referring
+% to the very small amount of material ejected in the jet, but that's not what
+% people usually mean when talking about material unbound by the merger).
 % 
@@
 nature of matter at nuclear densities', i.e. what is the equation of state
-(EoS) of neutron stars. This is what we will focus on in this paper.
-% 
-The gravitational-waves emitted by NSBH binaries carry subtle hints of the 
-matter structure of the neutron star. During early inspiral, the (weak) tidal
-field of the companion hole deforms the star, thereby exciting its resonant
-oscillation modes. The induced quadrupolar moment of the star depends on the
-tidal field $\mathcal{E}_{ij}$ due to its companion through the relation
-$Q_{ij} = -\lambda\mathcal{E}_{ij}$, where $\lambda$ is the EoS dependent
-tidal deformability parameter related to the neutron star's dimensionless
-Love number $k_2$ and radius $R_\mathrm{NS}$ as
-$\lambda:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$. This drains energy from the orbit
-resulting in an increase of the inspiral rate. We know the leading and next-to
+(EoS) of neutron star material. This is what we will focus on in this paper.
+% 
+GWs emitted by NSBH coalescences carry subtle hints of the material structure
+of NSs. During early inspiral, the (weak) tidal field of the companion BH
+deforms the star, thereby exciting its resonant oscillation modes. The induced
+quadrupolar moment $Q_{ij}$ of the star depends on the
+tidal field $\mathcal{E}_{ij}$ of its companion through the relation
+$Q_{ij} = -\lambda\mathcal{E}_{ij}$, where $\lambda$ is an EoS-dependent
+tidal deformability parameter that is related to NS's dimensionless Love number
+$k_2$ and radius $R_\mathrm{NS}$ as $\lambda:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$.
+The coherence of GW emission between stellar quadrupolar waves and orbital
+waves drains energy from the orbit and increases the inspiral rate~\cite{Flanagan2008}.
+We know the leading and next-to
 -leading order terms in post-Newtonian (PN) theory that capture this 
-effect~\cite{Vines2011}, entering binary phasing at $5$PN order. Close to
-merger, as the NS gets disrupted and forms long tails that feed into the 
-companion black hole, the quadrupole moment of the system falls monotonically.
-As a result, the emitted GWs are subdued significantly post-disruption
-(as compared to what they would be for a BBH),
-and the quasi-normal modes of the resulting black hole are not 
-excited~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart2015,
+effect~\cite{Vines2011}, entering binary phasing at $5$PN order. As the NS gets
+disrupted close to binary's merger, the quadrupole moment of the system falls
+monotonically. As a result, the emitted GWs are subdued significantly
+post-disruption (as compared to what they would be for a 
+BBH)~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart2015,
 Pannarale:2015jia}.
-% 
-% 
+
 The difficulty in discerning NS matter effects in an NSBH coalescence from its
@@
 As in rest of the paper, we restrict here our parameter space to span 
-mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin $\chibh\in[-0.5, +0.75]$, 
-and dimensionless neutron star tidal deformability 
+mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin (aligned with orbit)
+$\chibh\in[-0.5, +0.75]$, and dimensionless neutron star tidal deformability 
 $\lambdans:= G\left(\frac{c^2}{G \mns}\right)^5\lambda \in[500, 2000]$.
@@
 In order to keep the computational cost tractable, we make a further 
-approximation. We inject signals into zero noise at different SNRs and recover
+approximation.
+%
+\prayush{%
+We inject signals into zero noise at different SNRs and recover
 their parameters from Bayesian inferencing for a fixed set of binaries,
@@
 $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+% 
 Once the population parameters been selected, each event's parameters are
-replaced by their nearest neighbor in the above set. We make sure that this 
-approximation is conservative by keeping our population sampling ranges within the
-conservative side of the range spanned by the above set. With the caveats
-laid out, we report our results in Sec.~\ref{s1:multiple_observations}.
-In addition, we probe two paradigms, one that allows for BH masses to lie
-within the astrophysical mass-gap (paradigm A), and one that does not (paradigm
-B).
+replaced by their nearest neighbor in the above set.
+}
+We make sure that this approximation is conservative by keeping our population
+sampling ranges within the conservative side of the range spanned by the above
+set. With the caveats laid out, we report our results in
+Sec.~\ref{s1:multiple_observations}. In addition, we probe two paradigms, one
+that allows for BH masses to lie within the astrophysical mass-gap (paradigm
+A), and one that does not (paradigm B).
 % 

### 2016-05-24 Paper/paper.tex
fixed refs

@@
 \title{
-Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO: a Bayesian study
+Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of mixed binary observations
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 
-% Over the $3$-months of O1, aLIGO had access to more binary mergers than 
-% initial-LIGO over its entire coincident operation period.
 
@@
 % 
-During O1, aLIGO made the first terrestrial observation of gravitational 
+During O1, they made the first terrestrial observation of gravitational 
 waves~\cite{LIGOVirgo2016a}. Emitted by a pair of coalescing black holes, these
-waves
-heralded an era of observational GW astrophysics as they traveled through Earth.
+waves heralded an era of observational GW astrophysics as they traveled 
+through Earth.
+% 
+By the end of the decade, we expect aLIGO to reach its design sensitivity,
+as well as for Advanced Virgo and KAGRA to begin observing. 
 % 
@@
 which survives for $0.1-1$s, is hypothesized to drive SGRBs through the
-production of relativistic jets~\cite{Foucart2015,
+production of relativistic jets~\cite{Foucart:2015a,
 Lovelace:2013vma,Deaton2013,Foucart2012,Shibata:2005mz,Paschalidis2014}.
@@
 post-disruption (as compared to what they would be for a 
-BBH)~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart2015,
+BBH)~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart:2015a,
 Pannarale:2015jia}.
@@
 %
-\begin{figure}
-\centering    
-\includegraphics[width=1.05\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_N.pdf}
-\caption{%
-Assuming a power-law dependence of the measurement error on the number of
-events: $\delta\lambdans\propto 1/N^\alpha$, we show $\alpha$ in this figure
-as a function of the number of observed events $N$. Shown are five families
-of $100$ population draws each, with each family corresponding to one of
-$\lambdans=\{500,800,1000,1500,2000\}$. Each grey curve corresponds to one
-of these $100\times5 = 500$ populations. The thicker curves, one from each
-family, shows the population we discussed in
-Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
-We find that a power-law is a good approximation for the concerned dependence,
-and information accumulates {\it faster} than $1/\sqrt{N}$. We estimate
-$\alpha\simeq 0.7^{+0.2}_{-0.2}$.
-}
-\label{fig:TT_PowerLawLambdaErrorVsN}
-\end{figure}
-%
-% 
-\begin{figure}
-\centering    
-% \includegraphics[width=1.03\columnwidth]{plots/LambdaRelErrorBars_vs_Lambda_AllPopulations_N80_Log.pdf}
-\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N_AllPopulations.pdf}
-\caption{%
-In this figure, which is similar to Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN},
-we quantify the dependence of $\delta\lambdans$ on $\lambdans$ itself. Of 
-the five families of simulated NSBH populations, we construct $100$
-independent sets taking one population from each family. With each of 
-these $100$ sets, and assuming a power-law dependence:
-$\delta\lambdans\propto\lambdans^\beta$, we estimate $\beta$ and show it in
-this figure as a function of the number of observed events $N$.
-% 
-We find that $\beta$ can be estimated to lie within $[1/6,5/6]$ with a
-likely value close to $1/2$. Since $0<\beta<1$, the relative error
-$\delta\lambdans/\lambdans$ {\it decreases} as the star gets more 
-deformable, while the absolute error $\delta\lambdans$ {\it increases}.
-}
-\label{fig:TT_PowerLawLambdaErrorVsLambda}
-\end{figure}
 
@@
 
+% Finally, we quantitatively explore the dependence of our statistical
+% uncertainties for $\lambdans$ on the number of events, as well as on the true
+% NS deformability itself. First, we will focus on the dependence on $N$. We
+% assume a power-law dependence of the form
+% $\delta\lambdans\propto\ 1/N^\alpha$. For each of the $100$ populations 
+% for each of $\lambdans=500-2000$, we compute the exponent $\alpha$ as a
+% function of the number of observed events $N$, and show it in 
+% Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN}. There are $100\times5=500$ curves
+% on the figure, with one highlighted for each value of population's $\lambdans$.
+% These highlighted values are only special in the sense that they correspond to
+% populations discussed earlier in this section (c.f.
+% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}).
+% We immediately observe two things, (i) there is a globally similar dependence
+% on $N$ for all populations, and (ii) information accumulates {\it faster} than
+% $1/\sqrt{N}$. In fact, we find that if
+% $\delta\lambdans\propto\frac{1}{N^\alpha}$, $\alpha$ lines in the range
+% $0.7_{-0.2}^{+0.2}$.
+% % 
+% Next, we focus on the dependence of $\delta\lambdans$ on $\lambdans$ of the
+% population itself. As suggested by Fisher-matrix studies~\cite{Lackey:2013axa},
+% and as for $N$, we assume the form $\delta\lambdans\propto\lambdans^\beta$.
+% From each set of $100$ populations with a given $\lambdans$ value, we draw one
+% at random, and form a set of $5$ similarly drawn populations, one for each of
+% $\lambdans=\{500,800,1000,1500,2000\}$. With each set, we determine $\beta$
+% for different number of observed events $N$. In all, we make $100$ independent
+% $5-$population sets and show the value of $\beta$ measured from each in 
+% Fig.~\ref{fig:TT_PowerLawLambdaErrorVsLambda}. We find that the assumed
+% relation $\delta\lambdans\propto\lambdans^\beta$ gets fairly robust for 
+% larger values of $N$, with $\beta$ converging to $\beta=0.5^{+0.33}_{-0.33}$.
+% The fact that $0<\beta<1$ implies that the relative error
+% $\delta\lambdans/\lambdans$ {\it decreases} with increasing $\lambdans$, while
+% the absolute error {\it increases}.
+% % 
+% From these results, we conclude that the measurement uncertainty for
+% $\lambdans$ after $N$ observations is
+% \begin{equation}
+%  \delta\lambdans\propto \dfrac{\lambdans^{0.5^{+0.33}_{-0.33}}}{N^{0.7_{-0.2}^{+0.2}}}.
+% \end{equation}
+% We also find that while these results are inferred from paradigm A populations,
+% paradigm B gives very similar results.
+
+To summarize, in this section we study the improvement in our measurement of 
+NS deformability parameter $\lambdans$ with an increasing number of events. We
+do so by simulating plausible populations of disrupting NSBH binaries (with
+$\rho\geq 10$). We find that:
+(i) for more deformable neutron stars (harder equation of states), the median
+measured value of $\lambdans$ comes within $10\%$ of the true value with as 
+few as $10$ events, while achieving the same accuracy for softer equations of 
+state will take $15-20$ source detections; (ii) the statistical uncertainty
+associated with $\lambdans$ measurement shrinks to within $\pm50\%$ with
+$10-20$ events, and to within $\pm 25\%$ with $50+$ events, when source 
+$\lambdans\geq 1000$; (iii) for softer equations of state, the same could take
+$25-40$ and $50+$ events, respectively for the two uncertainty thresholds;
+and (iv) if BHs really do observe the astrophysical mass-gap, the information
+accumulation is somewhat slower than if they do not. We conclude that within
+$20-30$ observations, aLIGO would begin to place very interesting bounds on 
+the NS deformability, which would allow us to rule out or rank different
+equations of state for neutron star matter.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}\label{s1:discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The pioneering terrestrial observation of gravitational waves by Advanced LIGO
+harbingers the dawn of an era of gravitational-wave astronomy where 
+observations, more than theory, would play the driving role. Binaries of 
+compact stellar objects evolve under the radiation reaction force of emitting
+gravitational waves.
+
+Conventional astronomical methods have, to date, made observations of ~2500
+NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
+Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
+stellarcollapsemass}, although
+tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
+%
+Observations of over 2500 NSs have shown that the spin of NSs
+$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
+%
+On the other hand, indirect observations of stellar-mass BHs place their
+masses between $5-35M_\odot$, with their spin angular momenta 
+$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
+(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
+examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
+Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
+and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
+
+\prayush{[PK: To be continued after revision of the introduction.]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Acknowledgments
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgments}
+We thank Ben Lackey, Francesco Pannarale, Francois Foucart, and Duncan Brown
+    for helpful discussions. We gratefully acknowledge support
+  for this research at CITA from NSERC of Canada, the Ontario Early 
+  Researcher Awards Program, the Canada Research
+  Chairs Program, and the Canadian Institute for Advanced Research; at
+  Caltech from the Sherman Fairchild Foundation and NSF grants
+  PHY-1404569 and AST-1333520; at Cornell from the
+  Sherman Fairchild Foundation and NSF grants PHY-1306125 and
+  AST-1333129; and at Princeton from NSF grant PHY-1305682 and the
+  Simons Foundation.  Calculations were performed at the Vulcan
+  supercomputer at the Albert Einstein Institute;
+  H.P. and P.K. thank the Albert-Einstein Institute,
+  Potsdam, for hospitality during part of the time where this research
+  was completed.
+\end{acknowledgments}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{appendix}
+
+\section{Phenomenology of $\lambdans$ measurement errors}
+% 
+\begin{figure}
+\centering    
+\includegraphics[width=1.05\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_N.pdf}
+\caption{%
+Assuming a power-law dependence of the measurement error on the number of
+events: $\delta\lambdans\propto 1/N^\alpha$, we show $\alpha$ in this figure
+as a function of the number of observed events $N$. Shown are five families
+of $100$ population draws each, with each family corresponding to one of
+$\lambdans=\{500,800,1000,1500,2000\}$. Each grey curve corresponds to one
+of these $100\times5 = 500$ populations. The thicker curves, one from each
+family, shows the population we discussed in
+Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
+We find that a power-law is a good approximation for the concerned dependence,
+and information accumulates {\it faster} than $1/\sqrt{N}$. We estimate
+$\alpha\simeq 0.7^{+0.2}_{-0.2}$.
+}
+\label{fig:TT_PowerLawLambdaErrorVsN}
+\end{figure}
+%
+% 
+\begin{figure}
+\centering    
+% \includegraphics[width=1.03\columnwidth]{plots/LambdaRelErrorBars_vs_Lambda_AllPopulations_N80_Log.pdf}
+\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N_AllPopulations.pdf}
+\caption{%
+In this figure, which is similar to Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN},
+we quantify the dependence of $\delta\lambdans$ on $\lambdans$ itself. Of 
+the five families of simulated NSBH populations, we construct $100$
+independent sets taking one population from each family. With each of 
+these $100$ sets, and assuming a power-law dependence:
+$\delta\lambdans\propto\lambdans^\beta$, we estimate $\beta$ and show it in
+this figure as a function of the number of observed events $N$.
+% 
+We find that $\beta$ can be estimated to lie within $[1/6,5/6]$ with a
+likely value close to $1/2$. Since $0<\beta<1$, the relative error
+$\delta\lambdans/\lambdans$ {\it decreases} as the star gets more 
+deformable, while the absolute error $\delta\lambdans$ {\it increases}.
+}
+\label{fig:TT_PowerLawLambdaErrorVsLambda}
+\end{figure}
+% 
 Finally, we quantitatively explore the dependence of our statistical
@@
 
-To summarize, in this section we study the improvement in our measurement of 
-NS deformability parameter $\lambdans$ with an increasing number of events. We
-do so by simulating plausible populations of disrupting NSBH binaries (with
-$\rho\geq 10$). We find that:
-(i) for more deformable neutron stars (harder equation of states), the median
-measured value of $\lambdans$ comes within $10\%$ of the true value with as 
-few as $10$ events, while achieving the same accuracy for softer equations of 
-state will take $15-20$ source detections; (ii) the statistical uncertainty
-associated with $\lambdans$ measurement shrinks to within $\pm50\%$ with
-$10-20$ events, and to within $\pm 25\%$ with $50+$ events, when source 
-$\lambdans\geq 1000$; (iii) for softer equations of state, the same could take
-$25-40$ and $50+$ events, respectively for the two uncertainty thresholds;
-and (iv) if BHs really do observe the astrophysical mass-gap, the information
-accumulation is somewhat slower than if they do not. We conclude that within
-$20-30$ observations, aLIGO would begin to place very interesting bounds on 
-the NS deformability, which would allow us to rule out or rank different
-equations of state for neutron star matter.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}\label{s1:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The pioneering terrestrial observation of gravitational waves by Advanced LIGO
-harbingers the dawn of an era of gravitational-wave astronomy where 
-observations, more than theory, would play the driving role. Binaries of 
-compact stellar objects evolve under the radiation reaction force of emitting
-gravitational waves.
-
-Conventional astronomical methods have, to date, made observations of ~2500
-NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
-Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
-stellarcollapsemass}, although
-tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
-%
-Observations of over 2500 NSs have shown that the spin of NSs
-$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
-%
-On the other hand, indirect observations of stellar-mass BHs place their
-masses between $5-35M_\odot$, with their spin angular momenta 
-$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
-(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
-examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
-Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
-and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
-
-\prayush{[PK: To be continued after revision of the introduction.]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgments
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgments}
-We thank Ben Lackey, Francesco Pannarale, Francois Foucart, and Duncan Brown
-    for helpful discussions. We gratefully acknowledge support
-  for this research at CITA from NSERC of Canada, the Ontario Early 
-  Researcher Awards Program, the Canada Research
-  Chairs Program, and the Canadian Institute for Advanced Research; at
-  Caltech from the Sherman Fairchild Foundation and NSF grants
-  PHY-1404569 and AST-1333520; at Cornell from the
-  Sherman Fairchild Foundation and NSF grants PHY-1306125 and
-  AST-1333129; and at Princeton from NSF grant PHY-1305682 and the
-  Simons Foundation.  Calculations were performed at the Vulcan
-  supercomputer at the Albert Einstein Institute;
-  H.P. and P.K. thank the Albert-Einstein Institute,
-  Potsdam, for hospitality during part of the time where this research
-  was completed.
-\end{acknowledgments}
+\end{appendix}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2016-05-24 Paper/paper.tex
polished abstrsact a bit

@@
 \usepackage{placeins}
+\usepackage{lineno}
+\linenumbers
 
@@
 %
-In this paper, we present the first fully Bayesian study of the measurability 
-of neutron star (NS) tidal deformability $\lambdans\propto (R/M)_\mathrm{NS}^{5}$ from NSBH
-observation(s), where we
-take into account the crucial effect of BH spins as well.
+In this paper, we present a Bayesian study of the measurability of
+neutron star (NS) tidal deformability $\lambdans\propto (R/M)_\mathrm{NS}^{5}$
+from GW observation(s) of disruptive NSBH coalescences, taking into account the
+crucial effect of BH spins.
 % 
@@
 % 
-By $2018-19$, we expect the aLIGO instruments to reach their design
-sensitivity, at which point we may expect to observe $\sim 70$ binary mergers a
-year~\cite{Abadie:2010cf}.
+At its design sensitivity, we expect aLIGO to observe $\sim 70$ compact binary
+mergers a year~\cite{Abadie:2010cf}.
 
@@
 neutron stars (NS) are the primary target sources for 
-aLIGO~\cite{Abbott:2016nhf}. An important sub-class of these sources are NSBH
-binaries, of which we expect $\mathcal{O}(10)$ GW observations every 
-year~\cite{Abadie:2010cf}.
+aLIGO~\cite{Abbott:2016nhf,LIGOVirgo2016a}. An important sub-class of these
+are NSBH binaries, of which we expect $\mathcal{O}(10)$ GW observations
+every year~\cite{Abadie:2010cf}.
 NSBH binaries are of interest for multiple reasons. For instance,

### 2016-05-24 Paper/paper.tex
added HP's comments to intro

@@
 % 
-By the end of the decade, we expect aLIGO to reach its design sensitivity,
-as well as for Advanced Virgo and KAGRA to begin observing. 
-% 
-At its design sensitivity, we expect aLIGO to observe $\sim 70$ compact binary
-mergers a year~\cite{Abadie:2010cf}.
-
+Towards the end of this decade, we expect aLIGO to reach its design sensitivity.
+In addition to the US-based efforts, we also expect the French-Italian detector Advanced
+Virgo~\cite{aVIRGO,aVirgo2}, Japanese detctor KAGRA~\cite{kagra,Somiya:2011np},
+and LIGO-India~\cite{2013IJMPD..2241010U} to begin observing at comparable
+sensitivities within a few years. With a global network of sensitive GW
+observatories, we can expect GW astronomy to face significant developments over the
+coming years.
+% 
+% At design sensitivity, we expect aLIGO to observe $\sim 70$ compact binary
+% mergers a year~\cite{Abadie:2010cf}.
+
+\red{
+Text about CBCs. How was each type observed so far?
+}
+\prayush{
 Coalescing binaries of compact objects, i.e. black holes (BH) and/or
 neutron stars (NS) are the primary target sources for 
-aLIGO~\cite{Abbott:2016nhf,LIGOVirgo2016a}. An important sub-class of these
+aLIGO~\cite{Abbott:2016nhf,LIGOVirgo2016a}.
+% 
+Compact binary merger rate estimates suggest
+a GW detection rate of approximately a few tens of binary black holes
+(BBH) every year~\cite{Abadie:2010cfa}. The actual masses of
+astrophysical black holes are uncertain, but observations and
+population synthesis studies suggest that BHs formed from stellar
+core-collapse can have masses up to and higher than
+$34M_\odot$~\cite{2008ApJ...678L..17S,2010ApJ...714.1217B}.  Also,
+recent measurements using continuum fitting and X-ray reflection
+fitting suggest that black holes can have high spin, with the BH
+angular momentum in dimensionless units exceeding
+$0.8$~\cite{2011ApJ...742...85G,2012MNRAS.424..217F,Gou:2013dna,
+  2009ApJ...697..900M,McClintock:2006xd,Miller:2009cw,McClintock:2013vwa}.
+Therefore the observations of GWs emitted by spinning BBHs will allow us to
+understand the spin-spin and spin-orbit dynamics of the two-body
+system, apart from allowing us to test strong-field dynamics of
+General Relativity.
+}
+%
+
+\red{
+Text describing BHNS disruption and accretion dynamics.
+}
+An important sub-class of these
 are NSBH binaries, of which we expect $\mathcal{O}(10)$ GW observations
@@
 %
+
+\red{
+Text focusing on tides
+}
 Another question that NSBH mergers can help answer is `what is the 
@@
 (disruption and merger). The earlier in inspiral the NS disrupts, the more
-orbits there are over which its effect is visible. The mechanism of disruption
+orbits there are over which its effect is visible.
+% 
+\red{
+Text about disruption of NSBHs, MOVE EARLIER.
+}
+\prayush{
+The mechanism of disruption
 is governed by the nature of the BH, i.e. its mass and spin. More massive BHs 
@@
 Kawaguchi:2015}.
+}
+
+\red{
+Insert paragraph describing impact on GWs
+}
 
@@
 Lee1999a,Lee1999b,Lee2000,oechslin:07,Read:2008iy,Markakis:2010mp,Markakis:2011vd,
-stergioulas:11,East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}. While single 
-observations are not as informative, Bayesian studies have shown that with
-$20-30$ BNS observations, aLIGO would begin to distinguish between hard,
-moderate and soft candidate equations of state~\cite{DelPozzo:13,
-Chatziioannou:2015uea,Agathos:2015a}. These studies rely on the PN 
+stergioulas:11,East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}. 
+Studies~\cite{DelPozzo:13,Chatziioannou:2015uea,Agathos:2015a} have shown that
+with $20-30$ BNS observations, aLIGO would begin to distinguish between hard,
+moderate and soft candidate equations of state. These studies rely on the PN 
 description of BNS dynamics, which is reasonable since BNSs merge at very high
@@
 information for NSBHs, and (ii) the errors due to unknown vaccum terms in PN 
-waveforms could easily dominate over the tidal terms 
-themselves~\cite{Yagi:2014}, since there are unknown point-particle terms at 
-orders {\it lower} than that at which tidal effects enter. Other studies that
+waveforms could easily dominate over the tidal terms themselves, since there 
+are unknown point-particle terms at orders {\it lower} than that at which tidal
+effects enter~\cite{Yagi:2014}.
+\red{
+Make clear the distinction between Fischer matrix studies and Bayesian ones. 
+Explain the differences. this is important as it sets the stage for for what is
+new in this paper.
+} 
+Other studies that
 use complete numerical simulations or derived models either use Fisher matrix 
@@
 include tidal effects during inspiral and merger in a consistent way, by using
-the waveform model of Ref.~\cite{Lackey:2013axa} (henceforth ``LEA+''). Second,
-we include the effect
+the waveform model of Ref.~\cite{Lackey:2013axa} (henceforth ``LEA+'').
+\red{
+(describe meaning of the acronym LEA+)
+}
+Second, we include the effect
 of black hole spin on tidal GW signals, in addition to the effect of BH mass 
@@
 
-We first probe the effect of not including tidal effects in parameter 
-estimation templates on the recovery of non-tidal binary parameters, such as 
+We first probe the effect of ignoring tidal effects in parameter 
+estimation on the recovery of non-tidal binary parameters, such as 
 component masses and spins. This is the case of current and planned aLIGO
@@
 % 
+\red{
+The following numerical details disrupt the flow of the text. MOVE after
+description of analyses.
+}
+\prayush{
 As in rest of the paper, we restrict here our parameter space to span 
@@
 Each signal is rescaled to a range of SNR values $\rho\in[10, 70]$.
-% 
+}
+\red{
+MOVE the following to SUMMARY. Here, keep 1-2 sentences:
+For individual events, ignoring tidal effects will affect mass-estimation
+only marginally; only for very loud signals (SNRs $\gtrsim 30$) the systematic
+biases exceed the statistical uncertainties. Furthermore, detection searches
+can ignore tidal effects without loss of sensitivity.
+}
+\prayush{
 We find that non-inclusion of tidal effects leads to systematic biases in
@@
 choice of ignoring tidal effects in matched-filtering templates.
-
+}
+
+\red{
+Somewhat reduce discussion of results. Discuss in order of probability, starting
+with low SNR, ending with exceptional events.
+}
+\prayush{
 Next, we study the ability of aLIGO to constrain neutron star tidal 
@@
 value. For more realistic SNRs and BH parameters, a single measurement is 
-for $\lambdans$ is accompanied with a $\sim\pm150+\%$. uncertainty. While these
-results are expected, multiple realistic observations is the next window that 
-we explore in order to obtain more information on $\lambdans$. These results 
-are detailed in Sec.~\ref{s1:PEwithNS}.
-
+for $\lambdans$ is accompanied with a $\sim\pm150+\%$. uncertainty.
+% While these
+% results are expected, multiple realistic observations is the next window that 
+% we explore in order to obtain more information on $\lambdans$. These results 
+% are detailed in Sec.~\ref{s1:PEwithNS}.
+}
+
+\red{
+BEgin this paragraph by pointing out that EOS is universal among all NSs. 
+Therefore, one universal $\lambdans$ vs $\mns$ curve. Therefore, by combining
+multiple BHNS events, expect to measure $\lambdans$ better than with any 
+single BHNS event.
+}
+\prayush{
 Multiple realistic observations deliver far more information than what a single
@@
 analysis of how our estimation of $\lambdans$ changes as we accumulate more and
-more detections. The choices that determine the population are important here,
+more detections.
+}
+\red{
+MOVE into later sections. Keep here only 1-2 informational sentences.
+}
+\prayush{
+The choices that determine the population are important here,
 and were made as follows. Neutron star mass is fixed for signals at
@@
 %
-\prayush{%
+\textbf{%
 We inject signals into zero noise at different SNRs and recover
@@
 set. With the caveats laid out, we report our results in
-Sec.~\ref{s1:multiple_observations}. In addition, we probe two paradigms, one
+Sec.~\ref{s1:multiple_observations}.
+}
+\red{
+MOVE to SUMMARY, only keeping 2-4 sentences here. Also, paradigm A/B barely
+matters. Therefore, first describe the common results, then point out how the
+presence/absence of mass gap modifies them.
+}
+\prayush{
+In addition, we probe two paradigms, one
 that allows for BH masses to lie within the astrophysical mass-gap (paradigm
@@
 time-scale.
+}
 
 Finally, we note that our results are applicable for the design sensitivity
-LIGO instruments, that are expected to come online in $2018-19$. Also note
+LIGO instruments, that are expected to come online in $2018-19$.
+\red{
+MOVE towards end of SUMMARY, into ``Future work''.
+}
+\prayush{
+Also note
 that more a recent work that improves upon the waveform model we use in this
@@
 not change our qualitative conclusions.
+}
 

### 2016-05-24 Paper/paper.tex
text abt cbcs

@@
 }
+Coalescing compact binaries (both monolithic and mixed) of stellar-mass black
+holes (BH) and neutron stars (NS) are the primary target sources for the second
+generation GW detectors~\cite{Timmes:1995kp,Fryer:1999mi,RevModPhys.74.1015,
+2010ApJ...714.1217B,2010ApJ...715L.138B,Dominik:2014yma,Belczynski:2006zi,
+2012ApJ...749...91F,
+Wex:1998wt,1991ApJ...379L..17N,Mandel:2015spa,Abbott:2016nhf}.
+A monolithic binary of black holes was recently observed by 
+aLIGO~\cite{LIGOVirgo2016a}. Previously, stellar-mass black holes had only 
+been seen in binaries with stellar companion (inferred from the observed spectrum
+of electromagnetic waves emitted by the companion)~\cite{Lewin2010,
+Remillard:2006fc,Fragos:2010tm}.
+Neutron stars, on the other hand, have had numerous sightings. Thousands have
+electromagnetically emitting neutron stars, or pulsars, have been documented,
+in varied situations: as radio pulsars~\cite{Lattimer:2012nd,Manchester:2004bp},
+in binary systems with stellar companions~\cite{1971ApJ...169L..23M,
+Bond:2002eh,Lattimer:2012nd,Manchester:2004bp},
+and in monolithic neutron star binaries~\cite{Hulse:1975uf,Taylor:1982wi,
+Weisberg:2010zz,Lattimer:2012nd,Manchester:2004bp}.
+Mixed binaries of black holes and neutron stars, is a yet undetected but 
+astrophysically interesting class of systems~\cite{Wex:1998wt,
+1991ApJ...379L..17N,Janka1999,Fryer:2015jpa},
+which advanced GW detectors are likely to observe~\cite{Abadie:2010cfa}.
+
 \prayush{
@@
   2009ApJ...697..900M,McClintock:2006xd,Miller:2009cw,McClintock:2013vwa}.
-Therefore the observations of GWs emitted by spinning BBHs will allow us to
-understand the spin-spin and spin-orbit dynamics of the two-body
-system, apart from allowing us to test strong-field dynamics of
-General Relativity.
 }

### 2016-05-25 Paper/paper.tex
text on nsbhs

--- 2016-05-24 Paper/paper.tex
@@
 
-\red{
-Text about CBCs. How was each type observed so far?
-}
-\prayush{
-Coalescing binaries of compact objects, i.e. black holes (BH) and/or
-neutron stars (NS) are the primary target sources for 
-aLIGO~\cite{Abbott:2016nhf,LIGOVirgo2016a}.
-% 
-Compact binary merger rate estimates suggest
-a GW detection rate of approximately a few tens of binary black holes
-(BBH) every year~\cite{Abadie:2010cfa}. The actual masses of
-astrophysical black holes are uncertain, but observations and
-population synthesis studies suggest that BHs formed from stellar
-core-collapse can have masses up to and higher than
-$34M_\odot$~\cite{2008ApJ...678L..17S,2010ApJ...714.1217B}.  Also,
-recent measurements using continuum fitting and X-ray reflection
-fitting suggest that black holes can have high spin, with the BH
-angular momentum in dimensionless units exceeding
-$0.8$~\cite{2011ApJ...742...85G,2012MNRAS.424..217F,Gou:2013dna,
-  2009ApJ...697..900M,McClintock:2006xd,Miller:2009cw,McClintock:2013vwa}.
-Therefore the observations of GWs emitted by spinning BBHs will allow us to
-understand the spin-spin and spin-orbit dynamics of the two-body
-system, apart from allowing us to test strong-field dynamics of
-General Relativity.
-}
-%
-
-\red{
-Text describing BHNS disruption and accretion dynamics.
-}
-An important sub-class of these
-are NSBH binaries, of which we expect $\mathcal{O}(10)$ GW observations
-every year~\cite{Abadie:2010cf}.
-NSBH binaries are of interest for multiple reasons. For instance,
-they have been long associated with (as possible progenitors of) short
-Gamma-ray Bursts (GRBs)~\cite{eichler:89,1992ApJ...395L..83N,moch:93,Barthelmy:2005bx,
-2005Natur.437..845F,2005Natur.437..851G,Shibata:2005mz,Paschalidis2014,
-Tanvir:2013}. Under favorable conditions, the NS can get disrupted by
-the tidal field of its companion BH. Once disrupted, most of the NS
-material falls into the BH over an $\mathcal{O}(1$ms$)$ time-scale,
-with the rest forming an accretion disk around the BH. This disk-BH system,
-which survives for $0.1-1$s, is hypothesized to drive SGRBs through the
-production of relativistic jets~\cite{Foucart:2015a,
-Lovelace:2013vma,Deaton2013,Foucart2012,Shibata:2005mz,Paschalidis2014}.
-A {\it coincident} detection of both gravitational and electromagnetic
-waves (gamma rays) from an NSBH system is therefore a {\it unique}
-possibility to confirm this hypothesis~\cite{Abbott:2016wya}.
-% 
+% \red{
+% Text about CBCs. How was each type observed so far?
+% }
+Coalescing compact binaries (both monolithic and mixed) of stellar-mass black
+holes (BH) and neutron stars (NS) are the primary target sources for the second
+generation GW detectors~\cite{Timmes:1995kp,Fryer:1999mi,RevModPhys.74.1015,
+2010ApJ...714.1217B,2010ApJ...715L.138B,Dominik:2014yma,Belczynski:2006zi,
+2012ApJ...749...91F,
+Wex:1998wt,1991ApJ...379L..17N,Mandel:2015spa,Abbott:2016nhf}.
+A monolithic binary of black holes was recently observed by 
+aLIGO~\cite{LIGOVirgo2016a}. Previously, stellar-mass black holes had only 
+been seen in binaries with stellar companion (inferred from the observed spectrum
+of electromagnetic waves emitted by the companion)~\cite{Lewin2010,
+Remillard:2006fc,Fragos:2010tm}.
+Neutron stars, on the other hand, have had numerous sightings. Thousands have
+electromagnetically emitting neutron stars, or pulsars, have been 
+documented~\cite{Manchester:2004bp},
+in varied situations: as radio pulsars~\cite{Lattimer:2012nd,Manchester:2004bp},
+in binary systems with stellar companions~\cite{1971ApJ...169L..23M,
+Bond:2002eh,Lattimer:2012nd,Manchester:2004bp},
+and in monolithic neutron star binaries~\cite{Hulse:1975uf,Taylor:1982wi,
+Weisberg:2010zz,Lattimer:2012nd,Manchester:2004bp}.
+Mixed binaries of black holes and neutron stars, is an astrophysically
+interesting class of systems~\cite{Wex:1998wt,
+1991ApJ...379L..17N,Janka1999,Fryer:2015jpa}, that is yet undetected.
+We expect to observe $\mathcal{O}(10)$ mixed binaries per year with
+aLIGO~\cite{Abadie:2010cf}.
+
+% \prayush{
+% The actual masses of
+% astrophysical black holes are uncertain, but observations and
+% population synthesis studies suggest that BHs formed from stellar
+% core-collapse can have masses up to and higher than
+% $34M_\odot$~\cite{2008ApJ...678L..17S,2010ApJ...714.1217B}.  Also,
+% recent measurements using continuum fitting and X-ray reflection
+% fitting suggest that black holes can have high spin, with the BH
+% angular momentum in dimensionless units exceeding
+% $0.8$~\cite{2011ApJ...742...85G,2012MNRAS.424..217F,Gou:2013dna,
+%   2009ApJ...697..900M,McClintock:2006xd,Miller:2009cw,McClintock:2013vwa}.
+% }
+% %
+
+% \red{
+% Text describing BHNS disruption and accretion dynamics.
+% }
+
 % Most of the matter falls into the BH in ~1ms. But the disk itself has an
@@
 
+NSBH binaries are of interest for multiple reasons. For instance,
+they have been long associated with (as possible progenitors of) short
+Gamma-ray Bursts (SGRBs)~\cite{eichler:89,1992ApJ...395L..83N,moch:93,Barthelmy:2005bx,
+2005Natur.437..845F,2005Natur.437..851G,Shibata:2005mz,Paschalidis2014,
+Tanvir:2013}. Depending on their equation of state (EoS), NSs can get disrupted by
+the tidal field of their companion BHs. Once disrupted, most of the NS
+material falls into the hole over an $\mathcal{O}(1$ms$)$ time-scale,
+with the rest partly getting ejected as unbound material
+% (kilonovae, r-process material),
+and partly forming an accretion disk around the BH.
+% 
+This short lived ($0.1-1s$) disk-BH system is hypothesized to drive SGRBs
+through the production of relativistic jets~\cite{Foucart:2015a,
+Lovelace:2013vma,Deaton2013,Foucart2012,Shibata:2005mz,Paschalidis2014}.
+% 
+However, whether or not such a system forms depends also on the nature of
+the BH. For e.g., massive BHs (with $\mbh\gtrsim 10M_\odot$), as well as BHs with
+large retrograde spins, tend to swallow the NS whole without forming a
+disk~\cite{Foucart:2013psa}. 
+% 
+On the other hand, {\it low-mass} BHs with $\mbh\in[3M_\odot, 7M_\odot]$, 
+can disrupt their companion NSs much before merger, forming long-sustained disks
+that are required to sustain 
+SGRBs~\cite{Shibata:2007zm,2010PhRvD..81f4026F,Lovelace:2013vma,Kawaguchi:2015}.
+% 
+% Such systems additionally leave a strong imprint on the emitted GW signal.
+A {\it coincident} detection of both GWs and gamma-rays from an NSBH merger will
+provide us with a uniquely opportunity to confirm this hypothesized link between
+NSBH mergers and GRBs~\cite{Abbott:2016wya}.
+
 \red{
@@
 }
+
 Another question that NSBH mergers can help answer is `what is the 
@@
 % 
-GWs emitted by NSBH coalescences carry subtle hints of the material structure
-of NSs. During early inspiral, the (weak) tidal field of the companion BH
+During early inspiral, the (weak) tidal field of the companion BH
 deforms the star, thereby exciting its resonant oscillation modes. The induced
-quadrupolar moment $Q_{ij}$ of the star depends on the
-tidal field $\mathcal{E}_{ij}$ of its companion through the relation
-$Q_{ij} = -\lambda\mathcal{E}_{ij}$, where $\lambda$ is an EoS-dependent
-tidal deformability parameter that is related to NS's dimensionless Love number
-$k_2$ and radius $R_\mathrm{NS}$ as $\lambda:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$.
+quadrupolar moment of the star is proportional to the tidal field of its companion,
+% through the relation $Q_{ij} = -\lambda\mathcal{E}_{ij}$,
+with $\lambdans$ the EoS-dependent proportionality constant that is also the
+tidal deformability parameter, related to the NS's dimensionless Love number
+$k_2$ and radius $R_\mathrm{NS}$ as $\lambdans:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$.
+% 
 The coherence of GW emission between stellar quadrupolar waves and orbital
@@
 
+\red{
+Insert paragraph describing impact on GWs
+}
+
+GWs emitted by NSBH coalescences carry subtle hints of the material structure
+of NSs.
+
 The difficulty in discerning NS matter effects in an NSBH coalescence from its
@@
 % 
-\red{
-Text about disruption of NSBHs, MOVE EARLIER.
-}
-\prayush{
-The mechanism of disruption
-is governed by the nature of the BH, i.e. its mass and spin. More massive BHs 
-(with $m_\mathrm{BH}\gtrsim 7\, m_\mathrm{NS}$), as well as BHs with
-(large) retrograde spins $(\chi_\mathrm{BH}\in[-1. 0])$, tend to swallow the 
-NS whole due to a rapid merger (i.e. without it disrupting before the
-inner-most stable circular orbit (ISCO) is reached). The quasi-normal
-oscillation modes of the final black hole's horizon are also 
-excited~\cite{Foucart:2013psa}. Therefore, the only available information 
-comes from the inspiral rate modification. On the other hand, less massive
-black holes spinning in the prograde sense (with mass-ratio 
-$q\in[2, 5]$, and $\chi_\mathrm{BH}\geq 0$) can cause their companion star to 
-disrupt much before merger, leaving a strong imprint on the emitted GW 
-signal~\cite{Shibata:2007zm,2010PhRvD..81f4026F,Lovelace:2013vma,
-Kawaguchi:2015}.
-}
-
-\red{
-Insert paragraph describing impact on GWs
-}
-
 A large fraction of past work aimed at measuring NS matter effects from GW

### 2016-05-25 Paper/paper.tex
text on tides

@@
 \usepackage{placeins}
+% \usepackage[switch,columnwise]{lineno} 
 \usepackage{lineno}
-\linenumbers
-
 \newcommand{\roughly}{\mathchar"5218\relax} % Different from \sim in spacing
@@
 \begin{document}
+\linenumbers
 
@@
 
-\red{
-Text focusing on tides
-}
+% \red{
+% Text focusing on tides
+% }
 
@@
 % 
-During early inspiral, the (weak) tidal field of the companion BH
-deforms the star, thereby exciting its resonant oscillation modes. The induced
-quadrupolar moment of the star is proportional to the tidal field of its companion,
-% through the relation $Q_{ij} = -\lambda\mathcal{E}_{ij}$,
-with $\lambdans$ the EoS-dependent proportionality constant that is also the
-tidal deformability parameter, related to the NS's dimensionless Love number
-$k_2$ and radius $R_\mathrm{NS}$ as $\lambdans:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$.
-% 
-The coherence of GW emission between stellar quadrupolar waves and orbital
-waves drains energy from the orbit and increases the inspiral rate~\cite{Flanagan2008}.
-We know the leading and next-to
--leading order terms in post-Newtonian (PN) theory that capture this 
-effect~\cite{Vines2011}, entering binary phasing at $5$PN order. As the NS gets
-disrupted close to binary's merger, the quadrupole moment of the system falls
-monotonically. As a result, the emitted GWs are subdued significantly
-post-disruption (as compared to what they would be for a 
-BBH)~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart:2015a,
+During the course of early inspiral, the tidal field of the BH produces a
+deformation in its companion NS. The quadrupolar moment of the star associated
+with this deformation also depends on its material properties, through an EoS-dependent
+tidal deformability parameter $\lambdans$. This induced quadrupolar moment
+changes over the orbital time-scale, resulting in the emission of GWs in {\it
+coherence} with the orbital waves.
+These waves draw energy from the orbit and increase the inspiral rate (as
+compared to an equivalent BBH)~\cite{Flanagan2008}.
+% The coherence of GW emission between stellar and orbital waves drains energy 
+% more rapidly from the orbit and increases the inspiral rate (compared to a
+% BBH)~\cite{Flanagan2008}.
+% 
+We know the leading and next-to-leading order terms in post-Newtonian (PN)
+theory that capture this effect~\cite{Vines2011}, entering binary phasing at
+$5$PN order.
+% 
+Closer to merger, the strong tidal field of the BH can disrupt the NS. The
+quadrupolar moment of the disrupted binary system falls monotonically over a
+millisecond time-scale~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart:2015a,
 Pannarale:2015jia}.
+% 
+This penultimate stage carries the strongest signature in the GW spectrum, and
+can possibly be accompanied by SGRBs. 
+% The final stage is a BH, whose quasi-normal
+% modes may (or not) be strongly excited~\cite{FoucartEtAl:2011,Lackey:2013axa}.
+
+% During early inspiral, the (weak) tidal field of the BH deforms its companion
+% NS.
+% % , thereby exciting its resonant oscillation modes.
+% The resulting induced quadrupolar moment of the star is proportional to the
+% tidal field of its companion,
+% % through the relation $Q_{ij} = -\lambda\mathcal{E}_{ij}$,
+% with $\lambdans$ the EoS-dependent proportionality constant. $\lambdans$ is the
+% NS's tidal deformability parameter, related to its dimensionless Love number
+% $k_2$ and radius $R_\mathrm{NS}$ as $\lambdans:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$.
+% % 
+% Over the course of inspiral, the change in NS's quadrupole moment results in 
+% the emission of GWs that are coherent with the orbital waves.
+% % 
+%
 

### 2016-05-25 Paper/paper.tex
text on first part

@@
 % 
-We know the leading and next-to-leading order terms in post-Newtonian (PN)
-theory that capture this effect~\cite{Vines2011}, entering binary phasing at
-$5$PN order.
+% We know the leading and next-to-leading order terms in post-Newtonian (PN)
+% theory that capture this effect~\cite{Vines2011}, entering binary phasing at
+% $5$PN order.
 % 
@@
 % 
-This penultimate stage carries the strongest signature in the GW spectrum, and
-can possibly be accompanied by SGRBs. 
+This penultimate stage also depends strongly on the internal structure and energy
+transport mechanism of the NS~\cite{}.
+
+% This penultimate stage carries the strongest signature in the GW spectrum, and
+% can possibly be accompanied by SGRBs. 
 % The final stage is a BH, whose quasi-normal
@@
 
-\red{
-Insert paragraph describing impact on GWs
-}
-
-GWs emitted by NSBH coalescences carry subtle hints of the material structure
-of NSs.
-
-The difficulty in discerning NS matter effects in an NSBH coalescence from its
-GW signal lies in distinguishing it reliably from a BBH signal, since the 
-difference between the two is either subtle (during inspiral) or short-lived
-(disruption and merger). The earlier in inspiral the NS disrupts, the more
-orbits there are over which its effect is visible.
-% 
-A large fraction of past work aimed at measuring NS matter effects from GW
-signals has consisted of inquiries about binary neutron stars (BNS)~\cite{
-Lee1999a,Lee1999b,Lee2000,oechslin:07,Read:2008iy,Markakis:2010mp,Markakis:2011vd,
-stergioulas:11,East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}. 
-Studies~\cite{DelPozzo:13,Chatziioannou:2015uea,Agathos:2015a} have shown that
-with $20-30$ BNS observations, aLIGO would begin to distinguish between hard,
-moderate and soft candidate equations of state. These studies rely on the PN 
-description of BNS dynamics, which is reasonable since BNSs merge at very high
-frequencies and most of their signal power resides in inspiral.
-% 
-On the other hand, NSBHs merge at relatively lower frequencies than BNS, 
-and their disruption during late-inspiral occurs at frequencies where aLIGO
-has relatively more resolving power. This, combined with a shorter inspiral,
-means that the merger portion of NSBH signals is relatively more valuable for
-measuring $\lambdans$. Some past NSBH studies aimed at measuring tidal effects,
-however, have used PN inspiral-only waveforms~\cite{Maselli:2013rza}. In doing
-so, however, (i) they ignore the merger signal which could contain significant
+% \red{
+% Insert paragraph describing impact on GWs
+% }
+% 
+% \red{
+% Make clear the distinction between Fischer matrix studies and Bayesian ones. 
+% Explain the differences. this is important as it sets the stage for for what is
+% new in this paper.
+% } 
+% - Describe how waves are affected in each stage for a NSBH. 
+% - Describe what past work has studied and how?
+% - Do we need to compare with BNS.. Maybe in passing, say `` In comparison, for BNS we expect to measure EoSs within 20 detections.
+
+Gravitational waves emitted by coalescing NSBH binaries carry subtle hints of
+the NS EoS from inspiral through to merger. During early inspiral, the tidal
+dephasing is relatively weak and appears as a $5^{th}$ Post-Newtonian (PN)
+order effect~\cite{Vines2011}. Closer to merger, a disruptive fate of the NS
+can result in a strong suppression of GW emission above a cut-off 
+frequency~\cite{Pannarale:2015jia}. Some past studies of tidal measurements
+with NSBH binaries have used PN inspiral-only waveforms~\cite{Maselli:2013rza}.
+In doing so, however, (i) they ignore the merger signal which could contain significant
 information for NSBHs, and (ii) the errors due to unknown vaccum terms in PN 
-waveforms could easily dominate over the tidal terms themselves, since there 
-are unknown point-particle terms at orders {\it lower} than that at which tidal
-effects enter~\cite{Yagi:2014}.
-\red{
-Make clear the distinction between Fischer matrix studies and Bayesian ones. 
-Explain the differences. this is important as it sets the stage for for what is
-new in this paper.
-} 
-Other studies that
-use complete numerical simulations or derived models either use Fisher matrix 
-to estimate $\lambdans$ measurement error~\cite{Lackey2011,Lackey:2013axa},
-which is not reliable at realistic signal-to-noise ratios 
-(SNR)~\cite{Vallisneri:2007ev}, or are limited in the binary parameter space
-they sample~\cite{Foucart:2013psa}.
+waveforms could dominate over the tidal terms themselves~\cite{Barkett2015,
+Yagi:2014}.
+% 
+Some other studies that account for merger effects via the use of complete
+numerical simulations~\cite{Foucart:2013psa}, are limited in the binary
+parameter space they sample.
+% 
+Others, that do the same through the use of phenomenological waveform
+models~\cite{Lackey2011,Lackey:2013axa} use the Fisher matrix to estimate
+$\lambdans$ measurement errors. Fisher matrix estimates are known to be
+unreliable at realistic signal-to-noise ratios (SNR)~\cite{Vallisneri:2007ev},
+such as those as we might expect in the upcoming observing runs of GW
+detectors~\cite{Abadie:2010cf}.
+
+% A large fraction of past work aimed at measuring NS matter effects from GW
+% signals has consisted of inquiries about binary neutron stars (BNS)~\cite{
+% Lee1999a,Lee1999b,Lee2000,oechslin:07,Read:2008iy,Markakis:2010mp,Markakis:2011vd,
+% stergioulas:11,East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}.
+% % 
+% % It has shown~\cite{DelPozzo:13,Chatziioannou:2015uea,Agathos:2015a} that
+% % with $20-30$ BNS observations, aLIGO would begin to distinguish between hard,
+% % moderate and soft candidate equations of state.
+% % 
+% % These studies rely on the PN 
+% % description of BNS dynamics, which is reasonable since BNSs merge at very high
+% % frequencies and most of their signal power resides in inspiral.
+% % 
+% During early inspiral the tidally induced waves are weak in strength,
+% as appropriate for an effect that enters binary phasing at $5^{th}$
+% Post-Newtonian (PN) order~\cite{Vines2011}. During late-inspiral and merger,
+% a disruptive fate of the NS could result in significant suppression of the
+% emitted GWs~\cite{}.
+% % 
+% % The difficulty in discerning NS matter effects in an NSBH coalescence from its
+% % GW signal lies in distinguishing it reliably from a BBH signal, since the 
+% % difference between the two is either subtle (during inspiral) or short-lived
+% % (disruption and merger).
+% % 
+% The earlier in inspiral the NS disrupts, the more orbits there are over which
+% its effects are manifest in the GW signal.
+% % 
+% In addition, NSBHs merge at relatively lower frequencies than BNS, and their
+% disruption during late-inspiral occurs at frequencies where aLIGO has
+% relatively more resolving power. This, combined with a shorter inspiral,
+% means that the merger portion of NSBH signals is relatively more valuable for
+% measuring $\lambdans$.
 
 In this paper we study the measurability of neutron star's tidal deformability
-from GWs emitted by {\it disruptive} NSBH mergers using aLIGO. We also probe how tidal 
-effects affect the estimation of other binary parameters for the same class
-of systems. This study improves upon previous work in many ways. First, we
-include tidal effects during inspiral and merger in a consistent way, by using
-the waveform model of Ref.~\cite{Lackey:2013axa} (henceforth ``LEA+'').
-\red{
-(describe meaning of the acronym LEA+)
-}
-Second, we include the effect
-of black hole spin on tidal GW signals, in addition to the effect of BH mass 
-and NS's deformability itself. Third, we perform a fully Bayesian 
-analysis, instead of using a Fisher matrix approximation. Fourth, we explore 
-how our measurement errors decrease as we gain information from multiple 
-(realistic) events.
+from realistic binaries of {\it low}-mass BHs and NSs by aLIGO. We also probe
+how tidal effects affect the estimation of other binary parameters for the same
+class of systems. This study improves upon previous work in the following ways.
+% 
+First, we include tidal effects during inspiral and merger in a consistent
+way, by using the waveform model of Lackey et al.~\cite{Lackey:2013axa}
+(abbreviated henceforth to ``LEA+'').
+% 
+Second, we include the effect of black hole spin on tidal GW signals, in
+addition to the effect of BH mass, NS's deformability, and the SNR.
+% 
+Third, we perform a complete Bayesian analysis, instead of using the Fisher matrix
+approximation.
+% 
+Fourth, we explore how our measurement errors decrease as we gain information
+from multiple (realistic) events.
 
@@
 efforts.
-To do so, we use the LEA+ model to generate a set of realistic signals, and
-use that as data from which we estimate binary mass and spin parameters with
-a Bayesian Monte-Carlo algorithm, using non-tidal (BBH) waveforms as filters.
-% 
-\red{
-The following numerical details disrupt the flow of the text. MOVE after
-description of analyses.
-}
-\prayush{
-As in rest of the paper, we restrict here our parameter space to span 
-mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin (aligned with orbit)
-$\chibh\in[-0.5, +0.75]$, and dimensionless neutron star tidal deformability 
-$\lambdans:= G\left(\frac{c^2}{G \mns}\right)^5\lambda \in[500, 2000]$.
-The mass of NS for signals (and not templates) is fixed to $1.35M_\odot$ 
-throughout~\cite{stellarcollapsemass}, and is taken to be 
-non-spinning~\cite{Miller:2014aaa}.
-Each signal is rescaled to a range of SNR values $\rho\in[10, 70]$.
-}
-\red{
-MOVE the following to SUMMARY. Here, keep 1-2 sentences:
-For individual events, ignoring tidal effects will affect mass-estimation
-only marginally; only for very loud signals (SNRs $\gtrsim 30$) the systematic
-biases exceed the statistical uncertainties. Furthermore, detection searches
-can ignore tidal effects without loss of sensitivity.
-}
-\prayush{
-We find that non-inclusion of tidal effects leads to systematic biases in
-the measurement of non-tidal parameters. Consider first the binary chirp 
-mass $\mchirp:=M\eta^{3/5}$ where $M$ is the binary total mass, and 
-$\eta:=\mbh\mns/M^2$ is the dimensionless mass-ratio. We find that for
-realistic (moderate) SNRs, i.e. $\rho\lesssim 30$, not including tidal
-effects introduces a barely measurable bias that remains below $50\%$
-of the statistical uncertainty of the $\mchirp$ measurement itself.
-For higher signal strengths, we find that the underlying statistical
-uncertainty decreases enough for the systematic bias to become comparable
-to it for $\rho=50-70$ and higher. This is expected because chirp mass is fairly
-accurately determined by the inspiral, where tidal effects are weak.
-Next, we consider the mass-ratio $\eta$. We find that only if the NS is quite
-deformable (with $\lambdans=2000$), {\it and} has a companion BH whose mass
-is $\mbh\lesssim 4.5M_\odot$ (i.e. in the astrophysical ``mass-gap''~\cite{
-Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}),
-will the systematic bias in $\eta$ measurement become comparable to its
-statistical uncertainties at SNRs $\rho\leq 30$. For SNRs $\rho\gtrsim30$,
-however, not including tidal effects is very likely to {\it significantly} bias
-the mass-ratio estimate for NSBH binaries. Finally, we consider the spin on the
-black hole $\chibh$. We find that $\chibh$ measurements are affected in a similar
-way to $\eta$. By not including tidal effects, we compromise our measurement of
-$\chibh$ significantly only if signal $\rho\gtrsim 30$. 
-%
-Overall, we learn an important thing. Bayesian parameter estimation algorithms
-aimed at identifying NSBH binaries with low latency in order to alert 
-electromagnetic (EM) observers in time to establish a possible coincident 
-observation of a GRB~\cite{2012A&A...541A.155A,Singer:2014qca,Singer:2015ema,
-Pankow:2015cra} can use BBH templates to look for NSBH systems with aLIGO.
-This is so because the primary requirement of the EM follow-up effort is rapid
-classification of a binary as an NSBH system, which can be achieved just as 
-easily with BBH templates, purely on the basis of the smaller component's mass.
-Another conclusion we draw is that detection searches are unaffected by the
-choice of ignoring tidal effects in matched-filtering templates.
-}
+To do so, we first use the LEA+ model to generate a set of realistic signals;
+and then use non-tidal (BBH) waveform filters to estimate the underlying
+binary masses and spins with a Markov-chain Monte Carlo.
+% 
+% \red{
+% MOVE the following to SUMMARY. Here, keep 1-2 sentences:
+We find that, for individual events, ignoring tidal effects will affect
+mass-estimation only marginally; only for very loud signals (SNRs $\gtrsim 30$)
+will the systematic biases be large enough to exceed the underlying
+statistical uncertainty. Furthermore, detection searches can ignore tidal
+effects without loss of sensitivity.
+% }
 
@@
 not change our qualitative conclusions.
+}
+
+\red{
+The following numerical details disrupt the flow of the text. MOVE after
+description of analyses.
+}
+\prayush{
+As in rest of the paper, we restrict here our parameter space to span 
+mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin (aligned with orbit)
+$\chibh\in[-0.5, +0.75]$, and dimensionless neutron star tidal deformability 
+$\lambdans:= G\left(\frac{c^2}{G \mns}\right)^5\lambda \in[500, 2000]$.
+The mass of NS for signals (and not templates) is fixed to $1.35M_\odot$ 
+throughout~\cite{stellarcollapsemass}, and is taken to be 
+non-spinning~\cite{Miller:2014aaa}.
+Each signal is rescaled to a range of SNR values $\rho\in[10, 70]$.
 }
@@
 
+\prayush{
+We find that non-inclusion of tidal effects leads to systematic biases in
+the measurement of non-tidal parameters. Consider first the binary chirp 
+mass $\mchirp:=M\eta^{3/5}$ where $M$ is the binary total mass, and 
+$\eta:=\mbh\mns/M^2$ is the dimensionless mass-ratio. We find that for
+realistic (moderate) SNRs, i.e. $\rho\lesssim 30$, not including tidal
+effects introduces a barely measurable bias that remains below $50\%$
+of the statistical uncertainty of the $\mchirp$ measurement itself.
+For higher signal strengths, we find that the underlying statistical
+uncertainty decreases enough for the systematic bias to become comparable
+to it for $\rho=50-70$ and higher. This is expected because chirp mass is fairly
+accurately determined by the inspiral, where tidal effects are weak.
+Next, we consider the mass-ratio $\eta$. We find that only if the NS is quite
+deformable (with $\lambdans=2000$), {\it and} has a companion BH whose mass
+is $\mbh\lesssim 4.5M_\odot$ (i.e. in the astrophysical ``mass-gap''~\cite{
+Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}),
+will the systematic bias in $\eta$ measurement become comparable to its
+statistical uncertainties at SNRs $\rho\leq 30$. For SNRs $\rho\gtrsim30$,
+however, not including tidal effects is very likely to {\it significantly} bias
+the mass-ratio estimate for NSBH binaries. Finally, we consider the spin on the
+black hole $\chibh$. We find that $\chibh$ measurements are affected in a similar
+way to $\eta$. By not including tidal effects, we compromise our measurement of
+$\chibh$ significantly only if signal $\rho\gtrsim 30$. 
+%
+Overall, we learn an important thing. Bayesian parameter estimation algorithms
+aimed at identifying NSBH binaries with low latency in order to alert 
+electromagnetic (EM) observers in time to establish a possible coincident 
+observation of a GRB~\cite{2012A&A...541A.155A,Singer:2014qca,Singer:2015ema,
+Pankow:2015cra} can use BBH templates to look for NSBH systems with aLIGO.
+This is so because the primary requirement of the EM follow-up effort is rapid
+classification of a binary as an NSBH system, which can be achieved just as 
+easily with BBH templates, purely on the basis of the smaller component's mass.
+Another conclusion we draw is that detection searches are unaffected by the
+choice of ignoring tidal effects in matched-filtering templates.
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-05-25 Paper/paper.tex
text on second part

@@
 
-\red{
-Somewhat reduce discussion of results. Discuss in order of probability, starting
-with low SNR, ending with exceptional events.
-}
-\prayush{
-Next, we study the ability of aLIGO to constrain neutron star tidal 
-deformability with a single observation of coalescing NSBHs. For this, we
-use the same setup and parameter choices for signal waveforms as before, but
-replace the template model with one that includes tidal effects from inspiral
-through to merger~\cite{Lackey:2013axa} (i.e. LEA+). First of all, this reduces
-the biases in the measurement of non-tidal parameters. Second, as we can see
-from Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, at signal-to-noise
-ratios $\simeq 20-30$, only under the most favorable circumstance (including a
-BH with mass within the ``mass-gap'') can we put a factor of $1.5-2$ bound on
-$\lambdans$ at $90\%$ confidence level. For similar signals with 
-$\rho\gtrsim 30$ we can constrain $\lambdans$ within $\pm 50\%$ of its true
-value. For more realistic SNRs and BH parameters, a single measurement is 
-for $\lambdans$ is accompanied with a $\sim\pm150+\%$. uncertainty.
-% While these
-% results are expected, multiple realistic observations is the next window that 
-% we explore in order to obtain more information on $\lambdans$. These results 
-% are detailed in Sec.~\ref{s1:PEwithNS}.
-}
+% \red{
+% Somewhat reduce discussion of results. Discuss in order of probability, starting
+% with low SNR, ending with exceptional events.
+% }
+
+% \prayush{
+Second, we study the ability of aLIGO to constrain neutron star tidal 
+deformability with a single observation of NSBH merger. For this, we
+use the same setup for signal waveforms as before, but replace the filter
+template model with one that includes tidal effects from inspiral
+through to merger (i.e. LEA+)~\cite{Lackey:2013axa}. For most binaries with
+BH masses outside of the mass-gap and/or realistic signal-to-noise
+ratios (SNR), we find it difficult to put better than a factor of $2$ bound
+on $\lambdans$ with a single observation. As we can see from
+Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, it is only at SNRs 
+$\rho\gtrsim 20-30$ (under otherwise favorable circumstances) that we are able
+to bring this down to a $\pm 75\%$ bound on $\lambdans$. For signals louder
+than $\rho =30$, we can constrain $\lambdans$ to a much more meaningful degree
+(within $\pm 50\%$ of its true value).
+% For more realistic SNRs and BH parameters, a single measurement is 
+% for $\lambdans$ is accompanied with a $\sim\pm150+\%$. uncertainty.
+% }
+While this is discouraging at first, we turn to ask: what if we combine
+information from a population of low-SNR observations?
 
@@
 }
-\prayush{
+% \prayush{
 Multiple realistic observations deliver far more information than what a single
@@
 more detections.
-}
+% }
+% 
 \red{

### 2016-05-25 Paper/paper.tex
intro first pass

@@
 
-\red{
-BEgin this paragraph by pointing out that EOS is universal among all NSs. 
-Therefore, one universal $\lambdans$ vs $\mns$ curve. Therefore, by combining
-multiple BHNS events, expect to measure $\lambdans$ better than with any 
-single BHNS event.
-}
+% \red{
+% BEgin this paragraph by pointing out that EOS is universal among all NSs. 
+% Therefore, one universal $\lambdans$ vs $\mns$ curve. Therefore, by combining
+% multiple BHNS events, expect to measure $\lambdans$ better than with any 
+% single BHNS event.
+% }
+% % \prayush{
+The EoS of matter at nuclear densities is believed to be universal among all
+neutron stars. The TOV equations therefore furnish a unique dependence of
+$\lambdans$ on $\mns$. This fact enables us to go further - combine
+information from multiple observations of realistic NSBH systems and
+perform a fully-Bayesian analysis of how our estimation of $\lambdans$ changes
+as we accumulate detections. This is similar to an earlier
+study~\cite{DelPozzo:13} aimed at binary neutron stars.
+% % }
+% % 
+% \red{
+% MOVE into later sections. Keep here only 1-2 informational sentences.
+% }
+% % 
+We restrict ourselves to a population of NSs with masses clustered very tightly
+around $1.35M_\odot$ (with a negligible variance), and negligible spins. We
+sample different nuclear EoSs by sampling entire populations with different
+values set for the tidal deformability of the NS.
+% These values correspond to a hard EoS at the lower end, and a soft EoS at the upper.
+Finally, we take source
+locations to be uniformly distributed in spatial volume, and source
+orientations to be uniform on a $2-$sphere. From an ensemble of such
+populations, we learn the following.
+% 
+Our best estimate for $\lambdans$ improves much faster than our statistical
+uncertainties. Within $10-20$ detections, the median measured value of
+$\lambdans$ lies within $10\%$ of its actual value.
+Measurement uncertainties for $\lambdans$, on the other hand, depend on
+$\lambdans$ itself. We find that for hard equations of state (with 
+$\lambdans\geq 1000$), $10-20$ observations are sufficient to constrain
+$\lambdans$ within $\pm 50\%$. For softer equations of state, the same level
+of certainty would require substantially more ($25-40$) observations.
+% 
+Further, if the astrophysical ``mass-gap'' is real, we find that $20-50\%$
+additional observations would be required to attain the same measurement
+accuracy as above. Putter tighter constraints on the $\lambdans$ of a
+population would require $50+$ NSBH observations, in either scenario.
+% 
+All of the above is possible within a few years of design
+aLIGO operation~\cite{Abadie:2010cfa}.
+
+% \red{
+% MOVE to SUMMARY, only keeping 2-4 sentences here. Also, paradigm A/B barely
+% matters. Therefore, first describe the common results, then point out how the
+% presence/absence of mass gap modifies them.
+% }
+% 
+% % \prayush{
+% In addition, we probe two paradigms, one
+% that allows for BH masses to lie within the astrophysical mass-gap (paradigm
+% A), and one that does not (paradigm B).
+% % 
+% For paradigm A, we find the following: (i) for the softer equations of state
+% that result in less deformable neutron stars, $15-20$ detections are sufficient
+% for our median measured value of $\lambdans$ to lie within $10\%$ of the true
+% value. (ii) For NSBH populations with more deformable NSs
+% ($\lambdans\geq 1000$), the same is achievable within as few as $10$ realistic
+% observations. (iii) The statistical uncertainty associated with $\lambdans$
+% measurement can be restricted to be within $\pm50\%$ using $10-20$ observations
+% when $\lambdans\geq 1000$), and using $25-40$ observations for softer equations
+% of state. All of the above is possible within a few years of design
+% aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
+% masses between $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
+% restricting $\lambdans$ will require $50+$ NSBH observations.
+% % 
+% For paradigm B, we find the information accumulation to be somewhat slower.
+% While the quantitative inferences for $\lambdans\geq1000$ populations are
+% not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
+% required to attain the same measurement accuracy as under paradigm A.
+% % 
+% We conclude that within as few as $20-30$ observations of disruptive NSBH
+% mergers, aLIGO will begin to place interesting bounds on NS deformability.
+% This, amongst other things, will allow us to rank different equations of 
+% state for neutron star matter from most to least likely, with a few years'
+% time-scale.
+% % }
+
+% \red{
+% The following numerical details disrupt the flow of the text. MOVE after
+% description of analyses.
+% }
+
 % \prayush{
-Multiple realistic observations deliver far more information than what a single
-observation furnishes, as has been shown the case to be for binary neutron
-stars~\cite{DelPozzo:13}. As our final study, we perform a fully-Bayesian 
-analysis of how our estimation of $\lambdans$ changes as we accumulate more and
-more detections.
-% }
-% 
-\red{
-MOVE into later sections. Keep here only 1-2 informational sentences.
-}
-\prayush{
-The choices that determine the population are important here,
-and were made as follows. Neutron star mass is fixed for signals at
-$1.35M_\odot$, and its spin $\chins=0$. Black hole mass is sampled uniformly
-from the range $[2,5]\times 1.35=[2.7, 6.75]M_\odot$. Our choice here is given
-by the intersection set of the mass range that allows for neutron star disruption
-and the range supported by the waveform model LEA+~\cite{Foucart2012,
-Foucart:2013a,Lackey:2013axa}. Black hole spin is sampled uniformly from the
-{\it aligned} range, i.e. $\chibh\in[0,+1]$. The true deformability is fixed
-once per population, sampled uniformly from the range $\lambdans\in[0,2000]$.
-In order to keep the computational cost tractable, we make a further 
-approximation.
-%
-\textbf{%
-We inject signals into zero noise at different SNRs and recover
-their parameters from Bayesian inferencing for a fixed set of binaries,
-whose parameters lie on a uniform grid given by:
-$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
-% 
-Once the population parameters been selected, each event's parameters are
-replaced by their nearest neighbor in the above set.
-}
-We make sure that this approximation is conservative by keeping our population
-sampling ranges within the conservative side of the range spanned by the above
-set. With the caveats laid out, we report our results in
-Sec.~\ref{s1:multiple_observations}.
-}
-\red{
-MOVE to SUMMARY, only keeping 2-4 sentences here. Also, paradigm A/B barely
-matters. Therefore, first describe the common results, then point out how the
-presence/absence of mass gap modifies them.
-}
-\prayush{
-In addition, we probe two paradigms, one
-that allows for BH masses to lie within the astrophysical mass-gap (paradigm
-A), and one that does not (paradigm B).
-% 
-For paradigm A, we find the following: (i) for the softer equations of state
-that result in less deformable neutron stars, $15-20$ detections are sufficient
-for our median measured value of $\lambdans$ to lie within $10\%$ of the true
-value. (ii) For NSBH populations with more deformable NSs
-($\lambdans\geq 1000$), the same is achievable within as few as $10$ realistic
-observations. (iii) The statistical uncertainty associated with $\lambdans$
-measurement can be restricted to be within $\pm50\%$ using $10-20$ observations
-when $\lambdans\geq 1000$), and using $25-40$ observations for softer equations
-of state. All of the above is possible within a few years of design
-aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
-masses between $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
-restricting $\lambdans$ will require $50+$ NSBH observations.
-% 
-For paradigm B, we find the information accumulation to be somewhat slower.
-While the quantitative inferences for $\lambdans\geq1000$ populations are
-not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
-required to attain the same measurement accuracy as under paradigm A.
-% 
-We conclude that within as few as $20-30$ observations of disruptive NSBH
-mergers, aLIGO will begin to place interesting bounds on NS deformability.
-This, amongst other things, will allow us to rank different equations of 
-state for neutron star matter from most to least likely, with a few years'
-time-scale.
-}
-
-Finally, we note that our results are applicable for the design sensitivity
-LIGO instruments, that are expected to come online in $2018-19$.
-\red{
-MOVE towards end of SUMMARY, into ``Future work''.
-}
-\prayush{
-Also note
-that more a recent work that improves upon the waveform model we use in this
-study has been published~\cite{Pannarale:2015jka}, but it provides only an 
-amplitude model which needs to be (in future) augmented with a compatible
-phase model. In addition, accuracy of the underlying numerical simulations 
-used to calibrate the waveform model used here have not been verified against 
-independent codes so far.
-It is therefore difficult to assess the combined modeling error and its effect
-on our results. they are, therefore, limited by the limitations of our
-waveform model. However, we expect the combined effect of modeling errors to
-not change our qualitative conclusions.
-}
-
-\red{
-The following numerical details disrupt the flow of the text. MOVE after
-description of analyses.
-}
-\prayush{
-As in rest of the paper, we restrict here our parameter space to span 
+In rest of the paper, we restrict our parameter space to span 
 mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin (aligned with orbit)
@@
 Each signal is rescaled to a range of SNR values $\rho\in[10, 70]$.
-}
+% }
+% 
+Finally, we note that our results are applicable for the design sensitivity
+LIGO instruments, that are expected to come online in $2018-19$.
 
@@
 
+\prayush{
+The choices that determine the population are important here,
+and were made as follows. Neutron star mass is fixed for signals at
+$1.35M_\odot$, and its spin $\chins=0$. Black hole mass is sampled uniformly
+from the range $[2,5]\times 1.35=[2.7, 6.75]M_\odot$. Our choice here is given
+by the intersection set of the mass range that allows for neutron star disruption
+and the range supported by the waveform model LEA+~\cite{Foucart2012,
+Foucart:2013a,Lackey:2013axa}. Black hole spin is sampled uniformly from the
+{\it aligned} range, i.e. $\chibh\in[0,+1]$. The true deformability is fixed
+once per population, sampled uniformly from the range $\lambdans\in[0,2000]$.
+In order to keep the computational cost tractable, we make a further 
+approximation.
+%
+\textbf{%
+We inject signals into zero noise at different SNRs and recover
+their parameters from Bayesian inferencing for a fixed set of binaries,
+whose parameters lie on a uniform grid given by:
+$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+% 
+Once the population parameters been selected, each event's parameters are
+replaced by their nearest neighbor in the above set.
+}
+We make sure that this approximation is conservative by keeping our population
+sampling ranges within the conservative side of the range spanned by the above
+set. With the caveats laid out, we report our results in
+Sec.~\ref{s1:multiple_observations}.
+}
+
+\prayush{
+In addition, we probe two paradigms, one
+that allows for BH masses to lie within the astrophysical mass-gap (paradigm
+A), and one that does not (paradigm B).
+% 
+For paradigm A, we find the following: (i) for the softer equations of state
+that result in less deformable neutron stars, $15-20$ detections are sufficient
+for our median measured value of $\lambdans$ to lie within $10\%$ of the true
+value. (ii) For NSBH populations with more deformable NSs
+($\lambdans\geq 1000$), the same is achievable within as few as $10$ realistic
+observations. (iii) The statistical uncertainty associated with $\lambdans$
+measurement can be restricted to be within $\pm50\%$ using $10-20$ observations
+when $\lambdans\geq 1000$), and using $25-40$ observations for softer equations
+of state. All of the above is possible within a few years of design
+aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
+masses between $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
+restricting $\lambdans$ will require $50+$ NSBH observations.
+% 
+For paradigm B, we find the information accumulation to be somewhat slower.
+While the quantitative inferences for $\lambdans\geq1000$ populations are
+not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
+required to attain the same measurement accuracy as under paradigm A.
+% 
+We conclude that within as few as $20-30$ observations of disruptive NSBH
+mergers, aLIGO will begin to place interesting bounds on NS deformability.
+This, amongst other things, will allow us to rank different equations of 
+state for neutron star matter from most to least likely, with a few years'
+time-scale.
+}
+
+\red{
+MOVE towards end of SUMMARY, into ``Future work''.
+}
+\prayush{
+Also note
+that more a recent work that improves upon the waveform model we use in this
+study has been published~\cite{Pannarale:2015jka}, but it provides only an 
+amplitude model which needs to be (in future) augmented with a compatible
+phase model. In addition, accuracy of the underlying numerical simulations 
+used to calibrate the waveform model used here have not been verified against 
+independent codes so far.
+It is therefore difficult to assess the combined modeling error and its effect
+on our results. they are, therefore, limited by the limitations of our
+waveform model. However, we expect the combined effect of modeling errors to
+not change our qualitative conclusions.
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-05-25 Paper/paper.tex
tov text

@@
 The EoS of matter at nuclear densities is believed to be universal among all
-neutron stars. The TOV equations therefore furnish a unique dependence of
-$\lambdans$ on $\mns$. This fact enables us to go further - combine
-information from multiple observations of realistic NSBH systems and
-perform a fully-Bayesian analysis of how our estimation of $\lambdans$ changes
-as we accumulate detections. This is similar to an earlier
+neutron stars. The Tolman-Oppenheimer-Volkoff equation would then predict that all
+NSs lie on a universal $\lambdans$ versus $\mns$ curve. As the final part of
+this paper, we combine information from multiple observations of realistic NSBH
+systems and perform a fully-Bayesian analysis of how our estimation of
+$\lambdans$ changes as we accumulate detections. This is similar to an earlier
 study~\cite{DelPozzo:13} aimed at binary neutron stars.
@@
 around $1.35M_\odot$ (with a negligible variance), and negligible spins. We
-sample different nuclear EoSs by sampling entire populations with different
-values set for the tidal deformability of the NS.
+sample different nuclear EoSs by sampling entire populations fixing different
+values for the NS tidal deformability.
 % These values correspond to a hard EoS at the lower end, and a soft EoS at the upper.
-Finally, we take source
-locations to be uniformly distributed in spatial volume, and source
-orientations to be uniform on a $2-$sphere. From an ensemble of such
-populations, we learn the following.
-% 
-Our best estimate for $\lambdans$ improves much faster than our statistical
+For all populations, we take source locations to be uniformly distributed in
+spatial volume, and source orientations to be uniform on a $2-$sphere. To
+summarize, we find the following:
+% 
+(a) Our best estimate for $\lambdans$ improves much faster than our statistical
 uncertainties. Within $10-20$ detections, the median measured value of
-$\lambdans$ lies within $10\%$ of its actual value.
-Measurement uncertainties for $\lambdans$, on the other hand, depend on
+$\lambdans$ lies within $10\%$ of the actual value.
+(b) Measurement uncertainties for $\lambdans$, on the other hand, depend on
 $\lambdans$ itself. We find that for hard equations of state (with 
@@
 % 
-Further, if the astrophysical ``mass-gap'' is real, we find that $20-50\%$
+(c) Further, if the astrophysical ``mass-gap'' is real, we find that $20-50\%$
 additional observations would be required to attain the same measurement
-accuracy as above. Putter tighter constraints on the $\lambdans$ of a
-population would require $50+$ NSBH observations, in either scenario.
+accuracy as above. And, (d) putting tighter constraints on the $\lambdans$ of a
+population would require $50+$ NSBH observations, in any scenario.
 % 
@@
 LIGO instruments, that are expected to come online in $2018-19$.
+\prayush{INSERT TEXT ON LIMITATIONS}
 

### 2016-05-25 Paper/paper.tex
intro better

@@
 Coalescing compact binaries (both monolithic and mixed) of stellar-mass black
-holes (BH) and neutron stars (NS) are the primary target sources for the second
+holes (BH) and neutron stars (NS) are the primary targets for the second
 generation GW detectors~\cite{Timmes:1995kp,Fryer:1999mi,RevModPhys.74.1015,
@@
 aLIGO~\cite{LIGOVirgo2016a}. Previously, stellar-mass black holes had only 
-been seen in binaries with stellar companion (inferred from the observed spectrum
-of electromagnetic waves emitted by the companion)~\cite{Lewin2010,
+been observed by inference in mixed binaries with stellar companion (through
+electromagnetic observations of the companion)~\cite{Lewin2010,
 Remillard:2006fc,Fragos:2010tm}.
-Neutron stars, on the other hand, have had numerous sightings. Thousands have
+Neutron stars, on the other hand, have had numerous sightings. Thousands of
 electromagnetically emitting neutron stars, or pulsars, have been 
@@
 % Such systems additionally leave a strong imprint on the emitted GW signal.
-A {\it coincident} detection of both GWs and gamma-rays from an NSBH merger will
+A {\it coincident} detection of both GWs and gamma-rays from an NSBH merger,
+whenever it happens, will
 provide us with a uniquely opportunity to confirm this hypothesized link between
@@
 Another question that NSBH mergers can help answer is `what is the 
-nature of matter at nuclear densities', i.e. what is the equation of state
-(EoS) of neutron star material. This is what we will focus on in this paper.
+nature of matter at nuclear densities', i.e. `what is the EoS of NS material'?
+This is what we will focus on in this paper.
 % 
@@
 coherence} with the orbital waves.
-These waves draw energy from the orbit and increase the inspiral rate (as
+These waves draw more energy from the orbit and increase the inspiral rate (as
 compared to an equivalent BBH)~\cite{Flanagan2008}.
@@
 The EoS of matter at nuclear densities is believed to be universal among all
-neutron stars. The Tolman-Oppenheimer-Volkoff equation would then predict that all
-NSs lie on a universal $\lambdans$ versus $\mns$ curve. As the final part of
+neutron stars. The Tolman-Oppenheimer-Volkoff equation would then predict that NS
+properties satisfy a universal relationship between $\lambdans$ and $\mns$. As
+the final part of
 this paper, we combine information from multiple observations of realistic NSBH
@@
 % \prayush{
-In rest of the paper, we restrict our parameter space to span 
-mass-ratios $q:=\mbh/\mns\in[2,5]$, black hole spin (aligned with orbit)
-$\chibh\in[-0.5, +0.75]$, and dimensionless neutron star tidal deformability 
+In this paper, we restrict our parameter space to span mass-ratios
+$q:=\mbh/\mns\in[2,5]$, BH spin (aligned with orbit)
+$\chibh\in[-0.5, +0.75]$, and dimensionless NS tidal deformability 
 $\lambdans:= G\left(\frac{c^2}{G \mns}\right)^5\lambda \in[500, 2000]$.
 The mass of NS for signals (and not templates) is fixed to $1.35M_\odot$ 
-throughout~\cite{stellarcollapsemass}, and is taken to be 
-non-spinning~\cite{Miller:2014aaa}.
-Each signal is rescaled to a range of SNR values $\rho\in[10, 70]$.
+throughout~\cite{stellarcollapsemass}. Similarly, their spin is taken to be
+negligible~\cite{Miller:2014aaa}. Signal strength is sampled over a range of
+SNR $\rho\in[10, 70]$.
 % }
 % 
-Finally, we note that our results are applicable for the design sensitivity
-LIGO instruments, that are expected to come online in $2018-19$.
-\prayush{INSERT TEXT ON LIMITATIONS}
+Further, we note that our results are applicable for the design sensitivity
+LIGO instruments ($2018-19$).
+We also note that the parameter space probed is somewhat restricted by the
+domain of applicability of the LEA+ model which we use as filters.
+Finally, the accuracy of our quantitative results depends on the reliability of
+the same waveform model, which is the only one available of its kind in 
+current literature. A more recent work~\cite{Pannarale:2015jka}, improves on
+the GW amplitude model, which however needs to be augmented with a compatible
+phase model. 
+% In addition, accuracy of the underlying numerical simulations 
+% used to calibrate the waveform model used here have not been verified against 
+% independent codes so far.
+% It is therefore difficult to assess the combined modeling error and its effect
+% on our results. they are, therefore, limited by the limitations of our
+% waveform model. 
+However, we expect the combined effect of modeling errors to not change our 
+broad qualitative conclusions, which are based on vague statistical trends 
+with large errors themselves.
 

### 2016-05-26 Paper/paper.tex
text in s2a

@@
 
-In this section we summarize various technical aspects of this paper. The model
-used to obtain tidal waveforms is described in Sec.~\ref{s2:waveforms}. The 
-process of inferring source parameters from a GW signal is summarized in 
-Sec.~\ref{s2:bayesian}. The process of generating and combining multiple events
-is instead detailed in the corresponding results section,
-Sec.~\ref{s2:astro_multiple}.
+% In this section we summarize various technical aspects of this paper. The model
+% used to obtain tidal waveforms is described in Sec.~\ref{s2:waveforms}. The 
+% process of inferring source parameters from a GW signal is summarized in 
+% Sec.~\ref{s2:bayesian}. The process of generating and combining multiple events
+% is instead detailed in the corresponding results section,
+% Sec.~\ref{s2:astro_multiple}.
 
@@
 
-In Ref.~\cite{Lackey:2013axa}, the authors presented a suite of $134$ numerical
-relativity simulations of neutron stars inspiraling into spinning black holes. Amongst
-the varied quantities were (a) the equation of state for the NS, (b) NS mass, 
-(c) mass-ratio and (d) black hole spins, taken as aligned with the orbit.
-A total of $21$ two-parameter EoSs were used for these simulations, which also 
-had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
-$m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
-authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
-NSBH coalescences. The GW strain predicted by this model (henceforth ``LEA''),
-can be written as
-% 
+% \red{
+% Begin with introducing sentence:
+% }
+
+Lackey et al.~(LEA)~\cite{Lackey:2013axa} developed a complete inspiral-merger
+waveform model for disrupting NSBHs. Theirs is a frequency-domain
+phenomenological model that includes the effect of BH and NS masses and spins
+$\{\mbh, \chibh, \mns\}\equiv\vec{\theta}$ and NS tidal deformability
+$\lambdans$. It was calibrated to a suite of $134$ numerical relativity (NR)
+simulations of NSs inspiraling into spinning BHs. The parameter space these
+simulations span, which we assume to be the domain of validity for the LEA
+model, includes NS masses $1.2M_\odot\leq\mns\leq 1.45M_\odot$,
+mass-ratios $2\leq q\leq 5$, and BH spins $-0.5\leq\chibh\leq+0.75$.
+They also sample a total of $21$ two-parameter nuclear EoSs to cover the
+spectrum of NS deformability.
+% 
+% Amongst the varied quantities were (a) the
+% equation of state for the NS, (b) NS mass, (c) mass-ratio and (d) black hole
+% spins, taken as aligned with the orbit.
+% A total of $21$ two-parameter EoSs were used for these simulations, which also 
+% had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
+% $m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
+% authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
+% NSBH coalescences.
+% 
+The GW strain $\tilde{h}(f)$ per the LEA model can be written as
 \begin{equation}
@@
 \end{equation}
-where $\vec{\theta}:=\{\mbh, \mns, \chibh,\chins=0\}$ are the intrinsic binary
-parameters. The first term on the right hand side is an underlying
-BBH waveform model that is used as a base. In the original construction of LEA,
-this was taken to be the SEOBNRv1 model~\cite{Taracchini:2012} of the 
-Effective-one-body (EOB) family~\cite{Buonanno99}. The second term changes the 
-amplitude of the BBH model
-to match that of an NSBH merger of otherwise identical parameters, with NS-matter
-effects parametrized by $\lambdans$. During inspiral this term is set to
-unity, but is a sensitive function of $\lambdans$ close to merger. The third 
-term does the same as the second, but to the waveform phasing. During inspiral,
-$\Delta\Phi(f, \vec{\theta}, \lambdans)$ is set to the PN tidal phasing corrections,
+with NS spin $\chins=0$ identically. Here, $\tilde{h}_\mathrm{BBH}$ is
+an underlying BBH waveform model. In the original LEA model,
+this was taken to be the SEOBNRv1
+model~\cite{Taracchini:2012} of the Effective-one-body (EOB)
+family~\cite{Buonanno99}. The factor $A(\cdot)$ adjusts
+the amplitude of the BBH model to match that of an NSBH merger of otherwise
+identical parameters, with NS-matter effects parametrized by $\lambdans$.
+During early inspiral this term is set to
+unity, but is a sensitive function of $\lambdans$ close to merger. The term with
+$\Delta\Phi$ corrects the waveform phasing. During inspiral,
+$\Delta\Phi$ is set to the PN tidal phasing corrections,
 at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
-additional phenomenological terms are added and calibrated to the $134$ available
-NR simulations.
+additional phenomenological terms are needed. Both $A$ and $\Delta\Phi$ are
+calibrated to all $134$ available NR simulations.
 
 In this paper we use LEA for our signal and template modeling, but switch the 
-underlying BBH model to SEOBNRv2~\cite{Taracchini:2013rva},
-using its reduced-order frequency-domain version~\cite{Purrer:2015tud}. We
-expect this change to make our conclusions more robust because: (a) the 
+underlying BBH model to SEOBNRv2 (and refer to it as enhanced-LEA or
+``LEA+'')~\cite{Taracchini:2013rva}. We using the reduced-order
+frequency-domain version of SEOBNRv2, which has the additional benefit of
+reducing computational cost~\cite{Purrer:2015tud}. We expect this change
+(LEA$\rightarrow$LEA+) to make our conclusions more robust because: (a) the 
 SEOBNRv2 model is more accurate~\cite{Kumar:2015tha,Kumar:2016dhh}, and (b)
@@
 merger~\cite{Kumar:2015tha}.
-Since LEA only augments the inspiral phasing with PN tidal terms, our
-change in the underlying BBH model does not change LEA's construction, but
-increases the overall model accuracy during inspiral. Therefore the model used 
-here (``LEA+'') is a more robust version of LEA.
+Since LEA only augments inspiral phasing with PN tidal terms, our
+change in the underlying BBH model does not change LEA's construction, {\it and}
+increases the overall model accuracy during inspiral.
+% We will refer to this enhanced-LEA model henceforth as ``LEA+''.
 

### 2016-05-26 Paper/paper.tex
starting s2b

@@
 frequency-domain version of SEOBNRv2, which has the additional benefit of
-reducing computational cost~\cite{Purrer:2015tud}. We expect this change
-(LEA$\rightarrow$LEA+) to make our conclusions more robust because: (a) the 
+reducing computational cost~\cite{Purrer:2015tud}. We expect this enhancement
+from LEA$\rightarrow$LEA+ to make our conclusions more robust because: (a) the 
 SEOBNRv2 model is more accurate~\cite{Kumar:2015tha,Kumar:2016dhh}, and (b)
@@
 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
+% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
 \subsection{Bayesian methods}\label{s2:bayesian}
 % % Describe the emcee-based PE code
+\red{
+Section is immensely technical. Perform careful polishing pass to 
+improve flow and to make it more accessible.
+}
 
@@
 where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
-and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
-as
+% and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
+% as
+and $\langle\cdot|\cdot\rangle$ is the detector noise weighted 
+inner-product~\footnote{%
+This inner product is defined as
 \begin{equation}
@@
 where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
-and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. The 
+and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}
+The 
 definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument

### 2016-05-26 Paper/paper.tex
s2b cleaned

@@
 % % Describe the emcee-based PE code
-\red{
-Section is immensely technical. Perform careful polishing pass to 
-improve flow and to make it more accessible.
-}
-
-We perform Bayesian parameter estimation for tidal signals, using
-both tidal and non-tidal templates. The signal parameters are selected 
-to lie on a grid given by:
-$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
-Corresponding signal waveforms are generated using the model described in the 
-previous sub-section, and each injected into a separate zero noise data segment
-$d_n$. Templates can be either tidal or non-tidal (LEA+ with $\lambdans=0$).
-
-Inferred from $d_n$, the joint probability distribution for binary parameters
-$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ is given by
+% \red{
+% Section is immensely technical. Perform careful polishing pass to 
+% improve flow and to make it more accessible.
+% }
+
+The process of measuring systematic and statistical measurement errors
+involves simulating many artificial GW signals, and inferring source binary
+parameters from them.
+We start with generating a signal waveform, using the model LEA+, and injecting
+it in zero noise to obtain a stretch of data $d_n$. Source binary parameters
+$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ are reconstructed from this
+injected signal. Using Bayes' theorem, their joint inferred probability
+distribution can be calculated as
 \begin{equation}\label{eq:postprob}
@@
 \end{equation}
-Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of
-taking particular values. In this paper, we use a uniform prior on the
-individual masses within pre-chosen ranges, as well as a uniform prior on the
-spin of the black hole. We set neutron star's spin $=0$, and its tidal
-deformability parameter $\lambdans$ is sampled uniformly from $[0, 4000]$. In
-addition, we exclude mass-ratios between $[1,2]$ from our prior. This is done
-because of model constraints for LEA+. The prior probability of obtaining a 
-particular realization of data $p(d_n|K)$ is absorbed into the overall 
-normalization. Finally, the first term in the numerator
-$p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given
-stretch of data $d_n$ if we assume that a signal parameterized by
-$\vec{\theta}\cup\{\lambdans\}$ is buried in it, i.e.
+Here, $p(\vec{\Theta} | K)$ is the a-priori probability of binary parameters
+taking particular values. Throughout this paper, we impose a prior uniform on
+individual component masses, BH spin, and NS's tidal deformability. In
+addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not valid for 
+$1\leq q\leq 2$. $p(d_n|\vec{\Theta}, K)$ is the
+likelihood of obtaining the given stretch of data $d_n$ if we assume that a
+signal parameterized by $\vec{\Theta}$ is buried in it, and is given by
 \begin{equation}\label{eq:likelihood}
- \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+ \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N}\, \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
 \end{equation}
-where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
-% and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
-% as
-and $\langle\cdot|\cdot\rangle$ is the detector noise weighted 
-inner-product~\footnote{%
-This inner product is defined as
+where $h\equiv h(\vec{\Theta})$ is a filter template with parameters 
+$\vec{\Theta}$, and $\langle\cdot|\cdot\rangle$ is a suitably defined
+detector-noise weighted inner-product~\footnote{The inner product
+$\langle\cdot|\cdot\rangle$  is defined as
 \begin{equation}
@@
 \end{equation}
-% 
 where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
 and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}
-The 
-definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
-noise is colored Gaussian. The marginalized probability distribution for any
-single binary parameter (say $\alpha$) can be obtained by integrating
-Eq.~\ref{eq:postprob} over all other parameters, i.e. 
-\begin{equation}
+The denominator in Eq.~\ref{eq:postprob} is the a-priori probability of finding
+the particular signal in $d_n$ and we assume that each injected signal is as
+likely as any other. Having constructed the probability distribution function
+$p(\vec{\Theta} | d_n, K)$, extracting the measured probability distribution
+for a single parameter (say $\alpha$) involves integrating
+\begin{equation}\label{eq:marginalize}
  p(\alpha | d_n, K) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, K),
@@
 
-We sample the probability distribution of Eq.~\ref{eq:postprob} using
-an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}
-package~. A total of $100$ independent walkers were used, each of which
-was allowed to collect $100,000$ samples. We measure the correlation time and keep
-only $20,000$ independent posterior points to further analyze the calculated 
-posterior probability distributions. One simplifying assumption made to mitigate 
+We use an ensemble sampler Markov-chain Monte-Carlo algorithm based on the
+{\tt emcee} package~\cite{emcee}, to sample the probability distribution 
+$p(\vec{\Theta} | d_n, K)$. We start with $100$ independent walkers, each
+of which is allowed to collect $100,000$ samples.
+% 
+% We perform Bayesian parameter estimation for tidal signals, using
+% both tidal and non-tidal templates. The signal parameters are selected 
+% to lie on a grid given by:
+% $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+% Corresponding signal waveforms are generated using the model described in the 
+% previous sub-section, and each injected into a separate zero noise data segment
+% $d_n$. Templates can be either tidal or non-tidal (LEA+ with $\lambdans=0$).
+% 
+% Inferred from $d_n$, the joint probability distribution for binary parameters
+% $\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ is given by
+% \begin{equation}\label{eq:postprob}
+%  p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
+% \end{equation}
+% Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of
+% taking particular values. In this paper, we use a uniform prior on the
+% individual masses within pre-chosen ranges, as well as a uniform prior on the
+% spin of the black hole. We set neutron star's spin $=0$, and its tidal
+% deformability parameter $\lambdans$ is sampled uniformly from $[0, 4000]$. In
+% addition, we exclude mass-ratios between $[1,2]$ from our prior. This is done
+% because of model constraints for LEA+. The prior probability of obtaining a 
+% particular realization of data $p(d_n|K)$ is absorbed into the overall 
+% normalization. Finally, the first term in the numerator
+% $p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given
+% stretch of data $d_n$ if we assume that a signal parameterized by
+% $\vec{\theta}\cup\{\lambdans\}$ is buried in it, i.e.
+% \begin{equation}\label{eq:likelihood}
+%  \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+% \end{equation}
+% where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
+% % and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
+% % as
+% and $\langle\cdot|\cdot\rangle$ is the detector noise weighted 
+% inner-product~\footnote{%
+% This inner product is defined as
+% \begin{equation}
+% \langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
+% \end{equation}
+% % 
+% where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
+% and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}
+% The 
+% definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
+% noise is colored Gaussian. The marginalized probability distribution for any
+% single binary parameter (say $\alpha$) can be obtained by integrating
+% Eq.~\ref{eq:postprob} over all other parameters, i.e. 
+% \begin{equation}
+%  p(\alpha | d_n, K) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, K),
+% \end{equation}
+% where $\vec{\Theta}_\alpha$ is the set of remaining parameters, i.e.
+% $\vec{\Theta}_\alpha:=\vec{\Theta} - \{\alpha\}$.
+% 
+% 
+% 
+% We sample the probability distribution of Eq.~\ref{eq:postprob} using
+% an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}
+% package~. A total of $100$ independent walkers were used, each of which
+% was allowed to collect $100,000$ samples. We measure the correlation time and keep
+% only $20,000$ independent posterior points to further analyze the calculated 
+% posterior probability distributions. 
+% 
+One simplifying assumption made to mitigate 
 computational cost was setting the frequency sampling rate $\Delta f=0.4$~Hz, which
 we found to be sufficient to calculate likelihoods robustly~\cite{Purrer:2015nkh}.
-\prayush{TODO: MICHAEL}.
+% \prayush{TOCheck: MICHAEL}.
+Once more than $10,000$ independent samples have been collected, we integrate
+Eq.~\ref{eq:marginalize} to obtain the marginalized probability distribution
+for the tidal deformability of the NS $p(\lambdans|d_n,K)$.
+The most likely value of $\lambdans$ for this distribution is our 
+{\it measured} value for the parameter, and the $90\%$ confidence intervals
+associated with the distribution serve as statistical error-bars.
+% 
+% We use this
+% distribution to put confidence bounds on our measurement of $\lambdans$.
+% 
+% of this distribution to estimate our statistical and
 

### 2016-06-08 Paper/paper.tex
adding figures for sec 3

@@
 involves simulating many artificial GW signals, and inferring source binary
-parameters from them.
+parameters from them using Bayesian statistics.
 We start with generating a signal waveform, using the model LEA+, and injecting
@@
 injected signal. Using Bayes' theorem, their joint inferred probability
-distribution can be calculated as
+distribution of $\vec{\Theta}$ can be written as
 \begin{equation}\label{eq:postprob}
@@
 % 
-One simplifying assumption made to mitigate 
-computational cost was setting the frequency sampling rate $\Delta f=0.4$~Hz, which
-we found to be sufficient to calculate likelihoods robustly~\cite{Purrer:2015nkh}.
-% \prayush{TOCheck: MICHAEL}.
+One simplification we make to mitigate computational cost is to set the
+frequency sampling interval to $\Delta f=0.4$~Hz, which we find to be
+sufficient for robust likelihoods calculations~\cite{Purrer:2015nkh}.
 Once more than $10,000$ independent samples have been collected, we integrate
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How is PE affected by not including NS matter effects?}\label{s1:PEwithnoNS}
+\section{How is PE affected if we ignore NS matter effects?}\label{s1:PEwithnoNS}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 study of the effects of this assumption on the recovery of other physical parameters
-of NS-BH sources. We do so by comparing the systematic biases introduced by this 
-assumption to the overall statistical uncertainties in the measurement of binary 
-mass and spin parameters. We find that unless the single-detector SNR of the signal
-exceeds $30$ (for design Advanced LIGO sensitivity), the bias accurued by ignoring
-tidal effects in filtering templates will remain smaller than the statistical 
-uncertainty of the same measurement.
-
+of NS-BH sources.
+% We do so by comparing the systematic biases introduced by this 
+% assumption to the overall statistical uncertainties in the measurement of binary 
+% mass and spin parameters. We find that unless the single-detector SNR of the signal
+% exceeds $30$ (for design Advanced LIGO sensitivity), the bias accurued by ignoring
+% tidal effects in filtering templates will remain smaller than the statistical 
+% uncertainty of the same measurement.
+
+\red{Reorder:
+We construct NSBH waveforms with LEA+. We inject them, and run MCMC on the injections
+using BBH templates. We perform this analysis for [... parameters ...]. We always
+use $\mns=1.35M_\odot$, but explore $\lambdans=\{500,1000,1500,2000\}$.
+}
+
+\prayush{
 We first take a set of NSBH binary waveforms, which include the modeling of spin and
-tidal effects. Binary mass-ratio is sampled
+tidal effects.
+For all of the resulting signals, we perform a fully Bayesian parameter estimation
+aimed at measuring binary masses and spins from the signal. Our analysis
+templates differ from the signal in only one sense: all tidal terms are set to zero
+in the former, i.e. they correspond to BHBH mergers. Binary mass-ratio is sampled
 uniformly over the range $q\in[2,5]$ taking $q=\{2,3,4,5\}$ as our test points. The
@@
 $\lambdans=800$ case. Waveform are generated for all possible combinations of these
-parameters, and rescaled to different strengths with $\rho=\{20, 30, 50, 70\}$. For all
-of the resulting signals, we perform a fully Bayesian parameter estimation aimed at
-measuring binary masses and spins from the signal. Our analysis
-templates differ from the signal in only one sense: all tidal terms are set to zero
-in the former, i.e. they correspond to BHBH mergers.
+parameters, and rescaled to different strengths with $\rho=\{20, 30, 50, 70\}$.
+}
 

### 2016-06-08 emcee_PE/scripts/AnalyzeSamples.py

@@ -657,9 +657,10 @@ def get_results(data, q=None, chiBH=None, NSLmbd=None, SNR=None,\
     raise IOError("Need all of (q,chi,NSL,SNR) parameters")
   l1grp, l2grp, l3grp = \
         ['q%.1f.dir' % q, 'chiBH%.2f.dir' % chiBH, 'LambdaNS%.1f.dir' % NSLmbd]
-  dset = 'SNR%.1f.dat' % SNR
+  l4grp = 'SNR%.1f.dir' % SNR
+  dset  = 'summary.dat'
   if vverbose: print "Group names = ", l1grp, l2grp, l3grp, dset
-  alldata = data[l1grp][l2grp][l3grp][dset].value
+  alldata = data[l1grp][l2grp][l3grp][l4grp][dset].value
   
   num_of_data_fields = 5
   if 'm1' in p: pidx = 2 + 0*num_of_data_fields
@@ -953,12 +954,11 @@ etavec = qvec / (1. + qvec)**2
 mtotalvec = mNS + mNS * qvec
 mchirpvec = mtotalvec * etavec**0.6
 Lambdavec = np.array([0])#[500, 800, 1000])
-SNRvec = np.array([20, 30, 50, 70, 90, 120])
+SNRvec = np.array([20, 30, 50, 70])
 if inject_tidal: Lambdavec = np.array([500, 800, 1000, 1500, 2000])
 #if inject_tidal: Lambdavec = np.array([500, 800, 1000])
 
 ######################################################
-# TEMPLATE PARAMETERS
 ######################################################
 # Taking an uninformed prior
 m1min = 1.2
@@ -1011,7 +1011,7 @@ plot_ChiCriticalCurves    = False
 plot_QCriticalCurves      = False
 
 # Reduced ranges for plots
-plotSNRvec    = np.array([20, 30, 50, 70, 90, 120])
+plotSNRvec    = np.array([20, 30, 50, 70])
 plotqvec      = qvec
 plotchi2vec   = np.array([0, 0.5, 0.74999])
 plotLambdavec = np.array([800, 1000, 1500])
@@ -1023,21 +1023,34 @@ error_p         = error_threshold * 100.
 ######################################################
 # READ DATA
 ######################################################
-datadir  = '/home/prayush/research/NSBH/TidalParameterEstimation/FinalCombinedPEData/'
[…]

### 2016-06-08 emcee_PE/scripts/SingleObservation.py

@@ -98,7 +98,7 @@ col corresponds to the lower level group
   """
   if colorbartype != 'simple':
     raise IOError("Colorbar type %s not supported" % colorbartype)
-  colwidth = 1.7
+  #
   if np.shape(X)[:2] != np.shape(Y)[:2] or np.shape(X)[:2] != np.shape(Z2d)[:2]:
     raise IOError("X, Y and Z arrays have different number of sets to plot")
   
@@ -112,15 +112,17 @@ col corresponds to the lower level group
     'ytick.labelsize':16,\
     'figure.subplot.bottom':0.2,\
     'figure.figsize':figsize, \
-    'savefig.dpi': 300.0, \
-    'figure.autolayout': True})
+    'savefig.dpi': 300.0})#, \
+  #'figure.autolayout': True})
 
   nrow, ncol, _ = np.shape(X)
   if vverbose: print "Making plot with %d rows, %d cols" % (nrow, ncol)
   pltid = 0
   
+  colwidth = 1.3
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((1.5*gmean*ncol+1.25)*colwidth, 1.2*colwidth*nrow))
+              figsize=((1.3*gmean*ncol+1.25)*colwidth, 1.65*colwidth*nrow),\
+              dpi=100)
   fig.clf()
   grid = ImageGrid(fig, 111, nrows_ncols=(nrow, ncol), \
             share_all=True,\
@@ -187,20 +189,22 @@ col corresponds to the lower level group
       if jdx == 0: ax.set_ylabel(ylabel)
       #if np.shape(titles) == (nrow, ncol): ax.set_title(titles[idx][jdx])
       if np.shape(titles) == (nrow, ncol):
-        ax.text(.5, .9, titles[idx][jdx], horizontalalignment='center', transform=ax.transAxes)
+        ax.text(.5, .8, titles[idx][jdx], horizontalalignment='center', transform=ax.transAxes)
       #if idx == 0 and jdx==(ncol/2) and title != '':
       #  ax.set_title(titles[idx][jdx]+'\n '+title)  
       if idx == 0 and jdx==(ncol/2) and title != '':
-        ax.text(.5, .9, titles[idx][jdx]+'\n '+title, horizontalalignment='center', transform=ax.transAxes)
[…]

### 2016-06-08 Paper/paper.tex
adding stat err 1-d, 2-d figures
updated first two plots

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% #################
+\begin{figure*}
+\centering 
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
+\caption{We show here the measured probability distribution for the
+dimensionless mass ratio $\eta$ of a binary with
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and
+$\lambdans=2000$. The templates used here {\it ignore} tidal effects.
+These figures show that this approximation holds up to (SNR)
+$\rho\simeq 30-50$.
+% 
+In each panel- the dashed red line marks the {\it measured} value of $\eta$,
+which is given by the median of its probability distribution.
+The {\it true} value is marked by the dashed green line. The difference
+between the true and measured values is the systematic bias introduced
+by ignoring tidal effects in templates. The darker shading shows
+the recovered $90\%$ confidence interval for $\eta$. The width of this
+region is a direct measure of our statistical uncertainty.
+% 
+Comparing systematic and statistical errors, we find that:
+at $\rho=20$, $\eta$ measurement is dominated by statistical
+errors; at $\rho=30$, the two become comparable; and 
+for louder signals ($\rho\simeq50$), the systematic errors dominate.
+}
+\label{fig:SingleSystemEtaPDFvsSNR}
+\end{figure*}
+% #################
+\begin{figure*}
+\centering 
+\includegraphics[width=1.9\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}
+\caption{
+We show here the statistical uncertainty of our measurement of the 
+dimensionless mass ratio $\eta$ (at $90\%$ confidence), over the NSBH
+parameter space.
+Individual panels show the same as a function of BH mass and spin.
+Across each row, we see the effect of increasing signal strength (i.e.
+SNR) with the NS's tidal deformability $\lambdans$ fixed. Down each
+column, we see the effect of increasing $\lambdans$, at fixed SNR.
+}
+\label{fig:SingleSystemEtaPDFvsSNR}
+\end{figure*}
 % #################
@@
 have used (or plan to) BH-BH waveform templates for NS-BH mergers, ignoring
-the effect of the tidal deformation of the neutron star due to its companion hole
-on the emitted GWs. In this section we present the first fully Bayesian Monte-Carlo
-study of the effects of this assumption on the recovery of other physical parameters
-of NS-BH sources.
+NS tidal effects on the emitted GWs. In this section we present a fully
+Bayesian analysis of the effects of this assumption on the recovery of NSBH
+masses and spins from their GW signals.
 % We do so by comparing the systematic biases introduced by this 

### 2016-06-09 Paper/paper.tex
working on s3

--- 2016-06-08 Paper/paper.tex
@@
 \newcommand{\deff}{D_\mathrm{eff}}
+\newcommand{\arr}{\mathcal{r}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% #################
+% THESE FIGURES ARE FOR THE NEXT SECTION
+\begin{figure*}
+\centering 
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
+\caption{We show here the measured probability distribution for the
+dimensionless mass ratio $\eta$ of a binary with
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and
+$\lambdans=2000$. The templates used here {\it ignore} tidal effects.
+These figures show that this approximation holds up to (SNR)
+$\rho\simeq 30-50$.
+% 
+In each panel- the dashed red line marks the {\it measured} value of $\eta$,
+which is given by the median of its probability distribution.
+The {\it true} value is marked by the dashed green line. The difference
+between the true and measured values is the systematic bias introduced
+by ignoring tidal effects in templates. The darker shading shows
+the recovered $90\%$ confidence interval for $\eta$. The width of this
+region is a direct measure of our statistical uncertainty.
+% 
+Comparing systematic and statistical errors, we find that:
+at $\rho=20$, $\eta$ measurement is dominated by statistical
+errors; at $\rho=30$, the two become comparable; and 
+for louder signals ($\rho\simeq50$), the systematic errors dominate.
+}
+\label{fig:SingleSystemEtaPDFvsSNR}
+\end{figure*}
+% #################
+\begin{figure*}
+\centering 
+\includegraphics[width=1.9\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
+\includegraphics[width=1.9\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
+\includegraphics[width=1.9\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+\caption{
+We show here the statistical uncertainty of our measurement of the 
+dimensionless mass ratio $\eta$ (at $90\%$ confidence), over the NSBH
+parameter space.
+Individual panels show the same as a function of BH mass and spin.
+Across each row, we see the effect of increasing signal strength (i.e.
+SNR) with the NS's tidal deformability $\lambdans$ fixed. Down each
+column, we see the effect of increasing $\lambdans$, at fixed SNR.
+}
+\label{fig:CIWidths90_Lambda_SNR}
+\end{figure*}
+% ###############################
 % In this section we summarize various technical aspects of this paper. The model
@@
 Once more than $10,000$ independent samples have been collected, we integrate
-Eq.~\ref{eq:marginalize} to obtain the marginalized probability distribution
-for the tidal deformability of the NS $p(\lambdans|d_n,K)$.
-The most likely value of $\lambdans$ for this distribution is our 
-{\it measured} value for the parameter, and the $90\%$ confidence intervals
-associated with the distribution serve as statistical error-bars.
-% 
-% We use this
-% distribution to put confidence bounds on our measurement of $\lambdans$.
-% 
-% of this distribution to estimate our statistical and
+Eq.~\ref{eq:marginalize} to obtain marginalized probability distributions
+for the NS tidal deformability parameter: $p(\lambdans|d_n,K)$.
+The median value of this distribution is our {\it measured} value for
+$\lambdans$, and the $90\%$ confidence intervals associated with the distribution
+serve as statistical error-bars.
+%
 
@@
 
-\red{Reorder:
-We construct NSBH waveforms with LEA+. We inject them, and run MCMC on the injections
-using BBH templates. We perform this analysis for [... parameters ...]. We always
-use $\mns=1.35M_\odot$, but explore $\lambdans=\{500,1000,1500,2000\}$.
-}
-
-\prayush{
-We first take a set of NSBH binary waveforms, which include the modeling of spin and
-tidal effects.
-For all of the resulting signals, we perform a fully Bayesian parameter estimation
-aimed at measuring binary masses and spins from the signal. Our analysis
-templates differ from the signal in only one sense: all tidal terms are set to zero
-in the former, i.e. they correspond to BHBH mergers. Binary mass-ratio is sampled
-uniformly over the range $q\in[2,5]$ taking $q=\{2,3,4,5\}$ as our test points. The
-mass of the NS is fixed to $\mns=1.35M_\odot$, and therefore knowing $q$ gives $\mbh$.
-The spin of the NS is fixed to zero. The spin of the BH is sampled uniformly over the
-range $\chibh\in[-0.75, +0.75]$, choosing $\chibh=\{-0.5, 0, +0.5, +0.75\}$ as our
-test points. Finally, we vary the deformability parameter uniformly over 
-$\lambdans\in[0,2000]$, choosing $\lambdans=\{500, 1000, 1500, 2000\}$ plus an additional
-$\lambdans=800$ case. Waveform are generated for all possible combinations of these
-parameters, and rescaled to different strengths with $\rho=\{20, 30, 50, 70\}$.
-}
+% \red{Reorder:
+% We construct NSBH waveforms with LEA+. We inject them, and run MCMC on the injections
+% using BBH templates. We perform this analysis for [... parameters ...]. We always
+% use $\mns=1.35M_\odot$, but explore $\lambdans=\{500,1000,1500,2000\}$.
+% }
+
+We construct NSBH waveforms using the LEA+ model. We inject these into zero
+noise at different SNRs, and run MCMC on the injections using BBH 
+templates (LEA+ with tidal terms set to zero). We perform this analysis for mass
+ratios $q\in\{2,3,4,5\}$, i.e.
+$\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$, and BH spins
+$\chibh\in\{-0.5, 0, +0.5, +0.75\}$. We always use $\mns=1.35M_\odot$ and
+$\chins=0$, but explore $\lambdans\in\{500, 800, 1000, 1500, 2000\}$.
+
+% We simulate NSBH signal waveforms using the LEA+ model, inject them into zero
+% noise, and perform fully Bayesian parameter estimation analyses on them to
+% estimate their source masses and spins. We set tidal terms in search
+% templates to zero, ensuring that they mimic BBH mergers. We perform this
+% analysis for a range of systems: with mass ratios $q\in\{2,3,4,5\}$, BH spins
+% $\chibh\in\{-0.5, 0, +0.5, +0.75\}$, and NS tidal deformability
+% $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. We fix NS mass to $1.35M_\odot$
+% and its spin to zero (for signals). In addition, signals are simulated at
+% different strengths, i.e. $\rho=\{20, 30, 50, 70\}$, in order to estimate the
+% dependence of parameter measurement errors on the SNR.
+
+As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show the
+probability distribution for binary mass ratio $\eta$ recovered from a GW
+injection with intrinsic parameters: $\mns=1.35M_\odot$, $\chins=0$,
+$\mbh=5.4M_\odot$, $\chibh=+0.5$ and $\lambdans=2000$, and SNRs ranging
+from $\rho=20$ (left panel) to $\rho=50$ (right panel). In each panel, the
+{\it true} value of $\eta$ is marked by a dashed green line, and its measured
+(median) value by a dashed red line. The effect of ignoring tidal corrections
+in templates manifests as a systematic shift of the measured $\eta$ value away
+from its true value. The associated {\it systematic} bias/error is given by the
+difference between measured and true $\eta$ values. Darker shading of the
+probability distribution marks the $90\%$ confidence interval for $\eta$, whose
+width is a direct measure of the {\it statistical} measurement uncertainty/error.
+% 
+We clearly see that even when the signal is moderately loud, with $\rho=20$,
+statistical errors dominate over systematics completely for $\eta$.
+As we turn up the SNR
+further, we find that the two error sources become comparable at $\rho\sim 30$,
+and systematic errors dominate finally when $\rho\simeq 50$.
+
+Next, to get a better understanding of how {\it precisely} we measure NSBH
+binary parameters given BBH waveforms, we show the measured $90\%$ confidence
+intervals for binary masses and spins in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
+The three panels, corresponding to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
+(bottom), show the width of these confidence intervals as a function of BH
+mass/spin (within each sub-panel), and NS properties, i.e. $\lambdans$
+(downwards in each column).~\footnote{We restrict NS mass to $1.35M_\odot$ and
+its spin to zero. Varying its tidal deformability $\lambdans$ does not change
+the measurement uncertainties for non-tidal binary parameters, as is evident
+from comparing the two rows in each panel of Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}
+% 
+From the left-most column, we find that: (i) at $\rho=20, \mchirp$ is already
+measured remarkably well - to a precision of $0.16\%$ of its true value, and
+(ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined somewhat worse,
+with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
+all three measurements gain further precision, with $\eta$ errors shrinking down
+to single-digit percents of its true value.
+% 
+However, we remind ourselves that there are physical effects that have not been
+incorporated in our (BBH) search templates which we use here to decipher NSBH
+signals. Therefore, we cannot know the {\it accuracy} of our parameter
+measurements unless we know how large the systematic biases are due to an
+incomplete template model. To quantify this, we find it useful to define $\arr$
+as the ratio of the systematic biases induced by unmodeled tidal effects in
+templates, to the statistical uncertainties associated with the measurement
+itself. If $\arr\gtrsim 1$, including the effect of NS's
+tidal distortion (during inspiral) and possible disruption (close to merger)
+would be imperative to yield reliable estimates of binary masses and spins
+from its measured GW signal.
 

### 2016-06-09 Paper/paper.tex
polishing s3 more

@@
 \newcommand{\deff}{D_\mathrm{eff}}
-\newcommand{\arr}{\mathcal{r}}
+\newcommand{\arr}{\mathtt{r}}
 
@@
 \centering 
-%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR}
-\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
+\includegraphics[width=1.8\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
 \caption{We show here the ratio of systematic and statistical
@@
 % \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
-\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
+\includegraphics[width=1.8\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
 \caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
@@
 \centering
-\includegraphics[trim={1.3cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
+\includegraphics[width=1.8\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
 % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
@@
 We construct NSBH waveforms using the LEA+ model. We inject these into zero
-noise at different SNRs, and run MCMC on the injections using BBH 
-templates (LEA+ with tidal terms set to zero). We perform this analysis for mass
-ratios $q\in\{2,3,4,5\}$, i.e.
-$\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$, and BH spins
-$\chibh\in\{-0.5, 0, +0.5, +0.75\}$. We always use $\mns=1.35M_\odot$ and
-$\chins=0$, but explore $\lambdans\in\{500, 800, 1000, 1500, 2000\}$.
+noise and run MCMC on the injections using BBH templates (LEA+ with tidal terms
+set to zero). We perform this analysis for a variety of mass ratios
+$q\in\{2,3,4,5\}$, i.e. $\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$,
+and BH spins $\chibh\in\{-0.5, 0, +0.5, +0.75\}$. We always use $\mns=1.35M_\odot$ and
+$\chins=0$, but explore $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. The signal
+strength is varied from $\rho=20\rightarrow70$. Our choice of this SNR range is
+motivated by previous studies which suggest that signatures of NS tidal effects
+become observable only at relatively large SNRs~\cite{Lackey:2013axa}.
 
@@
 Next, to get a better understanding of how {\it precisely} we measure NSBH
-binary parameters given BBH waveforms, we show the measured $90\%$ confidence
-intervals for binary masses and spins in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
+binary parameters given BBH waveforms, we show the $90\%$ confidence intervals
+associated with NSBH mass and spin measurements in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
 The three panels, corresponding to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
@@
 mass/spin (within each sub-panel), and NS properties, i.e. $\lambdans$
-(downwards in each column).~\footnote{We restrict NS mass to $1.35M_\odot$ and
+(downwards in each column)~\footnote{We restrict NS mass to $1.35M_\odot$ and
 its spin to zero. Varying its tidal deformability $\lambdans$ does not change
 the measurement uncertainties for non-tidal binary parameters, as is evident
-from comparing the two rows in each panel of Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}
+from comparing the two rows in each panel of Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
 % 
@@
 measured remarkably well - to a precision of $0.16\%$ of its true value, and
-(ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined somewhat worse,
+(ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined more loosely,
 with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
@@
 signals. Therefore, we cannot know the {\it accuracy} of our parameter
-measurements unless we know how large the systematic biases are due to an
-incomplete template model. To quantify this, we find it useful to define $\arr$
-as the ratio of the systematic biases induced by unmodeled tidal effects in
+measurements unless we know how large the systematic biases due to an
+incomplete template model are. To quantify this, we find it useful to define
+$\arr$, the ratio of the systematic biases induced by unmodeled tidal effects in
 templates, to the statistical uncertainties associated with the measurement
-itself. If $\arr\gtrsim 1$, including the effect of NS's
-tidal distortion (during inspiral) and possible disruption (close to merger)
-would be imperative to yield reliable estimates of binary masses and spins
-from its measured GW signal.
-
-In Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR} we show the ratio of the systematic
-bias in the measured binary chirp mass $\mchirp$ versus its true value, to the 
-statistical uncertainty of the measurement itself. $\mchirp$ is the mass combination
-that occurs at leading order in the perturbative expression for GW strain from 
-compact binaries, and is therefore very precisely measurable. Even so, we find that
-for $\rho=30$ the signal is not loud enough for the systematic bias in $\mchirp$
-to exceed a fraction of the statistical uncertainty. When we increase the SNR to 
-$\rho=50$ (middle column), we find that if we have two favorable conditions present:
-(a) that the BH spin be $\gtrsim 0.4$, {\it and} (b) the NS's true deformability
-be large enough that its $\lambdans \gtrsim 1000$, the systematic biases from
-ignoring tidal effects in templates reach between $50-100\%$ of the statistical
-uncertainty. Therefore at this signal strength, including terms involving the neutron
-star's internal structure in waveform models becomes increasingly gainful. At 
-even higher SNRs, say $\rho=70$ (right column), under the same favorable conditions
-as listed above, the systematic bias in $\mchirp$ measurement {\it exceeds} its 
-statistical measurement uncertainty. If the BH is even mildly spinning, say
-$\chibh\gtrsim 0.2$, with the NS mildly deformable, say $\lambdans\gtrsim 800$, 
-we still are affected in our $\mchirp$ measurement at the $50\%$ level by the 
-use of point-particle waveform templates. Finally, we note that the measured (median)
-$\mchirp$ is always higher than its true value. This is reasonable since the tidal 
-deformation of NSs close to merger reduces the GW signal power in high frequencies, so
-it would look like a BH-BH signal of a higher mass (and therefore lower merger or
-peak GW-power frequency).
-
-In the next figure, Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the same
-information (i.e. the ratio of systematic bias to statistical uncertainty in parameter
-measurement) for the symmetric mass-ratio $\eta$. Note that $\eta$
+itself. If $\arr\gtrsim 1$, including the effect of NS's tidal distortion
+(during inspiral) and possible disruption (close to merger) would be imperative
+to obtain {\it accurate} estimates of source masses and spins from observed GW
+signals.
+
+We show this ratio of errors $\arr$ associated with chirp mass measurements in
+Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the mass combination
+that occurs at leading order in the perturbative expansion of GW strain emitted
+by compact binaries as they spiral in, and is therefore determined the most
+precisely. From the left-most column of the figure, we find that even at $\rho=30$, 
+the systematic bias in $\mchirp$ measurement is much smaller than its statistical
+uncertainty. For signals with $\rho\leq 30$, therefore, we are justified in using
+BBH templates to estimate binary chirp mass from NSBH signals. For louder and less
+realistic SNRs (say $\rho=50$), we find that $\arr$ can become comparable to unity
+if additionally: (i) the BH has prograde spin $\chibh\gtrsim 0.4$, {\it and} (ii)
+the NS's true deformability is large enough, s.t. $\lambdans \gtrsim 1000$.
+% 
+At even higher SNRs, say $\rho=70$, the ratio of errors $\arr$ is large
+($\simeq 0.5-1.5$) across most of the binary parameter space; and $\mchirp$ measurement
+is affected at a $50\%$ or higher level by the use of point-particle waveforms as
+templates.
+% 
+We conclude that only for $\rho\gtrsim 50$ do we need to include in our templates
+terms that capture the effect of neutron star's structural dynamics on emitted GWs.
+For lower SNRs, improvements in template model will be washed out by the statistical
+measurement uncertainty for $\mchirp$.
+% 
+As a last note, we find that $\arr$ is always positive for $\mchirp$, suggesting that
+it is always over-estimated.
+This is understandable since the tidal deformation of NSs close to merger reduces
+the GW signal power at high frequencies, which would make the resulting signal
+resemble a BBH signal of higher mass (and therefore lower merger frequency).
+
+In the next figure, Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
+measurement errors $\arr$ for the symmetric mass-ratio $\eta$. Note that $\eta$
 is bounded from above and below to within $\eta\in[0, 1/4]$. We immediately observe
 that the measured (median) $\eta$ is always lowered below its true value due to the 
-lack of $\lambdans$-dependent terms in the filter template model. This bias, as for 
-the $\mchirp$ bias, decreases the merger frequency for the BH-BH template, which must
-better match the NS-BH merger with its quietened merger and ringdown. At SNR $\rho=30$
+lack of $\lambdans$-dependent terms in the filter template model. This bias,
+similar to the $\mchirp$ bias, decreases the merger frequency for BBH templates,
+in order to better match the relatively quiet merger and ringdown of NSBHs.
+At SNR $\rho=30$
 the systematic bias in $\eta$ measurement stays mostly around $10-50\%$ of its statistical

### 2016-06-09 Paper/paper.tex
polishing s3 mchirp

@@
 
-We show this ratio of errors $\arr$ associated with chirp mass measurements in
+We first calculate the ratio of errors $\arr$ associated with chirp mass measurements
+and show it in
 Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the mass combination
-that occurs at leading order in the perturbative expansion of GW strain emitted
+that affects at leading order the GW strain emitted
 by compact binaries as they spiral in, and is therefore determined the most
@@
 realistic SNRs (say $\rho=50$), we find that $\arr$ can become comparable to unity
-if additionally: (i) the BH has prograde spin $\chibh\gtrsim 0.4$, {\it and} (ii)
+if additionally: (i) the BH has a prograde spin $\chibh\gtrsim 0.4$, {\it and} (ii)
 the NS's true deformability is large enough, s.t. $\lambdans \gtrsim 1000$.
 % 
-At even higher SNRs, say $\rho=70$, the ratio of errors $\arr$ is large
-($\simeq 0.5-1.5$) across most of the binary parameter space; and $\mchirp$ measurement
-is affected at a $50\%$ or higher level by the use of point-particle waveforms as
-templates.
-% 
-We conclude that only for $\rho\gtrsim 50$ do we need to include in our templates
+It is only at SNRs as high as $\rho=70$ that the systematic biases for $\mchirp$
+exceed the underlying statistical uncertainty to become the dominant source of
+errors.
+% 
+We conclude that only for $\rho\gtrsim 50-70$ do we need to include in our templates
 terms that capture the effect of neutron star's structural dynamics on emitted GWs.

### 2016-06-09 Paper/paper.tex
s3 first pass

@@
 \centering
-%\includegraphics[trim={11cm 0 0 0},width=2.7\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI68_3_Lambda_SNR.pdf}\\
-% \includegraphics[trim={3.6cm 0 0 0},width=2.25\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR.pdf}\\
 \includegraphics[width=1.8\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
@@
 \centering
-\includegraphics[width=1.8\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
+\includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
 % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
@@
 masses and spins from their GW signals.
-% We do so by comparing the systematic biases introduced by this 
-% assumption to the overall statistical uncertainties in the measurement of binary 
-% mass and spin parameters. We find that unless the single-detector SNR of the signal
-% exceeds $30$ (for design Advanced LIGO sensitivity), the bias accurued by ignoring
-% tidal effects in filtering templates will remain smaller than the statistical 
-% uncertainty of the same measurement.
-
-% \red{Reorder:
-% We construct NSBH waveforms with LEA+. We inject them, and run MCMC on the injections
-% using BBH templates. We perform this analysis for [... parameters ...]. We always
-% use $\mns=1.35M_\odot$, but explore $\lambdans=\{500,1000,1500,2000\}$.
-% }
 
@@
 motivated by previous studies which suggest that signatures of NS tidal effects
-become observable only at relatively large SNRs~\cite{Lackey:2013axa}.
-
-% We simulate NSBH signal waveforms using the LEA+ model, inject them into zero
-% noise, and perform fully Bayesian parameter estimation analyses on them to
-% estimate their source masses and spins. We set tidal terms in search
-% templates to zero, ensuring that they mimic BBH mergers. We perform this
-% analysis for a range of systems: with mass ratios $q\in\{2,3,4,5\}$, BH spins
-% $\chibh\in\{-0.5, 0, +0.5, +0.75\}$, and NS tidal deformability
-% $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. We fix NS mass to $1.35M_\odot$
-% and its spin to zero (for signals). In addition, signals are simulated at
-% different strengths, i.e. $\rho=\{20, 30, 50, 70\}$, in order to estimate the
-% dependence of parameter measurement errors on the SNR.
+become observable only at relatively large SNRs~\cite{Lackey:2013axa}, and is validated
+by the results to follow. At design sensitivity, if we expect $10-300$ NSBH detections 
+a year~\cite{Abadie:2010cfa}, and if BH masses are uniformly likely up to
+$\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we would expect to detect $0.3-7$ disruptive
+NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with $\rho\geq 30$, and $0.04-1$ a
+year with $\rho\geq 70$.
 
@@
 % 
-It is only at SNRs as high as $\rho=70$ that the systematic biases for $\mchirp$
+It is only at SNRs as high as $\rho\simeq70$ that the systematic biases for $\mchirp$
 exceed the underlying statistical uncertainty to become the dominant source of
@@
 % 
-As a last note, we find that $\arr$ is always positive for $\mchirp$, suggesting that
+Lastly, we note that $\arr$ is always positive for $\mchirp$, suggesting that
 it is always over-estimated.
@@
 
-In the next figure, Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
-measurement errors $\arr$ for the symmetric mass-ratio $\eta$. Note that $\eta$
-is bounded from above and below to within $\eta\in[0, 1/4]$. We immediately observe
-that the measured (median) $\eta$ is always lowered below its true value due to the 
-lack of $\lambdans$-dependent terms in the filter template model. This bias,
-similar to the $\mchirp$ bias, decreases the merger frequency for BBH templates,
-in order to better match the relatively quiet merger and ringdown of NSBHs.
-At SNR $\rho=30$
-the systematic bias in $\eta$ measurement stays mostly around $10-50\%$ of its statistical
-measurement uncertainty. For the sole case of a very deformable NS with $\lambdans=2000$,
-the missing tidal pieces in template models could cause errors that are comparable to 
-that due to noise (i.e. statistical uncertainties). For a still stronger signal with 
-$\rho=50$, we find that 
-when the following conditions are fulfilled, the measurement of $\eta$ is {\it completely
-compromised} due to template biases: (a) BH mass $\mbh\leq 5M_\odot$, (b) BH spin
-$\chibh\gtrsim +0.4$, and (c) $\lambdans\gtrsim 1500$. Even under more moderate restrictions
-on BH and NS parameters, we find that $\rho=50$ is loud enough to necessitate the 
-use of tidal-modeled templates. This observation is further strengthened if we divert 
-our attention to the third column in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR} that presents
+Next, in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
+measurement errors $\arr$ for the symmetric mass-ratio $\eta$. We immediately
+observe that the measured (median) $\eta$ is always lowered below its true
+value due to the lack of $\lambdans$-dependent terms in the filter template
+model. This bias, similar to the $\mchirp$ bias, reduces the merger frequency
+for BBH templates, in order to better match the relatively quiet merger and
+ringdown of NSBHs. 
+% 
+At $\rho=30$, the systematic bias in $\eta$ measurement stays mostly around
+$10-50\%$ of its statistical measurement uncertainty. For the sole case of a
+very deformable NS with $\lambdans=2000$, can the effect of missing tidal pieces
+in the template model become comparable to that of noise (which is usually
+more restrictive on measurements).
+% 
+For a still stronger signal with $\rho=50$, we find that when the following
+conditions are fulfilled, the measurement of $\eta$ is unreliable due to
+template biases: (a) BH mass $\mbh\leq 5M_\odot$, (b) BH spin
+$\chibh\gtrsim +0.4$, and (c) $\lambdans\gtrsim 1500$. Even under more
+moderate restrictions on BH and NS parameters, we find that $\rho=50$ is 
+loud enough to motivate the use of tidal-modeled templates. This observation
+is further strengthened if we divert our attention to the third column in
+Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR} that presents
 the case for even higher SNRs ($\rho=70$).

### 2016-06-10 Paper/paper.tex
added s4 figs

@@
 \centering 
-\includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
-%\includegraphics[trim={3cm 0 0 0},width=2.3\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
+\caption{We show here the probability distribution recovered for the NS tidal
+deformability parameter $\lambdans$ from a GW injection, with the following
+parameters characterizing the binary:
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and
+$\lambdans=2000$. The templates used here {\it ignore} tidal effects.
+These figures show that this approximation holds up to (SNR)
+$\rho\simeq 30-50$.
+% 
+In each panel- the dashed red line marks the {\it measured} value of
+$\lambdans$,
+which is given by the median of its probability distribution.
+The {\it true} value is marked by the dashed green line. The difference
+between the true and measured values is the systematic bias introduced
+by ignoring tidal effects in templates. The darker shading shows
+the recovered $90\%$ confidence interval for $\lambdans$. The width of this
+region is a direct measure of our statistical uncertainty.
+% 
+Comparing systematic and statistical errors, we find that:
+at $\rho=20$, $\lambdans$ measurement is dominated by statistical
+errors; at $\rho=30$, the two become comparable; and 
+for louder signals ($\rho\simeq50$), the systematic errors dominate.
+}
+\label{fig:SingleSystemLambdaPDFvsSNR}
+\end{figure*}
+% #################
+\begin{figure*}
+\centering    
+\includegraphics[trim={3cm 0 0 0},width=1.9\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
+% \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
 \caption{This figure shows the statistical uncertainty in the measurement of

### 2016-06-10 emcee_PE/scripts/SingleObservation.py

@@ -80,10 +80,11 @@ plt.rcParams.update({\
 
 # Plotting functions
 def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
-        title='', titles=[], levellines=[100,200], levelspacing=0.25,\
+        title='', titles=[], xticks=None, yticks=None,\
+	levellinesscale=None, levellines=[100,200], levellabels=True, levelspacing=0.25,\
         vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f', lfmt='%.1f',\
         xmin=None, xmax=None, ymin=None, ymax=None, alpha=0.9,\
-        colorbartype='simple', figname='plot.png'):
+        colorticks=None, colorbartype='simple', figname='plot.png'):
   """
   Function to plot arbitrary numbers of contour plots in a single figure
   """
@@ -121,11 +122,13 @@ col corresponds to the lower level group
   
   colwidth = 1.3
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((1.3*gmean*ncol+1.25)*colwidth, 1.65*colwidth*nrow),\
+              figsize=((1.3*gmean*ncol+1.75)*colwidth, 1.2*colwidth*nrow),\
               dpi=100)
   fig.clf()
-  grid = ImageGrid(fig, 111, nrows_ncols=(nrow, ncol), \
+  grid = ImageGrid(fig, 111,\
+            nrows_ncols=(nrow, ncol), \
             share_all=True,\
+            axes_pad=0.05,\
             cbar_mode="single", cbar_location="right",\
             cbar_pad=0.05, cbar_size="2%", \
             aspect=True,\
@@ -147,12 +150,19 @@ col corresponds to the lower level group
   
   ## FIXME
   #VMIN = 0.01
-
+  
   # Now make contour plots
   contours_all = {}
   for idx in range(nrow):
     contours_all[idx] = {}
     for jdx in range(ncol):
+      levline_scaling = 1
+      try:
+        if levellinesscale is not None:
+          levline_scaling = levellinesscale[idx][jdx]
+      except: pass
+      #
+      print "Scaling level lines with %f" % levline_scaling
       try: xx, yy, zz = X[idx][jdx], Y[idx][jdx], Z2d[idx][jdx]
       except: 
         print "Array shapes = ", np.shape(xx), np.shape(yy), np.shape(zz)
[…]

### 2016-06-10 Paper/paper.tex
updated figs

@@
 \centering
-\includegraphics[width=1.8\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear002}
+\includegraphics[width=2\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_70_linear}
 \caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
@@
 \centering    
-\includegraphics[trim={3cm 0 0 0},width=1.9\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
+\includegraphics[trim={2cm 0 0 0},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
 % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
@@
 i.e. (c), simply makes the NS more easily deformable. All of them reduce the onset 
-frequency for NS's disruption close to merger. Therefore, systems that fulfil the 
+frequency for NS's disruption close to merger. Therefore, systems that fulfill the 
 above conditions have considerably (or, at least, the most) different GW morphology
@@
 % 
+\red{Move to discussion}
+\prayush{%
 We find that when the BH is sufficiently spinning in the prograde sense,
@@
 matter, and possibly narrow the range they span.
-
+}
   `
-
+\red{Is this the LEA+ mode? If yes, say so}
+\prayush{%
 We use the same set of systems as in the previous section, with $q=\{2,3,4,5\}$,
@@
 We use the same waveform model as input to our Bayesian parameter estimation
-method in order to obtain estimates for different binary parameters alongwith 
-their full probability distributions, including for the combined NS tidal
+method in order to obtain estimates for different binary parameters along with 
+their full probability distributions, including for the NS tidal
 deformability parameter $\lambdans$.
-
+}
+
+\red{Present posterior distributions of selected configurations first..}
+
+\red{%
+\begin{enumerate}
+ \item State the prior on $\lambdans$ because $\lambdans$ weakly deforms waveforms,
+I would expect that measurement uncertainty in $\lambdans$ is approx. independent
+of $\lambdans$, except for running into the $\lambdans\geq 0$ prior.
+ \item Is $\lambdans$ measurement degenerate with other parameters? Either cite 
+ known studies [if any], or provide 2-d posterior plots.
+\end{enumerate}
+}
 The precision with which $\lambdans$ can be measured is given by the width of 

### 2016-06-10 emcee_PE/scripts/SingleObservation.py

@@ -109,8 +109,8 @@ col corresponds to the lower level group
     'axes.labelsize':16,\
     'font.family':'serif',\
     'font.size':16,\
-    'xtick.labelsize':16,\
-    'ytick.labelsize':16,\
+    'xtick.labelsize':14,\
+    'ytick.labelsize':14,\
     'figure.subplot.bottom':0.2,\
     'figure.figsize':figsize, \
     'savefig.dpi': 300.0})#, \
@@ -128,7 +128,7 @@ col corresponds to the lower level group
   grid = ImageGrid(fig, 111,\
             nrows_ncols=(nrow, ncol), \
             share_all=True,\
-            axes_pad=0.05,\
+            axes_pad=0.2,\
             cbar_mode="single", cbar_location="right",\
             cbar_pad=0.05, cbar_size="2%", \
             aspect=True,\
@@ -185,7 +185,7 @@ col corresponds to the lower level group
       contours_tmp = {}
       for lev in levellines:
         cset = ax.contour(xx, yy, zz, levels=[lev * levline_scaling],\
-				colors='k', ls='--', linewidths=2, hold="on")
+			alpha=0.8, colors='k', ls='--', linewidths=2, hold="on")
         for c in cset.collections: c.set_linestyle('dotted')
         if levellabels:
           label_dict = {}
@@ -197,7 +197,7 @@ col corresponds to the lower level group
         print "for %s" % titles[idx][jdx], contours_tmp
         print "ymin, ymax = ", ymin, ymax
       #
-      ax.grid(True)
+      ax.grid(True, color='k')
       if xticks is not None: ax.get_xaxis().set_ticks(xticks)
       if yticks is not None: ax.get_yaxis().set_ticks(yticks)
       ax.set_xlim([xmin, xmax])

### 2016-06-10 Paper/paper.tex
updated TN figs

@@
 \centering 
-\includegraphics[width=1.9\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
-\includegraphics[width=1.9\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
-\includegraphics[width=1.9\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
 \caption{
@@
 % \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
-\caption{In these figures we highlight the regions of BH mass and spin plane where we can 
+\caption{\red{Correct caption.}
+In these figures we highlight the regions of BH mass and spin plane where we can 
 constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
@@
 set by the detectors).
+\red{Use same colorbar bounds in both panels}.
 }
@@
 we show the lowest SNR value $\rho_c$ in order for the signal to have enough
-information that we can constrain $\dlambda$ below $200\%$. Note that 
-being able to constrain $\lambdans$ within a factor of $2$ interval 
+information that we can constrain $\dlambda$ below $200\%$.
+\red{%
+\begin{enumerate}
+ \item Inconsistent notation $\delta$ vs $\dlambda$. Unify.
+ \item Explain precise definition of $\delta\lambdans/\dlambda$: Is this the 
+ full width or half width of the posterior. Such explanation fit in nicely into
+ the discussion of the new $\lambdans$ posterior plots like Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}.
+\end{enumerate}
+}
+\hspace{-3mm}Note that being able to constrain $\lambdans$ within a factor of $2$ interval 
 (i.e. within $\pm 100\%$) is a fairly modest requirement. We find promising 
@@
 wide (but non-trivial) range of BH masses, within
-which if we get a moderately loud NSBH signal with $\rho\approx 20$, we constrain
+which if we get a moderately loud NSBH signal with $\rho\approx 23$, we constrain
 $\lambdans$ within a factor of $2$ interval (i.e. within $\pm 100\%$) around
@@
 If we want to constrain $\lambdans$ better with a single observation, however,
-we need SNRs of $30$ or higher in each detector. This can be seen from the right
+we need SNRs of $30$ or higher. This can be seen from the right
 panel of Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}, which is similar
@@
 Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
-conditions that BH parameters must fulfil beside having an SNR of $20$ (left 
-panel) or $30$ (right panel), to obtain meaningful contraints on $\lambdans$.
+conditions that BH parameters must fulfill beside having an SNR of $20$ (left 
+panel) or $30$ (right panel), to obtain meaningful constraints on $\lambdans$.
 At $\rho=20$, the left panel shows that we need high BH spins 
@@
 $2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
-NS with $\lambdans=200$ one could even put a $\pm 50\%$ errorbar on 
+NS with $\lambdans=200$ one could even put a $\pm 50\%$ error-bar on 
 $\lambdans$ measurement.

### 2016-06-10 emcee_PE/scripts/SingleObservation.py

@@ -122,7 +122,7 @@ col corresponds to the lower level group
   
   colwidth = 1.3
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((1.3*gmean*ncol+1.75)*colwidth, 1.2*colwidth*nrow),\
+              figsize=((1.5*gmean*ncol+1.75)*colwidth, 1.5*colwidth*nrow),\
               dpi=100)
   fig.clf()
   grid = ImageGrid(fig, 111,\
@@ -185,7 +185,7 @@ col corresponds to the lower level group
       contours_tmp = {}
       for lev in levellines:
         cset = ax.contour(xx, yy, zz, levels=[lev * levline_scaling],\
-			alpha=0.8, colors='k', ls='--', linewidths=2, hold="on")
+			alpha=1., colors='k', ls='--', linewidths=2, hold="on")
         for c in cset.collections: c.set_linestyle('dotted')
         if levellabels:
           label_dict = {}

### 2016-06-10 Paper/paper.tex
updating more figs

@@
 % }
+
+\red{SOMEHOW INTRODUCE SNR AND MENTION IF IT IS A NETWORK SNR OR WHAT KIND OF 
+'SNR' IS IT?}
 
@@
 \centering 
-\includegraphics[width=1.8\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
+\includegraphics[width=1.6\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
 \caption{We show here the ratio of systematic and statistical
@@
 % 
-Contours are drawn in each panel demarkating regions where we can constrain the
+Contours are drawn in each panel demarcating regions where we can constrain the
 $\lambdans$ parameter well (within a factor of two of the injected value).
@@
 \centering    
-\includegraphics[width=0.9025\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1000_0_CI90_0}
-\includegraphics[width=0.9025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
+\includegraphics[trim={0 0 0 2cm}, width=0.95\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0.pdf}
+\includegraphics[width=0.95\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0.pdf}
 % \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
@@
 \includegraphics[width=0.9025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
-\caption{In these figures we highlight the regions of BH mass and spin plane 
+\caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}
+In these figures we highlight the regions of BH mass and spin plane 
 where we can constrain the $\lambdans$ of its companion NS within a factor of 
@@
 constraint requirement that $\lambdans$ be measured within a factor of $1$ of 
-its true value (i.e. with $\pm 50\%$ errors, instead of $\pm 100\%$). Even for
+its true value (i.e. with $\pm 50\%$ errors, instead of $\pm 100\%$).
+\red{Dont use two different thresholds in figure.} Even for
 the most conducive cases, we {\it need} the SNR to be above $30$ to reach our 
@@
 
+\red{%
+(i) Rework text, so that you slowly build up the information needed. First, explain
+the general idea pursued, then discuss one panel first, then introduce second
+panel, finally, contrast both panels.
+(ii) High information density. Distinction between left and right panels notation
+not needed in the first sentence. deferring this info by a few sentences will make
+it easier for the reader to follow the discussion.
+}
+\prayush{%
 Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
@@
 $2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
-NS with $\lambdans=200$ one could even put a $\pm 50\%$ error-bar on 
+NS with $\lambdans=2000$ one could even put a $\pm 50\%$ error-bar on 
 $\lambdans$ measurement.
+}
 

### 2016-06-10 emcee_PE/scripts/SingleObservation.py

@@ -122,7 +122,7 @@ col corresponds to the lower level group
   
   colwidth = 1.3
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((1.5*gmean*ncol+1.75)*colwidth, 1.5*colwidth*nrow),\
+              figsize=((1.3*gmean*ncol+1.75)*colwidth, 1.2*colwidth*nrow),\
               dpi=100)
   fig.clf()
   grid = ImageGrid(fig, 111,\
@@ -185,13 +185,13 @@ col corresponds to the lower level group
       contours_tmp = {}
       for lev in levellines:
         cset = ax.contour(xx, yy, zz, levels=[lev * levline_scaling],\
-			alpha=1., colors='k', ls='--', linewidths=2, hold="on")
+			colors='k', ls='--', linewidths=1.2, hold="on")
         for c in cset.collections: c.set_linestyle('dotted')
         if levellabels:
           label_dict = {}
           for ll in [lev * levline_scaling]: label_dict[ll] = str(int(lev*100))
           lfmt = label_dict
-        plt.clabel(cset, colors='brown', inline=1, fmt=lfmt, fontsize=12)
+        plt.clabel(cset, colors='r', inline=1, fmt=lfmt, fontsize=10)
         contours_tmp[lev] = cset.collections[0].get_paths()
       if vverbose:
         print "for %s" % titles[idx][jdx], contours_tmp
@@ -320,11 +320,12 @@ def make_contour(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', title='',\
                 levelspacing=0.25, vmin=None, vmax=None, cmap=cm.Reds_r,\
                 xmin=None, xmax=None, ymin=None, ymax=None,\
                 cbfmt='%.1f', figname='plot.png'):
-  colwidth = 4.8
-  plt.figure(int(1e7 * np.random.random()), figsize=(1.*gmean*colwidth, colwidth))
+  colwidth = 4.
+  plt.figure(int(1e7 * np.random.random()), figsize=(0.75*gmean*colwidth, 0.75*colwidth))
   norm = cm.colors.Normalize(vmax=Z2d.max(), vmin=Z2d.min())
   #cmap = cm.rainbow
   levels = np.arange(Z2d.min(), Z2d.max(), levelspacing)
+  print np.min(Z2d), np.max(Z2d), levelspacing
   plt.contourf( X, Y, Z2d,\
               levels=levels, \
[…]

### 2016-06-10 Paper/paper.tex
augmenting auxiliary figs

@@
 \centering    
-\includegraphics[width=0.9025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
-\includegraphics[width=0.9025\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
+\includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
+\includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
 \caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}

### 2016-06-10 emcee_PE/scripts/SingleObservation.py

@@ -50,7 +50,7 @@ ErrThresh=[100, 200]
 # PLOTTING :
 # Function to make parameter bias plots
 ######################################################
-linestyles = ['dotted', 'dashed', 'solid', 'dashed', 'dotted', 'dashdot', 'dashed', 'dotted']
+linestyles = ['solid', 'dashed', 'dashdot', 'dashed', 'dotted', 'dashdot', 'dashed', 'dotted']
 #linecolors = ['crimson', 'olivedrab', 'k', 'b', 'm', 'y']
 linecolors = ['crimson', 'darkorange', 'olivedrab', 'royalblue', 'purple', 'k']
 linemarkers= ['', 'x', 'o', '^']

### 2016-06-12 Paper/paper.tex
added bunch of 2d histograms, and 1-d pdfs

@@
 \includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
-\caption{We show here the measured probability distribution for the
-dimensionless mass ratio $\eta$ of a binary with
-$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and
-$\lambdans=2000$. The templates used here {\it ignore} tidal effects.
-These figures show that this approximation holds up to (SNR)
-$\rho\simeq 30-50$.
-% 
-In each panel- the dashed red line marks the {\it measured} value of $\eta$,
-which is given by the median of its probability distribution.
-The {\it true} value is marked by the dashed green line. The difference
-between the true and measured values is the systematic bias introduced
+\caption{We show here probability distributions for mass ratio $\eta$ as measured
+for the same injection at different SNRs. The intrinsic parameters of the source
+are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
+and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
+{\it ignore} tidal effects.
+% 
+In each panel: the dashed red line marks the median {\it measured} value
+$\eta^\mathrm{Median}$, while the dashed green line show the true value
+$\eta^\mathrm{Injected}$. The difference between the true and measured values
+is the systematic bias introduced
 by ignoring tidal effects in templates. The darker shading shows
-the recovered $90\%$ confidence interval for $\eta$. The width of this
-region is a direct measure of our statistical uncertainty.
+the recovered $90\%$ confidence interval for $\eta$, $(\Delta\eta)^{90\%}$, 
+which is a direct measure of our statistical uncertainty on $\eta$.
 % 
@@
 \centering 
-\includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
-\caption{We show here the probability distribution recovered for the NS tidal
-deformability parameter $\lambdans$ from a GW injection, with the following
-parameters characterizing the binary:
+% \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50}
+\caption{We illustrate here probability distributions recovered for the NS tidal
+deformability parameter $\lambdans$ from three GW injections, with parameters:
 $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
-$\lambdans=2000$. The templates used here {\it ignore} tidal effects.
-These figures show that this approximation holds up to (SNR)
-$\rho\simeq 30-50$.
-% 
-In each panel- the dashed red line marks the {\it measured} value of
-$\lambdans$,
-which is given by the median of its probability distribution.
-The {\it true} value is marked by the dashed green line. The difference
-between the true and measured values is the systematic bias introduced
-by ignoring tidal effects in templates. The darker shading shows
-the recovered $90\%$ confidence interval for $\lambdans$. The width of this
-region is a direct measure of our statistical uncertainty.
-% 
-Comparing systematic and statistical errors, we find that:
-at $\rho=20$, $\lambdans$ measurement is dominated by statistical
-errors; at $\rho=30$, the two become comparable; and 
-for louder signals ($\rho\simeq50$), the systematic errors dominate.
+$\lambdans=\{1000,1500,2000\}$ from left to right. The injection SNR is fixed at
+$\rho=50$. The templates {\it include} tidal effects, with a prior $0\leq\lambdans\leq 4000$.
+% These figures show that this approximation holds up to (SNR)
+% $\rho\simeq 30-50$.
+% 
+In each panel- the dashed red line marks the median {\it measured} value for
+$\lambdans$, and the dashed green line marks its {\it true} value.
+The darker shading shows the $90\%$ confidence interval, whose width
+$(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
+By comparing the measurement uncertainty for these three injections, we see
+that $(\Delta\lambdans)^{90\%}$ grows very slowly with $\lambdans$. Therefore,
+the fractional measurement error - $(\Delta\lambdans)^{90\%}/\lambdans$ -
+decreases monotonically as $\lambdans$ increases (with signal strength fixed).
+% 
+% Comparing systematic and statistical errors, we find that:
+% at $\rho=20$, $\lambdans$ measurement is dominated by statistical
+% errors; at $\rho=30$, the two become comparable; and 
+% for louder signals ($\rho\simeq50$), the systematic errors dominate.
 }
@@
 % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
-\caption{This figure shows the statistical uncertainty in the measurement of
-$\lambdans$, as a percentage of the injected signal's $\lambdans$ value. In each panel,
-the same is shown as a function of the BH mass and spin, keeping the 
-$\lambdans$ and injection's SNR $\rho$ fixed (given in the panel). Each row contains
-panels with the same value of $\lambdans$, with $\rho$ increasing from left to right.
+\caption{Here we show the statistical uncertainty in the measurement of
+$\lambdans$, as a percentage of the injected/true value. In each panel, the
+same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
+injection's SNR $\rho$ fixed (noted in the panel). Each row contains panels
+with the same value of $\lambdans$, with $\rho$ increasing from left to right.
 Each column contains panels with the same value of $\rho$, with $\lambdans$ 
@@
 % 
-Contours are drawn in each panel demarcating regions where we can constrain the
+Contours demarcate regions where we can constrain the
 $\lambdans$ parameter well (within a factor of two of the injected value).
@@
 % 
-\begin{figure*}
+\begin{figure}
 \centering    
-\includegraphics[trim={0 0 0 2cm}, width=0.95\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0.pdf}
-\includegraphics[width=0.95\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0.pdf}
+% \includegraphics[trim={0 0 0 2cm}, width=0.95\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
+\includegraphics[width=\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}
 % \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
-\caption{\red{Correct caption.}
-In these figures we highlight the regions of BH mass and spin plane where we can 
-constrain the $\lambdans$ of its companion NS within a factor of $1-2$ of its true value.
-Panels correspond to different signal strength, with $\rho$ increasing from left to right.
-Within each panel, $\lambdans$ measurement uncertainty contours are shown as functions of BH 
-mass and spin, with line-types corresponding to different uncertainty levels, and
-colors showing the injected $\lambdans$.
+\caption{
+We show here, as a function of BH mass and spin, the {\it minimum} signal
+strength (SNR) required to constrain $\lambdans$ of the companion NS within a factor
+of $100\%$ of its true value, i.e. with $\pm 50\%$ error-bars. The NS mass is fixed
+at $1.35M_\odot$, spin at zero, and $\lambdans=2000$.
+% 
+We can see that, irrespective of BH properties, we cannot constrain $\lambdans$
+to better than $\pm 50\%$ if the SNR is below $\approx 29$.
+% 
 % We note that this information is, in principle, contained in 
@@
 % the effect of NS's deformability itself on its measurement. 
-As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
-of NS matter
-effects improves with all factors that enhances the signature of the NS's disruption on
-the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
-set by the detectors).
-\red{Use same colorbar bounds in both panels}.
+% 
+% As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
+% of NS matter effects improves with all factors that enhances the signature of the NS's disruption on
+% the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
+% set by the detectors).
 }
 \label{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}
-\end{figure*}
-%
-\begin{figure*}
-\centering    
-\includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
-\includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
-\caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}
-In these figures we highlight the regions of BH mass and spin plane 
-where we can constrain the $\lambdans$ of its companion NS within a factor of 
-$1-2$ of its true value. Panels correspond to different signal strength, with 
-$\rho$ increasing from left to right. Within each panel, $\lambdans$ 
-measurement uncertainty contours are shown as functions of BH mass and spin,
-with line-types corresponding to different uncertainty levels, and colors 
-showing the injected $\lambdans$. 
-% We note that this information is, in principle, contained in 
-% Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
-% the effect of NS's deformability itself on its measurement. 
-As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, the 
-measurability of NS tidal deformability improves with all factors that 
-enhances the signature of the NS's disruption on the \textit{detectable} 
-portion of the emitted GW signal (implying, within a frequency band set by the
-detectors).
-}
-\label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
-\end{figure*}
+\end{figure}
+%
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
+% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
+% \caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}
+% \prayush{TRY AND MAKE THIS POINT IN THE FIGURE ABOVE ONLY.}
+% In these figures we highlight the regions of BH mass and spin plane 
+% where we can constrain the $\lambdans$ of its companion NS within a factor of 
+% $1-2$ of its true value. Panels correspond to different signal strength, with 
+% $\rho$ increasing from left to right. Within each panel, $\lambdans$ 
+% measurement uncertainty contours are shown as functions of BH mass and spin,
+% with line-types corresponding to different uncertainty levels, and colors 
+% showing the injected $\lambdans$. 
+% % We note that this information is, in principle, contained in 
+% % Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
+% % the effect of NS's deformability itself on its measurement. 
+% As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, the 
+% measurability of NS tidal deformability improves with all factors that 
+% enhances the signature of the NS's disruption on the \textit{detectable} 
+% portion of the emitted GW signal (implying, within a frequency band set by the
+% detectors).
+% }
+% \label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
+% \end{figure*}
 %
@@
 %
-\red{Move to discussion}
-\prayush{%
-We find that when the BH is sufficiently spinning in the prograde sense,
-with $\chibh\gtrsim +0.7$, even at a relatively possible signal strengths ($\rho=20$),
-we can begin to constraint $\lambdans$ at the level of $\pm 75\%$ errors with a $90\%$
-confidence level. If the signal is stronger, say $\rho=30$, we can contraint
-$\lambdans$ better with only $\pm 50\%$ errors, for binaries with large BH spins.
-This trend continues as we vary the SNR from $\rho=30-50$ upwards. In other words,
-with a single but moderately loud NSBH signal observation, Advanced LIGO will begin
-to put constraints on NS matter properties. These constraints can consequently
-be used to assess the likelihood of different equations of state for nuclear
-matter, and possibly narrow the range they span.
-}
+
   `
-\red{Is this the LEA+ mode? If yes, say so}
-\prayush{%
-We use the same set of systems as in the previous section, with $q=\{2,3,4,5\}$,
-$\chibh=\{-0.5,0,+0.5,+0.75\}$, $\lambdans=\{500, 800, 1000, 1500, 2000\}$ and SNRs
-$\rho=\{20,30,50,70\}$. For each unique combination of these parameters, we obtain a
-signal waveform using the tidal model of Lackey et al~\cite{Lackey:2013axa}.
-We use the same waveform model as input to our Bayesian parameter estimation
-method in order to obtain estimates for different binary parameters along with 
-their full probability distributions, including for the NS tidal
-deformability parameter $\lambdans$.
-}
+% \red{Is this the LEA+ mode? If yes, say so}
+% \prayush{%
+We sample the same set of systems as in the previous section, i.e. those with
+$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and
+$\lambdans=\{500, 800, 1000, 1500, 2000\}$. For each unique combination of
+these parameters, we generate separate injections at SNRs
+$\rho=\{20,30,50,70\}$, using the LEA+ model~\cite{Lackey:2013axa}. We perform a
+full Bayesian analyses on each of these injections using LEA+ templates, and obtain
+probabilistic estimates for source binary parameters from the GW information
+alone, including for the NS tidal deformability parameter $\lambdans$.
 
@@
 
+\red{Move to discussion}
+\prayush{%
+We find that when the BH is sufficiently spinning in the prograde sense,
+with $\chibh\gtrsim +0.7$, even at a relatively possible signal strengths ($\rho=20$),
+we can begin to constraint $\lambdans$ at the level of $\pm 75\%$ errors with a $90\%$
+confidence level. If the signal is stronger, say $\rho=30$, we can contraint
+$\lambdans$ better with only $\pm 50\%$ errors, for binaries with large BH spins.
+This trend continues as we vary the SNR from $\rho=30-50$ upwards. In other words,
+with a single but moderately loud NSBH signal observation, Advanced LIGO will begin
+to put constraints on NS matter properties. These constraints can consequently
+be used to assess the likelihood of different equations of state for nuclear
+matter, and possibly narrow the range they span.
+}
+
 \red{

### 2016-06-12 emcee_PE/scripts/SingleObservation.py

@@ -120,9 +120,9 @@ col corresponds to the lower level group
   if vverbose: print "Making plot with %d rows, %d cols" % (nrow, ncol)
   pltid = 0
   
-  colwidth = 1.3
+  colwidth = 1.75
   fig = plt.figure(int(1e7 * np.random.random()), \
-              figsize=((1.3*gmean*ncol+1.75)*colwidth, 1.2*colwidth*nrow),\
+              figsize=((1.3*gmean*ncol+1.3)*colwidth, 1.2*colwidth*nrow),\
               dpi=100)
   fig.clf()
   grid = ImageGrid(fig, 111,\

### 2016-06-12 Paper/paper.tex
many new figs

@@
 % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
-\includegraphics[width=1.8\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50}
+\includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
 \caption{We illustrate here probability distributions recovered for the NS tidal
@@
 \label{fig:SingleSystemLambdaPDFvsSNR}
+\end{figure*}
+%%
+\begin{figure*}
+\centering 
+\includegraphics[width=0.65\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.65\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.65\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}
+\caption{\prayush{INCOMPLETE..}We illustrate here two-dimensional probability distributions, showing
+the correlation between the NS tidal deformability parameter $\lambdans$ 
+and binary mass and spin parameters.
+}
+\label{fig:SingleSystemLambda2DPDFs}
 \end{figure*}

### 2016-06-12 Paper/paper.tex
adding some nice figs to paper tex

@@
 \centering 
+\includegraphics[width=0.65\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.65\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.65\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
 \includegraphics[width=0.65\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
 \includegraphics[width=0.65\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.65\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.65\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+\includegraphics[width=0.65\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=0.65\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=0.65\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
 \caption{\prayush{INCOMPLETE..}We illustrate here two-dimensional probability distributions, showing

### 2016-06-22 Paper/paper.tex
start of s4

--- 2016-06-12 Paper/paper.tex
@@
 \centering 
-\includegraphics[width=0.65\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.65\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.65\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.62\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.62\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.62\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+\includegraphics[width=0.62\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.62\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.62\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+\includegraphics[width=0.6\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=0.6\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=0.6\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
 \caption{\prayush{INCOMPLETE..}We illustrate here two-dimensional probability distributions, showing
@@
 
-In the previous section we showed that we begin to care for the effects of NS's
-deformation due to its companion hole's tidal field, if (a) the BH is sufficiently
-small, (b) BH spin is aligned and $\gtrsim +0.4$, and (c) the NS is
-not very compact, with $\lambdans\gtrsim 1000$. The first two conditions
-enhance tidal effects on emitted GWs by increasing the number of orbits at small
-separation, where the deviation of NSBH and BHBH dynamics is the largest. The third,
-i.e. (c), simply makes the NS more easily deformable. All of them reduce the onset 
-frequency for NS's disruption close to merger. Therefore, systems that fulfill the 
-above conditions have considerably (or, at least, the most) different GW morphology
-when compared to equivalent BHBH coalescences. Ultimately, the extent of 
-tidal modulations of emitted GWs depends on the equation of state of neutron star
-matter, which becomes an indirect measurable in GW astrophysics.
-% 
-However, if we do indeed receive GWs from such a binary source, what information can we
-possibly extract from it about the internal structure of neutron stars?
-In this section we consider this question in perspective of measuring the tidal 
-deformability parameter of neutron stars purely from gravitational waves they emits 
-while orbiting spinning black holes.
-%
+In the previous section, we showed that the effects of NS's deformation due to
+its companion hole's tidal field becomes discernible in GW spectrum only when
+the following hold: (a) the BH has a sufficiently small mass, (b) BH spin
+is positive aligned, i.e. $\chibh\gtrsim +0.4$, (c) the NS is not very compact,
+with $\lambdans\gtrsim 1000$, and finally, (d) the source is close enough
+such that SNR $\gtrsim 30$.
+% 
+The first two conditions both enhance the tidal distortion of the star and increase
+the number of orbits the system goes through at small separation, where the
+differences between NSBH and BHBH signals is maximal.
+% 
+The third simply makes the NS more easily deformable. All of them also reduce
+the onset frequency of NS's disruption, allowing for it to happen earlier in
+the orbit. 
+% 
+In addition, a SNR of $30$ corresponds to a $1.35+5M_\odot$ binary with 
+$\chibh = 0.4$ at a luminosity distance of $XXX$~Mpc, if optimally aligned
+with LIGO. If, instead, it subtends a non-zero angle with LIGO's line of sight
+(say $45^o$), the binary would need to be at $\approx XXX$~Mpc to achieve the
+same signal strength.
+
+Ultimately, the extent of tidal modulations of emitted GWs depends on the
+equation of state of neutron star interior, which becomes an indirect
+measurable in GW astrophysics.
+% 
+We therefore turn to the question: If we do observe GWs from disruptive 
+coalescing NSBHs, what information can we extract from them about the internal
+structure of neutron stars?
+% 
+We calculate here the accuracy with which we measure $\lambdans$ from a 
+{\it single} GW observation. We will focus on combining multiple observations
+in the following section.
 

### 2016-06-22 Paper/paper.tex
added description of fig.6,7

--- 2016-06-12 Paper/paper.tex
@@
 become observable only at relatively large SNRs~\cite{Lackey:2013axa}, and is validated
-by the results to follow. At design sensitivity, if we expect $10-300$ NSBH detections 
+by results that follow. At design sensitivity, if we expect $10-300$ NSBH detections 
 a year~\cite{Abadie:2010cfa}, and if BH masses are uniformly likely up to
@@
 NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with $\rho\geq 30$, and $0.04-1$ a
-year with $\rho\geq 70$.
+year with $\rho\geq 70$. For our priors, we choose uniform distribution for 
+both component masses as well as for black hole spins, i.e. $\mbh\in[2,20]M_\odot$,
+$\mns\in[1,3]M_\odot$ and $-0.5\leq \chibh\leq +0.75$. \prayush{FIXME}
 
@@
 \centering 
-\includegraphics[width=0.65\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.65\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.65\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}
-\caption{\prayush{INCOMPLETE..}We illustrate here two-dimensional probability distributions, showing
-the correlation between the NS tidal deformability parameter $\lambdans$ 
-and binary mass and spin parameters.
+\includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+\includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+\includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+\includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
+\caption{%
+We illustrate here two-dimensional joint probability distributions, as
+recovered for the intrinsic parameters of a binary with
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
+$\lambdans=2000$, and $\rho=30$ (same as in
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
+The top row shows the same for combinations of 
+non-tidal parameters, i.e. binary masses and spins, while the middle row 
+shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
+is similar to the middle, with the only difference being that the SNR has been
+turned up to $\rho=50$.
 }
@@
 
-In the previous section we showed that we begin to care for the effects of NS's
-deformation due to its companion hole's tidal field, if (a) the BH is sufficiently
-small, (b) BH spin is aligned and $\gtrsim +0.4$, and (c) the NS is
-not very compact, with $\lambdans\gtrsim 1000$. The first two conditions
-enhance tidal effects on emitted GWs by increasing the number of orbits at small
-separation, where the deviation of NSBH and BHBH dynamics is the largest. The third,
-i.e. (c), simply makes the NS more easily deformable. All of them reduce the onset 
-frequency for NS's disruption close to merger. Therefore, systems that fulfill the 
-above conditions have considerably (or, at least, the most) different GW morphology
-when compared to equivalent BHBH coalescences. Ultimately, the extent of 
-tidal modulations of emitted GWs depends on the equation of state of neutron star
-matter, which becomes an indirect measurable in GW astrophysics.
-% 
-However, if we do indeed receive GWs from such a binary source, what information can we
-possibly extract from it about the internal structure of neutron stars?
-In this section we consider this question in perspective of measuring the tidal 
-deformability parameter of neutron stars purely from gravitational waves they emits 
-while orbiting spinning black holes.
-%
-
-  `
-% \red{Is this the LEA+ mode? If yes, say so}
-% \prayush{%
-We sample the same set of systems as in the previous section, i.e. those with
-$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and
-$\lambdans=\{500, 800, 1000, 1500, 2000\}$. For each unique combination of
-these parameters, we generate separate injections at SNRs
-$\rho=\{20,30,50,70\}$, using the LEA+ model~\cite{Lackey:2013axa}. We perform a
-full Bayesian analyses on each of these injections using LEA+ templates, and obtain
-probabilistic estimates for source binary parameters from the GW information
-alone, including for the NS tidal deformability parameter $\lambdans$.
-
-\red{Present posterior distributions of selected configurations first..}
-
-\red{%
-\begin{enumerate}
- \item State the prior on $\lambdans$ because $\lambdans$ weakly deforms waveforms,
-I would expect that measurement uncertainty in $\lambdans$ is approx. independent
-of $\lambdans$, except for running into the $\lambdans\geq 0$ prior.
- \item Is $\lambdans$ measurement degenerate with other parameters? Either cite 
- known studies [if any], or provide 2-d posterior plots.
-\end{enumerate}
-}
+In the previous section, we showed that the effects of NS's deformation due to
+its companion hole's tidal field become discernible in the GW spectrum only when
+the following hold: (a) the BH has a sufficiently small mass, (b) BH spin
+is positive aligned, i.e. $\chibh\gtrsim +0.4$, (c) the NS is not very compact,
+with $\lambdans\gtrsim 1000$, and finally, (d) the source is close enough
+such that SNR $\gtrsim 30$.
+% 
+The first two conditions both enhance the tidal distortion of the star and increase
+the number of orbits the system goes through at small separation, where the
+differences between NSBH and BHBH signals are maximal.
+% 
+The first three also reduce the onset frequency of NS's disruption, allowing
+for it to happen earlier in the orbit. 
+% 
+We expect that these conditions are also the ones which would maximize the
+likelihood of {\it measuring} tidal effects in NSBH signals. We, therefore,
+%
+%In addition, a SNR of $30$ corresponds to a $1.35+5M_\odot$ binary with 
+%$\chibh = 0.4$ at a luminosity distance of $XXX$~Mpc, if optimally aligned
+%with LIGO ($XXX$~Mpc if the binary subtends a $45^o$ angle with LIGO's line of
+%sight.
+%
+%These conditions are strong, but not unrealistic. For e.g., a detection rate
+%of $20$ NSBHs a year implies $1-2$ of them will have SNRs$\gtrsim 30$, since
+%the number of expected events scales as $1/\rho^2$ if we assume that sources
+%are distributed uniformly in spacial volume.
+%
+turn the question around and ask: under similarly favorable circumstances, can
+we gain insights about the internal structure of neutron stars from GW observations.
+
+%
+%Ultimately, the extent of tidal moduilations of emitted GWs depends on the
+%equation of state of neutron star matter, which becomes an indirect
+%measurable in GW astrophysics.
+% 
+%We therefore turn to the question: If we do observe GWs from disruptive 
+%coalescing NSBHs, what information can we extract from them about the internal
+%structure of neutron stars?
+% 
+In order to explore this, we calculate here the accuracy with which we measure
+$\lambdans$ from {\it single} GW observations. We sample the same set of
+disruptive NSBH systems as in the previous section, i.e. those with
+$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$.
+For each unique combination of these parameters, we generate separate injections
+at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model~\cite{Lackey:2013axa}. We
+perform full Bayesian analyses on each of these injections using LEA+ templates,
+and obtain probabilistic estimates for source binary parameters using GW information
+alone (including the NS tidal deformability parameter $\lambdans$). Our priors
+on component masses and spins remain as in the previous section, while $\lambdans$
+is sampled uniformly from $[0,4000]$.
+
+As an illustration, we show the recovered
+probability distribution for $\lambdans$ for three specific configurations in 
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. We fix $q = m_1/m_2 =
+5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, while varying $\lambdans$
+over $\{1000, 1500, 2000\}$ between the three panels. The SNR is fixed at $\rho=50$.
+The true value of $\lambdans$ is given the green (dashed) vertical line,
+while the median recovered value is marked by a red (dashed) line. The darker
+shaded region marks the $90\%$ confidence interval on $\lambdans$. One thing
+to note in this figure is that the measurement uncertainty on $\lambdans$, i.e.
+the full width of its $90\%$ confidence interval, grows slower than $\lambdans$
+as $\lambdans$ is turned up (while the fractional uncertainty 
+$(\Delta\lambdans)^{90\%}/\lambdans$ decreases with increasing $\lambdans$).
+%
+Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
+mass, spin, and tidal parameter measurements. We keep the binary parameters
+as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
+$\rho=30$ (instead of $50$). In the top panel, we show the correlation between
+non-tidal parameters, as manifested in their recovered joint probability
+distributions.
+In the middle row, we show the same between $\lambdans$ and non-tidal 
+parameters, i.e. chirp mass (left panel), mass ratio $\eta$ (middle panel), and
+BH spin $\chibh$ (right). The same is shown in the row below, with the
+difference that the SNR has been turned up to $50$ to bring out the intrinsic
+ambiguity contours.
+%
+We find that the measurement of $\lambdans$ is only weakly degenerate with
+other parameters, and may not improve much if we knew non-tidal parameters
+to better accuracy. The predominant factor that would improve the measurement
+accuracy for $\lambdans$ is the signal strength, SNR.
+
+%\red{Present posterior distributions of selected configurations first..}
+
+%\red{%
+%\begin{enumerate}
+% \item State the prior on $\lambdans$ because $\lambdans$ weakly deforms waveforms,
+%I would expect that measurement uncertainty in $\lambdans$ is approx. independent
+%of $\lambdans$, except for running into the $\lambdans\geq 0$ prior.
+% \item Is $\lambdans$ measurement degenerate with other parameters? Either cite 
+% known studies [if any], or provide 2-d posterior plots.
+%\end{enumerate}
+%}
+
 The precision with which $\lambdans$ can be measured is given by the width of 

### 2016-06-22 Paper/paper.tex
removed fig.10 and discussion -- uninteresting

@@
 \newcommand{\lambdans}{\Lambda_\mathrm{NS}}
-\newcommand{\dlambda}{\delta\Lambda}
+\newcommand{\dlambda}{\delta\lambdans^{90\%}}
 
@@
 We show here, as a function of BH mass and spin, the {\it minimum} signal
-strength (SNR) required to constrain $\lambdans$ of the companion NS within a factor
-of $100\%$ of its true value, i.e. with $\pm 50\%$ error-bars. The NS mass is fixed
-at $1.35M_\odot$, spin at zero, and $\lambdans=2000$.
-% 
-We can see that, irrespective of BH properties, we cannot constrain $\lambdans$
-to better than $\pm 50\%$ if the SNR is below $\approx 29$.
+strength (SNR) required to constrain $\lambdans$ within an interval of width
+equal to $100\%$ of its true value, i.e. with $\pm 50\%$ error-bars. The NS mass
+is fixed at $1.35M_\odot$, spin at zero, and $\lambdans=2000$.
+% 
+We can see that, even in the most conducive circumstances with large aligned 
+$\chibh$ and a comparable mass BH, we can only constrain $\lambdans$ to better
+than $\pm 50\%$ {\it if} the SNR is $\gtrsim 29$. In the era of design
+sensitivity LIGO instruments, we expect this to happen approximately once in a
+year of observation.
 % 
@@
 as $\lambdans$ is turned up (while the fractional uncertainty
-$(\Delta\lambdans)^{90\%}/\lambdans$ decreases with increasing $\lambdans$).
+\begin{equation}
+\delta\lambdans^{90\%}:= (\Delta\lambdans)^{90\%}/\lambdans
+\end{equation}
+decreases with increasing $\lambdans$).
 %
@@
 We find that the measurement of $\lambdans$ is only weakly degenerate with
-other parameters, and may not improve much if we knew non-tidal parameters
+other parameters, and may only marginally improve if we knew non-tidal parameters
 to better accuracy. The predominant factor that would improve the measurement
@@
 
-The precision with which $\lambdans$ can be measured is given by the width of 
-its $90\%$ confidence interval, as a fraction of the true value. 
+%The precision with which $\lambdans$ can be measured is given by the full 
+%width of its $90\%$ confidence interval, as a fraction of the true value. 
 In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the same as a function
@@
 information that we can constrain $\dlambda$ below $200\%$.
-\red{%
-\begin{enumerate}
- \item Inconsistent notation $\delta$ vs $\dlambda$. Unify.
- \item Explain precise definition of $\delta\lambdans/\dlambda$: Is this the 
- full width or half width of the posterior. Such explanation fit in nicely into
- the discussion of the new $\lambdans$ posterior plots like Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}.
-\end{enumerate}
-}
-\hspace{-3mm}Note that being able to constrain $\lambdans$ within a factor of $2$ interval 
+%\red{%
+%\begin{enumerate}
+% \item Inconsistent notation $\delta$ vs $\dlambda$. Unify.
+% \item Explain precise definition of $\delta\lambdans/\dlambda$: Is this the 
+% full width or half width of the posterior. Such explanation fit in nicely into
+% the discussion of the new $\lambdans$ posterior plots like Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}.
+%\end{enumerate}
+%}
+%\hspace{-3mm}
+Note that being able to constrain $\lambdans$ within a factor of $2$ interval 
 (i.e. within $\pm 100\%$) is a fairly modest requirement. We find promising 
@@
 
-\red{%
-(i) Rework text, so that you slowly build up the information needed. First, explain
-the general idea pursued, then discuss one panel first, then introduce second
-panel, finally, contrast both panels.
-(ii) High information density. Distinction between left and right panels notation
-not needed in the first sentence. deferring this info by a few sentences will make
-it easier for the reader to follow the discussion.
-}
-\prayush{%
-Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
-conditions that BH parameters must fulfill beside having an SNR of $20$ (left 
-panel) or $30$ (right panel), to obtain meaningful constraints on $\lambdans$.
-At $\rho=20$, the left panel shows that we need high BH spins 
-$\chibh\gtrsim +0.6$ and a very deformable NS with $\lambdans\geq 1500$, in 
-order to put a factor of $2$ bound on $\lambdans$. At $\rho=30$, with
-$\lambdans\gtrsim 1000$ and $\chibh\geq+0.6$ we can put a similar factor of
-$2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
-NS with $\lambdans=2000$ one could even put a $\pm 50\%$ error-bar on 
-$\lambdans$ measurement.
-}
+%\red{%
+%(i) Rework text, so that you slowly build up the information needed. First, explain
+%the general idea pursued, then discuss one panel first, then introduce second
+%panel, finally, contrast both panels.
+%(ii) High information density. Distinction between left and right panels notation
+%not needed in the first sentence. deferring this info by a few sentences will make
+%it easier for the reader to follow the discussion.
+%}
+%\prayush{%
+%Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
+%conditions that BH parameters must fulfill beside having an SNR of $20$ (left 
+%panel) or $30$ (right panel), to obtain meaningful constraints on $\lambdans$.
+%At $\rho=20$, the left panel shows that we need high BH spins 
+%$\chibh\gtrsim +0.6$ and a very deformable NS with $\lambdans\geq 1500$, in 
+%order to put a factor of $2$ bound on $\lambdans$. At $\rho=30$, with
+%$\lambdans\gtrsim 1000$ and $\chibh\geq+0.6$ we can put a similar factor of
+%$2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
+%NS with $\lambdans=2000$ one could even put a $\pm 50\%$ error-bar on 
+%$\lambdans$ measurement.
+%}
 

### 2016-06-28 Paper/paper.tex
s4 done

@@
 \usepackage{lineno}
+\interfootnotelinepenalty=10000
+
 \newcommand{\roughly}{\mathchar"5218\relax} % Different from \sim in spacing
@@
 
-In the previous section, we showed that the effects of NS's deformation due to
-its companion hole's tidal field become discernible in the GW spectrum only when
-the following hold: (a) the BH has a sufficiently small mass, (b) BH spin
+In the previous section, we showed that the effect of NS's deformation due to
+its companion BH's quadrupolar tidal field becomes discernible in the
+GW spectrum only when the following hold:
+(a) the BH has a sufficiently small mass, (b) BH spin
 is positive aligned, i.e. $\chibh\gtrsim +0.4$, (c) the NS is not very compact,
-with $\lambdans\gtrsim 1000$, and finally, (d) the source is close enough
-such that SNR $\gtrsim 30$.
+with $\lambdans\gtrsim 1000$, {\it and} (d) the source is close enough
+such that its GW SNR $\gtrsim 30$.
 % 
@@
 We expect that these conditions are also the ones which would maximize the
-likelihood of {\it measuring} tidal effects in NSBH signals. We, therefore,
+likelihood of {\it measuring} tidal effects in NSBH signals. In this section,
 %
@@
 %
-turn the question around and ask: under similarly favorable circumstances, can
+we turn the question around and ask: under similarly favorable circumstances, can
 we gain insights about the internal structure of neutron stars from GW observations.
@@
 % 
-In order to explore this, we calculate here the accuracy with which we measure
+We begin by calculating the accuracy with which we measure
 $\lambdans$ from {\it single} GW observations. We sample the same set of
 disruptive NSBH systems as in the previous section, i.e. those with
-$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$.
-For each unique combination of these parameters, we generate separate injections
-at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model~\cite{Lackey:2013axa}. We
+$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
+fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$.
+For each unique combination of these parameters, we generate separate 
+zero-noise injections at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model. We
 perform full Bayesian analyses on each of these injections using LEA+ templates,
@@
 parameters, i.e. chirp mass (left panel), mass ratio $\eta$ (middle panel), and
-BH spin $\chibh$ (right). The same is shown in the row below, with the
-difference that the SNR has been turned up to $50$ to bring out the intrinsic
-ambiguity contours.
+BH spin $\chibh$ (right panel). The same is shown in the row below, with the
+difference that the SNR has been turned up to $\rho=50$, which brings out the
+intrinsic ambiguity contours more clearly.
 %
@@
 
-%\red{Present posterior distributions of selected configurations first..}
-
-%\red{%
-%\begin{enumerate}
-% \item State the prior on $\lambdans$ because $\lambdans$ weakly deforms waveforms,
-%I would expect that measurement uncertainty in $\lambdans$ is approx. independent
-%of $\lambdans$, except for running into the $\lambdans\geq 0$ prior.
-% \item Is $\lambdans$ measurement degenerate with other parameters? Either cite 
-% known studies [if any], or provide 2-d posterior plots.
-%\end{enumerate}
-%}
-
-%The precision with which $\lambdans$ can be measured is given by the full 
-%width of its $90\%$ confidence interval, as a fraction of the true value. 
-In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the same as a function
-of binary parameters and signal strength. Each panel shows the percent statistical
-uncertainty $\dlambda$ in the measurement of $\lambdans$ as a function of BH 
-masses and spins, with NS parameters and signal strength being held fixed. 
-In each row, the signal strength is increased from
-left to right with $\rho = 20\rightarrow 70$. In each column, the deformability
-of the NS is increased from top to bottom, with $\lambdans=500\rightarrow 2000$.
-% 
-We find that at SNRs around $\rho=30$, we begin to constrain $\lambdans$ 
-meaningfully with $\dlambda\approx 100-150\%$. Amongst other parameters, 
-the BH spin and mass play a big role. A smaller BH with a larger spin always
-allows for a more precise measurement on $\lambdans$. Visually we can see that
-in each panel the bottom right corner, which corresponds to low-mass BHs
-with large spins, is the region of smallest measurement errors on $\lambdans$.
+In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the main results
+of this section. In each panel, as a function of black hole mass and spin,
+we show the measured $90\%$ confidence intervals for $\lambdans$, i.e.
+$(\Delta\lambdans)^{90\%}$. These correspond to the full width of the dark
+shaded region in the illustrative figure,
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. The effect of signal strength,
+i.e. SNR, can be seen as we go from left to right in each row. The
+effect of NS's tidal deformability $\lambdans$ on its own measurability
+can be seen by comparing panels within each column. The NS becomes more 
+deformable as we go from top to bottom. In addition, contours of
+$\dlambda=\{50\%,75\%,100\%,150\%,200\%\}$ are also shown in all panels.
+%
+A clear pattern emerges in the left-most column, which corresponds to
+$\rho=20$. We find that at this signal strength, our measurement of 
+$\lambdans$ is dominated by the width of our prior on $\lambdans$. In other
+words, our measured confidence interval is restricted by the
+allowed range of $\lambdans$. Increasing the signal strength to
+$\rho=30$ gives somewhat similar results, as can be seen from the second-from-left
+column of the figure. We do, however, find more encouraging results
+for high BH spins, in which case we can put $\pm 100\%$ to $\pm 75\%$
+bounds on $\lambdans$~\footnote{The symmetric error-bars of
+$\pm\mathrm{X}\%$ correspond to $\dlambda = 2\mathrm{X}\%$.}.
+In addition to large BH spins, this level of
+precision will also require rather deformable neutron stars.
+%
+At higher SNRs, $\rho\gtrsim 50$, we can put stronger constraints
+on $\lambdans$. With a {\it single} observation of a $q=4$ binary with
+$\chibh\geq 0.6$ and $\rho = 50$~\footnote{For an optimally oriented
+source with $q=4, \mns=1.35M_\odot, \chibh=0.6$, an SNR of 
+$\rho = 50$ corresponds to a luminosity distance of $\approx 113$Mpc.},
+we would be able to estimate 
+$\lambdans$ to within $\pm 40\%$ of its true value 
+(which is equivalent to measuring the ratio of its radius to mass
+with an uncertainty of about $\pm 10\%$).
+
+Amongst other source parameters, BH mass and spin play a dominant role. A smaller
+BH with a larger spin always allows for a more precise measurement on $\lambdans$.
+Visually we can see this in the bottom right corner of each panel in
+Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which corresponds to low-mass BHs
+with large spins, and is simultaneously the region of smallest measurement errors on $\lambdans$.
 The actual deformability of the NS also plays an important role on its own
 measurability. For e.g., when $\lambdans\leq 1000$, it is fairly difficult
-to meaningfully constrain $\lambdans$ without getting remarkably lucky and
-finding such a binary with $\rho\gtrsim 50$.
-
-In Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we take the case of
-moderately deformable neutron stars, with actual $\lambdans = 1000$ (left 
-panel) and $1500$ (right panel), and ask how loud such a signal will have
-to be for us to put meaningful constraints on the tidal deformability of NSs.
-In the left panel, as a function BH mass and spin, 
-we show the lowest SNR value $\rho_c$ in order for the signal to have enough
-information that we can constrain $\dlambda$ below $200\%$.
-%\red{%
-%\begin{enumerate}
-% \item Inconsistent notation $\delta$ vs $\dlambda$. Unify.
-% \item Explain precise definition of $\delta\lambdans/\dlambda$: Is this the 
-% full width or half width of the posterior. Such explanation fit in nicely into
-% the discussion of the new $\lambdans$ posterior plots like Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}.
-%\end{enumerate}
-%}
-%\hspace{-3mm}
-Note that being able to constrain $\lambdans$ within a factor of $2$ interval 
-(i.e. within $\pm 100\%$) is a fairly modest requirement. We find promising 
-results, which suggest that there is a fairly wide range of BH spins, less
-wide (but non-trivial) range of BH masses, within
-which if we get a moderately loud NSBH signal with $\rho\approx 23$, we constrain
-$\lambdans$ within a factor of $2$ interval (i.e. within $\pm 100\%$) around
-the true value. While this is not very precise, it can possibly allow us to
-rule out candidates for the equation of state of neutron star matter.
-% 
-If we want to constrain $\lambdans$ better with a single observation, however,
-we need SNRs of $30$ or higher. This can be seen from the right
-panel of Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}, which is similar
-to the left but for a higher $\lambdans (=1500)$ system, with a more stringent
-constraint requirement that $\lambdans$ be measured within a factor of $1$ of 
-its true value (i.e. with $\pm 50\%$ errors, instead of $\pm 100\%$).
-\red{Dont use two different thresholds in figure.} Even for
-the most conducive cases, we {\it need} the SNR to be above $30$ to reach our 
-accuracy threshold (which is more than $3\times$ less likely to happen than
-getting a signal with $\rho=20$).
-
-%\red{%
-%(i) Rework text, so that you slowly build up the information needed. First, explain
-%the general idea pursued, then discuss one panel first, then introduce second
-%panel, finally, contrast both panels.
-%(ii) High information density. Distinction between left and right panels notation
-%not needed in the first sentence. deferring this info by a few sentences will make
-%it easier for the reader to follow the discussion.
-%}
-%\prayush{%
-%Lastly, in Fig.~\ref{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0} we show the
-%conditions that BH parameters must fulfill beside having an SNR of $20$ (left 
-%panel) or $30$ (right panel), to obtain meaningful constraints on $\lambdans$.
-%At $\rho=20$, the left panel shows that we need high BH spins 
-%$\chibh\gtrsim +0.6$ and a very deformable NS with $\lambdans\geq 1500$, in 
-%order to put a factor of $2$ bound on $\lambdans$. At $\rho=30$, with
-%$\lambdans\gtrsim 1000$ and $\chibh\geq+0.6$ we can put a similar factor of
-%$2$ bound on $\lambdans$, but with $\chibh\sim+0.75$ and an extremely deformable
-%NS with $\lambdans=2000$ one could even put a $\pm 50\%$ error-bar on 
-%$\lambdans$ measurement.
-%}
-
-In summary, with a single moderately loud BHNS observation, we could constrain
+to meaningfully constrain $\lambdans$ without requiring the source to be
+close ($\approx 100$Mpc) with a GW SNR $\rho\gtrsim 50$. Quantifying this further,
+in Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we show the minimum
+signal strength required to attain a certain level of confidence in our
+$\lambdans$ measurement, as a function of BH properties. The NS is allowed
+the most favorable (hardest) EoS considered, with $\lambdans^\mathrm{true}=2000$.
+% 
+We first note that, even with the most favorable BH and NS properties, achieving
+a $\pm 50\%$ measurement certainty on $\lambdans$ will require a GW SNR
+$\rho\gtrsim 30$. If we additionally restrict BH masses to lie outside of the so-called
+astrophysical mass-gap~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,
+Littenberg:2015tpa}, we will simultaneously need to restrict BH spins
+to $\chibh\gtrsim +0.5$ to obtain the same measurement confidence at the same
+source location.
+
+In summary, with a single moderately loud ($\rho\lesssim 30$) GW signal from
+a disruptive BHNS coalescence, we can constrain
 the NS compactness parameter $\lambdans$ within $\pm 100\%$ of its true value.
-To measure better with one observation, we will need to get lucky with signals
-with $\rho\geq 30$ and high BH spins, or $\rho\geq 50$.
+To measure better with one observation, we will need a more fine-tuned source, with
+$\rho\geq 30$ and high BH spins, or $\rho\geq 50$.
+% 
+Finally, we note that these results are {\it conservative}, and 
+BHs with spins $\chibh > 0.75$ will prove to be even more favorable laboratories
+for $\lambdans$ measurement. However, we are presently unable to explore this case
+in quantitative detail due to waveform model restrictions~\cite{Lackey:2013axa},
+which will restrict our analyses of actual GW signals as well.
 

### 2016-06-30 Paper/paper.tex
adding script ot sample astrophys pop

@@
 
+\red{Add segway into multiple observations. Point out that all NSs share the same
+$\lambdans(\mns)$, so multiple signals can be combined the measure the common $\lambdans$.}
+
 \textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
+% 
+\red{Simplify presentation: remove (or 16) and (or $\rho=70$) etc. ADd sentence:
+similarly, $16,\rho=10$ events are comparable to one $\rho=70$ event.}
 % 
@@
 
+\dred{%
 The approximation of using (only) the dominant $l=|m|=2$ multipoles of
@@
 distributed in effective volume, i.e. within a sphere of radius 
-$\deff^\mathrm{max}$. The radius of this sphere is set by the lowest SNR that 
+$\deff^\mathrm{max}$.}
+% 
+\red{These sentences are distracting. I suggest to simply say 'sources
+uniform in volume'. Probably not necessary to discuss effective distance
+at all. If so, then do so in a footnote.}
+\red{Explain (or state, if explanation is trivial} that uniform in volume,
+uniform in orientation, is equivalent to uniform in effective volume.}
+% 
+\dred{%
+The radius of this sphere is set by the lowest SNR that 
 is distinguishable from noise by LIGO searches. Now, divide the sphere into
@@
 measurements of $\lambdans$ for identical sources at different distances is
-given by
+given by}
+\red{..into $I$ shells of same radial thickness. The radius at the middle of the}
 \begin{equation}\label{eq:1oversigma}

### 2016-06-30 emcee_PE/scripts/MultipleEventGeneration-N80-20-AstroPop.py

@@ -1,0 +1,1325 @@
+#!/usr/bin/env python
+# Tools for analysis of MCMC samples from emcee for spin estimation project
+# Copyright 2015
+# 
+__author__ = "Prayush Kumar <prayush.kumar@ligo.org>"
+
+import os, sys, time
+import commands as cmd
+import numpy as np
+from matplotlib import mlab, cm, use
+use('Agg')
+from matplotlib.pyplot import *
+import matplotlib.pyplot as plt
+plt.rcParams.update({'text.usetex' : True})
+from mpl_toolkits.axes_grid1 import ImageGrid
+
+
+from pydoc import help
+from scipy import stats
+from scipy.stats.stats import pearsonr
+from scipy.stats import gaussian_kde
+from scipy.stats import norm
+from scipy.interpolate import UnivariateSpline
+from scipy.optimize import minimize_scalar
+import scipy.integrate as si
+
+from sklearn.neighbors import KernelDensity
+from sklearn.grid_search import GridSearchCV
+
+from statsmodels.nonparametric.kde import KDEUnivariate
+from statsmodels.nonparametric.kernel_density import KDEMultivariate
+
+try:
+    sys.path.append( os.path.dirname(os.path.realpath(__file__)) + '/../src/' )
+    sys.path.append( os.path.dirname(os.path.realpath(__file__)) )
+except:
+    sys.path.append( os.path.dirname(cmd.getoutput('pwd -P')) + '/../src/' )
+    sys.path.append( os.path.dirname(cmd.getoutput('pwd -P')) )
+
+
+sys.path.append( '/home/prayush/src/UseNRinDA/scripts/' )
+sys.path.append( '/home/prayush/src/UseNRinDA/plotting' )
+sys.path.append( '/home/prayush/src/nsbh_tidal/emcee_PE/scripts/' )
+sys.path.append( '/home/prayush/src/nsbh_tidal/emcee_PE/src/' )
+
+from utils import *
+import pycbc.pnutils as pnutils
+from PlotOverlapsNR import make_contour_plot_multrow
+import h5py
+
+import MultipleObservations as MO
+#reload(MO)
+
+# Constants
+verbose  = True
+vverbose = True
+debug    = True
+
+quantiles_68 = [0.16, 0.5, 0.84] # 1-sigma
+quantiles_95 = [0.0228, 0.5, 0.9772] # 2-sigma ~ 95.4%
+percentiles_68 = 100*np.array(quantiles_68)
[…]

### 2016-06-30 emcee_PE/scripts/MultipleObservations.py

@@ -147,6 +147,19 @@ output:-
     #
     return [llimit, median, ulimit]
 
+
+def find_closest_match(tmp_rand, vec):
+    '''
+Take a random number in the range spanned by vec, and find the
+element in vec closest to it.
+    '''
+    diffs = np.abs(np.array(vec) - tmp_rand)
+    idx_min = np.where(diffs == diffs.min())[0][0]
+    if idx_min < 0 or idx_min >= len(vec):
+        raise RuntimeError("Could not find closest match!")
+    return vec[idx_min]
+    
+
 def sample_snr_from_volume(SNRvec):
     '''
 Sample the SNR of signals, as if they are distributed uniformly in
@@ -164,19 +177,124 @@ the observable volume, i.e. 4/3 pi D_Luminosity^3
         if cumsum > rnd_tmp: return SNRvec[i]
     #
     raise RuntimeError("SNR sampling not being done properly")
-        
 
-def find_closest_match(tmp_rand, vec):
+
+def sample_angles_on_S2():
+    """
+Samples a point uniformly from a 2-sphere. Returns its spherical angles.
+first -- polar angle
+second-- azimuthal angle
+    """
+    # First sample in a unit cube, with origin at centroid
+    r = 100
+    while r > 1:
+        x = np.random.uniform(-1, 1)
+        y = np.random.uniform(-1, 1)
+        z = np.random.uniform(-1, 1)
+        r = (x**2 + y**2 + z**2)**0.5
+    x /= r
+    y /= r
+    z /= r
+    #
+    theta = np.arccos(z) # polar angle     (w.r.t. z-axis)
+    if np.isnan(theta): print "Theta is NAN for x, y, z, r, z/r = %f,%f,%f,%f,%f" % (x,y,z,r,z/r)
+    phi   = np.arctan2(y, x) # azimuthal angle (in x-y plane)
+    #
+    return [theta, phi]
+
+
+def sample_snr_from_volume_inclination(SNRvec, q, chi,\
+                        snr_threshold_low=10, snr_threshold_high=100,\
+                        time_length = 256, sample_rate = 16384, f_lower = 15,\
+                        psd = None, verbose=False, vverbose=False):
     '''
-Take a random number in the range spanned by vec, and find the
-element in vec closest to it.
+Sample the SNR of signals, as if they are distributed uniformly in
+the spacial volume, i.e. 4/3 pi D^3, and simultaneously have 
[…]

### 2016-06-30 emcee_PE/src/utils.py

@@ -13,7 +13,7 @@ where rho is the desired snr.
 
 __author__ = "Michael Puerrer, Prayush Kumar"
 __copyright__ = "Copyright 2015"
-__email__ = "Michael.Puerrer@ligo.org"
+__email__ = "Michael.Puerrer@ligo.org; prayush.kumar@ligo.org"
 
 import sys,os
 import lal, lalsimulation
@@ -36,6 +36,24 @@ import scipy.interpolate as ip
 from match import *
 from injection import *
 
+
+################################################################################
+################################################################################
+def Fplus(theta, phi, psi):
+  """
+LIGO's sensitivity function, Eq. 57 of http://relativity.livingreviews.org/Articles/lrr-2009-2/download/lrr-2009-2Color.pdf
+  """
+  t1 = (1. + np.cos(theta)**2) * np.cos(2 * phi) * np.cos(2 * psi) / 2
+  t2 = np.cos(theta) * np.sin(2 * phi) * np.sin(2 * psi)
+  return t1 - t2
+
+def Fcross(theta, phi, psi):
+  """
+LIGO's sensitivity function, Eq. 57 of http://relativity.livingreviews.org/Articles/lrr-2009-2/download/lrr-2009-2Color.pdf
+  """
+  t1 = (1. + np.cos(theta)**2) * np.cos(2 * phi) * np.sin(2 * psi) / 2
+  t2 = np.cos(theta) * np.sin(2 * phi) * np.cos(2 * psi)
+  return t1 + t2
 
 ################################################################################
 # Function to compute percentile intervals from samples

### 2016-06-30 emcee_PE/scripts/MultipleObservations.py
increasing accuracy of kde

@@ -118,7 +118,7 @@ output:-
             return 0
         elif L > xulimit:
             L = xulimit
-        return si.quad(kde_func, xllimit, L, epsabs=1.49e-12, epsrel=1.49e-12)[0] / denom
+        return si.quad(kde_func, xllimit, L, epsabs=1.49e-16, epsrel=1.49e-16)[0] / denom
     #
     def tmp_rootfunc(L, eps): return np.abs(tmp_intfunc(L) - eps)
     def tmp_medianfunc(L): return np.abs(tmp_rootfunc(L, 0.5))
@@ -473,7 +473,7 @@ What it does:
             ## Normalize the chain's integral
             xllimit, xulimit = [0, 4000]
             lambda_kde_norm = si.quad(lambda_kde.evaluate, xllimit, xulimit,\
-                                        epsabs=1.e-12, epsrel=1.e-12)[0]
+                                        epsabs=1.e-16, epsrel=1.e-16)[0]
             #lambda_gaussian_kde = gaussian_kde(lambda_chain)
             #lambda_multivariate_kde = KDEMultivariate(lambda_chain, bw=0.2*np.ones_like(lambda_chain), var_type='c')
             #lambda_skl_kde = KernelDensity(bandwidth=0.2)
@@ -575,7 +575,7 @@ by the constructor.
             if precalculate_norms:
                 xllimit, xulimit = [0, 4000]
                 lambda_kde_norm = si.quad(lambda_kde.evaluate, xllimit, xulimit,\
-                                            epsabs=1.e-12, epsrel=1.e-12)[0]
+                                            epsabs=1.e-16, epsrel=1.e-16)[0]
             else: lambda_kde_norm = None
             #lambda_gaussian_kde = gaussian_kde(lambda_chain)
             #lambda_multivariate_kde = KDEMultivariate(lambda_chain, bw=0.2*np.ones_like(lambda_chain), var_type='c')

### 2016-07-01 Paper/paper.tex
simple calc done

--- 2016-06-28 Paper/paper.tex
@@
 
+\red{SAY HERE THAT WE ONLY USE THE $l=|m|=2$ modes.}
+% 
 \subsection{Waveform Models}\label{s2:waveforms}
@@
 %
-We find that the measurement of $\lambdans$ is only weakly degenerate with
-other parameters, and may only marginally improve if we knew non-tidal parameters
-to better accuracy. The predominant factor that would improve the measurement
-accuracy for $\lambdans$ is the signal strength, SNR.
+We find that the measurement of $\lambdans$ is weakly degenerate with
+other parameters, and at realistic SNRs it would improve by a few tens of 
+percent if we knew non-tidal parameters to better accuracy. The predominant
+factor that would enhance the measurement accuracy for $\lambdans$ is the
+signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
+measurement to be limited by its degeneracy with non-tidal parameters at 
+a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
 
@@
 
+In the previous section, we showed that single observations of NSBH coalescences
+at moderate SNRs have little information about the internal structure of neutron
+stars that will be accessible to Advanced LIGO at its design sensitivity. We
+expect all neutron stars to share the same equation of state, and hence the same
+$\lambdans(\mns)$. In addition, we know that the mass distribution of (most) NSs
+that have not been spun up to millisecond periods (which are what we focus on in 
+this paper, by approximating $\chins\approx 0$) is narrowly peaked around
+$\sim 1.35M_\odot$~\cite{Kiziltan2013}. Therefore, information from multiple NSBH
+observations can be combined to improve our estimation of $\lambdans$ for the
+population. This is what we explore in this section, within a fully Bayesian 
+framework.
+
 \textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
 % 
-We start with a crude calculation which tells us that approximately $8$ (or $16$) 
-observations
-of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
-give us similarly accurate a measurement of $\lambdans$ as would a single detection
-with $\rho=50$ (or $\rho=70$).
-
-The approximation of using (only) the dominant $l=|m|=2$ multipoles of
-gravitational-wave strain to construct the signal, ignoring other $l\neq 2$
-multipoles, has the advantage of making various angle parameters 
-associated with the source degenerate with its distance. As a result, source
-distance, source sky location angles, and the inclination angle of
-the binary's orbital angular momentum with respect to the detector,
-all combine into a single ``effective distance'' $\deff$ that scales
-out as the constant amplitude factor $1/\deff$ (constant because none
-of the above parameters vary temporally for aligned-spin binaries).
-% 
-In order to consider multiple events, lets imagine a population uniformly
-distributed in effective volume, i.e. within a sphere of radius 
-$\deff^\mathrm{max}$. The radius of this sphere is set by the lowest SNR that 
-is distinguishable from noise by LIGO searches. Now, divide the sphere into
-$I$ shells of equal thickness. The radius of the $i$-th sphere would then
-be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The measurement uncertainty in
-$\lambdans$ scales inversely with SNR, and hence directly with the 
-effective distance to the source. Therefore, if we have a measurement error 
-$\sigma_0$ for a source located at $\deff = D_0$, the same error for the same
-source located within the $i$-th shell would be 
-$\sigma_i = \sigma_0 \dfrac{D_i}{D_0}$. The combined error from independent
-measurements of $\lambdans$ for identical sources at different distances is
-given by
-\begin{equation}\label{eq:1oversigma}
-\frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
-\end{equation}
-where $N_i$ is the number of sources detected in the $i$-th shell.
-% , and is a random
-% variable, with its probability proportional to the volume of the shell, i.e.
-% $$
-% p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
-% $$
-% Therefore, the expected number of detections in the $i$-th shell would be
-% $$
-% \langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
-% $$
-% where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
-% to be Poisson distributed around the mean detection rate
-% $\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
-% $\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
-% a year per unit effective volume, then
-% \begin{equation}
-% \langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
+An intuitive understanding of the problem is gained by considering first
+multiple {\it identical} sources at realistic SNRs. Let us consider the case
+a population of optimally oriented binaries~\footnote{An optimally oriented
+binary is one which is located directly overhead the detector, with the 
+orbital angular momentum parallel to the line joining the detector to the
+source. Such a configuration maximizes the observed GW signal strength in 
+the detector.}, distributed uniformly in spacial volume out to a maximum
+distance $D^\mathrm{max}$. $D^\mathrm{max}$ is set by the minimum SNR 
+threshold at which a source is considered detectable~\footnote{which
+we take as $\rho_\mathrm{min}=10$ throughout.}. Next, we divide this volume into $I$
+concentric shells, with radii $D_i$. If we have a measurement error
+$\sigma_0$ for $\lambdans$, associated with a source located at $D=D_0$,
+the same error for the same source located within the $i-$th shell would
+be $\sigma_i=\sigma_0 \dfrac{D_i}{D_0}$. Ref.~\cite{Markakis:2010mp}
+calculated that the combined error $\sigma$ from $N$ independent
+measurements of $\lambdans$ in such a setting to be
+% 
+\begin{align}\label{eq:1oversigma}
+\frac{1}{\sigma^2} =& \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2}\\ \nonumber =& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{D^\mathrm{max}} \dfrac{4\pi D^2 n}{D^2}\D D = \left(\frac{D_0}{\sigma_0}\right)^2 \dfrac{3N}{(D^\mathrm{max})^2},
+\end{align}
+% 
+where $N_i$ is the number of sources within the $i-$th shell (such that
+$N:=\sum N_i$), and $n$ is the number density of source distribution in volume.
+% 
+% \begin{equation}\label{eq:rmsSigmaIdenticalSources}
+%  \sigma_{avg} := \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}},
 % \end{equation}
-% and
-% \begin{eqnarray}
-%  \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
-%  &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
-% \end{eqnarray}
-% where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
-% to an integral. 
-Continuing on this line of inquiry, Ref.~\cite{Markakis:2010mp} found the 
-root-mean-square averaged measurement error from 
-$\langle N\rangle:=\langle\sum N_i\rangle$ sources distributed uniformly in
-effective volume within a sphere of radius $\deff^\mathrm{max}$ to be
+% 
+% 
+% 
+% 
+% \prayush{%
+% The approximation of using (only) the dominant $l=|m|=2$ multipoles of
+% gravitational-wave strain to construct the signal, ignoring other $l\neq 2$
+% multipoles, has the advantage of making various angle parameters 
+% associated with the source degenerate with its distance. As a result, source
+% distance, source sky location angles, and the inclination angle of
+% the binary's orbital angular momentum with respect to the detector,
+% all combine into a single ``effective distance'' $\deff$ that scales
+% out as the constant amplitude factor $1/\deff$ (constant because none
+% of the above parameters vary temporally for aligned-spin binaries).
+% In order to consider multiple events, lets imagine a population uniformly
+% distributed in effective volume, i.e. within a sphere of radius 
+% $\deff^\mathrm{max}$.
+% }
+% % 
+% \red{These sentences are distracting. I suggest to simply say 'sources
+% uniform in volume'. Probably not necessary to discuss effective distance
+% at all. If so, then do so in a footnote.}
+% \red{Explain (or state, if explanation is trivial) that uniform in volume,
+% uniform in orientation, is equivalent to uniform in effective volume.}
+% % 
+% \prayush{%
+% The radius of this sphere is set by the lowest SNR that 
+% is distinguishable from noise by LIGO searches. Now, divide the sphere into
+% $I$ shells of equal thickness. The radius of the $i$-th sphere would then
+% be $D_i = \deff^\mathrm{max} (i - 1/2)/I$.}
+% \red{..into $I$ shells of same radial thickness. The radius at the middle of
+% the $i-$th shell $(i=1,2,...,I)$ is $D_i:=\frac{i-\frac{1}{2}}{I}\,D_\mathrm{max}$,
+% and the number of sources in this shell is $N_i=4\pi\beta D_i^2 \frac{D_\mathrm{max}}{I}$,
+% where $\beta$ is the volume density of sources.
+% }
+% The measurement uncertainty in
+% $\lambdans$ scales inversely with SNR, and hence directly with the 
+% effective distance to the source. Therefore, if we have a measurement error 
+% $\sigma_0$ for a source located at $\deff = D_0$, the same error for the same
+% source located within the $i$-th shell would be 
+% $\sigma_i = \sigma_0 \dfrac{D_i}{D_0}$. The combined error from independent
+% measurements of $\lambdans$ for identical sources at different distances is
+% given by
+% % 
+% \begin{equation}\label{eq:1oversigma}
+% \frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
+% \end{equation}
+% where $N_i$ is the number of sources detected in the $i$-th shell.
+% % , and is a random
+% % variable, with its probability proportional to the volume of the shell, i.e.
+% % $$
+% % p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
+% % $$
+% % Therefore, the expected number of detections in the $i$-th shell would be
+% % $$
+% % \langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
+% % $$
+% % where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
+% % to be Poisson distributed around the mean detection rate
+% % $\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
+% % $\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
+% % a year per unit effective volume, then
+% % \begin{equation}
+% % \langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
+% % \end{equation}
+% % and
+% % \begin{eqnarray}
+% %  \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
+% %  &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
+% % \end{eqnarray}
+% % where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
+% % to an integral. 
+The root-mean-square (RMS) averaged measurement error from $N$ sources is 
+then~\cite{Markakis:2010mp}
 \begin{equation}\label{eq:rmsSigmaIdenticalSources}
- \sigma_{avg} := \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}},
+ \sigma_{avg} := \frac{1}{\sqrt{1/\sigma^{2}}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3 N}},
 \end{equation}
 given a fiducial pair $(\sigma_0, D_0)$. It is straightforward to deduce from
-Eq.~\ref{eq:rmsSigmaIdenticalSources}
-that the same measurement certainty as afforded by a single
-observation with a high SNR $\rho_c$ can be obtained from $N = \rho_c^2/300$ observations
-uniformly distributed in effective volume with SNRs $\geq 10$. E.g., to get to the level 
-of certainty afforded at SNR$=70$, we would need $49/3\approx 16-17$ realistic detections.
-
-There are two main caveats to this rudimentary calculation, first outlined in 
-Ref.~\cite{Markakis:2010mp}: (i) it applies to identical sources,
-which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
-approximately; (ii) the combined average measurement error $\sigma_{avg}$ has been defined
-as in the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}. To overcome both,
-we next perform a Bayesian population analysis.
+Eq.~\ref{eq:rmsSigmaIdenticalSources} that measurement uncertainty scales as 
+$1/\sqrt{N}$, and the uncertainty afforded by a single observation with a high
+SNR $\rho_c$ can be attained with $N = \rho_c^2/300$ realistic observations
+that have $\rho\geq\rho_\mathrm{min}$. E.g., to get to the
+level of certainty afforded by a single observation with $\rho=70$, we would
+need $49/3\approx 16-17$ realistic (low SNR) detections.
+
+% There are two main caveats to this rudimentary calculation, first outlined in 
+% Ref.~\cite{Markakis:2010mp}: (i) it applies to identical sources,
+% which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
+% approximately; (ii) the combined average measurement error $\sigma_{avg}$ has been defined
+% as in the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}. To overcome both,
+% we next perform a Bayesian population analysis.
+%
 

### 2016-07-01 Paper/paper.tex
simple calc more

@@
 
-\textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
+% \textbf{Multiple identical sources at low SNR: }
+\label{s2:identical_multiple}
 % 
@@
 distance $D^\mathrm{max}$. $D^\mathrm{max}$ is set by the minimum SNR 
-threshold at which a source is considered detectable~\footnote{which
+threshold $\rho_\mathrm{min}$ at which a source is considered
+detectable~\footnote{which
 we take as $\rho_\mathrm{min}=10$ throughout.}. Next, we divide this volume into $I$
@@
 % 
-where $N_i$ is the number of sources within the $i-$th shell (such that
-$N:=\sum N_i$), and $n$ is the number density of source distribution in volume.
+where $N_i$ is the number of sources within the $i-$th shell (s.t.
+$N:=\sum N_i$), and $n$ is the number density of sources in volume.
 % 
@@
 
+While we discussed Eq.~\ref{eq:rmsSigmaIdenticalSources} for a population
+of optimally oriented sources, it is valid more generally if the population is
+distributed uniformly in {\it effective} volume~\footnote{{\it effective} 
+distance $D_\mathrm{eff}$ 
+is a combination of distance to the source, its orientation, and its sky
+location angles, and has a one-to-one correspondence with SNR for non-precessing
+systems, for which these angles are constant in time. Effective volume is 
+proportional to the $D_\mathrm{eff}^3$}~\cite{Markakis:2010mp}. However, it 
+still only applies to sources with identical masses and spins, and we 
+overcome this limitation by performing a fully Bayesian analysis next.
+
 % There are two main caveats to this rudimentary calculation, first outlined in 

### 2016-07-01 Paper/paper.tex
s6 prelim done, need new figs

@@
 \red{SAY HERE THAT WE ONLY USE THE $l=|m|=2$ modes.}
+\prayush{
+This is possible because of our simplification of 
+using only the dominant $l=|m|=2$ modes of the GW strain~\footnote{The 
+$l=|m|=2$ multipoles of the GW strain, when decomposed in a spin $-2$ weighted
+spherical harmonic basis, contain more than $99\%$ of the total signal power}
+that makes various angle parameters completely degenerate with distance. 
+}
 % 
@@
 
-% \textbf{Multiple identical sources at low SNR: }
-\label{s2:identical_multiple}
+% \textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
 % 
@@
 While we discussed Eq.~\ref{eq:rmsSigmaIdenticalSources} for a population
-of optimally oriented sources, it is valid more generally if the population is
+of optimally oriented sources, it is valid more generally for a population
 distributed uniformly in {\it effective} volume~\footnote{{\it effective} 
@@
 systems, for which these angles are constant in time. Effective volume is 
-proportional to the $D_\mathrm{eff}^3$}~\cite{Markakis:2010mp}. However, it 
+$\frac{4\pi}{3}D_\mathrm{eff}^3$.}~\cite{Markakis:2010mp}. However, it 
 still only applies to sources with identical masses and spins, and we 
@@
 
-% There are two main caveats to this rudimentary calculation, first outlined in 
-% Ref.~\cite{Markakis:2010mp}: (i) it applies to identical sources,
-% which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
-% approximately; (ii) the combined average measurement error $\sigma_{avg}$ has been defined
-% as in the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}. To overcome both,
-% we next perform a Bayesian population analysis.
-%
-
 \textbf{Astrophysical source population: }\label{s2:astro_multiple}
@@
 containing a single signal emitted by an NSBH binary. Each of these signals can
-be characterized by the following binary parameters:
-$\,\vec{\theta} = \{\mbh, \mns, \chibh, \chins, \rho\}\cup\{\lambdans\}$. Note
-that we have folded in the luminosity distance, inclination angle of the 
-binary's orbital angular momentum with respect to the plane of the detector,
-and source's sky location angles into a single parameter $\rho$, which is the 
-SNR of the received signal. This is possible because of our simplification of 
-using only the dominant $l=|m|=2$ modes of the GW strain~\footnote{The 
-$l=|m|=2$ multipoles of the GW strain, when decomposed in a spin $-2$ weighted
-spherical harmonic basis, contain more than $99\%$ of the total signal power}
-that makes various angle parameters completely degenerate with distance. 
+be characterized by the non-tidal source parameters
+$\vec{\theta} := \{\mbh, \mns, \chibh, \chins, \vec{\alpha}\}$,
+and $\{\lambdans\}$, where $\vec{\alpha}$ contains extrinsic parameters,
+such as source distance, inclination, and sky location angles.
+% 
 Further, let $K$ denote all of our collective prior knowledge, except for 
@@
 % 
-Using Bayes' theorem, the measured probability distribution function for 
-$\lambdans$ from $N$ observations is given by~\footnote{%
-Eq.~\ref{eq:lambdaMultiple} includes the following implicit factor to normalize
-its right-hand side: $\dfrac{\prod_i p(d_i|K)}{\int p(\lambdans)\,p(d_1, d_2, \cdots, d_N | \lambdans; K)\,\D\lambdans}$}
+The probability distribution for $\lambdans$, given $N$ unique and
+independent events, is
+% 
+\begin{eqnarray}
+ p(\lambdans |&& \hspace{-4mm}d_1, d_2, \cdots, d_N, K)\hspace{50mm}\nonumber\\ &=& \dfrac{p(d_1,d_2,\cdots,d_N |\lambdans , K)\,p(\lambdans|K)}{\int p(\lambdans |K) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans},\label{eq:p11}\\
+  &=& \dfrac{p(\lambdans|K) \prod_{i} p(d_i|\lambdans, K)}{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans},\label{eq:p12} \\
+  &=& \dfrac{p(\lambdans|K) \prod_{i} \left( p(\lambdans |d_i, K)\dfrac{p(d_i)}{p(\lambdans|K)} \right)}{\int\, p(\lambdans|K )\, p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans}\label{eq:p13};
+\end{eqnarray}
+% 
+where Eq.~\ref{eq:p11} and Eq.~\ref{eq:p13} are application of Bayes' theorem,
+while Eq.~\ref{eq:p12} comes from the mutual independence of all events.
+Assuming in addition that all events are {\it equally likely}: 
+$p(d_i) = p(d_j) = p(d)$, we get
+% 
+\begin{eqnarray}
+ p(&&\hspace{-4mm}\lambdans | d_1, d_2, \cdots, d_N, K)\hspace{50mm}\nonumber\\
+  &=& \dfrac{p(\lambdans) \left(\dfrac{p(d)}{p(\lambdans)}\right)^N\prod_{i} p(\lambdans |d_i, K) }{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans}\label{eq:p21},\\
+  &=& p(\lambdans)^{1-N}\times \dfrac{p(d)^N}{\int p(\lambdans) p(d_1, d_2, \cdots, d_N |\lambdans, K)\D\lambdans}\nonumber\\ &&\hspace{3mm}\times\prod_i p(\lambdans |d_i, K)\label{eq:p22}.
+\end{eqnarray}
+% 
+The first two factors in Eq.~\ref{eq:p22} can be absorbed into a
+normalization factor $\mathcal{N}$, simplifying it to
+% 
 \begin{equation}\label{eq:lambdaMultiple}
- p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{1-N}\prod_{i=1}^N p(\lambdans | d_i, K),
+ p(\lambdans | d_1, d_2, \cdots, d_N; K) = \mathcal{N}\prod_{i=1}^N p(\lambdans | d_i, K).
 \end{equation}
-where $p(\lambdans|K)$ is the prior expectation for $\lambdans$, and all $N$ 
-observations are taken to be mutually independent. A priori, we assume
-that no particular value of $\lambdans$ is preferred over another, and 
-$\lambdans\leq 4000$, i.e.
+% 
+% 
+% % 
+% Using Bayes' theorem, the measured probability distribution function for 
+% $\lambdans$ from $N$ observations is given by~\footnote{%
+% Eq.~\ref{eq:lambdaMultiple} includes the following implicit factor to normalize
+% its right-hand side: $\dfrac{\prod_i p(d_i|K)}{\int p(\lambdans)\,p(d_1, d_2, \cdots, d_N | \lambdans; K)\,\D\lambdans}$}
+% \begin{equation}\label{eq:lambdaMultiple}
+%  p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{1-N}\prod_{i=1}^N p(\lambdans | d_i, K),
+% \end{equation}
+A priori, we assume that no particular value of $\lambdans$ is preferred
+over another, and restrict $\lambdans$ to $[0, 4000]$, i.e.
 \begin{equation}\label{eq:lprior}
@@
 $p(\lambdans | d_i, K)$), each is the probability distribution for $\lambdans$
-inferred {\it a posteriori} from the \textit{i}-th observation in isolation. 
-They are calculated by integrating
+inferred {\it a posteriori} from the \textit{i}-th observation by marginalizing
 \begin{equation}\label{eq:margpost}
- p(\lambdans | d_i, K) = \int \D \vec{\theta}\, p(\vec{\theta}, \lambdans | d_i, K),
+ p(\lambdans | d_i, K) = \int\, p(\vec{\theta}, \lambdans | d_i, K)\, \D \vec{\theta},
 \end{equation}
-where $p(\vec{\theta}, \lambdans | d_n, K)$ is the inferred joint probability 
+where $p(\vec{\theta}, \lambdans | d_i, K)$ is the inferred joint probability 
 distribution of all source parameters $\vec{\theta}\cup\{\lambdans\}$ for the 
-$i$-th event, as given by Eq.~\ref{eq:postprob}. By substituting
+$i$-th event, as given by Eq.~\ref{eq:postprob}. We note that 
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR} illustrates $p(\lambdans | d_i, K)$
+for three individual events. By substituting
 Eq.~\ref{eq:lprior}-\ref{eq:margpost} into Eq.~\ref{eq:lambdaMultiple}, we
@@
 Our goal is to understand the improvement in our measurement of $\lambdans$
-with the number of recorded events. To do so, we simulate populations of $N$
-observations each, and quantify what we learn from each successive observation 
-using Eq.~\ref{eq:lambdaMultiple}. For each population, we determine how
+with the number of recorded events. To do so, we simulate a population~\footnote{%
+A population here is an ordered set of events, and an event itself is the 
+set of parameters describing one astrophysical NSBH binary.}
+of $N$ events, and quantify what we learn from each successive observation 
+using Eq.~\ref{eq:lambdaMultiple}. This allows us to quantify how
 rapidly our median estimate for $\lambdans$ converges to the true value,
 and how rapidly our confidence intervals for the same shrink, with increasing
-$N$. In order to generate each population, we first fix
+$N$. Finally, we generate and analyze an ensemble of populations in order to
+average over the stochastic process of population generation itself.
+
+In order to generate each population, the first step is to fix
 the NS properties: (i) NS mass $\mns=1.35M_\odot$, (ii) NS spin $\chins=0$
 and (iii) NS tidal deformability $\lambdans=$ fixed value chosen from
-$\{500,800,1000,1500,2000\}$. Each event is generated independently, sampling
-the remaining source parameters uniformly from the following ranges: (i)
-mass-ratio $q\in[2,5]$, (ii) BH spin $\chibh\in[0, 0.75]$, (iii) effective 
-distance from within a spherical volume whose radius is given by the lowest 
-detectable SNR (which we take as $\rho=10$). We remind ourselves that
-once binary parameters are fixed, there is a one-to-one mapping between
-effective distance and SNR. Additionally, we consider an alternate population
-paradigm (paradigm B, with paradigm A being the standard one described 
-above), where populations are generated as above but further restricting
-mass-ratios such that source BH masses lie {\it outside} the astrophysical 
-mass-gap~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}.
-Finally, in order to mitigate computational cost, we make an additional 
-simplification. We perform full Bayesian parameter estimation for a set of 
-simulated signals whose parameters are the vertices of the following regular
-hypercubic grid (henceforth ``G''):
-$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
-The final population is given by substituting all of the $N$ events, generated
-as above, with their respective nearest neighbours on the parameter grid G.
+$\{500,800,1000,1500,2000\}$. Next, we generate events, by sampling BH mass
+(uniformly) from $\mbh\in[3M_\odot,6.75M_\odot]$, BH spin (uniformly)
+from $\chibh\in[0, 0.75]$, source location uniformly in spatial volume,
+and its inclination from $\iota\in[0,\pi]$. \prayush{Note that we only
+sample positive aligned BH spins, since binaries with anti-aligned spins have
+very little information to add at realistic SNRs, as shown in 
+Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}.
+}
+As noted previously, we also restrict events to have
+$\rho\geq\rho_\mathrm{min}$, where $\rho_\mathrm{min}=10$.
+Finally, we repeat the process till we have an ordered set of $N$ events.
+% 
+Since we want to analyze not just a single realization of an astrophysical
+population, but an ensemble of them, we make an additional approximation to
+mitigate computational cost. Complete Bayesian parameter estimation is
+performed for a set of simulated signals whose parameters are the vertices
+of a regular hypercubic grid (henceforth ``G'') in the space of
+$\{q\}\times\{\chibh\}\times\{\rho\}$, with each sampled at $q=\{2,3,4,5\}$,
+$\chibh=\{-0.5,0,0.5,0.75\}$, and $\rho=\{10,20,30,50,70\}$.
+All events in each population draw are substituted by their
+respective nearest neighbours on the grid G.
 
 In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show illustrative results
-for a single 
-population of $N=80$ events. The neutron star deformability is fixed at
-$\lambdans=800$ for all events. In the left panel, each curve shows the 
+for a single population with neutron star deformability $\lambdans=800$.
+In the left panel, each curve shows the 
 probability distributions for $\lambdans$ as inferred from $N$ events, with $N$
 ranging from $1-80$. We also mark the $90\%$ confidence intervals associated
-with each of the probability distribution curves. From the figure, we
-can visualize information accumulation. The first few observations do not have
-enough information to bound $\lambdans$ (with $90\%$ confidence) much more than
+with each of the probability distribution curves. The first few observations
+do not have
+enough information to bound $\lambdans$ much more than
 our prior from Eq.~\ref{eq:lprior} does. However, the median of these
@@
 In the right panel, we present information derived from the left panel.
-The line-circle curve shows the measured median from $N$ observations, with
-the latter on the $x$-axis. The pair of dashed (dotted) horizontal lines mark
+The line-circle curve shows the measured median from $N$ observations.
+The pair of dashed (dotted) horizontal lines mark
 $\pm25\%$ ($\pm50\%$) error bars. At each $N$, the range spanned by the 
@@
 % 
-In Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0} we show further results from $4$
-independent populations, with the source $\lambdans$ alone changing between
-them. The sampled values, i.e.$\lambdans=\{500,1000,1500,2000\}$, 
-range from the least to the most deformable neutron stars. As in the 
+In Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0} we show further results from four
+independent populations for $\lambdans=\{500,1000,1500,2000\}$. As in the 
 right panel of Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}, the line-circle curves
@@
 the following: (i) the shrinkage of confidence interval widths with increasing
-$N$ happens in a similar manner for all 
-populations, (ii) within $5-10$ observations, median measured $\lambdans$ for
-all populations lie within $\pm10\%$ of the true values, and (iii) it takes
+$N$ happens in a similar manner for each $\lambdans$, (ii) within $\sim 10$
+observations, median measured $\lambdans$ for all populations lie within
+$10\%$ of the true values, and (iii) it takes
 approximately $20$ events to distinguish definitively (with $90\%$ confidence)
@@
 nuclear equations of state. This is comparable to what has been found for
-monolithic binaries of neutron stars~\cite{DelPozzo:13,Chatziioannou:2015uea,
-Agathos:2015a}.
+binary neutron stars~\cite{DelPozzo:13,Chatziioannou:2015uea,Agathos:2015a}.
 
@@
 From the figure, we can conclude with a high likelihood that our measured
-$\lambdans$ values will be within $10\%$ of the real value in $15-20$ events
-for less deformable neutron stars with $\lambdans\leq 1000$, and in as few as
-$10$ events for more deformable stars with $\lambdans\geq 1500$.
+$\lambdans$ values will be within $10\%$ of the real value after about
+$20$ events for less deformable neutron stars with $\lambdans\leq 1000$,
+and in as few as $10$ events for more deformable stars with
+$\lambdans\geq 1500$.
+% 
+The results discussed in Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0},
+\ref{fig:TT_Lambda_vs_N_CI90_0} and the left panel
+of Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} all apply to the parameter distribution spanned by 
+the grid G. This distribution has $\mbh$ as low as $2.7M_\odot$ (i.e. $q=2$).
+Given that disruptive signatures are strongest for small $\mbh$, we now
+investigate an alternate paradigm~\footnote{We denote our standard paradigm
+which does not respect the mass-gap as paradigm A; with paradigm B being
+the scenario where no black hole masses fall within the mass gap.}
+with $\mbh\geq 5M_\odot$, consistent
+with the ``mass-gap'' suggested by astronomical observations~\cite{Bailyn:1997xt,
+Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}
 % 
@@
 measured $\lambdans$ to lie within $10\%$ of the true value (fewer for 
-more deformable NSs, as one would expect).
-% 
-This is a promising deduction, as one might expect to see $\mathcal{O}(10)$
-events over a few years' timescale with design aLIGO~\cite{Abadie:2010cfa}.
-
-Further, we investigate the statistical uncertainties associated with
+more deformable NSs).
+% % 
+% This is a promising deduction, as one might expect to see $\mathcal{O}(10)$
+% events over a few years' timescale with design aLIGO~\cite{Abadie:2010cfa}.
+
+Finally, we investigate the statistical uncertainties associated with
 $\lambdans$ measurements. We use $90\%$ confidence intervals as our measure of
@@
 Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne} shows results
-under paradigm A, where BH mass priors include the mass-gap. In each panel
+under paradigm A. In each panel
 we show the width of the $90\%$ confidence interval measured from $N$ events
 (with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
-the population. In grey are shown results from each of $100$ population
+the population. In gray are shown results from each of $100$ population
 realizations, with the single population highlighted in color corresponding
@@
 %
-These are some of the primary findings of this paper.
 
@@
 the NS deformability, which would allow us to rule out or rank different
-equations of state for neutron star matter.
+equations of state for neutron star matter. Our key findings are 
+summarized in Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} and
+\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}.
 
@@
 and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
-
-\prayush{[PK: To be continued after revision of the introduction.]}
 
@@
 %
-\textbf{%
 We inject signals into zero noise at different SNRs and recover
@@
 replaced by their nearest neighbor in the above set.
-}
+% 
 We make sure that this approximation is conservative by keeping our population
@@
 
-\red{Move to discussion}
 \prayush{%
@@
 
-\red{
-MOVE towards end of SUMMARY, into ``Future work''.
-}
 \prayush{

### 2016-07-11 Paper/paper.tex
added new figures

@@
 % 
+% \begin{figure*}
+% \centering
+% \includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
+% \includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}
+% \caption{%
+% {\it Left:} This figure shows the median value of the recovered
+% probability distribution for $\lambdans$, as a function of the number of events
+% in the population $N$. There are four family of curves, one corresponding each
+% to $\lambdans=\{500,1000,1500,2000\}$, with $100$ independent populations
+% within each family. One curve in each family is highlighted in color, 
+% representing that it is the same population as was illustrated in
+% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
+% In the same color we show $\pm 10\%$ error-bounds on $\lambdans$ with
+% horizontal dash-dotted lines. We observe that within $10-25$ observations, 
+% the median of the measured cumulative probability distribution for $\lambdans$
+% will lie within $10\%$ of the true value.
+% {\it Right:} This figure is identical to the left panel, with the only
+% difference that the BH masses in each population are restricted to lie
+% {\it outside} the astrophysical mass-gap (i.e. paradigm B). The difference that
+% we observe under this paradigm is that we need more ($30+$) events to achieve 
+% the same ($10\%$) measurement accuracy for populations with $\lambdans<1000$.
+% For more deformable neutron stars, $10-25$ events would suffice.
+% }
+% \label{fig:TT_LambdaMedian_vs_N_AllInOne}
+% \end{figure*}
+% %
+% % 
+% \begin{figure*}
+% \centering    
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L2000.pdf}\\
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1500.pdf}\\
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1000.pdf}\\
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L800.pdf}\\
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
+% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L500.pdf}
+% \caption{%
+% These figures show the width of the $90\%$ confidence interval for $\lambdans$
+% (normalized by its true value), as a function of the number of observed events
+% $N$. The left column panels show populations sampled under paradigm A, which
+% allows BH masses to fall within the astrophysical mass-gap, while those on the
+% right show populations drawn under paradigm B which respects the mass-gap.
+% Each panel corresponds to a unique value of populations' $\lambdans$,
+% decreasing from $2000\rightarrow 500$ as we go from top to bottom. In each panel,
+% $100$ curves are shown, each corresponding to an independent population draw. The
+% one colored curve in each panel highlights the population used in
+% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
+% The insets in each panel show the same information, with the ordinate {\it not}
+% normalized by the true value of $\lambdans$.
+% % 
+% We find that with approximately $25$ or so events, we begin to put
+% statistically meaningful constraints on $\lambdans$, restricting it to within
+% $\pm 50\%$ of the true value. We can expect to achieve this with a few years
+% of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of the
+% confidence intervals will require $40+$ events.
+% }
+% \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}
+% \end{figure*}
+% % 
 \begin{figure*}
@@
 \includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
-\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}
+\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}\\
+\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian90pc_vs_N_AllPopulation}
+\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian90pc_vs_N_AstroPopulation}\\
 \caption{%
@@
 % 
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=1.03\columnwidth]{plots/RelErrorLambdaMedian_vs_N.pdf}
-% \includegraphics[width=1.03\columnwidth]{plots/LambdaCIWidths_vs_N.pdf}
-% \caption{{\it Left}: In the main panel is shown the width of the measured $90\%$ 
-% confidence interval for $\lambdans$, normalized by its true value,
-%  as a function of the number of events observed $N$. In the inset, we
-%  show the same measurement uncertainty, but not normalized, and on a 
-%  logarithmic axis for $N$ as well.
-% % 
-%  From both the inset and the main plot, we find that for the first $3-4$ events
-%  our measurement is prior limited. As $N$ increases, the uncertainty falls 
-%  roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
-%  (than $1/\sqrt{N}$) narrowing of the confidence interval.
-% %  
-%  {\it Right}: This figure shows the fractional difference between the median
-%  recovered value from the posterior probability distribution for $\lambdans$
-%  as a function of $N$. Different curves correspond to different injected values
-%  of $\lambdans$, and are shaded from light to dark in direct proportion. 
-%  The horizontal dashed brown line marks $10\%$ deviation of the median from the 
-%  true value.
-% %  
-% We find that within $\approx 10$ detections, the median measured value for
-% $\lambdans$ would estimate the true value for the star with the remarkable 
-% accuracy of $10\%$.
-% }
-% \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
-% \end{figure*}
-%LambdaCIWidths_vs_N_AllPopulations_Log_L
-% 
-%
-% 
 \begin{figure*}
 \centering    
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L2000.pdf}\\
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1500.pdf}\\
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1000.pdf}\\
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L800.pdf}\\
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
-\includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L500.pdf}
+\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AllPopulation.pdf}
+\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation.pdf}
 \caption{%
@@
 \end{figure*}
-% 
-% % 
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
-% \includegraphics[width=.9\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}\\
-% \includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
-% \includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
-% \includegraphics[width=.67\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
-% \caption{ In the main panel is shown the width of the measured $90\%$ 
-% confidence interval for $\lambdans$, normalized by its true value,
-%  as a function of the number of events observed $N$. In the inset, we
-%  show the same measurement uncertainty, but not normalized, and on a 
-%  logarithmic axis for $N$ as well.
-% % 
-%  From both the inset and the main plot, we find that for the first $3-4$ events
-%  our measurement is prior limited. As $N$ increases, the uncertainty falls 
-%  roughly as $1/\sqrt{N}$ with intermittent loud signals causing more rapid 
-%  (than $1/\sqrt{N}$) narrowing of the confidence interval.
-% }
-% \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}
-% \end{figure*}
-% % 
 %

### 2016-07-11 emcee_PE/scripts/MultipleObservations.py

@@ -85,7 +85,6 @@ plt.rcParams.update({\
 
 
 
-
 #######################################################
 ### FUNCTIONS 
 #######################################################
@@ -147,7 +146,25 @@ output:-
     #
     return [llimit, median, ulimit]
 
+#
+def ensemble_percentiles(ensembles):
+    NEVENTS = np.shape(ensembles)[1]
+    Narray = np.arange(1, NEVENTS+1)
+    #
+    plot_Lambdavec = [500, 1000, 1500, 2000]
+    #
+    yy = np.array( [ [ensemble_array[j] for ensemble_array in ensembles] for j in range(len(Narray))] )
+    ensemble_lowlims, ensemble_uplims = [], []
+    #
+    for i in range( NEVENTS ):
+        zz = np.sort(yy[i])
+        low_lim, up_lim = np.percentile(zz, [5., 95.], interpolation='nearest')
+        ensemble_lowlims.append(low_lim)
+        ensemble_uplims.append(up_lim)
+    #
+    return np.array(ensemble_lowlims), np.array(ensemble_uplims)
 
+#
 def find_closest_match(tmp_rand, vec):
     '''
 Take a random number in the range spanned by vec, and find the

### 2016-07-11 Paper/paper.tex
minor changes

@@
 \newcommand{\AEI}{\affiliation{Albert Einstein Institute,
-Am M\"uhlenberg, G\"olm, Germany}} %
+Am M\"uhlenberg, Golm, Germany}} %
 \newcommand{\CIFAR}{\affiliation{Canadian Institute for Advanced Research, 
@@
 % THESE FIGURES ARE FOR THE NEXT SECTION
-\begin{figure*}
-\centering 
-\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
-\caption{We show here probability distributions for mass ratio $\eta$ as measured
-for the same injection at different SNRs. The intrinsic parameters of the source
-are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
-and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
-{\it ignore} tidal effects.
-% 
-In each panel: the dashed red line marks the median {\it measured} value
-$\eta^\mathrm{Median}$, while the dashed green line show the true value
-$\eta^\mathrm{Injected}$. The difference between the true and measured values
-is the systematic bias introduced
-by ignoring tidal effects in templates. The darker shading shows
-the recovered $90\%$ confidence interval for $\eta$, $(\Delta\eta)^{90\%}$, 
-which is a direct measure of our statistical uncertainty on $\eta$.
-% 
-Comparing systematic and statistical errors, we find that:
-at $\rho=20$, $\eta$ measurement is dominated by statistical
-errors; at $\rho=30$, the two become comparable; and 
-for louder signals ($\rho\simeq50$), the systematic errors dominate.
-}
-\label{fig:SingleSystemEtaPDFvsSNR}
-\end{figure*}
-% #################
-\begin{figure*}
-\centering 
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
-\caption{
-We show here the statistical uncertainty of our measurement of the 
-dimensionless mass ratio $\eta$ (at $90\%$ confidence), over the NSBH
-parameter space.
-Individual panels show the same as a function of BH mass and spin.
-Across each row, we see the effect of increasing signal strength (i.e.
-SNR) with the NS's tidal deformability $\lambdans$ fixed. Down each
-column, we see the effect of increasing $\lambdans$, at fixed SNR.
-}
-\label{fig:CIWidths90_Lambda_SNR}
-\end{figure*}
+
 % ###############################
@@
 
-\red{SAY HERE THAT WE ONLY USE THE $l=|m|=2$ modes.}
-\prayush{
-This is possible because of our simplification of 
-using only the dominant $l=|m|=2$ modes of the GW strain~\footnote{The 
-$l=|m|=2$ multipoles of the GW strain, when decomposed in a spin $-2$ weighted
-spherical harmonic basis, contain more than $99\%$ of the total signal power}
-that makes various angle parameters completely degenerate with distance. 
-}
-% 
 \subsection{Waveform Models}\label{s2:waveforms}
@@
 increases the overall model accuracy during inspiral.
-% We will refer to this enhanced-LEA model henceforth as ``LEA+''.
+% 
+Finally, we note that we approximate the full GW signal with its dominant
+$l=|m|=2$ modes, that are modeled by LEA+.
+% ~\footnote{The 
+% $l=|m|=2$ multipoles of the GW strain, when decomposed in a spin $-2$ weighted
+% spherical harmonic basis, contain more than $99\%$ of the total signal power}
+% This approximation has the added advantage of making various angle parameters
+% degenerate with source distance.
 
@@
 % }
-
-\red{SOMEHOW INTRODUCE SNR AND MENTION IF IT IS A NETWORK SNR OR WHAT KIND OF 
-'SNR' IS IT?}
 
@@
 individual component masses, BH spin, and NS's tidal deformability. In
-addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not valid for 
+addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not calibrated for 
 $1\leq q\leq 2$. $p(d_n|\vec{\Theta}, K)$ is the
@@
 \begin{equation}\label{eq:likelihood}
- \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N}\, \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+ p(d_n| \vec{\Theta}, K) \equiv \LL(\vec{\Theta}) = \mathcal{N}\, \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
 \end{equation}
@@
 where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
-and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}
+and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}.
+\red{[MP: INSERT A LINE ABOUT APPROXIMATING $\LL \approx e^{\rho^2/2}$}
 The denominator in Eq.~\ref{eq:postprob} is the a-priori probability of finding
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\begin{figure*}
+\centering 
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
+\caption{We show here probability distributions for mass ratio $\eta$ as measured
+for the same injection at different SNRs. The intrinsic parameters of the source
+are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
+and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
+{\it ignore} tidal effects.
+% 
+In each panel: the dashed red line marks the median {\it measured} value
+$\eta^\mathrm{Median}$, while the dashed green line show the true value
+$\eta^\mathrm{Injected}$. The difference between the true and measured values
+is the systematic bias introduced
+by ignoring tidal effects in templates. The darker shading shows
+the recovered $90\%$ confidence interval for $\eta$, $(\Delta\eta)^{90\%}$, 
+which is a direct measure of our statistical uncertainty on $\eta$.
+% 
+Comparing systematic and statistical errors, we find that:
+at $\rho=20$, $\eta$ measurement is dominated by statistical
+errors; at $\rho=30$, the two become comparable; and 
+for louder signals ($\rho\simeq50$), the systematic errors dominate.
+}
+\label{fig:SingleSystemEtaPDFvsSNR}
+\end{figure*}
+% #################
+\begin{figure*}
+\centering 
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+\caption{
+We show here the statistical uncertainty of our measurement of the 
+dimensionless mass ratio $\eta$ (at $90\%$ confidence), over the NSBH
+parameter space.
+Individual panels show the same as a function of BH mass and spin.
+Across each row, we see the effect of increasing signal strength (i.e.
+SNR) with the NS's tidal deformability $\lambdans$ fixed. Down each
+column, we see the effect of increasing $\lambdans$, at fixed SNR.
+}
+\label{fig:CIWidths90_Lambda_SNR}
+\end{figure*}
 % #################

### 2016-07-11 Paper/paper.tex
description of new lambda median plots

@@
 underlying stochasticity of the population generation process makes it
-difficult to draw generalized inferences (from a single realization) about the
-measurability of $\lambdans$. We mitigate this by drawing multiple
-realizations of population generation (i.e. multiple identically constructed
-populations). Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} shows the first
-results for multiple populations. Lets focus on the left panel first. We
-show the median measured value of $\lambdans$ from $100$ independent 
-populations with fixed $\lambdans=2000$ as the top set of curves. The one 
-highlighted (in green) curve corresponds to the population discussed above.
-Similarly, we generate sets of $100$ populations with $\lambdans=500-1500$, and
-show the median measured $\lambdans$ as four distinct groups of curves in
-Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} (with the single highlighted
-curve in each group corresponding to
-Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}).
-Dash-dotted horizontal lines demarcate $\pm50\%$ error intervals around
-the true $\lambdans$ values. All of these populations are drawn under
-paradigm A (i.e. with BH masses allowed to sample the mass-gap that spans
-$2-5M_\odot$).
-% 
-From the figure, we can conclude with a high likelihood that our measured
-$\lambdans$ values will be within $10\%$ of the real value after about
-$20$ events for less deformable neutron stars with $\lambdans\leq 1000$,
-and in as few as $10$ events for more deformable stars with
-$\lambdans\geq 1500$.
+difficult to draw generalized inferences (from a single realization of an
+NSBH population) about the measurability of $\lambdans$. In order to mitigate
+this, we discuss ensembles of population draws next. In
+Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} we show the measured
+({median}) $\lambdans$ as a function of the number of observed
+events, for four population ensembles, with a hundred population draws
+in each ensemble. Lets focus on the {\it top left} panel first. In it, we
+show the measured (median) $\lambdans$ for all populations in four ensembles,
+with true $\lambdans=\{2000,1500,1000,500\}$ from top to bottom.
+Populations highlighted in color are simply those that we discussed in
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}. Dash-dotted horizontal lines
+demarcate $\pm10\%$ error intervals around the true $\lambdans$ values.
+The panel just below it shows the range of $\lambdans$ that encloses the
+median measured $\lambdans$ for $90\%$ of the populations in {\it each}
+ensemble. In other words, this panel shows the range of $\lambdans$ within which
+the median measured $\lambdans$ value for $90\%$ of NSBH populations is
+expected to lie. From these panels, we conclude that our measured $\lambdans$
+values will be within $10\%$ of the {\it true} value after $\sim 25$
+detections of less deformable neutron stars ($\lambdans\leq 1000$), or
+after as few as $15$ detections of more deformable neutron stars
+($\lambdans\geq 1500$).
 % 
 The results discussed in Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0},
-\ref{fig:TT_Lambda_vs_N_CI90_0} and the left panel
-of Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} all apply to the parameter distribution spanned by 
-the grid G. This distribution has $\mbh$ as low as $2.7M_\odot$ (i.e. $q=2$).
+\ref{fig:TT_Lambda_vs_N_CI90_0} and the left two panels
+of Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} apply to the parameter
+distribution spanned by the grid G. This distribution allows for $\mbh$
+as low as $2.7M_\odot$ (i.e. $q=2$).
 Given that disruptive signatures are strongest for small $\mbh$, we now
@@
 which does not respect the mass-gap as paradigm A; with paradigm B being
-the scenario where no black hole masses fall within the mass gap.}
+the scenario where no black hole masses fall within the mass gap $2-5M_\odot$.}
 with $\mbh\geq 5M_\odot$, consistent
@@
 % 
-The right panel of the figure is identical to the left panel, but drawn under
-population paradigm B (no BHs alloewd within the mass-gap). Under this
-paradigm, we expectedly find that information accumulation is slower. It would
-take $20-40$ detections with $\rho\geq10$ under this paradigm, for our (median)
-measured $\lambdans$ to lie within $10\%$ of the true value (fewer for 
-more deformable NSs).
+Both right panels of the figure are identical to their corresponding left
+panels, but drawn under population paradigm B. Under this paradigm, we
+expectedly find that information accumulation is much slower. It would
+take $25-40$ detections with $\rho\geq10$ under this paradigm, for our (median)
+measured $\lambdans$ to converge within $10\%$ of its true value.
 % % 

### 2016-07-11 Paper/paper.tex
description of new lambda CI plots

@@
 $\lambdans$ measurements. We use $90\%$ confidence intervals as our measure of
-the same. The left column of 
-Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne} shows results
-under paradigm A. In each panel
-we show the width of the $90\%$ confidence interval measured from $N$ events
-(with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
-the population. In gray are shown results from each of $100$ population
-realizations, with the single population highlighted in color corresponding
-to the one we focused on in 
-Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
-The inset shows the same, except that the ordinate is {\it not} normalized by
-the true value of $\lambdans$. From top to bottom, the population $\lambdans$
+the same. First, we draw an ensemble of a hundred populations each for
+$\lambdans=\{500,1000,1500,2000\}$. For each population $i$ in each ensemble,
+we construct
+its $90\%$ confidence interval $[{\lambdans^{90\%}}_{i-},{\lambdans^{90\%}}_{i+}]$.
+Next, we construct the interval $[X^-,Y^-]$ that contains ${\lambdans^{90\%}}_{i-}$
+for $90\%$ of the populations in each ensemble; and similarly $[X^+,Y^+]$
+for ${\lambdans^{90\%}}_{i+}$. Finally, in the left panel of 
+Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}, we show the
+conservative width $|Y^+ - X^-|$ that contains the $90\%$ confidence
+intervals for $90\%$ of all populations in each ensemble~\footnote{Drawn
+under paradigm A.}.
+% 
+% In each panel
+% we show the width of the $90\%$ confidence interval measured from $N$ events
+% (with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
+% the population. In gray are shown results from each of $100$ population
+% realizations, with the single population highlighted in color corresponding
+% to the one we focused on in 
+% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
+% The inset shows the same, except that the ordinate is {\it not} normalized by
+% the true value of $\lambdans$. 
+% 
+From top to bottom, the population $\lambdans$
 decreases from $\lambdans=2000\rightarrow 500$, corresponding to decreasingly
@@
 we can constrain $\lambdans$ within $\pm 50\%$ using only $10-20$ events, and
-within $\pm 25\%$ with $25-40$ events; (ii) for softer equations of state with
+within $\pm 25\%$ (marked by black circles) with $25-40$ events; (ii) for softer 
+equations of state with
 $\lambdans<1000$, we will achieve the same accuracy with $20-30$ and $50+$ 
 events, respectively; and (iii) for the first $5$ or so observations, our
-measurement confidence spans the entire prior allowed range 
-$\lambdans\in[0,4000]$, as shown by the plateauing towards the left edge in
-each panel's inset.
-% 
-In the right column, each panel panel is identical to its corresponding left
-one, with the difference that populations are drawn under paradigm B, which
-does {\it not} allow for BH masses to fall within the mass-gap. We find that
+measurement spans the entire prior allowed range:
+$\lambdans\in[0,4000]$, as shown by the plateauing of the $90\%$ 
+confidence intervals towards the left edge to $90\%$ of $4000$, i.e. $3600$.
+% 
+The right panel in Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}
+is identical to the left one, with the difference that populations are drawn
+under paradigm B, which does {\it not} allow for BH masses to fall within
+the mass-gap. We find that
 for NSs with $\lambdans\leq 1000$, it would take $25-40$ events to
@@
 % 
+Under either paradigms, we find that measuring $\lambdans$ better than
+$25\%$ will require $\mathcal{O}(10^2)$ observations of disruptive NSBH
+mergers.
 

### 2016-07-11 Paper/paper.tex
update fig captions

@@
 \caption{%
-{\it Left:} This figure shows the median value of the recovered
+{\it Top left}: This figure shows the median value of the recovered
 probability distribution for $\lambdans$, as a function of the number of events
-in the population $N$. There are four family of curves, one corresponding each
-to $\lambdans=\{500,1000,1500,2000\}$, with $100$ independent populations
-within each family. One curve in each family is highlighted in color, 
-representing that it is the same population as was illustrated in
+in the population $N$. There are four ensembles of curves,
+corresponding to $\lambdans=\{500,1000,1500,2000\}$, with a hundred
+independent population draws within each ensemble. One curve in each ensemble
+is highlighted in color, representing only that it is the same population as 
+was discussed in
 Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
 In the same color we show $\pm 10\%$ error-bounds on $\lambdans$ with
-horizontal dash-dotted lines. We observe that within $10-25$ observations, 
-the median of the measured cumulative probability distribution for $\lambdans$
-will lie within $10\%$ of the true value.
-{\it Right:} This figure is identical to the left panel, with the only
-difference that the BH masses in each population are restricted to lie
-{\it outside} the astrophysical mass-gap (i.e. paradigm B). The difference that
+horizontal dash-dotted lines.
+{\it Bottom left}: Here we show the interval of $\lambdans$ values within
+which the median measured $\lambdans$ lies for $90\%$ of the populations in
+each ensemble shown in the top left panel.
+% 
+We observe that within $10-25$ observations, the median of the measured 
+cumulative probability distribution for $\lambdans$ converges to within $10\%$
+of its true value.
+% 
+{\it Right:} These panels are identical to their counterparts on the left,
+with the only difference that the BH masses in each population are restricted
+to lie {\it outside} the astrophysical mass-gap (i.e. paradigm B). The
+difference that
 we observe under this paradigm is that we need more ($30+$) events to achieve 
@@
 \centering    
-\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AllPopulation.pdf}
-\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation.pdf}
+\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AllPopulation}
+\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation}
 \caption{%
-These figures show the width of the $90\%$ confidence interval for $\lambdans$
-(normalized by its true value), as a function of the number of observed events
-$N$. The left column panels show populations sampled under paradigm A, which
-allows BH masses to fall within the astrophysical mass-gap, while those on the
-right show populations drawn under paradigm B which respects the mass-gap.
+{\it Left}: This panel shows the width of $\lambdans$ interval within
+which the $90\%$ confidence intervals for $\lambdans$ lie, for $90\%$ of 
+the populations in each ensemble, as a function of the number of observed events
+$N$. Details of how this is calculated are given in the text.
+% 
+The populations are sampled under paradigm A, which allows BH masses to
+fall within the astrophysical mass-gap.
 Each panel corresponds to a unique value of populations' $\lambdans$,
-decreasing from $2000\rightarrow 500$ as we go from top to bottom. In each panel,
-$100$ curves are shown, each corresponding to an independent population draw. The
-one colored curve in each panel highlights the population used in
-Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
-The insets in each panel show the same information, with the ordinate {\it not}
-normalized by the true value of $\lambdans$.
+decreasing from $2000\rightarrow 500$ as we go from top to bottom.
+% 
+{\it Right}: This panel shows populations drawn under paradigm B, which
+respects the mass-gap.
 % 
@@
 $\pm 50\%$ of the true value. We can expect to achieve this with a few years
-of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of the
-confidence intervals will require $40+$ events.
+of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of 
+$\lambdans$ confidence intervals will require $40+$ events.
 }
@@
 $\delta\lambdans\propto\lambdans^\beta$, we estimate $\beta$ and show it in
-this figure as a function of the number of observed events $N$.
+this figure as a function of the number of observed events $N$. The thicker
+curve corresponds to the populations discussed in
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}.
 % 
@@
 % 
-Finally, we quantitatively explore the dependence of our statistical
+Here, we quantitatively explore the dependence of our statistical
 uncertainties for $\lambdans$ on the number of events, as well as on the true

### 2016-07-29 Paper/paper.tex
incorporating comments from mp

@@
 \title{
-Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of mixed binary observations
+Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis
+of neutron star - black hole binary observations
 }
@@
 GWs from NSBH sources carry signatures of (a) the star's tidal distortion by
-the companion black hole (BH) during inspiral, and (b) its tidal disruption
-(if that happens) near merger.
+the companion black hole (BH) during inspiral, and (b) its potential tidal
+disruption near merger.
 %
@@
 % 
-Second, we find that if we are to get a loud signal with SNR $\gtrsim 23$, 
-we can put a factor of two bound on $\lambdans$.
+Second, for (loud) signals with SNR $\gtrsim 23$ we can bound  $\lambdans$
+by a factor of two.
 % 
@@
 $10\%$ of the true value in approximately $20$ observations; (ii) in the same
-number of observations, the associated $90\%$ confidence intervals shrink to
+number of observations, the associated $90\%$ credible intervals shrink to
 $\pm 50\%$ of the true value.
@@
 waves~\cite{LIGOVirgo2016a}. Emitted by a pair of coalescing black holes, these
-waves heralded an era of observational GW astrophysics as they traveled 
+waves have heralded an era of observational GW astrophysics as they traveled 
 through Earth.
@@
 % }
-Coalescing compact binaries (both monolithic and mixed) of stellar-mass black
-holes (BH) and neutron stars (NS) are the primary targets for the second
+Coalescing compact binaries of stellar-mass black holes (BH) and/or
+neutron stars (NS) are the primary targets for the second
 generation GW detectors~\cite{Timmes:1995kp,Fryer:1999mi,RevModPhys.74.1015,
@@
 Wex:1998wt,1991ApJ...379L..17N,Mandel:2015spa,Abbott:2016nhf}.
-A monolithic binary of black holes was recently observed by 
+A binary system of black holes was recently observed by 
 aLIGO~\cite{LIGOVirgo2016a}. Previously, stellar-mass black holes had only 
@@
 in varied situations: as radio pulsars~\cite{Lattimer:2012nd,Manchester:2004bp},
-in binary systems with stellar companions~\cite{1971ApJ...169L..23M,
+in binary systems with a stellar companion~\cite{1971ApJ...169L..23M,
 Bond:2002eh,Lattimer:2012nd,Manchester:2004bp},
-and in monolithic neutron star binaries~\cite{Hulse:1975uf,Taylor:1982wi,
+and in neutron star binaries~\cite{Hulse:1975uf,Taylor:1982wi,
 Weisberg:2010zz,Lattimer:2012nd,Manchester:2004bp}.
@@
 interesting class of systems~\cite{Wex:1998wt,
-1991ApJ...379L..17N,Janka1999,Fryer:2015jpa}, that is yet undetected.
+1991ApJ...379L..17N,Janka1999,Fryer:2015jpa}, that has not yet been detected.
 We expect to observe $\mathcal{O}(10)$ mixed binaries per year with
@@
 However, whether or not such a system forms depends also on the nature of
-the BH. For e.g., massive BHs (with $\mbh\gtrsim 10M_\odot$), as well as BHs with
+the BH. Massive BHs (with $\mbh\gtrsim 10M_\odot$), as well as BHs with
 large retrograde spins, tend to swallow the NS whole without forming a
@@
 A {\it coincident} detection of both GWs and gamma-rays from an NSBH merger,
-whenever it happens, will
-provide us with a uniquely opportunity to confirm this hypothesized link between
+will
+provide us with a unique opportunity to confirm this hypothesized link between
 NSBH mergers and GRBs~\cite{Abbott:2016wya}.
@@
 information for NSBHs, and (ii) the errors due to unknown vaccum terms in PN 
-waveforms could dominate over the tidal terms themselves~\cite{Barkett2015,
+waveforms which could dominate over the tidal terms themselves~\cite{Barkett2015,
 Yagi:2014}.
@@
 Second, we include the effect of black hole spin on tidal GW signals, in
-addition to the effect of BH mass, NS's deformability, and the SNR.
+addition to the effect of BH mass, tidal deformability of the NS, and the SNR.
 % 
@@
 
-We first probe the effect of ignoring tidal effects in parameter 
+We now outline the main questions and results discussed in this paper.
+First, we probe the effect of ignoring tidal effects in parameter 
 estimation on the recovery of non-tidal binary parameters, such as 
@@
 Second, we study the ability of aLIGO to constrain neutron star tidal 
-deformability with a single observation of NSBH merger. For this, we
+deformability with a single observation of an NSBH merger. For this, we
 use the same setup for signal waveforms as before, but replace the filter
@@
 For all populations, we take source locations to be uniformly distributed in
-spatial volume, and source orientations to be uniform on a $2-$sphere. To
+spatial volume, and source orientations to be uniform on the $2-$sphere. To
 summarize, we find the following:
@@
 taking particular values. Throughout this paper, we impose a prior uniform on
-individual component masses, BH spin, and NS's tidal deformability. In
+individual component masses, BH spin, and the tidal deformability of the NS. In
 addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not calibrated for 
@@
 and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}.
-\red{[MP: INSERT A LINE ABOUT APPROXIMATING $\LL \approx e^{\rho^2/2}$}
+% \red{[MP: INSERT A LINE ABOUT APPROXIMATING $\LL \approx e^{\rho^2/2}$}
+As in Ref.~\cite{Purrer:2015nkh} we use a likelihood that is maximized over
+over the template norm
+and we ignore extrinsic parameters that only enter in the template norm such
+as distance, orientation and sky location. In addition, we also maximize over
+coalescence time and phase.
 The denominator in Eq.~\ref{eq:postprob} is the a-priori probability of finding
@@
 
-We use an ensemble sampler Markov-chain Monte-Carlo algorithm based on the
-{\tt emcee} package~\cite{emcee}, to sample the probability distribution 
-$p(\vec{\Theta} | d_n, K)$. We start with $100$ independent walkers, each
-of which is allowed to collect $100,000$ samples.
+We use the ensemble sampler Markov-chain Monte-Carlo algorithm implemented in
+the {\tt emcee} package~\cite{emcee}, to sample the probability distribution 
+$p(\vec{\Theta} | d_n, K)$. We run 100 independent chains, each of which is
+allowed to collect 100, 000 samples and combine samples from chains that have
+a Gelman-Rubin statistic~\cite{gelman1992} close to unity. This procedure yields
+about 10,000 independent samples.
 % 
@@
 frequency sampling interval to $\Delta f=0.4$~Hz, which we find to be
-sufficient for robust likelihoods calculations~\cite{Purrer:2015nkh}.
-Once more than $10,000$ independent samples have been collected, we integrate
+sufficient for robust likelihoods calculations in zero noise~\cite{Purrer:2015nkh}.
+We integrate
 Eq.~\ref{eq:marginalize} to obtain marginalized probability distributions
@@
 The median value of this distribution is our {\it measured} value for
-$\lambdans$, and the $90\%$ confidence intervals associated with the distribution
+$\lambdans$, and the $90\%$ credible intervals associated with the distribution
 serve as statistical error-bars.
@@
 by ignoring tidal effects in templates. The darker shading shows
-the recovered $90\%$ confidence interval for $\eta$, $(\Delta\eta)^{90\%}$, 
+the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
 which is a direct measure of our statistical uncertainty on $\eta$.
@@
 We show here the statistical uncertainty of our measurement of the 
-dimensionless mass ratio $\eta$ (at $90\%$ confidence), over the NSBH
+dimensionless mass ratio $\eta$ (at $90\%$ credibility), over the NSBH
 parameter space.
@@
 Across each row, we see the effect of increasing signal strength (i.e.
-SNR) with the NS's tidal deformability $\lambdans$ fixed. Down each
+SNR) with the tidal deformability of the NS $\lambdans$ fixed. Down each
 column, we see the effect of increasing $\lambdans$, at fixed SNR.
@@
 each row, we can see the effect of increasing the signal strength (i.e. SNR)
-with the NS's deformability fixed. Down each column, we can see 
-the effect of the increasing NS's tidal deformability at fixed SNR. We also
+with the tidal deformability of the NS fixed. Down each column, we can see 
+the effect of the increasing tidal deformability of the NS at fixed SNR. We also
 show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
@@
 mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
-not including the effects of NS's tidal deformation on GW emission can become the dominant
+not including the effects of tidal deformation of the NS on GW emission can become the dominant
 source of error for astrophysical searches with Advanced LIGO. However,
@@
 Past and future GW search and source characterization efforts with Advanced LIGO
-have used (or plan to) BH-BH waveform templates for NS-BH mergers, ignoring
+have used (or plan to use) BH-BH waveform templates for NS-BH mergers, ignoring
 NS tidal effects on the emitted GWs. In this section we present a fully
@@
 We construct NSBH waveforms using the LEA+ model. We inject these into zero
-noise and run MCMC on the injections using BBH templates (LEA+ with tidal terms
+noise and run an MCMC sampler on the injections using BBH templates (LEA+ with tidal terms
 set to zero). We perform this analysis for a variety of mass ratios
@@
 difference between measured and true $\eta$ values. Darker shading of the
-probability distribution marks the $90\%$ confidence interval for $\eta$, whose
+probability distribution marks the $90\%$ credible interval for $\eta$, whose
 width is a direct measure of the {\it statistical} measurement uncertainty/error.
@@
 Next, to get a better understanding of how {\it precisely} we measure NSBH
-binary parameters given BBH waveforms, we show the $90\%$ confidence intervals
+binary parameters given BBH waveforms, we show the $90\%$ credible intervals
 associated with NSBH mass and spin measurements in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
 The three panels, corresponding to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
-(bottom), show the width of these confidence intervals as a function of BH
+(bottom), show the width of these credible intervals as a function of BH
 mass/spin (within each sub-panel), and NS properties, i.e. $\lambdans$
@@
 templates, to the statistical uncertainties associated with the measurement
-itself. If $\arr\gtrsim 1$, including the effect of NS's tidal distortion
+itself. If $\arr\gtrsim 1$, including the effect of the tidal distortion of the NS
 (during inspiral) and possible disruption (close to merger) would be imperative
@@
 if additionally: (i) the BH has a prograde spin $\chibh\gtrsim 0.4$, {\it and} (ii)
-the NS's true deformability is large enough, s.t. $\lambdans \gtrsim 1000$.
+the true tidal deformability of the NS is large enough, s.t. $\lambdans \gtrsim 1000$.
 % 
@@
 $\lambdans$, and the dashed green line marks its {\it true} value.
-The darker shading shows the $90\%$ confidence interval, whose width
+The darker shading shows the $90\%$ credible interval, whose width
 $(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
@@
 
-In the previous section, we showed that the effect of NS's deformation due to
+In the previous section, we showed that the effect of deformation of the NS due to
 its companion BH's quadrupolar tidal field becomes discernible in the
@@
 % 
-The first three also reduce the onset frequency of NS's disruption, allowing
-for it to happen earlier in the orbit. 
+The first three also reduce the onset frequency of the disruption of the NS,
+allowing for it to happen earlier in the orbit. 
 % 
@@
 while the median recovered value is marked by a red (dashed) line. The darker
-shaded region marks the $90\%$ confidence interval on $\lambdans$. One thing
+shaded region marks the $90\%$ credible interval on $\lambdans$. One thing
 to note in this figure is that the measurement uncertainty on $\lambdans$, i.e.
-the full width of its $90\%$ confidence interval, grows slower than $\lambdans$
+the full width of its $90\%$ credible interval, grows slower than $\lambdans$
 as $\lambdans$ is turned up (while the fractional uncertainty
@@
 of this section. In each panel, as a function of black hole mass and spin,
-we show the measured $90\%$ confidence intervals for $\lambdans$, i.e.
+we show the measured $90\%$ credible intervals for $\lambdans$, i.e.
 $(\Delta\lambdans)^{90\%}$. These correspond to the full width of the dark
@@
 i.e. SNR, can be seen as we go from left to right in each row. The
-effect of NS's tidal deformability $\lambdans$ on its own measurability
+effect of the tidal deformability of the NS $\lambdans$ on its own measurability
 can be seen by comparing panels within each column. The NS becomes more 
@@
 $\lambdans$ is dominated by the width of our prior on $\lambdans$. In other
-words, our measured confidence interval is restricted by the
+words, our measured credible interval is restricted by the
 allowed range of $\lambdans$. Increasing the signal strength to
@@
 in Fig.~\ref{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0} we show the minimum
-signal strength required to attain a certain level of confidence in our
+signal strength required to attain a certain level of credibility in our
 $\lambdans$ measurement, as a function of BH properties. The NS is allowed
@@
 Littenberg:2015tpa}, we will simultaneously need to restrict BH spins
-to $\chibh\gtrsim +0.5$ to obtain the same measurement confidence at the same
+to $\chibh\gtrsim +0.5$ to obtain the same measurement credibility at the same
 source location.
@@
 {\it Left:} Posterior probability distributions for $\lambdans$ (colored 
-curves), and associated $90\%$ confidence intervals (grey vertical lines),
+curves), and associated $90\%$ credible intervals (grey vertical lines),
 shown for different number of accumulated observations $N$. Distributions are
@@
 {\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
-associated $90\%$ confidence intervals (as the vertical extent of filled
+associated $90\%$ credible intervals (as the vertical extent of filled
 region), shown as a function of number of observations $N$. Solid horizontal
@@
 % \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
-% \caption{Filled-region plots showing the median and $90\%$ confidence intervals
+% \caption{Filled-region plots showing the median and $90\%$ credible intervals
 % for $\lambdans$ measurement, as a function of the number of observed events $N$.
@@
 % the value shown in its title. In each panel, the measured median for $\lambdans$
-% is shown by filled circles, and the $90\%$ confidence interval is shown by the 
+% is shown by filled circles, and the $90\%$ credible interval is shown by the 
 % height of the filled region. In addition, the dashed and 
@@
 measured $\lambdans$ value (as filled circles), as well as the associated
-$90\%$ confidence intervals for the measurement (as the vertical extent of the
+$90\%$ credible intervals for the measurement (as the vertical extent of the
 filled region about the median), as functions of $N$.
@@
 % \caption{%
-% These figures show the width of the $90\%$ confidence interval for $\lambdans$
+% These figures show the width of the $90\%$ credible interval for $\lambdans$
 % (normalized by its true value), as a function of the number of observed events
@@
 % of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of the
-% confidence intervals will require $40+$ events.
+% credible intervals will require $40+$ events.
 % }
@@
 {\it Left}: This panel shows the width of $\lambdans$ interval within
-which the $90\%$ confidence intervals for $\lambdans$ lie, for $90\%$ of 
+which the $90\%$ credible intervals for $\lambdans$ lie, for $90\%$ of 
 the populations in each ensemble, as a function of the number of observed events
@@
 of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of 
-$\lambdans$ confidence intervals will require $40+$ events.
+$\lambdans$ credible intervals will require $40+$ events.
 }
@@
 calculation explicitly. For instance, $K$ includes our assumption that all
-NS's in a single population have the same deformability parameter $\lambdans$, 
+NSs in a single population have the same deformability parameter $\lambdans$, 
 and its cumulative measurement is therefore possible.
@@
 rapidly our median estimate for $\lambdans$ converges to the true value,
-and how rapidly our confidence intervals for the same shrink, with increasing
+and how rapidly our credible intervals for the same shrink, with increasing
 $N$. Finally, we generate and analyze an ensemble of populations in order to
@@
 probability distributions for $\lambdans$ as inferred from $N$ events, with $N$
-ranging from $1-80$. We also mark the $90\%$ confidence intervals associated
+ranging from $1-80$. We also mark the $90\%$ credible intervals associated
 with each of the probability distribution curves. The first few observations
@@
 $\pm25\%$ ($\pm50\%$) error bars. At each $N$, the range spanned by the 
-filled region is the $90\%$ confidence interval deduced from the same 
+filled region is the $90\%$ credible interval deduced from the same 
 events. This figure somewhat quantifies the qualitative deductions we made
@@
 quickly, reaching $10\%$ accuracy with $10-15$ observations. With the same
-information, our confidence intervals also shrink to $\pm 25\%$.
+information, our credible intervals also shrink to $\pm 25\%$.
 % 
@@
 track the median measured $\lambdans$, while the filled regions show
-the associated $90\%$ confidence intervals. From the figure, we observe
-the following: (i) the shrinkage of confidence interval widths with increasing
+the associated $90\%$ credible intervals. From the figure, we observe
+the following: (i) the shrinkage of credible interval widths with increasing
 $N$ happens in a similar manner for each $\lambdans$, (ii) within $\sim 10$
@@
 $10\%$ of the true values, and (iii) it takes
-approximately $20$ events to distinguish definitively (with $90\%$ confidence)
+approximately $20$ events to distinguish definitively (with $90\%$ credibility)
 between deformable NSs with $\lambdans=2000$ and compact NSs with 
@@
 Finally, we investigate the statistical uncertainties associated with
-$\lambdans$ measurements. We use $90\%$ confidence intervals as our measure of
+$\lambdans$ measurements. We use $90\%$ credible intervals as our measure of
 the same. First, we draw an ensemble of a hundred populations each for
@@
 we construct
-its $90\%$ confidence interval $[{\lambdans^{90\%}}_{i-},{\lambdans^{90\%}}_{i+}]$.
+its $90\%$ credible interval $[{\lambdans^{90\%}}_{i-},{\lambdans^{90\%}}_{i+}]$.
 Next, we construct the interval $[X^-,Y^-]$ that contains ${\lambdans^{90\%}}_{i-}$
@@
 Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}, we show the
-conservative width $|Y^+ - X^-|$ that contains the $90\%$ confidence
+conservative width $|Y^+ - X^-|$ that contains the $90\%$ credible
 intervals for $90\%$ of all populations in each ensemble~\footnote{Drawn
@@
 % In each panel
-% we show the width of the $90\%$ confidence interval measured from $N$ events
+% we show the width of the $90\%$ credible interval measured from $N$ events
 % (with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
@@
 $\lambdans\in[0,4000]$, as shown by the plateauing of the $90\%$ 
-confidence intervals towards the left edge to $90\%$ of $4000$, i.e. $3600$.
+credible intervals towards the left edge to $90\%$ of $4000$, i.e. $3600$.
 % 
@@
 we can begin to constraint $\lambdans$ at the level of $\pm 75\%$ errors with a $90\%$
-confidence level. If the signal is stronger, say $\rho=30$, we can contraint
+credible level. If the signal is stronger, say $\rho=30$, we can contraint
 $\lambdans$ better with only $\pm 50\%$ errors, for binaries with large BH spins.
@@
   Potsdam, for hospitality during part of the time where this research
-  was completed.
+  was completed. M.P. thanks CITA for hospitality where part of the work
+  was carried out.
 \end{acknowledgments}

### 2016-08-30 Paper/paper.tex
more changes as per MP's comments

@@
 from multiple (realistic) events.
+%
 
@@
 and then use non-tidal (BBH) waveform filters to estimate the underlying
-binary masses and spins with a Markov-chain Monte Carlo.
+binary masses and spins with a Markov-chain Monte Carlo. Here and throughout,
+we use the zero-detuning high-power design sensitivity curve~\cite{Shoemaker2009}
+to characterize the expected detector noise.
 % 
@@
 Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, it is only at SNRs 
-$\rho\gtrsim 20-30$ (under otherwise favorable circumstances) that we are able
+$\rho\gtrsim 20-30$ (under otherwise favorable circumstances, such as a stiff
+equation of state) that we are able
 to bring this down to a $\pm 75\%$ bound on $\lambdans$. For signals louder
@@
 The EoS of matter at nuclear densities is believed to be universal among all
-neutron stars. The Tolman-Oppenheimer-Volkoff equation would then predict that NS
-properties satisfy a universal relationship between $\lambdans$ and $\mns$. As
-the final part of
+neutron stars. The Tolman-Oppenheimer-Volkoff equation~\cite{Tolman:1939jz,
+Oppenheimer:1939ne,1934PNAS...20..169T}
+would then predict that NS properties satisfy a universal relationship
+between $\lambdans$ and $\mns$. As the final part of
 this paper, we combine information from multiple observations of realistic NSBH
@@
 $\Delta\Phi$ is set to the PN tidal phasing corrections,
-at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
+at the leading and next-to-leading orders~\cite{Vines2011}; close to merger,
 additional phenomenological terms are needed. Both $A$ and $\Delta\Phi$ are
@@
 Finally, we note that we approximate the full GW signal with its dominant
-$l=|m|=2$ modes, that are modeled by LEA+.
+$l=|m|=2$ modes, that are modeled by LEA+. For use in future LIGO science
+efforts, we have implemented the LEA+ model in the LIGO Algorithms
+Library~\cite{LAL}.
 % ~\footnote{The 
@@
 Here, $p(\vec{\Theta} | K)$ is the a-priori probability of binary parameters
-taking particular values. Throughout this paper, we impose a prior uniform on
+$\vec{\Theta}$
+taking particular values, given $K$ - which denotes all our collective knowledge,
+except for expectations on binary parameters that enter
+our calculations explicitly. Throughout this paper, we impose a prior that is
+uniform in
 individual component masses, BH spin, and the tidal deformability of the NS. In
@@
 where $h\equiv h(\vec{\Theta})$ is a filter template with parameters 
-$\vec{\Theta}$, and $\langle\cdot|\cdot\rangle$ is a suitably defined
+$\vec{\Theta}$, $\langle\cdot|\cdot\rangle$ is a suitably defined
 detector-noise weighted inner-product~\footnote{The inner product
-$\langle\cdot|\cdot\rangle$  is defined as
+$\langle\cdot|\cdot\rangle$  is defined as~\cite{Shoemaker2009}
 \begin{equation}
@@
 where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
-and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}.
+and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}, and
+$\mathcal{N}$ is an SNR-dependent normalization constant.
 % \red{[MP: INSERT A LINE ABOUT APPROXIMATING $\LL \approx e^{\rho^2/2}$}
@@
 \centering 
+\textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
+\textbf{at different SNR values (shown in legend).}\par\medskip
 \includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
@@
 \centering 
+\textbf{Statistical measurement uncertainty for NSBH parameters $\mchirp$ (top),
+$\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
 \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
@@
 \centering 
+\textbf{Ratio of systematic to statistical errors in measuring $\mchirp$;}\par
+\textbf{ignoring tidal effects.}\par\medskip
 \includegraphics[width=1.6\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
@@
 \centering
+\textbf{Ratio of systematic to statistical errors in measuring $\eta$;}\par
+\textbf{ignoring tidal effects.}\par\medskip
 \includegraphics[width=2\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_70_linear}
@@
 \centering
+\textbf{Ratio of systematic to statistical errors in measuring $\chibh$;}\par
+\textbf{ignoring tidal effects.}\par\medskip
 \includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
@@
 year with $\rho\geq 70$. For our priors, we choose uniform distribution for 
-both component masses as well as for black hole spins, i.e. $\mbh\in[2,20]M_\odot$,
-$\mns\in[1,3]M_\odot$ and $-0.5\leq \chibh\leq +0.75$. \prayush{FIXME}
+both component masses as well as for black hole spins, i.e. $\mbh\in[1.2,25]M_\odot$,
+$\mns\in[1.2,3]M_\odot$ and $-0.75\leq \chibh\leq +0.75$.
+% \prayush{FIXME}
 
@@
 
-We first calculate the ratio of errors $\arr$ associated with chirp mass measurements
-and show it in
+We first calculate the ratio of errors $\arr$ associated with the measurement of
+chirp mass, $\mchirp := (m_1 m_2)^{3/5} (m_1 + m_2)^{-1/5}$, and show it in
 Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the mass combination
@@
 extremes pushes towards milder spins, since the prior excludes any more extreme
-spins than $[-0.75, +0.75]$ and that pushes the bulk of the posterior inwards.
+spins than $[-0.75, +0.75]$ and that pushes the median of the posterior inwards.
 However as the signal's $\lambdans$ is increased from $500\rightarrow 2000$,
@@
 \centering 
+\textbf{Illustrative posterior probability distributions for NS tidal
+deformability $\lambdans$,}\par
+\textbf{for signals with different $\lambdans$ values (shown in legend).}\par\medskip
 % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
@@
 In the previous section, we showed that the effect of deformation of the NS due to
-its companion BH's quadrupolar tidal field becomes discernible in the
+the quadrupolar tidal field of its companion BH becomes discernible in the
 GW spectrum only when the following hold:
@@
 multiple {\it identical} sources at realistic SNRs. Let us consider the case
-a population of optimally oriented binaries~\footnote{An optimally oriented
+of a population of optimally oriented binaries~\footnote{An optimally oriented
 binary is one which is located directly overhead the detector, with the 
@@
 source. Such a configuration maximizes the observed GW signal strength in 
-the detector.}, distributed uniformly in spacial volume out to a maximum
-distance $D^\mathrm{max}$. $D^\mathrm{max}$ is set by the minimum SNR 
+the detector.}, distributed uniformly in spatial volume out
+to a maximum {\it effective} distance~\footnote{{\it effective} distance $D$ 
+is a combination of distance to the source, its orientation, and its sky
+location angles, and has a one-to-one correspondence with SNR for non-precessing
+systems, for which these angles are constant in time.}.
+$D^\mathrm{max}$. $D^\mathrm{max}$ is set by the minimum SNR 
 threshold $\rho_\mathrm{min}$ at which a source is considered
@@
 \begin{equation}\label{eq:rmsSigmaIdenticalSources}
- \sigma_{avg} := \frac{1}{\sqrt{1/\sigma^{2}}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3 N}},
+ \sigma_{avg} := \frac{1}{\sqrt{1/\sigma^{2}}} = \frac{\sigma_0}{D_0} D^\mathrm{max} \frac{1}{\sqrt{3 N}},
 \end{equation}
@@
 While we discussed Eq.~\ref{eq:rmsSigmaIdenticalSources} for a population
-of optimally oriented sources, it is valid more generally for a population
-distributed uniformly in {\it effective} volume~\footnote{{\it effective} 
-distance $D_\mathrm{eff}$ 
-is a combination of distance to the source, its orientation, and its sky
-location angles, and has a one-to-one correspondence with SNR for non-precessing
-systems, for which these angles are constant in time. Effective volume is 
-$\frac{4\pi}{3}D_\mathrm{eff}^3$.}~\cite{Markakis:2010mp}. However, it 
+of optimally oriented sources, it is valid for a more general population
+distributed uniformly in effective volume~\cite{Markakis:2010mp}
+($\propto D^3$).
+% ~\footnote{{\it effective} distance $D$ 
+% is a combination of distance to the source, its orientation, and its sky
+% location angles, and has a one-to-one correspondence with SNR for non-precessing
+% systems, for which these angles are constant in time. Effective volume is 
+% $\frac{4\pi}{3}D^3$.}
+However, it 
 still only applies to sources with identical masses and spins, and we 
@@
 % 
-Further, let $K$ denote all of our collective prior knowledge, except for 
-expectations on binary parameters themselves, which enter the following 
-calculation explicitly. For instance, $K$ includes our assumption that all
-NSs in a single population have the same deformability parameter $\lambdans$, 
-and its cumulative measurement is therefore possible.
+As before, let $K$ denote all of our collective prior knowledge; for instance,
+$K$ includes our assumption that all NSs in a single population have the same
+deformability parameter $\lambdans$, and that its cumulative measurement is
+therefore possible.
 % 
@@
  p(&&\hspace{-4mm}\lambdans | d_1, d_2, \cdots, d_N, K)\hspace{50mm}\nonumber\\
-  &=& \dfrac{p(\lambdans) \left(\dfrac{p(d)}{p(\lambdans)}\right)^N\prod_{i} p(\lambdans |d_i, K) }{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans}\label{eq:p21},\\
-  &=& p(\lambdans)^{1-N}\times \dfrac{p(d)^N}{\int p(\lambdans) p(d_1, d_2, \cdots, d_N |\lambdans, K)\D\lambdans}\nonumber\\ &&\hspace{3mm}\times\prod_i p(\lambdans |d_i, K)\label{eq:p22}.
+%   &=& \dfrac{p(\lambdans) \left(\dfrac{p(d)}{p(\lambdans)}\right)^N\prod_{i} p(\lambdans |d_i, K) }{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans}\label{eq:p21},\\
+  &=& p(\lambdans)^{1-N}\times \dfrac{p(d)^N}{\int p(\lambdans) p(d_1, d_2, \cdots, d_N |\lambdans, K)\D\lambdans}\nonumber\\ &&\hspace{3mm}\times\prod_i p(\lambdans |d_i, K)\label{eq:p22},
 \end{eqnarray}
 % 
+where $p(\lambdans|K)$ is written $p(\lambdans)$ for brevity.
 The first two factors in Eq.~\ref{eq:p22} can be absorbed into a
@@
 Given that disruptive signatures are strongest for small $\mbh$, we now
-investigate an alternate paradigm~\footnote{We denote our standard paradigm
-which does not respect the mass-gap as paradigm A; with paradigm B being
-the scenario where no black hole masses fall within the mass gap $2-5M_\odot$.}
-with $\mbh\geq 5M_\odot$, consistent
-with the ``mass-gap'' suggested by astronomical observations~\cite{Bailyn:1997xt,
-Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}
+investigate an alternate paradigm in which no black hole masses fall within
+the mass gap $2-5M_\odot$ suggested by astronomical
+observations~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,
+Littenberg:2015tpa}. We will henceforth denote our standard paradigm, which
+does not respect the mass-gap, as paradigm A; with paradigm B being
+this alternate scenario.
 % 
@@
 It is therefore difficult to assess the combined modeling error and its effect
-on our results. they are, therefore, limited by the limitations of our
+on our results. They are, therefore, limited by the limitations of our
 waveform model. However, we expect the combined effect of modeling errors to

### 2016-09-01 Paper/paper.tex
updated fig 7 to corner plots

@@
 NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with $\rho\geq 30$, and $0.04-1$ a
-year with $\rho\geq 70$. For our priors, we choose uniform distribution for 
-both component masses as well as for black hole spins, i.e. $\mbh\in[1.2,25]M_\odot$,
-$\mns\in[1.2,3]M_\odot$ and $-0.75\leq \chibh\leq +0.75$.
+year with $\rho\geq 70$. For our priors, we choose uniform distributions for 
+both component masses, and black hole spin: $\mbh\in[1.2,25]M_\odot$;
+$\mns\in[1.2,3]M_\odot$; and $-0.75\leq \chibh\leq +0.75$.
 % \prayush{FIXME}
@@
 deformability $\lambdans$,}\par
-\textbf{for signals with different $\lambdans$ values (shown in legend).}\par\medskip
+\textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
+from left to right.}\par\medskip
 % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
@@
 \centering
-\includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-\includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-\includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-\includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
-\includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
-\includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
+\textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
+\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
+\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
+% \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+% \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+% \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
 \caption{%
@@
 on component masses and spins remain as in the previous section, while $\lambdans$
-is sampled uniformly from $[0,4000]$.
+is sampled uniformly from $[0,4000]$, and mass-ratio is additionally restricted to
+$2\leq q\leq 6$.
 

### 2016-09-02 Paper/paper.tex
updated and annotated figures
moving fig2 to appendix, smoothening text

@@
 \newcommand{\deff}{D_\mathrm{eff}}
-\newcommand{\arr}{\mathtt{r}}
+\newcommand{\arr}{\mathtt{R}}
 
@@
 \begin{document}
-\linenumbers
+% \linenumbers
 
@@
 \end{figure*}
-% #################
-\begin{figure*}
-\centering 
-\textbf{Statistical measurement uncertainty for NSBH parameters $\mchirp$ (top),
-$\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
-\caption{
-We show here the statistical uncertainty of our measurement of the 
-dimensionless mass ratio $\eta$ (at $90\%$ credibility), over the NSBH
-parameter space.
-Individual panels show the same as a function of BH mass and spin.
-Across each row, we see the effect of increasing signal strength (i.e.
-SNR) with the tidal deformability of the NS $\lambdans$ fixed. Down each
-column, we see the effect of increasing $\lambdans$, at fixed SNR.
-}
-\label{fig:CIWidths90_Lambda_SNR}
-\end{figure*}
+% ################# FIXME FIXME
+% \begin{figure*}
+% \centering 
+% \textbf{Statistical measurement uncertainty for NSBH parameters $\mchirp$ (top),
+% $\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
+% \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
+% \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
+% \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+% \caption{
+% We show here the statistical uncertainty of our measurement of the 
+% dimensionless mass ratio $\eta$ (at $90\%$ credibility), over the NSBH
+% parameter space.
+% Individual panels show the same as a function of BH mass and spin.
+% Across each row, we see the effect of increasing signal strength (i.e.
+% SNR) with the tidal deformability of the NS $\lambdans$ fixed. Down each
+% column, we see the effect of increasing $\lambdans$, at fixed SNR.
+% }
+% \label{fig:CIWidths90_Lambda_SNR}
+% \end{figure*}
 % #################
@@
 % 
-Past and future GW search and source characterization efforts with Advanced LIGO
-have used (or plan to use) BH-BH waveform templates for NS-BH mergers, ignoring
-NS tidal effects on the emitted GWs. In this section we present a fully
-Bayesian analysis of the effects of this assumption on the recovery of NSBH
-masses and spins from their GW signals.
-
-We construct NSBH waveforms using the LEA+ model. We inject these into zero
-noise and run an MCMC sampler on the injections using BBH templates (LEA+ with tidal terms
-set to zero). We perform this analysis for a variety of mass ratios
-$q\in\{2,3,4,5\}$, i.e. $\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$,
-and BH spins $\chibh\in\{-0.5, 0, +0.5, +0.75\}$. We always use $\mns=1.35M_\odot$ and
-$\chins=0$, but explore $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. The signal
-strength is varied from $\rho=20\rightarrow70$. Our choice of this SNR range is
-motivated by previous studies which suggest that signatures of NS tidal effects
-become observable only at relatively large SNRs~\cite{Lackey:2013axa}, and is validated
-by results that follow. At design sensitivity, if we expect $10-300$ NSBH detections 
-a year~\cite{Abadie:2010cfa}, and if BH masses are uniformly likely up to
-$\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we would expect to detect $0.3-7$ disruptive
-NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with $\rho\geq 30$, and $0.04-1$ a
-year with $\rho\geq 70$. For our priors, we choose uniform distributions for 
-both component masses, and black hole spin: $\mbh\in[1.2,25]M_\odot$;
-$\mns\in[1.2,3]M_\odot$; and $-0.75\leq \chibh\leq +0.75$.
+Past (and future) efforts with Advanced LIGO have used (or plan to use) BHBH
+waveform templates to search for and characterize NSBH sources. In doing so,
+they ignore the signature of NS tidal effects on the emitted GWs. In this
+section we present a fully Bayesian analysis of the effects of this
+simplification on the recovery of non-tidal parameters from NSBH signals.
+
+We inject LEA+ NSBH signals into zero noise, and run an MCMC sampler on
+them using equivalent BBH templates (same model, tidal terms $\rightarrow 0$).
+We fix $\mns=1.35M_\odot$ and $\chins=0$, and explore a range of nuclear
+equations of state via the single tidal deformability parameter
+$\lambdans\in\{500, 800, 1000, 1500, 2000\}$. Our injections also span a
+hypercubic regular grid in the BH parameter space, with vertices at
+$q\in\{2,3,4,5\}$, i.e.
+$\mbh\in\{2.7M_\odot, 4.05M_\odot, 5.4M_\odot, 6.75M_\odot\}$, and BH spins
+$\chibh\in\{-0.5, 0, +0.5, +0.75\}$. 
+Finally, we sample all other source-related parameters that determine the
+signal strength but not character\footnote{For aligned-spin signals and
+aligned-spin templates both, we only consider the contribution of the dominant
+$l=|m|=2$ waveform multipoles. This approximation has the additional benefit
+of combining the dependence of the waveforms on inclination, polarization
+and sky location angles, as well as on distance, into the luminosity
+or {\it effective} distance. This quantity only appears as an overall scaling
+factor, and therefore only affects signal strength~\cite{Sathyaprakash:2009xs}.
+}, by sampling the SNR $\rho\in\{20,30,50,70\}$. Our choice of parameters here
+is motivated by two factors: (i) previous studies of the signatures of NS tidal
+effects on gravitational waves~\cite{FoucartEtAl:2011,Foucart:2013psa,
+Foucart:2014nda} (which suggest that necessary conditions for the observation
+of tidal effects with aLIGO include high SNRs and a low-mass spinning companion
+BH); and (ii) technical constraints of our chosen LEA+ model~\cite{Lackey:2013axa}.
+At design sensitivity, if we expect $10-300$ NSBH detections a 
+year~\cite{Abadie:2010cfa}, {\it and} if BH mass values are uniformly likely
+up to $\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we can expect to detect $0.3-7$
+disruptive NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with
+$\rho\geq 30$, and $0.04-1$ a year with $\rho\geq 70$. Therefore, the parameter
+range spanned by our injections represents a physically interesting class of
+systems that is {\it also} likely observable in the near future. 
+For our Bayesian priors, we choose uniform distributions for both component
+masses and black hole spin: $\mbh\in[1.2,25]M_\odot$; $\mns\in[1.2,3]M_\odot$;
+and $-0.75\leq \chibh\leq +0.75$.
 % \prayush{FIXME}
@@
 As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show the
-probability distribution for binary mass ratio $\eta$ recovered from a GW
-injection with intrinsic parameters: $\mns=1.35M_\odot$, $\chins=0$,
-$\mbh=5.4M_\odot$, $\chibh=+0.5$ and $\lambdans=2000$, and SNRs ranging
-from $\rho=20$ (left panel) to $\rho=50$ (right panel). In each panel, the
-{\it true} value of $\eta$ is marked by a dashed green line, and its measured
-(median) value by a dashed red line. The effect of ignoring tidal corrections
-in templates manifests as a systematic shift of the measured $\eta$ value away
-from its true value. The associated {\it systematic} bias/error is given by the
-difference between measured and true $\eta$ values. Darker shading of the
-probability distribution marks the $90\%$ credible interval for $\eta$, whose
-width is a direct measure of the {\it statistical} measurement uncertainty/error.
-% 
-We clearly see that even when the signal is moderately loud, with $\rho=20$,
-statistical errors dominate over systematics completely for $\eta$.
-As we turn up the SNR
+probability distribution for binary mass ratio $\eta$ recovered for an NSBH
+injection with $\mns=1.35M_\odot$, $\chins=0$, $\mbh=5.4M_\odot$, $\chibh=+0.5$
+and $\lambdans=2000$, with the SNR ranging from $\rho=20$ (left) to
+$\rho=50$ (right). In each panel, both the {\it true} and the {\it 
+recovered} (median) values of $\eta$ are marked. The effect of ignoring tidal
+corrections in templates manifests as a systematic shift of the measured $\eta$
+value away from its true value. The associated {\it systematic} bias/error is
+measured by the difference between the measured and true $\eta$ values. Darker
+shading of the probability distributions marks the $90\%$ credible interval,
+whose width $(\Delta\eta)^{90.0\%}$ is a direct measure of the
+{\it statistical} measurement uncertainty/error\footnote{We generalize the
+notation $(\Delta X)^{90.0\%}$ to mean the $90\%$ credible interval width
+for the measured source parameter $X$.}.
+% 
+We see clearly that even when the signal is moderately loud, with $\rho=20$,
+statistical errors dominate over systematics for $\eta$. As we turn up the SNR
 further, we find that the two error sources become comparable at $\rho\sim 30$,
@@
 
+\textcolor{gray}{%
 Next, to get a better understanding of how {\it precisely} we measure NSBH
-binary parameters given BBH waveforms, we show the $90\%$ credible intervals
-associated with NSBH mass and spin measurements in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
-The three panels, corresponding to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
-(bottom), show the width of these credible intervals as a function of BH
-mass/spin (within each sub-panel), and NS properties, i.e. $\lambdans$
-(downwards in each column)~\footnote{We restrict NS mass to $1.35M_\odot$ and
-its spin to zero. Varying its tidal deformability $\lambdans$ does not change
-the measurement uncertainties for non-tidal binary parameters, as is evident
-from comparing the two rows in each panel of Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
+parameters using BBH templates, we show $90\%$ credible intervals associated
+with NSBH mass and spin measurements in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
+The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
+(bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
+as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
+$\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
+$1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
+does not significantly change the measurement uncertainties for non-tidal
+binary parameters, as is evident from comparing the two rows in each panel of
+Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
 % 
@@
 % 
-However, we remind ourselves that there are physical effects that have not been
-incorporated in our (BBH) search templates which we use here to decipher NSBH
-signals. Therefore, we cannot know the {\it accuracy} of our parameter
-measurements unless we know how large the systematic biases due to an
-incomplete template model are. To quantify this, we find it useful to define
-$\arr$, the ratio of the systematic biases induced by unmodeled tidal effects in
-templates, to the statistical uncertainties associated with the measurement
-itself. If $\arr\gtrsim 1$, including the effect of the tidal distortion of the NS
-(during inspiral) and possible disruption (close to merger) would be imperative
-to obtain {\it accurate} estimates of source masses and spins from observed GW
+We remind ourselves that this {\it precision}, however, is only
+meaningful so long as the measurement is {\it accurate} to begin with. Due to
+unmodeled tidal effects, parameter measurements performed on NSBH signals with
+BBH templates will have systematic biases. If these biases are too large,
+using non-tidal BBH templates should severely impair our inferencing ability.
+In order to quantify the relevance of tidal signatures, we define $\arr_X$ as
+the ratio between systematic biases and statistical uncertainties associated
+with the measurement of parameter $X$. If $\arr_X\gtrsim 1$, incorporating
+tidal effects in inspiral and merger modeling would non-trivially improve our
+estimation of parameter $X$ from NSBH signals.
+% 
+% there are physical effects that have not been
+% incorporated in our (BBH) search templates which we use here to decipher NSBH
+% signals. Therefore, we cannot know the {\it accuracy} of our parameter
+% measurements unless we know how large the systematic biases due to an
+% incomplete template model are. To quantify this, we find it useful to define
+% $\arr_X$, the ratio of the systematic biases induced by unmodeled tidal effects in
+% templates, to the statistical uncertainties associated with the measurement
+% itself. If $\arr_X\gtrsim 1$, including the effects of NS tidal distortion
+% (during inspiral) and possible disruption (close to merger) would be imperative
+% to obtain {\it accurate} estimates of source masses and spins from observed GW
+% signals.}
+}
+
+\textcolor{blue}{%
+Credible intervals $(\Delta X)^{90\%}$ measure the precision with which a 
+quantity $X$ can be measured. We show credible intervals for
+$X=\{\mchirp,\eta,\chibh\}$ in Appendix~\ref{as1:illustrations}. This {\it 
+precision}, however, is only meaningful so long as the measurement is {\it
+accurate} to begin with. Due to unmodeled tidal effects, parameter measurements
+performed on NSBH signals using BBH templates will have systematic biases. If
+these biases are too large, using non-tidal BBH templates should severely impair
+our
+inferencing ability. In order to quantify the relevance of tidal signatures,
+we define $\arr_X$ as the ratio between systematic biases and statistical
+uncertainties associated with the measurement of parameter $X$. If
+$\arr_X\gtrsim 1$, incorporating tidal effects in inspiral and merger modeling
+would non-trivially improve our estimation of parameter $X$ from NSBH
 signals.
-
-We first calculate the ratio of errors $\arr$ associated with the measurement of
-chirp mass, $\mchirp := (m_1 m_2)^{3/5} (m_1 + m_2)^{-1/5}$, and show it in
+}
+
+We first calculate the ratio of errors $\arr_{\mchirp}$ , and show it in
 Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the mass combination
@@
 \centering    
+\textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
 \includegraphics[trim={2cm 0 0 0},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
@@
 \centering    
-\includegraphics[width=1.05\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}
-\includegraphics[width=1.\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+\textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
+\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
+\includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
 \caption{%
@@
 \centering    
-\includegraphics[width=1.05\columnwidth]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
+\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
 \caption{%
@@
 \centering
-\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
-\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}\\
-\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian90pc_vs_N_AllPopulation}
-\includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian90pc_vs_N_AstroPopulation}\\
+\textbf{Paradigm A \hspace{6cm} Paradigm B}\par\medskip
+\includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
+\includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}\\
+\includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{plots/LambdaMedian90pc_vs_N_AllPopulation}
+\includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{plots/LambdaMedian90pc_vs_N_AstroPopulation}\\
 \caption{%
@@
 \centering    
-\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AllPopulation}
-\includegraphics[width=\columnwidth]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation}
+\textbf{Paradigm A \hspace{6cm} Paradigm B}\par\medskip
+\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{plots/LambdaCIWidths90pc_vs_N_AllPopulation}
+\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation}
 \caption{%
@@
 is a combination of distance to the source, its orientation, and its sky
-location angles, and has a one-to-one correspondence with SNR for non-precessing
-systems, for which these angles are constant in time.}.
+location angles; and has a one-to-one correspondence with SNR for non-precessing
+sources. This is so because for such sources, their location and orientation
+remain constant over the timescales within which they sweep through
+aLIGO's sensitive frequency band.}.
 $D^\mathrm{max}$. $D^\mathrm{max}$ is set by the minimum SNR 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{appendix}
 
+\section{Illustrations of Bayesian posteriors}\label{as1:illustrations}
+% 
+% #################
+\begin{figure*}
+\centering 
+\textbf{Statistical measurement uncertainty for NSBH parameters $\mchirp$ (top),
+$\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+\caption{
+We show here the statistical uncertainty of our measurement of the 
+dimensionless mass ratio $\eta$ (at $90\%$ credibility), over the NSBH
+parameter space.
+Individual panels show the same as a function of BH mass and spin.
+Across each row, we see the effect of increasing signal strength (i.e.
+SNR) with the tidal deformability of the NS $\lambdans$ fixed. Down each
+column, we see the effect of increasing $\lambdans$, at fixed SNR.
+}
+\label{fig:CIWidths90_Lambda_SNR}
+\end{figure*}
+% #################
+%%
+\begin{figure*}
+\centering
+\textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
+\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
+\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
+% \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+% \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+% \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+% \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
+\caption{%
+We illustrate here two-dimensional joint probability distributions, as
+recovered for the intrinsic parameters of a binary with
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
+$\lambdans=2000$, and $\rho=30$ (same as in
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
+The top row shows the same for combinations of 
+non-tidal parameters, i.e. binary masses and spins, while the middle row 
+shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
+is similar to the middle, with the only difference being that the SNR has been
+turned up to $\rho=50$.
+}
+\label{fig:SingleSystemLambda2DPDFs}
+\end{figure*}
+% #################
+
+% #################
+\begin{figure*}
+\centering    
+\textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
+\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
+\includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+\caption{%
+These panels show the accumulation of information from a population whose
+$\lambdans=800$.
+{\it Left:} Posterior probability distributions for $\lambdans$ (colored 
+curves), and associated $90\%$ credible intervals (grey vertical lines),
+shown for different number of accumulated observations $N$. Distributions are
+normalized to unit area.
+{\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
+associated $90\%$ credible intervals (as the vertical extent of filled
+region), shown as a function of number of observations $N$. Solid horizontal
+line indicates the true value of $\lambdans$. Dashed and dotted horizontal
+lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
+bounds.
+}
+\label{fig:TT_Lambda_vs_N_L800_CI90_0}
+\end{figure*}
+%
+% #################
+
+% ###########################################################
+% ###########################################################
 \section{Phenomenology of $\lambdans$ measurement errors}
+% ###########################################################
 % 

### 2016-09-03 Paper/paper.tex
s3 smoothened

@@
 % 
-Past (and future) efforts with Advanced LIGO have used (or plan to use) BHBH
+Past (and future) efforts with Advanced LIGO have used (or plan to use) BBH
 waveform templates to search for and characterize NSBH sources. In doing so,
@@
 them using equivalent BBH templates (same model, tidal terms $\rightarrow 0$).
-We fix $\mns=1.35M_\odot$ and $\chins=0$, and explore a range of nuclear
+We fix $\mns=1.35M_\odot$ and $\chins=0$, and explore a range of NS
 equations of state via the single tidal deformability parameter
 $\lambdans\in\{500, 800, 1000, 1500, 2000\}$. Our injections also span a
-hypercubic regular grid in the BH parameter space, with vertices at
+rectangular grid in the BH parameter space, with vertices at
 $q\in\{2,3,4,5\}$, i.e.
@@
 factor, and therefore only affects signal strength~\cite{Sathyaprakash:2009xs}.
-}, by sampling the SNR $\rho\in\{20,30,50,70\}$. Our choice of parameters here
+}, by sampling the SNR $\rho\in\{20,30,50,70\}$. Our choice of injection
+parameters here
 is motivated by two factors: (i) previous studies of the signatures of NS tidal
@@
 up to $\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we can expect to detect $0.3-7$
-disruptive NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with
+{\it disruptive} NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with
 $\rho\geq 30$, and $0.04-1$ a year with $\rho\geq 70$. Therefore, the parameter
@@
 
-As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show the
-probability distribution for binary mass ratio $\eta$ recovered for an NSBH
-injection with $\mns=1.35M_\odot$, $\chins=0$, $\mbh=5.4M_\odot$, $\chibh=+0.5$
+As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show 
+recovered probability distributions for binary mass ratio $\eta$, for NSBH
+injections with $\mns=1.35M_\odot$, $\chins=0$, $\mbh=5.4M_\odot$, $\chibh=+0.5$
 and $\lambdans=2000$, with the SNR ranging from $\rho=20$ (left) to
@@
 notation $(\Delta X)^{90.0\%}$ to mean the $90\%$ credible interval width
-for the measured source parameter $X$.}.
+for any measured source parameter $X$.}.
 % 
@@
 \textcolor{gray}{%
-Next, to get a better understanding of how {\it precisely} we measure NSBH
-parameters using BBH templates, we show $90\%$ credible intervals associated
-with NSBH mass and spin measurements in Fig.~\ref{fig:CIWidths90_Lambda_SNR}.
+Let us first delve into statistical errors. In
+Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
+measure NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
 The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
@@
 to single-digit percents of its true value.
-% 
-We remind ourselves that this {\it precision}, however, is only
-meaningful so long as the measurement is {\it accurate} to begin with. Due to
-unmodeled tidal effects, parameter measurements performed on NSBH signals with
-BBH templates will have systematic biases. If these biases are too large,
-using non-tidal BBH templates should severely impair our inferencing ability.
-In order to quantify the relevance of tidal signatures, we define $\arr_X$ as
-the ratio between systematic biases and statistical uncertainties associated
-with the measurement of parameter $X$. If $\arr_X\gtrsim 1$, incorporating
-tidal effects in inspiral and merger modeling would non-trivially improve our
-estimation of parameter $X$ from NSBH signals.
+% % 
+% We remind ourselves that this {\it precision}, however, is only
+% meaningful so long as the measurement is {\it accurate} to begin with.
+% We define $\arr_X$ as the ratio between systematic and statistical
+% errors associated with the measurement of parameter $X$,
+% \begin{equation}\label{eq:arr}
+% \arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
+% \end{equation}
+% in order to compare the relative importance of both. Only when
+% $|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
+% will not hamper the measurement of non-tidal parameters from NSBH signals.
+% Due to
+% unmodeled tidal effects, parameter measurements performed on NSBH signals with
+% BBH templates will have systematic biases. If these biases are too large,
+% using non-tidal BBH templates should severely impair our inferencing ability.
 % 
@@
 \textcolor{blue}{%
-Credible intervals $(\Delta X)^{90\%}$ measure the precision with which a 
-quantity $X$ can be measured. We show credible intervals for
-$X=\{\mchirp,\eta,\chibh\}$ in Appendix~\ref{as1:illustrations}. This {\it 
-precision}, however, is only meaningful so long as the measurement is {\it
-accurate} to begin with. Due to unmodeled tidal effects, parameter measurements
-performed on NSBH signals using BBH templates will have systematic biases. If
-these biases are too large, using non-tidal BBH templates should severely impair
-our
-inferencing ability. In order to quantify the relevance of tidal signatures,
-we define $\arr_X$ as the ratio between systematic biases and statistical
-uncertainties associated with the measurement of parameter $X$. If
-$\arr_X\gtrsim 1$, incorporating tidal effects in inspiral and merger modeling
-would non-trivially improve our estimation of parameter $X$ from NSBH
-signals.
-}
-
-We first calculate the ratio of errors $\arr_{\mchirp}$ , and show it in
-Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the mass combination
-that affects at leading order the GW strain emitted
-by compact binaries as they spiral in, and is therefore determined the most
-precisely. From the left-most column of the figure, we find that even at $\rho=30$, 
-the systematic bias in $\mchirp$ measurement is much smaller than its statistical
-uncertainty. For signals with $\rho\leq 30$, therefore, we are justified in using
-BBH templates to estimate binary chirp mass from NSBH signals. For louder and less
-realistic SNRs (say $\rho=50$), we find that $\arr$ can become comparable to unity
-if additionally: (i) the BH has a prograde spin $\chibh\gtrsim 0.4$, {\it and} (ii)
-the true tidal deformability of the NS is large enough, s.t. $\lambdans \gtrsim 1000$.
-% 
-It is only at SNRs as high as $\rho\simeq70$ that the systematic biases for $\mchirp$
-exceed the underlying statistical uncertainty to become the dominant source of
-errors.
-% 
-We conclude that only for $\rho\gtrsim 50-70$ do we need to include in our templates
-terms that capture the effect of neutron star's structural dynamics on emitted GWs.
-For lower SNRs, improvements in template model will be washed out by the statistical
-measurement uncertainty for $\mchirp$.
-% 
-Lastly, we note that $\arr$ is always positive for $\mchirp$, suggesting that
-it is always over-estimated.
-This is understandable since the tidal deformation of NSs close to merger reduces
-the GW signal power at high frequencies, which would make the resulting signal
-resemble a BBH signal of higher mass (and therefore lower merger frequency).
+Credible intervals $(\Delta X)^{90\%}$ measuring the precision with which
+$X=\{\mchirp,\eta,\chibh\}$ can be measured, are discussed in detail in
+Appendix~\ref{as1:illustrations}.
+}
+We remind ourselves that this {\it precision} is only meaningful so long as the
+measurement is {\it accurate} to begin with. Therefore, we define $\arr_X$ as
+the ratio between systematic and statistical errors associated with the
+measurement of parameter $X$,
+\begin{equation}\label{eq:arr}
+\arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
+\end{equation}
+in order to compare the relative importance of both. Only when
+$|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
+will not hamper the measurement of non-tidal parameters from NSBH signals.
+% 
+We start with calculating $\arr_{\mchirp}$, and show it in
+Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the leading order
+mass combination that affects the GW strain emitted by compact binaries as they
+spiral in, and is therefore determined the most precisely. We notice
+immediately that for $\rho\leq 30$ the systematics are well under control
+and so we can obtain reliable chirp mass estimates for NSBH signals using BBH
+templates.
+% 
+For louder and less likely SNRs ($\rho\simeq 50$), we show that
+$\arr_{\mchirp}$ can become comparable to unity, but only if the BH has
+prograde spin $\chibh\gtrsim 0.4$, {\it and} the true NS tidal deformability
+is large enough, s.t. $\lambdans \gtrsim 1000$.
+% 
+We therefore conclude that only for very loud signals, with
+$\rho\gtrsim 50-70$, will the inclusion of tidal terms in our template
+models improve our $\mchirp$ estimation. For lower SNRs, inclusion of new
+physical content in templates will instead get washed out by detector noise.
+% 
+In addition to everything, we also note that $\arr_{\mchirp}\geq 0$ always,
+i.e. $\mchirp$ is always being over-estimated. This is to be expected since
+the tidal deformation and disruption of NSs close to merger reduces GW signal
+power at high frequencies, making the resulting signal resemble a BBH signal of
+lower merger frequency - and therefore of higher mass.
 
 Next, in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
-measurement errors $\arr$ for the symmetric mass-ratio $\eta$. We immediately
-observe that the measured (median) $\eta$ is always lowered below its true
-value due to the lack of $\lambdans$-dependent terms in the filter template
-model. This bias, similar to the $\mchirp$ bias, reduces the merger frequency
-for BBH templates, in order to better match the relatively quiet merger and
-ringdown of NSBHs. 
-% 
-At $\rho=30$, the systematic bias in $\eta$ measurement stays mostly around
-$10-50\%$ of its statistical measurement uncertainty. For the sole case of a
-very deformable NS with $\lambdans=2000$, can the effect of missing tidal pieces
-in the template model become comparable to that of noise (which is usually
-more restrictive on measurements).
-% 
-For a still stronger signal with $\rho=50$, we find that when the following
-conditions are fulfilled, the measurement of $\eta$ is unreliable due to
-template biases: (a) BH mass $\mbh\leq 5M_\odot$, (b) BH spin
-$\chibh\gtrsim +0.4$, and (c) $\lambdans\gtrsim 1500$. Even under more
-moderate restrictions on BH and NS parameters, we find that $\rho=50$ is 
-loud enough to motivate the use of tidal-modeled templates. This observation
-is further strengthened if we divert our attention to the third column in
-Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR} that presents
-the case for even higher SNRs ($\rho=70$).
-
-Moving on from mass to spin parameters, we now consider the measurement of spin
-angular momentum of the BH $\chibh$. The ratio of the systematic bias introduced into
-and the statistical uncertainty associated with $\chibh$ measurement is shown in
+measurement errors $\arr_\eta$ for the symmetric mass-ratio. We straight-away
+notice that, unlike for $\mchirp$, the measured (median) value of $\eta$ is
+always {\it lower} than its true value due to the lack of $\lambdans$-dependent
+terms in filter templates. This also is so because templates with $\arr_\eta<0$
+will have a lower merger frequency than the one with $\arr_\eta=0$, and will
+therefore better fit a disruptive NSBH merger.
+% 
+Going through the figure from left to right, we find that for realistic
+SNRs ($\rho\leq 30$) the systematics remain below statistical errors for
+$\eta$ measurement. The worst case is for the most deformable NSs
+($\lambdans = 2000$), but even for them systematics in $\eta$ are $2\times$
+smaller than the statistical measurement errors.
+% 
+Moving to louder signals with $\rho\simeq 50$, we find that for binaries of
+fairly deformable NSs ($\lambdans\gtrsim 1500$), and low-mass BHs
+($\mbh\leq 5M_\odot$) that have prograde spins ($\chibh\gtrsim +0.4$), our
+measurement of mass-ratio can be seriously compromised by ignoring tidal
+physics in template models. 
+% 
+This pattern is continued to even higher SNRs, as we can see from
+Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. We therefore conclude that, even if
+under moderate restrictions on BH and NS parameters, $\rho=50$ is loud enough
+to motivate the use of tidal templates in aLIGO data analyses.
+
+Moving on from mass to spin parameters, we now consider the measurement of BH
+spin angular momentum $\chibh$. The ratio of systematic and statistical errors
+associated with $\chibh$ measurement is shown in
 Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}. The presentation of information in this 
-figure is identical to that in Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
-and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. A diverging colormap is used because both 
+figure is identical to that of Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
+and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. A diverging colormap is used so that both 
 extremes of the colorbar range point to large systematic biases, while its zero (or
-small) value lies in the middle. For the lowest considered SNR $\rho=30$, $\chibh$
-bias is negligible, being below $50\%$ of its statistical measurement uncertainty.
-For the most deformable NS structure {\it and} with a low-mass BH, the statistical
-bias in $\chibh$ does become an important error. This is furthered at higher SNR, 
-considering $\rho=50$ next (middle column of Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}).
-We find that for low-mass BHs (with masses below $4.5M_\odot$), the systematic 
-measurement bias for BH spin exceeds its statistical uncertainty to become the 
-dominant source of error. Not surprisingly, the systematic bias at both spin 
-extremes pushes towards milder spins, since the prior excludes any more extreme
-spins than $[-0.75, +0.75]$ and that pushes the median of the posterior inwards.
-However as the signal's $\lambdans$ is increased from $500\rightarrow 2000$,
-low-mass BH-NS binaries with large spins on the BH show very large systematic biases
-in spin values, that become larger than the statistical uncertainty for $\rho\geq 30$
-and reach up to $4\times$ the same quantity as $\rho\rightarrow 70$.
+small) value lies in the middle.
+% 
+For the lowest considered SNR ($\rho=30$), $\chibh$ bias is about $2\times$
+smaller than the statistical measurement uncertainty and is therefore mostly
+negligible. Both can become comparable only when we have the most deformable NSs
+in orbit around low-mass BHs. 
+% 
+Moving to higher SNRs, at $\rho\simeq50$ we find that systematics in $\chibh$
+measurement begin to dominate for $\mbh\leq 4.5M_\odot$.
+This suggests that we would need tidal templates to accurately
+characterize astrophysical mass-gap violating NSBH binaries.
+% 
+At even higher SNRs, the systematics can completely dominate BH spin
+measurement.
+% 
+From Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR} we also note that when signal
+spins are the largest, $\chibh\times\arr_{\chibh}<0$. This is to be
+expected because the median of the recovered posterior distribution for $\chibh$
+can only get pushed inwards from the boundaries.
 
@@
 signal-to-noise ratio of $30$, our measurements of mass and spin parameters of 
-astrophysical BHNS binaries will remain limited by the intrinsic uncertainty due to 
+astrophysical NSBH binaries will remain limited by the intrinsic uncertainty due to 
 instrument noise, and do not depend on whether we include tidal effects in template
 models. However, when the signal-to-noise ratio exceeds $30$ the systematic bias in
-binary mass and spin measurements become comparable to and exceed the uncertainty
-due to noise. Of the different parameters considered, we find that the measurement 
-of $\eta$ degrades worst (in a relative-error sense) amongst all other parameters
-of BHNS binaries, due to the use of BH-BH waveforms in deciphering an NSBH signal.
+binary mass and spin measurements become comparable to and can exceed the
+uncertainty due to noise. Of the different parameters considered, we find that
+amongst all non-tidal parameters, the measurement of $\eta$ degrades worst (in a 
+relative-error sense) due to the use of BBH templates in deciphering an NSBH
+signal.
 
@@
 
+Let us first delve into statistical errors. In
+Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
+measure NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
+The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
+(bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
+as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
+$\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
+$1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
+does not significantly change the measurement uncertainties for non-tidal
+binary parameters, as is evident from comparing the two rows in each panel of
+Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
+% 
+From the left-most column, we find that: (i) at $\rho=20, \mchirp$ is already
+measured remarkably well - to a precision of $0.16\%$ of its true value, and
+(ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined more loosely,
+with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
+all three measurements gain further precision, with $\eta$ errors shrinking down
+to single-digit percents of its true value.
+% % 
+% We remind ourselves that this {\it precision}, however, is only
+% meaningful so long as the measurement is {\it accurate} to begin with.
+% We define $\arr_X$ as the ratio between systematic and statistical
+% errors associated with the measurement of parameter $X$,
+% \begin{equation}\label{eq:arr}
+% \arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
+% \end{equation}
+% in order to compare the relative importance of both. Only when
+% $|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
+% will not hamper the measurement of non-tidal parameters from NSBH signals.
+% Due to
+% unmodeled tidal effects, parameter measurements performed on NSBH signals with
+% BBH templates will have systematic biases. If these biases are too large,
+% using non-tidal BBH templates should severely impair our inferencing ability.
+% 
+% there are physical effects that have not been
+% incorporated in our (BBH) search templates which we use here to decipher NSBH
+% signals. Therefore, we cannot know the {\it accuracy} of our parameter
+% measurements unless we know how large the systematic biases due to an
+% incomplete template model are. To quantify this, we find it useful to define
+% $\arr_X$, the ratio of the systematic biases induced by unmodeled tidal effects in
+% templates, to the statistical uncertainties associated with the measurement
+% itself. If $\arr_X\gtrsim 1$, including the effects of NS tidal distortion
+% (during inspiral) and possible disruption (close to merger) would be imperative
+% to obtain {\it accurate} estimates of source masses and spins from observed GW
+% signals.}
+
 % ###########################################################

### 2016-09-05 Paper/paper.tex
improved s3

@@
 they ignore the signature of NS tidal effects on the emitted GWs. In this
-section we present a fully Bayesian analysis of the effects of this
+section we present a fully Bayesian analysis of the effect of this
 simplification on the recovery of non-tidal parameters from NSBH signals.
@@
 $\chibh\in\{-0.5, 0, +0.5, +0.75\}$. 
-Finally, we sample all other source-related parameters that determine the
+Finally, we sample all other source-related parameters, that determine the
 signal strength but not character\footnote{For aligned-spin signals and
@@
 {\it disruptive} NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with
-$\rho\geq 30$, and $0.04-1$ a year with $\rho\geq 70$. Therefore, the parameter
-range spanned by our injections represents a physically interesting class of
-systems that is {\it also} likely observable in the near future. 
+$\rho\geq 30$, and $0.04-1$ a year with $\rho\geq 70$. Therefore, our injection
+parameters span a physically interesting subset of NSBH binaries, that is
+{\it also} likely observable in the near future. 
 For our Bayesian priors, we choose uniform distributions for both component
@@
 
-As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show 
-recovered probability distributions for binary mass ratio $\eta$, for NSBH
-injections with $\mns=1.35M_\odot$, $\chins=0$, $\mbh=5.4M_\odot$, $\chibh=+0.5$
-and $\lambdans=2000$, with the SNR ranging from $\rho=20$ (left) to
-$\rho=50$ (right). In each panel, both the {\it true} and the {\it 
-recovered} (median) values of $\eta$ are marked. The effect of ignoring tidal
+As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show the
+recovered probability distributions for binary mass ratio $\eta$ for
+three NSBH injections, with $\rho=20$ (left), $30$ (middle), and $\rho=50$
+(right), and other parameters held fixed ($\mns=1.35M_\odot$, $\chins=0$,
+$\mbh=5.4M_\odot$, $\chibh=+0.5$ and $\lambdans=2000$). In each panel, both
+the {\it true} and the {\it measured} (i.e. median) values of $\eta$ are marked.
+The effect of ignoring tidal
 corrections in templates manifests as a systematic shift of the measured $\eta$
@@
 \textcolor{blue}{%
-Credible intervals $(\Delta X)^{90\%}$ measuring the precision with which
-$X=\{\mchirp,\eta,\chibh\}$ can be measured, are discussed in detail in
+Credible intervals $(\Delta X)^{90\%}$ showing the precision with which
+$X=\{\mchirp,\eta,\chibh\}$ can be measured, are presented in
 Appendix~\ref{as1:illustrations}.
@@
 \end{equation}
-in order to compare the relative importance of both. Only when
-$|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
-will not hamper the measurement of non-tidal parameters from NSBH signals.
-% 
-We start with calculating $\arr_{\mchirp}$, and show it in
-Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}. $\mchirp$ is the leading order
+in order to compare the relative magnitude of both. Only when
+$|\arr_X| < 1$ can we ignore tidal effects in our templates
+without hampering the measurement of non-tidal parameters from NSBH signals.
+% 
+We start with calculating $\arr_{\mchirp}$ as a function of various source
+parameters and show it in Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}.
+$\mchirp$ is the leading order
 mass combination that affects the GW strain emitted by compact binaries as they
@@
 % 
-For louder and less likely SNRs ($\rho\simeq 50$), we show that
+For louder and less likely SNRs ($\rho\simeq 50$), we find that
 $\arr_{\mchirp}$ can become comparable to unity, but only if the BH has
@@
 % 
-We therefore conclude that only for very loud signals, with
+We can therefore conclude that only for very loud signals, with
 $\rho\gtrsim 50-70$, will the inclusion of tidal terms in our template
@@
 Next, in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
-measurement errors $\arr_\eta$ for the symmetric mass-ratio. We straight-away
-notice that, unlike for $\mchirp$, the measured (median) value of $\eta$ is
-always {\it lower} than its true value due to the lack of $\lambdans$-dependent
-terms in filter templates. This also is so because templates with $\arr_\eta<0$
-will have a lower merger frequency than the one with $\arr_\eta=0$, and will
-therefore better fit a disruptive NSBH merger.
+measurement errors $\arr_\eta$ for the symmetric mass-ratio.
 % 
@@
 SNRs ($\rho\leq 30$) the systematics remain below statistical errors for
-$\eta$ measurement. The worst case is for the most deformable NSs
+$\eta$ measurement. The worst case is of the most deformable NSs
 ($\lambdans = 2000$), but even for them systematics in $\eta$ are $2\times$
@@
 Moving to louder signals with $\rho\simeq 50$, we find that for binaries of
-fairly deformable NSs ($\lambdans\gtrsim 1500$), and low-mass BHs
+fairly deformable NSs ($\lambdans\gtrsim 1500$) and low-mass BHs
 ($\mbh\leq 5M_\odot$) that have prograde spins ($\chibh\gtrsim +0.4$), our
@@
 % 
-This pattern is continued to even higher SNRs, as we can see from
+This pattern is continued at even higher SNRs, as we can see from
 Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. We therefore conclude that, even if
@@
 to motivate the use of tidal templates in aLIGO data analyses.
+% 
+In addition, we also notice that, unlike for $\mchirp$, the measured value of
+$\eta$ is always {\it lower} than its true value, which is what we expect if we
+want BBH templates to fit NSBHs that disrupt and merge at lower frequencies.
 
@@
 % 
-For the lowest considered SNR ($\rho=30$), $\chibh$ bias is about $2\times$
-smaller than the statistical measurement uncertainty and is therefore mostly
-negligible. Both can become comparable only when we have the most deformable NSs
-in orbit around low-mass BHs. 
-% 
-Moving to higher SNRs, at $\rho\simeq50$ we find that systematics in $\chibh$
-measurement begin to dominate for $\mbh\leq 4.5M_\odot$.
-This suggests that we would need tidal templates to accurately
-characterize astrophysical mass-gap violating NSBH binaries.
-% 
-At even higher SNRs, the systematics can completely dominate BH spin
-measurement.
-% 
-From Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR} we also note that when signal
-spins are the largest, $\chibh\times\arr_{\chibh}<0$. This is to be
-expected because the median of the recovered posterior distribution for $\chibh$
-can only get pushed inwards from the boundaries.
+For the lowest SNR considered ($\rho=30$), $\chibh$ bias is about $2\times$
+smaller than its statistical measurement uncertainty, and is therefore mostly
+negligible. Both do become somewhat comparable, but only when we have the most
+deformable NSs in orbit around low-mass BHs. 
+% 
+At higher SNRs $(\rho\simeq50-70)$, we find that the systematics in $\chibh$
+measurement can dominate completely, especially for binaries containing
+mass-gap violating BHs and/or deformable NSs with $\lambdans\geq1000$.
+% 
+From Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR} we additionally note that when
+the source spin magnitudes approach the highest allowed, i.e. at both extremes
+of the $x$-axes, $\chibh\times\arr_{\chibh}<0$. This is to be expected because
+the median of the recovered posterior distributions for $\chibh$ can only get
+pushed inwards from the boundaries.
 

### 2016-09-05 Paper/paper.tex
improved s3 more

@@
 Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. We therefore conclude that, even if
-under moderate restrictions on BH and NS parameters, $\rho=50$ is loud enough
+under moderate restrictions on BH and NS parameters, $\rho=30-50$ is loud enough
 to motivate the use of tidal templates in aLIGO data analyses.
@@
 spin angular momentum $\chibh$. The ratio of systematic and statistical errors
-associated with $\chibh$ measurement is shown in
+for $\chibh$ are shown in
 Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}. The presentation of information in this 
@@
 
-Summarizing these results, we find that irrespective of system parameters, below a
-signal-to-noise ratio of $30$, our measurements of mass and spin parameters of 
-astrophysical NSBH binaries will remain limited by the intrinsic uncertainty due to 
-instrument noise, and do not depend on whether we include tidal effects in template
-models. However, when the signal-to-noise ratio exceeds $30$ the systematic bias in
-binary mass and spin measurements become comparable to and can exceed the
-uncertainty due to noise. Of the different parameters considered, we find that
-amongst all non-tidal parameters, the measurement of $\eta$ degrades worst (in a 
-relative-error sense) due to the use of BBH templates in deciphering an NSBH
-signal.
+Summarizing these results, we find that irrespective of system parameters,
+below a signal-to-noise ratio of $30$, our measurements of mass and spin
+parameters of astrophysical NSBH binaries will remain limited by the intrinsic
+uncertainty due to instrument noise, and do not depend on whether we include
+tidal effects in template models. However, when the signal-to-noise ratio
+exceeds $30$ the systematic bias in binary mass and spin measurements become
+comparable to and can exceed the uncertainty due to noise. Of the different
+non-tidal parameters considered, we find that the measurement of $\eta$
+degrades worst (in a relative-error sense) due to the use of BBH templates in
+deciphering an NSBH signal. Of all the sub-categories, we find that tidal
+templates could especially help with the parameter estimation of astrophysical
+{\it mass-gap violating} NSBH binaries,
 

### 2016-09-05 Paper/paper.tex
improving s4

@@
 % THESE FIGURES ARE FOR THE NEXT SECTION
+\begin{figure*}
+\centering 
+\textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
+\textbf{at different SNR values (shown in legend).}\par\medskip
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
+\caption{We show here probability distributions for mass ratio $\eta$ as measured
+for the same injection at different SNRs. The intrinsic parameters of the source
+are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
+and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
+{\it ignore} tidal effects.
+% 
+In each panel: the dashed red line marks the median {\it measured} value
+$\eta^\mathrm{Median}$, while the dashed green line show the true value
+$\eta^\mathrm{Injected}$. The difference between the true and measured values
+is the systematic bias introduced
+by ignoring tidal effects in templates. The darker shading shows
+the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
+which is a direct measure of our statistical uncertainty on $\eta$.
+% 
+Comparing systematic and statistical errors, we find that:
+at $\rho=20$, $\eta$ measurement is dominated by statistical
+errors; at $\rho=30$, the two become comparable; and 
+for louder signals ($\rho\simeq50$), the systematic errors dominate.
+}
+\label{fig:SingleSystemEtaPDFvsSNR}
+\end{figure*}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\centering 
-\textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
-\textbf{at different SNR values (shown in legend).}\par\medskip
-\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
-\caption{We show here probability distributions for mass ratio $\eta$ as measured
-for the same injection at different SNRs. The intrinsic parameters of the source
-are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
-and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
-{\it ignore} tidal effects.
-% 
-In each panel: the dashed red line marks the median {\it measured} value
-$\eta^\mathrm{Median}$, while the dashed green line show the true value
-$\eta^\mathrm{Injected}$. The difference between the true and measured values
-is the systematic bias introduced
-by ignoring tidal effects in templates. The darker shading shows
-the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
-which is a direct measure of our statistical uncertainty on $\eta$.
-% 
-Comparing systematic and statistical errors, we find that:
-at $\rho=20$, $\eta$ measurement is dominated by statistical
-errors; at $\rho=30$, the two become comparable; and 
-for louder signals ($\rho\simeq50$), the systematic errors dominate.
-}
-\label{fig:SingleSystemEtaPDFvsSNR}
-\end{figure*}
+% \begin{figure*}
+% \centering 
+% \textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
+% \textbf{at different SNR values (shown in legend).}\par\medskip
+% \includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
+% \caption{We show here probability distributions for mass ratio $\eta$ as measured
+% for the same injection at different SNRs. The intrinsic parameters of the source
+% are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
+% and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
+% {\it ignore} tidal effects.
+% % 
+% In each panel: the dashed red line marks the median {\it measured} value
+% $\eta^\mathrm{Median}$, while the dashed green line show the true value
+% $\eta^\mathrm{Injected}$. The difference between the true and measured values
+% is the systematic bias introduced
+% by ignoring tidal effects in templates. The darker shading shows
+% the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
+% which is a direct measure of our statistical uncertainty on $\eta$.
+% % 
+% Comparing systematic and statistical errors, we find that:
+% at $\rho=20$, $\eta$ measurement is dominated by statistical
+% errors; at $\rho=30$, the two become comparable; and 
+% for louder signals ($\rho\simeq50$), the systematic errors dominate.
+% }
+% \label{fig:SingleSystemEtaPDFvsSNR}
+% \end{figure*}
 % ################# FIXME FIXME
@@
 \end{figure*}
+%
+% ##############################################################
+% 
+\begin{figure*}
+\centering 
+\textbf{Illustrative posterior probability distributions for NS tidal
+deformability $\lambdans$,}\par
+\textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
+from left to right.}\par\medskip
+% \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
+\includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
+\caption{We illustrate here probability distributions recovered for the NS tidal
+deformability parameter $\lambdans$ from three GW injections, with parameters:
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
+$\lambdans=\{1000,1500,2000\}$ from left to right. The injection SNR is fixed at
+$\rho=50$. The templates {\it include} tidal effects, with a prior $0\leq\lambdans\leq 4000$.
+% These figures show that this approximation holds up to (SNR)
+% $\rho\simeq 30-50$.
+% 
+In each panel- the dashed red line marks the median {\it measured} value for
+$\lambdans$, and the dashed green line marks its {\it true} value.
+The darker shading shows the $90\%$ credible interval, whose width
+$(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
+By comparing the measurement uncertainty for these three injections, we see
+that $(\Delta\lambdans)^{90\%}$ grows very slowly with $\lambdans$. Therefore,
+the fractional measurement error - $(\Delta\lambdans)^{90\%}/\lambdans$ -
+decreases monotonically as $\lambdans$ increases (with signal strength fixed).
+% 
+% Comparing systematic and statistical errors, we find that:
+% at $\rho=20$, $\lambdans$ measurement is dominated by statistical
+% errors; at $\rho=30$, the two become comparable; and 
+% for louder signals ($\rho\simeq50$), the systematic errors dominate.
+}
+\label{fig:SingleSystemLambdaPDFvsSNR}
+\end{figure*}
+% %% FIXME FIXME
+% \begin{figure*}
+% \centering
+% \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
+% \textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
+% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
+% % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+% % \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+% % \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% % \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+% % \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+% % \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% % \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+% % \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+% % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
+% \caption{%
+% We illustrate here two-dimensional joint probability distributions, as
+% recovered for the intrinsic parameters of a binary with
+% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
+% $\lambdans=2000$, and $\rho=30$ (same as in
+% Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
+% The top row shows the same for combinations of 
+% non-tidal parameters, i.e. binary masses and spins, while the middle row 
+% shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
+% is similar to the middle, with the only difference being that the SNR has been
+% turned up to $\rho=50$.
+% }
+% \label{fig:SingleSystemLambda2DPDFs}
+% \end{figure*}
+% #################
+% \begin{figure*}
+% \centering    
+% \textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
+% \includegraphics[trim={2cm 0 0 0},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
+% % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
+% \caption{Here we show the statistical uncertainty in the measurement of
+% $\lambdans$, as a percentage of the injected/true value. In each panel, the
+% same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
+% injection's SNR $\rho$ fixed (noted in the panel). Each row contains panels
+% with the same value of $\lambdans$, with $\rho$ increasing from left to right.
+% Each column contains panels with the same value of $\rho$, with $\lambdans$ 
+% increasing from top to bottom.
+% % 
+% Contours demarcate regions where we can constrain the
+% $\lambdans$ parameter well (within a factor of two of the injected value).
+% % 
+% We note that, as expected, the measurement accuracy for $\lambdans$ improves with (i) increasing
+% SNR, (ii) decreasing BH mass, (iii) increasing BH spin, and 
+% (iv) increasing $\lambdans$, i.e. the tidal deformability of the neutron star.
+% }
+% \label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
+% \end{figure*}
+% %
 % 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 
-\begin{figure*}
-\centering 
-\textbf{Illustrative posterior probability distributions for NS tidal
-deformability $\lambdans$,}\par
-\textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
-from left to right.}\par\medskip
-% \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
-\includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
-\caption{We illustrate here probability distributions recovered for the NS tidal
-deformability parameter $\lambdans$ from three GW injections, with parameters:
-$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
-$\lambdans=\{1000,1500,2000\}$ from left to right. The injection SNR is fixed at
-$\rho=50$. The templates {\it include} tidal effects, with a prior $0\leq\lambdans\leq 4000$.
-% These figures show that this approximation holds up to (SNR)
-% $\rho\simeq 30-50$.
-% 
-In each panel- the dashed red line marks the median {\it measured} value for
-$\lambdans$, and the dashed green line marks its {\it true} value.
-The darker shading shows the $90\%$ credible interval, whose width
-$(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
-By comparing the measurement uncertainty for these three injections, we see
-that $(\Delta\lambdans)^{90\%}$ grows very slowly with $\lambdans$. Therefore,
-the fractional measurement error - $(\Delta\lambdans)^{90\%}/\lambdans$ -
-decreases monotonically as $\lambdans$ increases (with signal strength fixed).
-% 
-% Comparing systematic and statistical errors, we find that:
-% at $\rho=20$, $\lambdans$ measurement is dominated by statistical
-% errors; at $\rho=30$, the two become comparable; and 
-% for louder signals ($\rho\simeq50$), the systematic errors dominate.
-}
-\label{fig:SingleSystemLambdaPDFvsSNR}
-\end{figure*}
-%%
-\begin{figure*}
-\centering
-\textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
-\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
-\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
-\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
-% \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
-% \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
-% \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
-\caption{%
-We illustrate here two-dimensional joint probability distributions, as
-recovered for the intrinsic parameters of a binary with
-$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
-$\lambdans=2000$, and $\rho=30$ (same as in
-Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
-The top row shows the same for combinations of 
-non-tidal parameters, i.e. binary masses and spins, while the middle row 
-shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
-is similar to the middle, with the only difference being that the SNR has been
-turned up to $\rho=50$.
-}
-\label{fig:SingleSystemLambda2DPDFs}
-\end{figure*}
+% % 
+% \begin{figure*}
+% \centering 
+% \textbf{Illustrative posterior probability distributions for NS tidal
+% deformability $\lambdans$,}\par
+% \textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
+% from left to right.}\par\medskip
+% % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
+% \includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
+% \caption{We illustrate here probability distributions recovered for the NS tidal
+% deformability parameter $\lambdans$ from three GW injections, with parameters:
+% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
+% $\lambdans=\{1000,1500,2000\}$ from left to right. The injection SNR is fixed at
+% $\rho=50$. The templates {\it include} tidal effects, with a prior $0\leq\lambdans\leq 4000$.
+% % These figures show that this approximation holds up to (SNR)
+% % $\rho\simeq 30-50$.
+% % 
+% In each panel- the dashed red line marks the median {\it measured} value for
+% $\lambdans$, and the dashed green line marks its {\it true} value.
+% The darker shading shows the $90\%$ credible interval, whose width
+% $(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
+% By comparing the measurement uncertainty for these three injections, we see
+% that $(\Delta\lambdans)^{90\%}$ grows very slowly with $\lambdans$. Therefore,
+% the fractional measurement error - $(\Delta\lambdans)^{90\%}/\lambdans$ -
+% decreases monotonically as $\lambdans$ increases (with signal strength fixed).
+% % 
+% % Comparing systematic and statistical errors, we find that:
+% % at $\rho=20$, $\lambdans$ measurement is dominated by statistical
+% % errors; at $\rho=30$, the two become comparable; and 
+% % for louder signals ($\rho\simeq50$), the systematic errors dominate.
+% }
+% \label{fig:SingleSystemLambdaPDFvsSNR}
+% \end{figure*}
+% % %% FIXME FIXME
+% % \begin{figure*}
+% % \centering
+% % \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
+% % \textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
+% % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+% % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
+% % % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
+% % % \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
+% % % \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
+% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
+% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
+% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
+% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
+% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
+% % \caption{%
+% % We illustrate here two-dimensional joint probability distributions, as
+% % recovered for the intrinsic parameters of a binary with
+% % $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
+% % $\lambdans=2000$, and $\rho=30$ (same as in
+% % Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
+% % The top row shows the same for combinations of 
+% % non-tidal parameters, i.e. binary masses and spins, while the middle row 
+% % shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
+% % is similar to the middle, with the only difference being that the SNR has been
+% % turned up to $\rho=50$.
+% % }
+% % \label{fig:SingleSystemLambda2DPDFs}
+% % \end{figure*}
 % #################
@@
 \end{figure*}
-% 
+% % 
 \begin{figure}
@@
 % the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
-% set by the detectors).
+% set by the detectorTT_SNRThresholds_BHspin_BHmass_CI90_0s).
 }
-\label{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}
+\label{fig:}
 \end{figure}
-%
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
-% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
-% \caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}
-% \prayush{TRY AND MAKE THIS POINT IN THE FIGURE ABOVE ONLY.}
-% In these figures we highlight the regions of BH mass and spin plane 
-% where we can constrain the $\lambdans$ of its companion NS within a factor of 
-% $1-2$ of its true value. Panels correspond to different signal strength, with 
-% $\rho$ increasing from left to right. Within each panel, $\lambdans$ 
-% measurement uncertainty contours are shown as functions of BH mass and spin,
-% with line-types corresponding to different uncertainty levels, and colors 
-% showing the injected $\lambdans$. 
-% % We note that this information is, in principle, contained in 
-% % Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
-% % the effect of NS's deformability itself on its measurement. 
-% As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, the 
-% measurability of NS tidal deformability improves with all factors that 
-% enhances the signature of the NS's disruption on the \textit{detectable} 
-% portion of the emitted GW signal (implying, within a frequency band set by the
-% detectors).
-% }
-% \label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}
-% \end{figure*}
-%
-
-In the previous section, we showed that the effect of deformation of the NS due to
-the quadrupolar tidal field of its companion BH becomes discernible in the
-GW spectrum only when the following hold:
-(a) the BH has a sufficiently small mass, (b) BH spin
-is positive aligned, i.e. $\chibh\gtrsim +0.4$, (c) the NS is not very compact,
-with $\lambdans\gtrsim 1000$, {\it and} (d) the source is close enough
-such that its GW SNR $\gtrsim 30$.
-% 
-The first two conditions both enhance the tidal distortion of the star and increase
+% %
+
+%
+
+In the previous section, we showed that the effects of the deformation of NSs
+due to the quadrupolar tidal field of their companion BHs become discernible in
+the GW spectrum under certain favorable conditions, including (a) BH mass is
+sufficiently small, (b) BH spin is positive aligned, i.e. $\chibh\gtrsim +0.4$,
+(c) the NS is not very compact, with $\lambdans\gtrsim 1000$, and (d) the
+source location and orientation are such that its GW SNR $\gtrsim 30$.
+% 
+Both condition (a) and (b) enhance the tidal distortion of the star and increase
 the number of orbits the system goes through at small separation, where the
-differences between NSBH and BHBH signals are maximal.
-% 
-The first three also reduce the onset frequency of the disruption of the NS,
+differences between NSBH and BBH signals are maximal.
+% 
+Conditions (a)-(c) also reduce the onset frequency of the disruption of the NS,
 allowing for it to happen earlier in the orbit. 
@@
 
-%
-%Ultimately, the extent of tidal moduilations of emitted GWs depends on the
-%equation of state of neutron star matter, which becomes an indirect
-%measurable in GW astrophysics.
-% 
-%We therefore turn to the question: If we do observe GWs from disruptive 
-%coalescing NSBHs, what information can we extract from them about the internal
-%structure of neutron stars?
-% 
-We begin by calculating the accuracy with which we measure
-$\lambdans$ from {\it single} GW observations. We sample the same set of
-disruptive NSBH systems as in the previous section, i.e. those with
-$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
-fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$.
-For each unique combination of these parameters, we generate separate 
-zero-noise injections at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model. We
-perform full Bayesian analyses on each of these injections using LEA+ templates,
-and obtain probabilistic estimates for source binary parameters using GW information
-alone (including the NS tidal deformability parameter $\lambdans$). Our priors
-on component masses and spins remain as in the previous section, while $\lambdans$
-is sampled uniformly from $[0,4000]$, and mass-ratio is additionally restricted to
-$2\leq q\leq 6$.
-
-As an illustration, we show the recovered
-probability distribution for $\lambdans$ for three specific configurations in 
-Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. We fix $q = m_1/m_2 =
-5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, while varying $\lambdans$
-over $\{1000, 1500, 2000\}$ between the three panels. The SNR is fixed at $\rho=50$.
-The true value of $\lambdans$ is given the green (dashed) vertical line,
-while the median recovered value is marked by a red (dashed) line. The darker
-shaded region marks the $90\%$ credible interval on $\lambdans$. One thing
-to note in this figure is that the measurement uncertainty on $\lambdans$, i.e.
-the full width of its $90\%$ credible interval, grows slower than $\lambdans$
-as $\lambdans$ is turned up (while the fractional uncertainty
+In this section, we calculate the accuracy with which we measure $\lambdans$ from
+{\it single} GW observations. We sample the same set of disruptive NSBH mergers
+as in the previous section, i.e. those with $q=\{2,3,4,5\}$,
+$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
+fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$. For each unique
+combination of these parameters, we inject LEA+ signals into zero noise and
+perform a fully Bayesian parameter estimation analysis of each with LEA+
+templates. Our priors on component masses and spins remain as in the previous
+section, with mass-ratio additionally restricted to $2\leq q\leq 6$, and
+$\lambdans$ sampled uniformly from $[0,4000]$.
+% 
+% 
+%
+As an illustration, we show the recovered probability distribution for
+$\lambdans$ for three specific configurations in 
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. We fix
+$q = \mbh /\mns = 5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, while
+varying $\lambdans$ over $\{1000, 1500, 2000\}$ between the three panels.
+The SNR is fixed at $\rho=50$. Both the true and measured values of
+$\lambdans$ are marked. The darker shaded region marks the $90\%$ credible
+interval on $\lambdans$. We note that $\lambdans$ is estimated to within
+$\pm 2000$ of its true value at this SNR. Another interesting thing to note 
+is that while $(\Delta\lambdans)^{90\%}$ slowly grows with $\lambdans$, the
+fractional uncertainty
 \begin{equation}
@@
 \end{equation}
-decreases with increasing $\lambdans$).
-%
+decreases instead.
+%
+\textcolor{blue}{%
+Further illustrations of the measurements of mass, spin and tidal parameters,
+as well as their correlations, are shown in Appendix~\ref{as1:illustrations}.
+}
+% 
+\textcolor{gray}{%
 Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
@@
 as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
-$\rho=30$ (instead of $50$). In the top panel, we show the correlation between
-non-tidal parameters, as manifested in their recovered joint probability
-distributions.
-In the middle row, we show the same between $\lambdans$ and non-tidal 
-parameters, i.e. chirp mass (left panel), mass ratio $\eta$ (middle panel), and
-BH spin $\chibh$ (right panel). The same is shown in the row below, with the
-difference that the SNR has been turned up to $\rho=50$, which brings out the
-intrinsic ambiguity contours more clearly.
+set $\rho=30$ (left panel) or $\rho=50$ (right panel).
 %
@@
 a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
-
-In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the main results
-of this section. In each panel, as a function of black hole mass and spin,
-we show the measured $90\%$ credible intervals for $\lambdans$, i.e.
-$(\Delta\lambdans)^{90\%}$. These correspond to the full width of the dark
-shaded region in the illustrative figure,
-Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. The effect of signal strength,
-i.e. SNR, can be seen as we go from left to right in each row. The
-effect of the tidal deformability of the NS $\lambdans$ on its own measurability
-can be seen by comparing panels within each column. The NS becomes more 
-deformable as we go from top to bottom. In addition, contours of
-$\dlambda=\{50\%,75\%,100\%,150\%,200\%\}$ are also shown in all panels.
-%
-A clear pattern emerges in the left-most column, which corresponds to
-$\rho=20$. We find that at this signal strength, our measurement of 
-$\lambdans$ is dominated by the width of our prior on $\lambdans$. In other
-words, our measured credible interval is restricted by the
-allowed range of $\lambdans$. Increasing the signal strength to
-$\rho=30$ gives somewhat similar results, as can be seen from the second-from-left
-column of the figure. We do, however, find more encouraging results
-for high BH spins, in which case we can put $\pm 100\%$ to $\pm 75\%$
-bounds on $\lambdans$~\footnote{The symmetric error-bars of
-$\pm\mathrm{X}\%$ correspond to $\dlambda = 2\mathrm{X}\%$.}.
-In addition to large BH spins, this level of
-precision will also require rather deformable neutron stars.
-%
-At higher SNRs, $\rho\gtrsim 50$, we can put stronger constraints
-on $\lambdans$. With a {\it single} observation of a $q=4$ binary with
-$\chibh\geq 0.6$ and $\rho = 50$~\footnote{For an optimally oriented
-source with $q=4, \mns=1.35M_\odot, \chibh=0.6$, an SNR of 
-$\rho = 50$ corresponds to a luminosity distance of $\approx 113$Mpc.},
-we would be able to estimate 
-$\lambdans$ to within $\pm 40\%$ of its true value 
-(which is equivalent to measuring the ratio of its radius to mass
-with an uncertainty of about $\pm 10\%$).
+}
+
+In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the main results of
+this section. In each panel, as a function of black hole mass and spin, we show
+the measured $90\%$ credible interval widths $(\Delta\lambdans)^{90\%}$. These
+correspond to the full width of the dark shaded regions in the illustrative
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. The effect of increasing signal
+strength can be seen as we go from left to right in each row. The effect of the
+NS tidal deformability parameter $\lambdans$ on its own measurability can be
+seen by comparing panels within each column, with the NS becoming more
+deformable from top to bottom. 
+%
+A uniform pattern emerges in the left-most column, which corresponds to $\rho=20$.
+We find that at this signal strength, our measurement of $\lambdans$ is
+dominated by the width of our prior on it. The $90\%$ credible intervals span
+the entire allowed range for $\lambdans$, making a reasonable estimation of
+$\lambdans$ at $\rho\simeq20$ difficult.
+% 
+Increasing the signal strength to $\rho=30$ gives marginally better results,
+bringing down the statistical uncertainties to within $\pm 75-100\%$ of the
+true $\lambdans$ value~\footnote{The symmetric error-bars of $\pm\mathrm{X}\%$
+correspond to $\dlambda = 2\mathrm{X}\%$.}.
+%
+It is not until we reach an SNR as high as $\rho\gtrsim 50$, can we put
+meaningful (i.e. $\mathcal{O}(10\%)$) constraints on $\lambdans$. For e.g.,
+with a {\it single} observation of a $q=4$ binary with $\chibh\geq 0.6$ and
+$\rho = 50$~\footnote{For an optimally oriented source with
+$q=4, \mns=1.35M_\odot, \chibh=0.6$, an SNR of $\rho = 50$ corresponds to
+a luminosity distance of $\approx 113$Mpc.}, we would be able to estimate 
+$\lambdans$ to within $\pm 40\%$ of its true value (which is equivalent to
+measuring the ratio of NS radius to mass with an uncertainty of about
+$\pm 10\%$).
+% 
+The results presented so far are quite in agreement with Sec.~\ref{s1:PEwithnoNS}.
 
@@
 % #################
-\begin{figure*}
-\centering    
-\textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
-\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
-\includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
-\caption{%
-These panels show the accumulation of information from a population whose
-$\lambdans=800$.
-{\it Left:} Posterior probability distributions for $\lambdans$ (colored 
-curves), and associated $90\%$ credible intervals (grey vertical lines),
-shown for different number of accumulated observations $N$. Distributions are
-normalized to unit area.
-{\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
-associated $90\%$ credible intervals (as the vertical extent of filled
-region), shown as a function of number of observations $N$. Solid horizontal
-line indicates the true value of $\lambdans$. Dashed and dotted horizontal
-lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
-bounds.
-}
-\label{fig:TT_Lambda_vs_N_L800_CI90_0}
-\end{figure*}
-%
+% \begin{figure*}
+% \centering    
+% \textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
+% \includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
+% \includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+% \caption{%
+% These panels show the accumulation of information from a population whose
+% $\lambdans=800$.
+% {\it Left:} Posterior probability distributions for $\lambdans$ (colored 
+% curves), and associated $90\%$ credible intervals (grey vertical lines),
+% shown for different number of accumulated observations $N$. Distributions are
+% normalized to unit area.
+% {\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
+% associated $90\%$ credible intervals (as the vertical extent of filled
+% region), shown as a function of number of observations $N$. Solid horizontal
+% line indicates the true value of $\lambdans$. Dashed and dotted horizontal
+% lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
+% bounds.
+% }
+% \label{fig:TT_Lambda_vs_N_L800_CI90_0}
+% \end{figure*}
+% %
 % #################
@@
 % signals.}
+
+Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
+mass, spin, and tidal parameter measurements. We keep the binary parameters
+as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
+set $\rho=30$ (left panel) or $\rho=50$ (right panel).
+%
+We find that the measurement of $\lambdans$ is weakly degenerate with
+other parameters, and at realistic SNRs it would improve by a few tens of 
+percent if we knew non-tidal parameters to better accuracy. The predominant
+factor that would enhance the measurement accuracy for $\lambdans$ is the
+signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
+measurement to be limited by its degeneracy with non-tidal parameters at 
+a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
 

### 2016-09-06 Paper/paper.tex
shrunk fig 9

@@
 }
-\label{fig:}
+\label{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}
 \end{figure}
@@
 
-In the previous section, we showed that the effects of the deformation of NSs
-due to the quadrupolar tidal field of their companion BHs become discernible in
+In the previous section, we showed that the effects of the tidal deformation of
+NSs by their companion BHs become discernible in
 the GW spectrum under certain favorable conditions, including (a) BH mass is
@@
 % 
-We expect that these conditions are also the ones which would maximize the
+We expect that these conditions are also the ones which should maximize the
 likelihood of {\it measuring} tidal effects in NSBH signals. In this section,
@@
 %
-we turn the question around and ask: under similarly favorable circumstances, can
-we gain insights about the internal structure of neutron stars from GW observations.
+we turn the question around to ask: under similarly favorable circumstances,
+can we gain insights about the internal structure of neutron stars from GW
+observations?
 
@@
 perform a fully Bayesian parameter estimation analysis of each with LEA+
-templates. Our priors on component masses and spins remain as in the previous
-section, with mass-ratio additionally restricted to $2\leq q\leq 6$, and
-$\lambdans$ sampled uniformly from $[0,4000]$.
-% 
-% 
-%
-As an illustration, we show the recovered probability distribution for
-$\lambdans$ for three specific configurations in 
+templates. Our  priors on component masses and spins remain as in the
+previous section, with mass-ratio additionally restricted to $2\leq q\leq 6$,
+and $\lambdans$ sampled uniformly from $[0,4000]$.
+% 
+%
+As an illustration of individual injections, we show the recovered probability
+distribution for $\lambdans$ for three specific configurations in 
 Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. We fix
-$q = \mbh /\mns = 5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, while
-varying $\lambdans$ over $\{1000, 1500, 2000\}$ between the three panels.
-The SNR is fixed at $\rho=50$. Both the true and measured values of
-$\lambdans$ are marked. The darker shaded region marks the $90\%$ credible
+$q = \mbh /\mns = 5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, and
+vary $\lambdans$ over $\{1000, 1500, 2000\}$ between the three panels.
+The SNR is fixed at $\rho=50$. The darker shaded regions mark the $90\%$ credible
 interval on $\lambdans$. We note that $\lambdans$ is estimated to within
@@
 \textcolor{blue}{%
-Further illustrations of the measurements of mass, spin and tidal parameters,
+Further illustrations of the measurements of non-tidal parameters,
 as well as their correlations, are shown in Appendix~\ref{as1:illustrations}.
+We will continue here to focus on the measurement of $\lambdans$ itself.
 }
@@
 % 
-The results presented so far are quite in agreement with Sec.~\ref{s1:PEwithnoNS}.
+The results presented so far are in agreement with Sec.~\ref{s1:PEwithnoNS}.
 
@@
 BH with a larger spin always allows for a more precise measurement on $\lambdans$.
-Visually we can see this in the bottom right corner of each panel in
+We can see this in the bottom right corner of each panel in
 Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which corresponds to low-mass BHs
@@
 % 
-\begin{figure*}
+\begin{figure}
 \centering    
-\textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
-\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
-\includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+\textbf{Improvement in $\lambdans$ measurement accuracy\\ with multiple observations}\par\medskip
+\includegraphics[width=0.8\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}\\
+\includegraphics[width=0.8\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
 \caption{%
@@
 \label{fig:TT_Lambda_vs_N_L800_CI90_0}
-\end{figure*}
+\end{figure}
 %

### 2016-09-06 Paper/paper.tex
fixed figure captions

@@
 \caption{We show here probability distributions for mass ratio $\eta$ as measured
-for the same injection at different SNRs. The intrinsic parameters of the source
+for the same signal at different SNRs. The intrinsic parameters of the source
 are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
@@
 % 
-In each panel: the dashed red line marks the median {\it measured} value
+In each panel: the dashed red line marks the median or {\it measured} value
 $\eta^\mathrm{Median}$, while the dashed green line show the true value
-$\eta^\mathrm{Injected}$. The difference between the true and measured values
-is the systematic bias introduced
-by ignoring tidal effects in templates. The darker shading shows
-the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
-which is a direct measure of our statistical uncertainty on $\eta$.
+$\eta^\mathrm{Injected}$. The darker shading shows
+the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$.
 % 
@@
 measurement uncertainties for the binary chirp mass over the NSBH parameter 
-space. Each panel shows the same as a function of BH mass and spin. Across
-each row, we can see the effect of increasing the signal strength (i.e. SNR)
-with the tidal deformability of the NS fixed. Down each column, we can see 
-the effect of the increasing tidal deformability of the NS at fixed SNR. We also
-show dashed contours for where the ratio equals $10\%$ (labelled ``-1'')
-or $100\%$ (labelled ``0'').
-% 
-Generally, these two sources of uncertainty are different and for BBHs, the
-statistical errors dominate systematic ones~\cite{Kumar:2016dhh}. We find
-that for BHNS binaries, its not much different until we get to high SNRs
-$\rho\gtrsim 70$ for which if the BH is spinning rapidly enough, the 
-measurement uncertainty for $\mchirp$ could become dominated by the exclusion
-of tidal effects in filtering templates.
+space. Each panel shows the same as a function of BH mass and spin. NS mass
+is fixed at $\mns=1.35M_\odot$, and its spin is set to zero. Down each column,
+we can see the effect of the increasing tidal deformability of the NS at fixed
+SNR. Across each row, we can see the effect of increasing the signal strength
+(SNR), with the tidal deformability of the NS fixed. We also show dashed
+contours for $\arr_{\mchirp}=10\%$, $50\%$, or $100\%$ (labeled
+without the ``$\%$'' symbol).
+% 
+For BBHs, the statistical errors dominate systematic ones for contemporary
+waveform models~\cite{Inprerp-LVC-WaveModels:2016,Kumar:2016dhh}. We find that
+its not much different for NSBH binaries, until we get to very high SNRs
+$\rho\gtrsim 70$.
 }
@@
 \caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
-with the difference that the quantity in question here is the symmetric 
-mass-ratio $\eta$. We find that for fairly loud GW signals, with $\rho\simeq 50$,
-not including the effects of tidal deformation of the NS on GW emission can become the dominant
-source of error for astrophysical searches with Advanced LIGO. However,
-for quieter signals with $\rho\leq 30$, it will have minimal effects on the measurement
-of $\eta$. We remind the reader that the SNRs here are all single detector values.
+with the difference that here we show the ratio of systematic and statistical
+error sources for the symmetric mass-ratio $\eta$ and not chirp mass. We find
+that for fairly loud GW signals, at $\rho\simeq 50$, not including the
+effects of tidal deformation of the NS on GW emission can become the dominant
+source of error for astrophysical searches with Advanced LIGO. However, for
+quieter signals with $\rho\leq 30$, it will have a negligible effect on the
+measurement of $\eta$. We remind the reader that the SNRs here are always
+single detector values.
 }
@@
 \caption{This figure shows the ratio of the systematic and statistical
-measurement errors for BH spins, with other attributed identical to 
+measurement errors for BH spins, with other attributes identical to 
 Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}.
-Similar to the case of mass parameters, we find that below
-$\rho\approx 30$, ignoring tidal effects in templates introduces minor systematic effects,
-which remain subdominant to the statistical measurement uncertainties.
+Similar to the case of mass parameters, we find that below $\rho\approx 30$,
+ignoring tidal effects in templates introduces minor systematic effects,
+which remain sub-dominant to the statistical measurement uncertainties.
 }
@@
 \includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
-\caption{We illustrate here probability distributions recovered for the NS tidal
+\caption{We show here probability distributions recovered for the NS tidal
 deformability parameter $\lambdans$ from three GW injections, with parameters:
@@
 % same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
-% injection's SNR $\rho$ fixed (noted in the panel). Each row contains panels
+% injection's SNR $\rho$ fixed (noted in the panel). Rows contain panels
 % with the same value of $\lambdans$, with $\rho$ increasing from left to right.
@@
 $X=\{\mchirp,\eta,\chibh\}$ can be measured, are presented in
-Appendix~\ref{as1:illustrations}.
+Appendix~\ref{as1:nontidalerrors}.
 }
@@
 same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
-injection's SNR $\rho$ fixed (noted in the panel). Each row contains panels
+injection's SNR $\rho$ fixed (noted in the panel). Rows contain panels
 with the same value of $\lambdans$, with $\rho$ increasing from left to right.
-Each column contains panels with the same value of $\rho$, with $\lambdans$ 
+Columns contain panels with the same value of $\rho$, with $\lambdans$ 
 increasing from top to bottom.
 % 
-Contours demarcate regions where we can constrain the
+Contours at $(\Delta\lambdans)^{90\%}=\{50\%, 75\%, 100\%, 150\%, 200\%\}\times\lambdans^\mathrm{Injected}$ demarcate regions where we can constrain the
 $\lambdans$ parameter well (within a factor of two of the injected value).
 % 
-We note that, as expected, the measurement accuracy for $\lambdans$ improves with (i) increasing
-SNR, (ii) decreasing BH mass, (iii) increasing BH spin, and 
-(iv) increasing $\lambdans$, i.e. the tidal deformability of the neutron star.
+We note that, as expected, the measurement accuracy for $\lambdans$ improves
+with (i) increasing SNR, (ii) increasing $\lambdans$, (iii) increasing BH spin,
+and (iv) decreasing BH mass.
 }
@@
 sensitivity LIGO instruments, we expect this to happen approximately once in a
-year of observation.
+year of observation~\cite{Abadie:2010cfa}.
 % 
@@
 \textcolor{blue}{%
-Further illustrations of the measurements of non-tidal parameters,
-as well as their correlations, are shown in Appendix~\ref{as1:illustrations}.
+Further illustrations, showing the correlation between tidal and non-tidal
+parameters, are presented in Appendix~\ref{as1:illustrations}.
 We will continue here to focus on the measurement of $\lambdans$ itself.
@@
 from $\chibh\in[0, 0.75]$, source location uniformly in spatial volume,
-and its inclination from $\iota\in[0,\pi]$. \prayush{Note that we only
+and its inclination from $\iota\in[0,\pi]$. Note that we only
 sample positive aligned BH spins, since binaries with anti-aligned spins have
 very little information to add at realistic SNRs, as shown in 
-Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}.
-}
+Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}. This further restricts the
+sub-space of all NSBH observations that our populations span, and will need to
+be take into account when the number of observations required for tidal
+measurements is related to detector observation time.
+% 
 As noted previously, we also restrict events to have
@@
 
-\section{Illustrations of Bayesian posteriors}\label{as1:illustrations}
+\section{Statistical uncertainty in measuring non-tidal parameters}\label{as1:nontidalerrors}
 % 
@@
 $\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
 \caption{
-We show here the statistical uncertainty of our measurement of the 
-dimensionless mass ratio $\eta$ (at $90\%$ credibility), over the NSBH
-parameter space.
-Individual panels show the same as a function of BH mass and spin.
-Across each row, we see the effect of increasing signal strength (i.e.
-SNR) with the tidal deformability of the NS $\lambdans$ fixed. Down each
-column, we see the effect of increasing $\lambdans$, at fixed SNR.
+We show here the statistical uncertainty associated with our measurement of
+non-tidal parameters $\mchirp, \eta,$ and $\chibh$ (at $90\%$ credibility),
+over the signal parameter space. Individual panels show the same as a function
+of BH mass and spin. Across each row, we see the effect of increasing signal
+strength (i.e. SNR) with the tidal deformability of the NS $\lambdans$ fixed.
+Down each column, we see the effect of increasing $\lambdans$, at fixed SNR.
+Tidal effects are ignored in templates.
 }
@@
 %%
+In Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
+measure non-tidal NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
+The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
+(bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
+as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
+$\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
+$1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
+does not significantly change the measurement uncertainties for non-tidal
+binary parameters, as is evident from comparing the two rows in each panel of
+Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
+% 
+From the left-most column, we find that: (i) at $\rho=20$ the chirp mass is
+measured remarkably well - to a precision of $0.16\%$ of its true value, and
+(ii) so is $\chibh$. (iii) The dimensionless mass-ratio $\eta$ is determined
+more loosely, with $25+\%$ uncertainty. If the signal is even louder
+($\rho\geq 30$), all three measurements gain further precision, especially
+$\eta$, for which the relative errors shrink down to single-digit percents.
+% % 
+We remind ourselves that these results do not tell the full story since the
+precision of a measurement is only meaningful if the measurement is accurate 
+to begin with. In our case there are tidal effects that have not been
+incorporated into our search (BBH) templates, which can lead to a systematic
+bias in parameter recovery. We refer the reader to Sec.~\ref{s1:PEwithnoNS} for
+a comparative study of both systematic and statistical errors.
+
+\section{Illustrations of Bayesian posteriors}\label{as1:illustrations}
+% 
 \begin{figure*}
@@
 \caption{%
-We illustrate here two-dimensional joint probability distributions, as
-recovered for the intrinsic parameters of a binary with
-$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
-$\lambdans=2000$, and $\rho=30$ (same as in
-Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
-The top row shows the same for combinations of 
-non-tidal parameters, i.e. binary masses and spins, while the middle row 
-shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
-is similar to the middle, with the only difference being that the SNR has been
-turned up to $\rho=50$.
+% We illustrate here two-dimensional joint probability distributions, as
+% recovered for the intrinsic parameters of a binary with
+% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
+% $\lambdans=2000$, and $\rho=30$ (same as in
+% Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
+% The top row shows the same for combinations of 
+% non-tidal parameters, i.e. binary masses and spins, while the middle row 
+% shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
+% is similar to the middle, with the only difference being that the SNR has been
+% turned up to $\rho=50$.
+We illustrate here two sets of two-dimensional joint probability distributions,
+differing only in signal strength, with $\rho=20$ in the left panel, and
+$\rho=50$ in the right. The injected parameters are 
+$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
+$\lambdans=2000$. Templates include tidal effects, as evident in the bottom rows
+of both panels which show the correlation of $\lambdans$ with non-tidal 
+parameters. Contrasting the two panels illustrates the effect of increasing the
+SNR on various parameter measurements.
 }
@@
 
-Let us first delve into statistical errors. In
-Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
-measure NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
-The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
-(bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
-as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
-$\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
-$1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
-does not significantly change the measurement uncertainties for non-tidal
-binary parameters, as is evident from comparing the two rows in each panel of
-Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
-% 
-From the left-most column, we find that: (i) at $\rho=20, \mchirp$ is already
-measured remarkably well - to a precision of $0.16\%$ of its true value, and
-(ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined more loosely,
-with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
-all three measurements gain further precision, with $\eta$ errors shrinking down
-to single-digit percents of its true value.
-% % 
-% We remind ourselves that this {\it precision}, however, is only
-% meaningful so long as the measurement is {\it accurate} to begin with.
-% We define $\arr_X$ as the ratio between systematic and statistical
-% errors associated with the measurement of parameter $X$,
-% \begin{equation}\label{eq:arr}
-% \arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
-% \end{equation}
-% in order to compare the relative importance of both. Only when
-% $|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
-% will not hamper the measurement of non-tidal parameters from NSBH signals.
-% Due to
-% unmodeled tidal effects, parameter measurements performed on NSBH signals with
-% BBH templates will have systematic biases. If these biases are too large,
-% using non-tidal BBH templates should severely impair our inferencing ability.
-% 
-% there are physical effects that have not been
-% incorporated in our (BBH) search templates which we use here to decipher NSBH
-% signals. Therefore, we cannot know the {\it accuracy} of our parameter
-% measurements unless we know how large the systematic biases due to an
-% incomplete template model are. To quantify this, we find it useful to define
-% $\arr_X$, the ratio of the systematic biases induced by unmodeled tidal effects in
-% templates, to the statistical uncertainties associated with the measurement
-% itself. If $\arr_X\gtrsim 1$, including the effects of NS tidal distortion
-% (during inspiral) and possible disruption (close to merger) would be imperative
-% to obtain {\it accurate} estimates of source masses and spins from observed GW
-% signals.}
-
-Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
+In Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
 mass, spin, and tidal parameter measurements. We keep the binary parameters
@@
 percent if we knew non-tidal parameters to better accuracy. The predominant
-factor that would enhance the measurement accuracy for $\lambdans$ is the
-signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
-measurement to be limited by its degeneracy with non-tidal parameters at 
-a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
+factor that would enhance the measurement accuracy for $\lambdans$ is 
+nevertheless the signal strength. Only when $\rho\gtrsim 50$ can we
+expect $\lambdans$ measurement to be limited by its degeneracy with 
+non-tidal parameters (at a factor of few level), as also reported by previous
+studies~\cite{Lackey:2013axa}.
 

### 2016-09-06 Paper/paper.tex
structuring s6

@@
 The pioneering terrestrial observation of gravitational waves by Advanced LIGO
-harbingers the dawn of an era of gravitational-wave astronomy where 
-observations, more than theory, would play the driving role. Binaries of 
-compact stellar objects evolve under the radiation reaction force of emitting
-gravitational waves.
-
-Conventional astronomical methods have, to date, made observations of ~2500
-NSs with masses between $1.25-2.1 M_\odot$~\cite{Lyne:2004cj,
-Demorest:2010bx,2013Sci...340..448A,atnfcatalog,mcgillmagnetarcatalog,
-stellarcollapsemass}, although
-tightly clustered within $1.35\pm0.5M_\odot$~\cite{stellarcollapsemass}.
-%
-Observations of over 2500 NSs have shown that the spin of NSs
-$|\vec{\chi}_\mathrm{NS}|$ have magnitudes that are $<1\%$~\cite{Miller:2014aaa}.
-%
-On the other hand, indirect observations of stellar-mass BHs place their
-masses between $5-35M_\odot$, with their spin angular momenta 
-$|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
-(see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
-examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
-Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
-and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
+harbingers the dawn of an era of gravitational-wave astronomy where observations
+would finally drive scientific discovery~\cite{Abbott:2016blz}. As confirmed by
+the first observations~\cite{Abbott:2016blz,Abbott:2016nmj,Abbott:2016nhf},
+stellar-mass compact binary mergers emit GWs right in the sensitive frequency
+band of the LIGO observatories, and are therefore their primary targets.
+Neutron star black hole binaries form a physically distinct sub-class of
+compact binaries, that we expect to detect the first of in the upcoming
+observing runs~\cite{Abbott:2016ymx}. And when the detectors reach their design
+sensitivity, we can expect to detect NSBHs at a healthy rate of $0.2-300$
+mergers a year~\cite{Abadie:2010cf}.
+
+\prayush{\bf Why NSBHs? 1. GRB, 2. tidal measurements.}
+
+\prayush{\bf what we do in this paper - brief summary}
+
+\prayush{\bf Sec 3 summary}
+% % On the known NS and BH properties
+% The number of neutron stars observed using conventional astronomical methods has
+% grown rapidly in the recent past, both in isolated and two-body
+% systems~\cite{Demorest:2010bx,Lyne:2004cj,2013Sci...340..448A,atnfcatalog,
+% mcgillmagnetarcatalog,stellarcollapsemass}. Their masses span the range $1.2-2M_\odot$
+% with an average of $\sim 1.35M_\odot$~\cite{Lattimer:2004sa,stellarcollapsemass}.
+% On the other hand, NS spins have been observed to have magnitudes below
+% $|\vec{S}_\mathrm{NS}|/\mns^2 < 0.01$~\cite{Miller:2014aaa}.
+% % 
+% On the other hand, indirect observations of stellar-mass BHs place their
+% masses between $5-35M_\odot$, with their spin angular momenta 
+% $|\vec{\chi}_\mathrm{BH}|$ ranging from small to nearly extremal (Kerr) values
+% (see, e.g., Refs.~\cite{McClintockEtAl:2006,Miller:2009cw,Gou:2014una} for 
+% examples of nearly extremal estimates of BH spins, Refs.~\cite{McClintock:2013vwa,
+% Reynolds:2013qqa} for recent reviews of astrophysical BH spin measurements,
+% and Figure 5 of Ref.~\cite{Miller:2014aaa} for a comparison of NS and BH spins).
 
@@
 $\chibh$ significantly only if signal $\rho\gtrsim 30$. 
-%
+}
+
+\prayush{
 Overall, we learn an important thing. Bayesian parameter estimation algorithms
@@
 }
+
+\prayush{\bf Sec 4 summary}
+
+\prayush{\bf Sec 5 summary}
 
@@
 
-\prayush{
-Also note
+\prayush{\bf Tidal measurements versus observation time - in how many years can
+we constrain $\lambdans$ reasonably?}
+
+Lastly, we note
 that more a recent work that improves upon the waveform model we use in this
@@
 not change our qualitative conclusions.
-}
 

### 2016-09-07 Paper/paper.tex
progress on discussion

@@
 stellar-mass compact binary mergers emit GWs right in the sensitive frequency
-band of the LIGO observatories, and are therefore their primary targets.
+band of the LIGO observatories, and are their primary targets.
 Neutron star black hole binaries form a physically distinct sub-class of
-compact binaries, that we expect to detect the first of in the upcoming
-observing runs~\cite{Abbott:2016ymx}. And when the detectors reach their design
-sensitivity, we can expect to detect NSBHs at a healthy rate of $0.2-300$
-mergers a year~\cite{Abadie:2010cf}.
-
-\prayush{\bf Why NSBHs? 1. GRB, 2. tidal measurements.}
-
-\prayush{\bf what we do in this paper - brief summary}
-
-\prayush{\bf Sec 3 summary}
+compact binaries. We expect to detect the first of them in the upcoming
+observing runs~\cite{Abbott:2016ymx}, and subsequently at a healthy rate of
+$0.2-300$ mergers a year when aLIGO detectors reach design
+sensitivity~\cite{Abadie:2010cf}.
+
+NSBH binaries are interesting for various reasons. Unlike BBHs, the presence of
+matter allows for richer phenomena to occur alongside the strong-field
+gravitational dynamics. The quadrupolar moment of the NS changes during the
+course of inspiral, which increases the inspiral rate of the binary and alters the
+form of the emitted gravitational waves. Close to merger, under restricted but
+plausible conditions, the neutron star is disrupted by the tidal field of its 
+companion black hole and forms an accretion disk around it. This disruption
+reduces the quadrupolar moment of the system, and decreases the amplitude of
+the emitted GWs from the time of disruption through to the end of ringdown.
+Both of these phenomena are discernible in their gravitational-wave signatures
+alone. In addition, if the neutron star matter is magnetized, the magnetic
+winding above the remnant black hole poles can build up magnetic fields
+sufficiently to power short gamma-ray bursts (SGRB)~\cite{}.
+Therefore a coincident observation of gravitational waves from an NSBH merger
+and a SGRB can potentially confirm the hypothesis that the former is a
+progenitor of the latter~\cite{}.
+
+In this paper we study the observability of tidal signatures in the
+gravitational-wave spectrum of NSBH binaries. More specifically, we investigate
+three questions. First, what is the effect of not including tidal effects in 
+templates while characterizing NSBH signals? Second, if we do include tidal 
+effects, how well can we measure the tidal deformability of the NS
+(parameterized by $\lambdans$) from individual NSBH signals? And third, as we
+observe more and more signals, how does our knowledge of $\lambdans$ improve?
+In the following, we summarize our main findings.
+
+First, we study the effects of not including tidal terms in our search
+templates while characterizing NSBH signals. We expect that the waveform
+template that best fits the signal would compensate for the reduced number of
+degrees of freedom in the template model by moving away from the true
+parameters of the binary. This should result in a {\it systematic} bias in 
+the recovered values of non-tidal source parameters, such as its masses 
+and spins. In order to quantify it, we inject tidal signals into zero noise,
+and perform a Bayesian parameter estimation analysis on them using templates
+{\it without} tidal terms.
+% 
+We use the LEA+ model to produce tidal waveforms that incorporate the effect
+of NS distortion during inspiral, and of its disruption close to merger. Our
+injected signals sample the region of NSBH parameter space where NS disruption
+prior to binary merger is likely {\it and} can be modeled using LEA+. Their
+parameters are given by combinations of $q=\mbh/\mns=\{2,3,4,5\},
+\chibh=\{-0.5, 0, 0.5, 0.75\}$ and $\lambdans=\{500,1000,1500,2000\}$.
+Other parameters, such as source location and orientation, that factor out of
+$h(t)$ as amplitude scaling are co-sampled by varying $\rho=\{20,30,50,70\}$.
+
+At low to moderate SNRs ($\rho\lesssim 30$), we find that using BBH templates
+does not significantly hamper our estimation of non-tidal parameters for NSBH
+signals. In the worst case, when the BH mass is within the astrophysical mass-gap
+and its spin is positive aligned, the systematic biases in $\eta$ and $\chibh$
+measurements do become somewhat comparable to statistical errors (ratio
+$\sim 0.5-0.8$) under very restrictive conditions~\footnote{requiring a
+companion BH with mass $\mbh\lesssim 4.5M_\odot$ (i.e. in the astrophysical
+``mass-gap''~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,
+Littenberg:2015tpa}), and the
+hardest NS EoS considered (with $\lambdans\simeq 2000$).}, but never exceed them.
+At high SNRs ($\rho\gtrsim 50$), systematic biases in $\mchirp$ become larger
+than the statistical uncertainties. For $\eta$ and $\chibh$ the difference
+is more drastic with the systematics reaching up to $4\times$ the statistical
+errors. We therefore conclude that $\rho\simeq 30-50$ is loud enough to
+motivate the use of tidal templates for even the estimation of non-tidal
+parameters from NSBH signals.
+% 
+We also conclude that low-latency parameter estimation algorithms, designed to
+classify GW signals into electromagnetically active (NSBH and NSNS) and
+inactive (BBH) sources, can use BBH templates to trigger GRB 
+alerts~\cite{2012A&A...541A.155A,Singer:2014qca,Singer:2015ema,Pankow:2015cra,
+Abbott:2016wya,Abbott:2016gcq} for NSBH signals with low to moderate SNRs
+($\rho\lesssim 30$).
+This is so because the primary requirement of identifying NS-X binaries (X =
+\{NS, BH\}) can be achieved just as easily with BBH templates, on the basis of
+the smaller component's mass\footnote{The smaller component mass is unlikely
+to be significantly biased by missing tidal effects in filter templates below
+$\rho\simeq 30$, as we show above.}.
+% 
+We also speculate that NSBH detection searches are unlikely to be
+affected by the choice of ignoring tidal effects in matched-filtering
+templates, if these effects are too subtle to manifest in parameter estimation
+below $\rho\simeq 30$.
+
+% 
 % % On the known NS and BH properties
@@
 
-\prayush{
-We find that non-inclusion of tidal effects leads to systematic biases in
-the measurement of non-tidal parameters. Consider first the binary chirp 
-mass $\mchirp:=M\eta^{3/5}$ where $M$ is the binary total mass, and 
-$\eta:=\mbh\mns/M^2$ is the dimensionless mass-ratio. We find that for
-realistic (moderate) SNRs, i.e. $\rho\lesssim 30$, not including tidal
-effects introduces a barely measurable bias that remains below $50\%$
-of the statistical uncertainty of the $\mchirp$ measurement itself.
-For higher signal strengths, we find that the underlying statistical
-uncertainty decreases enough for the systematic bias to become comparable
-to it for $\rho=50-70$ and higher. This is expected because chirp mass is fairly
-accurately determined by the inspiral, where tidal effects are weak.
-Next, we consider the mass-ratio $\eta$. We find that only if the NS is quite
-deformable (with $\lambdans=2000$), {\it and} has a companion BH whose mass
-is $\mbh\lesssim 4.5M_\odot$ (i.e. in the astrophysical ``mass-gap''~\cite{
-Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}),
-will the systematic bias in $\eta$ measurement become comparable to its
-statistical uncertainties at SNRs $\rho\leq 30$. For SNRs $\rho\gtrsim30$,
-however, not including tidal effects is very likely to {\it significantly} bias
-the mass-ratio estimate for NSBH binaries. Finally, we consider the spin on the
-black hole $\chibh$. We find that $\chibh$ measurements are affected in a similar
-way to $\eta$. By not including tidal effects, we compromise our measurement of
-$\chibh$ significantly only if signal $\rho\gtrsim 30$. 
-}
-
-\prayush{
-Overall, we learn an important thing. Bayesian parameter estimation algorithms
-aimed at identifying NSBH binaries with low latency in order to alert 
-electromagnetic (EM) observers in time to establish a possible coincident 
-observation of a GRB~\cite{2012A&A...541A.155A,Singer:2014qca,Singer:2015ema,
-Pankow:2015cra} can use BBH templates to look for NSBH systems with aLIGO.
-This is so because the primary requirement of the EM follow-up effort is rapid
-classification of a binary as an NSBH system, which can be achieved just as 
-easily with BBH templates, purely on the basis of the smaller component's mass.
-Another conclusion we draw is that detection searches are unaffected by the
-choice of ignoring tidal effects in matched-filtering templates.
-}
-
-\prayush{\bf Sec 4 summary}
+Second, we turn the question around to ask: can we measure the tidal effects if
+our template models did account for them? Tidal effects in our waveform model
+are parameterized using a single deformability parameter 
+$\lambdans\propto (R/M)_\mathrm{NS}^5$. In order to quantify the 
+measurability of $\lambdans$, we inject the same tidal signals as before, and
+this time perform a Bayesian analysis on them using {\it tidal} templates. 
+The results are detailed in Sec.~\ref{s1:PEwithNS}.
+% 
+At low SNRs ($\rho\simeq 20$), we find that the best we can do is to constrain
+$\lambdans$ within $\pm 75\%$ of its true value at $90\%$ credible level. This
+too only if the BH is spinning sufficiently rapidly, with $\chibh\gtrsim +0.7$,
+and the NS has $\lambdans\gtrsim 1000$. At moderate SNRs ($\rho\simeq 30$), we
+can constrain $\lambdans$ a little better, i.e. within $\pm 50\%$ of its true
+value. This level of accuracy, however, again requires that BH spin
+$\chibh\gtrsim+0.7$ and $\lambdans\gtrsim 1000$. Binaries with smaller BH spins
+and/or softer NS EoSs will furnish worse than $\pm 75\%-\pm 100\%$ errors for
+$\lambdans$. This trend continues as we increase the SNR from $\rho=30-50$. It
+is not before we reach an SNRs as high as $\rho\simeq 70$ that we can shrink
+$\lambdans$ errors substantially with a single observation (i.e. within
+$\pm 25\%$ of its true value).
+% 
+In summary, we find that with a single but moderately loud NSBH signal,
+Advanced LIGO can begin to put a factor of $1-2\times$ constraints on NS tidal
+deformability parameter. These constraints can subsequently be used to assess
+the likelihood of various candidate equations of state for nuclear matter, and
+possibly to narrow the range they span.
 
@@
 time-scale.
-}
-
-\prayush{%
-We find that when the BH is sufficiently spinning in the prograde sense,
-with $\chibh\gtrsim +0.7$, even at a relatively possible signal strengths ($\rho=20$),
-we can begin to constraint $\lambdans$ at the level of $\pm 75\%$ errors with a $90\%$
-credible level. If the signal is stronger, say $\rho=30$, we can contraint
-$\lambdans$ better with only $\pm 50\%$ errors, for binaries with large BH spins.
-This trend continues as we vary the SNR from $\rho=30-50$ upwards. In other words,
-with a single but moderately loud NSBH signal observation, Advanced LIGO will begin
-to put constraints on NS matter properties. These constraints can consequently
-be used to assess the likelihood of different equations of state for nuclear
-matter, and possibly narrow the range they span.
 }

### 2016-09-07 Paper/paper.tex
gone through s6

@@
 quadrupolar moment of the disrupted binary system falls monotonically over a
-millisecond time-scale~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,Foucart:2015a,
-Pannarale:2015jia}.
+millisecond time-scale~\cite{Kyutoku:2010zd,Lackey:2013axa,Lovelace:2013vma,
+Foucart:2015a,Pannarale:2015jia}, resulting in the damping of GW amplitude.
 % 
 This penultimate stage also depends strongly on the internal structure and energy
-transport mechanism of the NS~\cite{}.
+transport mechanism of the NS, and carries the strongest tidal signature in the
+GW spectrum~\cite{Foucart:2014nda,Deaton2013}.
 
@@
 (uniformly) from $\mbh\in[3M_\odot,6.75M_\odot]$, BH spin (uniformly)
-from $\chibh\in[0, 0.75]$, source location uniformly in spatial volume,
+from $\chibh\in[0, 1]$, source location uniformly in spatial volume,
 and its inclination from $\iota\in[0,\pi]$. Note that we only
@@
 winding above the remnant black hole poles can build up magnetic fields
-sufficiently to power short gamma-ray bursts (SGRB)~\cite{}.
+sufficiently to power short gamma-ray bursts (SGRB)~\cite{Foucart:2015a,
+Lovelace:2013vma,Deaton2013,Foucart2012,Shibata:2005mz,Paschalidis2014}.
 Therefore a coincident observation of gravitational waves from an NSBH merger
 and a SGRB can potentially confirm the hypothesis that the former is a
-progenitor of the latter~\cite{}.
+progenitor of the latter~\cite{eichler:89,1992ApJ...395L..83N,moch:93,
+Barthelmy:2005bx,2005Natur.437..845F,2005Natur.437..851G,Shibata:2005mz,
+Tanvir:2013,Paschalidis2014}.
 
@@
 % 
-We use the LEA+ model to produce tidal waveforms that incorporate the effect
+We use the LEA+ model~\cite{Lackey:2013axa} to produce tidal waveforms that
+incorporate the effect
 of NS distortion during inspiral, and of its disruption close to merger. Our
@@
 parameters are given by combinations of $q=\mbh/\mns=\{2,3,4,5\},
-\chibh=\{-0.5, 0, 0.5, 0.75\}$ and $\lambdans=\{500,1000,1500,2000\}$.
+\chibh=\{-0.5, 0, 0.5, 0.75\}$ and $\lambdans=\{500,800,1000,1500,2000\}$.
 Other parameters, such as source location and orientation, that factor out of
@@
 does not significantly hamper our estimation of non-tidal parameters for NSBH
-signals. In the worst case, when the BH mass is within the astrophysical mass-gap
+signals. In the worst case, when the BH mass is within the astrophysical 
+mass-gap~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}
 and its spin is positive aligned, the systematic biases in $\eta$ and $\chibh$
@@
 companion BH with mass $\mbh\lesssim 4.5M_\odot$ (i.e. in the astrophysical
-``mass-gap''~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,
-Littenberg:2015tpa}), and the
+mass-gap), and the
 hardest NS EoS considered (with $\lambdans\simeq 2000$).}, but never exceed them.
@@
 
-\prayush{\bf Sec 5 summary}
-
-\prayush{
-The choices that determine the population are important here,
-and were made as follows. Neutron star mass is fixed for signals at
-$1.35M_\odot$, and its spin $\chins=0$. Black hole mass is sampled uniformly
-from the range $[2,5]\times 1.35=[2.7, 6.75]M_\odot$. Our choice here is given
+% \prayush{\bf Sec 5 summary}\\
+% \prayush{\bf Tidal measurements versus observation time - in how many years can
+% we constrain $\lambdans$ reasonably?}\\
+% % 
+Third, knowing that single observations can furnish only so much information
+about the NS equation of state, we move on to investigate how well do we do with
+multiple signals. In order to quantify how $\lambdans$ measurement improves
+with the number of observed events $N$, we generate populations of NSBH signals
+and combine the information extracted from each event.
+% 
+The population generation procedure is as follows. The neutron star mass is
+held fixed at $1.35M_\odot$, its spin at $\chins=0$, and its tidal
+deformability is fixed to each of $\lambdans=\{500,1000,1500,2000\}$. Black hole
+mass is sampled uniformly from the range $[2,5]\times 1.35=[2.7, 6.75]M_\odot$,
+and spin from $\chibh\in[0,1]$. As before, our parameter choice here is given
 by the intersection set of the mass range that allows for neutron star disruption
-and the range supported by the waveform model LEA+~\cite{Foucart2012,
-Foucart:2013a,Lackey:2013axa}. Black hole spin is sampled uniformly from the
-{\it aligned} range, i.e. $\chibh\in[0,+1]$. The true deformability is fixed
-once per population, sampled uniformly from the range $\lambdans\in[0,2000]$.
-In order to keep the computational cost tractable, we make a further 
-approximation.
-%
-We inject signals into zero noise at different SNRs and recover
-their parameters from Bayesian inferencing for a fixed set of binaries,
-whose parameters lie on a uniform grid given by:
-$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
-% 
-Once the population parameters been selected, each event's parameters are
-replaced by their nearest neighbor in the above set.
-% 
-We make sure that this approximation is conservative by keeping our population
-sampling ranges within the conservative side of the range spanned by the above
-set. With the caveats laid out, we report our results in
-Sec.~\ref{s1:multiple_observations}.
-}
-
-\prayush{
-In addition, we probe two paradigms, one
-that allows for BH masses to lie within the astrophysical mass-gap (paradigm
-A), and one that does not (paradigm B).
-% 
-For paradigm A, we find the following: (i) for the softer equations of state
+and the range supported by LEA+~\cite{Foucart2012,Foucart:2013a,Lackey:2013axa}.
+In order to keep the computational cost reasonable, we make an additional
+approximation. For every population generated, we replace each event's
+parameters by their nearest neighbor on the uniform grid G, which has vertices
+at: $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,
+1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
+% 
+This way, we only have to run full Bayesian parameter estimation analysis on
+this fixed set of signals. 
+% 
+There are two sources of error that will enter the deductions we will make from
+a a single population generated in the manner described above. First, since the
+injection parameters are pushed to their nearest neighbor on a grid, we would 
+find discrete jumps in $\lambdans$ errors as a function of $N$. And second, an
+individual population is one particular realization of a stochastic process and
+could have excursions that may never be found in another population. To
+account for both of these limitations, we generate an ensemble of populations,
+and conservatively combine information from all of them\footnote{See 
+Sec.~\ref{s1:multiple_observations} for further details.}.
+
+We probe two astrophysical paradigms, one that allows for BH masses to lie
+within the astrophysical mass-gap (paradigm A), and one that does not (paradigm
+B).
+% 
+{\it For paradigm A}, we find the following: (i) for the softer equations of state
 that result in less deformable neutron stars, $15-20$ detections are sufficient
@@
 aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
-masses between $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
+masses $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
 restricting $\lambdans$ will require $50+$ NSBH observations.
 % 
-For paradigm B, we find the information accumulation to be somewhat slower.
-While the quantitative inferences for $\lambdans\geq1000$ populations are
+{\it For paradigm B}, we find the information accumulation to be somewhat slower.
+While the quantitative inferences for populations with $\lambdans\geq1000$ are
 not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
@@
 time-scale.
-}
-
-\prayush{\bf Tidal measurements versus observation time - in how many years can
-we constrain $\lambdans$ reasonably?}
-
-Lastly, we note
-that more a recent work that improves upon the waveform model we use in this
-study has been published~\cite{Pannarale:2015jka}, but it provides only an 
-amplitude model which needs to be (in future) augmented with a compatible
-phase model. In addition, accuracy of the underlying numerical simulations 
-used to calibrate the waveform model used here have not been verified against 
-independent codes so far.
-It is therefore difficult to assess the combined modeling error and its effect
-on our results. They are, therefore, limited by the limitations of our
-waveform model. However, we expect the combined effect of modeling errors to
-not change our qualitative conclusions.
+% 
+Our methods and results are detailed in Sec.~\ref{s1:multiple_observations}.
+
+Finally, we note that a more recent work that improves upon LEA+ has been
+published~\cite{Pannarale:2015jka}, but it
+provides only an amplitude model which needs to be (in future) augmented with a
+compatible phase model. In addition, accuracy of the underlying numerical
+simulations used to calibrate the waveform model used here have not been
+verified against independent codes so far. It is therefore difficult to assess
+the combined modeling error and its effect on our results. Our results here
+are, therefore, limited by the limitations of our waveform model, and presented
+with this caveat. However, we expect the combined effect of modeling errors to
+{\it not} affect our qualitative conclusions.
 

### 2016-09-21 Paper/paper.tex
more polishing

@@
 \begin{document}
-% \linenumbers
+\linenumbers
 
@@
 for our median measured value of $\lambdans$ to lie within $10\%$ of the true
-value. (ii) For NSBH populations with more deformable NSs
-($\lambdans\geq 1000$), the same is achievable within as few as $10$ realistic
+value. (ii) For NSBH populations with more deformable NSs ($\lambdans> 1000$),
+the same is achievable within as few as $10$ (or $15$ at most) realistic
 observations. (iii) The statistical uncertainty associated with $\lambdans$
 measurement can be restricted to be within $\pm50\%$ using $10-20$ observations
-when $\lambdans\geq 1000$), and using $25-40$ observations for softer equations
+when $\lambdans> 1000$), and using $25-40$ observations for softer equations
 of state. All of the above is possible within a few years of design
 aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
-masses $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
+masses $< 5M_\odot$ (i.e. in the mass-gap). However, further
 restricting $\lambdans$ will require $50+$ NSBH observations.
@@
 {\it For paradigm B}, we find the information accumulation to be somewhat slower.
-While the quantitative inferences for populations with $\lambdans\geq1000$ are
+While the quantitative inferences for populations with $\lambdans>1000$ are
 not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are

### 2016-09-21 Paper/paper.tex
mp comments v2

@@
 \usepackage{placeins}
+\usepackage[utf8]{inputenc}
 % \usepackage[switch,columnwise]{lineno} 
@@
 Bond:2002eh,Lattimer:2012nd,Manchester:2004bp},
-and in neutron star binaries~\cite{Hulse:1975uf,Taylor:1982wi,
+and in binary neutrons (BNS)~\cite{Hulse:1975uf,Taylor:1982wi,
 Weisberg:2010zz,Lattimer:2012nd,Manchester:2004bp}.
@@
 
-Another question that NSBH mergers can help answer is `what is the 
+Another question that compact object mergers can help answer is `what is the 
 nature of matter at nuclear densities', i.e. `what is the EoS of NS material'?
-This is what we will focus on in this paper.
+A large fraction of past work aimed at measuring NS matter effects from GW
+signals has consisted of inquiries about BNSs~\cite{Lee1999a,Lee1999b,Lee2000,
+oechslin:07,Read:2008iy,Markakis:2010mp,Markakis:2011vd,stergioulas:11,
+East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}. In this paper, we will
+instead focus on NSBHs.
 % 
@@
 \end{figure*}
-
+% 
 % ###############################
@@
 % #################
+% 
 \begin{figure*}
@@
 % 
-In addition to everything, we also note that $\arr_{\mchirp}\geq 0$ always,
+In addition, we also note that $\arr_{\mchirp}\geq 0$ always,
 i.e. $\mchirp$ is always being over-estimated. This is to be expected since
@@
 We expect that these conditions are also the ones which should maximize the
-likelihood of {\it measuring} tidal effects in NSBH signals. In this section,
+likelihood of {\it measuring} tidal effects in NSBH signals. Here,
 %
@@
 \end{figure*}
-%
-
-In the previous section, we showed that single observations of NSBH coalescences
-at moderate SNRs have little information about the internal structure of neutron
-stars that will be accessible to Advanced LIGO at its design sensitivity. We
-expect all neutron stars to share the same equation of state, and hence the same
-$\lambdans(\mns)$. In addition, we know that the mass distribution of (most) NSs
-that have not been spun up to millisecond periods (which are what we focus on in 
-this paper, by approximating $\chins\approx 0$) is narrowly peaked around
-$\sim 1.35M_\odot$~\cite{Kiziltan2013}. Therefore, information from multiple NSBH
-observations can be combined to improve our estimation of $\lambdans$ for the
-population. This is what we explore in this section, within a fully Bayesian 
-framework.
+
+In the previous section, we showed that single observations of NSBH
+coalescences at moderate SNRs have little information about the internal
+structure of neutron stars that will be accessible to Advanced LIGO at its
+design sensitivity. We expect all neutron stars to share the same equation of
+state, and hence the same $\lambdans(\mns)$. In addition, we know that the mass
+distribution of (most) NSs that have not been spun up to millisecond periods
+(which are the ones we focus on in this paper, by setting $\chins\approx 0$) is
+narrowly peaked around $\sim 1.35M_\odot$~\cite{Kiziltan2013}. Therefore,
+information from multiple NSBH observations can be combined to improve our
+estimation of $\lambdans$. We explore the same in this section within a fully
+Bayesian framework. We refer the reader to Ref.~\cite{Mandel:2009pc,Lackey2014,
+Wade:2014vqa} for similar analyses of BNS inspirals.
 
@@
 An intuitive understanding of the problem is gained by considering first
-multiple {\it identical} sources at realistic SNRs. Let us consider the case
+multiple {\it identical} sources with realistic but different SNRs. Let us consider the case
 of a population of optimally oriented binaries~\footnote{An optimally oriented
@@
 (uniformly) from $\mbh\in[3M_\odot,6.75M_\odot]$, BH spin (uniformly)
-from $\chibh\in[0, 1]$, source location uniformly in spatial volume,
-and its inclination from $\iota\in[0,\pi]$. Note that we only
-sample positive aligned BH spins, since binaries with anti-aligned spins have
-very little information to add at realistic SNRs, as shown in 
-Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}. This further restricts the
-sub-space of all NSBH observations that our populations span, and will need to
-be take into account when the number of observations required for tidal
-measurements is related to detector observation time.
-% 
-As noted previously, we also restrict events to have
-$\rho\geq\rho_\mathrm{min}$, where $\rho_\mathrm{min}=10$.
-Finally, we repeat the process till we have an ordered set of $N$ events.
+from $\chibh\in[0, 1]$, orbital inclination from $\iota\in[0,\pi]$, and 
+source location uniform in spatial volume\footnote{with a minimum SNR 
+$\rho_\mathrm{min}=10$}.
+% 
+We restrict ourselves to positive aligned BH spins, since binaries with
+anti-aligned spins have very little information to add at realistic SNRs,
+as demonstrated in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}. This is
+to be taken into account when the number of observations is related to detector
+operation time. 
+% 
+We repeat this process till we have an ordered set of $N$ events.
 % 
@@
 performed for a set of simulated signals whose parameters are the vertices
-of a regular hypercubic grid (henceforth ``G'') in the space of
+of a regular hyper-cubic grid (henceforth ``G'') in the space of
 $\{q\}\times\{\chibh\}\times\{\rho\}$, with each sampled at $q=\{2,3,4,5\}$,
 $\chibh=\{-0.5,0,0.5,0.75\}$, and $\rho=\{10,20,30,50,70\}$.
-All events in each population draw are substituted by their
-respective nearest neighbours on the grid G.
+All events in each population draw are substituted by their respective nearest
+neighbours on the grid G.
+% 
+Our chosen signal parameter distribution is different from some other studies
+in literature, which often sample from more astrophysically informed population
+distribution functions~\cite{Mandel:2009pc}.
 
@@
 % 
-We use the LEA+ model~\cite{Lackey:2013axa} to produce tidal waveforms that
-incorporate the effect
+We use the LEA+ model (c.f. Sec.~\ref{s2:waveforms}) to produce tidal waveforms
+that incorporate the effect
 of NS distortion during inspiral, and of its disruption close to merger. Our
@@
 Third, knowing that single observations can furnish only so much information
-about the NS equation of state, we move on to investigate how well do we do with
+about the NS equation of state, we move on to investigate how well we do with
 multiple signals. In order to quantify how $\lambdans$ measurement improves
@@
 In order to keep the computational cost reasonable, we make an additional
-approximation. For every population generated, we replace each event's
-parameters by their nearest neighbor on the uniform grid G, which has vertices
+approximation. For every population generated, we replace the parameters of each
+event by their nearest neighbor on the uniform grid G, which has vertices
 at: $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,
@@
 % 
-There are two sources of error that will enter the deductions we will make from
-a a single population generated in the manner described above. First, since the
-injection parameters are pushed to their nearest neighbor on a grid, we would 
+There are two sources of error that enter the deductions we make from
+a single population generated in the manner described above. First, since the
+injection parameters are pushed to their nearest neighbor on a grid, we
 find discrete jumps in $\lambdans$ errors as a function of $N$. And second, an
@@
 
-Finally, we note that a more recent work that improves upon LEA+ has been
-published~\cite{Pannarale:2015jka}, but it
-provides only an amplitude model which needs to be (in future) augmented with a
-compatible phase model. In addition, accuracy of the underlying numerical
-simulations used to calibrate the waveform model used here have not been
-verified against independent codes so far. It is therefore difficult to assess
-the combined modeling error and its effect on our results. Our results here
-are, therefore, limited by the limitations of our waveform model, and presented
-with this caveat. However, we expect the combined effect of modeling errors to
-{\it not} affect our qualitative conclusions.
+Finally, we note that the underlying numerical simulations used to calibrate
+the waveform model used here have not been verified against
+independent codes so far.
+% 
+It is therefore difficult to assess the combined modeling error of LEA+ and its
+effect
+on our results. Our results here are, therefore, limited by the limitations of
+our waveform model, and presented with this caveat. However, we do expect the
+combined effect of modeling errors to {\it not} affect our {\it qualitative}
+conclusions.
+% 
+In future, we plan to further the results presented here by using more recent
+tidal models~\cite{Pannarale:2015jka,Hinderer:2016eia}, that
+may improve upon LEA+\footnote{One of them~\cite{Pannarale:2015jka} is only an
+amplitude model though, which has to be augmented with a compatible phase model
+first.}.
 

### 2016-09-21 Paper/paper.tex
updated Fig.13

@@
 Our chosen signal parameter distribution is different from some other studies
-in literature, which often sample from more astrophysically informed population
-distribution functions~\cite{Mandel:2009pc}.
+in literature, which often sample from more astrophysically motivated population
+distribution functions~\cite{Mandel:2009pc}. We chose one that is sufficiently
+agnostic in absence of actual known NSBHs, and pragmatic enough for generating
+population ensembles.
 
@@
 of state. All of the above is possible within a few years of design
-aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
+aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs are allowed
 masses $< 5M_\odot$ (i.e. in the mass-gap). However, further
@@
 While the quantitative inferences for populations with $\lambdans>1000$ are
-not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
-required to attain the same measurement accuracy as under paradigm A.
+not affected significantly, we find that $\lambdans< 1000$ populations require 
+$10-20\%$ more events to attain the same measurement accuracy as under
+paradigm A. In either case, the accumulation of information from NSBH signals
+is likely slower than from BNS inspirals~\cite{Mandel:2009pc,Lackey2014,
+Wade:2014vqa}, although template models for the latter may be more
+uncertain due to missing point-particle PN terms at orders comparable to
+the tidal terms.
 % 
@@
 This, amongst other things, will allow us to rank different equations of 
-state for neutron star matter from most to least likely, with a few years'
-time-scale.
+state for neutron star matter from most to least likely, within a few years'
+detector operation.
 % 
@@
 combined effect of modeling errors to {\it not} affect our {\it qualitative}
-conclusions.
+conclusions, especially since the underlying point-particle component of LEA+
+includes all high-order terms, unlike past BNS studies~\cite{Lackey2014,
+Wade:2014vqa}
 % 

### 2016-09-26 Paper/paper.tex
forgot to comment out the gray text

@@
 \begin{document}
-\linenumbers
+%\linenumbers
 
@@
 
-\textcolor{gray}{%
-Let us first delve into statistical errors. In
-Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
-measure NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
-The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
-(bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
-as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
-$\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
-$1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
-does not significantly change the measurement uncertainties for non-tidal
-binary parameters, as is evident from comparing the two rows in each panel of
-Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
-% 
-From the left-most column, we find that: (i) at $\rho=20, \mchirp$ is already
-measured remarkably well - to a precision of $0.16\%$ of its true value, and
-(ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined more loosely,
-with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
-all three measurements gain further precision, with $\eta$ errors shrinking down
-to single-digit percents of its true value.
-% % 
-% We remind ourselves that this {\it precision}, however, is only
-% meaningful so long as the measurement is {\it accurate} to begin with.
-% We define $\arr_X$ as the ratio between systematic and statistical
-% errors associated with the measurement of parameter $X$,
-% \begin{equation}\label{eq:arr}
-% \arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
-% \end{equation}
-% in order to compare the relative importance of both. Only when
-% $|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
-% will not hamper the measurement of non-tidal parameters from NSBH signals.
-% Due to
-% unmodeled tidal effects, parameter measurements performed on NSBH signals with
-% BBH templates will have systematic biases. If these biases are too large,
-% using non-tidal BBH templates should severely impair our inferencing ability.
-% 
-% there are physical effects that have not been
-% incorporated in our (BBH) search templates which we use here to decipher NSBH
-% signals. Therefore, we cannot know the {\it accuracy} of our parameter
-% measurements unless we know how large the systematic biases due to an
-% incomplete template model are. To quantify this, we find it useful to define
-% $\arr_X$, the ratio of the systematic biases induced by unmodeled tidal effects in
-% templates, to the statistical uncertainties associated with the measurement
-% itself. If $\arr_X\gtrsim 1$, including the effects of NS tidal distortion
-% (during inspiral) and possible disruption (close to merger) would be imperative
-% to obtain {\it accurate} estimates of source masses and spins from observed GW
-% signals.}
-}
-
-\textcolor{blue}{%
+% \textcolor{gray}{%
+% Let us first delve into statistical errors. In
+% Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
+% measure NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
+% The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
+% (bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
+% as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
+% $\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
+% $1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
+% does not significantly change the measurement uncertainties for non-tidal
+% binary parameters, as is evident from comparing the two rows in each panel of
+% Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
+% % 
+% From the left-most column, we find that: (i) at $\rho=20, \mchirp$ is already
+% measured remarkably well - to a precision of $0.16\%$ of its true value, and
+% (ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined more loosely,
+% with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
+% all three measurements gain further precision, with $\eta$ errors shrinking down
+% to single-digit percents of its true value.
+% % % 
+% % We remind ourselves that this {\it precision}, however, is only
+% % meaningful so long as the measurement is {\it accurate} to begin with.
+% % We define $\arr_X$ as the ratio between systematic and statistical
+% % errors associated with the measurement of parameter $X$,
+% % \begin{equation}\label{eq:arr}
+% % \arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
+% % \end{equation}
+% % in order to compare the relative importance of both. Only when
+% % $|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
+% % will not hamper the measurement of non-tidal parameters from NSBH signals.
+% % Due to
+% % unmodeled tidal effects, parameter measurements performed on NSBH signals with
+% % BBH templates will have systematic biases. If these biases are too large,
+% % using non-tidal BBH templates should severely impair our inferencing ability.
+% % 
+% % there are physical effects that have not been
+% % incorporated in our (BBH) search templates which we use here to decipher NSBH
+% % signals. Therefore, we cannot know the {\it accuracy} of our parameter
+% % measurements unless we know how large the systematic biases due to an
+% % incomplete template model are. To quantify this, we find it useful to define
+% % $\arr_X$, the ratio of the systematic biases induced by unmodeled tidal effects in
+% % templates, to the statistical uncertainties associated with the measurement
+% % itself. If $\arr_X\gtrsim 1$, including the effects of NS tidal distortion
+% % (during inspiral) and possible disruption (close to merger) would be imperative
+% % to obtain {\it accurate} estimates of source masses and spins from observed GW
+% % signals.}
+% }
+
+%\textcolor{blue}{%
 Credible intervals $(\Delta X)^{90\%}$ showing the precision with which
@@
 Appendix~\ref{as1:nontidalerrors}.
-}
+%}
 We remind ourselves that this {\it precision} is only meaningful so long as the
@@
 %
-\textcolor{blue}{%
+% \textcolor{blue}{%
 Further illustrations, showing the correlation between tidal and non-tidal
@@
 We will continue here to focus on the measurement of $\lambdans$ itself.
-}
-% 
-\textcolor{gray}{%
-Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
-mass, spin, and tidal parameter measurements. We keep the binary parameters
-as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
-set $\rho=30$ (left panel) or $\rho=50$ (right panel).
-%
-We find that the measurement of $\lambdans$ is weakly degenerate with
-other parameters, and at realistic SNRs it would improve by a few tens of 
-percent if we knew non-tidal parameters to better accuracy. The predominant
-factor that would enhance the measurement accuracy for $\lambdans$ is the
-signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
-measurement to be limited by its degeneracy with non-tidal parameters at 
-a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
-}
+% }
+% 
+% \textcolor{gray}{%
+% Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
+% mass, spin, and tidal parameter measurements. We keep the binary parameters
+% as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
+% set $\rho=30$ (left panel) or $\rho=50$ (right panel).
+% %
+% We find that the measurement of $\lambdans$ is weakly degenerate with
+% other parameters, and at realistic SNRs it would improve by a few tens of 
+% percent if we knew non-tidal parameters to better accuracy. The predominant
+% factor that would enhance the measurement accuracy for $\lambdans$ is the
+% signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
+% measurement to be limited by its degeneracy with non-tidal parameters at 
+% a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
+% }
 

### 2016-09-26 Paper/paper.tex
improve intro

@@
 waves~\cite{LIGOVirgo2016a}. Emitted by a pair of coalescing black holes, these
-waves have heralded an era of observational GW astrophysics as they traveled 
+waves heralded an era of observational GW astrophysics as they traveled 
 through Earth.
@@
 Bond:2002eh,Lattimer:2012nd,Manchester:2004bp},
-and in binary neutrons (BNS)~\cite{Hulse:1975uf,Taylor:1982wi,
+and in binary neutron stars (BNS)~\cite{Hulse:1975uf,Taylor:1982wi,
 Weisberg:2010zz,Lattimer:2012nd,Manchester:2004bp}.
@@
 with NSBH binaries have used PN inspiral-only waveforms~\cite{Maselli:2013rza}.
-In doing so, however, (i) they ignore the merger signal which could contain significant
+In doing so, however, they ignore (i) the merger signal which could contain significant
 information for NSBHs, and (ii) the errors due to unknown vaccum terms in PN 
-waveforms which could dominate over the tidal terms themselves~\cite{Barkett2015,
+waveforms, which could dominate over the tidal terms themselves~\cite{Barkett2015,
 Yagi:2014}.
@@
 way, by using the waveform model of Lackey et al.~\cite{Lackey:2013axa}
-(abbreviated henceforth to ``LEA+'').
+(abbreviated henceforth to ``LEA'').
 % 
@@
 % 
-Fourth, we explore how our measurement errors decrease as we gain information
+And fourth, we explore how our measurement errors decrease as we gain information
 from multiple (realistic) events.
@@
 We now outline the main questions and results discussed in this paper.
-First, we probe the effect of ignoring tidal effects in parameter 
-estimation on the recovery of non-tidal binary parameters, such as 
-component masses and spins. This is the case of current and planned aLIGO
+First, we probe the effect of ignoring tidal effects in
+the recovery of non-tidal binary parameters, such as 
+component masses and spins. This is the case for current and planned aLIGO
 efforts.
-To do so, we first use the LEA+ model to generate a set of realistic signals;
+To do so, we first use the enhanced-LEA (or ``LEA+'', see Sec.~\ref{s2:waveforms})
+model to generate a set of realistic signals;
 and then use non-tidal (BBH) waveform filters to estimate the underlying
@@
 % MOVE the following to SUMMARY. Here, keep 1-2 sentences:
-We find that, for individual events, ignoring tidal effects will affect
-mass-estimation only marginally; only for very loud signals (SNRs $\gtrsim 30$)
+We find that, for individual events, ignoring tidal effects will affect mass
+and spin-estimation only marginally; only for very loud signals (SNRs $\gtrsim 30$)
 will the systematic biases be large enough to exceed the underlying
@@
 the same waveform model, which is the only one available of its kind in 
-current literature. A more recent work~\cite{Pannarale:2015jka}, improves on
-the GW amplitude model, which however needs to be augmented with a compatible
-phase model. 
-% In addition, accuracy of the underlying numerical simulations 
-% used to calibrate the waveform model used here have not been verified against 
-% independent codes so far.
-% It is therefore difficult to assess the combined modeling error and its effect
-% on our results. they are, therefore, limited by the limitations of our
-% waveform model. 
-However, we expect the combined effect of modeling errors to not change our 
-broad qualitative conclusions, which are based on vague statistical trends 
+current literature. A more recent work~\cite{Pannarale:2015jka} improves on
+the GW amplitude model for NSBHs, but it needs to be augmented with a compatible
+phase model. However, we expect the combined effect of modeling errors to not
+change our broad qualitative conclusions, which are based on statistical trends
 with large errors themselves.

### 2016-10-05 Paper/paper.tex
addressing ben's comments

@@
 multiple observations. Using an approximately astrophysical population, 
-we find that (i) the median measured $\lambdans$ catches up to within
-$10\%$ of the true value in approximately $20$ observations; (ii) in the same
+we find that (i) the median $\lambdans$ converges within $10\%$ of the true
+value in approximately $20$ observations; (ii) in the same
 number of observations, the associated $90\%$ credible intervals shrink to
@@
 % 
-On the other hand, {\it low-mass} BHs with $\mbh\in[3M_\odot, 7M_\odot]$, 
+On the other hand, {\it low-mass} BHs with $\mbh\in[3M_\odot, 8M_\odot]$, 
 can disrupt their companion NSs much before merger, forming long-sustained disks
@@
 First, we include tidal effects during inspiral and merger in a consistent
-way, by using the waveform model of Lackey et al.~\cite{Lackey:2013axa}
+way, by using the waveform model of Lackey {\it et al.}~\cite{Lackey:2013axa}
 (abbreviated henceforth to ``LEA'').
@@
 through to merger (i.e. LEA+)~\cite{Lackey:2013axa}. For most binaries with
-BH masses outside of the mass-gap and/or realistic signal-to-noise
+BH masses outside of the mass-gap $(2-5M_\odot)$~\cite{Bailyn:1997xt,
+Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa} and/or realistic signal-to-noise
 ratios (SNR), we find it difficult to put better than a factor of $2$ bound
@@
 (a) Our best estimate for $\lambdans$ improves much faster than our statistical
-uncertainties. Within $10-20$ detections, the median measured value of
+uncertainties. Within $10-20$ detections, the median value of
 $\lambdans$ lies within $10\%$ of the actual value.
@@
 % 
-(c) Further, if the astrophysical ``mass-gap'' is real, we find that $20-50\%$
+(c) Further, if the astrophysical ``mass-gap''~\cite{Bailyn:1997xt,
+Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa} is real, we find that $20-50\%$
 additional observations would be required to attain the same measurement
@@
 % that result in less deformable neutron stars, $15-20$ detections are sufficient
-% for our median measured value of $\lambdans$ to lie within $10\%$ of the true
+% for our median value of $\lambdans$ to lie within $10\%$ of the true
 % value. (ii) For NSBH populations with more deformable NSs
@@
 We also note that the parameter space probed is somewhat restricted by the
-domain of applicability of the LEA+ model which we use as filters.
-Finally, the accuracy of our quantitative results depends on the reliability of
-the same waveform model, which is the only one available of its kind in 
-current literature. A more recent work~\cite{Pannarale:2015jka} improves on
-the GW amplitude model for NSBHs, but it needs to be augmented with a compatible
-phase model. However, we expect the combined effect of modeling errors to not
-change our broad qualitative conclusions, which are based on statistical trends
-with large errors themselves.
+domain of applicability of the LEA+ model which we use as filters. Finally,
+the accuracy of our quantitative results depends on the reliability of the same
+waveform model, which is the only one available of its kind in current
+literature. A more recent work~\cite{Pannarale:2015jka} improves upon the 
+amplitude description of LEA+, but needs to be augmented with a compatible
+phase model. However, we expect our broad conclusions here to hold despite
+the intrinsic inaccuracies of LEA+ (with errors not exceeding
+$\mathcal{O}(10\%)$~\cite{Pannarale:2015jka}).
+% 
+% we expect the combined effect of modeling errors to not
+% change our broad qualitative conclusions, which are based on statistical trends
+% with large errors themselves.
 
@@
 % 
-In each panel: the dashed red line marks the median or {\it measured} value
+In each panel: the dashed red line marks the median value
 $\eta^\mathrm{Median}$, while the dashed green line show the true value
@@
 
-Lackey et al.~(LEA)~\cite{Lackey:2013axa} developed a complete inspiral-merger
+Lackey {\it et al.}~(LEA)~\cite{Lackey:2013axa} developed a complete inspiral-merger
 waveform model for disrupting NSBHs. Theirs is a frequency-domain
@@
 \begin{equation}\label{eq:postprob}
- p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
+ p(\vec{\Theta} | d_n, H) = \dfrac{p(d_n|\vec{\Theta}, H)\,p(\vec{\Theta} | H)}{p(d_n|H)}.
 \end{equation}
-Here, $p(\vec{\Theta} | K)$ is the a-priori probability of binary parameters
+Here, $p(\vec{\Theta} | H)$ is the {\it a priori} probability of binary parameters
 $\vec{\Theta}$
-taking particular values, given $K$ - which denotes all our collective knowledge,
+taking particular values, given $H$ - which denotes all our collective knowledge,
 except for expectations on binary parameters that enter
@@
 addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not calibrated for 
-$1\leq q\leq 2$. $p(d_n|\vec{\Theta}, K)$ is the
+$1\leq q\leq 2$. $p(d_n|\vec{\Theta}, H)$ is the
 likelihood of obtaining the given stretch of data $d_n$ if we assume that a
@@
 \begin{equation}\label{eq:likelihood}
- p(d_n| \vec{\Theta}, K) \equiv \LL(\vec{\Theta}) = \mathcal{N}\, \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+ p(d_n| \vec{\Theta}, H) \equiv \LL(\vec{\Theta}) = \mathcal{N}\, \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
 \end{equation}
@@
 coalescence time and phase.
-The denominator in Eq.~\ref{eq:postprob} is the a-priori probability of finding
+The denominator in Eq.~\ref{eq:postprob} is the {\it a priori} probability of finding
 the particular signal in $d_n$ and we assume that each injected signal is as
 likely as any other. Having constructed the probability distribution function
-$p(\vec{\Theta} | d_n, K)$, extracting the measured probability distribution
+$p(\vec{\Theta} | d_n, H)$, extracting the measured probability distribution
 for a single parameter (say $\alpha$) involves integrating
 \begin{equation}\label{eq:marginalize}
- p(\alpha | d_n, K) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, K),
+ p(\alpha | d_n, H) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, H),
 \end{equation}
@@
 the {\tt emcee} package~\cite{emcee}, to sample the probability distribution 
-$p(\vec{\Theta} | d_n, K)$. We run 100 independent chains, each of which is
+$p(\vec{\Theta} | d_n, H)$. We run 100 independent chains, each of which is
 allowed to collect 100, 000 samples and combine samples from chains that have
@@
 % \begin{equation}\label{eq:postprob}
-%  p(\vec{\Theta} | d_n, K) = \dfrac{p(d_n|\vec{\Theta}, K)\,p(\vec{\Theta} | K)}{p(d_n|K)}.
+%  p(\vec{\Theta} | d_n, H) = \dfrac{p(d_n|\vec{\Theta}, H)\,p(\vec{\Theta} | H)}{p(d_n|H)}.
 % \end{equation}
-% Here, $p(\vec{\Theta} | K)$ is the prior probability for binary parameters of
+% Here, $p(\vec{\Theta} | H)$ is the prior probability for binary parameters of
 % taking particular values. In this paper, we use a uniform prior on the
@@
 % because of model constraints for LEA+. The prior probability of obtaining a 
-% particular realization of data $p(d_n|K)$ is absorbed into the overall 
+% particular realization of data $p(d_n|H)$ is absorbed into the overall 
 % normalization. Finally, the first term in the numerator
-% $p(d_n|\vec{\theta}, \lambdans, K)$ is the likelihood of obtaining the given
+% $p(d_n|\vec{\theta}, \lambdans, H)$ is the likelihood of obtaining the given
 % stretch of data $d_n$ if we assume that a signal parameterized by
@@
 % \begin{equation}\label{eq:likelihood}
-%  \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, K) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+%  \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, H) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
 % \end{equation}
@@
 % \begin{equation}
-%  p(\alpha | d_n, K) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, K),
+%  p(\alpha | d_n, H) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, H),
 % \end{equation}
@@
 Eq.~\ref{eq:marginalize} to obtain marginalized probability distributions
-for the NS tidal deformability parameter: $p(\lambdans|d_n,K)$.
-The median value of this distribution is our {\it measured} value for
+for the NS tidal deformability parameter: $p(\lambdans|d_n,H)$. We will quote
+the median value of this distribution as our {\it measured} value for
 $\lambdans$, and the $90\%$ credible intervals associated with the distribution
-serve as statistical error-bars.
+as the statistical error-bars.
 %
@@
 % % 
-% In each panel: the dashed red line marks the median {\it measured} value
+% In each panel: the dashed red line marks the median value
 % $\eta^\mathrm{Median}$, while the dashed green line show the true value
-% $\eta^\mathrm{Injected}$. The difference between the true and measured values
+% $\eta^\mathrm{Injected}$. The difference between the true and median values
 % is the systematic bias introduced
@@
 % 
-In each panel- the dashed red line marks the median {\it measured} value for
+In each panel- the dashed red line marks the median value for
 $\lambdans$, and the dashed green line marks its {\it true} value.
@@
 BH); and (ii) technical constraints of our chosen LEA+ model~\cite{Lackey:2013axa}.
-At design sensitivity, if we expect $10-300$ NSBH detections a 
-year~\cite{Abadie:2010cfa}, {\it and} if BH mass values are uniformly likely
-up to $\sim 35M_\odot$~\cite{LIGOVirgo2016a}, we can expect to detect $0.3-7$
-{\it disruptive} NSBH mergers a year with $\rho\geq 20$, $0.07-2$ a year with
-$\rho\geq 30$, and $0.04-1$ a year with $\rho\geq 70$. Therefore, our injection
-parameters span a physically interesting subset of NSBH binaries, that is
-{\it also} likely observable in the near future. 
+At design sensitivity, if we expect $0.2-300$ NSBH detections a 
+year~\cite{Abadie:2010cfa}, we can expect to see $0.02-25$ 
+{\it disruptive}\footnote{We assume here that BH mass values are {\it uniformly}
+likely from $2M_\odot$ to $\sim 35M_\odot$~\cite{LIGOVirgo2016a}, but NSs are
+disrupted in NSBH mergers only if $q\leq 6$ and $\chibh\geq 0$~\cite{Foucart:2014nda,
+Foucart:2013psa}.
+} NSBH mergers a year, of which we will have $0.005-7$ observations with 
+$\rho\geq 20$, and $0.002-3$ a year with $\rho\geq 30$.
+Therefore, our injection parameters span a physically interesting subset of NSBH
+binaries, that is {\it also} likely observable in the near future. 
 For our Bayesian priors, we choose uniform distributions for both component
@@
 and $-0.75\leq \chibh\leq +0.75$.
-% \prayush{FIXME}
 
@@
 $\mbh=5.4M_\odot$, $\chibh=+0.5$ and $\lambdans=2000$). In each panel, both
-the {\it true} and the {\it measured} (i.e. median) values of $\eta$ are marked.
-The effect of ignoring tidal
-corrections in templates manifests as a systematic shift of the measured $\eta$
-value away from its true value. The associated {\it systematic} bias/error is
-measured by the difference between the measured and true $\eta$ values. Darker
+the {\it true} and the median values of $\eta$ are marked.
+The effect of ignoring tidal corrections in templates manifests as a systematic
+shift of the median $\eta$ value away from its true value. We measure
+the associated {\it systematic} bias/error by the differences between the
+true and median $\eta$ values. Darker
 shading of the probability distributions marks the $90\%$ credible interval,
@@
 in order to compare the relative magnitude of both. Only when
-$|\arr_X| < 1$ can we ignore tidal effects in our templates
+$|\arr_X| \ll 1$ can we ignore tidal effects in our templates
 without hampering the measurement of non-tidal parameters from NSBH signals.
-% 
+When $\arr_X$ approaches a few tens of percent of unity, we can begin to favor
+tidal templates for NSBH studies.
+
 We start with calculating $\arr_{\mchirp}$ as a function of various source
@@
 % 
-In addition, we also notice that, unlike for $\mchirp$, the measured value of
+In addition, we also notice that, unlike for $\mchirp$, the median value of
 $\eta$ is always {\it lower} than its true value, which is what we expect if we
@@
 % % 
-% In each panel- the dashed red line marks the median {\it measured} value for
+% In each panel- the dashed red line marks the median value for
 % $\lambdans$, and the dashed green line marks its {\it true} value.
@@
 % Each panel corresponds a unique population with injected $\lambdans$ fixed to
-% the value shown in its title. In each panel, the measured median for $\lambdans$
+% the value shown in its title. In each panel, the recovered median for $\lambdans$
 % is shown by filled circles, and the $90\%$ credible interval is shown by the 
@@
 $\lambdans$ given in the legend. For each population, we show the median 
-measured $\lambdans$ value (as filled circles), as well as the associated
+$\lambdans$ value (as filled circles), as well as the associated
 $90\%$ credible intervals for the measurement (as the vertical extent of the
@@
 {\it Bottom left}: Here we show the interval of $\lambdans$ values within
-which the median measured $\lambdans$ lies for $90\%$ of the populations in
+which the median $\lambdans$ lies for $90\%$ of the populations in
 each ensemble shown in the top left panel.
@@
 % 
-As before, let $K$ denote all of our collective prior knowledge; for instance,
-$K$ includes our assumption that all NSs in a single population have the same
+As before, let $H$ denote all of our collective prior knowledge; for instance,
+$H$ includes our assumption that all NSs in a single population have the same
 deformability parameter $\lambdans$, and that its cumulative measurement is
@@
 \begin{eqnarray}
- p(\lambdans |&& \hspace{-4mm}d_1, d_2, \cdots, d_N, K)\hspace{50mm}\nonumber\\ &=& \dfrac{p(d_1,d_2,\cdots,d_N |\lambdans , K)\,p(\lambdans|K)}{\int p(\lambdans |K) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans},\label{eq:p11}\\
-  &=& \dfrac{p(\lambdans|K) \prod_{i} p(d_i|\lambdans, K)}{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans},\label{eq:p12} \\
-  &=& \dfrac{p(\lambdans|K) \prod_{i} \left( p(\lambdans |d_i, K)\dfrac{p(d_i)}{p(\lambdans|K)} \right)}{\int\, p(\lambdans|K )\, p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans}\label{eq:p13};
+ p(\lambdans |&& \hspace{-4mm}d_1, d_2, \cdots, d_N, H)\hspace{50mm}\nonumber\\ &=& \dfrac{p(d_1,d_2,\cdots,d_N |\lambdans , H)\,p(\lambdans|H)}{\int p(\lambdans |H) p(d_1,d_2,\cdots,d_N |\lambdans , H)\D\lambdans},\label{eq:p11}\\
+  &=& \dfrac{p(\lambdans|H) \prod_{i} p(d_i|\lambdans, H)}{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , H)\D\lambdans},\label{eq:p12} \\
+  &=& \dfrac{p(\lambdans|H) \prod_{i} \left( p(\lambdans |d_i, H)\dfrac{p(d_i)}{p(\lambdans|H)} \right)}{\int\, p(\lambdans|H )\, p(d_1,d_2,\cdots,d_N |\lambdans , H)\D\lambdans}\label{eq:p13};
 \end{eqnarray}
@@
 \begin{eqnarray}
- p(&&\hspace{-4mm}\lambdans | d_1, d_2, \cdots, d_N, K)\hspace{50mm}\nonumber\\
-%   &=& \dfrac{p(\lambdans) \left(\dfrac{p(d)}{p(\lambdans)}\right)^N\prod_{i} p(\lambdans |d_i, K) }{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , K)\D\lambdans}\label{eq:p21},\\
-  &=& p(\lambdans)^{1-N}\times \dfrac{p(d)^N}{\int p(\lambdans) p(d_1, d_2, \cdots, d_N |\lambdans, K)\D\lambdans}\nonumber\\ &&\hspace{3mm}\times\prod_i p(\lambdans |d_i, K)\label{eq:p22},
+ p(&&\hspace{-4mm}\lambdans | d_1, d_2, \cdots, d_N, H)\hspace{50mm}\nonumber\\
+%   &=& \dfrac{p(\lambdans) \left(\dfrac{p(d)}{p(\lambdans)}\right)^N\prod_{i} p(\lambdans |d_i, H) }{\int p(\lambdans ) p(d_1,d_2,\cdots,d_N |\lambdans , H)\D\lambdans}\label{eq:p21},\\
+  &=& p(\lambdans)^{1-N}\times \dfrac{p(d)^N}{\int p(\lambdans) p(d_1, d_2, \cdots, d_N |\lambdans, H)\D\lambdans}\nonumber\\ &&\hspace{3mm}\times\prod_i p(\lambdans |d_i, H)\label{eq:p22},
 \end{eqnarray}
 % 
-where $p(\lambdans|K)$ is written $p(\lambdans)$ for brevity.
-The first two factors in Eq.~\ref{eq:p22} can be absorbed into a
-normalization factor $\mathcal{N}$, simplifying it to
+where the prior probability $p(\lambdans|H)$ is written $p(\lambdans)$ for
+brevity. {\it A priori}, we assume that no particular value of $\lambdans$ is
+preferred over another within the range $[0, 4000]$, i.e.
+\begin{equation}\label{eq:lprior}
+ p(\lambdans | H) = \dfrac{1}{4000}\,\mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right).
+\end{equation}
+With a uniform prior, the first two factors in Eq.~\ref{eq:p22} can be
+absorbed into a normalization factor $\mathcal{N}$, simplifying it to
 % 
 \begin{equation}\label{eq:lambdaMultiple}
- p(\lambdans | d_1, d_2, \cdots, d_N; K) = \mathcal{N}\prod_{i=1}^N p(\lambdans | d_i, K).
+ p(\lambdans | d_1, d_2, \cdots, d_N; H) = \mathcal{N}\prod_{i=1}^N p(\lambdans | d_i, H).
 \end{equation}
@@
 % Eq.~\ref{eq:lambdaMultiple} includes the following implicit factor to normalize
-% its right-hand side: $\dfrac{\prod_i p(d_i|K)}{\int p(\lambdans)\,p(d_1, d_2, \cdots, d_N | \lambdans; K)\,\D\lambdans}$}
+% its right-hand side: $\dfrac{\prod_i p(d_i|H)}{\int p(\lambdans)\,p(d_1, d_2, \cdots, d_N | \lambdans; H)\,\D\lambdans}$}
 % \begin{equation}\label{eq:lambdaMultiple}
-%  p(\lambdans | d_1, d_2, \cdots, d_N; K) = p(\lambdans | K)^{1-N}\prod_{i=1}^N p(\lambdans | d_i, K),
+%  p(\lambdans | d_1, d_2, \cdots, d_N; H) = p(\lambdans | H)^{1-N}\prod_{i=1}^N p(\lambdans | d_i, H),
 % \end{equation}
-A priori, we assume that no particular value of $\lambdans$ is preferred
-over another, and restrict $\lambdans$ to $[0, 4000]$, i.e.
-\begin{equation}\label{eq:lprior}
- p(\lambdans | K) = \dfrac{1}{4000}\,\mathrm{Rect}\left(\frac{\lambdans-2000}{4000}\right).
-\end{equation}
+
 % 
 In the second set of terms in Eq.~\ref{eq:lambdaMultiple} (of the form 
-$p(\lambdans | d_i, K)$), each is the probability distribution for $\lambdans$
+$p(\lambdans | d_i, H)$), each is the probability distribution for $\lambdans$
 inferred {\it a posteriori} from the \textit{i}-th observation by marginalizing
 \begin{equation}\label{eq:margpost}
- p(\lambdans | d_i, K) = \int\, p(\vec{\theta}, \lambdans | d_i, K)\, \D \vec{\theta},
+ p(\lambdans | d_i, H) = \int\, p(\vec{\theta}, \lambdans | d_i, H)\, \D \vec{\theta},
 \end{equation}
-where $p(\vec{\theta}, \lambdans | d_i, K)$ is the inferred joint probability 
+where $p(\vec{\theta}, \lambdans | d_i, H)$ is the inferred joint probability 
 distribution of all source parameters $\vec{\theta}\cup\{\lambdans\}$ for the 
 $i$-th event, as given by Eq.~\ref{eq:postprob}. We note that 
-Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR} illustrates $p(\lambdans | d_i, K)$
+Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR} illustrates $p(\lambdans | d_i, H)$
 for three individual events. By substituting
@@
 In the right panel, we present information derived from the left panel.
-The line-circle curve shows the measured median from $N$ observations.
+The line-circle curve shows the measured median value from $N$ observations.
 The pair of dashed (dotted) horizontal lines mark
@@
 events. This figure somewhat quantifies the qualitative deductions we made
-from the left panel. We find that the measured median does track the true value
+from the left panel. We find that the median does track the true value
 quickly, reaching $10\%$ accuracy with $10-15$ observations. With the same
@@
 right panel of Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}, the line-circle curves
-track the median measured $\lambdans$, while the filled regions show
+track the median $\lambdans$, while the filled regions show
 the associated $90\%$ credible intervals. From the figure, we observe
@@
 $N$ happens in a similar manner for each $\lambdans$, (ii) within $\sim 10$
-observations, median measured $\lambdans$ for all populations lie within
+observations, median $\lambdans$ for all populations lie within
 $10\%$ of the true values, and (iii) it takes
@@
 this, we discuss ensembles of population draws next. In
-Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} we show the measured
-({median}) $\lambdans$ as a function of the number of observed
+Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} we show the median
+$\lambdans$ as a function of the number of observed
 events, for four population ensembles, with a hundred population draws
 in each ensemble. Lets focus on the {\it top left} panel first. In it, we
-show the measured (median) $\lambdans$ for all populations in four ensembles,
+show the median $\lambdans$ for all populations in four ensembles,
 with true $\lambdans=\{2000,1500,1000,500\}$ from top to bottom.
@@
 The panel just below it shows the range of $\lambdans$ that encloses the
-median measured $\lambdans$ for $90\%$ of the populations in {\it each}
+median $\lambdans$ for $90\%$ of the populations in {\it each}
 ensemble. In other words, this panel shows the range of $\lambdans$ within which
-the median measured $\lambdans$ value for $90\%$ of NSBH populations is
-expected to lie. From these panels, we conclude that our measured $\lambdans$
+the median $\lambdans$ value for $90\%$ of NSBH populations is
+expected to lie. From these panels, we conclude that our median $\lambdans$
 values will be within $10\%$ of the {\it true} value after $\sim 25$
@@
 expectedly find that information accumulation is much slower. It would
-take $25-40$ detections with $\rho\geq10$ under this paradigm, for our (median)
-measured $\lambdans$ to converge within $10\%$ of its true value.
+take $25-40$ detections with $\rho\geq10$ under this paradigm, for our median
+$\lambdans$ to converge within $10\%$ of its true value.
 % % 
@@
 (i) for more deformable neutron stars (harder equation of states), the median
-measured value of $\lambdans$ comes within $10\%$ of the true value with as 
+value of $\lambdans$ comes within $10\%$ of the true value with as 
 few as $10$ events, while achieving the same accuracy for softer equations of 
@@
 that result in less deformable neutron stars, $15-20$ detections are sufficient
-for our median measured value of $\lambdans$ to lie within $10\%$ of the true
+for our median value of $\lambdans$ to lie within $10\%$ of the true
 value. (ii) For NSBH populations with more deformable NSs ($\lambdans> 1000$),

### 2016-10-10 emcee_PE/scripts/MultipleEventGeneration-N80-20-AstroPop.py
improved basic scripts

@@ -138,7 +138,7 @@ Nburnin  = 2000
 # Read in parameter biases and other data
 ######################################################
 datafile_postfix = 'TT_ParameterBiasesAndConfidenceIntervals.h5'
-datafile_dir     = '/home/prayush/research/NSBH/TidalParameterEstimation/set006/'
+datafile_dir     = '/home/prayush/projects/nsbh/TidalParameterEstimation/nsbh_tidal/PEData/'
 
 # Load in all the Posterior distribution function samples
 all_posterior_chains   = {}
@@ -185,7 +185,14 @@ for q in qvec:
 # In[4]:
 
 ## INITIATE A SET OF MULTIPLE OBSERVATIONS
+
+# Constants
 NEVENTS = int(sys.argv[1])
+#
+RAND = 0
+if len(sys.argv > 2):
+  RAND = int(sys.argv[2])
+#
 INDEX   = 20
 
 
@@ -194,7 +201,14 @@ INDEX   = 20
 xx = {}
 for NSL in Lambdavec:
     print "Initiating event class with Lambda = %f" % NSL
-    xx[NSL] = MO.multiple_observation_results(lambda_posterior_chains,                                      N=NEVENTS, NSLambda=NSL,                                      chi2vec=chi2vec, qvec=qvec,                                      SNRvec=SNRvec,                                      source_distribution='UniformIn3DVolumeInclination',                                      kernel='gau',                                      verbose=True)
+    xx[NSL] = MO.multiple_observation_results(lambda_posterior_chains,\
+                            N=NEVENTS, NSLambda=NSL,\
+                            chi2vec=chi2vec, qvec=qvec,\
+                            SNRvec=SNRvec,\
+                            source_distribution='UniformIn3DVolumeInclination',\
+                            kernel='gau',\
+                            RND=RAND,\
+                            verbose=True)
 
 
 # In[6]:

### 2016-10-10 emcee_PE/scripts/MultipleObservations.py

@@ -284,7 +284,7 @@ Return the nearest neighbor of the drawn SNR from the list of SNR values given
       Fp = Fplus(sky_theta, sky_phi, polarization)
       Fc = Fcross(sky_theta, sky_phi, polarization)
       # Next, calculate the expected SNR of the event
-      hp, hc = get_fd_waveform(approximant='IMRPhenomD',\
+      hp, hc = get_fd_waveform(approximant='IMRPhenomC',\
                                 mass1=q*1.35, mass2=1.35, spin1z=chi,\
                                 distance=distance, inclination=inclination,\
                                 f_lower=f_lower, delta_f=1./time_length)

### 2016-10-10 emcee_PE/src/utils.py
bugfix in converting Mc,eta chains to m1,m2

@@ -448,8 +448,8 @@ def load_samples_safe(dataDir, SNR, burnin=1500, burnend=0,\
 
     s_q = qfun(s_eta)
     s_M = Mfun(s_Mc, s_eta)
-    s_m1 = m1fun(s_Mc, s_q)
-    s_m2 = m2fun(s_Mc, s_q)
+    s_m1 = m1fun(s_M, s_q)
+    s_m2 = m2fun(s_M, s_q)
     s_chieff = chi_eff(s_eta, s_chi1, s_chi2)
     s_chiPN = chi_PN(s_eta, s_chi1, s_chi2)
     

### 2016-10-11 Paper/paper.tex
first bout of changes from Bens comments

@@
 % 
-(a) Our best estimate for $\lambdans$ improves much faster than our statistical
-uncertainties. Within $10-20$ detections, the median value of
-$\lambdans$ lies within $10\%$ of the actual value.
+(a) Our median estimate for $\lambdans$ starts out prior dominated, but 
+converges to within $10\%$ of the true value within $10-20$ detections.
 (b) Measurement uncertainties for $\lambdans$, on the other hand, depend on
@@
 \textbf{at different SNR values (shown in legend).}\par\medskip
-\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
+\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}\\
 \caption{We show here probability distributions for mass ratio $\eta$ as measured
@@
 \includegraphics[width=1.6\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
+\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
+% \includegraphics[width=0.67\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
+\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
+\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
+\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
 \caption{We show here the ratio of systematic and statistical
@@
 \centering
-\textbf{Paradigm A \hspace{6cm} Paradigm B}\par\medskip
+\textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
 \includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
@@
 \centering    
-\textbf{Paradigm A \hspace{6cm} Paradigm B}\par\medskip
+\textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
 \includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{plots/LambdaCIWidths90pc_vs_N_AllPopulation}
@@
 for a single population with neutron star deformability $\lambdans=800$.
-In the left panel, each curve shows the 
+In the top panel, each curve shows the 
 probability distributions for $\lambdans$ as inferred from $N$ events, with $N$
@@
 with each of the probability distribution curves. The first few observations
-do not have
-enough information to bound $\lambdans$ much more than
-our prior from Eq.~\ref{eq:lprior} does. However, the median of these
-cumulative distributions appears to track the true value of $\lambdans$ 
-rapidly.
-% 
-In the right panel, we present information derived from the left panel.
+do not have enough information to bound $\lambdans$ much more than
+our prior from Eq.~\ref{eq:lprior} does. 
+% 
+In the bottom panel, we present information derived from the top panel.
 The line-circle curve shows the measured median value from $N$ observations.
@@
 events. This figure somewhat quantifies the qualitative deductions we made
-from the left panel. We find that the median does track the true value
-quickly, reaching $10\%$ accuracy with $10-15$ observations. With the same
-information, our credible intervals also shrink to $\pm 25\%$.
+from the left panel. We find that the median does track the true value quickly,
+reaching within its $10\%$ with $10-15$ observations. This is as one expects of 
+injections in zero noise where random fluctuations are unable to shift the
+median away from the true value, so long as the measurement is not restricted
+by the prior. With the same information, our credible intervals also shrink to $\pm 25\%$.
 % 
@@
 the following: (i) the shrinkage of credible interval widths with increasing
-$N$ happens in a similar manner for each $\lambdans$, (ii) within $\sim 10$
-observations, median $\lambdans$ for all populations lie within
-$10\%$ of the true values, and (iii) it takes
+$N$ happens in a similar manner for each $\lambdans$, 
+% (ii) within $\sim 10$ observations, median $\lambdans$ for all populations lie
+% within $10\%$ of the true values, 
+and (ii) it takes
 approximately $20$ events to distinguish definitively (with $90\%$ credibility)
@@
 the median $\lambdans$ value for $90\%$ of NSBH populations is
-expected to lie. From these panels, we conclude that our median $\lambdans$
+expected to lie. From these panels, we observe that our median $\lambdans$
 values will be within $10\%$ of the {\it true} value after $\sim 25$
@@
 after as few as $15$ detections of more deformable neutron stars
-($\lambdans\geq 1500$).
+($\lambdans\geq 1500$). This is not surprising because we inject simulated
+signals in {\it zero} noise, which ensures that the median not be shifted away
+from the true value. That it takes $15+$ events for the median to approach
+the true value is a manifestation of the fact that the measurement is limited
+by the prior on $\lambdans$ when we have fewer than $15$ events.
 % 
@@
 % 
-{\it For paradigm A}, we find the following: (i) for the softer equations of state
-that result in less deformable neutron stars, $15-20$ detections are sufficient
-for our median value of $\lambdans$ to lie within $10\%$ of the true
-value. (ii) For NSBH populations with more deformable NSs ($\lambdans> 1000$),
+{\it For paradigm A}, we find the following: (i) for the softer equations of
+state that result in less deformable neutron stars, $15-20$ detections bring
+the measured probability distribution for $\lambdans$ entirely within the prior,
+which ensures that the median $\lambdans$ tracks the true value to within $10\%$.
+(ii) For NSBH populations with more deformable NSs ($\lambdans> 1000$),
 the same is achievable within as few as $10$ (or $15$ at most) realistic

### 2016-10-11 emcee_PE/scripts/SingleObservation.py

@@ -52,8 +52,12 @@ ErrThresh=[100, 200]
 ######################################################
 linestyles = ['solid', 'dashed', 'dashdot', 'dashed', 'dotted', 'dashdot', 'dashed', 'dotted']
 #linecolors = ['crimson', 'olivedrab', 'k', 'b', 'm', 'y']
-linecolors = ['crimson', 'darkorange', 'olivedrab', 'royalblue', 'purple', 'k']
+linecolors = ['crimson', 'olivedrab', 'royalblue', 'purple', 'darkorange', 'k']
+linecolors = ['purple', 'royalblue', 'olivedrab', 'crimson', 'darkorange', 'k']
 linemarkers= ['', 'x', 'o', '^']
+linewidths = [1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5]
+linewidths = [0.5, 1, 2, 4, 8]
+#linewidths = [1, 2, 3, 4, 8]
 
 gmean = (5**0.5 + 1)/2.
 
@@ -440,6 +444,114 @@ def make_multilines(XY, labels=None, xlabel='SNR', ylabel='', clabel='', title='
   plt.savefig(figname)
   return
 
+
+## UNDER CONSTRUCTION
+def make_multilines_2(XY,\
+                sort_group1=True, label_group1_idx=0,\
+                sort_group2=True, label_group2_idx=0,\
+                labels=None, xlabel='SNR', ylabel='', clabel='',
+                title='', in_title='',\
+                levelspacing=0.25, vmin=None, vmax=None, cmap=cm.Reds_r, logY=True,\
+                pcolors=linecolors, plines=linestyles, pmarkers=linemarkers,\
+                alpha=0.8,\
+                xmin=None, xmax=None, ymin=None, ymax=None,\
+                constant_line=None,\
+                single_legend=True,\
+                cbfmt='%.1f', figname='plot.png'):
+  #
+  colwidth = 4.8  
+  plt.rcParams.update({\
+    'legend.fontsize':12, \
+    'text.fontsize':16,\
+    'axes.labelsize':16,\
+    'font.family':'serif',\
+    'font.size':16,\
+    'xtick.labelsize':16,\
+    'ytick.labelsize':16,\
+    'figure.subplot.bottom':0.2,\
+    'figure.figsize':figsize, \
+    'savefig.dpi': 300.0, \
+    'figure.autolayout': True})
+  
+  ngrp1 = len(XY.keys())
[…]

### 2016-10-11 emcee_PE/scripts/SingleObservation.py
updated fig.6, fixed bug in SO.py

@@ -787,11 +787,13 @@ def get_results(data, q=None, chiBH=None, NSLmbd=None, SNR=None,\
   * 'Lambda'
 
   Quantity requested in **qnt** has to be one of :
-  * 'median-val': X(median)
-  * 'fbias'     : (X(median) - X(injection)) / X(injection)
-  * 'CIlower'   : X[10%] or 100-x%
-  * 'CIhigher'  : X[90%] or x%
-  * 'CIfwidth'  : (X[90%] - X[10%]) / X(injection)
+  * 'median-val'    : X(median)
+  * 'maxLogL-val'   : X(MaxLikelihood)
+  * 'fbias'         : (X(median) - X(injection)) / X(injection)
+  * 'maxLogL-fbias' : (X(MaxLikelihood) - X(injection)) / X(injection)
+  * 'CIlower'       : X[10%] or 100-x%
+  * 'CIhigher'      : X[90%] or x%
+  * 'CIfwidth'      : (X[90%] - X[10%]) / X(injection)
   '''
   if q is None or chiBH is None or NSLmbd is None or SNR is None:
     raise IOError("Need all of (q,chi,NSL,SNR) parameters")
@@ -833,8 +835,8 @@ def get_results(data, q=None, chiBH=None, NSLmbd=None, SNR=None,\
   else:
     if 'median-val' in qnt: pidx += 0
     elif 'maxLogL-val' in qnt: pidx += 1
-    elif 'fbias' in qnt: pidx += 2
     elif 'maxLogL-fbias' in qnt: pidx += 3
+    elif 'fbias' in qnt: pidx += 2
     elif 'CIlower' in qnt: pidx += 4
     elif 'CIhigher' in qnt: pidx += 5
     elif 'CIfwidth' in qnt: pidx += 6

### 2016-10-11 Paper/paper.tex
improved fig13

@@
 \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
-\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
-\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+\textbf{for signals at different SNR values; $\rho = 30$ (left) and $50$ (right).}\par\medskip
+\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr30}%\\
 \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
@@
 We illustrate here two sets of two-dimensional joint probability distributions,
-differing only in signal strength, with $\rho=20$ in the left panel, and
+differing only in signal strength, with $\rho=30$ in the left panel, and
 $\rho=50$ in the right. The injected parameters are 

### 2016-10-12 Paper/paper.tex
improved fig13 caption
changed fig.13 again

--- 2016-10-11 Paper/paper.tex
@@
 \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr30}%\\
 \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
@@
 $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
-$\lambdans=2000$. Templates include tidal effects, as evident in the bottom rows
+$\lambdans=2000$. Contours are shown for $\{1-,2-,3-\}\sigma$ confidence levels.
+Templates include tidal effects, as evident in the bottom rows
 of both panels which show the correlation of $\lambdans$ with non-tidal 
@@
 as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
-set $\rho=30$ (left panel) or $\rho=50$ (right panel).
+set $\rho=20$ (left panel) or $\rho=50$ (right panel).
 %

### 2016-10-12 emcee_PE/scripts/SingleObservation.py
added 1D plots of sys bias vs SNR

@@ -66,18 +66,18 @@ ppi=72.0
 aspect=(5.**0.5 - 1) * 0.5
 size=4.0 * 2# was 6
 figsize=(size,aspect*size)
-plt.rcParams.update({\
-'legend.fontsize':11, \
-'text.fontsize':11,\
-'axes.labelsize':11,\
-'font.family':'serif',\
-'font.size':11,\
-'xtick.labelsize':11,\
-'ytick.labelsize':11,\
-'figure.subplot.bottom':0.2,\
-'figure.figsize':figsize, \
-'savefig.dpi': 300.0, \
-'figure.autolayout': True})
+#plt.rcParams.update({\
+#'legend.fontsize':11, \
+#'text.fontsize':11,\
+#'axes.labelsize':11,\
+#'font.family':'serif',\
+#'font.size':11,\
+#'xtick.labelsize':11,\
+#'ytick.labelsize':11,\
+#'figure.subplot.bottom':0.2,\
+#'figure.figsize':figsize, \
+#'savefig.dpi': 300.0, \
+#'figure.autolayout': True})
 
 # hist(PercentileInterval(s1[0], pc=99.0), 100);
 
@@ -108,13 +108,13 @@ col corresponds to the lower level group
     raise IOError("X, Y and Z arrays have different number of sets to plot")
   
   plt.rcParams.update({\
-    'legend.fontsize':16, \
-    'text.fontsize':16,\
-    'axes.labelsize':16,\
+    'legend.fontsize':18, \
+    'text.fontsize':18,\
+    'axes.labelsize':18,\
     'font.family':'serif',\
-    'font.size':16,\
-    'xtick.labelsize':14,\
-    'ytick.labelsize':14,\
+    'font.size':18,\
+    'xtick.labelsize':15,\
+    'ytick.labelsize':15,\
     'figure.subplot.bottom':0.2,\
     'figure.figsize':figsize, \
     'savefig.dpi': 300.0})#, \
@@ -124,7 +124,7 @@ col corresponds to the lower level group
   if vverbose: print "Making plot with %d rows, %d cols" % (nrow, ncol)
   pltid = 0
   
-  colwidth = 1.75
+  colwidth = 1.55
   fig = plt.figure(int(1e7 * np.random.random()), \
               figsize=((1.3*gmean*ncol+1.3)*colwidth, 1.2*colwidth*nrow),\
               dpi=100)

### 2016-10-17 Paper/paper.tex
insert 1d figures for param biases in paper
finalizing figures

@@
 \textbf{ignoring tidal effects.}\par\medskip
-\includegraphics[width=1.6\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
-\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
-% \includegraphics[width=0.67\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
-\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
-\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
-\includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
+\includegraphics[width=1.95\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
+% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
+% % \includegraphics[width=0.67\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
+% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
+% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
+% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
 \caption{We show here the ratio of systematic and statistical
@@
 \textbf{ignoring tidal effects.}\par\medskip
-\includegraphics[width=2\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_70_linear}
+\includegraphics[width=1.95\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_50_linear}
+% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
+% % \includegraphics[width=0.67\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
+% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
+% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
+% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
 \caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
@@
 \includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
-% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
-% \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
+% % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
+% % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
+% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
+% % \includegraphics[width=0.67\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
+% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
+% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
+% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
 \caption{This figure shows the ratio of the systematic and statistical
@@
 \textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
-\includegraphics[trim={2cm 0 0 0},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
+\includegraphics[trim={3cm 0 0 0.5cm},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
 % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}

### 2016-10-17 emcee_PE/scripts/SingleObservation.py

@@ -85,7 +85,8 @@ figsize=(size,aspect*size)
 # Plotting functions
 def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
         title='', titles=[], xticks=None, yticks=None,\
-	levellinesscale=None, levellines=[100,200], levellabels=True, levelspacing=0.25,\
+	    levellinesscale=None, levellines=[100,200], levellabels=True, levelspacing=0.25,\
+                       leveltextcolor='b', leveltextfontsize=10,\
         vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f', lfmt='%.1f',\
         xmin=None, xmax=None, ymin=None, ymax=None, alpha=0.9,\
         colorticks=None, colorbartype='simple', figname='plot.png'):
@@ -118,16 +119,16 @@ col corresponds to the lower level group
     'figure.subplot.bottom':0.2,\
     'figure.figsize':figsize, \
     'savefig.dpi': 300.0})#, \
-  #'figure.autolayout': True})
+    #'figure.autolayout': True})
 
   nrow, ncol, _ = np.shape(X)
   if vverbose: print "Making plot with %d rows, %d cols" % (nrow, ncol)
   pltid = 0
   
-  colwidth = 1.55
+  colwidth = 1.7
   fig = plt.figure(int(1e7 * np.random.random()), \
               figsize=((1.3*gmean*ncol+1.3)*colwidth, 1.2*colwidth*nrow),\
-              dpi=100)
+              dpi=300)
   fig.clf()
   grid = ImageGrid(fig, 111,\
             nrows_ncols=(nrow, ncol), \
@@ -193,9 +194,9 @@ col corresponds to the lower level group
         for c in cset.collections: c.set_linestyle('dotted')
         if levellabels:
           label_dict = {}
-          for ll in [lev * levline_scaling]: label_dict[ll] = str(int(lev*100))
+          for ll in [lev * levline_scaling]: label_dict[ll] = "{0:.0f}\%%".format(int(lev*100))
           lfmt = label_dict
-        plt.clabel(cset, colors='r', inline=1, fmt=lfmt, fontsize=10)
+        plt.clabel(cset, colors=leveltextcolor, inline=1, fmt=lfmt, fontsize=leveltextfontsize)
         contours_tmp[lev] = cset.collections[0].get_paths()
       if vverbose:
[…]

### 2016-10-17 Paper/paper.tex
improved abstract

@@
 % 
-GWs from NSBH sources carry signatures of (a) the star's tidal distortion by
-the companion black hole (BH) during inspiral, and (b) its potential tidal
-disruption near merger.
+GWs from NSBH sources carry signatures of (a) the tidal distortion of the
+neutron star by its companion black hole during inspiral, and (b) its
+potential tidal disruption near merger.
 %
 In this paper, we present a Bayesian study of the measurability of
-neutron star (NS) tidal deformability $\lambdans\propto (R/M)_\mathrm{NS}^{5}$
-from GW observation(s) of disruptive NSBH coalescences, taking into account the
-crucial effect of BH spins.
-% 
-First, we find that if BHBH templates are used to estimate source 
-parameters for an NSBH signal, a significant bias is introduced in the 
-inference if the signal-to-noise ratio (SNR) is greater than $\simeq30$.
-% 
-Second, for (loud) signals with SNR $\gtrsim 23$ we can bound  $\lambdans$
-by a factor of two.
-% 
-Finally, we study the improvement in our measurement of $\lambdans$ from 
-multiple observations. Using an approximately astrophysical population, 
-we find that (i) the median $\lambdans$ converges within $10\%$ of the true
-value in approximately $20$ observations; (ii) in the same
-number of observations, the associated $90\%$ credible intervals shrink to
-$\pm 50\%$ of the true value.
+neutron star tidal deformability $\lambdans\propto (R/M)_\mathrm{NS}^{5}$
+using observation(s) of inspiral-merger GW signals from disruptive NSBH
+coalescences, taking into account the crucial effect of black hole spins.
+% 
+First, we find that if non-tidal templates are used to estimate source 
+parameters for an NSBH signal, the bias introduced in the estimation of
+non-tidal physical parameters will only be significant for loud signals with
+signal-to-noise ratios greater than $\simeq30$.
+% 
+For similarly loud signals, we also find that we can begin to put
+interesting constraints on $\lambdans$ (factor of $1-2$) with individual
+observations.
+% 
+Next, we study how a population of realistic NSBH detections will improve our
+measurement of neutron star tidal deformability. For an astrophysically likely
+population of {\it disruptive} NSBH coalescences, we find that $20-35$ events
+are sufficient to constrain $\lambdans$ within $\pm 25-50\%$, depending on the
+neutron star equation of state. For these calculations we assume that LIGO will
+detect black holes with masses within the astrophysical {\it mass-gap}. In case
+the mass-gap remains {\it preserved} in NSBHs detected by LIGO, we estimate that
+approximately $25\%$ additional detections will furnish comparable $\lambdans$
+measurement accuracy.
 % 
 We find these results encouraging, and recommend that an effort to 
-measure $\lambdans$ be taken within the LIGO-Virgo Collaboration once
-we begin to detect NSBH coalescences.
+measure $\lambdans$ be planned for upcoming NSBH observations with the
+LIGO-Virgo instruments.
 \end{abstract}

### 2016-10-17 Paper/paper.tex
fix fig 2

@@
 
-% \red{
-% Text focusing on tides
-% }
-
 Another question that compact object mergers can help answer is `what is the 
-nature of matter at nuclear densities', i.e. `what is the EoS of NS material'?
+nature of matter at nuclear densities supported by NSs'?
 A large fraction of past work aimed at measuring NS matter effects from GW
@@
 
-% During early inspiral, the (weak) tidal field of the BH deforms its companion
-% NS.
-% % , thereby exciting its resonant oscillation modes.
-% The resulting induced quadrupolar moment of the star is proportional to the
-% tidal field of its companion,
-% % through the relation $Q_{ij} = -\lambda\mathcal{E}_{ij}$,
-% with $\lambdans$ the EoS-dependent proportionality constant. $\lambdans$ is the
-% NS's tidal deformability parameter, related to its dimensionless Love number
-% $k_2$ and radius $R_\mathrm{NS}$ as $\lambdans:=\frac{2}{3G}k_2 R_\mathrm{NS}^5$.
-% % 
-% Over the course of inspiral, the change in NS's quadrupole moment results in 
-% the emission of GWs that are coherent with the orbital waves.
-% % 
-%
-
-% \red{
-% Insert paragraph describing impact on GWs
-% }
-% 
-% \red{
-% Make clear the distinction between Fischer matrix studies and Bayesian ones. 
-% Explain the differences. this is important as it sets the stage for for what is
-% new in this paper.
-% } 
-% - Describe how waves are affected in each stage for a NSBH. 
-% - Describe what past work has studied and how?
-% - Do we need to compare with BNS.. Maybe in passing, say `` In comparison, for BNS we expect to measure EoSs within 20 detections.
-
 Gravitational waves emitted by coalescing NSBH binaries carry subtle hints of
@@
 $\lambdans$ measurement errors. Fisher matrix estimates are known to be
-unreliable at realistic signal-to-noise ratios (SNR)~\cite{Vallisneri:2007ev},
+unreliable at realistic signal-to-noise ratios (SNR)~\1.35cite{Vallisneri:2007ev},
 such as those as we might expect in the upcoming observing runs of GW
@@
 
-% \red{
-% MOVE to SUMMARY, only keeping 2-4 sentences here. Also, paradigm A/B barely
-% matters. Therefore, first describe the common results, then point out how the
-% presence/absence of mass gap modifies them.
-% }
-% 
-% % \prayush{
-% In addition, we probe two paradigms, one
-% that allows for BH masses to lie within the astrophysical mass-gap (paradigm
-% A), and one that does not (paradigm B).
-% % 
-% For paradigm A, we find the following: (i) for the softer equations of state
-% that result in less deformable neutron stars, $15-20$ detections are sufficient
-% for our median value of $\lambdans$ to lie within $10\%$ of the true
-% value. (ii) For NSBH populations with more deformable NSs
-% ($\lambdans\geq 1000$), the same is achievable within as few as $10$ realistic
-% observations. (iii) The statistical uncertainty associated with $\lambdans$
-% measurement can be restricted to be within $\pm50\%$ using $10-20$ observations
-% when $\lambdans\geq 1000$), and using $25-40$ observations for softer equations
-% of state. All of the above is possible within a few years of design
-% aLIGO operation~\cite{Abadie:2010cfa}, if astrophysical BHs do indeed have
-% masses between $\leq 5M_\odot$ (i.e. in the mass-gap). However, further
-% restricting $\lambdans$ will require $50+$ NSBH observations.
-% % 
-% For paradigm B, we find the information accumulation to be somewhat slower.
-% While the quantitative inferences for $\lambdans\geq1000$ populations are
-% not affected significantly, we find $\mathcal{O}(10-20)\%$ more events are
-% required to attain the same measurement accuracy as under paradigm A.
-% % 
-% We conclude that within as few as $20-30$ observations of disruptive NSBH
-% mergers, aLIGO will begin to place interesting bounds on NS deformability.
-% This, amongst other things, will allow us to rank different equations of 
-% state for neutron star matter from most to least likely, with a few years'
-% time-scale.
-% % }
-
-% \red{
-% The following numerical details disrupt the flow of the text. MOVE after
-% description of analyses.
-% }
-
-% \prayush{
 In this paper, we restrict our parameter space to span mass-ratios
@@
 $\lambdans:= G\left(\frac{c^2}{G \mns}\right)^5\lambda \in[500, 2000]$.
-The mass of NS for signals (and not templates) is fixed to $1.35M_\odot$ 
-throughout~\cite{stellarcollapsemass}. Similarly, their spin is taken to be
-negligible~\cite{Miller:2014aaa}. Signal strength is sampled over a range of
-SNR $\rho\in[10, 70]$.
-% }
-% 
-Further, we note that our results are applicable for the design sensitivity
-LIGO instruments ($2018-19$). 
-We also note that the parameter space probed is somewhat restricted by the
-domain of applicability of the LEA+ model which we use as filters. Finally,
-the accuracy of our quantitative results depends on the reliability of the same
-waveform model, which is the only one available of its kind in current
-literature. A more recent work~\cite{Pannarale:2015jka} improves upon the 
-amplitude description of LEA+, but needs to be augmented with a compatible
-phase model. However, we expect our broad conclusions here to hold despite
-the intrinsic inaccuracies of LEA+ (with errors not exceeding
-$\mathcal{O}(10\%)$~\cite{Pannarale:2015jka}).
-% 
-% we expect the combined effect of modeling errors to not
-% change our broad qualitative conclusions, which are based on statistical trends
-% with large errors themselves.
+These ranges are governed by the calibration of the LEA+ model which we use as
+filters. 
+% 
+Most of the disruptive NSBH simulations that LEA+ has been calibrated to
+involve $1.35M_\odot$ NSs, and it is unclear how reliable the model is 
+for different NS masses~\cite{Lackey:2013axa,Pannarale:2015jka}. This
+motivates us to conservatively fix NS masses to $1.35M_\odot$ in our 
+simulated signals (not templates). But, since the domain of calibration of
+LEA+ excludes NS spin completely, we fix $\chins=0$ in both signals as well as
+filter templates. We expect the effect of ignoring NS mass and spin variations
+in our NSBH populations to be less severe than for
+BNSs~\cite{Agathos:2015a}, considering the higher mass-ratios of NSBHs.
+% 
+The accuracy of our quantitative results depends on the reliability of LEA+,
+which is the only model of its kind in current literature. A more recent 
+work~\cite{Pannarale:2015jka} improves upon the amplitude description of LEA+,
+but needs to be augmented with a compatible phase model. Overall, we expect
+our broad conclusions here to hold despite modeling inaccuracies (with errors
+not exceeding $\mathcal{O}(10\%)$~\cite{Pannarale:2015jka}).
+% 
+Finally, our results apply to LIGO instruments at design sensitivity,
+which they are projected to attain by $2019$~\cite{Shoemaker2009,
+Abbott:2016wya}.
 
@@
 SNR. Across each row, we can see the effect of increasing the signal strength
-(SNR), with the tidal deformability of the NS fixed. We also show dashed
-contours for $\arr_{\mchirp}=10\%$, $50\%$, or $100\%$ (labeled
-without the ``$\%$'' symbol).
+(SNR), with the tidal deformability of the NS fixed. We show dashed contours
+for $\arr_{\mchirp}=10\%, 25\%, 50\%\cdots$, with interleaving color levels
+separated by $5\%$.
 % 
@@
 immediately that for $\rho\leq 30$ the systematics are well under control
-and so we can obtain reliable chirp mass estimates for NSBH signals using BBH
+and we can obtain reliable chirp mass estimates for NSBH signals using BBH
 templates.
@@
 % 
-We can therefore conclude that only for very loud signals, with
-$\rho\gtrsim 50-70$, will the inclusion of tidal terms in our template
-models improve our $\mchirp$ estimation. For lower SNRs, inclusion of new
+We therefore conclude that only for very loud signals, with
+$\rho\gtrsim 50-70$, will the inclusion of tidal terms in template
+models improve $\mchirp$ estimation. For lower SNRs, inclusion of new
 physical content in templates will instead get washed out by detector noise.

### 2016-10-17 emcee_PE/scripts/SingleObservation.py
improved fig 3,4

@@ -89,7 +89,8 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
                        leveltextcolor='b', leveltextfontsize=10,\
         vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f', lfmt='%.1f',\
         xmin=None, xmax=None, ymin=None, ymax=None, alpha=0.9,\
-        colorticks=None, colorbartype='simple', figname='plot.png'):
+        colorticks=None, colorbartype='simple', fix_one_end_of_colorbar_to_zero=True,\
+        figname='plot.png'):
   """
   Function to plot arbitrary numbers of contour plots in a single figure
   """
@@ -154,7 +155,9 @@ col corresponds to the lower level group
   if VMIN > VMAX: raise IOError("Cant fit all data on a single colorbar")
   
   ## FIXME
-  #VMIN = 0.01
+  if fix_one_end_of_colorbar_to_zero:
+    if VMIN < 0: VMAX = 0.0
+    else: VMIN = 0
   
   # Now make contour plots
   contours_all = {}
@@ -182,7 +185,7 @@ col corresponds to the lower level group
       norm = cm.colors.Normalize(vmax=VMAX, vmin=VMIN)#zz.min())
       #cmap = cm.rainbow
       #levels = np.arange(zz.min(), zz.max(), levelspacing)
-      levels = np.arange(VMIN, VMAX, levelspacing)
+      levels = np.arange(VMIN, VMAX+levelspacing, levelspacing)
       CS = ax.contourf( xx, yy, zz,\
               levels=levels, \
               cmap = cm.get_cmap(cmap, len(levels)-1), norm=norm,\

### 2016-10-17 Paper/paper.tex
improved fig 3,4
explanation of why we use injected values for median TT

@@
 (SNR), with the tidal deformability of the NS fixed. We show dashed contours
-for $\arr_{\mchirp}=10\%, 25\%, 50\%\cdots$, with interleaving color levels
-separated by $5\%$.
+for $\arr_{\mchirp}=10\%, 25\%, 50\%\cdots$, with interleaving filled color
+levels separated by $5\%$.
 % 
@@
 \caption{This figure shows the ratio of the systematic and statistical
-measurement errors for BH spins, with other attributes identical to 
-Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}.
+measurement errors for BH spins $\arr_{\chibh}$. Information is arranged identically
+to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}, 
+and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, with the level spacing of filled contours
+increased to $15\%$.
 Similar to the case of mass parameters, we find that below $\rho\approx 30$,
@@
 \end{figure*}
-% %% FIXME FIXME
-% \begin{figure*}
-% \centering
-% \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
-% \textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
-% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
-% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
-% % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
-% % \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
-% % \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% % \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-% % \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-% % \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% % \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
-% % \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
-% % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
-% \caption{%
-% We illustrate here two-dimensional joint probability distributions, as
-% recovered for the intrinsic parameters of a binary with
-% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
-% $\lambdans=2000$, and $\rho=30$ (same as in
-% Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
-% The top row shows the same for combinations of 
-% non-tidal parameters, i.e. binary masses and spins, while the middle row 
-% shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
-% is similar to the middle, with the only difference being that the SNR has been
-% turned up to $\rho=50$.
-% }
-% \label{fig:SingleSystemLambda2DPDFs}
-% \end{figure*}
-% #################
-% \begin{figure*}
-% \centering    
-% \textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
-% \includegraphics[trim={2cm 0 0 0},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
-% % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
-% \caption{Here we show the statistical uncertainty in the measurement of
-% $\lambdans$, as a percentage of the injected/true value. In each panel, the
-% same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
-% injection's SNR $\rho$ fixed (noted in the panel). Rows contain panels
-% with the same value of $\lambdans$, with $\rho$ increasing from left to right.
-% Each column contains panels with the same value of $\rho$, with $\lambdans$ 
-% increasing from top to bottom.
-% % 
-% Contours demarcate regions where we can constrain the
-% $\lambdans$ parameter well (within a factor of two of the injected value).
-% % 
-% We note that, as expected, the measurement accuracy for $\lambdans$ improves with (i) increasing
-% SNR, (ii) decreasing BH mass, (iii) increasing BH spin, and 
-% (iv) increasing $\lambdans$, i.e. the tidal deformability of the neutron star.
-% }
-% \label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
-% \end{figure*}
-% %
-% 
 % 
@@
 
+The effect of ignoring
+tidal corrections in templates will manifest as a systematic shift
+of recovered median parameter values away from what they would be if we had used
+tidal templates with identical priors. 
+In zero noise, we expect the probability distributions recovered using tidal
+templates to be multi-dimensional Gaussians with the maximum likelihood
+parameter values approaching their true values. If the priors are not
+restrictive, we expect the recovered median to also converge to the true value.
+However,
+the LEA model imposes significantly more restrictive priors (both mass-ratio
+and spin) than SEOBNRv2~\cite{Taracchini:2013rva,Lackey:2013axa}, which shifts
+the median value of parameters recovered using {\it our} tidal templates away
+from their true value. If we use LEA+ priors for our non-tidal templates, it
+would add a caveat to our original question 'can we estimate non-tidal NSBH
+parameters with equivalent BBH templates'. Instead, we approximate the median
+tidally recovered parameters by their true injected values, as one would expect
+to recover with an ideal model for tidally disruptive NSBH mergers. With this
+caveat, we estimate {\it systematic} measurement bias/errors as the
+differences between median and {\it injected} parameter values. 
+% 
 As an illustration, in Fig.~\ref{fig:SingleSystemEtaPDFvsSNR} we show the
@@
 $\mbh=5.4M_\odot$, $\chibh=+0.5$ and $\lambdans=2000$). In each panel, both
-the {\it true} and the median values of $\eta$ are marked.
-The effect of ignoring tidal corrections in templates manifests as a systematic
-shift of the median $\eta$ value away from its true value. We measure
-the associated {\it systematic} bias/error by the differences between the
-true and median $\eta$ values. Darker
-shading of the probability distributions marks the $90\%$ credible interval,
-whose width $(\Delta\eta)^{90.0\%}$ is a direct measure of the
+the {\it true} and median values of $\eta$ are marked, and we use
+the shift between the red and green vertical lines as our estimate of systematic
+measurement errors. Darker shading in all panels marks $90\%$ credible intervals,
+whose width $(\Delta\eta)^{90.0\%}$ we use as a direct measure of our
 {\it statistical} measurement uncertainty/error\footnote{We generalize the
@@
 % 
-We see clearly that even when the signal is moderately loud, with $\rho=20$,
+For the illustrated binary,
+we see clearly that even when the signal is moderately loud, with $\rho=20$,
 statistical errors dominate over systematics for $\eta$. As we turn up the SNR
-further, we find that the two error sources become comparable at $\rho\sim 30$,
+further, the two error sources become comparable at $\rho\sim 30$,
 and systematic errors dominate finally when $\rho\simeq 50$.

### 2016-10-17 Paper/paper.tex
typo fixed

@@
 $\lambdans$ measurement errors. Fisher matrix estimates are known to be
-unreliable at realistic signal-to-noise ratios (SNR)~\1.35cite{Vallisneri:2007ev},
+unreliable at realistic signal-to-noise ratios (SNR)~\cite{Vallisneri:2007ev},
 such as those as we might expect in the upcoming observing runs of GW

### 2016-10-17 Paper/paper.tex
added ref to agathos bns work

@@
 in quantitative detail due to waveform model restrictions~\cite{Lackey:2013axa},
-which will restrict our analyses of actual GW signals as well.
+which will also restrict our analyses of GW signals during the upcoming LIGO
+observing runs.
 
@@
 intervals for $90\%$ of all populations in each ensemble~\footnote{Drawn
-under paradigm A.}.
-% 
-% In each panel
-% we show the width of the $90\%$ credible interval measured from $N$ events
-% (with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
-% the population. In gray are shown results from each of $100$ population
-% realizations, with the single population highlighted in color corresponding
-% to the one we focused on in 
-% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
-% The inset shows the same, except that the ordinate is {\it not} normalized by
-% the true value of $\lambdans$. 
+under paradigm A (mass-gap not respected).}.
 % 
@@
 mergers.
-
-% Finally, we quantitatively explore the dependence of our statistical
-% uncertainties for $\lambdans$ on the number of events, as well as on the true
-% NS deformability itself. First, we will focus on the dependence on $N$. We
-% assume a power-law dependence of the form
-% $\delta\lambdans\propto\ 1/N^\alpha$. For each of the $100$ populations 
-% for each of $\lambdans=500-2000$, we compute the exponent $\alpha$ as a
-% function of the number of observed events $N$, and show it in 
-% Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN}. There are $100\times5=500$ curves
-% on the figure, with one highlighted for each value of population's $\lambdans$.
-% These highlighted values are only special in the sense that they correspond to
-% populations discussed earlier in this section (c.f.
-% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}).
-% We immediately observe two things, (i) there is a globally similar dependence
-% on $N$ for all populations, and (ii) information accumulates {\it faster} than
-% $1/\sqrt{N}$. In fact, we find that if
-% $\delta\lambdans\propto\frac{1}{N^\alpha}$, $\alpha$ lines in the range
-% $0.7_{-0.2}^{+0.2}$.
-% % 
-% Next, we focus on the dependence of $\delta\lambdans$ on $\lambdans$ of the
-% population itself. As suggested by Fisher-matrix studies~\cite{Lackey:2013axa},
-% and as for $N$, we assume the form $\delta\lambdans\propto\lambdans^\beta$.
-% From each set of $100$ populations with a given $\lambdans$ value, we draw one
-% at random, and form a set of $5$ similarly drawn populations, one for each of
-% $\lambdans=\{500,800,1000,1500,2000\}$. With each set, we determine $\beta$
-% for different number of observed events $N$. In all, we make $100$ independent
-% $5-$population sets and show the value of $\beta$ measured from each in 
-% Fig.~\ref{fig:TT_PowerLawLambdaErrorVsLambda}. We find that the assumed
-% relation $\delta\lambdans\propto\lambdans^\beta$ gets fairly robust for 
-% larger values of $N$, with $\beta$ converging to $\beta=0.5^{+0.33}_{-0.33}$.
-% The fact that $0<\beta<1$ implies that the relative error
-% $\delta\lambdans/\lambdans$ {\it decreases} with increasing $\lambdans$, while
-% the absolute error {\it increases}.
-% % 
-% From these results, we conclude that the measurement uncertainty for
-% $\lambdans$ after $N$ observations is
-% \begin{equation}
-%  \delta\lambdans\propto \dfrac{\lambdans^{0.5^{+0.33}_{-0.33}}}{N^{0.7_{-0.2}^{+0.2}}}.
-% \end{equation}
-% We also find that while these results are inferred from paradigm A populations,
-% paradigm B gives very similar results.
+% 
+These results show that our probed subset of disruptive NSBH mergers is as
+informative of NS tidal properties as are BNS populations, if we assume a uniform mass
+distribution for NSs and zero NS spins, and possibly better if NS masses are
+not distributed uniformly in astrophysical binaries~\cite{Agathos:2015a}.
 

### 2016-10-17 Paper/paper.tex
touches to conclusions

@@
 % 
-These results show that our probed subset of disruptive NSBH mergers is as
+These results demonstrate that our probed set of disruptive NSBH mergers is as
 informative of NS tidal properties as are BNS populations, if we assume a uniform mass
-distribution for NSs and zero NS spins, and possibly better if NS masses are
-not distributed uniformly in astrophysical binaries~\cite{Agathos:2015a}.
+distribution for NSs and zero NS spins, and possibly more if NS masses are
+not distributed uniformly~\cite{Agathos:2015a}.
+However, being a subset of all NSBH binaries, the accumulation of information
+from NSBH signals in general may be slower than from binary neutron stars,
+depending on the distribution of BH masses in coalescing NSBH binaries.
 
@@
 $10-20\%$ more events to attain the same measurement accuracy as under
-paradigm A. In either case, the accumulation of information from NSBH signals
-is likely slower than from BNS inspirals~\cite{Mandel:2009pc,Lackey2014,
-Wade:2014vqa}, although template models for the latter may be more
+paradigm A. In either case, the accumulation of information from the general
+NSBH population will likely be slower than from BNS inspirals~\cite{Mandel:2009pc,
+Lackey2014,Wade:2014vqa,Agathos:2015a}, depending on the mass distribution of 
+stellar-mass black holes. Though, template models for the latter may be more
 uncertain due to missing point-particle PN terms at orders comparable to
-the tidal terms.
+tidal terms~\cite{Lackey2014}.
 % 

### 2016-10-17 Paper/paper.tex
compare with Fisher matrix results
moved figure titles to captions

@@
 \centering 
-\textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
-\textbf{at different SNR values (shown in legend).}\par\medskip
+% \textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
+% \textbf{at different SNR values (shown in legend).}\par\medskip
 \includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}\\
-\caption{We show here probability distributions for mass ratio $\eta$ as measured
+\caption{{\bf Illustrative posterior probability distributions for mass-ratio $\eta$ at different SNR values:}
+We show here probability distributions for mass ratio $\eta$ as measured
 for the same signal at different SNRs. The intrinsic parameters of the source
@@
 \centering 
-\textbf{Ratio of systematic to statistical errors in measuring $\mchirp$;}\par
-\textbf{ignoring tidal effects.}\par\medskip
+% \textbf{Ratio of systematic to statistical errors in measuring $\mchirp$;}\par
+% \textbf{ignoring tidal effects.}\par\medskip
 \includegraphics[width=1.95\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
@@
 % \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
-\caption{We show here the ratio of systematic and statistical
+\caption{{\bf Ratio of systematic to statistical errors in measuring $\mchirp$, ignoring tidal effects:}
+We show here the ratio of systematic and statistical
 measurement uncertainties for the binary chirp mass over the NSBH parameter 
@@
 \centering
-\textbf{Ratio of systematic to statistical errors in measuring $\eta$;}\par
-\textbf{ignoring tidal effects.}\par\medskip
+% \textbf{Ratio of systematic to statistical errors in measuring $\eta$;}\par
+% \textbf{ignoring tidal effects.}\par\medskip
 \includegraphics[width=1.95\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_50_linear}
@@
 % \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
-\caption{This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
+\caption{{\bf Ratio of systematic to statistical errors in measuring $\eta$, ignoring tidal effects:}
+This figure is similar to Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
 with the difference that here we show the ratio of systematic and statistical
@@
 \centering
-\textbf{Ratio of systematic to statistical errors in measuring $\chibh$;}\par
-\textbf{ignoring tidal effects.}\par\medskip
+% \textbf{Ratio of systematic to statistical errors in measuring $\chibh$;}\par
+% \textbf{ignoring tidal effects.}\par\medskip
 \includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
@@
 % \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
-\caption{This figure shows the ratio of the systematic and statistical
+\caption{{\bf Ratio of systematic to statistical errors in measuring $\chibh$, ignoring tidal effects:}
+This figure shows the ratio of the systematic and statistical
 measurement errors for BH spins $\arr_{\chibh}$. Information is arranged identically
@@
 \centering 
-\textbf{Illustrative posterior probability distributions for NS tidal
-deformability $\lambdans$,}\par
-\textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
-from left to right.}\par\medskip
+% \textbf{Illustrative posterior probability distributions for NS tidal
+% deformability $\lambdans$,}\par
+% \textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
+% from left to right.}\par\medskip
 % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
 \includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
-\caption{We show here probability distributions recovered for the NS tidal
+\caption{{\bf Illustrative posterior probability distributions for NS tidal
+deformability $\lambdans$:}
+We show here probability distributions recovered for the NS tidal
 deformability parameter $\lambdans$ from three GW injections, with parameters:
@@
 \centering    
-\textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
+% \textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
 \includegraphics[trim={3cm 0 0 0.5cm},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
 % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
-\caption{Here we show the statistical uncertainty in the measurement of
+\caption{{\bf Statistical uncertainty in $\lambdans$ measurement:}
+Here we show the statistical uncertainty in the measurement of
 $\lambdans$, as a percentage of the injected/true value. In each panel, the
@@
 % 
-The results presented so far are in agreement with Sec.~\ref{s1:PEwithnoNS}.
+These results agree well with Sec.~\ref{s1:PEwithnoNS}, and are consistent with
+Fisher matrix estimates at high SNRs~\cite{Lackey:2013axa}.
 
@@
 \centering    
-\textbf{Improvement in $\lambdans$ measurement accuracy\\ with multiple observations}\par\medskip
+% \textbf{Improvement in $\lambdans$ measurement accuracy\\ with multiple observations}\par\medskip
 \includegraphics[width=0.8\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}\\
 \includegraphics[width=0.8\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
-\caption{%
+\caption{{\bf Improvement in $\lambdans$ measurement accuracy with multiple observations:}
 These panels show the accumulation of information from a population whose
@@
 \includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
-\caption{%
+\caption{{\bf Improvement in $\lambdans$ measurement accuracy for different NS EoS:}
 In this figure, the filled regions show how our measurement of $\lambdans$
@@
 \centering
-\textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
+% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
 \includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
@@
 \includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{plots/LambdaMedian90pc_vs_N_AstroPopulation}\\
-\caption{%
-{\it Top left}: This figure shows the median value of the recovered
-probability distribution for $\lambdans$, as a function of the number of events
-in the population $N$. There are four ensembles of curves,
+\caption{
+{\bf No Mass-Gap}, {\it top left}: The top figure shows the median value of the recovered
+probability distribution for $\lambdans$, as a function of the number of observed 
+events $N$. There are four ensembles of curves,
 corresponding to $\lambdans=\{500,1000,1500,2000\}$, with a hundred
@@
 horizontal dash-dotted lines.
-{\it Bottom left}: Here we show the interval of $\lambdans$ values within
+{\bf No Mass-Gap}, {\it bottom left}: Here we show the interval of $\lambdans$ values within
 which the median $\lambdans$ lies for $90\%$ of the populations in
@@
 % 
-{\it Right:} These panels are identical to their counterparts on the left,
+{\bf Mass-Gap}, {\it right column}: These panels are identical to their counterparts on the left,
 with the only difference that the BH masses in each population are restricted
@@
 \centering    
-\textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
+% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
 \includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{plots/LambdaCIWidths90pc_vs_N_AllPopulation}
 \includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation}
-\caption{%
-{\it Left}: This panel shows the width of $\lambdans$ interval within
+\caption{{\bf No Mass-Gap} ({\it left}): This panel shows the width of $\lambdans$ interval within
 which the $90\%$ credible intervals for $\lambdans$ lie, for $90\%$ of 
@@
 % 
-{\it Right}: This panel shows populations drawn under paradigm B, which
+{\bf Mass-Gap} ({\it right}): This panel shows populations drawn under paradigm B, which
 respects the mass-gap.
@@
 \begin{acknowledgments}
-We thank Ben Lackey, Francesco Pannarale, Francois Foucart, and Duncan Brown
+  We thank Benjamin Lackey, Francesco Pannarale, Francois Foucart, and Duncan Brown
   for helpful discussions. We gratefully acknowledge support
@@
 \centering 
-\textbf{Statistical measurement uncertainty for NSBH parameters $\mchirp$ (top),
-$\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
 \includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
@@
 \includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
-\caption{
+\caption{{\bf Statistical measurement uncertainty for NSBH parameters,
+ignoring tidal effects:}
 We show here the statistical uncertainty associated with our measurement of
@@
 \centering
-\textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
-\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
 \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
@@
 % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
-\caption{%
-% We illustrate here two-dimensional joint probability distributions, as
-% recovered for the intrinsic parameters of a binary with
-% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
-% $\lambdans=2000$, and $\rho=30$ (same as in
-% Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
-% The top row shows the same for combinations of 
-% non-tidal parameters, i.e. binary masses and spins, while the middle row 
-% shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
-% is similar to the middle, with the only difference being that the SNR has been
-% turned up to $\rho=50$.
+\caption{{\bf Illustrative posterior probability distributions for NSBH parameters,
+for signals at different SNRs:}
 We illustrate here two sets of two-dimensional joint probability distributions,

### 2016-10-18 Paper/paper.tex
incorporated all of harald's comments but for fig10

@@
 \centering 
+% \fbox{
 % \textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
 % \textbf{at different SNR values (shown in legend).}\par\medskip
-\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}\\
+\includegraphics[trim=20 18 18 18 0,clip=true,width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}\\
 \caption{{\bf Illustrative posterior probability distributions for mass-ratio $\eta$ at different SNR values:}
@@
 }
+% }
 \label{fig:SingleSystemEtaPDFvsSNR}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{figure*}
-% \centering 
-% \textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
-% \textbf{at different SNR values (shown in legend).}\par\medskip
-% \includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
-% \caption{We show here probability distributions for mass ratio $\eta$ as measured
-% for the same injection at different SNRs. The intrinsic parameters of the source
-% are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
-% and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
-% {\it ignore} tidal effects.
-% % 
-% In each panel: the dashed red line marks the median value
-% $\eta^\mathrm{Median}$, while the dashed green line show the true value
-% $\eta^\mathrm{Injected}$. The difference between the true and median values
-% is the systematic bias introduced
-% by ignoring tidal effects in templates. The darker shading shows
-% the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
-% which is a direct measure of our statistical uncertainty on $\eta$.
-% % 
-% Comparing systematic and statistical errors, we find that:
-% at $\rho=20$, $\eta$ measurement is dominated by statistical
-% errors; at $\rho=30$, the two become comparable; and 
-% for louder signals ($\rho\simeq50$), the systematic errors dominate.
-% }
-% \label{fig:SingleSystemEtaPDFvsSNR}
-% \end{figure*}
-% ################# FIXME FIXME
-% \begin{figure*}
-% \centering 
-% \textbf{Statistical measurement uncertainty for NSBH parameters $\mchirp$ (top),
-% $\eta$ (middle) and $\chibh$ (bottom);\newline ignoring tidal effects.}\par\medskip
-% \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
-% \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
-% \includegraphics[trim = {2cm 0 0 0},width=2.1\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
-% \caption{
-% We show here the statistical uncertainty of our measurement of the 
-% dimensionless mass ratio $\eta$ (at $90\%$ credibility), over the NSBH
-% parameter space.
-% Individual panels show the same as a function of BH mass and spin.
-% Across each row, we see the effect of increasing signal strength (i.e.
-% SNR) with the tidal deformability of the NS $\lambdans$ fixed. Down each
-% column, we see the effect of increasing $\lambdans$, at fixed SNR.
-% }
-% \label{fig:CIWidths90_Lambda_SNR}
-% \end{figure*}
-% #################
 % 
@@
 \centering 
-% \textbf{Ratio of systematic to statistical errors in measuring $\mchirp$;}\par
-% \textbf{ignoring tidal effects.}\par\medskip
-\includegraphics[width=1.95\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
-% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
-% % \includegraphics[width=0.67\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
-% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
-% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
-% \includegraphics[width=0.8\columnwidth]{plots/TNMchirpBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
+\includegraphics[trim=20 20 18 22 0,clip=true,width=1.95\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
 \caption{{\bf Ratio of systematic to statistical errors in measuring $\mchirp$, ignoring tidal effects:}
@@
 \centering
-% \textbf{Ratio of systematic to statistical errors in measuring $\eta$;}\par
-% \textbf{ignoring tidal effects.}\par\medskip
-\includegraphics[width=1.95\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_50_linear}
-% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
-% % \includegraphics[width=0.67\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
-% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
-% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
-% \includegraphics[width=0.8\columnwidth]{plots/TNEtaBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
+\includegraphics[trim=20 20 18 21 0,clip=true,width=1.95\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_50_linear}
 \caption{{\bf Ratio of systematic to statistical errors in measuring $\eta$, ignoring tidal effects:}
@@
 \centering
-% \textbf{Ratio of systematic to statistical errors in measuring $\chibh$;}\par
-% \textbf{ignoring tidal effects.}\par\medskip
-\includegraphics[width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
-% % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI68_3.pdf}
-% % \includegraphics[width=\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_All_CI90_0.pdf}
-% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda500_CI90_0}
-% % \includegraphics[width=0.67\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda800_CI90_0}
-% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda1000_CI90_0}\\
-% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda1500_CI90_0}
-% \includegraphics[width=0.8\columnwidth]{plots/TNChiBHBiasesOverCIWidthsVsSNR_Lambda2000_CI90_0}
+\includegraphics[trim=20 20 18 20 0,clip=true,width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
 \caption{{\bf Ratio of systematic to statistical errors in measuring $\chibh$, ignoring tidal effects:}
@@
 \centering 
-% \textbf{Illustrative posterior probability distributions for NS tidal
-% deformability $\lambdans$,}\par
-% \textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
-% from left to right.}\par\medskip
-% \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
-\includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
+\includegraphics[trim=10 20 8 18 0,clip=true,width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
 \caption{{\bf Illustrative posterior probability distributions for NS tidal
@@
 decreases monotonically as $\lambdans$ increases (with signal strength fixed).
-% 
-% Comparing systematic and statistical errors, we find that:
-% at $\rho=20$, $\lambdans$ measurement is dominated by statistical
-% errors; at $\rho=30$, the two become comparable; and 
-% for louder signals ($\rho\simeq50$), the systematic errors dominate.
 }
@@
 % % 
-% \begin{figure*}
-% \centering 
-% \textbf{Illustrative posterior probability distributions for NS tidal
-% deformability $\lambdans$,}\par
-% \textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
-% from left to right.}\par\medskip
-% % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
-% \includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
-% \caption{We illustrate here probability distributions recovered for the NS tidal
-% deformability parameter $\lambdans$ from three GW injections, with parameters:
-% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
-% $\lambdans=\{1000,1500,2000\}$ from left to right. The injection SNR is fixed at
-% $\rho=50$. The templates {\it include} tidal effects, with a prior $0\leq\lambdans\leq 4000$.
-% % These figures show that this approximation holds up to (SNR)
-% % $\rho\simeq 30-50$.
-% % 
-% In each panel- the dashed red line marks the median value for
-% $\lambdans$, and the dashed green line marks its {\it true} value.
-% The darker shading shows the $90\%$ credible interval, whose width
-% $(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
-% By comparing the measurement uncertainty for these three injections, we see
-% that $(\Delta\lambdans)^{90\%}$ grows very slowly with $\lambdans$. Therefore,
-% the fractional measurement error - $(\Delta\lambdans)^{90\%}/\lambdans$ -
-% decreases monotonically as $\lambdans$ increases (with signal strength fixed).
-% % 
-% % Comparing systematic and statistical errors, we find that:
-% % at $\rho=20$, $\lambdans$ measurement is dominated by statistical
-% % errors; at $\rho=30$, the two become comparable; and 
-% % for louder signals ($\rho\simeq50$), the systematic errors dominate.
-% }
-% \label{fig:SingleSystemLambdaPDFvsSNR}
-% \end{figure*}
-% % %% FIXME FIXME
-% % \begin{figure*}
-% % \centering
-% % \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
-% % \textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
-% % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
-% % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
-% % % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
-% % % \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
-% % % \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
-% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
-% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
-% % \caption{%
-% % We illustrate here two-dimensional joint probability distributions, as
-% % recovered for the intrinsic parameters of a binary with
-% % $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
-% % $\lambdans=2000$, and $\rho=30$ (same as in
-% % Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
-% % The top row shows the same for combinations of 
-% % non-tidal parameters, i.e. binary masses and spins, while the middle row 
-% % shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
-% % is similar to the middle, with the only difference being that the SNR has been
-% % turned up to $\rho=50$.
-% % }
-% % \label{fig:SingleSystemLambda2DPDFs}
-% % \end{figure*}
 % #################
@@
 \centering    
-% \textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
-\includegraphics[trim={3cm 0 0 0.5cm},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
+\includegraphics[trim=35 21 15 21 0,clip=true,width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
 % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
@@
 Here we show the statistical uncertainty in the measurement of
-$\lambdans$, as a percentage of the injected/true value. In each panel, the
+$\lambdans$. In each panel, the
 same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
@@
 \centering    
-% \includegraphics[trim={0 0 0 2cm}, width=0.95\columnwidth]{plots/TTSNRThresholdFor200LambdaMeasurement_BHspin_BHmass_Lambda1500_0_CI90_0}
-\includegraphics[width=\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}
+\includegraphics[trim=10 10 0 10 0,clip=true,width=1.05\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}
 % \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
@@
 year of observation~\cite{Abadie:2010cfa}.
-% 
-% We note that this information is, in principle, contained in 
-% Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
-% the effect of NS's deformability itself on its measurement. 
-% 
-% As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, we note that the measurability
-% of NS matter effects improves with all factors that enhances the signature of the NS's disruption on
-% the \textit{detectable} portion of the emitted GW signal (implying, within a frequency band
-% set by the detectorTT_SNRThresholds_BHspin_BHmass_CI90_0s).
 }
@@
 \centering    
-% \textbf{Improvement in $\lambdans$ measurement accuracy\\ with multiple observations}\par\medskip
-\includegraphics[width=0.8\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}\\
-\includegraphics[width=0.8\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
-\caption{{\bf Improvement in $\lambdans$ measurement accuracy with multiple observations:}
-These panels show the accumulation of information from a population whose
-$\lambdans=800$.
-{\it Left:} Posterior probability distributions for $\lambdans$ (colored 
-curves), and associated $90\%$ credible intervals (grey vertical lines),
-shown for different number of accumulated observations $N$. Distributions are
-normalized to unit area.
-{\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
-associated $90\%$ credible intervals (as the vertical extent of filled
-region), shown as a function of number of observations $N$. Solid horizontal
-line indicates the true value of $\lambdans$. Dashed and dotted horizontal
-lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
-bounds.
+\includegraphics[trim=18 18 18 10 0,clip=true,width=\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}\\
+\includegraphics[trim=18 18 18 10 0,clip=true,width=\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+\caption{{\bf Recovery of $\lambdans$ for an increasingly large population of BH-NS signals.}
+{\it Top}: Posterior probability distributions for $\lambdans$ (colored curves), and
+associated $90\%$ credible intervals (grey vertical lines), shown for different number
+of accumulated observations N. Distributions are normalized to unit area.  
+{\it Bottom}: Measured median value of $\lambdans$ (as solid circles) and the
+associated $90\%$ credible intervals (as the vertical extent of filled region), shown as
+a function of number of observations N. Solid horizontal line indicates the true value of
+$\lambdans=800$. Dashed and dotted horizontal lines (a pair for each line-style) demarcate
+$\pm 25\%$ and $\pm 50\%$ error bounds.
 }
@@
 \end{figure}
-%
-% % 
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L2000.pdf}
-% \includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1500.pdf}\\
-% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1000.pdf}
-% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
-% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
-% \caption{Filled-region plots showing the median and $90\%$ credible intervals
-% for $\lambdans$ measurement, as a function of the number of observed events $N$.
-% Each panel corresponds a unique population with injected $\lambdans$ fixed to
-% the value shown in its title. In each panel, the recovered median for $\lambdans$
-% is shown by filled circles, and the $90\%$ credible interval is shown by the 
-% height of the filled region. In addition, the dashed and 
-% dotted lines (2 of each line-style) show the $\pm 25\%$ and $\pm 50\%$ error
-% bounds, with respect to the true value.
-% }
-% \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}
-% \end{figure*}
 %
@@
 % 
-% \begin{figure*}
-% \centering
-% \includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
-% \includegraphics[trim=20 0 0 0, width=1.02\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}
-% \caption{%
-% {\it Left:} This figure shows the median value of the recovered
-% probability distribution for $\lambdans$, as a function of the number of events
-% in the population $N$. There are four family of curves, one corresponding each
-% to $\lambdans=\{500,1000,1500,2000\}$, with $100$ independent populations
-% within each family. One curve in each family is highlighted in color, 
-% representing that it is the same population as was illustrated in
-% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
-% In the same color we show $\pm 10\%$ error-bounds on $\lambdans$ with
-% horizontal dash-dotted lines. We observe that within $10-25$ observations, 
-% the median of the measured cumulative probability distribution for $\lambdans$
-% will lie within $10\%$ of the true value.
-% {\it Right:} This figure is identical to the left panel, with the only
-% difference that the BH masses in each population are restricted to lie
-% {\it outside} the astrophysical mass-gap (i.e. paradigm B). The difference that
-% we observe under this paradigm is that we need more ($30+$) events to achieve 
-% the same ($10\%$) measurement accuracy for populations with $\lambdans<1000$.
-% For more deformable neutron stars, $10-25$ events would suffice.
-% }
-% \label{fig:TT_LambdaMedian_vs_N_AllInOne}
-% \end{figure*}
-% %
-% % 
-% \begin{figure*}
-% \centering    
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L2000.pdf}
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L2000.pdf}\\
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1500.pdf}
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1500.pdf}\\
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L1000.pdf}
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L1000.pdf}\\
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L800.pdf}
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L800.pdf}\\
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AllPopulations_Log_L500.pdf}
-% \includegraphics[width=0.8\columnwidth]{plots/LambdaCIWidths_vs_N_AstroPopulations_Log_L500.pdf}
-% \caption{%
-% These figures show the width of the $90\%$ credible interval for $\lambdans$
-% (normalized by its true value), as a function of the number of observed events
-% $N$. The left column panels show populations sampled under paradigm A, which
-% allows BH masses to fall within the astrophysical mass-gap, while those on the
-% right show populations drawn under paradigm B which respects the mass-gap.
-% Each panel corresponds to a unique value of populations' $\lambdans$,
-% decreasing from $2000\rightarrow 500$ as we go from top to bottom. In each panel,
-% $100$ curves are shown, each corresponding to an independent population draw. The
-% one colored curve in each panel highlights the population used in
-% Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
-% The insets in each panel show the same information, with the ordinate {\it not}
-% normalized by the true value of $\lambdans$.
-% % 
-% We find that with approximately $25$ or so events, we begin to put
-% statistically meaningful constraints on $\lambdans$, restricting it to within
-% $\pm 50\%$ of the true value. We can expect to achieve this with a few years
-% of design aLIGO operation~\cite{Abadie:2010cfa}. Further tightening of the
-% credible intervals will require $40+$ events.
-% }
-% \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}
-% \end{figure*}
-% % 
 \begin{figure*}
@@
   Researcher Awards Program, the Canada Research
-  Chairs Program, and the Canadian Institute for Advanced Research; at
-  Caltech from the Sherman Fairchild Foundation and NSF grants
-  PHY-1404569 and AST-1333520; at Cornell from the
-  Sherman Fairchild Foundation and NSF grants PHY-1306125 and
-  AST-1333129; and at Princeton from NSF grant PHY-1305682 and the
-  Simons Foundation.  Calculations were performed at the Vulcan
+  Chairs Program, and the Canadian Institute for Advanced Research.
+  Calculations were performed at the Vulcan
   supercomputer at the Albert Einstein Institute;
@@
 % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr30}%\\
-\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
+\includegraphics[width=1.05\columnwidth,trim=0.7cm 0 1cm 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
 % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
@@
 $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
-$\lambdans=2000$. Contours are shown for $\{1-,2-,3-\}\sigma$ confidence levels.
+$\lambdans=2000$. Contours are shown for $\{1-,2-,3-,\cdots\}\sigma$ confidence levels.
 Templates include tidal effects, as evident in the bottom rows

### 2016-10-18 emcee_PE/scripts/SingleObservation.py

@@ -90,6 +90,7 @@ def make_contour_array(X, Y, Z2d, xlabel='Time (s)', ylabel='', clabel='', \
         vmin=None, vmax=None, cmap=cm.rainbow, cfmt='%.1f', lfmt='%.1f',\
         xmin=None, xmax=None, ymin=None, ymax=None, alpha=0.9,\
         colorticks=None, colorbartype='simple', fix_one_end_of_colorbar_to_zero=True,\
+        gridded=True, grid_color='lightgrey',\
         figname='plot.png'):
   """
   Function to plot arbitrary numbers of contour plots in a single figure
@@ -190,6 +191,7 @@ col corresponds to the lower level group
               levels=levels, \
               cmap = cm.get_cmap(cmap, len(levels)-1), norm=norm,\
               alpha=alpha, vmin=VMIN, vmax=VMAX)
+      ax.tick_params('both', length=6, width=1, which='major')
       contours_tmp = {}
       for lev in levellines:
         cset = ax.contour(xx, yy, zz, levels=[lev * levline_scaling],\
@@ -205,7 +207,8 @@ col corresponds to the lower level group
         print "for %s" % titles[idx][jdx], contours_tmp
         print "ymin, ymax = ", ymin, ymax
       #
-      ax.grid(True, color='k')
+      if gridded: ax.grid(True, color=grid_color)
+      else: ax.grid(False)
       if xticks is not None: ax.get_xaxis().set_ticks(xticks)
       if yticks is not None: ax.get_yaxis().set_ticks(yticks)
       ax.set_xlim([xmin, xmax])

### 2016-10-19 Paper/journal_submission/paper.tex
files going on arxiv

--- 2016-10-18 Paper/paper.tex
@@
 
-% \prayush{
-% The actual masses of
-% astrophysical black holes are uncertain, but observations and
-% population synthesis studies suggest that BHs formed from stellar
-% core-collapse can have masses up to and higher than
-% $34M_\odot$~\cite{2008ApJ...678L..17S,2010ApJ...714.1217B}.  Also,
-% recent measurements using continuum fitting and X-ray reflection
-% fitting suggest that black holes can have high spin, with the BH
-% angular momentum in dimensionless units exceeding
-% $0.8$~\cite{2011ApJ...742...85G,2012MNRAS.424..217F,Gou:2013dna,
-%   2009ApJ...697..900M,McClintock:2006xd,Miller:2009cw,McClintock:2013vwa}.
-% }
-% %
-
-% \red{
-% Text describing BHNS disruption and accretion dynamics.
-% }
-
-% Most of the matter falls into the BH in ~1ms. But the disk itself has an
-% accretion timescale of ~100ms-1s. The disk-BH system is the expected engine
-% for a SGRB, through the production of a relativistic jet. The unbound material,
-% on the other hand, is responsible for an EM signal in the infrared/optical, 
-% powered by radioactive decays in the ejecta. And possibly a radio afterglow
-% through interactions with the ISM. But not gamma-rays (unless you are referring
-% to the very small amount of material ejected in the jet, but that's not what
-% people usually mean when talking about material unbound by the merger).
-%
-
 NSBH binaries are of interest for multiple reasons. For instance,
@@
 % BBH)~\cite{Flanagan2008}.
-% 
 % We know the leading and next-to-leading order terms in post-Newtonian (PN)
@@
 
-% A large fraction of past work aimed at measuring NS matter effects from GW
-% signals has consisted of inquiries about binary neutron stars (BNS)~\cite{
-% Lee1999a,Lee1999b,Lee2000,oechslin:07,Read:2008iy,Markakis:2010mp,Markakis:2011vd,
-% stergioulas:11,East:2011xa,Lackey2014,Wade:2014vqa,Bauswein:2014qla}.
-% % 
-% % It has shown~\cite{DelPozzo:13,Chatziioannou:2015uea,Agathos:2015a} that
-% % with $20-30$ BNS observations, aLIGO would begin to distinguish between hard,
-% % moderate and soft candidate equations of state.
-% % 
-% % These studies rely on the PN 
-% % description of BNS dynamics, which is reasonable since BNSs merge at very high
-% % frequencies and most of their signal power resides in inspiral.
-% % 
-% During early inspiral the tidally induced waves are weak in strength,
-% as appropriate for an effect that enters binary phasing at $5^{th}$
-% Post-Newtonian (PN) order~\cite{Vines2011}. During late-inspiral and merger,
-% a disruptive fate of the NS could result in significant suppression of the
-% emitted GWs~\cite{}.
-% % 
-% % The difficulty in discerning NS matter effects in an NSBH coalescence from its
-% % GW signal lies in distinguishing it reliably from a BBH signal, since the 
-% % difference between the two is either subtle (during inspiral) or short-lived
-% % (disruption and merger).
-% % 
-% The earlier in inspiral the NS disrupts, the more orbits there are over which
-% its effects are manifest in the GW signal.
-% % 
-% In addition, NSBHs merge at relatively lower frequencies than BNS, and their
-% disruption during late-inspiral occurs at frequencies where aLIGO has
-% relatively more resolving power. This, combined with a shorter inspiral,
-% means that the merger portion of NSBH signals is relatively more valuable for
-% measuring $\lambdans$.
-
 In this paper we study the measurability of neutron star's tidal deformability
@@
 % 
-% \red{
-% MOVE the following to SUMMARY. Here, keep 1-2 sentences:
 We find that, for individual events, ignoring tidal effects will affect mass
@@
 effects without loss of sensitivity.
-% }
-
-% \red{
-% Somewhat reduce discussion of results. Discuss in order of probability, starting
-% with low SNR, ending with exceptional events.
-% }
-
-% \prayush{
+%
+
 Second, we study the ability of aLIGO to constrain neutron star tidal 
@@
 (within $\pm 50\%$ of its true value).
-% For more realistic SNRs and BH parameters, a single measurement is 
-% for $\lambdans$ is accompanied with a $\sim\pm150+\%$. uncertainty.
-% }
 While this is discouraging at first, we turn to ask: what if we combine
@@
 
-% \red{
-% BEgin this paragraph by pointing out that EOS is universal among all NSs. 
-% Therefore, one universal $\lambdans$ vs $\mns$ curve. Therefore, by combining
-% multiple BHNS events, expect to measure $\lambdans$ better than with any 
-% single BHNS event.
-% }
-% % \prayush{
 The EoS of matter at nuclear densities is believed to be universal among all
@@
 study~\cite{DelPozzo:13} aimed at binary neutron stars.
-% % }
-% % 
-% \red{
-% MOVE into later sections. Keep here only 1-2 informational sentences.
-% }
 % % 
@@
 \centering 
-% \fbox{
-% \textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
-% \textbf{at different SNR values (shown in legend).}\par\medskip
-\includegraphics[trim=20 18 18 18 0,clip=true,width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}\\
+\includegraphics[trim=20 18 18 18 0,clip=true,width=1.8\columnwidth]{SingleSystemEta_q4_0_mc2_25_chi0_50}\\
 \caption{{\bf Illustrative posterior probability distributions for mass-ratio $\eta$ at different SNR values:}
@@
 }
-% }
 \label{fig:SingleSystemEtaPDFvsSNR}
@@
 \subsection{Waveform Models}\label{s2:waveforms}
-
-% \red{
-% Begin with introducing sentence:
-% }
 
@@
 % 
-% Amongst the varied quantities were (a) the
-% equation of state for the NS, (b) NS mass, (c) mass-ratio and (d) black hole
-% spins, taken as aligned with the orbit.
-% A total of $21$ two-parameter EoSs were used for these simulations, which also 
-% had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
-% $m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
-% authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
-% NSBH coalescences.
 % 
@@
 Library~\cite{LAL}.
-% ~\footnote{The 
-% $l=|m|=2$ multipoles of the GW strain, when decomposed in a spin $-2$ weighted
-% spherical harmonic basis, contain more than $99\%$ of the total signal power}
-% This approximation has the added advantage of making various angle parameters
-% degenerate with source distance.
 
@@
 \subsection{Bayesian methods}\label{s2:bayesian}
-% % Describe the emcee-based PE code
-% \red{
-% Section is immensely technical. Perform careful polishing pass to 
-% improve flow and to make it more accessible.
-% }
 
@@
 $\mathcal{N}$ is an SNR-dependent normalization constant.
-% \red{[MP: INSERT A LINE ABOUT APPROXIMATING $\LL \approx e^{\rho^2/2}$}
 As in Ref.~\cite{Purrer:2015nkh} we use a likelihood that is maximized over
@@
 % 
-% We perform Bayesian parameter estimation for tidal signals, using
-% both tidal and non-tidal templates. The signal parameters are selected 
-% to lie on a grid given by:
-% $q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
-% Corresponding signal waveforms are generated using the model described in the 
-% previous sub-section, and each injected into a separate zero noise data segment
-% $d_n$. Templates can be either tidal or non-tidal (LEA+ with $\lambdans=0$).
-% 
-% Inferred from $d_n$, the joint probability distribution for binary parameters
-% $\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ is given by
-% \begin{equation}\label{eq:postprob}
-%  p(\vec{\Theta} | d_n, H) = \dfrac{p(d_n|\vec{\Theta}, H)\,p(\vec{\Theta} | H)}{p(d_n|H)}.
-% \end{equation}
-% Here, $p(\vec{\Theta} | H)$ is the prior probability for binary parameters of
-% taking particular values. In this paper, we use a uniform prior on the
-% individual masses within pre-chosen ranges, as well as a uniform prior on the
-% spin of the black hole. We set neutron star's spin $=0$, and its tidal
-% deformability parameter $\lambdans$ is sampled uniformly from $[0, 4000]$. In
-% addition, we exclude mass-ratios between $[1,2]$ from our prior. This is done
-% because of model constraints for LEA+. The prior probability of obtaining a 
-% particular realization of data $p(d_n|H)$ is absorbed into the overall 
-% normalization. Finally, the first term in the numerator
-% $p(d_n|\vec{\theta}, \lambdans, H)$ is the likelihood of obtaining the given
-% stretch of data $d_n$ if we assume that a signal parameterized by
-% $\vec{\theta}\cup\{\lambdans\}$ is buried in it, i.e.
-% \begin{equation}\label{eq:likelihood}
-%  \LL(\vec{\Theta}) \equiv p(d_n| \vec{\Theta}, H) = \mathcal{N} \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
-% \end{equation}
-% where $h\equiv h(\vec{\Theta})$ is a template given by our waveform model LEA+,
-% % and the detector noise weighted inner-product $\langle a|b\rangle$ is defined 
-% % as
-% and $\langle\cdot|\cdot\rangle$ is the detector noise weighted 
-% inner-product~\footnote{%
-% This inner product is defined as
-% \begin{equation}
-% \langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
-% \end{equation}
-% % 
-% where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
-% and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}
-% The 
-% definition of $\LL$ in Eq.~\ref{eq:likelihood} assumes that the instrument
-% noise is colored Gaussian. The marginalized probability distribution for any
-% single binary parameter (say $\alpha$) can be obtained by integrating
-% Eq.~\ref{eq:postprob} over all other parameters, i.e. 
-% \begin{equation}
-%  p(\alpha | d_n, H) = \int\D \vec{\Theta}_\alpha\, p(\vec{\Theta} | d_n, H),
-% \end{equation}
-% where $\vec{\Theta}_\alpha$ is the set of remaining parameters, i.e.
-% $\vec{\Theta}_\alpha:=\vec{\Theta} - \{\alpha\}$.
-% 
-% 
-% 
-% We sample the probability distribution of Eq.~\ref{eq:postprob} using
-% an ensemble sampler Markov-chain Monte-Carlo algorithm based on the {\tt emcee}
-% package~. A total of $100$ independent walkers were used, each of which
-% was allowed to collect $100,000$ samples. We measure the correlation time and keep
-% only $20,000$ independent posterior points to further analyze the calculated 
-% posterior probability distributions. 
 % 
@@
 \centering 
-\includegraphics[trim=20 20 18 22 0,clip=true,width=1.95\columnwidth]{plots/TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
+\includegraphics[trim=20 20 18 22 0,clip=true,width=1.95\columnwidth]{TNMchirpBiasesOverCIWidths_CI90_0_Lambda_SNR30_70_linear}
 \caption{{\bf Ratio of systematic to statistical errors in measuring $\mchirp$, ignoring tidal effects:}
@@
 \centering
-\includegraphics[trim=20 20 18 21 0,clip=true,width=1.95\columnwidth]{plots/TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_50_linear}
+\includegraphics[trim=20 20 18 21 0,clip=true,width=1.95\columnwidth]{TNEtaBiasesOverCIWidths_CI90_0_Lambda_SNR20_50_linear}
 \caption{{\bf Ratio of systematic to statistical errors in measuring $\eta$, ignoring tidal effects:}
@@
 \centering
-\includegraphics[trim=20 20 18 20 0,clip=true,width=1.95\columnwidth]{plots/TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
+\includegraphics[trim=20 20 18 20 0,clip=true,width=1.95\columnwidth]{TNChiBHBiasesOverCIWidths_CI90_0_Lambda_SNR_linear}
 \caption{{\bf Ratio of systematic to statistical errors in measuring $\chibh$, ignoring tidal effects:}
@@
 \centering 
-\includegraphics[trim=10 20 8 18 0,clip=true,width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
+\includegraphics[trim=10 20 8 18 0,clip=true,width=1.9\columnwidth]{SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
 \caption{{\bf Illustrative posterior probability distributions for NS tidal
@@
 
-% \textcolor{gray}{%
-% Let us first delve into statistical errors. In
-% Fig.~\ref{fig:CIWidths90_Lambda_SNR}, we show how {\it precisely} can we
-% measure NSBH parameters $X=\{\mchirp,\eta,\chibh\}$ using BBH templates.
-% The three panels correspond to $\mchirp$ (top), $\eta$ (middle), and $\chibh$
-% (bottom), and show the width of these credible intervals $(\Delta X)^{90\%}$
-% as a function of BH mass/spin (within each sub-panel), and NS properties, i.e.
-% $\lambdans$ (downwards in each column)~\footnote{We restrict NS mass to
-% $1.35M_\odot$ and its spin to zero. Varying its tidal deformability $\lambdans$
-% does not significantly change the measurement uncertainties for non-tidal
-% binary parameters, as is evident from comparing the two rows in each panel of
-% Fig.~\ref{fig:CIWidths90_Lambda_SNR}.}.
-% % 
-% From the left-most column, we find that: (i) at $\rho=20, \mchirp$ is already
-% measured remarkably well - to a precision of $0.16\%$ of its true value, and
-% (ii) so is $\chibh$. (iii) Mass ratio $\eta$ is determined more loosely,
-% with $\gtrsim 25\%$ uncertainty. If the signal is even louder ($\rho\geq 30$),
-% all three measurements gain further precision, with $\eta$ errors shrinking down
-% to single-digit percents of its true value.
-% % % 
-% % We remind ourselves that this {\it precision}, however, is only
-% % meaningful so long as the measurement is {\it accurate} to begin with.
-% % We define $\arr_X$ as the ratio between systematic and statistical
-% % errors associated with the measurement of parameter $X$,
-% % \begin{equation}\label{eq:arr}
-% % \arr_X = \dfrac{(X^\mathrm{Median} - X^\mathrm{Injected})}{(\Delta X)^{90\%}},
-% % \end{equation}
-% % in order to compare the relative importance of both. Only when
-% % $|\arr_X| < 1$ can we hope that ignoring tidal effects in our templates
-% % will not hamper the measurement of non-tidal parameters from NSBH signals.
-% % Due to
-% % unmodeled tidal effects, parameter measurements performed on NSBH signals with
-% % BBH templates will have systematic biases. If these biases are too large,
-% % using non-tidal BBH templates should severely impair our inferencing ability.
-% % 
-% % there are physical effects that have not been
-% % incorporated in our (BBH) search templates which we use here to decipher NSBH
-% % signals. Therefore, we cannot know the {\it accuracy} of our parameter
-% % measurements unless we know how large the systematic biases due to an
-% % incomplete template model are. To quantify this, we find it useful to define
-% % $\arr_X$, the ratio of the systematic biases induced by unmodeled tidal effects in
-% % templates, to the statistical uncertainties associated with the measurement
-% % itself. If $\arr_X\gtrsim 1$, including the effects of NS tidal distortion
-% % (during inspiral) and possible disruption (close to merger) would be imperative
-% % to obtain {\it accurate} estimates of source masses and spins from observed GW
-% % signals.}
-% }
-
-%\textcolor{blue}{%
 Credible intervals $(\Delta X)^{90\%}$ showing the precision with which
@@
 \centering    
-\includegraphics[trim=35 21 15 21 0,clip=true,width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
-% \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
+\includegraphics[trim=35 21 15 21 0,clip=true,width=2.2\columnwidth]{TTLambdaRawCIWidths90_0_Lambda_SNR}
+% \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{TTLambdaCIWidths90_0_Lambda_SNR.pdf}
 \caption{{\bf Statistical uncertainty in $\lambdans$ measurement:}
@@
 \centering    
-\includegraphics[trim=10 10 0 10 0,clip=true,width=1.05\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}
-% \includegraphics[width=1.025\columnwidth]{plots/TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
+\includegraphics[trim=10 10 0 10 0,clip=true,width=1.05\columnwidth]{TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}
+% \includegraphics[width=1.025\columnwidth]{TTSNRThresholdFor100LambdaMeasurement_BHspin_BHmass_Lambda2000_0_CI90_0}\\
 \caption{
@@
 %
-%In addition, a SNR of $30$ corresponds to a $1.35+5M_\odot$ binary with 
-%$\chibh = 0.4$ at a luminosity distance of $XXX$~Mpc, if optimally aligned
-%with LIGO ($XXX$~Mpc if the binary subtends a $45^o$ angle with LIGO's line of
-%sight.
-%
-%These conditions are strong, but not unrealistic. For e.g., a detection rate
-%of $20$ NSBHs a year implies $1-2$ of them will have SNRs$\gtrsim 30$, since
-%the number of expected events scales as $1/\rho^2$ if we assume that sources
-%are distributed uniformly in spacial volume.
 %
@@
 We will continue here to focus on the measurement of $\lambdans$ itself.
-% }
-% 
-% \textcolor{gray}{%
-% Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
-% mass, spin, and tidal parameter measurements. We keep the binary parameters
-% as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
-% set $\rho=30$ (left panel) or $\rho=50$ (right panel).
-% %
-% We find that the measurement of $\lambdans$ is weakly degenerate with
-% other parameters, and at realistic SNRs it would improve by a few tens of 
-% percent if we knew non-tidal parameters to better accuracy. The predominant
-% factor that would enhance the measurement accuracy for $\lambdans$ is the
-% signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
-% measurement to be limited by its degeneracy with non-tidal parameters at 
-% a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
-% }
 
@@
 \centering    
-\includegraphics[trim=18 18 18 10 0,clip=true,width=\columnwidth]{plots/pdfLambda_vs_N_L800.pdf}\\
-\includegraphics[trim=18 18 18 10 0,clip=true,width=\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
+\includegraphics[trim=18 18 18 10 0,clip=true,width=\columnwidth]{pdfLambda_vs_N_L800.pdf}\\
+\includegraphics[trim=18 18 18 10 0,clip=true,width=\columnwidth]{FillBetweenErrorBarsLambda_vs_N_L800.pdf}
 \caption{{\bf Recovery of $\lambdans$ for an increasingly large population of BH-NS signals.}
@@
 \centering    
-\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
+\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{FillBetweenRelErrorBarsLambda_vs_NShifted_AllLambda.pdf}
 \caption{{\bf Improvement in $\lambdans$ measurement accuracy for different NS EoS:}
@@
 % \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
-\includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{plots/LambdaMedian_vs_N_AllPopulation}
-\includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{plots/LambdaMedian_vs_N_AstroPopulation}\\
-\includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{plots/LambdaMedian90pc_vs_N_AllPopulation}
-\includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{plots/LambdaMedian90pc_vs_N_AstroPopulation}\\
+\includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{LambdaMedian_vs_N_AllPopulation}
+\includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{LambdaMedian_vs_N_AstroPopulation}\\
+\includegraphics[trim=1cm 0 0 0, width=1.025\columnwidth]{LambdaMedian90pc_vs_N_AllPopulation}
+\includegraphics[trim=0 0 1cm 0, width=1.025\columnwidth]{LambdaMedian90pc_vs_N_AstroPopulation}\\
 \caption{
@@
 independent population draws within each ensemble. One curve in each ensemble
-is highlighted in color, representing only that it is the same population as 
-was discussed in
-Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}-\ref{fig:TT_Lambda_vs_N_CI90_0}.
+is highlighted in color, representing the realizations already plotted in
+Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}.
 In the same color we show $\pm 10\%$ error-bounds on $\lambdans$ with
@@
 % \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
-\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{plots/LambdaCIWidths90pc_vs_N_AllPopulation}
-\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{plots/LambdaCIWidths90pc_vs_N_AstroPopulation}
+\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{LambdaCIWidths90pc_vs_N_AllPopulation}
+\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{LambdaCIWidths90pc_vs_N_AstroPopulation}
 \caption{{\bf No Mass-Gap} ({\it left}): This panel shows the width of $\lambdans$ interval within
@@
 % 
+One curve in each ensemble is highlighted in color (thin lines), representing the 
+realizations already plotted in Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}.
 {\bf Mass-Gap} ({\it right}): This panel shows populations drawn under paradigm B, which
@@
 % 
-% \begin{equation}\label{eq:rmsSigmaIdenticalSources}
-%  \sigma_{avg} := \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}},
-% \end{equation}
-% 
-% 
-% 
-% 
-% \prayush{%
-% The approximation of using (only) the dominant $l=|m|=2$ multipoles of
-% gravitational-wave strain to construct the signal, ignoring other $l\neq 2$
-% multipoles, has the advantage of making various angle parameters 
-% associated with the source degenerate with its distance. As a result, source
-% distance, source sky location angles, and the inclination angle of
-% the binary's orbital angular momentum with respect to the detector,
-% all combine into a single ``effective distance'' $\deff$ that scales
-% out as the constant amplitude factor $1/\deff$ (constant because none
-% of the above parameters vary temporally for aligned-spin binaries).
-% In order to consider multiple events, lets imagine a population uniformly
-% distributed in effective volume, i.e. within a sphere of radius 
-% $\deff^\mathrm{max}$.
-% }
-% % 
-% \red{These sentences are distracting. I suggest to simply say 'sources
-% uniform in volume'. Probably not necessary to discuss effective distance
-% at all. If so, then do so in a footnote.}
-% \red{Explain (or state, if explanation is trivial) that uniform in volume,
-% uniform in orientation, is equivalent to uniform in effective volume.}
-% % 
-% \prayush{%
-% The radius of this sphere is set by the lowest SNR that 
-% is distinguishable from noise by LIGO searches. Now, divide the sphere into
-% $I$ shells of equal thickness. The radius of the $i$-th sphere would then
-% be $D_i = \deff^\mathrm{max} (i - 1/2)/I$.}
-% \red{..into $I$ shells of same radial thickness. The radius at the middle of
-% the $i-$th shell $(i=1,2,...,I)$ is $D_i:=\frac{i-\frac{1}{2}}{I}\,D_\mathrm{max}$,
-% and the number of sources in this shell is $N_i=4\pi\beta D_i^2 \frac{D_\mathrm{max}}{I}$,
-% where $\beta$ is the volume density of sources.
-% }
-% The measurement uncertainty in
-% $\lambdans$ scales inversely with SNR, and hence directly with the 
-% effective distance to the source. Therefore, if we have a measurement error 
-% $\sigma_0$ for a source located at $\deff = D_0$, the same error for the same
-% source located within the $i$-th shell would be 
-% $\sigma_i = \sigma_0 \dfrac{D_i}{D_0}$. The combined error from independent
-% measurements of $\lambdans$ for identical sources at different distances is
-% given by
-% % 
-% \begin{equation}\label{eq:1oversigma}
-% \frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
-% \end{equation}
-% where $N_i$ is the number of sources detected in the $i$-th shell.
-% % , and is a random
-% % variable, with its probability proportional to the volume of the shell, i.e.
-% % $$
-% % p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
-% % $$
-% % Therefore, the expected number of detections in the $i$-th shell would be
-% % $$
-% % \langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
-% % $$
-% % where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
-% % to be Poisson distributed around the mean detection rate
-% % $\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
-% % $\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
-% % a year per unit effective volume, then
-% % \begin{equation}
-% % \langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
-% % \end{equation}
-% % and
-% % \begin{eqnarray}
-% %  \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
-% %  &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
-% % \end{eqnarray}
-% % where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
-% % to an integral. 
 The root-mean-square (RMS) averaged measurement error from $N$ sources is 
@@
 ($\propto D^3$).
-% ~\footnote{{\it effective} distance $D$ 
-% is a combination of distance to the source, its orientation, and its sky
-% location angles, and has a one-to-one correspondence with SNR for non-precessing
-% systems, for which these angles are constant in time. Effective volume is 
-% $\frac{4\pi}{3}D^3$.}
 However, it 
@@
 %
-% % 
-% Using Bayes' theorem, the measured probability distribution function for 
-% $\lambdans$ from $N$ observations is given by~\footnote{%
-% Eq.~\ref{eq:lambdaMultiple} includes the following implicit factor to normalize
-% its right-hand side: $\dfrac{\prod_i p(d_i|H)}{\int p(\lambdans)\,p(d_1, d_2, \cdots, d_N | \lambdans; H)\,\D\lambdans}$}
-% \begin{equation}\label{eq:lambdaMultiple}
-%  p(\lambdans | d_1, d_2, \cdots, d_N; H) = p(\lambdans | H)^{1-N}\prod_{i=1}^N p(\lambdans | d_i, H),
-% \end{equation}
 
@@
 % %
-% This is a promising deduction, as one might expect to see $\mathcal{O}(10)$
-% events over a few years' timescale with design aLIGO~\cite{Abadie:2010cfa}.
 
@@
 
-% \prayush{\bf Sec 5 summary}\\
-% \prayush{\bf Tidal measurements versus observation time - in how many years can
-% we constrain $\lambdans$ reasonably?}\\
-% % 
 Third, knowing that single observations can furnish only so much information
@@
 \centering 
-\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNMchirpCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNEtaCIWidths90_0_Lambda_SNR}\\
-\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{plots/TNChiBHCIWidths90_0_Lambda_SNR}
+\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{TNMchirpCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{TNEtaCIWidths90_0_Lambda_SNR}\\
+\includegraphics[trim = {2cm 0 0 0},width=2.\columnwidth]{TNChiBHCIWidths90_0_Lambda_SNR}
 \caption{{\bf Statistical measurement uncertainty for NSBH parameters,
@@
 \centering
-\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
-% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr30}%\\
-\includegraphics[width=1.05\columnwidth,trim=0.7cm 0 1cm 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
-% \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
-% \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
-% \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
-% \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
-% \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
+\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
+\includegraphics[width=1.05\columnwidth,trim=0.7cm 0 1cm 0]{AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
 \caption{{\bf Illustrative posterior probability distributions for NSBH parameters,
@@
 
-% #################
-% \begin{figure*}
-% \centering    
-% \textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
-% \includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
-% \includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
-% \caption{%
-% These panels show the accumulation of information from a population whose
-% $\lambdans=800$.
-% {\it Left:} Posterior probability distributions for $\lambdans$ (colored 
-% curves), and associated $90\%$ credible intervals (grey vertical lines),
-% shown for different number of accumulated observations $N$. Distributions are
-% normalized to unit area.
-% {\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
-% associated $90\%$ credible intervals (as the vertical extent of filled
-% region), shown as a function of number of observations $N$. Solid horizontal
-% line indicates the true value of $\lambdans$. Dashed and dotted horizontal
-% lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
-% bounds.
-% }
-% \label{fig:TT_Lambda_vs_N_L800_CI90_0}
-% \end{figure*}
-% %
-% #################
-
 In Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
@@
 \centering    
-\includegraphics[width=1.05\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsN_vs_N.pdf}
+\includegraphics[width=1.05\columnwidth]{PowerLawCoefficient_LambdaErrorvsN_vs_N.pdf}
 \caption{%
@@
 \centering    
-% \includegraphics[width=1.03\columnwidth]{plots/LambdaRelErrorBars_vs_Lambda_AllPopulations_N80_Log.pdf}
-\includegraphics[width=\columnwidth]{plots/PowerLawCoefficient_LambdaErrorvsLambda_vs_N_AllPopulations.pdf}
+\includegraphics[width=\columnwidth]{PowerLawCoefficient_LambdaErrorvsLambda_vs_N_AllPopulations.pdf}
 \caption{%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{References/References}
+\bibliography{paper}
 

### 2016-12-17 Paper/journal_submission/paper.tex
added phenomc vs eob based lea comparison

@@
 
+% ###########################################################
+% ###########################################################
+\section{Choice of underlying BBH model in LEA}
+% ###########################################################
+% 
+\begin{figure}
+\centering    
+\includegraphics[width=1.05\columnwidth]{match-q-spin1-PhenomC-mf0_01.pdf}
+\caption{%
+We compare two alternatives of the tidal NSBH model from Ref.~\cite{
+Lackey:2013axa}, which differ in their underlying BBH prescriptions. One which
+we use in this study uses SEOBNRv2, while the other uses IMRPhenomC as its
+base. In this figure, we show the normalized overlap (match) between the two for
+$2,000,000$ points sampled uniformly in the NSBH parameter space. We find that
+for $q\gtrsim 4$ the discrepancies between IMRPhenomC and SEOBNRv2 as reported
+in~\cite{Kumar:2015tha} dominate over tidal terms.
+}
+\label{fig:PhenomC_vs_SEOBNRv2_LEA}
+\end{figure}
+%
+The waveform model used in this paper is a variant of those calibrated
+in Ref~\cite{Lackey:2013axa}. In that work, the authors also calibrate
+a tidal prescription with the phenomenological model IMRPhenomC~\cite{
+Santamaria:2010yb}
+as the base BBH model. Previous work~\cite{Kumar:2015tha} has shown that 
+IMRPhenomC can exhibit pathological behavior for mass-ratios ever so
+$q\gtrsim 4$ and/or non-zero black hole spins. We compute noise-weighted
+inner-products between the two variants for $2,000,000$ points sampled
+over the NSBH parameter space, and show the results in Fig.~\ref{
+fig:PhenomC_vs_SEOBNRv2_LEA}. We restrict the comparison to frequencies
+that are affected by the tidal disruption of the NS, by integrating
+the inner-products from $f = \mathrm{max}(15, 0.01/M)$~Hz
+(where $M$ is expressed in seconds ($1M_\odot \simeq 4.925\mu$S, see
+Eq.~(32-34) of~\cite{Lackey:2013axa}).
+% 
+We find that the differences between the two BBH approximants dominate
+over the tidal terms that are added to their phasing. As SEOBNRv2 is
+more reliable than IMRPhenomC, we recommend the use of SEOBNRv2-based
+LEA+ in upcoming LIGO-Virgo analyses.
+
 \end{appendix}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{paper}
+\bibliography{References}
 

### 2017-01-03 emcee_PE/scripts/MultipleObservations.py
improvements to cum stat reading func, new plots

@@ -1,4 +1,4 @@
-# 
+#
 __author__ = "Prayush Kumar <prayush.kumar@ligo.org>"
 
 import os, sys, time
@@ -86,7 +86,7 @@ plt.rcParams.update({\
 
 
 #######################################################
-### FUNCTIONS 
+### FUNCTIONS
 #######################################################
 def obtain_statistical_information_from_kde(\
                 kde_func=None,\
@@ -175,7 +175,7 @@ element in vec closest to it.
     if idx_min < 0 or idx_min >= len(vec):
         raise RuntimeError("Could not find closest match!")
     return vec[idx_min]
-    
+
 
 def sample_snr_from_volume(SNRvec):
     '''
@@ -226,12 +226,12 @@ def sample_snr_from_volume_inclination(SNRvec, q, chi,\
                         psd = None, verbose=False, vverbose=False):
     '''
 Sample the SNR of signals, as if they are distributed uniformly in
-the spacial volume, i.e. 4/3 pi D^3, and simultaneously have 
+the spacial volume, i.e. 4/3 pi D^3, and simultaneously have
 a. inclination angle with respect to our line of sight uniformly distributed \in [0, pi]
 b. polarization angle
 c. sky location angles on a 2-sphere
 
-Inputs are : 
+Inputs are :
 1. list of SNR values for Nearest-neighbor interpolation
 2,3. Chosen Q and CHI values
 
@@ -304,13 +304,13 @@ Return the nearest neighbor of the drawn SNR from the list of SNR values given
       #
       signal_snr = sigma(h, psd=psd, low_frequency_cutoff=f_lower)
       if verbose: print "SNR this iteration.. = %.2f" % signal_snr
-    #  
+    #
     SNRvec    = np.array(SNRvec)
     if verbose: print "\n\n"
     return find_closest_match(signal_snr, SNRvec)
     #
     raise RuntimeError("SNR sampling not being done properly")
-  
+
 
 #######################################################
 ### CALCULATION CLASSES
@@ -496,7 +496,7 @@ What it does:
             #lambda_gaussian_kde = gaussian_kde(lambda_chain)
             #lambda_multivariate_kde = KDEMultivariate(lambda_chain, bw=0.2*np.ones_like(lambda_chain), var_type='c')
             #lambda_skl_kde = KernelDensity(bandwidth=0.2)
-            #lambda_skl_kde.fit(lambda_chain[:, np.newaxis])    
[…]

### 2017-01-03 Paper/journal_submission/paper.tex
adding more figs to text

@@
 \end{figure*}
+% 
+\begin{figure*}
+\centering    
+% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
+\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{LambdaMedianCIWidths90pc_vs_N_AllPopulation_SNRSorted}
+\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{LambdaCIWidths90pc_vs_N_AstroPopulation_SNRSorted}
+\caption{This figure is similar to Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne},
+with the difference
+that events in each population have been sorted according to their
+signal strength (SNR). We note that most of the information about
+the NS tidal deformation comes from the loudest $5-10$ events, for
+both paradigms.
+}
+\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}
+\end{figure*}
 

### 2017-01-05 Paper/journal_submission/paper.tex
caption of new figs in text
addressed PRD comments partially

@@
 % 
-During O1, they made the first terrestrial observation of gravitational 
+During O1, they made the first direct observation of gravitational 
 waves~\cite{LIGOVirgo2016a}. Emitted by a pair of coalescing black holes, these
@@
 coming years.
-% 
-% At design sensitivity, we expect aLIGO to observe $\sim 70$ compact binary
-% mergers a year~\cite{Abadie:2010cf}.
-
-% \red{
-% Text about CBCs. How was each type observed so far?
-% }
+
 Coalescing compact binaries of stellar-mass black holes (BH) and/or
@@
 However, whether or not such a system forms depends also on the nature of
-the BH. Massive BHs (with $\mbh\gtrsim 10M_\odot$), as well as BHs with
+the BH. Massive BHs (with $\mbh\gtrsim 12M_\odot$), as well as BHs with
 large retrograde spins, tend to swallow the NS whole without forming a
@@
 % 
-On the other hand, {\it low-mass} BHs with $\mbh\in[3M_\odot, 8M_\odot]$, 
+On the other hand, {\it low-mass} BHs with $\mbh\in[3M_\odot, 12M_\odot]$\footnote{
+The upper limit on BH mass that allows for NS disruption may very well be
+higher, depending strongly on the magnitude of BH spin~\cite{Foucart:2014nda}.},
 can disrupt their companion NSs much before merger, forming long-sustained disks
-that are required to sustain 
-SGRBs~\cite{Shibata:2007zm,2010PhRvD..81f4026F,Lovelace:2013vma,Kawaguchi:2015}.
+that are required to sustain SGRBs~\cite{Shibata:2007zm,2010PhRvD..81f4026F,
+Lovelace:2013vma,Foucart:2014nda,Kawaguchi:2015}.
 % 
@@
 A {\it coincident} detection of both GWs and gamma-rays from an NSBH merger,
-will
-provide us with a unique opportunity to confirm this hypothesized link between
-NSBH mergers and GRBs~\cite{Abbott:2016wya}.
+will provide us with a unique opportunity to confirm this hypothesized link
+between NSBH mergers and GRBs~\cite{Abbott:2016wya}.
 
@@
 the NS EoS from inspiral through to merger. During early inspiral, the tidal
-dephasing is relatively weak and appears as a $5^{th}$ Post-Newtonian (PN)
-order effect~\cite{Vines2011}. Closer to merger, a disruptive fate of the NS
+dephasing is relatively weak and has a frequency dependence equivalent to a
+$5^{th}$ Post-Newtonian (PN) order effect~\cite{Vines2011}. Closer to merger,
+a disruptive fate of the NS
 can result in a strong suppression of GW emission above a cut-off 
@@
 models~\cite{Lackey2011,Lackey:2013axa} use the Fisher matrix to estimate
-$\lambdans$ measurement errors. Fisher matrix estimates are known to be
+$\lambdans$ measurement errors. Fisher matrix estimates may become
 unreliable at realistic signal-to-noise ratios (SNR)~\cite{Vallisneri:2007ev},
 such as those as we might expect in the upcoming observing runs of GW
-detectors~\cite{Abadie:2010cf}.
+detectors~\cite{Abadie:2010cf}, and we supplement such studies with a
+fully Bayesian treatment of the problem here.
 
@@
 In this paper, we restrict our parameter space to span mass-ratios
-$q:=\mbh/\mns\in[2,5]$, BH spin (aligned with orbit)
+$q:=\mbh/\mns\in[2,5]$, dimensionless BH spin (aligned with orbit)
 $\chibh\in[-0.5, +0.75]$, and dimensionless NS tidal deformability 
@@
 %
-% ###############################
-% In this section we summarize various technical aspects of this paper. The model
-% used to obtain tidal waveforms is described in Sec.~\ref{s2:waveforms}. The 
-% process of inferring source parameters from a GW signal is summarized in 
-% Sec.~\ref{s2:bayesian}. The process of generating and combining multiple events
-% is instead detailed in the corresponding results section,
-% Sec.~\ref{s2:astro_multiple}.
 
@@
 $\lambdans$. It was calibrated to a suite of $134$ numerical relativity (NR)
-simulations of NSs inspiraling into spinning BHs. The parameter space these
-simulations span, which we assume to be the domain of validity for the LEA
-model, includes NS masses $1.2M_\odot\leq\mns\leq 1.45M_\odot$,
+simulations of NSs inspiraling into spinning BHs, with
+NS masses ranging between $1.2M_\odot\leq\mns\leq 1.45M_\odot$,
 mass-ratios $2\leq q\leq 5$, and BH spins $-0.5\leq\chibh\leq+0.75$.
@@
 \begin{equation}\label{eq:likelihood}
- p(d_n| \vec{\Theta}, H) \equiv \LL(\vec{\Theta}) = \mathcal{N}\, \mathrm{exp}[- \langle d_n - h | d_n - h\rangle ],
+ p(d_n| \vec{\Theta}, H) \equiv \LL(\vec{\Theta}) = \mathcal{N}\, \mathrm{exp}[-\frac{1}{2} \langle d_n - h | d_n - h\rangle ],
 \end{equation}
@@
 detector-noise weighted inner-product~\footnote{The inner product
-$\langle\cdot|\cdot\rangle$  is defined as~\cite{Shoemaker2009}
+$\langle\cdot|\cdot\rangle$  is defined as
 \begin{equation}
-\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde(b)(f)^*}{S_n(|f|)}\,\D f\right],
+\langle a|b\rangle \equiv 4\,\mathrm{Re}\left[\int_0^\infty \dfrac{\tilde{a}(f) \tilde{b}(f)^*}{S_n(|f|)}\,\D f\right],
 \end{equation}
 where $\tilde{a}(f)$ is the Fourier transform of the finite time series $a(t)$,
-and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise.}, and
+and $S_n(|f|)$ is the one-sided amplitude spectrum of detector noise. In this
+work, we use the zero-detuning high-power design sensitivity curve~\cite{
+Shoemaker2009} for Advanced LIGO, with $15$Hz as the lower frequency cutoff.}, and
 $\mathcal{N}$ is an SNR-dependent normalization constant.
@@
 that events in each population have been sorted according to their
-signal strength (SNR). We note that most of the information about
-the NS tidal deformation comes from the loudest $5-10$ events, for
-both paradigms.
+signal strength (SNR). We note that information about
+the tidal deformability of neutron stars comes primarily 
+from the loudest $5-10$ events, whether we allow BH masses
+in the mass gap (left panel) or restrict them to
+$\mbh\geq 5M_\odot$ (right panel).
 }
@@
 
-The pioneering terrestrial observation of gravitational waves by Advanced LIGO
+The pioneering observation of gravitational waves by Advanced LIGO
 harbingers the dawn of an era of gravitational-wave astronomy where observations

### 2017-01-06 Paper/journal_submission/paper.tex
more changes as per referee

@@
 We start with generating a signal waveform, using the model LEA+, and injecting
-it in zero noise to obtain a stretch of data $d_n$. Source binary parameters
-$\vec{\Theta}:=\vec{\theta}\cup\{\lambdans\}$ are reconstructed from this
-injected signal. Using Bayes' theorem, their joint inferred probability
-distribution of $\vec{\Theta}$ can be written as
+it in zero noise to obtain a stretch of data $d_n$. Source intrinsic parameters
+$\vec{\Theta}:=\{\mbh,\mns,\chibh,\lambdans\}$ are reconstructed from this
+injected signal. Extrinsic parameters $\vec{\theta}:=\{t_c,\phi_c\}$ 
+representing the time of and phase at the arrival of signal are marginalized
+over numerically and analytically (respectively), while source location and
+orientation parameters such as its luminosity distance, sky location, inclination
+and polarization angles are absorbed into a normalization, as describe later,
+and subsequently maximized over. This is justified because in this paper we
+consider the {\it single-detector case}. Using Bayes' theorem, the joint inferred
+probability  distribution of $\vec{\Theta}$ can be evaluated as
 \begin{equation}\label{eq:postprob}
@@
 except for expectations on binary parameters that enter
-our calculations explicitly. Throughout this paper, we impose a prior that is
-uniform in
-individual component masses, BH spin, and the tidal deformability of the NS. In
-addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not calibrated for 
-$1\leq q\leq 2$. $p(d_n|\vec{\Theta}, H)$ is the
-likelihood of obtaining the given stretch of data $d_n$ if we assume that a
+our calculations explicitly. Throughout this paper, we impose priors that are
+uniform in individual component masses, BH spin, and the tidal deformability of
+the NS. In addition, we restrict mass-ratios to $q\geq 2$, as LEA+ is not
+calibrated for $1\leq q\leq 2$. $p(d_n|\vec{\Theta}, H)$ is the {\it likelihood}
+of obtaining the given stretch of data $d_n$ if we assume that a
 signal parameterized by $\vec{\Theta}$ is buried in it, and is given by
@@
 $\vec{\Theta}$, $\langle\cdot|\cdot\rangle$ is a suitably defined
-detector-noise weighted inner-product~\footnote{The inner product
+detector-noise weighted inner-product\footnote{The inner product
 $\langle\cdot|\cdot\rangle$  is defined as
@@
 work, we use the zero-detuning high-power design sensitivity curve~\cite{
-Shoemaker2009} for Advanced LIGO, with $15$Hz as the lower frequency cutoff.}, and
-$\mathcal{N}$ is an SNR-dependent normalization constant.
-As in Ref.~\cite{Purrer:2015nkh} we use a likelihood that is maximized over
-over the template norm
-and we ignore extrinsic parameters that only enter in the template norm such
-as distance, orientation and sky location. In addition, we also maximize over
-coalescence time and phase.
+Shoemaker2009} for Advanced LIGO, with $15$Hz as the lower frequency cutoff.},
+and $\mathcal{N}$ is the normalization constant that absorbs source distance,
+orientation and sky location parameters. As in Ref.~\cite{Purrer:2015nkh} we
+use a likelihood that is maximized over the template
+norm, allowing us to ignore the extrinsic parameters that only enter in the
+template norm through $\mathcal{N}$. As a result, we only need to sample over
+$\vec{\Theta}$ (or $\vec{\Theta} - \{\lambdans\}$ in the case of non-tidal
+templates).
+%In addition, we also maximize over coalescence time and phase.
 The denominator in Eq.~\ref{eq:postprob} is the {\it a priori} probability of finding
 the particular signal in $d_n$ and we assume that each injected signal is as
-likely as any other. Having constructed the probability distribution function
-$p(\vec{\Theta} | d_n, H)$, extracting the measured probability distribution
+likely as any other. From the joint probability distribution
+$p(\vec{\Theta} | d_n, H)$ so constructed, extracting the measured probability distribution
 for a single parameter (say $\alpha$) involves integrating
@@
 i.e. $\mchirp$ is always being over-estimated. This is to be expected since
-the tidal deformation and disruption of NSs close to merger reduces GW signal
-power at high frequencies, making the resulting signal resemble a BBH signal of
-lower merger frequency - and therefore of higher mass.
+the tidal deformation of the NS drains energy from the orbit faster than in
+the BBH case, and the disruption of NS close to merger reduces GW signal
+power at high frequencies. Both of these effects make the resulting signal
+resemble a BBH signal of higher chirp mass.
 
@@
 source location.
+
+It is interesting to note that the parameter ranges most favorable to
+the measurability of $\lambdans$ are also those which produce
+relatively more massive post-merger disks~\cite{Foucart2012}. That is, 
+the subset of NSBHs that potentially produce SGRBs (using a sufficiently-large
+disk mass as an indicator) would be the same subset most favorable
+for measurement of tidal effects. Therefore the rate of SGRBs in the 
+local universe (allowing for the fraction that are produced by NSBHs versus
+BNSs) would be an indicator of the rate of events most favorable for nuclear
+equation of state measurements.
 
@@
 \label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}
-\end{figure*}
-% 
-\begin{figure*}
-\centering    
-% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
-\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{LambdaMedianCIWidths90pc_vs_N_AllPopulation_SNRSorted}
-\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{LambdaCIWidths90pc_vs_N_AstroPopulation_SNRSorted}
-\caption{This figure is similar to Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne},
-with the difference
-that events in each population have been sorted according to their
-signal strength (SNR). We note that information about
-the tidal deformability of neutron stars comes primarily 
-from the loudest $5-10$ events, whether we allow BH masses
-in the mass gap (left panel) or restrict them to
-$\mbh\geq 5M_\odot$ (right panel).
-}
-\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}
 \end{figure*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 
+\begin{figure*}
+\centering    
+% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
+\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{LambdaMedianCIWidths90pc_vs_N_AllPopulation_SNRSorted}
+\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{LambdaCIWidths90pc_vs_N_AstroPopulation_SNRSorted}
+\caption{This figure is similar to Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne},
+with the difference
+that events in each population have been sorted according to their
+signal strength (SNR). We note that information about
+the tidal deformability of neutron stars comes primarily 
+from the loudest $5-10$ events, whether we allow BH masses
+in the mass gap (left panel) or restrict them to
+$\mbh\geq 5M_\odot$ (right panel).
+}
+\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}
+\end{figure*}
+%
 

### 2017-01-10 Paper/journal_submission/paper.tex
final version for resubmission

@@
 measurement accuracy.
+% 
+In both cases, we find that it is the loudest $5-10$ events that provide
+most of the tidal information, and not the combination of tens of low-SNR
+events, thereby facilitating targeted numerical-GR follow-ups of NSBHs.
 % 
@@
 additional observations would be required to attain the same measurement
-accuracy as above. And, (d) putting tighter constraints on the $\lambdans$ of a
+accuracy as above. (d) Putting tighter constraints on the $\lambdans$ of a
 population would require $50+$ NSBH observations, in any scenario.
+And, (e) it is the loudest $5-10$ events that will furnish the bulk
+of tidal information, and not the combination of a large number of 
+low-SNR events.
 % 
@@
 \end{figure*}
+% 
+\begin{figure*}
+\centering    
+% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
+\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{LambdaMedianCIWidths90pc_vs_N_AllPopulation_SNRSorted}
+\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{LambdaCIWidths90pc_vs_N_AstroPopulation_SNRSorted}
+\caption{This figure is similar to Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne},
+with the only difference being that events in each population
+have been sorted according to their signal strength (SNR), instead of
+their simulated chronology. We note that information about
+the tidal deformability of neutron stars comes primarily 
+from the loudest $5-10$ events, whether we allow BH masses
+in the mass gap (left panel) or restrict them to
+$\mbh\geq 5M_\odot$ (right panel).
+}
+\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}
+\end{figure*}
+%
 
@@
 
+{\it Do all events matter?} In Ref.~\cite{Lackey2014} the authors demonstrate
+that the overwhelming majority of information about the NS equation of state
+comes from the loudest $\sim 5$ events in the case of binary neutron star
+detections, and not from the combined effect of a large number of low-SNR
+systems.
+The question naturally arises if the same is true for NSBH sources as well.
+Therefore, in Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}
+we re-evaluate the accumulation of information with each successive NSBH 
+detection, sorting the events in each population by their SNR instead of 
+simulated chronology. We find the same qualitative behavior as in the case of 
+BNSs~\cite{Lackey2014}. 
+Whether or not there is an astrophysical mass gap, we find
+that the bulk of tidal information will be furnished by the loudest $5-10$
+NSBH detections of aLIGO detectors. This result is especially encouraging
+to NR follow-up efforts for GW detections, as we now know that only a handful
+of loudest NSBH events will merit full numerical-GR follow-up simulations.
+
 To summarize, in this section we study the improvement in our measurement of 
@@
 the NS deformability, which would allow us to rule out or rank different
-equations of state for neutron star matter. Our key findings are 
-summarized in Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} and
-\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}.
+equations of state for neutron star matter. Within this population, we also
+find that it will be the loudest $5-10$ events that will furnish most of the
+tidal information. Our key findings are summarized in 
+Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} -
+\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 
-\begin{figure*}
-\centering    
-% \textbf{No Mass-Gap \hspace{6cm} Mass-Gap}\par\medskip
-\includegraphics[width=1.025\columnwidth,trim=1cm 0 0 0]{LambdaMedianCIWidths90pc_vs_N_AllPopulation_SNRSorted}
-\includegraphics[width=1.025\columnwidth,trim=0 0 1cm 0]{LambdaCIWidths90pc_vs_N_AstroPopulation_SNRSorted}
-\caption{This figure is similar to Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne},
-with the difference
-that events in each population have been sorted according to their
-signal strength (SNR). We note that information about
-the tidal deformability of neutron stars comes primarily 
-from the loudest $5-10$ events, whether we allow BH masses
-in the mass gap (left panel) or restrict them to
-$\mbh\geq 5M_\odot$ (right panel).
-}
-\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne_SNRSorted}
-\end{figure*}
-%
 
@@
 detector operation.
+% 
+We also find that, within this population, the loudest $\sim 5-10$ events
+will provide us with most of the tidal information, and will therefore
+merit full NR follow-up.
 % 
@@
 in~\cite{Kumar:2015tha} dominate over tidal terms.
+\label{fig:PhenomC_vs_SEOBNRv2_LEA}
 }
-\label{fig:PhenomC_vs_SEOBNRv2_LEA}
 \end{figure}
@@
 inner-products between the two variants for $2,000,000$ points sampled
-over the NSBH parameter space, and show the results in Fig.~\ref{
-fig:PhenomC_vs_SEOBNRv2_LEA}. We restrict the comparison to frequencies
+over the NSBH parameter space, and show the results in Fig.~\ref{fig:PhenomC_vs_SEOBNRv2_LEA}. We restrict the comparison to frequencies
 that are affected by the tidal disruption of the NS, by integrating

### 2017-01-12 emcee_PE/scripts/MultipleObservations.py
updating snr-sorted figs with insets

@@ -807,6 +807,32 @@ READ the 'statistical_data' data structure to disk, in a self contained way.
         fin.close()
         return
     #
+    def read_snrs(self, chain_set_number=None,\
+                                filename=None, file_start_tag='',\
+                                NSLambda=None, N=None, sort=False,\
+                                verbose=False):
+        '''
+READ the 'statistical_data' data structure to disk, in a self contained way.
+        '''
+        if filename == None and chain_set_number == None:
+            raise IOError("Need to give either file name to chain set's RND'")
+        #
+        if filename is None or not os.path.exists(filename):
+            tmp_dir = '%s/L%d_N%d_' % (self.data_prefix, self.NSLambda, self.N)
+            self.RND = chain_set_number
+            self.TAG = tmp_dir + ('%d/' % self.RND)
+            finname = self.TAG + file_start_tag + 'chain_params.dat'
+        elif os.path.exists(filename):
+            finname = filename
+        #
+        if not hasattr(self, 'snr_data'): self.snr_data = np.array([])
+        fdata = np.loadtxt(finname)
+        for fline in fdata: self.snr_data = np.append(self.snr_data, fline[2])
+        if sort:
+            self.snr_data.sort()
+            self.snr_data = self.snr_data[::-1]
+        return
+    #
     def print_info(self):
         print """
 For each event:-

### 2017-01-13 Paper/journal_submission/paper.tex
incorporating MP's response to responses to PRD

@@
 i.e. $\mchirp$ is always being over-estimated. This is to be expected since
-the tidal deformation of the NS drains energy from the orbit faster than in
-the BBH case, and the disruption of NS close to merger reduces GW signal
-power at high frequencies. Both of these effects make the resulting signal
-resemble a BBH signal of higher chirp mass.
+the tidal deformation of the NS drains energy faster from the orbit during
+inspiral (as compared to the BBH case), and its disruption close to merger
+reduces GW signal power at high frequencies. Both of these effects make the
+resulting signal resemble a BBH signal of higher chirp (or total) mass, although
+we expect the latter effect to be dominant~\cite{Pannarale:2011pk}.
 
@@
 in the mass gap (left panel) or restrict them to
-$\mbh\geq 5M_\odot$ (right panel).
+$\mbh\geq 5M_\odot$ (right panel). Left inset zooms
+in on the main figure for the first $15$ events. Right inset
+shows the actual (ensemble mean) SNR value for each event. We find that
+events with $\rho\gtrsim 20-30$ provide the bulk of tidal information
+in our analysis.
 }
@@
 that the bulk of tidal information will be furnished by the loudest $5-10$
-NSBH detections of aLIGO detectors. This result is especially encouraging
+NSBH detections of aLIGO detectors. {\it This result is especially encouraging
 to NR follow-up efforts for GW detections, as we now know that only a handful
-of loudest NSBH events will merit full numerical-GR follow-up simulations.
+of loudest NSBH events (with SNRs $\rho\gtrsim 20-30$) are the ones that may merit
+full numerical-GR + magnetohydrodynamical follow-up simulations.}
 
@@
 % 
-We also find that, within this population, the loudest $\sim 5-10$ events
-will provide us with most of the tidal information, and will therefore
-merit full NR follow-up.
+We also find that, within this population, the loudest $5-10$ events (with SNRs
+$\rho\gtrsim 20-30$) will provide us with most of the tidal information,
+and will therefore merit full NR follow-up.
 % 

### 2017-01-18 Paper/journal_submission/paper.tex
final changes before resubmission

@@
 such as those as we might expect in the upcoming observing runs of GW
-detectors~\cite{Abadie:2010cf}, and we supplement such studies with a
+detectors~\cite{Abadie:2010cf}, and we improve such studies with a
 fully Bayesian treatment of the problem here.
@@
 Santamaria:2010yb}
-as the base BBH model. Previous work~\cite{Kumar:2015tha} has shown that 
-IMRPhenomC can exhibit pathological behavior for mass-ratios ever so
+as the base BBH model. Previous work~\cite{Kumar:2015tha,Kumar:2016dhh} has 
+shown that IMRPhenomC can exhibit pathological behavior for mass-ratios
 $q\gtrsim 4$ and/or non-zero black hole spins. We compute noise-weighted
 inner-products between the two variants for $2,000,000$ points sampled
-over the NSBH parameter space, and show the results in Fig.~\ref{fig:PhenomC_vs_SEOBNRv2_LEA}. We restrict the comparison to frequencies
+over the NSBH parameter space, and show the results in 
+Fig.~\ref{fig:PhenomC_vs_SEOBNRv2_LEA}. We restrict the comparison to frequencies
 that are affected by the tidal disruption of the NS, by integrating
@@
 % 
-We find that the differences between the two BBH approximants dominate
-over the tidal terms that are added to their phasing. As SEOBNRv2 is
-more reliable than IMRPhenomC, we recommend the use of SEOBNRv2-based
-LEA+ in upcoming LIGO-Virgo analyses.
+We find that the differences between the two BBH approximants cause mismatches
+of a few percent, while the tidal corrections contribute at a sub-percent
+level. We conclude that the differences of the underlying BBH model in LEA+
+dominates over their tidal calibration, and since SEOBNRv2 has been shown to
+be more reliable than IMRPhenomC~\cite{Kumar:2015tha,Kumar:2016dhh}, we 
+recommend the use of SEOBNRv2-based LEA+ in upcoming LIGO-Virgo analyses.
 

### 2017-01-18 Paper/journal_submission/paper.tex
resubmitted

@@
 % 
-We find that the differences between the two BBH approximants cause mismatches
+We find that the differences between the two variants of LEA+ have mismatches
 of a few percent, while the tidal corrections contribute at a sub-percent
 level. We conclude that the differences of the underlying BBH model in LEA+
-dominates over their tidal calibration, and since SEOBNRv2 has been shown to
+dominate over its tidal calibration, and since SEOBNRv2 has been shown to
 be more reliable than IMRPhenomC~\cite{Kumar:2015tha,Kumar:2016dhh}, we 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{References}
+\bibliography{paper}
 
